### 2019-02-12 chapter_00_open_science.tex
Add initial compilation including chapter from long fluiddyn metapaper

%% Journal of Open Research Software Latex template -- Created By Stephen
%% Bonner and John Brennan, Durham Universtiy, UK.
%% see http://openresearchsoftware.metajnl.com

\chapter{Towards reproducible, open science through open research software}

\section*{(1) Overview}

\vspace{0.5cm}

\section*{Title}

% \textcolor{blue}{The title of the software paper should focus on the software,
% e.g. “Text mining software from the X project”. If the software is closely
% linked to a specific research paper, then “Software from Paper Title” is
% appropriate. The title should be factual, relating to the functionality of the
% software and the area it relates to rather than making claims about the
% software, e.g. “Easy-to-use”.}

% old title

% FluidDyn: a Python open-source framework for research and teaching in fluid
% dynamics

% pa: this title may not be really good for this paper...
% - "research and teaching" may not be very important in the title
% - Recent progress in open-source software engineering and consequences on
%   science

% todo: find a better title!

Taking advantage of recent advances in open-source software engineering to study
fluid dynamics: the example of the FluidDyn project

\section*{Abstract}

% \textcolor{blue}{A short (ca. 100 word) summary of the software being
% described: what problem the software addresses, how it was implemented and
% architected, where it is stored, and its reuse potential.}

\href{http://fluiddyn.readthedocs.io}{FluidDyn} is a project to foster
open-science and open-source in the fluid dynamics community.  It is thought of
as a research project to channel open-source dynamics, methods and tools to do
better science.
%
We propose a set of Python packages forming a framework to study fluid dynamics
with different methods, in particular laboratory experiments (package
\fluidpack{lab}), simulations (packages \fluidpack{fft}, \fluidpack{sim} and
\fluidpack{foam}) and data processing (package \fluidpack{image}).
%
In the present article, we review the recent advancements in web-based software
development tools and Python which facilitated the execution of this project,
and establish a picture of software engineering in fluid mechanics.
%
We give an overview of the specialized packages of the project and then
focus on the base package called \fluidpack{dyn}, which contains common code
used in the specialized packages.  Packages \fluidpack{fft} and \fluidpack{sim}
are described with greater detail in two companion papers, \citet{fluidfft,
fluidsim}.
%
With the project FluidDyn, we demonstrate that specialized scientific code can
be written with methods and good practices of the open-source community. The
Mercurial repositories are available in Bitbucket
(\url{https://bitbucket.org/fluiddyn/}). All codes are documented using Sphinx
and Read the Docs, and tested with continuous integration run on Bitbucket
Pipelines and Travis.
%
To improve the reuse potential, the codes are as modular as possible, leveraging
the simple object-oriented programming model of Python.
%
All codes are also written to be highly efficient, using C++, Cython and
Pythran to speedup the performance of critical functions.

\section*{Keywords}

% \textcolor{blue}{keyword 1; keyword 2; etc. \\
% Keywords should make it easy to identify who and what the software will be
% useful for.}

Fluid dynamics research with Python. Free and open-source software, modular,
object-oriented, collaborative, efficient, tested, documented.

\section*{Introduction}

% \textcolor{blue}{An overview of the software, how it was produced, and the
% research for which it has been used, including references to relevant research
% articles. A short comparison with software which implements similar
% functionality should be included in this section. }

\subsection*{Science, software, open-source and the computer revolution}

Science is mainly a collective activity.  We can go further only by
\href{https://en.wikipedia.org/wiki/%
Standing_on_the_shoulders_of_giants}{``standing on the shoulders of giants''}
(and of a huge number of technicians and scientists).  Science is a lot about
how to build new knowledge from the work of others and hence, the exchange
of ideas is a fundamental aspect.
%
In the last decades, we have lived through a revolution on how people exchange
ideas.
%
Computers of all kinds (from smartphones to HPC clusters) connected by a world
wide web are used for human communication and also for many other applications.
%
It has become nearly effortless to reproduce and exchange ideas and data.
%
The set of intangibles that grows when shared and degrades when hoarded, such as
knowledge and love, has been somehow extended.
% Relevant quote: 
% O Goddess Saraswati, your Treasure is unique in nature.  It increases when it
% is freely spent and gets destroyed if it is hoarded (and not shared with
% others).
%
The information technology revolution open doors to fantastic opportunities for
human collaboration, and of course, for science.

Web related activities alone account for at least 5\% of GDP in the USA and the
European Union\footnote{See the report by Internet Association titled
\href{https://internetassociation.org/reports/refreshing-understanding-internet-economy-ia-report/}
{``Refreshing Our Understanding of the Internet Economy''}}.
% \cite{hooton_christopher_refreshing_2017}.
A huge amount of money (and work) is invested on the related technologies. We
are familiar with the most prominent companies involved in this dynamics
(Google, Facebook, etc.), but there are also several smaller and sometimes
lesser known organizations.  Most of these companies base part of their work on
the open-source paradigm contributing to open-source languages, libraries,
software and operating systems, while also using them --- a win-win situation
for both the corporations and the community.
%
This has lead to deep changes in software engineering, with a massive use of
open-source methods and tools, for example
\href{https://en.wikipedia.org/wiki/Distributed_version_control}{distributed
version control systems (DVCS)} and web-based source development platforms.

The computer performance continues to
\href{https://en.wikipedia.org/wiki/Moore\%27s_law}{increase exponentially},
now also with the help of Graphical Processing Units (GPU). This gave way to a
big boom in practical uses of data science and machine learning, which drives a
strong research on artificial intelligence.  This dynamics contributes to
progresses in open-source software.

These changes in our world also reflect in the way science is done.
%
Software and programming in science occupy a much bigger place than before.
%
% see http://sciencecodemanifesto.org/
% http://lorenabarba.com/gallery/reproducibility-pi-manifesto/
% https://www.numfocus.org/
%
The role of software in science has changed. In the past, coding was sometimes
considered as an inferior activity by some scientists.  The focus was on the
theory and the mathematical demonstration, which had to be elegant as it gets
included in the articles.  In contrast, it was normal to write crude code
% av: crappy = shitty!
% alternatives: shabby, inelegant, amateurish, unpolished, clumsy
% https://en.oxforddictionaries.com/thesaurus/amateurish
and to just show the results. Nowadays, codes tend to be at the heart of
research.

``Open-science'' is a new trend taking advantage of these new facts.  Pioneering
attempts are being made to do better science, improving reproducibility and
collective efficiency, by using the open-source methods and tools for science
and sharing and collaborating via the world wide web.

% http://www.nature.com/news/interactive-notebooks-sharing-the-code-1.16261

This paper presents FluidDyn, a project to foster open-science and open-source in
fluid mechanics. This first section ``Overview'' is organized into two large
parts.  A long introduction presents the context and describes our motivations and
the concept behind the project FluidDyn.  It is organized as follows:
\begin{itemize}
\item A scientific project based on the collaboration through open-source, such
as FluidDyn, is only possible thanks to the recent developments in methods and
tools for open-source software engineering. We start by presenting these
developments and discuss the implications for the scientists.

\item FluidDyn is based on the language Python. Incidently, it is the defining
characteristics of Python which make the eventual success of such a
project realizable.

Python is one of the most important tool in recent open-source dynamics and in
particular in science.
%
Python in fluid mechanics is growing fast, but is yet to reach widespread
adoption as the language of choice. Therefore, it is useful to present its
characteristics for readers who are unacquainted with its nuances.

\item We will discuss about the possible contradiction between productivity of
individuals and productivity at the community level.

\item We will establish a picture of software engineering in fluid mechanics
and present some of the main tools commonly used.

\item The introduction finally ends with a summary of the motivations for the
project FluidDyn and a presentation of its principles.
\end{itemize}

The section ``Overview'' is completed by a second subsection called
``Implementation and architecture'' devoted to present the organization of the
packages of the project and to describe in more detail the base package
\fluiddyn\footnote{We use FluidDyn (with capital letters) to name the project and
\fluiddyn for the base package.}.

\subsection*{Methods and tools for open-source software engineering}

\paragraph{Free and Open-Source Software (FOSS).} The term \emph{free} in
FOSS is a misnomer, as it actually stands for \emph{freedom} (to use, modify
and distribute). The FOSS movement has dramatically decreased the cost of using
computers, and is evident from the widespread use of GNU/Linux systems in
desktops, computing clusters and web-servers in academia and beyond.
%
The beginnings of FOSS culture can be traced back to the success for scientific
publication of the free type-setting standards \TeX\ (1977) and \LaTeX\ (1985).
%
Another founding moment for the FOSS movement was the launch of the GNU project
in 1983 by \href{https://en.wikipedia.org/wiki/Richard_Stallman}{Richard
Stallman}, to create a Unix-like computer operating system composed entirely of
free software\footnote{For completeness, see also the work done on
\href{https://www.levenez.com/unix/}{other Unix operating systems,
for example BSD}.}.
%
GNU is today known for its compiler collection (GCC) and a multitude of tools
which when combined with the Linux kernel (created by
\href{https://en.wikipedia.org/wiki/Linus_Torvalds}{Linus Torvalds} in 1991)
forms the GNU/Linux operating systems that we are familiar today.
%
FOSS movement has had huge successes in many frontiers --- Apache could be
termed as the ``first killer-app of Linux'' and now with an ever-increasing
suite of softwares including Firefox, LibreOffice, Gimp as solid alternatives
to proprietary offerings.
%
Linux has become the most widely used kernel, being deployed on servers,
personal computers, embedded devices, and also smart phones (with Android).

% remark Julien Salort: not interesting. Nothing on BSD.
% https://www.levenez.com/unix/

Over the years, FOSS development has transitioned from an organic community of
volunteers, towards an organized system with participation from industries,
non-profit organizations and government institutions. This could be regarded as
the second-age of FOSS movement~\cite{fitzgerald_transformation_2006}.

%todo: DONE
% Here, we have to add a paragraph on free and open-source (libre)
% software:

% The following facts can be mentioned:

% - tex (started in 1977), latex (1985)

% - GNU (started 1983), FSF in 1985, licence and philosophy...
% first complete system (with the Linux kernel) in 1991

% - 90s arrival of internet (soon mass market)

% Some huge open-source successes : Apache ("first killer app of Linux"),
% now Firefox, Open-office
% av: https://www.reddit.com/r/AskReddit/comments/7x639l/what_free_software_is_so_good_you_cant_believe/du6pw11/

% Linux kernel now widely used on servers, personal computers, embedded
% devices, smarth phone (with Android).
% https://en.wikipedia.org/wiki/Linux_kernel

% Now, new period for open-source: 
% see http://www.cepis.org/upgrade/files/full-2005-III.pdf

% - "Libre Software Movement: The Next Evolution of The IT Production
% Organization?"

% - ``the composition of development teams was changing, from all-volunteer
% teams to teams with paid participants from industry, government or
% not-for-profit organizations.''

% By the way, we also have to use the term "libre software" (?)
% av: Libre-software is not so mainstream. Even Linux kernel is not libre with
% binary blobs for drivers.

% We may have to change the order of this list?
% - git and mercurial 2005
% - software repositories ~ 10 years before
% so I would exchange the too paragraph...
% av:That is probably because we did not mention patches and CVS, subversion
% etc.

\paragraph{Distributed Version Control Systems (DVCS) and web-based source
development platforms.}  Collaboration was achieved in the early years of FOSS
development through emailing ``patches'', and centralized VCS near the turn of
the 21\textsuperscript{st} century.  The collective collaboration on
development is today streamlined with the emergence of DVCS (especially, Git
and Mercurial) and web-based source development platforms like GitHub,
Bitbucket and more recently GitLab.
%
These tools are designed to work on any kind of text files, not just codes, so
they could even be utilized to write scientific documents with multiple authors.
%
These tools are so useful for work, that using DVCS in science shall reap
benefits in terms of productivity in the long run~\cite{wilson_best_2014}.

Git has become the \textit{de facto} standard for source management. However,
for the FluidDyn project, we prefer to use Mercurial because we consider it to
be simpler for the beginners and just as powerful for expert users\footnote{See
\href{https://www.mercurial-scm.org/wiki/GitConcepts}{``a detailed
comparison''} on Mercurial website.}. Simplicity and lowering the learning
barrier for scientists who are not experts in software development is crucial
for our project.
%
Note that, it is now possible to use Mercurial to work with Git repositories,
for example hosted in GitHub and GitLab\footnote{By using the Mercurial
extension \href{http://hg-git.github.io/}{hg-git}.}.

\paragraph{Package management and third-party software repositories.} Python
as a medium for scientific computing owes its success to a rich ecosystem of
third-party packages.  Smooth deployment of these packages requires a package
manager and a supported repository. A credible hub to upload source codes as
compressed archives or as pre-compiled binary packages was therefore
indispensable. The use of repositories started with the conception of CPAN and
CRAN repositories for the languages Perl and R in 1993 --- based on the
Comprehensive TeX Archive Network (CTAN) model for \TeX\ packages.

Today, \href{https://pypi.org}{Python Package Index (PyPI)} and
\href{https://anaconda.org}{Anaconda Cloud} are major repositories for
delivering and downloading Python packages, assisted by package managers
\pack{pip} and \pack{conda} respectively.
% Almost all major 
% languages\footnote{\href{https://stackoverflow.com/questions/1693529/%
% list-of-top-repositories-by-programming-language}{https://stackoverflow.com/%
% questions/1693529/list-of-top-repositories-by-programming-language}}
% have similar repositories, with exceptions such as Fortran and Matlab.
Specialized package managers have also emerged to install scientific
libraries and compilers such as \pack{easybuild} and \pack{spack}.

\paragraph{Continuous integration (CI).} The need for strong reliability of
programs with increasing complexity has forced developers to use continuous
integration.
%
In CI, firstly the code is \emph{built} against a known, repeatable emulation
environment on a server. Thereafter, a set of \emph{unit tests} are run for each
commit (modification of the code) to check that no bugs have been introduced. It
is also important to analyse what portion of the source code is covered by these
unit tests --- a method referred to as \emph{code coverage}. In FluidDyn project
we stop at this level, but it is possible to further enhance it by adding code
style checks by running linters alongside unit tests; and also to add
\emph{continuous delivery} of documentation and package releases.

Almost all CI platforms are free for open-source projects --- Travis CI,
CircleCI and Appveyor % (sic) av: I have double checked the official names
to name a few. Bitbucket and GitLab have built-in CI
implementations.  Websites such as Codecov and Coveralls help to chart and
analyse the code coverage history, viewable by anyone.

When the code depends on a wide variety of external dependencies, CI is useful
to avoid \href{https://en.wikipedia.org/wiki/Software\_rot}{software rot},
i.e.\ to avoid becoming obsolete with newer versions of the compiler or its
dependencies.  CI is specifically important for dynamical
languages and open-source code since it allows open-source projects to welcome
participation from many people while maintaining a high degree of reliability.

\paragraph{Websites to share knowledge.} Mutual aid and knowledge sharing
have been important aspects of the web since its inception and these practices
are now of foremost importance for many human activities.
%
In particular, learning programming and computer science is now greatly aided
by websites like the community driven encyclopedia
\href{https://www.wikipedia.org/}{Wikipedia} or the community driven forum
\href{https://stackoverflow.com/}{Stack Overflow}, where people write thousands
of questions and answers on programming each week.
% todo: remark from Julien Salort: it's old.
% av: So are technologies like git and hg, but we are putting together what
% characterizes our workflow, aren't we? We can emphasize that these forums
% often work only when there is a "thriving community" like the Python
% userbase.
IRC protocol has continued to sustain the test of time. Instant messaging
channels such as \codeinline{\#python} on Freenode facilitate interactive
discussions and a form of community-level support system --- a role which used
to be filled by mailing lists alone in the past. Alternatives such as Riot
(Matrix protocol), Gitter, and Slack have gained traction offering different
integrations solutions on top of instant messaging.
%
Through instant messaging a geographically separated team, as is the case with
FluidDyn project, can achieve near real-time collaboration without the hassle
of setting up video conferences or meetings.

\paragraph{Automatic web documentations built on servers.}  Documentation
generators such as Sphinx (for Python) and Doxygen (for many languages
including C, C++, Fortran and Python) ``automagically'' convert in-line
comments into documentations as PDF, HTML and more. Read the Docs has in the
recent years become the main host for building and publishing documentation on
servers as web pages, thus easing the process for developers and considerably
improving access to information.

% We need transitions and some words at the end of the subsections to help the
% readers to understand why we present what we present and what is the link
% with the FluidDyn project...

We have presented new methods and tools used in software development.  Taken
all together, they greatly improve our efficiency for collective work and open
possibilities of interactions that were unthinkable only few years ago.
%
For example, the success of community-driven software like \pack{astropy} (a
subject-area research library with \href{https://github.com/astropy/astropy}{more
than 240 contributors}) and \pack{scikit-learn} (a machine learning toolkit with
\href{https://github.com/scikit-learn/scikit-learn}{more than 1000 contributors})
would not have been possible without these new software development methods.
%
% pa: done: we need to mention in this paragraph existing community-driven
% subject-area research libraries (see the file
% ./communitydriven_research_soft.md).
%
As a scientists, we can wonder how to fully exploit these new possibilities to do
science.  The project FluidDyn is an attempt to help the fluid dynamics community
to advance in this direction.

\subsection*{Python, a programming language adapted for open-science}

We now turn to the presentation of the
\href{https://www.python.org/}{programming language Python}, another
fundamental tool for the FluidDyn project.  Note that, this is not a detailed
technical presentation: we focus on the ideas necessary to understand why the
FluidDyn project is based on this language and on facts interesting for
potential FluidDyn users and developers.

Python community has thrived remarkably in the last few years, evident from the
fact that by the end of 2017, Python became the second most popular language in
terms of pull-requests in GitHub\footnote{See GitHub's annual report titled
\href{https://octoverse.github.com/}{ ``The State of the Octoverse 2017''}.} and
questions on StackOverflow\footnote{See
\href{https://stackoverflow.blog/2017/09/06/incredible-growth-python/}{ ``The
Incredible Growth of Python''}, a blog post by Stack Overflow.}.
%
TIOBE-index\footnote{\url{https://www.tiobe.com/tiobe-index}}, which assess the
popularity of a language based on number of skilled users, courses and
third-party vendors, ranks Python at number four. IEEE Spectrum\footnote{See
the analysis titled \href{https://spectrum.ieee.org/computing/software/%
the-2017-top-programming-languages}{``The 2017 Top Programming Languages''} on
IEEE spectrum.} ranks Python as the most popular language in 2017 with a metric
based on trends, demand by employers and users on open-source hubs.  All the
rankings do agree that Python usage has been steadily growing and is here to
stay.  Python largely outclasses languages like Fortran and Matlab according to
these statistics, and is now clearly one of the mainstream languages with C,
Java, C++ and Javascript.
%
Let us summarize some characteristics of the Python languages that have lead to
such incredible success.

% todo: DONE
% We need an introductory paragraph on how big is Python today
% + the fact that several scientists miss this fact.

% See the file ./python-one-main-languages.rst

The Python language has been designed to boost the communication of technical
ideas between humans. The resulting code is as simple as possible due to the
inherent emphasis on readability. Humans can focus on the ideas. Python differs
from other mainstream languages for a variety of reasons:

\begin{itemize}
\item \textbf{Aesthetics.} The syntax is nice and elegant. Blocks of code are
defined with the indentation.
\item \textbf{Code style.} There are explicit guidelines
(\href{https://www.python.org/dev/peps/pep-0008/}{PEP~8}) supporting the
regularity (and thereby readability) of the code and tools
(linters) to evaluate adherence to the style.
\item \textbf{Dynamic typing.} The variables (the ``names'' in the code) are
not attached for life to an object in memory\footnote{Opposite behavior
compared to static languages as Fortran, C and C++.}. The types of the objects
are inferred at run time from the code and the context, and in many situations,
one does not need to declare them explicitly.
%
% Being a strongly typed language\footnote{In contrast to weakly typed languages,
% such as JavaScript}, types of the objects are respected and implicit type
% conversions are rarely performed.
% av: "rarely / never"?
% eg: 'hello' * 2 -> 'hellohello'; 1/3 -> 0.3333333
%
% pa: I think we can not cover such complicate notion in this document... It is
% anyway not so important for our demonstration.
\item \textbf{Automatic memory management.}  The user need not handle the
memory allocation and deallocation by hand.  There are, of course, ways to
control the memory usage of a program, though not as precisely as for low-level
languages.
\item \textbf{Interpreted.} The standard way\footnote{There are also tools to
compile Python code (for example Cython and Pythran).} to run a Python code is
to ``interpret'' it.  A program called the interpreter executes the code
nearly\footnote{Actually, the execution of Python code is less simple and
involves an Abstract Syntax Tree.}  instruction-by-instruction, with very few
optimizations.
%
In contrast to other compiled languages such as Fortran, C or C++, there is no
proper compilation step: the code is not translated to optimized machine
instructions, but only to Python bytecode.
%
We will discuss shortly the disadvantages of this characteristic. Two evident
advantages are rapid development cycle and the possibility of interactive
workflow (usually with IPython or Jupyter notebooks).

Note that there are many implementations of Python interpreters\footnote{We can
cite CPython (written in C), Jython (Java), IronPython (C\#), PyPy (Rpython, a
subset of Python) and MicroPython/CircuitPython (C, targeted to
micro-controllers).}. The default and most widely used implementation is written
in C and is called CPython.

\item \textbf{Simple, yet powerful.}

\begin{itemize}
\item Python supports multiple programming paradigms, including imperative,
object-oriented and functional.

\item It is easy to interface with code written in other languages (in particular
C, C++ and Fortran).

\item Python comes with a large and high quality \href{https://docs.python.org%
/3/library/index.html}{standard library}.

\item Python can be deployed on many different machines with different operating
systems (Linux, Windows, macOS, Android) and architectures (from a microcontroller
--- with MicroPython or CircuitPython --- to a Blue Gene supercomputer).

\end{itemize}
\end{itemize}

Thus, Python is one of the best ``glue'' languages for fast prototyping.
Developing with Python is much faster and easier than with many other
languages.  Less bugs are introduced by the developers simply because there are
much less lines of code \cite{nanz_comparative_2015}.
%
Maintaining and modifying existing code is also eased by the very good code
readability associated with the language.

Another consequence of the apparent simplicity of Python is that the learning
curve has a nice shape. Learning Python is very easy at the beginning and the
complexity gradually increases as the level of the developer increases.
%
Therefore, Python is good for developers of all levels: very gentle for
beginners and very powerful for advanced users.

Python is an old language (first implementation in 1991!) but it continues to
evolve.
\begin{itemize}
\item A hard and controversial decision was taken to clean up Python from its
historical incoherences by having a new series of backwards-incompatible
versions of the language, Python 3.  Python 3.0 was released in December 2008.
After a long and difficult process, the transition from Python 2 to Python 3 is
now (2018) behind us for nearly all important packages\footnote{Close to
\href{https://python3wos.appspot.com/}{95\% of the most downloaded packages}
are Python 3 compatible now.}.  Especially for scientific applications, we can
now work only in Python 3. For instance, \href{https://github.com/%
numpy/numpy/blob/master/doc/neps/dropping-python2.7-proposal.rst}{future
releases of \Numpy} and of \href{http://www.python3statement.org/}{the main
scientific packages} will soon drop compatibility with Python 2. These changes
open doors to very clean, coherent and potentially faster Python
experience. Therefore, one should not use Python 2 for science anymore and
instead, adopt the newer versions ($>=3.5$ in 2018).

\item New versions of Python have nice new features to adapt to new usages and to
recent trends in computer science.  For example, three new features were
introduced in Python 3.5 (first released on September 2015): the \codeinline{@}
operator for matrix multiplication, the new \codeinline{async} and
\codeinline{await} keywords for concurrency and type hinting\footnote{See also our
paragraph on the issue of type checking in the following.} (with the module
\codeinline{typing} and an associated syntax presented in the
\href{https://www.python.org/dev/peps/pep-0484/}{PEP~484}).
\end{itemize}

\subsubsection*{An incredible success}

Being a very versatile language, Python is widely used for many different
applications:
\begin{itemize}
\item Simple scripting.

\item System, database and network administration.

\item GNU/Linux distribution software.

\item Python is now one of the main scripting languages to add programmability
to applications (for example Paraview, Visit, QGIS, Blender, ...).

\item Web servers.

\item Web scraping and data analysis.

\item Animation movies, game development and gaming.

\item Education. Python has recently become one of the most popular language
for teaching programming and computer science.

\item Science!
Python is now widely used in scientific applications. It has now a mature and
powerful scientific ecosystem with well-established based packages (\Numpy for
N-dimensional homogeneous arrays, \Scipy as the fundamental toolkit for
scientific computing, \pack{Matplotlib} for plotting and \pack{Pandas} for data
structures) and several more specialized packages (to name a few, \pack{h5py},
\pack{mpi4py}, \pack{skimage}, \pack{sklearn}, ...). There are now great (and
often shiny) tools for most of the applications. For example, the
\href{https://github.com/rougier/python-visualization-landscape}{%
landscape of visualization tools usable through Python} is impressive.
%
Python is one of the main languages for data science with packages such as
\pack{Pandas}, \pack{statmodels}, \pack{sklearn}, \pack{Keras}, \pack{TensorFlow}
and \pack{PyTorch}.

There are now ready-to-use ``batteries-included'' Python installers
(similar to Matlab), like in particular the
\href{https://en.wikipedia.org/wiki/Anaconda_(Python_distribution)}{%
freemium open-source distribution Anaconda}, so it is very easy to start using
Python for scientific purposes.

There are now nice Python integrated development environments (IDE) adapted for
scientists, for example \href{https://github.com/spyder-ide/spyder}{Spyder},
\href{http://www.pyzo.org/}{Pyzo} and
\href{https://jupyterlab.readthedocs.io}{Jupyterlab}.

\end{itemize}

Python has a very large and supportive community (see for example
the \href{http://stackoverflow.com/tags}{Stack Overflow tags}).
%
But Python is also supported by companies using Python and the open-source
dynamics.  Most of these companies that support Python are specialized in
web technologies and data-driven research.  But, recently, companies basing
their business
% av: earlier "computer and web technologies". Computer is quite obvious, isn't
% it?
on Python for science have started to become quite powerful in terms of
development. For example Anaconda Inc.\ is strongly implied in the development of
tools like \pack{Numba} and \pack{dask}. Hardware giants such as Intel and Nvidia
also contribute to the Python scientific ecosystem.

\subsubsection*{Some Python issues}

\paragraph{Since the standard Python interpreter does not do any proper
compilation,} Python code in some CPU bounded cases can be too slow.  Thus, one
has to use special tools to get optimized machine instructions to speedup
performance-critical code.

\textbf{Ahead Of Time compilation.}  Since Python is primarily a ``glue''
language, a common solution is to use code written in C, C++ or Fortran.
%
% todo: DONE
% add a reference to ctypes and cffi to call C with pure Python code
%
It is possible, without a compilation step, to call functions in C libraries
through pure python code using \pack{ctypes} standard module or the \pack{cffi}
package.
%
Libraries with Python bindings can also be used directly (for e.g., wrapped
using SWIG or Boost.Python) or one has to write extensions, i.e.\ Python
modules written in C or C++ using the CPython API that can be imported directly
from Python like any Python module.
%
However, to rely on such techniques to call C libraries would inhibit
development of extending functionalities, as they have a steeper learning
curve compared to Python.
%
Nevertheless, this strategy has given rise to the base modules of the scientific
Python stack, namely \Numpy, \Scipy and \pack{Matplotlib}.

Python developers do not actually have to write the compiled extensions in C,
C++ or Fortran, since there are tools to generate them automatically from
Python code (Pythran) or from Cython code. Cython \citep{behnel_cython2011} is a
smooth blend of two languages with possibility to add type declarations like in
C/C++, but with a syntax similar to Python.

Pythran \citep{guelton2015pythran, guelton2018pythran} is a quite recent Python
compiler which gives impressive results. It creates compiled extensions from pure
Python code with simple type annotations written as comments.
%
The resulting extensions are usually as fast as Fortran or C++ written by
non-specialists \cite[see, for example, benchmarks for the packages
\fluidpack{fft} and \fluidpack{sim},][]{fluidfft, fluidsim}.
%
They are created with a very interesting two-step compilations: first the code
is optimized at the Python level and then an automatically produced C++ code is
properly compiled.
%
This two-step compilation provides room for very clever optimizations as
demonstrated, for example, in
\href{http://serge-sans-paille.github.io/pythran-stories/%
being-more-than-a-translator.html}{%
this blog post by the main developer of Pythran}.
%
Pythran supports OpenMP pragma \citep{guelton2013compiling} and can use modern
vectorization with SIMD instructions \citep{guelton2014exploring}. Pythran
understands both Matlab-like vectorized code and C-like code with explicit loops
\citep{guelton2018pythran}.

We present in the two companion papers \citep{fluidfft, fluidsim} examples of
highly efficient Python codes, optimized using a combination of Cython and
Pythran extensions.

\textbf{Just In Time (JIT) compilation.} Another way to obtain optimized
machine instructions is by using JIT compilation, i.e.\ by compiling only the
critical code at run time.
%
This strategy can yield good results with other languages as for example Matlab
or Julia.
%
Pypy, an alternative interpreter written in Python, has a JIT compiler. However,
Pypy is not widely used for scientific applications mainly because of
compatibility problems with the extensions written for CPython.
%
However this could change since a recent version of Pypy (6.0) supports the main
packages of the Python scientific stack (\Numpy, \Scipy, \pack{Matplotlib},
\pack{Pandas}, etc.).

Adding a JIT to the interpreter CPython has been notoriously difficult (see
\href{https://faster-cpython.readthedocs.io/}{faster-cpython}). One of the
problem seems to be the C API provided in CPython and used by many extensions.
%
Python being first a glue language, a faster CPython with a JIT is actually not
so necessary for many purposes. Therefore, not as much money and work has been
put in accelerating Python than for example for Java and JavaScript.

Another strategy is to add a JIT to CPython through an external package
(\pack{Numba}) and to only compile the critical functions.  Moreover,
\pack{Numba} is particularly interesting because it can take advantage of the
GPU\footnote{See for example
\url{https://devblogs.nvidia.com/parallelforall/seven-things-numba/}}.

\paragraph{No type-checking.} In contrast to languages using Ahead Of Time
compilation, there is no type-checking in Python (and more generally in many
dynamical languages).  Type-checking is useful since it can avoid simple bugs.
However, it is not a good practice even with static languages to rely only on
type-checking to look for bugs in a code.
%
It is known that a rigorous set of unit tests is also needed to obtain more
reliable codes.
%
Dynamical languages rely a lot on unit tests so it is very important to put at
least a little bit of time and energy to write a decent battery of tests.
%
We have to mention the young and promising project
\href{http://mypy-lang.org/}{Mypy}\footnote{Guido van Rossum, the author and
``Benevolent Dictator For Life'' of Python, is involved in the project Mypy.},
which allows for Optional Static Typing and thus type-checking for
Python\footnote{We plan to add type hitting in the most important FluidDyn modules
to investigate how we can take advantage of type-checking with Mypy.}.

\paragraph{A lively, huge and thus complicated ecosystem.} As already
mentioned, Python is a versatile language. For many applications, the standard
library is insufficient and one needs external packages.
%
% For most applications there are several projects usable through Python.
Most of the time there are several Python packages for a certain application.
A user/developer has to make choices regarding the tools to use.  Especially
for the beginners who are not used to the open-source realm, it can be
difficult to understand a ``landscape of projects'' for a particular
application and to make good technological choices.

It is important to get a good introduction on how to work with the open-source
dynamics and the scientific Python ecosystem.  It can also be very useful to ask
to more advanced users and attend Python courses or workshops\footnote{This could
be regular academic courses, MOOCs or workshops like ones organized by
\href{https://software-carpentry.org/}{Software Carpentry}.}.
%
Of course, such abundance of projects can also be seen as an advantage.

\paragraph{Absence of multicore computational parallelism using threads.}
Threads behave like light subprograms. In principle, they can use the different
cores of the CPU at the same time but the CPython interpreter forbids that they
interpret python code at the same time\footnote{See
\href{https://opensource.com/article/17/4/grok-gil}{this blog article on ``How
to grok the GIL''} and
\href{https://faster-cpython.readthedocs.io/cpython37.html}{this technical
presentation by a CPython core developer}.}.
%
This limitation comes from an important detail in the implementation of the
CPython interpreter: a Global Interpreter Lock (GIL) to prevent race conditions
that could corrupt data.
%
It seems that the GIL greatly simplies the implementation of CPython and that it
is very difficult to remove it while keeping other nice technical properties of
CPython intact\footnote{See \url{https://wiki.python.org/moin/GlobalInterpreterLock}.}.
Note that Pypy also uses a GIL, while two other implementations of the Python
language, Jython and IronPython, do not have this limitation.

At any rate, threads using the interpreter do not use the CPU at the same time
so it is not possible to use threads to do multicore computational parallelism
with pure python code to accelerate CPU bounded tasks.
%
Note however, that threads can be (and are widely) used in Python for
concurrency, i.e.\ to perform I/O tasks (which does not need CPU) concurrently.
%
To do proper computational parallelism for CPU bounded tasks with the GIL, one has
to use other strategies: extensions (fine grain parallelism), multiprocessing
(coarse grain parallelism) and inter-process communication, for example, with MPI
(with mpi4py) or ZeroMQ.

\subsubsection*{Rich landscape of open-source languages useful for science}

So far, we have discussed in detail about Python. However, there are many
interesting open-source languages that can be used for scientific applications.
All have strengths and weaknesses.

Fortran remains widely used for computing codes, at least in fluid mechanics.
C++ is used in many recent developments. Its recent improvements (C++11, C++14)
make it a modern and very useful language.
%
JavaScript, Java, Scala, Smalltalk, Haskell, R, Julia, Perl and Lua are used for
some scientific codes. Go and Rust are quite young languages and are not (yet)
widely used for scientific applications.

It is of course very difficult to predict which technologies will be widely
used in even the near future (10 or 20 years for example).
%
From what we see, it seems that the idea of ``one language to do everything for
science'' will not succeed, at least soon.  This was maybe the project of Julia
creators, but despite the qualities of this language, it also has
limitations\footnote{For example C-like include instead of Python-like
import. From our point of view, a Pythonic flavor of Julia would be very
interesting.}.
%
It seems that a very important aspect in the future will be interoperability
between coexisting tools\footnote{See for example the cross-language development
platform \href{https://arrow.apache.org/}{Apache Arrow}.}.

So what about Python?  As already mentioned, it is today the language with the
strongest dynamics in science, data analysis and much more.  Several students
learn and will learn it. Many scientists in the future would, hopefully, be
able to appreciate these strengths and favour the Pythonic approaches.  With
such dynamics today, we can anticipate that the Python world will be able to
embrace new technological trends and to overcome some of its current
limitations.
%
It is reasonable to predict that Python will continue to shine as a language
(i) particularly good for human communication and to write ideas, (ii) for fast
prototyping and (iii) for interoperability with code written in other languages
(see for example the recent projects \href{https://cppyy.readthedocs.io}{cppyy}
and \href{https://github.com/pybind/pybind11/}{pybind11}, or
\href{https://github.com/PyO3}{the possibility to interact with Rust code}).

\subsubsection*{Summary: Python, a versatile Swiss Army Knife for the
scientists}

Python has become a great versatile tool for scientists.  The scientific
ecosystem of Python is nowadays really impressive.  The scientific Python
community is strong and quickly growing.
%
A scientist with good skills in Python can do most of what she/he needs to do in
programming (which does not mean that Python is the best language to do
anything!).
%
In contrast, no skill in Python is a real disadvantage for many tasks and for
employability.
%
Therefore, in most fields today, if a scientist has to acquire good skills in
one language, it is in Python.

Considering what we have presented on Python, we think that it is reasonable to
bet that it will become one of the \textit{lingua franca} in science and in
fluid dynamics.  This explains why we have chosen it to be the main language
for the project FluidDyn.

\subsection*{Productivity at individual, group and community levels}

% remark Julien Salort "It is well known that": bad
% pa: common sense says ?
% It is well known 
It is not surprising
% av: "It is not surprising"
%
that short-term efficiency and long-term efficiency are
sometimes incompatible.
%
Quick and dirty scripts can be efficient in the short term, but such a workflow
has a negative impact on a longer time scale.

Similarly, we can also differentiate productivity at different scales of
collaboration.
%
An individual can be very efficient with a particular tool, but the same tool
can be very inefficient to collaborate even with co-workers.
%
A group can be very efficient with a code which is closed-source.  If this
strategy can be efficient for the group (at least in the short term), it leads
to waste for the community. Other groups will need to develop codes with the
same features.  Good ideas will be spread in different codes and since the
number of users and developers are smaller for each of these codes, the code
quality will not improve as fast as it would with an open-source strategy. In a
nutshell, as this proverb says, \textit{``If you want to go fast, go alone. But
if you want to go far, go together.''}

Thus, a community using open-source methods can be efficient to co-develop its
tools. However, some of its members have to spend more energy to bear this
dynamics, which implies that they somehow sacrifice their individual efficiency.
We see that short-term efficiency, long-term efficiency and productivity at
different levels can often be incompatible and that we have to consider these
conflicts when choosing between different technologies.

\subsection*{Programming in the field of fluid mechanics}

We study fluid mechanics via laboratory experiments, in situ measurements,
analytic computation, numerical simulations and data processing.
%
Nowadays, programming is involved in all methods, and thus it is difficult to do
fluid mechanics without software development.
%
However, the level in software engineering is on average very low in the
community.
%
Even today, it is quite commonplace to start a PhD without any serious training in
GNU/Linux and modern programming tools.
%
A large majority of scientists and technicians are also unaware of the new
challenges and opportunities of open-source.

Since fluid dynamics is an engineering science, closed-source commercial software
have historically been dominant in the field.
%
Few people are aware of the vicious circle of the closed-source model for people
and groups relying on it.
%
This mechanism is nevertheless very damaging. In the closed-source model, a group
pays for a license or a new development. The group does not learn how to develop
what has been paid for. The company, on contrary, stands to gain from profits and
the feedback generated from users, through which it improves the product sold.
The group produces codes, books, courses using the closed-source product or
acquire knowledge on how to use the closed-source product, so that the group is
increasingly dependent on the product and is ready to pay more for it.
%
Such vicious circles are difficult to break, but not impossible. This can be
done by introducing alternative open-source solutions, bringing in positivity to
this circle.

We have seen that it was very common to rely on commercial companies and
closed-source software for most of the new developments in fluid mechanics.
%
Nevertheless there are also a lot of codes produced by scientists in the
field. But without the technical knowledge on how to work collectively on code,
this production is often of bad quality (compared to the information technology
industry standard) and it is basically doomed to be abandoned after its use.
%
However, it is natural to try to reuse code or at least to build the next steps
on what has been done before.  Thousands of hours of highly qualified people
are spent in trying to understand and reuse codes badly written with
inappropriate languages!

On a technical point of view, the majority of the coding in the field involve a
mix of Fortran/C or C++, shell languages (as Bash) and Matlab. For experiments,
the graphical programming environment
\href{http://www.ni.com/en-us/shop/labview.html}{Labview} is dominant for
control of physical objects and acquisition and Matlab is used for data
processing.
%
Often, languages are used for things for which they are not adapted.
%
Fortran, C or C++ have not been designed for fast prototyping of complex
programs.
%
It is known that for scientific purposes, one should restrict shell scripting
to extremely simple tasks.
%
Similarly, using Matlab for developing complex programs is technically a
questionable choice.
%
These technical solutions also represent a problem for code reuse, sharing and
collaborative development.
%
We are now going to present factual arguments explaining why we think that the
massive usage of Matlab decreases the collective efficiency of the community.

\subsubsection*{Discussion on Matlab, Labview and Mathematica}

% see http://www.pyzo.org/python_vs_matlab.html
% see https://www.linkedin.com/pulse/matlab-vs-python-jan-rhebergen

Matlab is a closed-source proprietary numerical computing environment.
%
It is a good tool for simple processing with matrices, image processing, data
visualization and certain niche applications. The language is well adapted for
these tasks. The development environment is nice and the interpreter is quite
fast, especially now that it has a JIT compiler.

However, the language suffers from serious technical issues which make Matlab
an inadequate tool
% \citationneeded[opinion] % pa: I think what we say is fine for an article. We
% are not writing a wikipedia page which has the ambition to be neutral. Here,
% we gives an argued opinion, which is totally normal in a scientific article.
% av: fair enough :)
for doing more than simple processing and data plotting. The comparison with
Python often debated upon. We list a few striking Matlab weaknesses:

\begin{itemize}
\item One file for each function (or class) available outside the file where it
is implemented. No notion of package or module.

\item No real organization of the standard library. All built-in and
user-defined functions are available in a huge flat namespace. No import
mechanism and as a result we cannot deduce, simply by looking at the code,
where a function comes from.

\item A standard way to organize multi-file code is to write scripts that
modify and define global variables (see for example
\href{https://www.damtp.cam.ac.uk/user/jrt51/files/diablo\_mat.tar.gz}{%
this Matlab version of the code Diablo}).  It is so simple to do this that we
can assert the language strongly encourages this practice.  Matlab files are
not self consistent, i.e.\ it is normal to use in a file a global variable
defined outside of the file.

\item Very bad default argument mechanism.

Default arguments for a function is a very common feature is many programming
languages. In Python, we can write:
\begin{minted}[fontsize=\footnotesize]{python}
def myfunc(a, b, c=1, has_to_print=True):
    if has_to_print:
        print('a =', a, 'b =', b, 'c =', c)
    return c * (a + b)
\end{minted}

An implementation for approximately the same behavior in Matlab could be (as
advised in the official Matlab documentation):
\begin{minted}[fontsize=\footnotesize]{matlab}
function ret = myfunc(a, b, varargin)
    if nargin < 2 | nargin > 4
        error(['The number of arguments has to be ' ...
               'greater than 2 and lower than 5'])
    end
    if nargin == 4
        has_to_print = varargin{2};
    else
        has_to_print = 1;
    end
    if nargin >= 3
        c = varargin{1};
    else
        c = 1;
    end

    if has_to_print
        disp(['a = ' num2str(a) '; b = ' num2str(b) '; c = ' num2str(c)])
    end
    ret = c * (a + b);
end
\end{minted}
Even for a very simple function, there are many places where bugs can be
introduced and the code is much less readable than the corresponding Python
code.

\item Less intuitive syntax for string operations in Matlab, compared to
Python. Take for example for string comparison, \codeinline{strcmp(s1, s2)} in
Matlab versus \codeinline{s1 == s2} in Python or for look up,
\codeinline{contains(s1, pattern)} in Matlab versus \codeinline{pattern in s1}
in Python.
% \item Less intuitive and clear syntax for exception
% handling. \citationneeded[opinionated].
% pa: Matlab has now a MException class. The syntax is not nice, but it is
% not so important.
\item Less clear and intuitive model and syntax for object oriented
programming, compared to Python. As for functions, one file per user-defined
class is necessary.
% \citationneeded[opinionated].
\item Parenthesis used for both function calls and indexing, causing ambiguity.
\item Matlab codes are usually full of \codeinline{;}, \codeinline{.*},
\codeinline{./}, \codeinline{\&\&} and \codeinline{||}, which make them quite
``noisy''.  Moreover, there is nothing like the Python PEP~8. This makes a
large proportion of the Matlab codes difficult to read and understand, due to
lack of a consistent code style.
\item The syntax \codeinline{a(100, 100) = 1;} to create and extend matrix is a
very good way to hide bugs. The code \codeinline{a = eye(2); a(i0, i1) = 1;}
will never raise any error regardless the values of \codeinline{i0} and
\codeinline{i1}! No error is raised if a user misspell the variable
\codeinline{a} and write something like \codeinline{aa(100, 100) = 1;}.
\end{itemize}

Now we turn our attention, from structural issues of Matlab syntax, to more
practical constraints its use brings about.  By definition of a closed-source
software, Matlab is a big black box: it is impossible to study the
implementation of its functions.

Matlab is not free. The price for one license for non-commercial use is not
very expensive for most research and teaching institutes. However, when we
start to add up the price of the toolboxes necessary to run most codes, it
starts to become substantially large for some institutions.  The price of
Matlab quickly escalates when deployed on a cluster with MPI, since a licence
is required per user, per node!
% av: not per process.
% see: https://www.nas.nasa.gov/hecc/support/kb/running-matlab-applications-in-parallel_522.html

It would be interesting to estimate what could be done with open-source
solutions with a part of the amount paid by public research institutes. How
many developers could be paid to improve the alternative open-source solutions?
%
This question raises the issue of funding scientific open-source software.

Moreover, it is well known that the lack of license for the personal computers
of staff is sometimes solved with cracked versions of Matlab. How can serious
research projects rely on such illegal methods?
%
Note that open-source Matlab interpreters do exist, such as Octave and Scilab.
%
% pa: warning! Octave has a JIT compiler!
%
% Unfortunately, these are not serious alternative today. Their user
% community is too small and they are not very efficient, especially now that
% Matlab uses a JIT compiler and that optimized Matlab codes rely on it.

The recent open-source and Python dynamics represent a hard punch for Matlab,
which is losing the dominant position it has in some domains.
%
Note that this is a very recent development. The Python scientific ecosystem can
now compete in nearly all domains, but this was not the case typically 5 years
ago.
%
There are still a large number of researchers and teachers in sciences who are
strongly attached to Matlab mainly because they have used it a lot.
%
It is, of course, difficult to change habits especially when those are efficient
(for a short time-span and on an individual or group scale) enough for what you
do.

It is normal that Mathworks, the company which makes money with Matlab, spreads
\href{https://www.mathworks.com/products/matlab/matlab-vs-python.html}{arguments
against Python and its scientific ecosystem}.  Some of them are valid, for
example the issue of the lively, huge and thus complicated scientific ecosystem
that we have already discussed.  Other arguments are unfair or no longer valid
nowadays.
% Here, we remove a long argument with Mathworks (see the file
% old_arguments_mathworks.tex)
% av: Less provocative, that is good. But we can still add a footnote linking
% the Pyzo page / LinkedIn page comparing python and matlab.
%
% We replace by this:
The scientific Python ecosystem is now very mature, robust, well documented and
fully integrated.  Many of this components are supported by professional
developers paid by companies or research institutes, for example \pack{Numba}
(Anaconda, Nvidia), \pack{TensorFlow} (Google), \pack{PyTorch} (Facebook, Twitter,
Nvidia), Scikit-learn (INRIA), Mercurial (Facebook).
%
The support by the Python community is both very reactive and of very good
level (for example in Stack Overflow, and issues page on major repositories)
and it is now possible to pay companies to get professional support on
scientific Python (for example
\href{https://software.intel.com/en-us/distribution-for-python/get-help}{Intel}
and \href{https://www.anaconda.com/enterprise/}{Anaconda}).
%
Finally, with the tools available today, with little effort we can get a very
efficient program using only Python code.

To conclude on Matlab, some people may use it efficiently for their work, but it
starts to be a problem for the collective efficiency of the community. Except
from the existence of a large code base difficult to translate in better
languages or for some very specific applications, it seems difficult today to
find a good technical reason to continue using Matlab.

The same remarks can be made for Labview. It is a graphical programming
language and the programs are saved as binary files. This is a problem because
first it is impossible to read a Labview program without Labview and, more
importantly, it forbids the use of source management tools which are so
important for collective work and open-source.

Matlab and Labview also lead to problems for students, who learn bad coding
habits and languages that are much less in demand by employers than for example
Python.

For symbolic computation, some specialists use Mathematica, which is a good and
very expensive proprietary closed-source mathematical symbolic computation
program.
%
Sage and Sympy are two complementary open-source alternative both based on
Python.

\subsubsection*{Different models for software development in fluid mechanics}

Proprietary codes tend to dominate the field of fluid mechanics. This is true
for industrial Computational Fluid Dynamics (CFD), for example with the Ansys
suite and also for acquisition and analysis of images of fluid with the
two companies Dantec and Lavision.
%
Even certain softwares developed by researchers in academic laboratories
also follow a closed-source model, for example
\href{http://www.damtp.cam.ac.uk/user/fdl/digiflow/index.htm}{Digiflow}.
%
Closed-source or even undistributed software is a widespread model in the
field.
%
The arguments against sharing are diverse:

% todo: DONE
% pa: I am really uncomfortable with this presentation which gives the
% impression that we think that the arguments against sharing are always wrong.
%
% We have to make it clear that it is not the case and that we do respect how
% other people think and work.
% av: I have commented out specific counter arguments, instead added two
% sentences after the bullets
\begin{itemize}
\item \textit{``We do not share to keep a comparative advantage.''}

%	Sharing code leads to reproducibility, making related
%	publications more attractive to readers. The authors gain
%	comparative advantage by earning citations.

\item \textit{``We do not share because we do not provide support.''}

%	Support and documentation is invaluable, of course. But sometimes, a code
%	with comments can be sufficient, which is a common practice.

\item \textit{``We do not share because people would not be able to
correctly use the code or interpret the results.''}

% pa: this one does not respond to the problem... (bad use of the program)

%	Poorly written code is infinitely more useful than no code at all.

\item \textit{``We do not share because we want to control industrial usage.''}

%	Proper choice of license can legally restrict and permit industrial
%	usage at the discretion of the developer.

\item \textit{``We do not share because we do not want people to review
and criticize our code.''}

%	People are, in general, kind and welcoming in the open-source community.

\item \textit{``We do not share, because in this way, people will think our
project has more value.''}

% pa: this one looks too "religious" to me :-) but ok...
%	True value is gained by furthering science and dissemination of ideas, not by
%	fabricating an appearance.
\end{itemize}

The flip-side involves releasing a well-written, documented, easy to use and
maintain source code with appropriate license to ensure credit where it is due.
This can potentially add value for the community through added transparency in
publications, and in return, value for the authors of software in terms of
recognition and citations. We have only looked at the most ideal scenarios and
the question of soundness of the above arguments is left as an open question to
the reader.

Between closed-source and open-source, is a grey area gathering different
practices.
%
It is quite common to share only for ``friends'' to control the dissemination
of the software.
%
This model can be institutionalized with an official group, for example with
the code \href{https://www.coria-cfd.fr/index.php/YALES2}{Yales2}.
%
% av: Institutionalized? I don't understand.
With such a model, the lack of a proper open-source license can naturally lead
to problems and is a limitation for users/developers.
%
Another common practice is to share without repository (for example
\href{http://www.damtp.cam.ac.uk/user/jrt51/files.html}{Diablo} or
\href{http://yakari.polytechnique.fr/people/deloncle/ns3d.html}{NS3D}) and/or
without license (\href{https://choosealicense.com/no-permission/}{meaning that the
code is actually under exclusive copyright}).

Finally, several scientists in the fluid dynamics community have used
open-source methods in their work for a long time.
%
There are a lot of open-source research codes written with proprietary tools
(typically written in Matlab, see for example
\href{http://pivlab.blogspot.de/}{PIVlab},
\href{http://servforge.legi.grenoble-inp.fr/projects/soft-uvmat/}{UVmat} and
\href{https://de.mathworks.com/matlabcentral/fileexchange/10902-pivmat-4-00}{PIVmat}).
%
Let us first say that these codes are a great wealth for the community at least
for three reasons: (i) they can be used as they are, (ii) people using these
codes can be more open to the open-source dynamics, and (iii) open-source
research projects using open-source tools can exploit the ideas in these codes.
%
Let us also point out a drawback of such a model. We do not pay the authors, but
people do need to pay to use the software. Even the authors of the software give
money to the same company to be able to use their software. Do they realize that
when they work on these programs, they work (often, for free) for the company that
sells the proprietary tool?

Some proper open-source codes have emerged, for example
\href{https://nek5000.mcs.anl.gov/}{NEK5000} (Fortran),
\href{https://www.openfoam.com/}{OpenFOAM} (C++),
\href{http://basilisk.fr}{Basilisk} (C) and
\href{http://channelflow.org/}{Channelflow} (C++) and
\href{https://www.code-saturne.org}{Code\_Saturn} (C/Fortran).
%
Big companies have started to use open-source development for fluid mechanics
application. For example, Volkswagen group (Volkswagen, Audi, Seat, Porsche,
Skoda, ...) has used OpenFOAM since
2006\footnote{\url{https://www.nas.nasa.gov/assets/pdf/ams/%
2014/AMS_20141202_Othmer.pdf}}.
%
EDF, the main French electric utility company has also a interesting
\href{http://linuxfr.org/news/strat\%C3\%A9gie-open-source-\%C3\%A0-edf-rd}{%
open-source strategy}.  It has produced many programs and in particular the CFD
solver \href{https://www.code-saturne.org}{Code\_Saturn}.
%
EDF is also involved in the development of
\href{https://www.paraview.org/}{Paraview}\footnote{\url{https://blog.kitware.com/introducing-some-new-paraview-5-0-features/}}.

Finally, scientists have started to use Python to study fluid mechanics,
especially through CFD codes (\href{http://dedalus-project.org/}{Dedalus},
\href{https://github.com/spectralDNS}{SpectralDNS},
\href{https://github.com/mikaem/Oasis}{Oasis}, \href{http://pyfr.org/}{PyFR}
and \href{https://fenicsproject.org/}{FEniCS}) or through data analysis
(\href{http://www.openptv.net/}{OpenPTV},
\href{https://github.com/jr7/pypiv}{PyPIV}).
%
The packages of the FluidDyn project are also part of this trend.

\subsection*{FluidDyn project: Breaking the status quo}

We have seen that there is a strong dynamics in play around the use of
computers (in particular with the web) and that this creates very efficient
tools and methods for collective work and software development.
%
The potential of these tools is not fully exploited yet, in the field of fluid
mechanics.

FluidDyn is a project to foster open-science and open-source coding in Python
in the field of fluid mechanics.
%
The project envisages to provide the technical framework to allow collaborative
development of tools useful for the fluid mechanics community, and to do better
science with open methods.
%
% It first started with one package.  Soon we had to prefer a decentralized
% organization with one common package and other specialized packages.
% av: mentioned in the next subsection
We provide examples of solutions for:
\begin{itemize}
\item Good coding practice with readable and easy to comprehend Python code
(PEP~8).
\item Source control management (Mercurial) and forge (Bitbucket) simple for
the new comers.
\item Packaging and installation procedure. All packages are available from
\href{https://pypi.org/}{Pypi} and can be install simply with
\href{https://pypi.org/project/pip/}{pip}, the standard Python installation
tool.
\item Licenses: depending on the packages, we choose to use the CeCILL-B or the
CeCILL licenses. These licenses are a BSD compatible and a GPL compatible
licenses adapted to both international and French legal
matters\footnote{\url{http://www.cecill.info/licences.en.html}}.
\item Documentation produced with standard and up-to-date tools: Sphinx,
Anaconda and Jupyter. Built and hosted online at
\href{https://readthedocs.org/}{Read the Docs}. Can also be generated offline.
\item Unittest and continuous integration with Bitbucket Pipelines and Travis.
\end{itemize}
We hope that such clean framework will allow contributions from scientists in
the field and that we can build together a nice, user-friendly and efficient
ecosystem specialized in research and teaching in fluid dynamics.

\section*{Implementation and architecture}

% \textcolor{blue}{How the software was implemented, with details of the
% architecture where relevant. Use of relevant diagrams is appropriate. Please
% also describe any variants and associated implementation differences.}

\subsection*{Organization of the code in packages}

FluidDyn was originally intended to be a single package to perform experiments
and simulations. Since a typical user may not be involved in both experiments
and simulations and also, due to increasing complexity as a virtue of rapid
development cycle made the need to decentralize FluidDyn evident.
Now, FluidDyn project hosts a number of specialized packages, namely:

\begin{itemize}
\item \fluidpack{dyn}: The base package which contains pure-python code that can
be reused in scripts or in specialized FluidDyn packages. It also contains codes
for miscellaneous command-line utilities useful for a typical fluid dynamics user.

The code of this package is presented in further detail in its documentation
(\url{https://fluiddyn.readthedocs.io}) and some prominent features are presented
in the following subsection.

\item \fluidpack{fft}~\cite[see the companion paper][]{fluidfft}: a package which
provides C++ and Python classes unifying various libaries to perform Fast Fourier
Transform (FFT) in sequential and in parallel.

\item \fluidpack{sim}~\cite[see the companion paper][]{fluidsim}: Numerically
oriented framework to run sequential and parallel Computational Fluid Dynamics
(CFD) simulations and on-the-fly post-processing for a variety of problems
(Navier-Stokes, Shallow Water, F\"oppl von K\'arm\'an equations, to name a few).
A study using \fluidpack{sim} has just been published in Physics of Fluids
\cite[]{LindborgMohanan2017}.

\item \fluidpack{lab}: Package to handle laboratory experiments. Primarily used to
communicate with various hardware devices such as motors and pumps, to handle I/O
between sensors, and to store data.
%
Experiments using \fluidpack{lab} have been carried out in the DAMTP fluid
laboratory \cite[Cambridge, UK. cf.][]{LeclercqPartridgeAugierDalzielKerswell2016}, in ENS Lyon laboratory
\cite[Lyon, France][]{salort2018} and at LEGI \cite[Grenoble, France][]{ISSF2016}.

\item \fluidpack{image}: Scalable image processing package which implements
various algorithms to calibrate cameras, to preprocess images, to do Particle
Image Velocimetry (PIV) and to postprocess data.
%
Fluidimage is used to process images taken during experiments performed in the
Coriolis platform at LEGI \cite[]{ISSF2016}.

\item \fluidpack{foam}: Small package to load OpenFoam data and plot them.

\item \fluidpack{coriolis}: Small package used to carry out experiments in the
\href{http://www.legi.grenoble-inp.fr/web/spip.php?article757}{Coriolis platform}
(a large rotating platform participating in the European consortiums Euhit and
Hydralab) and open the data obtained \cite[see, for example,][]{ISSF2016}. One of
the motivations behind creating this package is to study how to use open-source to
create and share open-data.

\end{itemize}

A detailed presentation on the above packages can be found in their respective
documentations on the web and for \fluidpack{fft} and \fluidpack{sim} in the
two companion papers \cite[]{fluidfft, fluidsim}.  This article will now focus
on the base package \fluiddyn.

All functions and classes defined in \fluiddyn are pure Python elements, meaning
that no extensions are implemented in \fluiddyn.  Thus the package \fluiddyn is
extremely easy to install with just a \codeinline{pip install} command and no
compilation.

The code base was designed to follow Python 2.7 syntax during its genesis. Now,
it has been made forward compatible with Python 3 through the use of external
package future.

\subsection*{API of the Python library \fluiddyn}

The package \fluiddyn is organized into five sub-packages: 
\begin{itemize}
\item \codeinline{fluiddyn.io}: This subpackage provides utilities for
input/output to different file formats.
\item \codeinline{fluiddyn.util}: Miscellaneous utilities.
\item \codeinline{fluiddyn.calcul}: Modules for numerical computations.
\item \codeinline{fluiddyn.clusters}: Classes to launch jobs on HPC clusters.
\item \codeinline{fluiddyn.output}: Utilities to produce scientific outputs
(figures and videos).
\end{itemize}

Sub-packages \codeinline{io}, \codeinline{util} and \codeinline{calcul} are the
largest in terms of lines of code and provides the Application Programming
Interfaces (API) to support \fluidpack{sim}, \fluidpack{lab} and
\fluidpack{image}.
%
For the sake of brevity, we shall only describe here some of the most important
modules.

\subsubsection*{Module \codeinline{fluiddyn.util.paramcontainer}}

Sub-package \codeinline{paramcontainer} defines class
\codeinline{ParamContainer} which is a hierarchical container for any type of
parameters. As shown in
\href{http://fluiddyn.readthedocs.io/en/latest/ipynb/tuto_paramscontainer.html}{%
this tutorial}, various strengths of an object of this class include:
\begin{itemize}
\item Support containing and printing documentation on the parameters.
\item Support printing to console as XML and saving as XML, hdf5 and netcdf4
files.
\item Evaluate data types automatically while loading from saved files.
\item Easy exploration in an interactive console.
\item Allows modification of default parameters through simple Python script
files. An error is raised if the user attempts to use an invalid parameter.
\item Graphical User Interface (GUI) frontend with PyQt.
\end{itemize}

Thus, it makes it a much more robust implementation for saving key parameters,
compared to conventional methods which rely on text or CSV files.

\subsubsection*{Module \codeinline{fluiddyn.util.serieofarrays}}

This modules provides classes to iterate over files. It is a common task in data
processing to understand how to organized a \textit{serie} of files from filenames
and formats. Then one has to form smaller sets of arrays contained in the
files. For example, we can have such \textit{serie}: \codeinline{im1\_1.png,
im1\_2.png, im1\_3.png, im2\_1.png, im2\_2.png, im2\_3.png} and we could create
subsets like \codeinline{((im1\_1.png, im1\_2.png, im1\_2.png), (im2\_1.png,
im2\_2.png, im2\_3.png))}. From the same \textit{serie} of images, we can also
create other subsets \codeinline{((im1\_1.png, im2\_1.png), (im1\_2.png,
im2\_2.png), (im1\_3.png, im2\_3.png))}.

The classes of this module allows one to do it with a quite simple and general
API as shown in
\href{http://fluiddyn.readthedocs.io/en/latest/ipynb/tuto_serieofarrays.html}{%
the tutorial on the module}.

\subsubsection*{Module \codeinline{fluiddyn.util.mpi}}

This simple module makes simulataneous sequential and MPI programming a breeze
by providing number of processes, \codeinline{nb\_proc = 1} and \codeinline{rank
= 0}, when used in sequential mode, otherwise providing the appropriate values
provided by \pack{mpi4py} package. Use of this module thwarts coding several if-else
clauses. If the program was using MPI, also defines the variable
\codeinline{comm} as an alias for the \codeinline{MPI.COMM\_WORLD} communicator.

\subsubsection*{Module \codeinline{fluiddyn.calcul.easyfft}}

Thin wrapper for an unified API using classes around the packages \pack{pyfftw}
and \pack{scipy.fftpack}.  It is very easy to perform forward and inverse Fast
Fourier Transforms (FFT) in one-, two- and three-dimensions.
The FFT can be multithreaded if the environment variable
\codeinline{OMP\_NUM\_THREADS} is defined.

\subsection*{The \fluiddyn command-line utilities}

The package \fluiddyn also provides few command-line utilities to perform
simple tasks useful for a scientist developing with Python and using the
FluidDyn packages.
\begin{itemize}

\item \codeinline{fluidinfo}

Displays important information related to software and hardware. It
includes detailed information such as currently installed FluidDyn packages,
other third-party packages, C compiler, MPI and \Numpy configuration.

\item \codeinline{fluiddump}

Utility to print the hierarchy of HDF5 and NetCDF files. It does not depend on
the NetCDF4 library.

\item \codeinline{fluidnbstripout}

Very simple layer to stripout Jupyter notebooks of output, which is useful to
keep the notebooks lightweight when included in a repository.
%
This tool is based on \href{https://github.com/kynan/nbstripout}{nbstripout}
but, in contrast to nbstripout,
%
by default the notebooks with a file name ending as `.nbconvert.ipynb' are
excluded.

\item \codeinline{fluidmat2py}

Utility to produce a strange code which is no longer Matlab and not yet Python.
%
This strange code is then much easier to translate into correct Python than the
original Matlab code.

\end{itemize}

\section*{Quality control}

% \textcolor{blue}{Detail the level of testing that has been carried out on the
% code (e.g. unit, functional, load etc.), and in which environments. If not
% already included in the software documentation, provide details of how a user
% could quickly understand if the software is working (e.g. providing examples
% of running the software with sample input and output data). }

The package \fluiddyn\ currently supplies unit tests covering around 70\% of its
code.  These unit tests are run regularly through continuous integration on Travis
CI with the most recent releases of \fluiddyn's dependencies and on Bitbucket
Pipelines inside a static \href{https://hub.docker.com/u/fluiddyn}{Docker
container}.  The tests are run using standard Python interpreter with all
supported versions.

For \fluiddyn, the code coverage results are displayed at
\href{https://codecov.io/gh/fluiddyn/fluiddyn}{Codecov}.  Using third-party
packages \pack{coverage} and \pack{tox}, it is straightforward to bootstrap the
installation with dependencies, test with multiple Python versions and combine the
code coverage report, ready for upload. It is also possible to run similar
isolated tests using \pack{tox} or coverage analysis using \pack{coverage} in a
local machine.  Up-to-date build status and coverage status are displayed on the
landing page of the Bitbucket repository.

We also try to follow a consistent code style as recomended by PEP (Python
enhancement proposals) 8 and 257. This is also inspected using lint checkers
such as \codeinline{flake8} and \codeinline{pylint} among the developers.  The
code is regularly cleaned up using the Python code formatter \codeinline{black}.

\section*{(2) Availability}
\vspace{0.5cm}
\section*{Operating system}

% \textcolor{blue}{Please include minimum version compatibility.}

Windows and any POSIX based OS, such as GNU/Linux and macOS.

\section*{Programming language}

% \textcolor{blue}{Please include minimum version compatibility.}

Python 2.7, 3.4 or above.

% \section*{Additional system requirements}

% \textcolor{blue}{E.g. memory, disk space, processor, input devices, output devices.}

\section*{Dependencies}

% \textcolor{blue}{E.g. libraries, frameworks, incl. minimum version
% compatibility.}

We list here only the dependencies of the base package \fluidpack{dyn}.

\begin{itemize}
\item {\bf Minimum:} \Numpy, \pack{Matplotlib}, \pack{psutil}, \pack{future},
\pack{subprocess32} (for Python 2.7 only).

\item {\bf Full functionality:} \pack{h5py}, \pack{h5netcdf}, \pack{mpi4py},
\pack{Scipy}, \pack{pyfftw} (requires FFTW library), \pack{pillow}.

\item {\bf Optional:} OpenCV with Python bindings, \pack{scikit-image}.
\end{itemize}
\section*{List of contributors}

% \textcolor{blue}{Please list anyone who helped to create the software (who may
% also not be an author of this paper), including their roles and affiliations.}

\begin{itemize}
\item Pierre Augier (LEGI): creator of the FluidDyn project, developer of
majority of the FluidDyn packages, future-proofing with Python 3 compatibility
and documentation.

\item Ashwin Vishnu Mohanan (KTH): developer of the packages
\fluidpack{sim}, \fluidpack{fft}, \fluidpack{image} and
\fluiddyn; documentation, code coverage and continuous integration (Docker,
Bitbucket pipeline and Travis).

\item Cyrille Bonamy (LEGI): developer of \fluidpack{fft}, \fluidpack{sim} and
\fluidpack{foam}. Main maintainer of \fluidpack{foam}.

\item Antoine Campagne (LEGI): developer of \fluidpack{image} and
\fluidpack{lab}.

\item Miguel Calpe Linares (LEGI): developer of \fluidpack{sim} and
\fluidpack{lab}.

\item Julien Salort (Laboratoire de physique, ENS de Lyon): developer of
\fluidpack{lab}.

\end{itemize}
%
% \section*{Software location:}
%
% % {\bf Archive} \textcolor{blue}{(e.g. institutional repository, general
% % repository) (required – please see instructions on journal website for
% % depositing archive copy of software in a suitable repository)}
%
% \begin{description}[noitemsep,topsep=0pt]
% \item[Name:] PyPI
% \item[Persistent identifier:] https://pypi.org/project/fluiddyn
% \item[Licence:] CeCILL-B, a BSD compatible French licence.
% \item[Publisher:] Pierre Augier
% \item[Version published:] 0.2.3
% \item[Date published:] 02/05/2018
% \end{description}
%
% {\bf Code repository}
% % \textcolor{blue}{(e.g. SourceForge, GitHub etc.) (required)}
%
% \begin{description}[noitemsep,topsep=0pt]
% \item[Name:] Bitbucket
% \item[Persistent identifier:] https://bitbucket.org/fluiddyn/fluiddyn
% \item[Licence:] CeCILL-B
% \item[Date published:] 2015
% \end{description}

% {\bf Emulation environment} \textcolor{blue}{(if appropriate)}

% \begin{description}[noitemsep,topsep=0pt]
% \item[Name:] \textcolor{blue}{The name of the archive.}
% \item[Persistent identifier:] \textcolor{blue}{e.g. DOI, handle, PURL, etc.}
% \item[Licence:] \textcolor{blue}{Open license under which the software is licensed.}
% \item[Date published:] \textcolor{blue}{dd/mm/yy}
% \end{description}

% \section*{Language}

% \textcolor{blue}{Language of repository, software and supporting files.}

% English

\section*{(3) Reuse potential}
%
% \textcolor{blue}{Please describe in as much detail as possible the ways in
% which the software could be reused by other researchers both within and outside
% of your field. This should include the use cases for the software, and also
% details of how the software might be modified or extended (including how
% contributors should contact you) if appropriate. Also you must include details
% of what support mechanisms are in place for this software (even if there is no
% support).}
%

As a library, \fluiddyn\ has been used in the project's specialized packages.
The common code base for packages with such varied applications is a proof of
\fluiddyn's versatility and generality. It can be used in other packages outside
the FluidDyn project easily depending on the need. The command-lines tools can
be useful to all scientists working with Python. Other use cases could be:
\begin{itemize}
	\item \codeinline{ParamContainer} as a generic parameter storage
		standard.
	\item Modules within \codeinline{fluiddyn.util} for miniature tasks,
		printing in terminal with colours, getting memory usage
		information, detecting if the current Python session is inside
		IPython, creating a string with time and date, etc.
	\item Modules within \codeinline{fluiddyn.io} to read
		and save images of various formats including TIFF files with
		multiple images; easily ask user with yes/no queries; handle
		different file formats using classes, including CSV, HDF5,
		Digiflow, Dantec formats.
	\item Modules inside \codeinline{fluiddyn.clusters} subpackage to
		make job submission scriptable in HPC clusters with OAR or
		SLURM job schedulers. It could be even used for non-Python
		jobs.
\end{itemize}

There is no formal support mechanism. However, bug reports can be submitted at the
\href{https://bitbucket.org/fluiddyn/fluiddyn/issues}{Issues page on Bitbucket}.
Discussions and questions can be aired on instant messaging channels in Riot (or
equivalent with Matrix protocol) at
\url{%
  https://matrix.to/#/#fluiddyn-users:matrix.org}
or via IRC protocol on Freenode at \codeinline{\#fluiddyn-users}.

\subsection*{Conclusions}

FluidDyn is an attempt to set off collaborative dynamics based on open-source
development in fluid dynamics research.
%
We shall try, with this project, to explore the possibilities of open-source in
science and fluid dynamics, by fully exploiting the new open-source tools and
methods presented in the first subsections of the introduction.

% todo: again mention some success of community-driven subject-area research
% libraries? See the file ./communitydriven_research_soft.md

The project is right now in a very preliminary stage.  Packages are actively
evolving with interesting features and a framework for collaborative
development, packaging, documenting and testing is now well set.
%
However, the community around the project is currently tiny and now, we have to
work on attracting users and developers since an active community is a
criteria for success and sustenance of an open-source project.

Will people use the FluidDyn tools and collaborate through the project
FluidDyn?
%
There are clearly many challenges and potential barriers:
\begin{itemize}
\item Some habits in the community.
\item Knowledge and skills in the community. For example, only few people use
issue tracker and pull requests.
% from Julien Salort: bad!
% \item Misunderstanding between colleagues and in particular colleagues with
% responsibilities and power.
\item Lack of a business model for open-source software in science.  A good
quality open-source software has a cost that institutions should be willing
to fund and support.
\item Lack of recognition of the work spent in open-source, in particular for
precarious scientists.
\end{itemize}

On our side, we also have positive points.
%
The quality of the tools we use (Python and its scientific ecosystem,
Mercurial, Read the Docs, Jupyter, ...) is impressive.
%
Scientific code is done to be read and to transmit ideas. To this effect, Python
is among the best languages today.
%
We can benefit from the dynamics of Python and of emerging subjects like deep
learning and the Internet of Things.
%
Finally, if we manage to gather a community of users and of developers, the
collective efficiency related to open-source methods and tools can be a strong
booster.

\section*{Acknowledgements}

% \textcolor{blue}{Please add any relevant acknowledgements to anyone else who
% supported the project in which the software was created, but did not work
% directly on the software itself.}

We thank the CNRS to finance the work of Pierre Augier while giving freedom in
terms of scientific project.
%
Similarly, Ashwin Vishnu Mohanan could not have been as involved in this
project without the kindness of Erik Lindborg.
%
We thank Cyrille Bonamy, Antoine Campagne, Miguel Calpe Linares and Julien
Salort for their implications in the development of some of the FluidDyn
packages.
%
We thank Bitbucket for providing us for free a high quality forge compatible with
Mercurial.
%
Finally, we thank Gabriel Moreau and Jo\"el Sommeria for their constant willing
to discuss and share their knowledge.

\section*{Funding statement}

% \textcolor{blue}{If the software resulted from funded research please give the
% funder and grant number.}

This project has indirectly benefit from funding from the foundation Simone et
Cino Del Duca de l'Institut de France, the European Research Council (ERC)
under the European Union's Horizon 2020 research and innovation program (grant
agreement No 647018-WATU and Euhit consortium) and the Swedish Research Council
(Vetenskapsr{\aa}det): 2013-5191.
%
We have also been able to use supercomputers of CIMENT/GRICAD, CINES/GENCI and
the Swedish National Infrastructure for Computing (SNIC).

% \section*{Competing interests}

% \textcolor{blue}{If any of the authors have any competing interests then these
% must be declared. The authors’ initials should be used to denote differing
% competing interests. For example: “BH has minority shares in [company name],
% which part funded the research grant for this project. All other authors have
% no competing interests."
% %
% If there are no competing interests, please add the statement: “The authors
% declare that they have no competing interests.” }

% The authors declare that they have no competing interests.

% already included by \bibliography{../bib}
% \section*{References}

% \textcolor{blue}{Please enter references in the Harvard style and include a DOI
% where available, citing them in the text with a number in square brackets,
% e.g.}

% \textcolor{blue}{[1] Piwowar, H A 2011 Who Shares? Who Doesn't? Factors
% Associated with Openly Archiving Raw Research Data. PLoS ONE 6(7): e18657. DOI:
% \\ http://dx.doi.org/10.1371/journal.pone.0018657.}
%
% \bibliographystyle{../agsm}
% \bibliography{../bib}

% \rule{\textwidth}{1pt}
%
% { \bf Copyright Notice} \\
% Authors who publish with this journal agree to the following terms: \\
%
% Authors retain copyright and grant the journal right of first publication with
% the work simultaneously licensed under a
% \href{http://creativecommons.org/licenses/by/3.0/}{Creative Commons Attribution
% License} that allows others to share the work with an acknowledgement of the
% work's authorship and initial publication in this journal.
%
% Authors are able to enter into separate, additional contractual arrangements
% for the non-exclusive distribution of the journal's published version of the
% work (e.g., post it to an institutional repository or publish it in a book),
% with an acknowledgement of its initial publication in this journal.
%
% By submitting this paper you agree to the terms of this Copyright Notice, which
% will apply to this submission if and when it is published by this journal.
%
%

### 2019-02-13 templates/paper.tex
Templating paper.tex and JFM CSL

% Define title, author(s), affiliation and publishing status
%
\papertitle[Title] % Short title used in headlines (optional)
{%
  Long title% THE COMMENT SYMBOL AT THE END OF THIS LINE IS NEEDED
}%
%
\papertoctitle{Long title} % Title for toc
%
% Short authors used in headlines and List Of Papers
\paperauthor[A. Beta, G. Delta \& E. Phi]
{%
  Alpha Beta$^1$, Gamma Delta$^2$ and Epsilon Phi$^2$] % Short authors used in headlines and List Of Papers
}%
%
% (optional) Short authors used in List Of Papers
\listpaperauthor[A. Beta, G. Delta \& E. Phi]
%
\paperaffiliation
{%
      $^1$ Linn\'e FLOW Centre, KTH Mechanics, S-100 44 Stockholm, Sweden \\
      $^2$ Ancient Rome University
}%
%
\paperjournal[Gal. Empire Pub.] % Short publish info used in List Of Papers
{%
	Galactic Empire Publications%
}%
%
\papervolume{42}%
%
\papernumber{2}%
%
\paperpages{1--10}%
%
\paperyear{3639}%
%
\papersummary%
{% Insert summary of the paper here (used in introduction) 
    The implications of concurrent archetypes have been far-reaching and
pervasive. Given the current status of heterogeneous technology,
cyberinformaticians daringly desire the key unification of the Turing
machine and erasure coding. We explore new decentralized information,
which we call Tuna.

}%
%
\graphicspath{{paper1/}}%
%
%
%===============================================================================
%                            BEGIN PAPER
%===============================================================================
%
\begin{paper}

\makepapertitle

%------------------------------------------------------------------------------
% Abstract
%------------------------------------------------------------------------------
%
\begin{paperabstract}
    The implications of concurrent archetypes have been far-reaching and
pervasive. Given the current status of heterogeneous technology,
cyberinformaticians daringly desire the key unification of the Turing
machine and erasure coding. We explore new decentralized information,
which we call Tuna.

\end{paperabstract}

%------------------------------------------------------------------------------
% Article
%------------------------------------------------------------------------------
%
\input{paper1/article.tex}

%------------------------------------------------------------------------------
% Bibliography
%------------------------------------------------------------------------------
%
%\clearpage
\bibliographystyle{jfm}
\bibliography{thesis}
%
\IfFileExists{paper1/paper.bbl}{\input{paper1/paper.bbl}}{}

%===============================================================================
%                            END PAPER
%===============================================================================
\end{paper}

### 2019-02-13 templates/utils_render.py

@@ -1,0 +1,121 @@
+#!/usr/bin/env python
+"""Render a latex jinja template
+
+Usage:
+    utils_render.py <options_file> <template_file>
+
+Notes:
+    See https://github.com/AKuederle/Py-Tex-automation-example
+
+"""
+import re
+import jinja2
+import os
+import sys
+from pathlib import Path
+
+
+def get_options_from_file(path):
+    """Read a options file.
+
+    The pattern in the file is '%<key>:<value>'
+    """
+    with open(path) as f:
+        content = f.read()
+        keys = re.findall(r"%(.+):", content)
+        values = re.findall(r":\s*([\w\W]+?)\s*(?:%|$)", content)
+
+    options = dict(zip(keys, values))
+    return options
+
+
+def name_in_initials(name):
+    """John Doe -> J. Doe"""
+    name = name.split()
+    first_name = name[:-1]
+    last_name = name[-1]
+    name = [n[0] + "." for n in first_name]
+    name.append(last_name)
+    return " ".join(name)
+
+
+def get_options_from_yaml(path):
+    """Read a options YAML file."""
+    import ruamel.yaml as yaml
+
+    yml = yaml.YAML()
+    with open(path) as f:
+        options = yml.load(f.read())
+
+    options["authors_short"] = [
+        name_in_initials(name) for name in options["authors"]
+    ]
+
+    uniq_affiliations = list(set(options["affiliations"]))  # unique
+    options["authors_aff"] = []
+    for author, affiliation in zip(options["authors"], options["affiliations"]):
+        options["authors_aff"].append(
+            f"{author}$^{uniq_affiliations.index(affiliation) + 1}$"
+        )
+    options["enum_affiliations"] = [
+        f"$^{i+1}$ {aff}" for i, aff in enumerate(uniq_affiliations)
+    ]
+    
+
+    return options
+
+
+def get_template(template_file):
+    """Get a jinja template with latex tags.
+
+    modified from http://eosrei.net/articles/2015/11/latex-templates-python-and-jinja2-generate-pdfs
+    """
+    latex_jinja_env = jinja2.Environment(
+        block_start_string=r"%:",
+        block_end_string=":%",
+        variable_start_string=r"\VAR{",
+        variable_end_string="}",
+        comment_start_string=r"\#{",
+        comment_end_string="}",
+        line_statement_prefix="%%",
+        line_comment_prefix="%#",
+        trim_blocks=True,
[…]

### 2019-02-13 paper_00_fluiddyn_meta/article.tex
Add fluiddyn papers nearly as it is

--- 2019-02-12 chapter_00_open_science.tex
@@
 %% Journal of Open Research Software Latex template -- Created By Stephen
-%% Bonner and John Brennan, Durham Universtiy, UK.
+%% Bonner and John Brennan, Durham University, UK.
 %% see http://openresearchsoftware.metajnl.com
 
-\chapter{Towards reproducible, open science through open research software}
+\documentclass{../jors}
+
+\begin{document}
+
+{\bf Software paper for submission to the Journal of Open Research Software} \\
+
+To complete this template, please replace the blue text with your own. The paper
+has three main sections: (1) Overview; (2) Availability; (3) Reuse potential.
+
+Please submit the completed paper to: editor.jors@ubiquitypress.com
+
+\rule{\textwidth}{1pt}
 
@@
 
-% pa: this title may not be really good for this paper...
-% - "research and teaching" may not be very important in the title
-% - Recent progress in open-source software engineering and consequences on
-%   science
-
-% todo: find a better title!
-
-Taking advantage of recent advances in open-source software engineering to study
-fluid dynamics: the example of the FluidDyn project
+% new title
+
+FluidDyn: a Python open-source framework for research and teaching in fluid
+dynamics by simulations, experiments and data processing
+
+\section*{Paper Authors}
+
+% \textcolor{blue}{1. Last name, first name; (Lead/corresponding author first) \\
+% 2. Last name, first name; etc.}
+
+1. AUGIER Pierre$^a$\\
+2. MOHANAN Ashwin Vishnu$^b$\\
+3. BONAMY Cyrille$^a$\\
+
+\smallskip
+
+$^a$ Univ. Grenoble Alpes, CNRS, Grenoble INP\footnote{Institute of Engineering
+Univ. Grenoble Alpes}, LEGI, 38000 Grenoble, France.\\
+$^b$ Linn\'e Flow Centre, Department of Mechanics, KTH, 10044 Stockholm, Sweden.
+
+\section*{Paper Author Roles and Affiliations}
+% \textcolor{blue}{1. First author role and affiliation \\
+% 2. Second author role and affiliation etc.}
+
+1. Researcher, LEGI, Universit\'e Grenoble Alpes, CNRS, France\\
+2. Ph.D. student, Linn\'e Flow Centre, KTH Royal Institute of Technology,
+Sweden; \\
+3. Research Engineer, LEGI, Universit\'e Grenoble Alpes, CNRS, France; \\
 
@@
 \href{http://fluiddyn.readthedocs.io}{FluidDyn} is a project to foster
-open-science and open-source in the fluid dynamics community.  It is thought of
-as a research project to channel open-source dynamics, methods and tools to do
-better science.
+open-science and open-source in the fluid dynamics community.  It is thought of as
+a research project to channel open-source dynamics, methods and tools to do
+science.
 %
@@
 %
-In the present article, we review the recent advancements in web-based software
-development tools and Python which facilitated the execution of this project,
-and establish a picture of software engineering in fluid mechanics.
-%
-We give an overview of the specialized packages of the project and then
-focus on the base package called \fluidpack{dyn}, which contains common code
-used in the specialized packages.  Packages \fluidpack{fft} and \fluidpack{sim}
-are described with greater detail in two companion papers, \citet{fluidfft,
-fluidsim}.
+In the present article, we give an overview of the specialized packages of the
+project and then focus on the base package called \fluidpack{dyn}, which contains
+common code used in the specialized packages.  Packages \fluidpack{fft} and
+\fluidpack{sim} are described with greater detail in two companion papers,
+\citet{fluidfft, fluidsim}.
 %
@@
 
-Fluid dynamics research with Python. Free and open-source software, modular,
-object-oriented, collaborative, efficient, tested, documented.
+Fluid dynamics research with Python. Numerical simulations. Laboratory
+experiments. Free and open-source software, modular, object-oriented,
+collaborative, efficient, tested, documented.
 
@@
 % functionality should be included in this section. }
-
-\subsection*{Science, software, open-source and the computer revolution}
 
@@
 big boom in practical uses of data science and machine learning, which drives a
-strong research on artificial intelligence.  This dynamics contributes to
+strong research on artificial intelligence. Such developments contribute to
 progresses in open-source software.
+%
+To summarize, there is a strong dynamics in play around the use of computers (in
+particular with the web) and this creates very efficient tools and methods
+for collective work and software development.
 
@@
 theory and the mathematical demonstration, which had to be elegant as it gets
-included in the articles.  In contrast, it was normal to write crude code
-% av: crappy = shitty!
-% alternatives: shabby, inelegant, amateurish, unpolished, clumsy
-% https://en.oxforddictionaries.com/thesaurus/amateurish
-and to just show the results. Nowadays, codes tend to be at the heart of
-research.
+included in the articles.  In contrast, it was normal to write crude code and to
+just show the results. Nowadays, codes tend to be at the heart of research.
 
@@
 
-This paper presents FluidDyn, a project to foster open-science and open-source in
-fluid mechanics. This first section ``Overview'' is organized into two large
-parts.  A long introduction presents the context and describes our motivations and
-the concept behind the project FluidDyn.  It is organized as follows:
-\begin{itemize}
-\item A scientific project based on the collaboration through open-source, such
-as FluidDyn, is only possible thanks to the recent developments in methods and
-tools for open-source software engineering. We start by presenting these
-developments and discuss the implications for the scientists.
-
-\item FluidDyn is based on the language Python. Incidently, it is the defining
-characteristics of Python which make the eventual success of such a
-project realizable.
-
-Python is one of the most important tool in recent open-source dynamics and in
-particular in science.
-%
-Python in fluid mechanics is growing fast, but is yet to reach widespread
-adoption as the language of choice. Therefore, it is useful to present its
-characteristics for readers who are unacquainted with its nuances.
-
-\item We will discuss about the possible contradiction between productivity of
-individuals and productivity at the community level.
-
-\item We will establish a picture of software engineering in fluid mechanics
-and present some of the main tools commonly used.
-
-\item The introduction finally ends with a summary of the motivations for the
-project FluidDyn and a presentation of its principles.
-\end{itemize}
-
-The section ``Overview'' is completed by a second subsection called
-``Implementation and architecture'' devoted to present the organization of the
-packages of the project and to describe in more detail the base package
-\fluiddyn\footnote{We use FluidDyn (with capital letters) to name the project and
-\fluiddyn for the base package.}.
-
-\subsection*{Methods and tools for open-source software engineering}
-
-\paragraph{Free and Open-Source Software (FOSS).} The term \emph{free} in
-FOSS is a misnomer, as it actually stands for \emph{freedom} (to use, modify
-and distribute). The FOSS movement has dramatically decreased the cost of using
-computers, and is evident from the widespread use of GNU/Linux systems in
-desktops, computing clusters and web-servers in academia and beyond.
-%
-The beginnings of FOSS culture can be traced back to the success for scientific
-publication of the free type-setting standards \TeX\ (1977) and \LaTeX\ (1985).
-%
-Another founding moment for the FOSS movement was the launch of the GNU project
-in 1983 by \href{https://en.wikipedia.org/wiki/Richard_Stallman}{Richard
-Stallman}, to create a Unix-like computer operating system composed entirely of
-free software\footnote{For completeness, see also the work done on
-\href{https://www.levenez.com/unix/}{other Unix operating systems,
-for example BSD}.}.
-%
-GNU is today known for its compiler collection (GCC) and a multitude of tools
-which when combined with the Linux kernel (created by
-\href{https://en.wikipedia.org/wiki/Linus_Torvalds}{Linus Torvalds} in 1991)
-forms the GNU/Linux operating systems that we are familiar today.
-%
-FOSS movement has had huge successes in many frontiers --- Apache could be
-termed as the ``first killer-app of Linux'' and now with an ever-increasing
-suite of softwares including Firefox, LibreOffice, Gimp as solid alternatives
-to proprietary offerings.
-%
-Linux has become the most widely used kernel, being deployed on servers,
-personal computers, embedded devices, and also smart phones (with Android).
-
-% remark Julien Salort: not interesting. Nothing on BSD.
-% https://www.levenez.com/unix/
-
-Over the years, FOSS development has transitioned from an organic community of
-volunteers, towards an organized system with participation from industries,
-non-profit organizations and government institutions. This could be regarded as
-the second-age of FOSS movement~\cite{fitzgerald_transformation_2006}.
-
-%todo: DONE
-% Here, we have to add a paragraph on free and open-source (libre)
-% software:
-
-% The following facts can be mentioned:
-
-% - tex (started in 1977), latex (1985)
-
-% - GNU (started 1983), FSF in 1985, licence and philosophy...
-% first complete system (with the Linux kernel) in 1991
-
-% - 90s arrival of internet (soon mass market)
-
-% Some huge open-source successes : Apache ("first killer app of Linux"),
-% now Firefox, Open-office
-% av: https://www.reddit.com/r/AskReddit/comments/7x639l/what_free_software_is_so_good_you_cant_believe/du6pw11/
-
-% Linux kernel now widely used on servers, personal computers, embedded
-% devices, smarth phone (with Android).
-% https://en.wikipedia.org/wiki/Linux_kernel
-
-% Now, new period for open-source: 
-% see http://www.cepis.org/upgrade/files/full-2005-III.pdf
-
-% - "Libre Software Movement: The Next Evolution of The IT Production
-% Organization?"
-
-% - ``the composition of development teams was changing, from all-volunteer
-% teams to teams with paid participants from industry, government or
-% not-for-profit organizations.''
-
-% By the way, we also have to use the term "libre software" (?)
-% av: Libre-software is not so mainstream. Even Linux kernel is not libre with
-% binary blobs for drivers.
-
-% We may have to change the order of this list?
-% - git and mercurial 2005
-% - software repositories ~ 10 years before
-% so I would exchange the too paragraph...
-% av:That is probably because we did not mention patches and CVS, subversion
-% etc.
-
-\paragraph{Distributed Version Control Systems (DVCS) and web-based source
-development platforms.}  Collaboration was achieved in the early years of FOSS
-development through emailing ``patches'', and centralized VCS near the turn of
-the 21\textsuperscript{st} century.  The collective collaboration on
-development is today streamlined with the emergence of DVCS (especially, Git
-and Mercurial) and web-based source development platforms like GitHub,
-Bitbucket and more recently GitLab.
-%
-These tools are designed to work on any kind of text files, not just codes, so
-they could even be utilized to write scientific documents with multiple authors.
-%
-These tools are so useful for work, that using DVCS in science shall reap
-benefits in terms of productivity in the long run~\cite{wilson_best_2014}.
-
-Git has become the \textit{de facto} standard for source management. However,
-for the FluidDyn project, we prefer to use Mercurial because we consider it to
-be simpler for the beginners and just as powerful for expert users\footnote{See
-\href{https://www.mercurial-scm.org/wiki/GitConcepts}{``a detailed
-comparison''} on Mercurial website.}. Simplicity and lowering the learning
-barrier for scientists who are not experts in software development is crucial
-for our project.
-%
-Note that, it is now possible to use Mercurial to work with Git repositories,
-for example hosted in GitHub and GitLab\footnote{By using the Mercurial
-extension \href{http://hg-git.github.io/}{hg-git}.}.
-
-\paragraph{Package management and third-party software repositories.} Python
-as a medium for scientific computing owes its success to a rich ecosystem of
-third-party packages.  Smooth deployment of these packages requires a package
-manager and a supported repository. A credible hub to upload source codes as
-compressed archives or as pre-compiled binary packages was therefore
-indispensable. The use of repositories started with the conception of CPAN and
-CRAN repositories for the languages Perl and R in 1993 --- based on the
-Comprehensive TeX Archive Network (CTAN) model for \TeX\ packages.
-
-Today, \href{https://pypi.org}{Python Package Index (PyPI)} and
-\href{https://anaconda.org}{Anaconda Cloud} are major repositories for
-delivering and downloading Python packages, assisted by package managers
-\pack{pip} and \pack{conda} respectively.
-% Almost all major 
-% languages\footnote{\href{https://stackoverflow.com/questions/1693529/%
-% list-of-top-repositories-by-programming-language}{https://stackoverflow.com/%
-% questions/1693529/list-of-top-repositories-by-programming-language}}
-% have similar repositories, with exceptions such as Fortran and Matlab.
-Specialized package managers have also emerged to install scientific
-libraries and compilers such as \pack{easybuild} and \pack{spack}.
-
-\paragraph{Continuous integration (CI).} The need for strong reliability of
-programs with increasing complexity has forced developers to use continuous
-integration.
-%
-In CI, firstly the code is \emph{built} against a known, repeatable emulation
-environment on a server. Thereafter, a set of \emph{unit tests} are run for each
-commit (modification of the code) to check that no bugs have been introduced. It
-is also important to analyse what portion of the source code is covered by these
-unit tests --- a method referred to as \emph{code coverage}. In FluidDyn project
-we stop at this level, but it is possible to further enhance it by adding code
-style checks by running linters alongside unit tests; and also to add
-\emph{continuous delivery} of documentation and package releases.
-
-Almost all CI platforms are free for open-source projects --- Travis CI,
-CircleCI and Appveyor % (sic) av: I have double checked the official names
-to name a few. Bitbucket and GitLab have built-in CI
-implementations.  Websites such as Codecov and Coveralls help to chart and
-analyse the code coverage history, viewable by anyone.
-
-When the code depends on a wide variety of external dependencies, CI is useful
-to avoid \href{https://en.wikipedia.org/wiki/Software\_rot}{software rot},
-i.e.\ to avoid becoming obsolete with newer versions of the compiler or its
-dependencies.  CI is specifically important for dynamical
-languages and open-source code since it allows open-source projects to welcome
-participation from many people while maintaining a high degree of reliability.
-
-\paragraph{Websites to share knowledge.} Mutual aid and knowledge sharing
-have been important aspects of the web since its inception and these practices
-are now of foremost importance for many human activities.
-%
-In particular, learning programming and computer science is now greatly aided
-by websites like the community driven encyclopedia
-\href{https://www.wikipedia.org/}{Wikipedia} or the community driven forum
-\href{https://stackoverflow.com/}{Stack Overflow}, where people write thousands
-of questions and answers on programming each week.
-% todo: remark from Julien Salort: it's old.
-% av: So are technologies like git and hg, but we are putting together what
-% characterizes our workflow, aren't we? We can emphasize that these forums
-% often work only when there is a "thriving community" like the Python
-% userbase.
-IRC protocol has continued to sustain the test of time. Instant messaging
-channels such as \codeinline{\#python} on Freenode facilitate interactive
-discussions and a form of community-level support system --- a role which used
-to be filled by mailing lists alone in the past. Alternatives such as Riot
-(Matrix protocol), Gitter, and Slack have gained traction offering different
-integrations solutions on top of instant messaging.
-%
-Through instant messaging a geographically separated team, as is the case with
-FluidDyn project, can achieve near real-time collaboration without the hassle
-of setting up video conferences or meetings.
-
-\paragraph{Automatic web documentations built on servers.}  Documentation
-generators such as Sphinx (for Python) and Doxygen (for many languages
-including C, C++, Fortran and Python) ``automagically'' convert in-line
-comments into documentations as PDF, HTML and more. Read the Docs has in the
-recent years become the main host for building and publishing documentation on
-servers as web pages, thus easing the process for developers and considerably
-improving access to information.
-
-% We need transitions and some words at the end of the subsections to help the
-% readers to understand why we present what we present and what is the link
-% with the FluidDyn project...
-
-We have presented new methods and tools used in software development.  Taken
-all together, they greatly improve our efficiency for collective work and open
-possibilities of interactions that were unthinkable only few years ago.
-%
-For example, the success of community-driven software like \pack{astropy} (a
-subject-area research library with \href{https://github.com/astropy/astropy}{more
-than 240 contributors}) and \pack{scikit-learn} (a machine learning toolkit with
-\href{https://github.com/scikit-learn/scikit-learn}{more than 1000 contributors})
-would not have been possible without these new software development methods.
-%
-% pa: done: we need to mention in this paragraph existing community-driven
-% subject-area research libraries (see the file
-% ./communitydriven_research_soft.md).
-%
-As a scientists, we can wonder how to fully exploit these new possibilities to do
-science.  The project FluidDyn is an attempt to help the fluid dynamics community
-to advance in this direction.
-
-\subsection*{Python, a programming language adapted for open-science}
-
-We now turn to the presentation of the
-\href{https://www.python.org/}{programming language Python}, another
-fundamental tool for the FluidDyn project.  Note that, this is not a detailed
-technical presentation: we focus on the ideas necessary to understand why the
-FluidDyn project is based on this language and on facts interesting for
-potential FluidDyn users and developers.
-
-Python community has thrived remarkably in the last few years, evident from the
-fact that by the end of 2017, Python became the second most popular language in
-terms of pull-requests in GitHub\footnote{See GitHub's annual report titled
-\href{https://octoverse.github.com/}{ ``The State of the Octoverse 2017''}.} and
-questions on StackOverflow\footnote{See
-\href{https://stackoverflow.blog/2017/09/06/incredible-growth-python/}{ ``The
-Incredible Growth of Python''}, a blog post by Stack Overflow.}.
-%
-TIOBE-index\footnote{\url{https://www.tiobe.com/tiobe-index}}, which assess the
-popularity of a language based on number of skilled users, courses and
-third-party vendors, ranks Python at number four. IEEE Spectrum\footnote{See
-the analysis titled \href{https://spectrum.ieee.org/computing/software/%
-the-2017-top-programming-languages}{``The 2017 Top Programming Languages''} on
-IEEE spectrum.} ranks Python as the most popular language in 2017 with a metric
-based on trends, demand by employers and users on open-source hubs.  All the
-rankings do agree that Python usage has been steadily growing and is here to
-stay.  Python largely outclasses languages like Fortran and Matlab according to
-these statistics, and is now clearly one of the mainstream languages with C,
-Java, C++ and Javascript.
-%
-Let us summarize some characteristics of the Python languages that have lead to
-such incredible success.
-
-% todo: DONE
-% We need an introductory paragraph on how big is Python today
-% + the fact that several scientists miss this fact.
-
-% See the file ./python-one-main-languages.rst
-
-The Python language has been designed to boost the communication of technical
-ideas between humans. The resulting code is as simple as possible due to the
-inherent emphasis on readability. Humans can focus on the ideas. Python differs
-from other mainstream languages for a variety of reasons:
-
-\begin{itemize}
-\item \textbf{Aesthetics.} The syntax is nice and elegant. Blocks of code are
-defined with the indentation.
-\item \textbf{Code style.} There are explicit guidelines
-(\href{https://www.python.org/dev/peps/pep-0008/}{PEP~8}) supporting the
-regularity (and thereby readability) of the code and tools
-(linters) to evaluate adherence to the style.
-\item \textbf{Dynamic typing.} The variables (the ``names'' in the code) are
-not attached for life to an object in memory\footnote{Opposite behavior
-compared to static languages as Fortran, C and C++.}. The types of the objects
-are inferred at run time from the code and the context, and in many situations,
-one does not need to declare them explicitly.
-%
-% Being a strongly typed language\footnote{In contrast to weakly typed languages,
-% such as JavaScript}, types of the objects are respected and implicit type
-% conversions are rarely performed.
-% av: "rarely / never"?
-% eg: 'hello' * 2 -> 'hellohello'; 1/3 -> 0.3333333
-%
-% pa: I think we can not cover such complicate notion in this document... It is
-% anyway not so important for our demonstration.
-\item \textbf{Automatic memory management.}  The user need not handle the
-memory allocation and deallocation by hand.  There are, of course, ways to
-control the memory usage of a program, though not as precisely as for low-level
-languages.
-\item \textbf{Interpreted.} The standard way\footnote{There are also tools to
-compile Python code (for example Cython and Pythran).} to run a Python code is
-to ``interpret'' it.  A program called the interpreter executes the code
-nearly\footnote{Actually, the execution of Python code is less simple and
-involves an Abstract Syntax Tree.}  instruction-by-instruction, with very few
-optimizations.
-%
-In contrast to other compiled languages such as Fortran, C or C++, there is no
-proper compilation step: the code is not translated to optimized machine
-instructions, but only to Python bytecode.
-%
-We will discuss shortly the disadvantages of this characteristic. Two evident
-advantages are rapid development cycle and the possibility of interactive
-workflow (usually with IPython or Jupyter notebooks).
-
-Note that there are many implementations of Python interpreters\footnote{We can
-cite CPython (written in C), Jython (Java), IronPython (C\#), PyPy (Rpython, a
-subset of Python) and MicroPython/CircuitPython (C, targeted to
-micro-controllers).}. The default and most widely used implementation is written
-in C and is called CPython.
-
-\item \textbf{Simple, yet powerful.}
-
-\begin{itemize}
-\item Python supports multiple programming paradigms, including imperative,
-object-oriented and functional.
-
-\item It is easy to interface with code written in other languages (in particular
-C, C++ and Fortran).
-
-\item Python comes with a large and high quality \href{https://docs.python.org%
-/3/library/index.html}{standard library}.
-
-\item Python can be deployed on many different machines with different operating
-systems (Linux, Windows, macOS, Android) and architectures (from a microcontroller
---- with MicroPython or CircuitPython --- to a Blue Gene supercomputer).
-
-\end{itemize}
-\end{itemize}
-
-Thus, Python is one of the best ``glue'' languages for fast prototyping.
-Developing with Python is much faster and easier than with many other
-languages.  Less bugs are introduced by the developers simply because there are
-much less lines of code \cite{nanz_comparative_2015}.
-%
-Maintaining and modifying existing code is also eased by the very good code
-readability associated with the language.
-
-Another consequence of the apparent simplicity of Python is that the learning
-curve has a nice shape. Learning Python is very easy at the beginning and the
-complexity gradually increases as the level of the developer increases.
-%
-Therefore, Python is good for developers of all levels: very gentle for
-beginners and very powerful for advanced users.
-
-Python is an old language (first implementation in 1991!) but it continues to
-evolve.
-\begin{itemize}
-\item A hard and controversial decision was taken to clean up Python from its
-historical incoherences by having a new series of backwards-incompatible
-versions of the language, Python 3.  Python 3.0 was released in December 2008.
-After a long and difficult process, the transition from Python 2 to Python 3 is
-now (2018) behind us for nearly all important packages\footnote{Close to
-\href{https://python3wos.appspot.com/}{95\% of the most downloaded packages}
-are Python 3 compatible now.}.  Especially for scientific applications, we can
-now work only in Python 3. For instance, \href{https://github.com/%
-numpy/numpy/blob/master/doc/neps/dropping-python2.7-proposal.rst}{future
-releases of \Numpy} and of \href{http://www.python3statement.org/}{the main
-scientific packages} will soon drop compatibility with Python 2. These changes
-open doors to very clean, coherent and potentially faster Python
-experience. Therefore, one should not use Python 2 for science anymore and
-instead, adopt the newer versions ($>=3.5$ in 2018).
-
-\item New versions of Python have nice new features to adapt to new usages and to
-recent trends in computer science.  For example, three new features were
-introduced in Python 3.5 (first released on September 2015): the \codeinline{@}
-operator for matrix multiplication, the new \codeinline{async} and
-\codeinline{await} keywords for concurrency and type hinting\footnote{See also our
-paragraph on the issue of type checking in the following.} (with the module
-\codeinline{typing} and an associated syntax presented in the
-\href{https://www.python.org/dev/peps/pep-0484/}{PEP~484}).
-\end{itemize}
-
-\subsubsection*{An incredible success}
-
-Being a very versatile language, Python is widely used for many different
-applications:
-\begin{itemize}
-\item Simple scripting.
-
-\item System, database and network administration.
-
-\item GNU/Linux distribution software.
-
-\item Python is now one of the main scripting languages to add programmability
-to applications (for example Paraview, Visit, QGIS, Blender, ...).
-
-\item Web servers.
-
-\item Web scraping and data analysis.
-
-\item Animation movies, game development and gaming.
-
-\item Education. Python has recently become one of the most popular language
-for teaching programming and computer science.
-
-\item Science!
-Python is now widely used in scientific applications. It has now a mature and
-powerful scientific ecosystem with well-established based packages (\Numpy for
-N-dimensional homogeneous arrays, \Scipy as the fundamental toolkit for
-scientific computing, \pack{Matplotlib} for plotting and \pack{Pandas} for data
-structures) and several more specialized packages (to name a few, \pack{h5py},
-\pack{mpi4py}, \pack{skimage}, \pack{sklearn}, ...). There are now great (and
-often shiny) tools for most of the applications. For example, the
-\href{https://github.com/rougier/python-visualization-landscape}{%
-landscape of visualization tools usable through Python} is impressive.
-%
-Python is one of the main languages for data science with packages such as
-\pack{Pandas}, \pack{statmodels}, \pack{sklearn}, \pack{Keras}, \pack{TensorFlow}
-and \pack{PyTorch}.
-
-There are now ready-to-use ``batteries-included'' Python installers
-(similar to Matlab), like in particular the
-\href{https://en.wikipedia.org/wiki/Anaconda_(Python_distribution)}{%
-freemium open-source distribution Anaconda}, so it is very easy to start using
-Python for scientific purposes.
-
-There are now nice Python integrated development environments (IDE) adapted for
-scientists, for example \href{https://github.com/spyder-ide/spyder}{Spyder},
-\href{http://www.pyzo.org/}{Pyzo} and
-\href{https://jupyterlab.readthedocs.io}{Jupyterlab}.
-
-\end{itemize}
-
-Python has a very large and supportive community (see for example
-the \href{http://stackoverflow.com/tags}{Stack Overflow tags}).
-%
-But Python is also supported by companies using Python and the open-source
-dynamics.  Most of these companies that support Python are specialized in
-web technologies and data-driven research.  But, recently, companies basing
-their business
-% av: earlier "computer and web technologies". Computer is quite obvious, isn't
-% it?
-on Python for science have started to become quite powerful in terms of
-development. For example Anaconda Inc.\ is strongly implied in the development of
-tools like \pack{Numba} and \pack{dask}. Hardware giants such as Intel and Nvidia
-also contribute to the Python scientific ecosystem.
-
-\subsubsection*{Some Python issues}
-
-\paragraph{Since the standard Python interpreter does not do any proper
-compilation,} Python code in some CPU bounded cases can be too slow.  Thus, one
-has to use special tools to get optimized machine instructions to speedup
-performance-critical code.
-
-\textbf{Ahead Of Time compilation.}  Since Python is primarily a ``glue''
-language, a common solution is to use code written in C, C++ or Fortran.
-%
-% todo: DONE
-% add a reference to ctypes and cffi to call C with pure Python code
-%
-It is possible, without a compilation step, to call functions in C libraries
-through pure python code using \pack{ctypes} standard module or the \pack{cffi}
-package.
-%
-Libraries with Python bindings can also be used directly (for e.g., wrapped
-using SWIG or Boost.Python) or one has to write extensions, i.e.\ Python
-modules written in C or C++ using the CPython API that can be imported directly
-from Python like any Python module.
-%
-However, to rely on such techniques to call C libraries would inhibit
-development of extending functionalities, as they have a steeper learning
-curve compared to Python.
-%
-Nevertheless, this strategy has given rise to the base modules of the scientific
-Python stack, namely \Numpy, \Scipy and \pack{Matplotlib}.
-
-Python developers do not actually have to write the compiled extensions in C,
-C++ or Fortran, since there are tools to generate them automatically from
-Python code (Pythran) or from Cython code. Cython \citep{behnel_cython2011} is a
-smooth blend of two languages with possibility to add type declarations like in
-C/C++, but with a syntax similar to Python.
-
-Pythran \citep{guelton2015pythran, guelton2018pythran} is a quite recent Python
-compiler which gives impressive results. It creates compiled extensions from pure
-Python code with simple type annotations written as comments.
-%
-The resulting extensions are usually as fast as Fortran or C++ written by
-non-specialists \cite[see, for example, benchmarks for the packages
-\fluidpack{fft} and \fluidpack{sim},][]{fluidfft, fluidsim}.
-%
-They are created with a very interesting two-step compilations: first the code
-is optimized at the Python level and then an automatically produced C++ code is
-properly compiled.
-%
-This two-step compilation provides room for very clever optimizations as
-demonstrated, for example, in
-\href{http://serge-sans-paille.github.io/pythran-stories/%
-being-more-than-a-translator.html}{%
-this blog post by the main developer of Pythran}.
-%
-Pythran supports OpenMP pragma \citep{guelton2013compiling} and can use modern
-vectorization with SIMD instructions \citep{guelton2014exploring}. Pythran
-understands both Matlab-like vectorized code and C-like code with explicit loops
-\citep{guelton2018pythran}.
-
-We present in the two companion papers \citep{fluidfft, fluidsim} examples of
-highly efficient Python codes, optimized using a combination of Cython and
-Pythran extensions.
-
-\textbf{Just In Time (JIT) compilation.} Another way to obtain optimized
-machine instructions is by using JIT compilation, i.e.\ by compiling only the
-critical code at run time.
-%
-This strategy can yield good results with other languages as for example Matlab
-or Julia.
-%
-Pypy, an alternative interpreter written in Python, has a JIT compiler. However,
-Pypy is not widely used for scientific applications mainly because of
-compatibility problems with the extensions written for CPython.
-%
-However this could change since a recent version of Pypy (6.0) supports the main
-packages of the Python scientific stack (\Numpy, \Scipy, \pack{Matplotlib},
-\pack{Pandas}, etc.).
-
-Adding a JIT to the interpreter CPython has been notoriously difficult (see
-\href{https://faster-cpython.readthedocs.io/}{faster-cpython}). One of the
-problem seems to be the C API provided in CPython and used by many extensions.
-%
-Python being first a glue language, a faster CPython with a JIT is actually not
-so necessary for many purposes. Therefore, not as much money and work has been
-put in accelerating Python than for example for Java and JavaScript.
-
-Another strategy is to add a JIT to CPython through an external package
-(\pack{Numba}) and to only compile the critical functions.  Moreover,
-\pack{Numba} is particularly interesting because it can take advantage of the
-GPU\footnote{See for example
-\url{https://devblogs.nvidia.com/parallelforall/seven-things-numba/}}.
-
-\paragraph{No type-checking.} In contrast to languages using Ahead Of Time
-compilation, there is no type-checking in Python (and more generally in many
-dynamical languages).  Type-checking is useful since it can avoid simple bugs.
-However, it is not a good practice even with static languages to rely only on
-type-checking to look for bugs in a code.
-%
-It is known that a rigorous set of unit tests is also needed to obtain more
-reliable codes.
-%
-Dynamical languages rely a lot on unit tests so it is very important to put at
-least a little bit of time and energy to write a decent battery of tests.
-%
-We have to mention the young and promising project
-\href{http://mypy-lang.org/}{Mypy}\footnote{Guido van Rossum, the author and
-``Benevolent Dictator For Life'' of Python, is involved in the project Mypy.},
-which allows for Optional Static Typing and thus type-checking for
-Python\footnote{We plan to add type hitting in the most important FluidDyn modules
-to investigate how we can take advantage of type-checking with Mypy.}.
-
-\paragraph{A lively, huge and thus complicated ecosystem.} As already
-mentioned, Python is a versatile language. For many applications, the standard
-library is insufficient and one needs external packages.
-%
-% For most applications there are several projects usable through Python.
-Most of the time there are several Python packages for a certain application.
-A user/developer has to make choices regarding the tools to use.  Especially
-for the beginners who are not used to the open-source realm, it can be
-difficult to understand a ``landscape of projects'' for a particular
-application and to make good technological choices.
-
-It is important to get a good introduction on how to work with the open-source
-dynamics and the scientific Python ecosystem.  It can also be very useful to ask
-to more advanced users and attend Python courses or workshops\footnote{This could
-be regular academic courses, MOOCs or workshops like ones organized by
-\href{https://software-carpentry.org/}{Software Carpentry}.}.
-%
-Of course, such abundance of projects can also be seen as an advantage.
-
-\paragraph{Absence of multicore computational parallelism using threads.}
-Threads behave like light subprograms. In principle, they can use the different
-cores of the CPU at the same time but the CPython interpreter forbids that they
-interpret python code at the same time\footnote{See
-\href{https://opensource.com/article/17/4/grok-gil}{this blog article on ``How
-to grok the GIL''} and
-\href{https://faster-cpython.readthedocs.io/cpython37.html}{this technical
-presentation by a CPython core developer}.}.
-%
-This limitation comes from an important detail in the implementation of the
-CPython interpreter: a Global Interpreter Lock (GIL) to prevent race conditions
-that could corrupt data.
-%
-It seems that the GIL greatly simplies the implementation of CPython and that it
-is very difficult to remove it while keeping other nice technical properties of
-CPython intact\footnote{See \url{https://wiki.python.org/moin/GlobalInterpreterLock}.}.
-Note that Pypy also uses a GIL, while two other implementations of the Python
-language, Jython and IronPython, do not have this limitation.
-
-At any rate, threads using the interpreter do not use the CPU at the same time
-so it is not possible to use threads to do multicore computational parallelism
-with pure python code to accelerate CPU bounded tasks.
-%
-Note however, that threads can be (and are widely) used in Python for
-concurrency, i.e.\ to perform I/O tasks (which does not need CPU) concurrently.
-%
-To do proper computational parallelism for CPU bounded tasks with the GIL, one has
-to use other strategies: extensions (fine grain parallelism), multiprocessing
-(coarse grain parallelism) and inter-process communication, for example, with MPI
-(with mpi4py) or ZeroMQ.
-
-\subsubsection*{Rich landscape of open-source languages useful for science}
-
-So far, we have discussed in detail about Python. However, there are many
-interesting open-source languages that can be used for scientific applications.
-All have strengths and weaknesses.
-
-Fortran remains widely used for computing codes, at least in fluid mechanics.
-C++ is used in many recent developments. Its recent improvements (C++11, C++14)
-make it a modern and very useful language.
-%
-JavaScript, Java, Scala, Smalltalk, Haskell, R, Julia, Perl and Lua are used for
-some scientific codes. Go and Rust are quite young languages and are not (yet)
-widely used for scientific applications.
-
-It is of course very difficult to predict which technologies will be widely
-used in even the near future (10 or 20 years for example).
-%
-From what we see, it seems that the idea of ``one language to do everything for
-science'' will not succeed, at least soon.  This was maybe the project of Julia
-creators, but despite the qualities of this language, it also has
-limitations\footnote{For example C-like include instead of Python-like
-import. From our point of view, a Pythonic flavor of Julia would be very
-interesting.}.
-%
-It seems that a very important aspect in the future will be interoperability
-between coexisting tools\footnote{See for example the cross-language development
-platform \href{https://arrow.apache.org/}{Apache Arrow}.}.
-
-So what about Python?  As already mentioned, it is today the language with the
-strongest dynamics in science, data analysis and much more.  Several students
-learn and will learn it. Many scientists in the future would, hopefully, be
-able to appreciate these strengths and favour the Pythonic approaches.  With
-such dynamics today, we can anticipate that the Python world will be able to
-embrace new technological trends and to overcome some of its current
-limitations.
-%
-It is reasonable to predict that Python will continue to shine as a language
-(i) particularly good for human communication and to write ideas, (ii) for fast
-prototyping and (iii) for interoperability with code written in other languages
-(see for example the recent projects \href{https://cppyy.readthedocs.io}{cppyy}
-and \href{https://github.com/pybind/pybind11/}{pybind11}, or
-\href{https://github.com/PyO3}{the possibility to interact with Rust code}).
-
-\subsubsection*{Summary: Python, a versatile Swiss Army Knife for the
-scientists}
-
-Python has become a great versatile tool for scientists.  The scientific
-ecosystem of Python is nowadays really impressive.  The scientific Python
-community is strong and quickly growing.
-%
-A scientist with good skills in Python can do most of what she/he needs to do in
-programming (which does not mean that Python is the best language to do
-anything!).
-%
-In contrast, no skill in Python is a real disadvantage for many tasks and for
-employability.
-%
-Therefore, in most fields today, if a scientist has to acquire good skills in
-one language, it is in Python.
-
-Considering what we have presented on Python, we think that it is reasonable to
-bet that it will become one of the \textit{lingua franca} in science and in
-fluid dynamics.  This explains why we have chosen it to be the main language
-for the project FluidDyn.
-
-\subsection*{Productivity at individual, group and community levels}
-
-% remark Julien Salort "It is well known that": bad
-% pa: common sense says ?
-% It is well known 
-It is not surprising
-% av: "It is not surprising"
-%
-that short-term efficiency and long-term efficiency are
-sometimes incompatible.
-%
-Quick and dirty scripts can be efficient in the short term, but such a workflow
-has a negative impact on a longer time scale.
-
-Similarly, we can also differentiate productivity at different scales of
-collaboration.
-%
-An individual can be very efficient with a particular tool, but the same tool
-can be very inefficient to collaborate even with co-workers.
-%
-A group can be very efficient with a code which is closed-source.  If this
-strategy can be efficient for the group (at least in the short term), it leads
-to waste for the community. Other groups will need to develop codes with the
-same features.  Good ideas will be spread in different codes and since the
-number of users and developers are smaller for each of these codes, the code
-quality will not improve as fast as it would with an open-source strategy. In a
-nutshell, as this proverb says, \textit{``If you want to go fast, go alone. But
-if you want to go far, go together.''}
-
-Thus, a community using open-source methods can be efficient to co-develop its
-tools. However, some of its members have to spend more energy to bear this
-dynamics, which implies that they somehow sacrifice their individual efficiency.
-We see that short-term efficiency, long-term efficiency and productivity at
-different levels can often be incompatible and that we have to consider these
-conflicts when choosing between different technologies.
-
-\subsection*{Programming in the field of fluid mechanics}
-
-We study fluid mechanics via laboratory experiments, in situ measurements,
-analytic computation, numerical simulations and data processing.
-%
-Nowadays, programming is involved in all methods, and thus it is difficult to do
-fluid mechanics without software development.
-%
-However, the level in software engineering is on average very low in the
-community.
-%
-Even today, it is quite commonplace to start a PhD without any serious training in
-GNU/Linux and modern programming tools.
-%
-A large majority of scientists and technicians are also unaware of the new
-challenges and opportunities of open-source.
-
-Since fluid dynamics is an engineering science, closed-source commercial software
-have historically been dominant in the field.
-%
-Few people are aware of the vicious circle of the closed-source model for people
-and groups relying on it.
-%
-This mechanism is nevertheless very damaging. In the closed-source model, a group
-pays for a license or a new development. The group does not learn how to develop
-what has been paid for. The company, on contrary, stands to gain from profits and
-the feedback generated from users, through which it improves the product sold.
-The group produces codes, books, courses using the closed-source product or
-acquire knowledge on how to use the closed-source product, so that the group is
-increasingly dependent on the product and is ready to pay more for it.
-%
-Such vicious circles are difficult to break, but not impossible. This can be
-done by introducing alternative open-source solutions, bringing in positivity to
-this circle.
-
-We have seen that it was very common to rely on commercial companies and
-closed-source software for most of the new developments in fluid mechanics.
-%
-Nevertheless there are also a lot of codes produced by scientists in the
-field. But without the technical knowledge on how to work collectively on code,
-this production is often of bad quality (compared to the information technology
-industry standard) and it is basically doomed to be abandoned after its use.
-%
-However, it is natural to try to reuse code or at least to build the next steps
-on what has been done before.  Thousands of hours of highly qualified people
-are spent in trying to understand and reuse codes badly written with
-inappropriate languages!
-
-On a technical point of view, the majority of the coding in the field involve a
-mix of Fortran/C or C++, shell languages (as Bash) and Matlab. For experiments,
-the graphical programming environment
-\href{http://www.ni.com/en-us/shop/labview.html}{Labview} is dominant for
-control of physical objects and acquisition and Matlab is used for data
-processing.
-%
-Often, languages are used for things for which they are not adapted.
-%
-Fortran, C or C++ have not been designed for fast prototyping of complex
-programs.
-%
-It is known that for scientific purposes, one should restrict shell scripting
-to extremely simple tasks.
-%
-Similarly, using Matlab for developing complex programs is technically a
-questionable choice.
-%
-These technical solutions also represent a problem for code reuse, sharing and
-collaborative development.
-%
-We are now going to present factual arguments explaining why we think that the
-massive usage of Matlab decreases the collective efficiency of the community.
-
-\subsubsection*{Discussion on Matlab, Labview and Mathematica}
-
-% see http://www.pyzo.org/python_vs_matlab.html
-% see https://www.linkedin.com/pulse/matlab-vs-python-jan-rhebergen
-
-Matlab is a closed-source proprietary numerical computing environment.
-%
-It is a good tool for simple processing with matrices, image processing, data
-visualization and certain niche applications. The language is well adapted for
-these tasks. The development environment is nice and the interpreter is quite
-fast, especially now that it has a JIT compiler.
-
-However, the language suffers from serious technical issues which make Matlab
-an inadequate tool
-% \citationneeded[opinion] % pa: I think what we say is fine for an article. We
-% are not writing a wikipedia page which has the ambition to be neutral. Here,
-% we gives an argued opinion, which is totally normal in a scientific article.
-% av: fair enough :)
-for doing more than simple processing and data plotting. The comparison with
-Python often debated upon. We list a few striking Matlab weaknesses:
-
-\begin{itemize}
-\item One file for each function (or class) available outside the file where it
-is implemented. No notion of package or module.
-
-\item No real organization of the standard library. All built-in and
-user-defined functions are available in a huge flat namespace. No import
-mechanism and as a result we cannot deduce, simply by looking at the code,
-where a function comes from.
-
-\item A standard way to organize multi-file code is to write scripts that
-modify and define global variables (see for example
-\href{https://www.damtp.cam.ac.uk/user/jrt51/files/diablo\_mat.tar.gz}{%
-this Matlab version of the code Diablo}).  It is so simple to do this that we
-can assert the language strongly encourages this practice.  Matlab files are
-not self consistent, i.e.\ it is normal to use in a file a global variable
-defined outside of the file.
-
-\item Very bad default argument mechanism.
-
-Default arguments for a function is a very common feature is many programming
-languages. In Python, we can write:
-\begin{minted}[fontsize=\footnotesize]{python}
-def myfunc(a, b, c=1, has_to_print=True):
-    if has_to_print:
-        print('a =', a, 'b =', b, 'c =', c)
-    return c * (a + b)
-\end{minted}
-
-An implementation for approximately the same behavior in Matlab could be (as
-advised in the official Matlab documentation):
-\begin{minted}[fontsize=\footnotesize]{matlab}
-function ret = myfunc(a, b, varargin)
-    if nargin < 2 | nargin > 4
-        error(['The number of arguments has to be ' ...
-               'greater than 2 and lower than 5'])
-    end
-    if nargin == 4
-        has_to_print = varargin{2};
-    else
-        has_to_print = 1;
-    end
-    if nargin >= 3
-        c = varargin{1};
-    else
-        c = 1;
-    end
-
-    if has_to_print
-        disp(['a = ' num2str(a) '; b = ' num2str(b) '; c = ' num2str(c)])
-    end
-    ret = c * (a + b);
-end
-\end{minted}
-Even for a very simple function, there are many places where bugs can be
-introduced and the code is much less readable than the corresponding Python
-code.
-
-\item Less intuitive syntax for string operations in Matlab, compared to
-Python. Take for example for string comparison, \codeinline{strcmp(s1, s2)} in
-Matlab versus \codeinline{s1 == s2} in Python or for look up,
-\codeinline{contains(s1, pattern)} in Matlab versus \codeinline{pattern in s1}
-in Python.
-% \item Less intuitive and clear syntax for exception
-% handling. \citationneeded[opinionated].
-% pa: Matlab has now a MException class. The syntax is not nice, but it is
-% not so important.
-\item Less clear and intuitive model and syntax for object oriented
-programming, compared to Python. As for functions, one file per user-defined
-class is necessary.
-% \citationneeded[opinionated].
-\item Parenthesis used for both function calls and indexing, causing ambiguity.
-\item Matlab codes are usually full of \codeinline{;}, \codeinline{.*},
-\codeinline{./}, \codeinline{\&\&} and \codeinline{||}, which make them quite
-``noisy''.  Moreover, there is nothing like the Python PEP~8. This makes a
-large proportion of the Matlab codes difficult to read and understand, due to
-lack of a consistent code style.
-\item The syntax \codeinline{a(100, 100) = 1;} to create and extend matrix is a
-very good way to hide bugs. The code \codeinline{a = eye(2); a(i0, i1) = 1;}
-will never raise any error regardless the values of \codeinline{i0} and
-\codeinline{i1}! No error is raised if a user misspell the variable
-\codeinline{a} and write something like \codeinline{aa(100, 100) = 1;}.
-\end{itemize}
-
-Now we turn our attention, from structural issues of Matlab syntax, to more
-practical constraints its use brings about.  By definition of a closed-source
-software, Matlab is a big black box: it is impossible to study the
-implementation of its functions.
-
-Matlab is not free. The price for one license for non-commercial use is not
-very expensive for most research and teaching institutes. However, when we
-start to add up the price of the toolboxes necessary to run most codes, it
-starts to become substantially large for some institutions.  The price of
-Matlab quickly escalates when deployed on a cluster with MPI, since a licence
-is required per user, per node!
-% av: not per process.
-% see: https://www.nas.nasa.gov/hecc/support/kb/running-matlab-applications-in-parallel_522.html
-
-It would be interesting to estimate what could be done with open-source
-solutions with a part of the amount paid by public research institutes. How
-many developers could be paid to improve the alternative open-source solutions?
-%
-This question raises the issue of funding scientific open-source software.
-
-Moreover, it is well known that the lack of license for the personal computers
-of staff is sometimes solved with cracked versions of Matlab. How can serious
-research projects rely on such illegal methods?
-%
-Note that open-source Matlab interpreters do exist, such as Octave and Scilab.
-%
-% pa: warning! Octave has a JIT compiler!
-%
-% Unfortunately, these are not serious alternative today. Their user
-% community is too small and they are not very efficient, especially now that
-% Matlab uses a JIT compiler and that optimized Matlab codes rely on it.
-
-The recent open-source and Python dynamics represent a hard punch for Matlab,
-which is losing the dominant position it has in some domains.
-%
-Note that this is a very recent development. The Python scientific ecosystem can
-now compete in nearly all domains, but this was not the case typically 5 years
-ago.
-%
-There are still a large number of researchers and teachers in sciences who are
-strongly attached to Matlab mainly because they have used it a lot.
-%
-It is, of course, difficult to change habits especially when those are efficient
-(for a short time-span and on an individual or group scale) enough for what you
-do.
-
-It is normal that Mathworks, the company which makes money with Matlab, spreads
-\href{https://www.mathworks.com/products/matlab/matlab-vs-python.html}{arguments
-against Python and its scientific ecosystem}.  Some of them are valid, for
-example the issue of the lively, huge and thus complicated scientific ecosystem
-that we have already discussed.  Other arguments are unfair or no longer valid
-nowadays.
-% Here, we remove a long argument with Mathworks (see the file
-% old_arguments_mathworks.tex)
-% av: Less provocative, that is good. But we can still add a footnote linking
-% the Pyzo page / LinkedIn page comparing python and matlab.
-%
-% We replace by this:
-The scientific Python ecosystem is now very mature, robust, well documented and
-fully integrated.  Many of this components are supported by professional
-developers paid by companies or research institutes, for example \pack{Numba}
-(Anaconda, Nvidia), \pack{TensorFlow} (Google), \pack{PyTorch} (Facebook, Twitter,
-Nvidia), Scikit-learn (INRIA), Mercurial (Facebook).
-%
-The support by the Python community is both very reactive and of very good
-level (for example in Stack Overflow, and issues page on major repositories)
-and it is now possible to pay companies to get professional support on
-scientific Python (for example
-\href{https://software.intel.com/en-us/distribution-for-python/get-help}{Intel}
-and \href{https://www.anaconda.com/enterprise/}{Anaconda}).
-%
-Finally, with the tools available today, with little effort we can get a very
-efficient program using only Python code.
-
-To conclude on Matlab, some people may use it efficiently for their work, but it
-starts to be a problem for the collective efficiency of the community. Except
-from the existence of a large code base difficult to translate in better
-languages or for some very specific applications, it seems difficult today to
-find a good technical reason to continue using Matlab.
-
-The same remarks can be made for Labview. It is a graphical programming
-language and the programs are saved as binary files. This is a problem because
-first it is impossible to read a Labview program without Labview and, more
-importantly, it forbids the use of source management tools which are so
-important for collective work and open-source.
-
-Matlab and Labview also lead to problems for students, who learn bad coding
-habits and languages that are much less in demand by employers than for example
-Python.
-
-For symbolic computation, some specialists use Mathematica, which is a good and
-very expensive proprietary closed-source mathematical symbolic computation
-program.
-%
-Sage and Sympy are two complementary open-source alternative both based on
-Python.
-
-\subsubsection*{Different models for software development in fluid mechanics}
-
-Proprietary codes tend to dominate the field of fluid mechanics. This is true
-for industrial Computational Fluid Dynamics (CFD), for example with the Ansys
-suite and also for acquisition and analysis of images of fluid with the
-two companies Dantec and Lavision.
-%
-Even certain softwares developed by researchers in academic laboratories
-also follow a closed-source model, for example
-\href{http://www.damtp.cam.ac.uk/user/fdl/digiflow/index.htm}{Digiflow}.
-%
-Closed-source or even undistributed software is a widespread model in the
-field.
-%
-The arguments against sharing are diverse:
-
-% todo: DONE
-% pa: I am really uncomfortable with this presentation which gives the
-% impression that we think that the arguments against sharing are always wrong.
-%
-% We have to make it clear that it is not the case and that we do respect how
-% other people think and work.
-% av: I have commented out specific counter arguments, instead added two
-% sentences after the bullets
-\begin{itemize}
-\item \textit{``We do not share to keep a comparative advantage.''}
-
-%	Sharing code leads to reproducibility, making related
-%	publications more attractive to readers. The authors gain
-%	comparative advantage by earning citations.
-
-\item \textit{``We do not share because we do not provide support.''}
-
-%	Support and documentation is invaluable, of course. But sometimes, a code
-%	with comments can be sufficient, which is a common practice.
-
-\item \textit{``We do not share because people would not be able to
-correctly use the code or interpret the results.''}
-
-% pa: this one does not respond to the problem... (bad use of the program)
-
-%	Poorly written code is infinitely more useful than no code at all.
-
-\item \textit{``We do not share because we want to control industrial usage.''}
-
-%	Proper choice of license can legally restrict and permit industrial
-%	usage at the discretion of the developer.
-
-\item \textit{``We do not share because we do not want people to review
-and criticize our code.''}
-
-%	People are, in general, kind and welcoming in the open-source community.
-
-\item \textit{``We do not share, because in this way, people will think our
-project has more value.''}
-
-% pa: this one looks too "religious" to me :-) but ok...
-%	True value is gained by furthering science and dissemination of ideas, not by
-%	fabricating an appearance.
-\end{itemize}
-
-The flip-side involves releasing a well-written, documented, easy to use and
-maintain source code with appropriate license to ensure credit where it is due.
-This can potentially add value for the community through added transparency in
-publications, and in return, value for the authors of software in terms of
-recognition and citations. We have only looked at the most ideal scenarios and
-the question of soundness of the above arguments is left as an open question to
-the reader.
-
-Between closed-source and open-source, is a grey area gathering different
-practices.
-%
-It is quite common to share only for ``friends'' to control the dissemination
-of the software.
-%
-This model can be institutionalized with an official group, for example with
-the code \href{https://www.coria-cfd.fr/index.php/YALES2}{Yales2}.
-%
-% av: Institutionalized? I don't understand.
-With such a model, the lack of a proper open-source license can naturally lead
-to problems and is a limitation for users/developers.
-%
-Another common practice is to share without repository (for example
-\href{http://www.damtp.cam.ac.uk/user/jrt51/files.html}{Diablo} or
-\href{http://yakari.polytechnique.fr/people/deloncle/ns3d.html}{NS3D}) and/or
-without license (\href{https://choosealicense.com/no-permission/}{meaning that the
-code is actually under exclusive copyright}).
-
-Finally, several scientists in the fluid dynamics community have used
-open-source methods in their work for a long time.
-%
-There are a lot of open-source research codes written with proprietary tools
-(typically written in Matlab, see for example
-\href{http://pivlab.blogspot.de/}{PIVlab},
-\href{http://servforge.legi.grenoble-inp.fr/projects/soft-uvmat/}{UVmat} and
-\href{https://de.mathworks.com/matlabcentral/fileexchange/10902-pivmat-4-00}{PIVmat}).
-%
-Let us first say that these codes are a great wealth for the community at least
-for three reasons: (i) they can be used as they are, (ii) people using these
-codes can be more open to the open-source dynamics, and (iii) open-source
-research projects using open-source tools can exploit the ideas in these codes.
-%
-Let us also point out a drawback of such a model. We do not pay the authors, but
-people do need to pay to use the software. Even the authors of the software give
-money to the same company to be able to use their software. Do they realize that
-when they work on these programs, they work (often, for free) for the company that
-sells the proprietary tool?
-
-Some proper open-source codes have emerged, for example
-\href{https://nek5000.mcs.anl.gov/}{NEK5000} (Fortran),
-\href{https://www.openfoam.com/}{OpenFOAM} (C++),
-\href{http://basilisk.fr}{Basilisk} (C) and
-\href{http://channelflow.org/}{Channelflow} (C++) and
-\href{https://www.code-saturne.org}{Code\_Saturn} (C/Fortran).
-%
-Big companies have started to use open-source development for fluid mechanics
-application. For example, Volkswagen group (Volkswagen, Audi, Seat, Porsche,
-Skoda, ...) has used OpenFOAM since
-2006\footnote{\url{https://www.nas.nasa.gov/assets/pdf/ams/%
-2014/AMS_20141202_Othmer.pdf}}.
-%
-EDF, the main French electric utility company has also a interesting
-\href{http://linuxfr.org/news/strat\%C3\%A9gie-open-source-\%C3\%A0-edf-rd}{%
-open-source strategy}.  It has produced many programs and in particular the CFD
-solver \href{https://www.code-saturne.org}{Code\_Saturn}.
-%
-EDF is also involved in the development of
-\href{https://www.paraview.org/}{Paraview}\footnote{\url{https://blog.kitware.com/introducing-some-new-paraview-5-0-features/}}.
-
-Finally, scientists have started to use Python to study fluid mechanics,
-especially through CFD codes (\href{http://dedalus-project.org/}{Dedalus},
-\href{https://github.com/spectralDNS}{SpectralDNS},
-\href{https://github.com/mikaem/Oasis}{Oasis}, \href{http://pyfr.org/}{PyFR}
-and \href{https://fenicsproject.org/}{FEniCS}) or through data analysis
-(\href{http://www.openptv.net/}{OpenPTV},
-\href{https://github.com/jr7/pypiv}{PyPIV}).
-%
-The packages of the FluidDyn project are also part of this trend.
-
-\subsection*{FluidDyn project: Breaking the status quo}
-
-We have seen that there is a strong dynamics in play around the use of
-computers (in particular with the web) and that this creates very efficient
-tools and methods for collective work and software development.
-%
-The potential of these tools is not fully exploited yet, in the field of fluid
-mechanics.
-
-FluidDyn is a project to foster open-science and open-source coding in Python
-in the field of fluid mechanics.
+FluidDyn is a project to foster open-science and open-source coding in Python in
+the field of fluid mechanics.
 %
 The project envisages to provide the technical framework to allow collaborative
-development of tools useful for the fluid mechanics community, and to do better
-science with open methods.
+development of tools useful for the fluid mechanics community, and to do science
+with open methods.
 %
@@
 \item Packaging and installation procedure. All packages are available from
-\href{https://pypi.org/}{Pypi} and can be install simply with
+\href{https://pypi.org/}{PyPI} and can be install simply with
 \href{https://pypi.org/project/pip/}{pip}, the standard Python installation
@@
 \item Licenses: depending on the packages, we choose to use the CeCILL-B or the
-CeCILL licenses. These licenses are a BSD compatible and a GPL compatible
-licenses adapted to both international and French legal
+CeCILL licenses. These licenses are compatible respectively with BSD and GPL
+licenses and adapted to both international and French legal
 matters\footnote{\url{http://www.cecill.info/licences.en.html}}.
@@
 \end{itemize}
-We hope that such clean framework will allow contributions from scientists in
-the field and that we can build together a nice, user-friendly and efficient
-ecosystem specialized in research and teaching in fluid dynamics.
+We hope that such a clean framework will facilitate contributions from
+scientists in the field and that we can build together a nice, user-friendly
+and efficient ecosystem specialized in research and teaching in fluid dynamics.
 
@@
 \begin{itemize}
-\item \fluidpack{dyn}: The base package which contains pure-python code that can
-be reused in scripts or in specialized FluidDyn packages. It also contains codes
-for miscellaneous command-line utilities useful for a typical fluid dynamics user.
+\item \fluidpack{dyn}\footnote{We use FluidDyn (with capital letters) to name the
+project and \fluiddyn for the base package.}: The base package which contains
+pure-python code that can be reused in scripts or in specialized FluidDyn
+packages. It also contains codes for miscellaneous command-line utilities useful
+for a typical fluid dynamics user.
 
@@
 
+\item \pack{Transonic}: a pure Python package to accelerate Python-Numpy code
+with Pythran and potentially other Python compilers.
+
 \item \fluidpack{fft}~\cite[see the companion paper][]{fluidfft}: a package which
-provides C++ and Python classes unifying various libaries to perform Fast Fourier
+provides C++ and Python classes unifying various libraries to perform Fast Fourier
 Transform (FFT) in sequential and in parallel.
@@
 Experiments using \fluidpack{lab} have been carried out in the DAMTP fluid
-laboratory \cite[Cambridge, UK. cf.][]{LeclercqPartridgeAugierDalzielKerswell2016}, in ENS Lyon laboratory
-\cite[Lyon, France][]{salort2018} and at LEGI \cite[Grenoble, France][]{ISSF2016}.
+laboratory \cite[Cambridge,
+UK. cf.][]{LeclercqPartridgeAugierDalzielKerswell2016}, in ENS Lyon laboratory
+\cite[Lyon, France,][]{salort2018} and at LEGI \cite[Grenoble,
+France,][]{ISSF2016}.
 
@@
 %
-Fluidimage is used to process images taken during experiments performed in the
+\fluidpack{image} is used to process images taken during experiments performed in the
 Coriolis platform at LEGI \cite[]{ISSF2016}.
@@
 documentations on the web and for \fluidpack{fft} and \fluidpack{sim} in the
-two companion papers \cite[]{fluidfft, fluidsim}.  This article will now focus
-on the base package \fluiddyn.
+two companion papers \cite[]{fluidfft, fluidsim}.
+%
+The code base was designed to follow Python 2.7 syntax during its genesis. Now,
+it has been made forward compatible with Python 3 through the use of external
+package \pack{future}.
+%
+We now tend to abandon Python 2.7 support for the next releases of the package.
+%
+This article will now focus on the base package \fluiddyn.
+
+\subsection*{API of the Python library \fluiddyn}
 
@@
 compilation.
-
-The code base was designed to follow Python 2.7 syntax during its genesis. Now,
-it has been made forward compatible with Python 3 through the use of external
-package future.
-
-\subsection*{API of the Python library \fluiddyn}
 
@@
 \item Support containing and printing documentation on the parameters.
-\item Support printing to console as XML and saving as XML, hdf5 and netcdf4
-files.
+\item Support printing to console as XML and saving as XML, HDF5 and NetCDF
+Files.
 \item Evaluate data types automatically while loading from saved files.
@@
 
-This modules provides classes to iterate over files. It is a common task in data
-processing to understand how to organized a \textit{serie} of files from filenames
-and formats. Then one has to form smaller sets of arrays contained in the
-files. For example, we can have such \textit{serie}: \codeinline{im1\_1.png,
-im1\_2.png, im1\_3.png, im2\_1.png, im2\_2.png, im2\_3.png} and we could create
-subsets like \codeinline{((im1\_1.png, im1\_2.png, im1\_2.png), (im2\_1.png,
-im2\_2.png, im2\_3.png))}. From the same \textit{serie} of images, we can also
-create other subsets \codeinline{((im1\_1.png, im2\_1.png), (im1\_2.png,
-im2\_2.png), (im1\_3.png, im2\_3.png))}.
+This module provides classes to iterate over files. It is a common task in data
+processing to understand how to organize a \textit{sequence} of files from
+filenames and formats. Typically, one has to form smaller sets of arrays
+contained in the files. For example, we may have a \textit{sequence} such as:
+
+\fbox{%
+\begin{minipage}{\textwidth}
+  \codeinline{%
+    (im1\_1.png, im1\_2.png, im1\_3.png, im2\_1.png, im2\_2.png, im2\_3.png)}
+\end{minipage}
+}
+
+from which we could create different subsets like,
+
+\fbox{%
+\begin{minipage}{0.5\textwidth}
+\codeinline{((im1\_1.png, im1\_2.png, im1\_3.png),\\
+  (im2\_1.png, im2\_2.png, im2\_3.png))}\\
+\end{minipage}
+}
+\fbox{%
+\begin{minipage}{0.47\textwidth}
+\codeinline{((im1\_1.png, im2\_1.png),\\
+  (im1\_2.png, im2\_2.png),\\
+  (im1\_3.png, im2\_3.png))}
+\end{minipage}
+}
 
@@
 
-This simple module makes simulataneous sequential and MPI programming a breeze
+This simple module makes simultaneous sequential and MPI programming a breeze
 by providing number of processes, \codeinline{nb\_proc = 1} and \codeinline{rank
@@
 excluded.
+
+\item \codeinline{fluidcluster-help}
+
+Tiny utility to print a short documentation on the most useful commands to
+interact with the scheduler of a HPC cluster.
 
@@
 For \fluiddyn, the code coverage results are displayed at
-\href{https://codecov.io/gh/fluiddyn/fluiddyn}{Codecov}.  Using third-party
+\href{https://codecov.io/bb/fluiddyn/fluiddyn}{Codecov}.  Using third-party
 packages \pack{coverage} and \pack{tox}, it is straightforward to bootstrap the
@@
 
-Python 2.7, 3.4 or above.
+Python 2.7, 3.4 or above. For the next versions, we will
+\href{https://python3statement.org/}{drop Python 2.7 support and Python $>=$
+3.6 will be required}.
 
@@
 \item {\bf Minimum:} \Numpy, \pack{Matplotlib}, \pack{psutil}, \pack{future},
-\pack{subprocess32} (for Python 2.7 only).
-
-\item {\bf Full functionality:} \pack{h5py}, \pack{h5netcdf}, \pack{mpi4py},
-\pack{Scipy}, \pack{pyfftw} (requires FFTW library), \pack{pillow}.
+\pack{subprocess32} (for Python 2.7 only), \pack{h5py}, \pack{h5netcdf}.
+
+\item {\bf Full functionality:} \pack{mpi4py}, \pack{Scipy}, \pack{pyfftw}
+(requires FFTW library), \pack{pillow}.
 
@@
 \fluiddyn; documentation, code coverage and continuous integration (Docker,
-Bitbucket pipeline and Travis).
-
-\item Cyrille Bonamy (LEGI): developer of \fluidpack{fft}, \fluidpack{sim} and
-\fluidpack{foam}. Main maintainer of \fluidpack{foam}.
+Bitbucket Pipelines and Travis CI).
+
+\item Cyrille Bonamy (LEGI): developer of \fluidpack{fft}, \fluidpack{sim},
+\fluidpack{image} and \fluidpack{foam}. Main maintainer of \fluidpack{foam}.
 
@@
 \end{itemize}
-%
-% \section*{Software location:}
-%
-% % {\bf Archive} \textcolor{blue}{(e.g. institutional repository, general
-% % repository) (required – please see instructions on journal website for
-% % depositing archive copy of software in a suitable repository)}
-%
-% \begin{description}[noitemsep,topsep=0pt]
-% \item[Name:] PyPI
-% \item[Persistent identifier:] https://pypi.org/project/fluiddyn
-% \item[Licence:] CeCILL-B, a BSD compatible French licence.
-% \item[Publisher:] Pierre Augier
-% \item[Version published:] 0.2.3
-% \item[Date published:] 02/05/2018
-% \end{description}
-%
-% {\bf Code repository}
-% % \textcolor{blue}{(e.g. SourceForge, GitHub etc.) (required)}
-%
-% \begin{description}[noitemsep,topsep=0pt]
-% \item[Name:] Bitbucket
-% \item[Persistent identifier:] https://bitbucket.org/fluiddyn/fluiddyn
-% \item[Licence:] CeCILL-B
-% \item[Date published:] 2015
-% \end{description}
+
+\section*{Software location:}
+
+% {\bf Archive} \textcolor{blue}{(e.g. institutional repository, general
+% repository) (required – please see instructions on journal website for
+% depositing archive copy of software in a suitable repository)}
+
+\begin{description}[noitemsep,topsep=0pt]
+\item[Name:] PyPI
+\item[Persistent identifier:] https://pypi.org/project/fluiddyn
+\item[Licence:] CeCILL-B, a BSD compatible French licence.
+\item[Publisher:] Pierre Augier
+\item[Version published:] 0.2.4
+\item[Date published:] 02/07/2018
+\end{description}
+
+{\bf Code repository}
+% \textcolor{blue}{(e.g. SourceForge, GitHub etc.) (required)}
+
+\begin{description}[noitemsep,topsep=0pt]
+\item[Name:] Bitbucket
+\item[Persistent identifier:] https://bitbucket.org/fluiddyn/fluiddyn
+\item[Licence:] CeCILL-B
+\item[Date published:] 2015
+\end{description}
 
@@
 
-% \section*{Language}
+\section*{Language}
 
@@
 
-% English
+English
 
@@
 \fluiddyn's versatility and generality. It can be used in other packages outside
-the FluidDyn project easily depending on the need. The command-lines tools can
+the FluidDyn project easily depending on the need. The command-line tools can
 be useful to all scientists working with Python. Other use cases could be:
@@
 Discussions and questions can be aired on instant messaging channels in Riot (or
-equivalent with Matrix protocol) at
+equivalent with Matrix protocol)\footnote{
 \url{%
-  https://matrix.to/#/#fluiddyn-users:matrix.org}
-or via IRC protocol on Freenode at \codeinline{\#fluiddyn-users}.
+  https://matrix.to/\#/\#fluiddyn-users:matrix.org}}
+or via IRC protocol on Freenode at \codeinline{\#fluiddyn-users}. Discussions
+can also be exchanged via the official mailing list\footnote{
+\url{https://www.freelists.org/list/fluiddyn}}.
 
@@
 We shall try, with this project, to explore the possibilities of open-source in
-science and fluid dynamics, by fully exploiting the new open-source tools and
-methods presented in the first subsections of the introduction.
-
-% todo: again mention some success of community-driven subject-area research
-% libraries? See the file ./communitydriven_research_soft.md
-
-The project is right now in a very preliminary stage.  Packages are actively
-evolving with interesting features and a framework for collaborative
-development, packaging, documenting and testing is now well set.
+science and fluid dynamics by fully exploiting the new open-source tools and
+methods.
+
+The project is right now in a preliminary stage.  Packages are actively evolving
+with interesting features and a framework for collaborative development,
+packaging, documenting and testing is now well set.
 %
@@
 \item Lack of recognition of the work spent in open-source, in particular for
-precarious scientists.
+scientists without permanent positions.
 \end{itemize}
@@
 %
-We can benefit from the dynamics of Python and of emerging subjects like deep
-learning and the Internet of Things.
+Python starts to be a standard tool in fluid dynamics, especially used for CFD
+(\href{http://dedalus-project.org/}{Dedalus},
+\href{https://github.com/spectralDNS}{SpectralDNS},
+\href{https://pypi.org/project/triflow/}{TriFlow},
+\href{http://pylbm.readthedocs.io}{PyLBM},
+\href{https://github.com/mikaem/Oasis}{Oasis}, \href{http://pyfr.org/}{PyFR},
+\href{https://fenicsproject.org/}{FEniCS},
+\href{http://elsa.onera.fr/Cassiopee/}{Cassiopee},
+\href{https://github.com/pyCGNS}{pyCGNS}, etc.) and data analysis
+(\href{http://www.openptv.net/}{OpenPTV},
+\href{https://github.com/jr7/pypiv}{PyPIV}).
+%
+Moreover, we can benefit from the dynamics of Python and of emerging subjects like
+deep learning and the Internet of Things.
 %
@@
 %
-We thank Cyrille Bonamy, Antoine Campagne, Miguel Calpe Linares and Julien
-Salort for their implications in the development of some of the FluidDyn
-packages.
-%
-We thank Bitbucket for providing us for free a high quality forge compatible with
-Mercurial.
-%
-Finally, we thank Gabriel Moreau and Jo\"el Sommeria for their constant willing
+We thank Antoine Campagne, Miguel Calpe Linares and Julien Salort for their
+implications in the development of some of the FluidDyn packages.
+%
+We are grateful to Bitbucket for providing us with a high quality forge
+compatible with Mercurial, free of cost.
+%
+Finally, we thank Gabriel Moreau and Jo\"el Sommeria for their constant will
 to discuss and share their knowledge.
@@
 
-This project has indirectly benefit from funding from the foundation Simone et
+This project has indirectly benefited from funding from the foundation Simone et
 Cino Del Duca de l'Institut de France, the European Research Council (ERC)
@@
 %
-We have also been able to use supercomputers of CIMENT/GRICAD, CINES/GENCI and
-the Swedish National Infrastructure for Computing (SNIC).
-
-% \section*{Competing interests}
+We have also been able to use supercomputers of CIMENT/GRICAD, CINES/GENCI
+(grant 2018-A0040107567) and the Swedish National Infrastructure for Computing
+(SNIC).
+
+\section*{Competing interests}
 
@@
 
-% The authors declare that they have no competing interests.
+The authors declare that they have no competing interests.
 
@@
 % \\ http://dx.doi.org/10.1371/journal.pone.0018657.}
-%
-% \bibliographystyle{../agsm}
-% \bibliography{../bib}
-
-% \rule{\textwidth}{1pt}
-%
-% { \bf Copyright Notice} \\
-% Authors who publish with this journal agree to the following terms: \\
-%
-% Authors retain copyright and grant the journal right of first publication with
-% the work simultaneously licensed under a
-% \href{http://creativecommons.org/licenses/by/3.0/}{Creative Commons Attribution
-% License} that allows others to share the work with an acknowledgement of the
-% work's authorship and initial publication in this journal.
-%
-% Authors are able to enter into separate, additional contractual arrangements
-% for the non-exclusive distribution of the journal's published version of the
-% work (e.g., post it to an institutional repository or publish it in a book),
-% with an acknowledgement of its initial publication in this journal.
-%
-% By submitting this paper you agree to the terms of this Copyright Notice, which
-% will apply to this submission if and when it is published by this journal.
-%
-%
+
+\bibliographystyle{../agsm}
+\bibliography{../bib}
+
+\rule{\textwidth}{1pt}
+
+{ \bf Copyright Notice} \\
+Authors who publish with this journal agree to the following terms: \\
+
+Authors retain copyright and grant the journal right of first publication with
+the work simultaneously licensed under a
+\href{http://creativecommons.org/licenses/by/3.0/}{Creative Commons Attribution
+License} that allows others to share the work with an acknowledgement of the
+work's authorship and initial publication in this journal.
+
+Authors are able to enter into separate, additional contractual arrangements
+for the non-exclusive distribution of the journal's published version of the
+work (e.g., post it to an institutional repository or publish it in a book),
+with an acknowledgement of its initial publication in this journal.
+
+By submitting this paper you agree to the terms of this Copyright Notice, which
+will apply to this submission if and when it is published by this journal.
+
+\end{document}

### 2019-02-13 paper_01_fluidfft/article.tex
Add fluiddyn papers nearly as it is

%% Journal of Open Research Software Latex template -- Created By Stephen
%% Bonner and John Brennan, Durham University, UK.
%% see http://openresearchsoftware.metajnl.com

\documentclass{../jors}

\begin{document}

{\bf Software paper for submission to the Journal of Open Research Software} \\

To complete this template, please replace the blue text with your own. The
paper has three main sections: (1) Overview; (2) Availability; (3) Reuse
potential.

Please submit the completed paper to: editor.jors@ubiquitypress.com

\rule{\textwidth}{1pt}

\section*{(1) Overview}

\vspace{0.5cm}

\section*{Title}

% \textcolor{blue}{The title of the software paper should focus on the software,
% e.g. “Text mining software from the X project”. If the software is closely
% linked to a specific research paper, then “Software from Paper Title” is
% appropriate. The title should be factual, relating to the functionality of the
% software and the area it relates to rather than making claims about the
% software, e.g. “Easy-to-use”.}

FluidFFT: common API (C++ and Python) for Fast Fourier Transform HPC libraries

\section*{Paper Authors}

% \textcolor{blue}{1. Last name, first name; (Lead/corresponding author first) \\
% 2. Last name, first name; etc.}

1. MOHANAN Ashwin Vishnu$^a$\\
2. BONAMY Cyrille$^b$\\
3. AUGIER Pierre$^b$\\

\smallskip

$^a$ Linn\'e Flow Centre, Department of Mechanics, KTH, 10044 Stockholm, Sweden.
$^b$ Univ. Grenoble Alpes, CNRS, Grenoble INP\footnote{Institute of Engineering
Univ. Grenoble Alpes}, LEGI, 38000 Grenoble, France.\\

\section*{Paper Author Roles and Affiliations}
% \textcolor{blue}{1. First author role and affiliation \\
% 2. Second author role and affiliation etc.}

1. Ph.D. student, Linn\'e Flow Centre, KTH Royal Institute of Technology,
Sweden; \\
2. Research Engineer, LEGI, Universit\'e Grenoble Alpes, CNRS, France; \\
3. Researcher, LEGI, Universit\'e Grenoble Alpes, CNRS, France

\section*{Abstract}

% \textcolor{blue}{A short (ca. 100 word) summary of the software being
% described: what problem the software addresses, how it was implemented and
% architected, where it is stored, and its reuse potential.}

The Python package \fluidpack{fft} provides a common Python API for performing
Fast Fourier Transforms (FFT) in sequential, in parallel and on GPU with different
FFT libraries (FFTW, P3DFFT, PFFT, cuFFT). \fluidpack{fft} is a comprehensive FFT
framework which allows Python users to easily and efficiently perform FFT and the
associated tasks, such as as computing linear operators and energy spectra.
%
We describe the architecture of the package composed of C++ and Cython FFT
classes, Python ``operator'' classes and Pythran functions.
%
The package supplies utilities to easily test itself and benchmark the different
FFT solutions for a particular case and on a particular machine.
%
We present a performance scaling analysis on three different computing clusters
and a microbenchmark showing that \fluidpack{fft} is an interesting solution to
write efficient Python applications using FFT.

\section*{Keywords}

% \textcolor{blue}{keyword 1; keyword 2; etc. \\
% Keywords should make it easy to identify who and what the software will be
% useful for.}

Free and open-source library; Python; Fast Fourier Transform; Distributed; MPI;
GPU; High performance computing%

\section*{Introduction}

% \textcolor{blue}{An overview of the software, how it was produced, and the
% research for which it has been used, including references to relevant research
% articles. A short comparison with software which implements similar
% functionality should be included in this section. }

Fast Fourier Transform (FFT) is a class of algorithms used to calculate the
discrete Fourier transform, which traces back its origin to the groundbreaking
work by \citet{cooley_tukey}.
%
Ever since then, FFT as a computational tool has been applied in multiple
facets of science and technology, including digital signal processing, image
compression, spectroscopy, numerical simulations and scientific computing in
general. There are many good libraries to perform FFT, in particular the
\emph{de-facto} standard \libpack{FFTW} \citep{frigo2005design}.\@ A challenge
is to efficiently scale FFT on clusters with the memory distributed over a
large number of cores using Message Passing Interface (MPI). This is imperative
to solve big problems faster and when the arrays do not fit in the memory of
single computational node.
%
A problem is that for one-dimensional FFT, all the data have to be located in the
memory of the process that perform the FFT, so a lot of communications between
processes are needed for 2D and 3D FFT.

To elaborate, there is only one way to apply domain decomposition for 2D FFT,
which is to split them into narrow strips across one dimension. However for 3D
FFT, there are two strategies to distribute an array in the memory, the 1D (or
\emph{slab}) decomposition and the 2D (or \emph{pencil}) decomposition. The 1D
decomposition is more efficient when only few processes are used but suffers
from an important limitation in terms of number of MPI processes that can be
used. Utilizing 2D decomposition overcomes this limitation.

Some of the well-known libraries are written in C, C++ and Fortran. The classical
\libpack{FFTW} library supports MPI using 1D decomposition and hybrid parallelism
using MPI and OpenMP. Other libraries, now implement the 2D decomposition for
FFT over 3D arrays: \libpack{PFFT} \citep{pippig_pfft2013}, \libpack{P3DFFT}
\citep{pekurovsky2012p3dfft}, \libpack{2decomp\&FFT} and so on. These libraries
rely on MPI for the communications between processes, are optimized for
supercomputers and scales well to hundreds of thousands of cores. However, since
there is no common API, it is not simple to write applications that are able to
use these libraries and to compare their performances. As a result, developers are
met with a hard decision, which is to choose a library before the code is
implemented.

Apart from CPU-based parallelism, General Purpose computing on Graphical
Processing Units (GPGPU) is also gaining traction in scientific computing.
Scalable libraries written for GPGPU such as OpenCL and CUDA have emerged, with
their own FFT implementations, namely \libpack{clFFT} and \libpack{cuFFT}
respectively.

% As explained in the companion paper \citet{fluiddyn},
Python can easily link these libraries through compiled extensions. For a Python
developer, the following packages leverage this approach to perform FFT:

\begin{outline}
  \1 sequential FFT, using:
    \2 \pack{numpy.fft} and \pack{scipy.fftpack} which are essentially
    C and Fortran extensions for \libpack{FFTPACK} library.
    \2 \pack{pyFFTW} which wraps \libpack{FFTW} library and provides interfaces similar to
    the \pack{numpy.fft} and \pack{scipy.fftpack} implementations.
    \2 \pack{mkl\_fft}, which wraps Intel's \libpack{MKL} library and exposes python
    interfaces to act as drop-in replacements for \pack{numpy.fft} and
    \pack{scipy.fftpack}.
  \1 FFT with MPI, using:
    \2 \pack{mpiFFT4py} and \pack{mpi4py-fft} built on top of \pack{pyFFTW} and
    \pack{numpy.fft}.
    \2 \pack{pfft-python} which provides extensions for
    PFFT library.
  \1 FFT with GPGPU, using:
    \2 \pack{Reikna}, a pure python package which depends on \pack{PyCUDA}
    and \pack{PyOpenCL}
    \2 \pack{pytorch-fft}: provides C extensions for cuFFT, meant to work with
    PyTorch, a tensor library similar to NumPy.
\end{outline}

Although these Python packages are under active development, they suffer from
certain drawbacks:

\begin{itemize}
  \item No effort so far to consolidate sequential, MPI and GPGPU based FFT
  libraries under a single package with similar syntax.

  \item Quite complicated even for the simplest use case scenarios. To
  understand how to use them, a novice user has to, at least, read the
  \libpack{FFTW} documentation.

  \item No benchmarks between libraries and between the Python
  solutions and solutions based only on a compiled language (as C, C++ or
  Fortran).

  \item Provides just the FFT and inverse FFT functions, no associated
  mathematical operators.

\end{itemize}

The Python package \fluidpack{fft} fills this gap by providing C++ classes and
their Python wrapper classes for performing simple and common tasks with different
FFT libraries. It has been written to make things easy while being as efficient as
possible. It provides:

\begin{itemize}
\item tests,

\item documentation and tutorials,

\item benchmarks,

\item operators for simple tasks (for example, compute the energy or the
gradient of a field).

\end{itemize}

In the present article, we shall start by describing the implementation of
\fluidpack{fft} including its design aspects and the code organization. Thereafter,
we shall compare the performance of different classes in \fluidpack{fft} in
three computing clusters, and also describe, using microbenchmarks, how a Python
function can be optimized to be as fast as a Fortran implementation. Finally,
we show how we test and maintain the quality of the code base through
continuous integration and mention some possible applications of
\fluidpack{fft}.

\section*{Implementation and architecture}

% \textcolor{blue}{How the software was implemented, with details of the
% architecture where relevant. Use of relevant diagrams is appropriate. Please
% also describe any variants and associated implementation differences.}
The two major design goals of \fluidpack{fft} are:
\begin{itemize}
 \item to support multiple FFT libraries under the same umbrella and expose the
 interface for both C++ and Python code development.
 \item to keep the design of the interfaces as human-centric and easy to use as
 possible, without sacrificing performance.
\end{itemize}

Both C++ and Python APIs provided by \fluidpack{fft} currently support linking
with \libpack{FFTW} (with and without MPI and OpenMP support enabled),
\libpack{MKL}, \libpack{PFFT}, \libpack{P3DFFT}, \libpack{cuFFT} libraries. The
classes in \fluidpack{fft} offers API for performing
double-precision\footnote{Most C++ classes also support single-precision.}
computation with real-to-complex FFT, complex-to-real inverse FFT, and additional
helper functions.

\subsection*{C++ API}

\begin{figure}[htp]
  \centering
  \includegraphics[width=\linewidth]{Pyfig/fig_classes}
  \caption{Class hierarchy demonstrating object-oriented approach. The
    sequential classes are shown in red, the CUDA-based classes in magenta and
    the MPI-based classes in green. The arrows represent inheritance from
    parent to child class.
  }\label{fig:classes}
\end{figure}

The C++ API is implemented as a hierarchy of classes as shown in
Fig.~\ref{fig:classes}.
%
The naming convention used for the classes (\codeinline{<Type of FFT>With<Name of
Library>}) is a cue for how these are functioning internally.
%
By utilizing inheritance, the classes share the same function names and syntax
defined in the \emph{base} classes, shown in white boxes in
Fig.~\ref{fig:classes}. Some examples of such functions are:

\begin{itemize}
  \item \codeinline{alloc\_array\_X}: Allocates array to store a physical array
    with real datatype for the current process.
  \item \codeinline{alloc\_array\_K}: Allocates array to store a spectral array
    with complex datatype  for the current process.
  \item \codeinline{init\_array\_X\_random}: Allocates and initializes a physical
    array with random values.
  \item \codeinline{test}: Run tests for a class by comparing mean and mean energy
    values in an array before and after a set of \codeinline{fft} and
    \codeinline{ifft} calls.
  \item \codeinline{bench}: Benchmark the \codeinline{fft} and
    \codeinline{ifft} methods for certain number of iterations.
\end{itemize}

Remaining methods which are specific to a library are defined in the
corresponding child classes, depicted in coloured boxes in
Fig.~\ref{fig:classes}, for example:

\begin{itemize}
  \item \codeinline{are\_parameters\_bad}: Verifies whether the global array
    shape can be decomposed with the number of MPI processes available or not.
    If the parameters are compatible, the method returns \codeinline{false}.
    This method is called prior to initializing the class.
  \item \codeinline{fft} and \codeinline{ifft}: Forward and inverse FFT
    methods.
\end{itemize}

Let us illustrate with a trivial example, wherein we initialize the FFT with a
random physical array, and perform a set of \codeinline{fft} and \codeinline{ifft}
operations.
\begin{minted}[fontsize=\footnotesize]{cpp}
#include <iostream>
using namespace std;

#include <fft3dmpi_with_fftwmpi3d.h>
// #include <fft3dmpi_with_p3dfft.h>

int main(int argc, char **argv) {
  int N0 = N1 = N2 = 32;
  // MPI-related
  int nb_procs = 4;
  MPI_Init(&argc, &argv);
  MPI_Comm_size(MPI_COMM_WORLD, &(nb_procs));

  myreal* array_X;
  mycomplex* array_K;

  FFT3DMPIWithFFTWMPI3D o(N0, N1, N2);
  // FFT3DMPIWithP3DFFT o(N0, N1, N2);

  o.init_array_X_random(array_X);
  o.alloc_array_K(array_K);
  o.fft(array_X, array_K);
  o.ifft(array_K, array_X);
  MPI_Finalize();
  return 0;
}
\end{minted}

As suggested through comments above, in order to switch the FFT library, the
user only needs to change the header file and the class name. An added
advantage is that, the user does not need to bother about the domain
decomposition while declaring and allocating the arrays. A few more helper
functions are available with the FFT classes, such as functions to compute the
mean value and energies in the array. These are demonstrated with examples in
the documentation.\footnote{%
\url{https://fluidfft.readthedocs.io/en/latest/examples/cpp.html}.}
%
Detailed information regarding the C++ classes and its member functions are
also included in the online documentation\footnote{%
\url{https://fluidfft.readthedocs.io/en/latest/doxygen/index.html}.}.

\subsection*{Python API} Similar to other packages in the FluidDyn project,
\fluidpack{fft} also uses an object-oriented approach, providing FFT classes.
%
This is in contrast with the approach adopted by \pack{numpy.fft} and \pack{%
scipy.fftpack} which provides functions instead, with which the user has to
figure out the procedure to design the input values and to use the return
values, from the documentation.
%
In \fluidpack{fft}, the Python API wraps all the functionalities of its C++
counterpart and offers a richer experience through an accompanying
operator class.

As a short example, let us try to calculate the gradient of a plane sine-wave
using spectral methods, mathematically described as follows:

\begin{align*}
  u(x,y) &=
    \sin(x + y) &\forall x,y \in \left[0, L \right] \\
  \hat u(k_x,k_y) &=
    \frac{1}{L^2}
    \int_0^{L}\int_0^{L}
    u(x,y) \exp(ik_x x + ik_y y) dx dy \\
  \nabla u(x,y) &=
    \sum_{k_x} \sum_{k_y}
    i\mathbf{k}
    \hat u(k_x,k_y) \exp(-ik_x x - ik_y y)
\end{align*}
%
where $k_x$, $k_y$ represent the wavenumber corresponding to $x$ and $y$ directions,
and $\mathbf{k}$ is the wavenumber vector.

The equivalent pseudo-spectral implementation in \fluidpack{fft} is as follows:
\begin{minted}[fontsize=\footnotesize]{python}
  from fluidfft.fft2d.operators import OperatorsPseudoSpectral2D, pi
  from numpy import sin

  nx = ny = 100
  lx = ly = 2 * pi

  oper = OperatorsPseudoSpectral2D(nx, ny, lx, ly, fft="fft2d.with_fftw2d")

  u = sin(oper.XX + oper.YY)
  u_fft = oper.fft(u)
  px_u_fft, py_u_fft = oper.gradfft_from_fft(u_fft)
  px_u = oper.ifft(px_u_fft)
  py_u = oper.ifft(py_u_fft)
  grad_u = (px_u, py_u)
\end{minted}

A parallelized version of the code above will work out of the box, simply by
replacing the FFT class with an MPI-based FFT class, for instance
\codeinline{fft2d.with\_fftwmpi2d}. One can also let \fluidpack{fft} automatically
choose an appropriate FFT class by instantiating the operator class with
\codeinline{fft=None} or \codeinline{fft="default"}. Even if one finds the methods
in the operator class to be lacking, one can inherit the class and easily create a
new method, for instance using the wavenumber arrays, \codeinline{oper.KX} and
\codeinline{oper.KY}.  Arguably, a similar implementation with other available
packages would require the know-how on how FFT arrays are allocated in the memory,
normalized, decomposed in parallel and so on.
%
Moreover, the FFT and the operator classes contain objects describing the shapes
of the real and complex arrays and how the data is shared between processes.
%
A more detailed introduction on how
to use \fluidpack{fft} and available functions can be found in the
tutorials\footnote{%
\url{https://fluidfft.readthedocs.io/en/latest/tutorials.html}.}.

Thus, we have demonstrated how, by using \fluidpack{fft}, a developer can
easily switch between FFT libraries.
%
Let us now turn our attention to how the code is organized. We shall also describe
how the source code is built, and linked with the supported libraries.

\subsection*{Code organization}
\begin{figure}[htp]
  \centering
  \includegraphics[width=0.96\linewidth]{Pyfig/fig_build_use}
  \caption{Flowchart illustrating how the C++ and Python API are built and used
  for one particular class, viz. \codeinline{FFT2DMPIWithFFTWMPI2D}. The dotted
  arrows in C++ part stand for include statements, demonstrating the class
  hierarchy and in the Python part indicate how different codes are imported. On
  the bottom, a smaller flowchart demonstrates how to use the API by writing user
  code.  }\label{fig:build_use}
\end{figure}

The internal architecture of \fluidpack{fft} can be visualized as layers.  Through
Fig.~\ref{fig:build_use}, we can see how these layers are linked together forming
the API for C++ and Python development. For simplicity, only one FFT class is
depicted in the figure, namely \codeinline{FFT2DMPIWithFFTWMPI2D}, which wraps
\libpack{FFTW}'s parallelized 2D FFT implementation. The C++ API is accessed by
importing the appropriate header file and building the user code with a Makefile,
an example of which is available \href{%
https://fluidfft.readthedocs.io/en/latest/examples/cpp.html}{%
in the documentation}.

The Python API is built automatically when \fluidpack{fft} is
installed\footnote{%
\href{https://fluidfft.readthedocs.io/en/latest/install.html}{Detailed steps
for installation} are provided in the documentation.}.
%
It first generates the Cython source code as a pair of \codeinline{.pyx} and
\codeinline{.pxd} files containing a class wrapping its C++
counterpart\footnote{Uses an approach similar to guidelines \href{%
    https://cython.readthedocs.io/en/latest/src/userguide/wrapping_CPlusPlus.html}{%
``Using C++ in Cython''} in the Cython documentation.}.
%
The Cython files are produced from template files (specialized for the 2D and
3D cases) using the template library \mako.
%
Thereafter, \pack{Cython} \citep{behnel_cython2011} generates C++ code with
necessary Python bindings, which are then built in the form of extensions or
dynamic libraries importable in Python code. All the built extensions are then
installed as a Python package named \fluidpack{fft}.

A helper function \codeinline{fluidfft.import\_fft\_class} is provided with the
package to simply import the FFT class. However, it is more convenient and
recommended to use an operator class, as described in the example for Python
API.\@ Although the operator classes can function as pure Python code, some of
its critical methods can be compiled, if \pack{Pythran}
\citep{guelton2018pythran} is available during installation of
\fluidpack{fft}. We will show towards the end of this section that by using
\pack{Pythran}, we reach the performance of the equivalent Fortran code.

To summarize, \fluidpack{fft} consists of the following layers:
\begin{itemize}

\item One C++ class per FFT library derived from a hierarchy of C++ classes
as shown in Fig.~\ref{fig:classes}.

\item \pack{Cython} wrappers of the C++ classes with their unit test cases.

\item Python operator classes (2D and 3D) to write code independently of the
library used for the computation of the FFT and with some mathematical helper
methods. These classes are accompanied by unit test cases.

\item \pack{Pythran} functions to speedup critical methods in the Python
operator classes.

\end{itemize}

Command-line utilities (\codeinline{fluidfft-bench} and
\codeinline{fluidfft-bench-analysis}) are also provided with the \fluidpack{fft}
installation to run benchmarks and plot the results. In the next subsection, we
shall look at some results by making use of these utilities on three computing
clusters.

\subsection*{Performance}

\subsubsection*{Scalability tests using \codeinline{fluidfft-bench}}

% Simple!! Few cases. Few clusters. Figures obtained with
% fluidfft-bench-analysis

Scalability of \fluidpack{fft} is measured in the form of strong scaling speedup,
defined in the present context as:
\begin{equation*}
S_\alpha(n_p) = \frac
{[\mathrm{Time\ elapsed\ for\ } N \mathrm{\ iterations\ with\ }n_{p,\min}\mathrm{\ processes}]_{\mathrm{fastest}}
\times n_{p,\min}}
{[\mathrm{Time\ elapsed\ for\ } N \mathrm{\ iterations\ with\ } n_p \mathrm{\
processes}]_\alpha}
\label{eq:speedup}
\end{equation*}

where $n_{p,\min}$ is the minimum number of processes employed for a specific
array size and hardware. The subscripts, $\alpha$ denotes the FFT class used and
``fastest'' corresponds to the fastest result among various FFT classes.

To compute strong scaling the utility \codeinline{fluidfft-bench} is launched
as scheduled jobs on HPC clusters, ensuring no interference from background
processes. No hyperthreading was used.
%
We have used $N=20$ iterations for each run, with which we obtain sufficiently
repeatable results.
%
For a particular choice of array size, every FFT class available are
benchmarked for the two tasks, forward and inverse FFT. Three different function
variants are compared (see the legend in subsequent figures):

\begin{itemize}

\item \codeinline{fft\_cpp}, \codeinline{ifft\_cpp} (continuous lines): benchmark
of the C++ function from the C++ code. An array is passed as an argument to store
the result. No memory allocation is performed inside these functions.

\item \codeinline{fft\_as\_arg}, \codeinline{ifft\_as\_arg} (dashed lines):
benchmark of a Python method from Python. Similar to the C++ code, the second
argument of this method is an array to contain the result of the transform, so no
memory allocation is needed.

\item \codeinline{fft\_return}, \codeinline{ifft\_return} (dotted lines):
benchmark of a Python method from Python. No array is provided to the function to
contain the result, and therefore a numpy array is created and then returned by
the function.

\end{itemize}

On big HPC clusters, we have only focussed on 3D array transforms as benchmark
problems, since these are notoriously expensive to compute and require massive
parallelization.  The physical arrays used in all four 3D MPI based FFT classes
are identical in structure.  However, there are subtle differences, in terms of
how the domain decomposition and the allocation of the transformed array in the
memory are handled\footnote{Detailed discussion on \href{%
https://fluidfft.readthedocs.io/en/latest/ipynb/executed/tuto_fft3d_mpi_domain_decomp.html}{%
``FFT 3D parallel (MPI): Domain decomposition''} tutorial}.

Hereafter, for the sake of brevity, the FFT classes will be named in terms of the
associated library (For example, the class \codeinline{FFT3DMPIWithFFTW1D} is
named \codeinline{fftw1d}).  Let us go through the results\footnote{Saved at
\url{%
https://bitbucket.org/fluiddyn/fluidfft-bench-results}} plotted using
\codeinline{fluidfft-bench-analysis}.

\paragraph{Benchmarks on Occigen}

\href{https://www.top500.org/system/178465}{Occigen} is a GENCI-CINES HPC
cluster which uses Intel Xeon CPU E5--2690 v3 (2.6 GHz) processors with 24 cores
per node. The installation was performed using Intel C++ 17.2 compiler, Python
3.6.5, and OpenMPI 2.0.2.

\begin{figure}[htp!]
\centering
\includegraphics[width=\linewidth]{tmp/fig_occigen_384x1152x1152}
\caption{Speedup computed from the median of the elapsed times for 3D fft
(384$\times$1152$\times$1152, left: fft and right: ifft) on Occigen.}%
\label{fig:occigen384x1152x1152}
\end{figure}

Fig.~\ref{fig:occigen384x1152x1152} demonstrates the strong scaling performance
of a cuboid array sized $384\times1152\times1152$. This case is particularly
interesting since for FFT classes implementing 1D domain decomposition
(\codeinline{fftw1d} and \codeinline{fftwmpi3d}), the processes are spread on
the first index for the physical input array. This restriction is as a result
of some \libpack{FFTW} library internals and design choices adopted in
\fluidpack{fft}. This limits \codeinline{fftw1d} (our own MPI implementation
using MPI types and 1D transforms from FFTW) to 192 cores and
\codeinline{fftwmpi3d} to 384 cores. The latter can utilize more cores since it
is capable of working with empty arrays, while sharing some of the
computational load.
%
The fastest methods for relatively
low and high number of processes are \codeinline{fftw1d} and
\codeinline{p3dfft} respectively for the present case.

The benchmark is not sufficiently accurate to measure the cost of calling the
functions from Python (difference between continuous and dashed lines,
i.e. between pure C++ and the \codeinline{as\_arg} Python method) and even the
creation of the numpy array (difference between the dashed and the dotted line,
i.e. between the \codeinline{as\_arg} and the \codeinline{return} Python
methods).

\begin{figure}[htp!]
\centering
\includegraphics[width=\linewidth]{tmp/fig_occigen_1152x1152x1152}
\caption{Speedup computed from the median of the elapsed times for 3D fft
(1152$\times$1152$\times$1152, left: fft and right: ifft) on Occigen.}
\label{fig:occigen1152x1152x1152}
\end{figure}

Fig.~\ref{fig:occigen1152x1152x1152} demonstrates the strong scaling
performance of a cubical array sized $1152\times1152\times1152$. For this
resolution as well, \codeinline{fftw1d} is the fastest method when using only
few cores and it can not be used for more that 192 cores. The faster library
when using more cores is also \codeinline{p3dfft}. This also shows that
\fluidpack{fft} can effectively scale for over 10,000 cores with a significant
increase in speedup.

\paragraph{Benchmarks on Beskow}

\href{ https://www.pdc.kth.se/hpc-services/computing-systems}{Beskow} is a Cray
machine maintained by SNIC at PDC, Stockholm. It runs on Intel(R) Xeon(R) CPU
E5-2695 v4 (2.1 GHz) processors with 36 cores per node. The installation was
done using Intel C++ 18 compiler, Python 3.6.5 and CRAY-MPICH 7.0.4.

\begin{figure}[htp!]
\centering
\includegraphics[width=\linewidth]{tmp/fig_beskow_384x1152x1152}
\caption{Speedup computed from the median of the elapsed times for 3D fft
(384$\times$1152$\times$1152, left: fft and right: ifft) on Beskow.}
\label{fig:beskow384x1152x1152}
\end{figure}

In Fig.~\ref{fig:beskow384x1152x1152}, the strong scaling results of the cuboid
array can be observed. In this set of results we have also included intra-node
scaling, wherein there is no latency introduced due to typically slower
node-to-node communication. The fastest library for very low (below 16) and
very high (above 384) number of processes in this configuration is
\codeinline{p3dfft}. For moderately high number of processes (16 and above) the
fastest library is \codeinline{fftwmpi3d}. Here too, we notice that
\codeinline{fftw1d} is limited to 192 cores and \codeinline{fftwmpi3d} to 384
cores, for reasons mentioned earlier.

A striking difference when compared with Fig.~\ref{fig:occigen384x1152x1152} is
that \codeinline{fftw1d} is not the fastest of the four classes in this machine.
One can only speculate that this could be a consequence of the differences in MPI
library and hardware which has been employed. This also emphasises the need to
perform benchmarks when using an entirely new configuration.

\begin{figure}[htp!]
\centering
\includegraphics[width=\linewidth]{tmp/fig_beskow_1152x1152x1152}
\caption{Speedup computed from the median of the elapsed times for 3D fft
(1152$\times$1152$\times$1152, left: fft and right: ifft) on Beskow.}
\label{fig:beskow1152x1152x1152}
\end{figure}

The strong scaling results of the cubical array on Beskow are displayed on
Fig.~\ref{fig:beskow1152x1152x1152}, wherein we restrict to inter-node
computation.  We observe that the fastest method for low number of processes is
again, \codeinline{fftwmpi3d}. When high number of processes (above 1000)
are utilized, initially \codeinline{p3dfft} is the faster methods as before,
but with 3000 and above processes, \codeinline{pfft} is comparable in speed and
sometimes faster.

\paragraph{Benchmarks on a LEGI cluster}

Let us also analyse how \fluidpack{fft} scales on a computing cluster
maintained at an institutional level, named Cluster8 at \href{%
http://www.legi.grenoble-inp.fr}{LEGI}, Grenoble. This cluster functions using
Intel Xeon CPU E5-2650 v3 (2.3 GHz) with 20 cores per node and \fluidpack{fft}
was installed using a toolchain which comprises of gcc 4.9.2, Python 3.6.4 and
OpenMPI 1.6.5 as key software components.

\begin{figure}[htp!]
\centering
\includegraphics[width=\linewidth]{tmp/fig_legi_cluster8_320x640x640}
\caption{Speedup computed from the median of the elapsed times for 3D fft
(320$\times$640$\times$640) at LEGI on cluster8.}
\label{fig:cluster8:320x640x640}
\end{figure}

In Fig.~\ref{fig:cluster8:320x640x640} we observe that the strong scaling for an
array shape of $320\times640\times640$ is not far from the ideal linear trend. The
fastest library is \codeinline{fftwmpi3d} for this case.  As expected from FFT
algorithms, there is a slight drop in speedup when the array size is not exactly
divisible by the number of processes, i.e.\ with 12 processes. The speedup
declines rapidly when more than one node is employed (above 20 processes). This
effect can be attributed to the latency introduced by inter-node communications, a
hardware limitation of this cluster (10 Gb/s).

\begin{figure}[htp!]
\centering
\includegraphics[width=\linewidth]{tmp/fig_legi_cluster8_2160x2160}
\caption{Speedup computed from the median of the elapsed times for 2D fft
(2160$\times$2160) at LEGI on cluster8.}
\label{fig:cluster8:2160x2160}
\end{figure}

We have also analysed the performance of 2D MPI enabled FFT classes on the same
machine using an array shaped $2160\times2160$ in
Fig.~\ref{fig:cluster8:2160x2160}. The fastest library is
\codeinline{fftwmpi2d}. Both \codeinline{fftw1d} and \codeinline{fftwmpi2d}
libraries display near-linear scaling, except when more than one node is used
and the performance tapers off.

As a conclusive remark on scalability, a general rule of thumb should be to use
1D domain decomposition when only very few processors are employed. For massive
parallelization, 2D decomposition is required to achieve good speedup without
being limited by the number of processors at disposal. We have thus shown that
overall performance of the libraries interfaced by \fluidpack{fft} are quite
good, and there is no noticeable drop in speedup when the Python API is used.
%
This benchmark analysis also shows that the fastest FFT implementation depends
on the size of the arrays and on the hardware.
%
Therefore, an application build upon \fluidpack{fft} can be efficient for
different sizes and machines.

\subsubsection*{Microbenchmark of critical ``operator'' functions}

As already mentioned, we use \pack{Pythran} \citep{guelton2018pythran} to
compile some critical ``operator'' functions.  In this subsection, we present a
microbenchmark for one simple task used in pseudo-spectral codes: projecting a
velocity field on a non-divergent velocity field.  It is performed in spectral
space, where it can simply be written as
\begin{minted}[fontsize=\footnotesize]{python}
# pythran export proj_out_of_place(
#     complex128[][][], complex128[][][], complex128[][][],
#     float64[][][], float64[][][], float64[][][], float64[][][])

def proj_out_of_place(vx, vy, vz, kx, ky, kz, inv_k_square_nozero):
    tmp = (kx * vx + ky * vy + kz * vz) * inv_k_square_nozero
    return vx - kx * tmp, vy - ky * tmp, vz - kz * tmp
\end{minted}
Note that, this implementation is ``out-of-place'', meaning that the result is
returned by the function and that the input velocity field (\codeinline{vx, vy,
vz}) is unmodified.
%
The comment above the function definition is a \pack{Pythran} annotation, which
serves as a type-hint for the variables used within the functions --- all
arguments being \pack{Numpy} arrays in this case.
%
\pack{Pythran} needs such annotation to be able to compile this code into
efficient machine instructions \emph{via} a C++ code.
%
Without \pack{Pythran} the annotation has no effect, and
of course, the function defaults to using Python with \pack{Numpy} to execute.

The array notation is well adapted and less verbose to express this simple
vector calculus.
%
Since explicit loops with indexing is not required, the computation with Python
and \pack{Numpy} is not extremely slow. Despite this being quite a favourable
case for \pack{Numpy}, the computation with \pack{Numpy} is not optimized
because, internally, it involves many loops (one per arithmetic operator) and
creation of temporary arrays.
%av: Have I understood correctly here with the clarifications: "internally" &
%   "arithmetic operator"?

\begin{figure}[htp]
\centering
\includegraphics[width=\linewidth]{tmp/fig_microbench}
\caption{Elapsed time (smaller is better) for the projection function for
different implementations and tools.  The shape of the arrays is
$(128,\ 128,\ 65)$. The dotted lines indicate the times for Fortran for better
comparison.}
\label{fig:microbench}
\end{figure}

In the top axis of Fig.~\ref{fig:microbench}, we compare the elapsed times for
different implementations of this function.
%
For this out-of-place version, we used three different codes:
\begin{enumerate}
\item a Fortran code (not shown\footnote{The codes and a Makefile used for this
benchmark study are available in \href{%
https://bitbucket.org/fluiddyn/fluiddyn_paper/src/default/fluidfft/microbench/}{%
the repository of the article}.}) written with three nested explicit loops (one
per dimension). Note that as in the Python version we also allocate the memory
where the result is stored.
\item the simplest Python version shown above.
\item a Python version with three nested explicit loops:
% (code not shown).
\begin{minted}[fontsize=\footnotesize]{python}
# pythran export proj_out_of_place_loop(
#     complex128[][][], complex128[][][], complex128[][][],
#     float64[][][], float64[][][], float64[][][], float64[][][])

def proj_out_of_place_loop(vx, vy, vz, kx, ky, kz, inv_k_square_nozero):

    rx = np.empty_like(vx)
    ry = np.empty_like(vx)
    rz = np.empty_like(vx)

    n0, n1, n2 = kx.shape

    for i0 in range(n0):
        for i1 in range(n1):
            for i2 in range(n2):
                tmp = (kx[i0, i1, i2] * vx[i0, i1, i2]
                       + ky[i0, i1, i2] * vy[i0, i1, i2]
                       + kz[i0, i1, i2] * vz[i0, i1, i2]
                ) * inv_k_square_nozero[i0, i1, i2]

                rx[i0, i1, i2] = vx[i0, i1, i2] - kx[i0, i1, i2] * tmp
                ry[i0, i1, i2] = vz[i0, i1, i2] - kx[i0, i1, i2] * tmp
                rz[i0, i1, i2] = vy[i0, i1, i2] - kx[i0, i1, i2] * tmp

    return rx, ry, rz
\end{minted}
\end{enumerate}
For the version without explicit loops, we present the elapsed time for two
cases: (i) simply using Python (yellow bar) and (ii) using the Pythranized
function (first cyan bar).
%
For the Python version with explicit loops, we only present the results for (i)
the Pythranized function (second cyan bar) and (ii) the result of \pack{Numba}
(blue bar).
%
We do not show the result for \pack{Numba} for the code without explicit loops
because it is slower than \pack{Numpy}. We have also omitted the result for
\pack{Numpy} for the code with explicit loops because it is very inefficient.
%
The timing is performed upon tuning the computer using the package
\href{https://pypi.org/project/perf/}{\pack{perf}}.

We see that \pack{Numpy} is approximately three time slower than the Fortran
implementation (which as already mentioned contains the memory allocation).
%
Just using \pack{Pythran} without changing the code (first cyan bar), we save
nearly 50\% of the execution time but we are still significantly slower than
the Fortran implementation.
%
We reach the Fortran performance (even slightly faster) only by using
\pack{Pythran} with the code with explicit loops.
%
With this code, \pack{Numba} is nearly as fast (but still slower) without
requiring any type annotation.

Note that the exact performance differences depend on the hardware, the software
versions\footnote{Here, we use Python~3.6.4 (packaged by conda-forge),
\pack{Numpy}~1.13.3, \pack{Pythran}~0.8.5, \pack{Numba}~0.38, gfortran~6.3 and
clang~6.0.}, the compilers and the compilation options.
%
We use \codeinline{gfortran -O3 -march=native} for Fortran and
\codeinline{clang++ -O3 -march=native} for \pack{Pythran}\footnote{The results
with \codeinline{g++ -O3 -march=native} are very similar but tend to be slightly
slower.}.
%

Since allocating memory is expensive and we do not need the non-projected
velocity field after the call of the function, an evident optimization is to
put the output in the input arrays.  Such an ``in-place'' version can be written
with \pack{Numpy} as:
\begin{minted}[fontsize=\footnotesize]{python}
# pythran export proj_in_place(
#     complex128[][][], complex128[][][], complex128[][][],
#     float64[][][], float64[][][], float64[][][], float64[][][])

def proj_in_place(vx, vy, vz, kx, ky, kz, inv_k_square_nozero):
    tmp = (kx * vx + ky * vy + kz * vz) * inv_k_square_nozero
    vx -= kx * tmp
    vy -= ky * tmp
    vz -= kz * tmp
\end{minted}

As in the first version, we have included the \pack{Pythran} annotation.
%
We also consider an ``in-place'' version with explicit loops:
\begin{minted}[fontsize=\footnotesize]{python}
# pythran export proj_in_place_loop(
#     complex128[][][], complex128[][][], complex128[][][],
#     float64[][][], float64[][][], float64[][][], float64[][][])

def proj_in_place_loop(vx, vy, vz, kx, ky, kz, inv_k_square_nozero):

    n0, n1, n2 = kx.shape

    for i0 in range(n0):
        for i1 in range(n1):
            for i2 in range(n2):
                tmp = (kx[i0, i1, i2] * vx[i0, i1, i2]
                       + ky[i0, i1, i2] * vy[i0, i1, i2]
                       + kz[i0, i1, i2] * vz[i0, i1, i2]
                ) * inv_k_square_nozero[i0, i1, i2]

                vx[i0, i1, i2] -= kx[i0, i1, i2] * tmp
                vy[i0, i1, i2] -= ky[i0, i1, i2] * tmp
                vz[i0, i1, i2] -= kz[i0, i1, i2] * tmp

\end{minted}
Note that this code is much longer and clearly less readable than the version
without explicit loops.  This is however the version which is used in
\pack{fluidfft} since it leads to faster execution.

The elapsed time for these in-place versions and for an equivalent Fortran
implementation are displayed in the bottom axis of Fig.~\ref{fig:microbench}.
%
The ranking is the same as for the out-of-place versions and \pack{Pythran} is also
the faster solution.
%
However, \pack{Numpy} is even more slower (7.8 times slower than \pack{Pythran}
with the explicit loops) than for the out-of-place versions.

From this short and simple microbenchmark, we can infer four main points:
\begin{itemize}
\item Memory allocation takes time!  In Python, memory management is automatic
and we tend to forget it.  An important rule to write efficient code is to
reuse the buffers already allocated as much as possible.

\item Even for this very simple case quite favorable for \pack{Numpy} (no indexing
or slicing), \pack{Numpy} is three to eight time slower than the Fortran
implementations. As long as the execution time is small or that the
function represents a small part of the total execution time, this is not an
issue. However, in other cases, Python-\pack{Numpy} users need to consider
other solutions.

\item \pack{Pythran} is able to speedup the \pack{Numpy} code without explicit
loops and is as fast as Fortran (even slightly faster in our case) for the
codes with explicit loops.

\item \pack{Numba} is unable to speedup the \pack{Numpy} code.
%
It gives very interesting performance for the version with explicit loops
without any type annotation but the result is significantly slower than with
\pack{Pythran} and Fortran.
\end{itemize}

For the aforementioned reasons, we have preferred \pack{Pythran} to compile
optimized ``operator'' functions that complement the FFT classes. Although with
this we obtain remarkable performance, there is still room for some
improvement, in terms of logical implementation and allocation of arrays. For
example, applications such as CFD simulations often deals with non-linear terms
which require dealiasing. The FFT classes of \fluidpack{fft}, currently
allocates the same number of modes in the spectral array so as to transform the
physical array. Thereafter, we apply dealiasing by setting zeros to wavenumbers
which are larger than, say, two-thirds of the maximum wavenumber. Instead, we
could take into account dealiasing in the FFT classes to save some memory and
computation time\footnote{See
\href{https://bitbucket.org/fluiddyn/fluidfft/issues/21/}{fluidfft issue 21}.}.

\section*{Quality control}

% \textcolor{blue}{Detail the level of testing that has been carried out on the
% code (e.g. unit, functional, load etc.), and in which environments. If not
% already included in the software documentation, provide details of how a user
% could quickly understand if the software is working (e.g. providing examples of
% running the software with sample input and output data). }

The package \fluidpack{fft} currently supplies unit tests covering 90\% of its
code.  These unit tests are run regularly through continuous integration on Travis
CI with the most recent releases of \fluidpack{fft}'s dependencies and on
Bitbucket Pipelines inside a static
\href{https://hub.docker.com/u/fluiddyn}{Docker container}.  The tests are run
using standard Python interpreter with all supported versions.

For \fluidpack{fft}, the code coverage results are displayed at
\href{https://codecov.io/bb/fluiddyn/fluidfft}{Codecov}.  Using third-party
packages \pack{coverage} and \pack{tox}, it is straightforward to bootstrap the
installation with dependencies, test with multiple Python versions and combine the
code coverage report, ready for upload. It is also possible to run similar
isolated tests using \pack{tox} or coverage analysis using \pack{coverage} in a
local machine.  Up-to-date build status and coverage status are displayed on the
landing page of the Bitbucket repository.  Instructions on how to run unit tests,
coverage and lint tests are included in the documentation.

We also try to follow a consistent code style as recommended by PEP (Python
enhancement proposals) 8 and 257. This is also inspected using lint checkers such
as \codeinline{flake8} and \codeinline{pylint} among the developers.  The Python
code is regularly cleaned up using the code formatter \codeinline{black}.

\section*{(2) Availability}
\vspace{0.5cm}
\section*{Operating system}

% \textcolor{blue}{Please include minimum version compatibility.}

Windows and any POSIX based OS, such as GNU/Linux and macOS.

\section*{Programming language}

% \textcolor{blue}{Please include minimum version compatibility.}

Python 2.7, 3.5 or above. For the next versions, we will
\href{https://python3statement.org/}{drop Python 2.7 support and Python $>=$
3.6 will be required}.
%
Note that while Cython and Pythran both use the C API of CPython, \fluidpack{fft}
has been successfully tested on PyPy 6.0.
%
A C++11 supporting compiler, while not mandatory for the C++ API or Cython
extensions of \fluidpack{fft}, is recommended to be able to use Pythran extensions.

\section*{Dependencies}

% \textcolor{blue}{E.g. libraries, frameworks, incl. minimum version
% compatibility.}
C++ API:
\begin{itemize}
  \item{\bf Optional:} \pack{OpenMPI} or equivalent, \libpack{FFTW},
    \libpack{P3DFFT}, \libpack{PFFT} and \libpack{cuFFT} libraries.
\end{itemize}

Python API:

\begin{itemize}
\item {\bf Minimum:} \fluidpack{dyn}, \pack{Numpy}, \pack{Cython}, and
  \pack{mako}\ or \pack{Jinja2}; \libpack{FFTW} library.
\item {\bf Optional:} \pack{mpi4py} and \pack{Pythran}; \libpack{P3DFFT},
  \libpack{PFFT} and \libpack{cuFFT} libraries.
\end{itemize}

\section*{List of contributors}

% \textcolor{blue}{Please list anyone who helped to create the software (who may
% also not be an author of this paper), including their roles and affiliations.}

\begin{itemize}
\item Pierre Augier (LEGI): creator of the FluidDyn project and of
\fluidpack{fft}.
\item Cyrille Bonamy (LEGI): C++ code and some methods in the operator classes.
\item Ashwin Vishnu Mohanan (KTH): command lines utilities, benchmarks, unit
  tests, continuous integration, and bug fixes.
\end{itemize}

\section*{Software location:}

% {\bf Archive} \textcolor{blue}{(e.g. institutional repository, general
% repository) (required – please see instructions on journal website for
% depositing archive copy of software in a suitable repository)}

\begin{description}[noitemsep,topsep=0pt]
\item[Name:] PyPI
\item[Persistent identifier:] https://pypi.org/project/fluidfft
\item[Licence:] CeCILL, a free software license adapted to both international
and French legal matters, in the spirit of and retaining compatibility with the
GNU General Public License (GPL).
\item[Publisher:] Pierre Augier
\item[Version published:] 0.2.4
\item[Date published:] 02/07/2018
\end{description}

{\bf Code repository}

\begin{description}[noitemsep,topsep=0pt]
\item[Name:] Bitbucket
\item[Persistent identifier:] https://bitbucket.org/fluiddyn/fluidfft
\item[Licence:] CeCILL
\item[Date published:] 2017
\end{description}

{\bf Emulation environment}

\begin{description}[noitemsep,topsep=0pt]
\item[Name:] Docker
\item[Persistent identifier:] https://hub.docker.com/r/fluiddyn/python3-stable
\item[Licence:] CeCILL-B, a BSD compatible French licence.
\item[Date published:] 02/10/2017
\end{description}

\section*{Language}

% \textcolor{blue}{Language of repository, software and supporting files.}

English

\section*{(3) Reuse potential}

% \textcolor{blue}{Please describe in as much detail as possible the ways in
% which the software could be reused by other researchers both within and outside
% of your field. This should include the use cases for the software, and also
% details of how the software might be modified or extended (including how
% contributors should contact you) if appropriate. Also you must include details
% of what support mechanisms are in place for this software (even if there is no
% support).}

\fluidpack{fft} is used by the Computational Fluid Mechanics framework
\fluidpack{sim} \citep{fluidsim}. It could be used by any C++ or Python project
where real-to-complex 2D or 3D FFTs are performed.

There is no formal support mechanism. However, bug reports can be submitted at
the \href{https://bitbucket.org/fluiddyn/fluidsim/issues}{Issues page on
Bitbucket}. Discussions and questions can be aired on instant messaging
channels in Riot (or equivalent with Matrix protocol)\footnote{
\url{%
  https://matrix.to/\#/\#fluiddyn-users:matrix.org}}
or via IRC protocol on Freenode at \codeinline{\#fluiddyn-users}. Discussions
can also be exchanged via the official mailing list\footnote{
\url{https://www.freelists.org/list/fluiddyn}}.

\section*{Acknowledgements}

% \textcolor{blue}{Please add any relevant acknowledgements to anyone else who
% supported the project in which the software was created, but did not work
% directly on the software itself.}

Ashwin Vishnu Mohanan could not have been as involved in this project without the
kindness of Erik Lindborg.
%
We are grateful to Bitbucket for providing us with a high quality forge
compatible with Mercurial, free of cost.

\section*{Funding statement}

% \textcolor{blue}{If the software resulted from funded research please give the
% funder and grant number.}

This project has indirectly benefited from funding from the foundation Simone et
Cino Del Duca de l'Institut de France, the European Research Council (ERC)
under the European Union's Horizon 2020 research and innovation program (grant
agreement No 647018-WATU and Euhit consortium) and the Swedish Research Council
(Vetenskapsr{\aa}det): 2013--5191.
%
We have also been able to use supercomputers of CIMENT/GRICAD, CINES/GENCI
(grant 2018-A0040107567) and the Swedish National Infrastructure for Computing
(SNIC).

\section*{Competing interests}

% \textcolor{blue}{If any of the authors have any competing interests then these
% must be declared. The authors’ initials should be used to denote differing
% competing interests. For example: “BH has minority shares in [company name],
% which part funded the research grant for this project. All other authors have
% no competing interests."
% %
% If there are no competing interests, please add the statement: “The authors
% declare that they have no competing interests.” }

The authors declare that they have no competing interests.

% \section*{References}

% \textcolor{blue}{Please enter references in the Harvard style and include a DOI
% where available, citing them in the text with a number in square brackets,
% e.g.}

\bibliographystyle{../agsm}
\bibliography{../bib}

\rule{\textwidth}{1pt}

{\bf Copyright Notice} \\
Authors who publish with this journal agree to the following terms: \\

Authors retain copyright and grant the journal right of first publication with
the work simultaneously licensed under a
\href{http://creativecommons.org/licenses/by/3.0/}{Creative Commons Attribution
License} that allows others to share the work with an acknowledgement of the
work's authorship and initial publication in this journal.

Authors are able to enter into separate, additional contractual arrangements
for the non-exclusive distribution of the journal's published version of the
work (e.g., post it to an institutional repository or publish it in a book),
with an acknowledgement of its initial publication in this journal.

By submitting this paper you agree to the terms of this Copyright Notice, which
will apply to this submission if and when it is published by this journal.

\end{document}

### 2019-02-13 paper_02_fluidsim/article.tex
Add fluiddyn papers nearly as it is

%% Journal of Open Research Software Latex template -- Created By Stephen
%% Bonner and John Brennan, Durham Universtiy, UK.
%% see http://openresearchsoftware.metajnl.com

\documentclass{../jors}

\begin{document}

{\bf Software paper for submission to the Journal of Open Research Software} \\

\rule{\textwidth}{1pt}

\section*{(1) Overview}

\vspace{0.5cm}

\section*{Title}

FluidSim: modular, object-oriented Python package for high-performance CFD
simulations

\section*{Paper Authors}

1. MOHANAN, Ashwin Vishnu$^a$ \\
2. BONAMY, Cyrille$^b$ \\
3. CALPE LINARES, Miguel$^b$ \\
4. AUGIER, Pierre$^b$

\smallskip

$^a$ Linn\'e Flow Centre, Department of Mechanics, KTH, 10044 Stockholm, Sweden.\\
$^b$ Univ. Grenoble Alpes, CNRS, Grenoble INP\footnote{Institute of Engineering
Univ. Grenoble Alpes}, LEGI, 38000 Grenoble, France.

\section*{Paper Author Roles and Affiliations}
1. Ph.D. student, Linn\'e Flow Centre, KTH Royal Institute of Technology,
Sweden; \\
2. Research Engineer, LEGI, Universit\'e Grenoble Alpes, CNRS, France; \\
3. Ph.D. student, LEGI, Universit\'e Grenoble Alpes, CNRS, France; \\
4. Researcher, LEGI, Universit\'e Grenoble Alpes, CNRS, France

\section*{Abstract}

%\textcolor{blue}{A short (ca. 100 word) summary of the software being
%described: what problem the software addresses, how it was implemented and
%architected, where it is stored, and its reuse potential.}

The Python package \fluidpack{sim} is introduced in this article as an extensible
framework for Computational Fluid Mechanics (CFD) solvers.
%
It is developed as a part of \href{https://fluiddyn.readthedocs.io}{FluidDyn}
project \citep{fluiddyn}, an effort to promote open-source and open-science
collaboration within fluid mechanics community and intended for both educational
as well as research purposes.
%
Solvers in \fluidpack{sim} are scalable, High-Performance Computing (HPC) codes
which are powered under the hood by the rich, scientific Python ecosystem and the
Application Programming Interfaces (API) provided by \fluidpack{dyn} and
\fluidpack{fft} packages \citep{fluidfft}.
%
The present article describes the design aspects of \fluidpack{sim}, which
includes use of Python as the main language; focus on the ease of use, reuse
and maintenance of the code without compromising performance.
%
The implementation details including optimization methods, modular organization
of features and object-oriented approach of using classes to implement solvers
are also briefly explained.
%
Currently, \fluidpack{sim} includes solvers for a variety of physical problems
using different numerical methods (including finite-difference methods).
%
However, this metapaper shall dwell only on the implementation and performance
of its pseudo-spectral solvers, in particular the two- and three-dimensional
Navier-Stokes solvers.
%
We investigate the performance and scalability of \fluidpack{sim} in a
state of the art HPC cluster.
%
Three similar pseudo-spectral CFD codes based on Python (Dedalus, SpectralDNS) and
Fortran (NS3D) are presented and qualitatively and quantitatively compared to
\fluidpack{sim}.
%
The source code is hosted at Bitbucket as a Mercurial repository
\href{https://bitbucket.org/fluiddyn/fluidsim}{bitbucket.org/fluiddyn/fluidsim}
and the documentation generated using Sphinx can be read online at
\href{https://fluidsim.readthedocs.io}{fluidsim.readthedocs.io}.

\section*{Keywords}

% Python;\ CFD;\ HPC;\ MPI;\ modular;\ object-oriented;\ tested;\ documented;\
% open-source
0D; 1D; 2D; 3D; pseudo-spectral; finite difference; computational fluid
dynamics; sequential; MPI; GPU

\section*{Introduction}

% \textcolor{blue}{An overview of the software, how it was produced, and the
% research for which it has been used, including references to relevant research
% articles. A short comparison with software which implements similar
% functionality should be included in this section. }

Designed as a specialized package of the
\href{https://fluiddyn.readthedocs.io}{FluidDyn} project for computational fluid
mechanics (CFD), \fluidpack{sim} is a comprehensive solution to address the needs
of a fluid mechanics student and researcher alike --- by providing scalable high
performance solvers, on-the-fly postprocessing, and plotting functionalities under
one umbrella.  In the open-science paradigm, scientists will be both users and
developers of the tools at the same time.
%
An advantage of \fluidpack{sim} is that, most of the users just have to read and
write Python code.  \fluidpack{sim} ensures that all critical modules, classes and
functions are well documented --- both as inline comments and as standalone
documentation, complete with examples and tutorials.  For these reasons
\fluidpack{sim} can become a true collaborative code and has the potential to
replace some in-house pseudo-spectral codes written in more conventional
languages.

\subsubsection*{Balance between runtime efficiency and cost of development}
In today's world where clusters are moving from petascale to exascale
performance, computing power is aplenty. In most research scenarios, it becomes
apparent that \emph{man-hours are more expensive than computing time}. In other
words, the cost of development outweighs the cost of computing time. Therefore,
developers should be willing to make small sacrifices in efficiency, to improve
development time, code maintainability and clarity in general. In exceptional
cases when computing time become more expensive, one can still use
existing optimized and performance-tuned compiled libraries implemented in
native languages as the core, and rely on high level programming languages to
extend them for specific applications\footnote{See
  \citet{ramachandran_pysph_2016} for a similar discussion}.

For the above reasons, majority of \fluidpack{sim}'s code-base, in terms of
line of code, is written using pure Python syntax. However, this is done
without compromising performance, by making use of libraries such as \Numpy,
and optimized compilers such as \pack{Cython} and \pack{Pythran}.

\Numpy functions and data types are sufficient for applications such as
initialization and postprocessing operations, since these functions are used
sparingly. Computationally intensive tasks such as time-stepping and linear
algebra operators which are used in every single iteration must be offloaded to
compiled extensions.
%
This optimization strategy can be considered as the computational equivalent of
the \href{https://en.wikipedia.org/wiki/Pareto_principle}{Pareto principle}, also
known as the 80/20 rule\footnote{See \citet{behnel_cython2011},
\href{https://wiki.haskell.org/Why_Haskell_matters}{%
wiki.haskell.org/Why\_Haskell\_matters}}.
%
The goal is to optimize such that ``80 percent of the runtime is spent in
20 percent of the code''~\cite[]{meyers2012effective}.
%
Here, \pack{Cython} \citep{behnel_cython2011} and \pack{Pythran}
\citep{guelton2018pythran} compilers comes in handy.
%
% These two compilers are presented in the companion paper on \fluidpack{dyn}
% \citep{fluiddyn} and
An example on how we use \pack{Pythran} to reach similar performance than with
Fortran by writing only Python code is described in the companion paper on
\fluidpack{fft} \citep{fluidfft}.

% There are some key differences between these packages.

% \pack{Cython}\~\cite{behnel_cython2011} is a very generic and mature library
% based on Pyrex language --- which is Python with some additional syntax.  It has
% the capability to interface between other C or C++ code by generating shared
% libraries which can be imported by Python.  \pack{Cython} can also generate such
% libraries from scratch and even has object-oriented syntax if required.  Some of
% the drawbacks of \pack{Cython} include use of Pyrex which can feel verbose and
% esoteric to a Python developer.  While it supports \Numpy\ arrays as inputs, it
% requires the developer to explicitly write loops using indices, which can be
% verbose and less generic than array notations.

% On the other hand, \pack{Pythran}\~\cite{guelton_pythran2013,
% guelton2018pythran} is a relatively recent development, specializing in
% scientific applications.  \pack{Pythran} can export simple Python functions into
% extensions with the help of type annotations written in Python comments.  When
% \pack{Pythran} is not being used these functions can work like plain Python
% functions.  \pack{Pythran} can also recognize \Numpy functions and optimizes
% code while building extensions internally.  However, \pack{Pythran} does not
% support classes and custom data types as function parameters.

The result of using such an approach is studied in the forthcoming sections by
measuring the performance of \fluidpack{sim}.
%
We will demonstrate that a very large percentage of the elapsed time is spent in
the execution of optimized compiled functions and thus that the ``Python cost'' is
negligible.

\subsubsection*{Target audiences}
\fluidpack{sim} is designed to cater to the needs of three kinds of audience.
\begin{itemize}
\item \emph{Users}, who run simulations with already available solvers. To do
this, one needs to have very basic skills in Python scripting.
\item \emph{Advanced users}, who may extend \fluidpack{sim} by developing
customized solvers for new problems by taking advantage of built-in operators and
time-stepping classes.  In practice, one can easily implement such solvers by
changing a few methods in base solver classes.  To do this, one needs to have
fairly good skills in Python and especially object-oriented programming.
\item \emph{Developers}, who develop the base classes, in particular, the
operators and time stepping classes.  One may also sometime need to write compiled
extensions to improve runtime performance. To do this, desirable traits include
good knowledge in Python, \Numpy, \pack{Cython} and \pack{Pythran}.
\end{itemize}

This metapaper is intended as a short introduction to \fluidpack{sim} and its
implementation, written mainly from a user-perspective. Nevertheless, we also discuss how
\fluidpack{sim} can be customized and extended with minimal effort to promote code
reuse.
%
A more comprehensive and hands-on look at how to use \fluidpack{sim} can be found
in the tutorials\footnote{See
\href{https://fluidsim.readthedocs.io/en/latest/tutorials.html}{fluidsim.readthedocs.io/en/latest/tutorials.html}},
both from a user's and a developer's perspective.
%
In the latter half of the paper, we shall also inspect the performance of
\fluidpack{sim} in large computing clusters and compare \fluidpack{sim} with three
different pseudo-spectral CFD codes.

\section*{Implementation and architecture}

% \textcolor{blue}{How the software was implemented, with details of the
% architecture where relevant. Use of relevant diagrams is appropriate. Please
% also describe any variants and associated implementation differences.}

New features were added over the years to the package whenever demanded by
research interests, thus making the code very user-centric and
function-oriented.  This aspect of code development is termed as YAGNI, one of
the principles of agile programming software development method, which
emphasizes not to spent a lot of time developing a functionality, because most
likely \emph{you aren't gonna need it}.

Almost all functionalities in \fluidpack{sim} are implemented as classes and its
methods are designed to be modular and generic.  This means that the user is free
to use inheritance to modify certain parts to suit one's needs, and avoiding the
risk of breaking a functioning code.

\subsubsection*{Package organization}

\fluidpack{sim} is meant to serve as a framework for numerical solvers using
different methods. For the present version of \fluidpack{sim} there is support for
finite difference and pseudo-spectral methods. An example of a finite difference
solver is \codeinline{fluidsim.solvers.ad1d} which solves the 1D advection
equation. There are also solvers which do not rely on most of the base classes,
such as \codeinline{fluidsim.base.basilisk} which implements a 2D adaptive meshing
solver based on the CFD code \href{http://basilisk.fr/}{Basilisk}. The collection
of solvers using pseudo-spectral methods are more feature-rich in comparison.

The code is organized into the following sub-packages:

\begin{itemize}
\item \codeinline{fluidsim.base}: contains all base classes and a solver for the
trivial equation $\partial_t \mathbf{\hat{u}} = 0 $.
\item \codeinline{fluidsim.operators}: specialized linear algebra and numerical
method operators (e.g., divergence, curl, variable transformations, dealiasing).

\item \codeinline{fluidsim.solvers}: solvers and postprocessing modules for
problems such as 1D advection, 2D and 3D Navier-Stokes equations (incompressible
and under the Boussinesq approximation, with and without density stratification
and system rotation), one-layer shallow water and F\"oppl-von K\'arm\'an
equations.

\item \codeinline{fluidsim.util}: utilities to load and modify an existing
simulation, to test, and to benchmark a solver.
\end{itemize}

Subpackages \codeinline{base} and \codeinline{operators} form the backbone of this
package, and are not meant to be used by the user explicitly.
%
In practice, one can make an entirely new solver for a new problem using this
framework by simply writing one or two importable files containing three classes:
\begin{itemize}
\item an \codeinline{InfoSolver} class\footnote{Inheriting from the base class
\codeinline{fluidsim.base.solvers.info\_base.InfoSolverBase}.}, containing the
information on which classes will be used for the different tasks in the solver
(time stepping, state, operators, output, etc.).
\item a \codeinline{Simulation} class\footnote{Inheriting from the base class
\codeinline{fluidsim.base.solvers.base.SimulBase}.} describing the equations to be
solved.
\item a \codeinline{State} class\footnote{Inheriting from the base class
\codeinline{fluidsim.base.state.StateBase}.} defining all physical variables and
their spectral counterparts being solved (for example: $u_x$ and $u_y$) and
methods to compute one variable from another.
\end{itemize}

We now turn our attention to the simulation object which illustrates how to
access the code from the user's perspective.

\subsubsection*{The simulation object}

The simulation object is instantiated with necessary parameters just before
starting the time stepping.  A simple 2D Navier-Stokes simulation can be
launched using the following Python script:

\begin{minted}[fontsize=\footnotesize]{python}
 from fluidsim.solvers.ns2d.solver import Simul

 params = Simul.create_default_params()
 # Modify parameters as needed
 sim = Simul(params)
 sim.time_stepping.start()
\end{minted}

The script first invokes the \codeinline{create\_default\_params}
\href{https://docs.python.org/3/library/functions.html#classmethod}{classmethod}
which returns a \codeinline{Parameters} object, typically named
\codeinline{params} containing all default parameters.
%
Modifications to simulation parameters are made after this step, to meet the
user's needs. The simulation object is then instantiated by passing
\codeinline{params} as the only argument, typically named \codeinline{sim},
ready to start the iterations.

As demonstrated above, parameters are stored in an object of the class
\codeinline{Parameters}, which uses the
\href{https://fluiddyn.readthedocs.io/en/latest/generated/fluiddyn.util.paramcontainer.html}{\codeinline{ParamsContainer}}
API\footnote{See
\href{https://fluiddyn.readthedocs.io/en/latest/generated/fluiddyn.util.paramcontainer.html}{fluiddyn.readthedocs.io/en/latest/generated/fluiddyn.util.paramcontainer.html}}
of \fluidpack{dyn} package \cite{fluiddyn}.
%
Parameters for all possible modifications to initialization, preprocessing,
forcing, time-stepping, and output of the solvers is incorporated into the
object \codeinline{params} in a hierarchical manner.
%
Once initialized, the ``public'' (not hidden) API does not allow to add new
parameters to this object and only modifications are permitted.\footnote{Example
on modifying the parameters for a simple simulation:
\href{https://fluidsim.readthedocs.io/en/latest/examples/running-simul-onlineplot.html}{%
fluidsim.readthedocs.io/en/latest/examples/running\_simul.html}}

This approach is different from conventional solvers reliant on
text-based input files to specify parameters, which is less robust and can
cause the simulation to crash due to human errors during runtime.
%
A similar, but less readable approach to \codeinline{ParamsContainer} is
adopted by OpenFOAM which relies on \codeinline{dictionaries} to customize
parameters.  The \codeinline{params} object can be printed on the Python or
IPython console and explored interactively using tab-completion, and can be
loaded from and saved into XML and HDF5 files, thus being very versatile.

Note that the same simulation object is used for the plotting and post-processing
tasks.  During or after the execution of a simulation, a simulation object can be
created with the following code (to restart a simulation, one would rather use the
function \codeinline{fluidsim.load\_state\_phys\_file}.):
\begin{minted}[fontsize=\footnotesize]{python}
 from fluidsim import load_sim_for_plot

 # in the directory of the simulation
 sim = load_sim_for_plot()
 # or with the path of the simulation
 # sim = load_sim_for_plot("~/Sim_data/NS2D.strat_240x240_S8x8_2018-04-20_13-45-54")

 # to retrieve the value of a parameter
 print(f"viscosity = {sim.params.nu_2}")

 # to plot the space averaged quantities versus time
 sim.output.spatial_means.plot()
 # to load the corresponding data
 data = sim.output.spatial_means.load()

 # for a 2d plot of the variable "b"
 sim.output.phys_fields.plot("b", time=2)
 # to save a 2d animation
 sim.output.phys_fields.animate("b", tmin=1, tmax=5, save_file=True)
\end{minted}

% \begin{figure}[htp]
%  \centering
%  \includegraphics[width=\linewidth]{Fig/fig_callgraph_inkscape_simple}
%  \caption{Simplified callgraph tracing the initialization of a simulation object}\label{fig:callgraph}
% \end{figure}
%
% Fig.~\ref{fig:callgraph} provides some insight into how the base simulation
% class of all pseudo-spectral solvers are intialized,
% i.e. \codeinline{fluidsim.base.solvers.SimulBasePseudoSpectral}.  All necessary
% sub-classes are then initialized by the \codeinline{\_\_init\_\_} function of
% the simulation class and customized as specified by the parameters.  The objects
% of all sub-classes are attached to the simulation object, as indicated with blue
% labels in Fig.~\ref{fig:callgraph}.  The sub-classes are responsible for a
% unique set of functionalities for the solver to work.

% todo: redo this figure (API changes)...

\begin{figure}[htp]
\centering
\includegraphics[width=\linewidth]{Fig/fig_simul_obj}
\caption{%
\href{https://www.uml-diagrams.org}{UML diagram} of the simulation
object (\codeinline{sim}) for the solver
\codeinline{fluidsim.solvers.ns3d}.
%
Each block represents an object or instance of a class, and the object name
and the class name are written as headings. The solid arrows show how objects are
associated with each other.
%
Methods and variables of significance to the user are displayed in the body of
each object block.}
\label{fig:simul}
\end{figure}

Fig.~\ref{fig:simul} demonstrates how an user can access different objects and
its associated methods through the simulation object for the solver
\codeinline{fluidsim.solvers.ns3d}. Do note that the object and method names
are similar, if not same, for other solvers. The purpose of the objects are
listed below in the order of instantiation:

\begin{outline}
\1 \codeinline{sim.params}: A copy of the \codeinline{params} object supplied
by the user, which holds all information critical to run the simulation and
generating the output.
%
\1 \codeinline{sim.info\_solver}: Contains all the subclass information,
including the module the class belongs to.
%
\1 \codeinline{sim.info}: A union of the \codeinline{sim.info\_solver} and the
\codeinline{sim.params} objects.
%
\1 \codeinline{sim.oper}: Responsible for generating the grid, and for
pseudo-spectral numerical methods such as FFT, IFFT, dealiasing, divergence,
curl, random arrays, etc.
%
\1 \codeinline{sim.output}: Takes care of all on-the-fly post-processing
outputs and functions to load and plot saved output files. Different objects
are assigned with tasks of loading, plotting and sometimes computing:
%
\2 \codeinline{sim.output.print\_stdout}: the mean energies, time elapsed and
time-step of the simulation printed as console output.
%
\2 \codeinline{sim.output.phys\_fields}: the state variables in the physical
plane.  It relies on \codeinline{sim.state} to load or compute the variables
into arrays.
%
\2 \codeinline{sim.output.spatial\_means}: mean quantities such as energy,
enstrophy, forcing power, dissipation.
%
\2 \codeinline{sim.output.spectra}: energy spectra as line plots (i.e.\ as
functions of the module or a component of the wavenumber).
%
\2 \codeinline{sim.output.spect\_energy\_budg}: spectral energy budget by
calculating the transfer term.
%
\2 \codeinline{sim.output.increments}: structure functions from physical
velocity fields.
%
\1 \codeinline{sim.state}: Defines the names of the physical variables being
solved for and their spectral equivalents, along with all required variable
transformations.
%
Also includes high-level objects, aptly named \codeinline{sim.state.state\_phys}
and \codeinline{sim.state.state\_spect} to hold the arrays.
%
\1 \codeinline{sim.time\_stepping}: Generic numeric time-integration object
which dynamically determines the time-step using the CFL criterion for specific
solver and advances the state variables using Runge-Kutta method of order 2 or
4.
%
\1 \codeinline{sim.init\_fields}: Used only once to initialize all state variables
from a previously generated output file or with simple kinds of flow structures,
for example a dipole vortex, base flow with constant value for all gridpoints,
grid of vortices, narrow-band noise, etc.
%
\1 \codeinline{sim.forcing}: Initialized only when
\codeinline{params.forcing.enable} is set as \codeinline{True} and it computes
the forcing variables, which is added on to right-hand-side of the equations
being solved.
%
\1 \codeinline{sim.preprocess}: Adjusts solver parameters such as the magnitude
of initialized fields, viscosity value and forcing rates after all other
subclasses are initialized, and just before the time-integration starts.
\end{outline}

Such a modular organization of the solver's features has several advantages.
The most obvious one, will be the ease of maintaining the code base. As opposed
to a monolithic solver, modular codes are well separated and leads to less
conflicts while merging changes from other developers. Secondly, with this
implementation, it is possible to extend or replace a particular set of
features by inheriting or defining a new class.

Modular codes can be difficult to navigate and understand the connection between
objects and the classes in static languages. It is much less a problem with
Python where one can easily decipher this information from object attributes or
using IPython's dynamic object information feature\footnote{See
\href{https://ipython.readthedocs.io/en/stable/interactive/%
reference.html\#dynamic-object-information}{ipython.readthedocs.io}.}.
%
Now, \fluidpack{sim} goes one step further and one can effortlessly print the
\codeinline{sim.info\_solver} object in the Python / IPython console, to get this
information.  A truncated example of the output is shown below.

\begin{minted}[fontsize=\footnotesize]{python}
>>> sim.info_solver
<fluidsim.solvers.ns3d.solver.InfoSolverNS3D object at 0x7fb6278263c8>

<solver class_name="Simul" module_name="fluidsim.solvers.ns3d.solver"
        short_name="ns3d">
  <classes>
    <Operators class_name="OperatorsPseudoSpectral3D"
               module_name="fluidsim.operators.operators3d"/>

    <State class_name="StateNS3D" keys_computable="['rotz']"
           keys_linear_eigenmodes="['rot_fft']" keys_phys_needed="['vx', 'vy',
           'vz']" keys_state_phys="['vx', 'vy', 'vz']"
           keys_state_spect="['vx_fft', 'vy_fft', 'vz_fft']"
           module_name="fluidsim.solvers.ns3d.state"/>

    <TimeStepping class_name="TimeSteppingPseudoSpectralNS3D"
                  module_name="fluidsim.solvers.ns3d.time_stepping"/>

    <InitFields class_name="InitFieldsNS3D"
                module_name="fluidsim.solvers.ns3d.init_fields">
      <classes>
        <from_file class_name="InitFieldsFromFile"
                   module_name="fluidsim.base.init_fields"/>

        <from_simul class_name="InitFieldsFromSimul"
                    module_name="fluidsim.base.init_fields"/>

        <in_script class_name="InitFieldsInScript"
                   module_name="fluidsim.base.init_fields"/>

        <constant class_name="InitFieldsConstant"
                  module_name="fluidsim.base.init_fields"/>

	<!--truncated output-->

  </classes>

</solver>
\end{minted}

Note that while the 3D Navier-Stokes solver relies on some generic \emph{base
classes}, such as \codeinline{OperatorsPseudoSpectral3D} and
\codeinline{TimeSteppingPseudoSpectral}, shared with other solvers; for other
purposes there are \emph{solver-specific classes}.  The latter is often inherited
from the base classes in \codeinline{fluidsim.base} or classes available in other
solvers --- this made possible by the use of an object-oriented approach.  This is
particularly advantageous while extending existing features or creating new
solvers to use
\href{https://docs.python.org/3/tutorial/classes.html#inheritance}{%
class inheritance}.

\subsection*{Performance}

% todo: we need to go back over this part after the fluidfft paper will be
% written and using the fluidfft paper. Some bits of what is here could go in
% the fluidfft paper.
%
% It is not interesting here to multiply the FFT comparisons and to run these
% benchmarks on many clusters. We have to be as simple as possible.
%
% We need to include the CFD code comparison here.

Performance of a code can be carefully measured by three different program
analysis methods: profiling, micro-benchmarking and scalability analysis.
%
\href{https://en.wikipedia.org/wiki/Profiling_(computer_programming)}{Profiling}
traces various function calls and records the cumulative time consumed and the
number of calls for each function.  Through profiling, we shed light on what
part of the code consumes the lion's share of the computational time and
observe the impact as number of processes and MPI communications increase.
%
\href{https://en.wiktionary.org/wiki/microbenchmark}{Micro-benchmarking} is the
process of timing a specific portion of the code and comparing different
implementations.  The aspect is addressed in greater detail in the companion
paper on \fluidpack{fft} \citep{fluidfft}.
%
On the other hand,
\href{https://en.wikipedia.org/wiki/Scalability#Performance_tuning_versus_hardware_scalability}{a
scalability study} measures how the speed of the code improves when it is deployed
with multiple CPUs operating in parallel.  To do this, the walltime required by
the whole code is measured to complete a certain number of iterations, given a
problem size.
%
Finally, performance can also be studied by comparing different codes on
representative problems. Since such comparisons should not focus only on
performance, we present a comparison study in a separate section.

\begin{table}[h]
\centering
\begin{tabular}{l p{8cm}}
  \toprule
  Cluster & Beskow (Cray XC40 system with Aries interconnect) \\
  CPU &  Intel Xeon CPU E5--2695v4, 2.1GHz \\
  Operating System & SUSE Linux Enterprise Server 11, Linux Kernel 3.0.101\\
  No.\ of cores per nodes used & 32 \\
  Maximum no.\ of nodes used & 32 (2D cases), 256 (3D cases) \\
  Compilers & CPython 3.6.5, Intel C++ Compiler (\pack{icpc}) 18.0.0 \\
  Python packages & \fluidpack{dyn} 0.2.3, \fluidpack{fft} 0.2.3,
  \fluidpack{sim} 0.2.1, \pack{numpy} (OpenBLAS) 1.14.2, \pack{Cython} 0.28.1,
  \pack{mpi4py} 3.0.0, \pack{pythran} 0.8.5 \\

  \bottomrule
\end{tabular}
\caption{Specifications of the supercomputing cluster and software used for profiling and benchmarking.}
\label{tab:specs}

\end{table}

The profiling and scaling tests were conducted in a supercomputing cluster of
the Swedish National Infrastructure for Computing (SNIC) namely Beskow (PDC,
Stockholm). Relevant details regarding software and hardware which affect
performance are summarised in Table~\ref{tab:specs}. Note that no
hyperthreading was used while carrying out the studies.
%
The code comparisons were made on a smaller machine.
%
Results from the three analyses are systematically studied in the following
sections.

\subsubsection*{Profiling}

It is straightforward to perform profiling with the help of the
\href{https://docs.python.org/3/library/profile.html}{cProfile} module, available
in the Python standard library.
%
For \fluidpack{sim}, this module has been conveniently packaged into a command
line utility, \codeinline{fluidsim-profile}.
%
Here, we have analyzed both 2D and 3D Navier-Stokes solvers in Beskow, and
plotted the results in Fig.~\ref{fig:profile2d} and Fig.~\ref{fig:profile3d}
respectively. Functions which consume less than $2\%$ of the total time are
displayed within a single category, \emph{other}.

% todo: redo these figures (API changes + better speed). Use with_pyfftw for
% sequential runs.

\begin{figure}[htp]
\centering
\includegraphics[width=\linewidth]{tmp/fig_profile2d}
\caption{Profiling analysis of the 2D Navier-Stokes
(\codeinline{fluidsim.solvers.ns2d}) solver using a grid sized $1024\times1024$
(a) in sequential with \codeinline{fft2d.with\_fftw1d} operator and (b) with 8
processes with \codeinline{fft2d.mpi\_with\_fftwmpi2d}
operator.}\label{fig:profile2d}
\end{figure}

In Fig.~\ref{fig:profile2d} both sequential and parallel profiles of the 2D
Navier-Stokes solver shows that majority of time is spent in inverse and forward
FFT calls (\codeinline{ifft\_as\_arg} and \codeinline{fft\_as\_arg}). For the
sequential case, approximately 0.14\% of the time is spent in pure Python
functions, i.e.\ functions not built using \pack{Cython} and \pack{Pythran}.
%
\pack{Cython} extensions are responsible for interfacing with FFT operators and
also for the time-step algorithm.  \pack{Pythran} extensions are used to translate
most of the linear algebra operations into optimized, statically compiled
extensions.
%
We also see that only 1.3\% of the time is not spent in the main six functions
(category \emph{other} in the figure).
%
With 8 processes deployed in parallel, time spent in pure Python function
increases to 1.1\% of the total time.
%
These results show that during the optimization process, we have to focus on a
very small number of functions.

\begin{figure}[htp]
\centering
\includegraphics[width=\linewidth]{tmp/fig_profile3d}
\caption{Profiling analysis of the 3D Navier-Stokes
(\codeinline{fluidsim.solvers.ns3d}) solver.
%
Top row: grid sized $128\times128\times128$ solved (a) sequentially using
\codeinline{fft3d.with\_fftw3d} operator and (b) with 8 processes using
\codeinline{fft3d.mpi\_with\_fftwmpi3d} operator.
%
Bottom row: grid sized $512\times512\times512$ using
\codeinline{fft3d.mpi\_with\_fftwmpi3d} operator (c) with 2 processes and
(d) with 128 processes.}
\label{fig:profile3d}
\end{figure}

From Fig.~\ref{fig:profile3d} it can be shown that, for the 3D Navier-Stokes
solver for all cases majority of time is attributed to FFT calls. The overall time
spent in pure Python function range from 0.001\% for $512^3$ grid points and 2
processes to 0.46\% for $128^3$ grid points and 8 processes.
%
This percentage tends to increase with the number of processes used since the real
calculation done in compiled extensions take less time.
%
This percentage is also higher for the coarser resolution for the same reason.
%
However, the time in pure Python remains for all cases largely negligible compared
to the time spent in compiled extensions.

\subsubsection*{Scalability}

Scalability can be quantified by speedup $S$ which is a measure of the time taken
to complete $N$ iterations for different number of processes, $n_p$.
%We shall
%refrain from comparing sequential runs in this context, since the operators used
%for the sequential mode differ from the parallel mode, especially the FFT
%class.
Speedup is formally defined here as:

\begin{equation}
S_\alpha(n_p) = \frac
{[\mathrm{Time\ elapsed\ for\ } N \mathrm{\ iterations\ with\ }n_{p,\min}\mathrm{\ processes}]_{\mathrm{fastest}}
\times n_{p,\min}}
{[\mathrm{Time\ elapsed\ for\ } N \mathrm{\ iterations\ with\ } n_p \mathrm{\
processes}]_\alpha}
\label{eq:speedup}
\end{equation}
where $n_{p,\min}$ is the minimum number of processes employed for a specific
array size and hardware, $\alpha$ denotes the FFT class used and ``fastest''
corresponds to the fastest result among various FFT classes.
%
In addition to number of processes, there is another important parameter, which
is the size of the problem; in other words, the number of grid points used to
discretize the problem at hand.
%
In \emph{strong scaling} analysis, we keep the global grid-size fixed and
increase the number of processes.

Ideally, this should yield a speedup which increases linearly with number of
processes.  Realistically, as the number of processes increase, so does the
number of MPI communications, contributing to some latency in the overall time
spent and thus resulting in less than ideal performance.
%
Also, as shown by profiling in the previous section, majority of the time is
consumed in making forward- and inverse-FFT calls, an inherent bottleneck of
the pseudo-spectral approach. The FFT function calls are the source of most of
the MPI calls during runtime, limiting the parallelism.

\subsubsection*{2D benchmarks}\label{sec:bench2d}

The Navier-Stokes 2D solver (\codeinline{fluidsim.solvers.ns2d}) solving an
initial value problem (with random fields) was chosen as the test case for strong
scaling analysis here. The physical grid was discretized
%
with $1024\times1024$ and $2048\times2048$ points.
%
Fourth-order Runge-Kutta (RK4) method with a constant time-step was used for
time-integration.
%
File input-output and the forcing term has been disabled so as to measure the
performance accurately.  The test case is then executed for 20 iterations.
%for one or more passes
The time elapsed was measured just before and after the
\codeinline{sim.time\_stepping.start()} function call, which was then utilized
to calculate the average walltime per iteration and speedup.
%
This process is repeated for two different FFT classes provided by
\fluidpack{fft}, namely \codeinline{fft2d.mpi\_with\_fftw1d} and
\codeinline{fft2d.mpi\_with\_fftwmpi2d}.

Although the FFT classes are different for sequential and parallel
runs, for the sake of comparison, we have tabulated the actual
walltime elapsed in Table~\ref{table:seqpar}. Clearly due to the differences
in the implementations, for \codeinline{fftw1d}-based classes it takes a
performance hit with parallelization, while for classes based on
\codeinline{fftw2d} the performance improves.

\begin{table}
\centering
\begin{tabular}{lrlr}
\hline
 FFT class &   Time ($n_p=1$) &   FFT class & Time ($n_p=2$)  \\
\hline
  \codeinline{fft2d.with\_fftw1d}  & 6.63   &
  \codeinline{fft2d.mpi\_with\_fftw1d}    & 7.63 \\
  \codeinline{fft2d.with\_fftw2d}  &  5.59  &
  \codeinline{fft2d.mpi\_with\_fftwmpi2d} &  3.91 \\
\hline
\end{tabular}
\caption{Elapsed times (in seconds) for twenty time steps of 1024 x 1024 case
with the 2D Navier-Stokes 2D (\codeinline{fluidsim.solvers.ns2d}) solver.}
\label{table:seqpar}
\end{table}

\begin{figure}[htp]
\centering
\includegraphics[width=\linewidth]{tmp/fig_bench_strong2d}
\caption{Strong scaling benchmarks of the 2D Navier-Stokes
(\codeinline{fluidsim.solvers.ns2d}) solver. The number of cores $n_p$ goes from 2
to $2^{10} = 1024$. Crosses and dots correspond to $1024\times1024$ and
$2048\times2048$ grid points, respectively.}
\label{fig:strong2d}
\end{figure}

In Fig.~\ref{fig:strong2d} we have analyzed the strong scaling speedup $S$ and
walltime per iteration. The fastest result for a particular case is assigned
the value $S=n_p$ as mentioned earlier in Eq.~\ref{eq:speedup}. Ideal speedup
is indicated with a dotted black line and it varies linearly with number of
processes.  We notice that for the $1024\times1024$ case there is an assured
increasing trend in speedup for intra-nodes computation.
%
Nevertheless, when this test case is solved with over a node ($n_p > 32$); the
speedup drops abruptly. While it may be argued that the speedup is impacted by
the cost of inter-node MPI communications via network interfaces, that is not
the case here. This is shown by speedup for the $2048\times2048$ case, where
speedup increases from $n_p = 32$ to $64$, after which it drops again. It is thus
important to remember that a decisive factor in pseudo-spectral simulations is
the choice of the grid size, both global and local (per-process), and for certain
shapes the FFT calls can be exceptionally fast or vice-versa.

%MPI operations are especially slower in inter-node computation, since nodes
%communicate over network interfaces.
%

From the above results, it may also be inferred that superior performance is
achieved through the use of \codeinline{fft2d.mpi\_with\_fftwmpi2d} as the FFT
method. The \codeinline{fft2d.mpi\_with\_fftw1d} method serves as a fallback
option when either FFTW library is not compiled using MPI bindings or the
domain decomposition results in zero-sized arrays. The latter is a known issue
with the current version of \fluidpack{fft} and \fluidpack{sim} and requires
further development\footnote{See
  \href{https://bitbucket.org/fluiddyn/fluidfft/issues/4}{FluidFFT issue 4}}.
This is the reason why the scalability of
\codeinline{fft2d.mpi\_with\_fftwmpi2d} seems to be restricted
in Fig.~\ref{fig:strong2d}, whereas in reality it can be scalable as
\codeinline{fft2d.mpi\_with\_fftw1d}.

To the right of Fig.~\ref{fig:strong2d}, the real-time or walltime required to
perform a single iteration in seconds is found to vary inversely proportional
to the number of processes, $n_p$. The walltime per iteration ranges from
$0.195$ to $0.023$ seconds for the $1024\times1024$ case, and from
$0.128$ to $0.051$ seconds for the $2048\times2048$ case. Thus it is indeed
feasible and scalable to use this particular solver.

\subsubsection*{3D benchmarks}

Using a similar process as described in the previous section,
%~\ref{sec:bench2d},
the Navier-Stokes 3D solver (\codeinline{fluidsim.solvers.ns3d}) is chosen to
perform 3D benchmarks.
%
As demonstrated in Fig.~\ref{fig:strong3d_beskow} two physical global grids
with $128\times128\times128$ and $1024\times1024\times1024$ are used to
discretize the domain.
%
Other parameters are identical to what was described for the 2D benchmarks.

Through \fluidpack{fft}, this solver has four FFT methods at disposal:

\begin{itemize}
 \item \codeinline{fft3d.mpi\_with\_fftw1d}
 \item \codeinline{fft3d.mpi\_with\_fftwmpi3d}
 \item \codeinline{fft3d.mpi\_with\_p3dfft}
 \item \codeinline{fft3d.mpi\_with\_pfft}
\end{itemize}

The first two methods implements a 1D or \emph{slab} decomposition, i.e.\ the
processes are distributed over one index of a 3D array. And the last two
methods implement a 2D or \emph{pencil} decomposition. For the sake of clarity,
we have restricted this analysis to the fastest FFT method of the two types in
this configuration, namely \codeinline{fft3d.mpi\_with\_fftwmpi3d} and
\codeinline{fft3d.mpi\_with\_p3dfft}. A more comprehensive study of the
performance of these FFT methods can be found in \citet{fluidfft}.

\begin{figure}[htp]
\centering
\includegraphics[width=\linewidth]{tmp/fig_bench_strong3d}
\caption{Strong scaling benchmarks of the 3D Navier-Stokes
(\codeinline{fluidsim.solvers.ns3d}) solver in Beskow. The number of cores $n_p$
goes from $2^5 = 32$ to $2^{13} = 8192$. Crosses and dots correspond to $128^3$
and $1024^3$ grid points, respectively.}
\label{fig:strong3d_beskow}
\end{figure}

In Fig.~\ref{fig:strong3d_beskow} the strong scaling speedup and walltime per
iteration are plotted from 3D benchmarks in Beskow.
%
The analysis here is limited to single-node and inter-node performance.
%
For both grid-sizes analyzed here, the \codeinline{fft3d.mpi\_with\_fftwmpi3d}
method is the fastest of all methods but limited in scalability because of the
1D domain decomposition strategy.  To utilize a large number of processors, one
requires the 2D decomposition approach. Also, note that for the
$1024\times1024\times1024$ case, a single-node measurement was not possible as
the size of the arrays required to run the solvers exceeds the available
memory. For the same case, a speedup reasonably close to linear variation is
observed with \codeinline{fft3d.mpi\_with\_p3dfft}.
%
It is also shown that the walltime per iteration improved from
%
$0.083$ to $0.027$ seconds for the $128\times128\times128$ case, and from
$31.078$ to $2.175$ seconds for the $1024\times1024\times1024$ case.

\subsection*{CFD pseudo-spectral code comparisons}

%TODO-DONE: Compare profiling of Dedalus SpectralDNS, NS3D
% See the file notes_compare_codes.md
% Global comparison, not only performance (which is only one aspect for a CFD
% code)

% First broad description of the three codes

As a general CFD framework, \fluidpack{sim} could be compared to OpenFOAM (a CFD
framework based on finite-volume methods).
%
However, in contrast to OpenFOAM, the current version of \fluidpack{sim} is highly
specialized in pseudo-spectral Fourier methods and it is not adapted for
industrial CFD.

In this subsection, we compare \fluidpack{sim} with three other open-source CFD
pseudo-spectral codes\footnote{For the sake of conciseness, we limit this
comparison to only four codes. We have also found the Julia code
\href{https://github.com/FourierFlows/FourierFlows.jl}{FourierFlows.jl} to
demonstrate interesting performance for 2D sequential runs, but without support
for 3D cases and MPI parallelization.}:
\begin{itemize}
\item \href{http://dedalus-project.org/}{Dedalus} \citep{burns_dedalus} is ``a
flexible framework for spectrally solving differential equations''. It is very
versatile and the user describes the problem to be solved symbolically.
%
This approach is very different than the one of \fluidpack{sim}, where the
equations are described with simple \Numpy code.  There is no equivalent of the
\fluidpack{sim} concept of a ``solver'', i.e.\ a class corresponding to a set of
equations with specialized outputs (with the corresponding plotting methods).  To
run a simulation with Dedalus, one has to describe the problem using mathematical
equations.  This can be very convenient because it is very versatile and it is not
necessary to understand how Dedalus works to define a new problem. However, this
approach has also drawbacks:
\begin{itemize}
\item Even for very standard problems, one needs to describe the problem in the
launching script.
\item There is a potentially long initialization phase during which Dedalus
processes the user input and prepares the ``solver''.
\item Even when a user knows how to define a problem symbolically, it is not
simple to understand how the problem is solved by Dedalus and how to interact with
the program with Python.
\item Since solvers are not implemented out-of-the-box in Dedalus,
specialized forcing scheme or outputs are absent. For example, the user has to
implement the computation, saving and plotting of standard outputs like energy
spectra.
\end{itemize}

\item \href{https://github.com/spectralDNS/spectralDNS}{SpectralDNS}
\citep{mortensen_spectraldns2016} is a ``high-performance pseudo-spectral
Navier-Stokes DNS solver for triply periodic domains. The most notable feature of
this solver is that it is written entirely in Python using \Numpy, MPI
for Python (\pack{mpi4py}) and \pack{pyFFTW}.''

Therefore, SpectralDNS is technically very similar to \fluidpack{sim}.
%
Some differences are that SpectralDNS has no object oriented API, and that the user
has to define output and forcing in the launching script\footnote{See
\href{https://github.com/spectralDNS/spectralDNS/tree/master/demo}{the demo
scripts of SpectralDNS}.}, which are thus usually much longer than for
\fluidpack{sim}.
%
Moreover, the parallel Fourier transforms are done using the Python package
\href{https://github.com/spectralDNS/mpiFFT4py}{\pack{mpiFFT4py}}, which is only
able to use the FFTW library and not other libraries as with \fluidpack{fft}
\citep[][]{fluidfft}.

\item \href{https://bitbucket.org/paugier/ns3d}{NS3D} is a highly efficient
pseudo-spectral Fortran code.
%
It has been written in the laboratory
\href{https://www.ladhyx.polytechnique.fr}{LadHyX} and used for several studies
involving simulations (in 3D and in 2D) of the Navier-Stokes equations under the
Boussinesq approximation with stratification and system
rotation~\cite[][]{DeloncleBillantChomaz2008}.It is in particular specialized in
stability studies~\cite[][]{BillantDeloncleChomazOtheguy2010}.
%
NS3D has been highly optimized and it is very efficient for sequential and
parallel simulations (using MPI and OpenMP).  However, the parallelization is
limited to 1D decomposition for the FFT~\cite[][]{fluidfft}.
%
Another weakness compared to \fluidpack{sim} is that NS3D uses simple binary files
instead of HDF5 and NetCDF4 files for \fluidpack{sim}.  Therefore, visualization
programs like Paraview or Visit cannot load NS3D data.

As with many Fortran codes, Bash and Matlab are used for launching and
post-processing, respectively.
%
In terms of user experience, this can be a drawback compared to the coherent
framework \fluidpack{sim} for which the user works only with Python.

\end{itemize}

\paragraph{Description of the test.} We have formulated the benchmarks tests
that all the aforementioned codes solve the Navier-Stokes equations expressed
in the
\begin{itemize}
  \item stream function-vorticity ($\psi-\zeta$) formulation for the
    bidimensional simulations,

    \begin{align*}
      \p_t \hat\zeta &= -(\widehat{\mathbf{u} \cdot \nabla \zeta}) - \nu
        \kappa^2 \hat \zeta \\
      -\kappa^2 \hat \psi &= \hat \zeta \\
      \mathbf{\hat u} &= i\mathbf{k} \times \mathbf{\hat e}_{k_z} \hat \psi
    \end{align*}

  where the hat ($\hat{}$) notation signifies spectral quantities,
  $\mathbf{u}$, $\nu$ is the viscosity, and $\mathbf{k}$ represents the
  velocity and wavenumber vectors,  and $\kappa = |\mathbf{k}|$ (See
  \href{https://fluidsim.readthedocs.io/en/latest/generated/fluidsim.solvers.ns2d.solver.html}{\codeinline{fluidsim.solvers.ns2d}
  solver documentation}).

\item rotational form for the tridimensional simulations,

    \begin{align*}
      \p_t \mathbf{\hat u} = P_\perp \widehat{(\mathbf{u} \times \boldsymbol{\zeta})}
        - \nu \kappa ^ 2 \mathbf{u}
    \end{align*}

    where the $P_\perp$ operator projects the vector field onto a divergence free
    plane (see
  \href{https://fluidsim.readthedocs.io/en/latest/generated/fluidsim.solvers.ns3d.solver.html}{\codeinline{fluidsim.solvers.ns3d}
  solver documentation} and \citet{canuto_algorithms_1988}) and is equivalent
  to a pressure correction step.

\end{itemize}
We compare the code with a very simple and standard task, running a solver for
ten time steps with the Runge-Kutta 4 method. The solvers in \fluidpack{sim}
relies on \fluidpack{fft} to perform the dealiasing operation applied to the
non-linear terms (say, $\mathbf{\hat N}$) to ensure that no aliasing errors emerge in
the course of the simulation. This is implemented by setting zeros according to
this expression

\begin{equation*}
  \mathbf{\hat N}(k_x, k_y) = 0
  \quad \text{if} \quad (k_x > C_d k_{max} \lor k_y > C_d  k_{max})
\end{equation*}

where $C_{d}$ stands for the coefficient of dealiasing and has a typical
value of $2/3$. Other codes, such as spectralDNS and Dedalus allows instead to
zero-pad the spectral arrays with typically $3/2$ times more modes along each
wavenumber axis, an optimization which can potentially save significant
computation time\footnote{See
\href{https://bitbucket.org/fluiddyn/fluidfft/issues/21/}{FluidFFT issue 21}.}.
%
For quantitative comparisons and for the sake of simplicity, we limit
ourselves to compare only sequential runs.  We have already discussed in
detail, the issue of the scalability of pseudo-spectral codes based on Fourier
transforms in the previous section and in the companion paper \citep{fluidfft}.

In contrast to the framework \fluidpack{sim} for which it is easy to define a new
solver for a new set of equations, NS3D is specialized in solving the
Navier-Stokes equations under the Boussinesq approximation. Using NS3D to solve a
new set of equations would require very deep changes in many places in the code.

Note that Dedalus does not implement the standard fully explicit RK4
method\footnote{See
\href{https://bitbucket.org/dedalus-project/dedalus/issues/38/%
slow-simulation-ns2d-over-a-biperiodic}{the Dedalus issue 38.}}. Thus for
Dedalus, we use the most similar time stepping scheme available, RK443, a
4-stage, third-order mixed implicit-explicit scheme described in
\citet{ascher1997implicit}.
%
Note that in the other codes, part of the linear terms are also treated
implicitly.
%
Also note that in several cases, the upper bound of time step is not first
limited by the stability of the time scheme, rather by other needs (to
resolve the fastest wave, accuracy, etc.), so these benchmarks are
representative of elapsed time for accurate real-life simulations.

\paragraph{Bidimensional simulations.}

\begin{table}
\centering
\begin{tabular}{lrrrr}
\hline
          &   \fluidpack{sim} &   Dedalus &   SpectralDNS &   NS3D \\
\hline
 512$^2$  &              0.51 &      1.53 &          0.92 &   0.82 \\
 1024$^2$ &              2.61 &      8.00 &          3.48 &   3.96 \\
\hline
\end{tabular}
\caption{Elapsed times (in seconds) for ten RK4 time steps for two bidimensional
cases and the four CFD codes.}
\label{table:compare}
\end{table}

We first compare the elapsed times for two resolutions (512$^2$ and 1024$^2$) over
a bidimensional space.  The results are summarized in Table~\ref{table:compare}.
%
The results are consistent for the two resolutions. \fluidpack{sim} is the
fastest code for these cases. Dedalus is three times slower than
\fluidpack{sim} but as discussed earlier, the time stepping method is
different. Another notable issue of Dedalus is a long initialization time (more
than 3 min for 1024$^2$ compared to approximately 5~s for \fluidpack{sim}).
Also note that Dedalus has been optimized for bounded domains with Chebyshev
methods.
%
The two other codes SpectralDNS and NS3D have similar performance: slightly slower
than \fluidpack{sim} and much faster than Dedalus.
%
The Fortran code NS3D is slower (by 47\%) than the Python code \fluidpack{sim}.
This can be explained by the fact that there is no specialized numerical scheme
for the 2D case in NS3D, so that more FFTs have to be performed compared to
SpectralDNS and \fluidpack{sim}.
%
This highlights the importance of implementing a well-adapted algorithm for
a class of problems, which is much easier with a highly modular code as
\fluidpack{sim} than with a specialized code as NS3D.

\paragraph{Tridimensional simulations.}

We now compare the elapsed times for ten RK4 time steps for a tridimensional case
with a resolution 128$^3$.
%
Dedalus is slow and does not seem to be adapted for this case so we do not give
exact elapsed time for this code.
%
SpectralDNS is slightly slower (11.55~s) than the two other codes (9.45~s for
\fluidpack{sim} and 9.52~s for NS3D). This difference is mainly explained by the
slower FFTs for SpectralDNS.

\begin{figure}[htp]
\centering
\includegraphics[width=\linewidth]{./tmp/fig_compare_with_ns3d}
\caption{Comparison of the execution times for a 3D case (128$^3$, 10 time steps)
between NS3D (blue bars) and \codeinline{fluidsim.solvers.ns3d} (yellow bars).
%
The first two bars correspond to the total time and the others to the main tasks
in terms of time consumption, namely FFT, Runge-Kutta 4, curl, vector product and
``projection''. }
\label{fig:compare:with:ns3d}
\end{figure}

Fig.~\ref{fig:compare:with:ns3d} presents a more detailed comparison between NS3D
(blue bars) and \fluidpack{sim} (yellow bars).
%
The total elapsed times is mainly spent in five tasks: FFTs, Runge-Kutta 4, curl,
vector product and ``projection''. The times spent to perform these tasks are
compared for the two codes.

We see that FFTs in NS3D are very fast: the FFT execution is 0.55 s longer for
\fluidpack{sim} (nearly 9\% longer). This difference is especially significant for
sequential run for which there is no communication cost involved in the FFT
computation, thus making it the least favorable case for \fluidpack{sim}.
Indeed, MPI communications are input-output bounded tasks which are not faster
in Fortran than in Python.

This difference can partially be explained by the fact that in NS3D, all FFTs are
inplace (so the input can be erased during the transform). On one hand, this
choice is good for performance and for a lower memory consumption.  On the other
hand, since the same variables are used to store the fields in real and in Fourier
spaces, it makes the code harder to write, to understand and to modify.  Since
memory consumption in clusters is much less of a problem than in the
past and that code simplicity is highly important for a framework like
\fluidpack{sim}, we choose to use out-of-place FFTs in \fluidpack{sim}.
%
Another factor is that the flag FFTW\_PATIENT is used in NS3D which leads to very
long initialization and sometimes faster FFTs. Since we did not see significant
speed-up by using this flag in \fluidpack{sim} and that we also care about
initialization time, this flag is not used and we prefer to use the flag
FFTW\_MEASURE, which usually leads to similar performance.

Time stepping in NS3D is significantly slower than in \fluidpack{sim} (0.34 s
$\simeq$ 20 \% slower). We did not find a performance issue in NS3D.
%
The linear operators are slightly faster in \fluidpack{sim} than in the Fortran
code NS3D.  This is because this corresponds to Pythran functions written with
explicit loops~\cite[see][]{fluidfft}.

Although the FFTs are a little bit faster for NS3D, the total time is slightly
smaller (less than 1\% of the total time) for \fluidpack{sim} for this case.

These examples do not prove that \fluidpack{sim} is always faster than NS3D or
is as fast as any very well optimized Fortran codes. However, they do
demonstrate that our very high-level and modular Python code is very efficient
and is not slower than a well-optimized Fortran code.

\section*{Quality control}

% \textcolor{blue}{Detail the level of testing that has been carried out on the
% code (e.g.\ unit, functional, load etc.), and in which environments. If not
% already included in the software documentation, provide details of how a user
% could quickly understand if the software is working (e.g.\ providing examples of
% running the software with sample input and output data). }

\fluidpack{sim} also packages unit tests to go alongside the related
modules. Throughout the development process it is made sure that all tests pass
on a priority-basis, especially prior to a release, to ensure that new changes
to package does not damage existing functionality.

A part of the unit tests checks the solvers for verifiable conservation laws
such as, enstrophy and energy conservation due to non-dissipative terms in the
governing equations. Thus we ensure the accuracy of the solutions and also to
some degree prevent arithmetic and logic errors. The remainder of tests are
meant to verify the robustness of the code for example, to avoid passing
unexpected inputs between functions. These tests can prevent common error
types such as syntax errors, I/O errors, run time errors and compilation errors
(for the extensions).

It is also important to quantify the completeness of the tests, and this
is done by calculating the code coverage. Code coverage is the ratio of the
number of lines tested by unit tests over the total number of lines in the whole
package. For the present version of \fluidpack{sim} the code coverage is
valued at approximately 70\%.
%
For \fluidpack{sim}, the code coverage results are displayed at
\href{https://codecov.io/bb/fluiddyn/fluidsim}{Codecov}.
%
\fluidpack{sim} has also been checked by comparing its results for wellknown
verification problems, as for example the Taylor-Green Vortices for the
incompressible Navier-Stokes equations\footnote{An demonstration is documented
  as \href{https://fluidsim.readthedocs.io/en/latest/examples.html}{an example
  in the online FluidSim documentation}.
}.

We also try to follow a consistent code style as recommended by PEP (Python
enhancement proposals) 8 and 257. This is also inspected using lint
checkers such as \codeinline{flake8} and \codeinline{pylint} among the
developers. The code is regularly cleaned up using the Python code formatter
\codeinline{black}.

All the above quality control techniques are implemented within the continuous
testing solutions, Travis CI and Bitbucket Pipelines. Instructions on how to
run unit tests, coverage and lint tests are included in the documentation.

\section*{(2) Availability}
\vspace{0.5cm}
\section*{Operating system}

Any POSIX based OS, such as GNU/Linux and macOS.

\section*{Programming language}

Python 2.7, 3.4 or above. For the next versions, we will
\href{https://python3statement.org/}{drop Python 2.7 support and Python $>=$
3.6 will be required}.

\section*{Dependencies}

\begin{itemize}
\item {\bf Minimum:} \fluidpack{dyn}, \Numpy, \pack{h5netcdf}, \fluidpack{fft}
\cite[and FFT libraries, see][]{fluidfft}.
\item {\bf Optional:} \Scipy, \pack{mpi4py}, \pack{Cython} and
\pack{Pythran}, \pack{pulp}.
\end{itemize}

\section*{List of contributors}

\begin{itemize}
\item Ashwin Vishnu Mohanan (KTH): Development of the shallow water equations
solver, \codeinline{fluidsim.solvers.sw1l}, testing, continuous integration,
code coverage and documentation.
\item Cyrille Bonamy (LEGI): Extending the sub-package
\codeinline{fluidsim.operators.fft} (currently deprecated) into a
dedicated package, \fluidpack{fft} used by \fluidpack{sim} solvers.
\item Miguel Calpe (LEGI): Development of the 2D Boussinesq equation solver,
\codeinline{fluidsim.solvers.ns2d.strat}.
\item Pierre Augier (LEGI): Creator of \fluidpack{sim} and FluidDyn project,
developer of majority of the modules and solvers, future-proofing with Python 3
compatibility and documentation.
\end{itemize}

\section*{Software location:}

{\bf Archive}

\begin{description}[noitemsep,topsep=0pt]
\item[Name:] PyPI
\item[Persistent identifier:] https://pypi.org/project/fluidsim
\item[Licence:] CeCILL, a free software license adapted to both international
and French legal matters, in the spirit of and retaining compatibility with the
GNU General Public License (GPL).
\item[Publisher:] Pierre Augier
\item[Version published:] 0.2.2
\item[Date published:] 02/07/2018
\end{description}

{\bf Code repository}

\begin{description}[noitemsep,topsep=0pt]
\item[Name:] Bitbucket
\item[Persistent identifier:] https://bitbucket.org/fluiddyn/fluidsim
\item[Licence:] CeCILL
\item[Date published:] 2015
\end{description}

{\bf Emulation environment}

\begin{description}[noitemsep,topsep=0pt]
\item[Name:] Docker
\item[Persistent identifier:] https://hub.docker.com/r/fluiddyn/python3-stable
\item[Licence:] CeCILL-B, a BSD compatible French licence.
\item[Date published:] 02/10/2017
\end{description}

\section*{Language}

English

\section*{(3) Reuse potential}

%\textcolor{blue}{Please describe in as much detail as possible the ways in
%which the software could be reused by other researchers both within and outside
%of your field. This should include the use cases for the software, and also
%details of how the software might be modified or extended (including how
%contributors should contact you) if appropriate. Also you must include details
%of what support mechanisms are in place for this software (even if there is no
%support).}

\fluidpack{sim} can be used in research and teaching to run numerical simulations
with existing solvers.
%
Its simplicity of use and its plotting capacities make it particularly adapted for
teaching.
%
\fluidpack{sim} is used at LEGI and at KTH for studies on geophysical turbulence
\cite[see for example][]{LindborgMohanan2017}.
%
Since it is easy to modify any characteristics of the existing solvers or to build
new solvers, \fluidpack{sim} is a good tool to carry out other types of
simulations for academic studies.
%
The qualities and advantages of \fluidpack{sim} (integration with the Python
ecosystem and the FluidDyn project, documentation, reliability --- thanks to unit tests
and continuous integration ---, versatility, efficiency and scalability) make us think
that \fluidpack{sim} can become a true collaborative code.

There is no formal support mechanism. However, bug reports can be submitted at
the \href{https://bitbucket.org/fluiddyn/fluidsim/issues}{Issues page on
Bitbucket}. Discussions and questions can be aired on instant messaging
channels in Riot (or equivalent with Matrix protocol)\footnote{
\url{%
  https://matrix.to/\#/\#fluiddyn-users:matrix.org}}
or via IRC protocol on Freenode at \codeinline{\#fluiddyn-users}. Discussions
can also be exchanged via the official mailing list\footnote{
\url{https://www.freelists.org/list/fluiddyn}}.

\section*{Acknowledgements}

% \textcolor{blue}{Please add any relevant acknowledgements to anyone else who
% supported the project in which the software was created, but did not work
% directly on the software itself.}

Ashwin Vishnu Mohanan could not have been as involved in this project without the
kindness of Erik Lindborg.
%
We are grateful to Bitbucket for providing us with a high quality forge
compatible with Mercurial, free of cost.

\section*{Funding statement}

% \textcolor{blue}{If the software resulted from funded research please give the
% funder and grant number.}

This project has indirectly benefited from funding from the foundation Simone et
Cino Del Duca de l'Institut de France, the European Research Council (ERC)
under the European Union's Horizon 2020 research and innovation program (grant
agreement No 647018-WATU and Euhit consortium) and the Swedish Research Council
(Vetenskapsr{\aa}det): 2013-5191.
%
We have also been able to use supercomputers of CIMENT/GRICAD, CINES/GENCI
(grant 2018-A0040107567) and the Swedish National Infrastructure for Computing
(SNIC).

\section*{Competing interests}

The authors declare that they have no competing interests.

% \section*{References}

% \textcolor{blue}{Please enter references in the Harvard style and include a DOI
% where available, citing them in the text with a number in square brackets,
% e.g.}
%
% \textcolor{blue}{[1] Piwowar, H A 2011 Who Shares? Who Doesn't? Factors
% Associated with Openly Archiving Raw Research Data. PLoS ONE 6(7): e18657. DOI:
% \\ http://dx.doi.org/10.1371/journal.pone.0018657.}

\bibliographystyle{../agsm}
\bibliography{../bib}

\rule{\textwidth}{1pt}

{\bf Copyright Notice} \\
Authors who publish with this journal agree to the following terms: \\

Authors retain copyright and grant the journal right of first publication with
the work simultaneously licensed under a
\href{http://creativecommons.org/licenses/by/3.0/}{Creative Commons Attribution
License} that allows others to share the work with an acknowledgement of the
work's authorship and initial publication in this journal.

Authors are able to enter into separate, additional contractual arrangements
for the non-exclusive distribution of the journal's published version of the
work (e.g., post it to an institutional repository or publish it in a book),
with an acknowledgement of its initial publication in this journal.

By submitting this paper you agree to the terms of this Copyright Notice, which
will apply to this submission if and when it is published by this journal.

\end{document}

### 2019-02-13 paper_01_fluidfft/microbench/bench.py

@@ -1,0 +1,69 @@
+
+from numpy.random import random
+
+from proj import (
+    proj as proj0,
+    proj_loop as proj_loop0,
+    proj_inplace as proj_inplace0,
+    proj_inplace_loop as proj_inplace_loop0
+)
+
+from proj_pythran import (
+    proj,
+    proj_loop,
+    proj_inplace,
+    proj_inplace_loop
+)
+
+import proj_pythran
+
+try:
+    from proj_numba import (
+        proj as proj_numba,
+        proj_loop as proj_loop_numba,
+        proj_inplace as proj_inplace_numba,
+        proj_inplace_loop as proj_inplace_loop_numba
+    )
+except ImportError:
+    pass
+
+assert hasattr(proj_pythran, '__pythran__')
+
+__all__ = [
+    'proj0',
+    'proj_loop0',
+    'proj_inplace0',
+    'proj_inplace_loop0',
+    'proj',
+    'proj_loop',
+    'proj_inplace',
+    'proj_inplace_loop',
+    # 'proj_numba',
+    # 'proj_loop_numba',
+    # 'proj_inplace_numba',
+    # 'proj_inplace_loop_numba',
+]
+
+
+n0 = n1 = n2 = 128
+
+shape = (n0, n1, n2//2+1)
+
+c0 = 1j*random(shape) + random(shape)
+c1 = 1j*random(shape) + random(shape)
+c2 = 1j*random(shape) + random(shape)
+
+a0 = random(shape)
+a1 = random(shape)
+a2 = random(shape)
+a3 = random(shape)
+
+try:
+    # JIT compiler
+    for func in (proj_numba,
+                 proj_loop_numba,
+                 proj_inplace_numba,
+                 proj_inplace_loop_numba):
+        func(c0, c1, c2, a0, a1, a2, a3)
+except NameError:
+    pass

### 2019-02-13 paper_01_fluidfft/microbench/bench_proj_fortran.f90

@@ -1,0 +1,86 @@
+! benchmark of the function proj
+
+program benchmark_proj
+  implicit none
+  integer, parameter :: N0=128, N1=128, N2=65, N=100
+  double precision, allocatable :: vx(:,:,:,:), vy(:,:,:,:), vz(:,:,:,:)
+  double precision, allocatable :: kx(:,:,:), ky(:,:,:), kz(:,:,:)
+  double precision, allocatable :: inv_k_square_nozero(:,:,:)
+  double precision, allocatable :: res(:,:,:,:,:)
+  real :: start, finish, cumtime
+  integer :: i
+
+  allocate(vx(2, N2, N1, N0), vy(2, N2, N1, N0), vz(2, N2, N1, N0))
+  allocate(kx(N2, N1, N0), ky(N2, N1, N0), kz(N2, N1, N0))
+  allocate(inv_k_square_nozero(N2, N1, N0))
+
+  call random_number(vx)
+  call random_number(vy)
+  call random_number(vz)
+  call random_number(kx)
+  call random_number(ky)
+  call random_number(kz)
+  call random_number(inv_k_square_nozero)
+
+  cumtime = 0
+
+  print*, "This program make some calculations."
+  do i = 1, N
+     call cpu_time(start)
+     allocate(res(2, 3, N2, N1, N0))
+     call proj(res, vx, vy, vz, kx, ky, kz, inv_k_square_nozero, N0, N1, N2)
+     deallocate(res)
+     call cpu_time(finish)
+     cumtime = cumtime + finish - start
+  enddo
+  print '("Mean Time = ",f6.3," ms")', 1000*cumtime/N
+
+  cumtime = 0
+
+  print*, "without allocate/deallocate."
+  allocate(res(2, 3, N2, N1, N0))
+  do i = 1, N
+     call cpu_time(start)
+     call proj(res, vx, vy, vz, kx, ky, kz, inv_k_square_nozero, N0, N1, N2)
+     call cpu_time(finish)
+     cumtime = cumtime + finish - start
+  enddo
+  print '("Mean Time = ",f6.3," ms")', 1000*cumtime/N
+  deallocate(res)
+
+  deallocate(vx, vy, vz)
+  deallocate(kx, ky, kz)
+  deallocate(inv_k_square_nozero)
+end program benchmark_proj
+
+
[…]

### 2019-02-13 paper_01_fluidfft/microbench/make_fig_bar.py

@@ -1,0 +1,102 @@
+import os
+import sys
+
+import numpy as np
+import matplotlib.pyplot as plt
+
+from fluiddyn.output.rcparams import set_rcparams
+
+set_rcparams(fontsize=14, for_article=True, for_beamer=False)
+
+here = os.path.abspath(os.path.dirname(__file__))
+here_tmp = os.path.join(here, '../tmp')
+
+if not os.path.exists(here_tmp):
+    os.mkdir(here_tmp)
+
+keys_outplace = (
+    'fortran',
+    'numpy',
+    'pythran',
+    'pythran\nloops',
+    'numba\nloops'
+)
+
+# on my KTH laptop
+times_outplace = np.array([22.1, 66.6, 36.2, 19.1, 24.4,])
+times_inplace = np.array([9., 54.2, 18.3, 8.5, 14.6,])
+
+# on my LEGI desktop with g++
+times_outplace = np.array([25.6, 74.2, 41.0, 23.0, 28.7,])
+times_inplace = np.array([7.3, 56.2, 17.5, 7.64, 14.7,])
+
+# on my LEGI desktop with clang 6.0
+times_outplace = np.array([25.6, 74.2, 41.0, 22.3, 28.7,])
+times_inplace = np.array([7.3, 56.2, 17.4, 7.19, 14.7,])
+
+left = [0, 1, 2, 3, 4]
+
+colors = ('w', 'y', 'c', 'c', 'b')
+
+fig, axes = plt.subplots(2, figsize=(8, 4))
+
+for ax in axes:
+    ax.set_ylim(0, 78)
+
+ax = axes[0]
+
+ax.bar(left, height=times_outplace, color=colors, edgecolor='k', yerr=2)
+
+ax.set_xticks([])
+ax.set_xticklabels([])
+ax.set_ylabel('elapsed time (ms)')
+ax.set_title('out-of-place (with memory allocation)')
+
+xlim = ax.get_xlim()
+ax.plot(xlim, (times_outplace[0],)*2, 'k:')
+ax.set_xlim(xlim)
+
+y = 55
+for x, s in zip(left, keys_outplace):
+    ax.text(x, y, s, rotation=20,
+            horizontalalignment='center',
+            verticalalignment='center')
+
+keys_inplace = (
+    'fortran',
+    'numpy',
+    'pythran',
+    'pythran\nloops',
+    'numba\nloops')
+
+colors = ('w', 'y', 'c', 'c', 'b')
+
[…]

### 2019-02-13 paper_01_fluidfft/microbench/proj.py

@@ -1,0 +1,72 @@
+
+import numpy as np
+
+# pythran export proj(
+#     complex128[][][], complex128[][][], complex128[][][],
+#     float64[][][], float64[][][], float64[][][], float64[][][])
+
+def proj(vx, vy, vz, kx, ky, kz, inv_k_square_nozero):
+    tmp = (kx * vx + ky * vy + kz * vz) * inv_k_square_nozero
+    return vx - kx * tmp, vy - ky * tmp, vz - kz * tmp
+
+# pythran export proj_loop(
+#     complex128[][][], complex128[][][], complex128[][][],
+#     float64[][][], float64[][][], float64[][][], float64[][][])
+
+
+def proj_loop(vx, vy, vz, kx, ky, kz, inv_k_square_nozero):
+
+    rx = np.empty_like(vx)
+    ry = np.empty_like(vx)
+    rz = np.empty_like(vx)
+
+    n0, n1, n2 = kx.shape
+
+    for i0 in range(n0):
+        for i1 in range(n1):
+            for i2 in range(n2):
+                tmp = (kx[i0, i1, i2] * vx[i0, i1, i2]
+                       + ky[i0, i1, i2] * vy[i0, i1, i2]
+                       + kz[i0, i1, i2] * vz[i0, i1, i2]
+                ) * inv_k_square_nozero[i0, i1, i2]
+
+                rx[i0, i1, i2] = vx[i0, i1, i2] - kx[i0, i1, i2] * tmp
+                ry[i0, i1, i2] = vz[i0, i1, i2] - kx[i0, i1, i2] * tmp
+                rz[i0, i1, i2] = vy[i0, i1, i2] - kx[i0, i1, i2] * tmp
+
+    return rx, ry, rz
+
+
+# pythran export proj_inplace(
+#     complex128[][][], complex128[][][], complex128[][][],
+#     float64[][][], float64[][][], float64[][][], float64[][][])
+
+
+def proj_inplace(vx, vy, vz, kx, ky, kz, inv_k_square_nozero):
+    tmp = (kx * vx + ky * vy + kz * vz) * inv_k_square_nozero
+
+    vx -= kx * tmp
+    vy -= ky * tmp
+    vz -= kz * tmp
+
+
+# pythran export proj_inplace_loop(
+#     complex128[][][], complex128[][][], complex128[][][],
+#     float64[][][], float64[][][], float64[][][], float64[][][])
+
+
[…]

### 2019-02-13 paper_02_fluidsim/Python/base.py

@@ -1,0 +1,24 @@
+from string import ascii_lowercase
+import os
+from fluiddyn.output.rcparams import set_rcparams
+import matplotlib.pyplot as plt
+
+
+titles = ['({})'.format(a) for a in ascii_lowercase]
+
+curdir = os.path.dirname(__file__)
+
+def matplotlib_rc(
+        fontsize=8, dpi=300, tex=True, interactive=False):
+
+    set_rcparams(fontsize, for_article=True, for_beamer=False)
+    plt.rc('legend', fontsize=fontsize-2)
+    plt.rc('font', size=fontsize)
+    plt.rc('figure', dpi=dpi)
+    plt.rc('xtick', direction='in')
+    plt.rc('ytick', direction='in')
+
+    if interactive:
+         plt.ion()
+    else:
+         plt.ioff()

### 2019-02-13 paper_02_fluidsim/Python/bench_analysis.py

@@ -1,0 +1,319 @@
+"""Load and plot benchmarks (:mod:`fluidsim.util.console.bench_analysis`)
+=========================================================================
+
+"""
+import os
+from glob import glob
+import json
+import sys
+
+import pandas as pd
+import numpy as np
+from scipy.signal import argrelmax, argrelmin, argrelextrema
+import matplotlib.pyplot as plt
+from matplotlib.ticker import MaxNLocator
+from fluidsim.util.console.util import ConsoleError
+
+
+def load_bench(path_dir, solver, hostname="any"):
+    """Load benchmarks results saved as JSON files."""
+
+    dicts = []
+    for path in glob(path_dir + "/result_bench_{}*.json".format(solver)):
+        with open(path) as f:
+            try:
+                d = json.load(f)
+            except json.decoder.JSONDecodeError:
+                print(path)
+                raise
+
+        if hostname != "any" and not d["hostname"].startswith(hostname):
+            continue
+
+        dicts.append(d)
+
+    df = pd.DataFrame(dicts)
+    df = df[df.columns.difference(["pid", "time_as_str"])]
+    return df
+
+
+def filter_by_shape(df, n0, n1, n2=None):
+    """Filters all results with the same `n0` and same `n1`."""
+
+    if n2 is None:
+        df = df[(df.n0 == n0) & (df.n1 == n1)]
+        return df[df.columns.difference(["n0", "n1"])]
+    else:
+        df = df[(df.n0 == n0) & (df.n1 == n1) & (df.n2 == n2)]
+        return df[df.columns.difference(["n0", "n1", "n2"])]
+
+
+def filter_by_shapeloc(df, n0_loc, n1_loc, n2_loc=None):
+    """Filters all results with the same `n0_loc * n1_loc`.
+
+    This implies shapeK_loc has same no. of points. This is a weak check and
+    the shapeK_loc may not have the same shape. Make sure all shapes are
+    estimated and tested apriori using:
+
+    >>> fluidsim bench -e
+    >>> mpirun -np {nb_proc} fluidsim bench -c
+
+    """
+    if n2_loc is None:
+        df = df[(df.n0_loc * df.n1_loc == n0_loc * n1_loc)]
+        return df[df.columns.difference(["n0_loc", "n1_loc"])]
+    else:
+        df = df[(df.n0_loc * df.n1_loc * df.n2_loc == n0_loc * n1_loc * n2_loc)]
[…]

### 2019-02-13 templates/utils_render.py
Avoid recursive make: because of shell escape? only compile overview

@@ -47,6 +47,9 @@ def get_options_from_yaml(path):
     with open(path) as f:
         options = yml.load(f.read())
 
+    if "dir" not in options or options["dir"] is None:
+        options["dir"] = path.parent
+
     options["authors_short"] = [
         name_in_initials(name) for name in options["authors"]
     ]

### 2019-02-21 paper_00_fluiddyn_meta/article.tex
Update template and python dependencies

@@
 
-\documentclass{../jors}
+% % \documentclass{../jors}
 
-\begin{document}
+% \begin{document}
 
-{\bf Software paper for submission to the Journal of Open Research Software} \\
+% {\bf Software paper for submission to the Journal of Open Research Software} \\
 
-To complete this template, please replace the blue text with your own. The paper
-has three main sections: (1) Overview; (2) Availability; (3) Reuse potential.
+% To complete this template, please replace the blue text with your own. The paper
+% has three main sections: (1) Overview; (2) Availability; (3) Reuse potential.
 
-Please submit the completed paper to: editor.jors@ubiquitypress.com
+% Please submit the completed paper to: editor.jors@ubiquitypress.com
 
-\rule{\textwidth}{1pt}
+% \rule{\textwidth}{1pt}
 
-\section*{(1) Overview}
+% \section*{(1) Overview}
 
-\vspace{0.5cm}
+% \vspace{0.5cm}
 
-\section*{Title}
+% \section*{Title}
 
@@
 
-FluidDyn: a Python open-source framework for research and teaching in fluid
-dynamics by simulations, experiments and data processing
+% FluidDyn: a Python open-source framework for research and teaching in fluid
+% dynamics by simulations, experiments and data processing
 
-\section*{Paper Authors}
+% \section*{Paper Authors}
 
@@
 % 2. Last name, first name; etc.}
-
-1. AUGIER Pierre$^a$\\
-2. MOHANAN Ashwin Vishnu$^b$\\
-3. BONAMY Cyrille$^a$\\
-
-\smallskip
-
-$^a$ Univ. Grenoble Alpes, CNRS, Grenoble INP\footnote{Institute of Engineering
-Univ. Grenoble Alpes}, LEGI, 38000 Grenoble, France.\\
-$^b$ Linn\'e Flow Centre, Department of Mechanics, KTH, 10044 Stockholm, Sweden.
-
-\section*{Paper Author Roles and Affiliations}
+%
+% 1. AUGIER Pierre$^a$\\
+% 2. MOHANAN Ashwin Vishnu$^b$\\
+% 3. BONAMY Cyrille$^a$\\
+%
+% \smallskip
+%
+% $^a$ Univ. Grenoble Alpes, CNRS, Grenoble INP\footnote{Institute of Engineering
+% Univ. Grenoble Alpes}, LEGI, 38000 Grenoble, France.\\
+% $^b$ Linn\'e Flow Centre, Department of Mechanics, KTH, 10044 Stockholm, Sweden.
+%
+% \section*{Paper Author Roles and Affiliations}
 % \textcolor{blue}{1. First author role and affiliation \\
@@
 
-1. Researcher, LEGI, Universit\'e Grenoble Alpes, CNRS, France\\
-2. Ph.D. student, Linn\'e Flow Centre, KTH Royal Institute of Technology,
-Sweden; \\
-3. Research Engineer, LEGI, Universit\'e Grenoble Alpes, CNRS, France; \\
+% 1. Researcher, LEGI, Universit\'e Grenoble Alpes, CNRS, France\\
+% 2. Ph.D. student, Linn\'e Flow Centre, KTH Royal Institute of Technology,
+% Sweden; \\
+% 3. Research Engineer, LEGI, Universit\'e Grenoble Alpes, CNRS, France; \\
 
-\section*{Abstract}
+% \section*{Abstract}
 
@@
 
-\href{http://fluiddyn.readthedocs.io}{FluidDyn} is a project to foster
-open-science and open-source in the fluid dynamics community.  It is thought of as
-a research project to channel open-source dynamics, methods and tools to do
-science.
-%
-We propose a set of Python packages forming a framework to study fluid dynamics
-with different methods, in particular laboratory experiments (package
-\fluidpack{lab}), simulations (packages \fluidpack{fft}, \fluidpack{sim} and
-\fluidpack{foam}) and data processing (package \fluidpack{image}).
-%
-In the present article, we give an overview of the specialized packages of the
-project and then focus on the base package called \fluidpack{dyn}, which contains
-common code used in the specialized packages.  Packages \fluidpack{fft} and
-\fluidpack{sim} are described with greater detail in two companion papers,
-\citet{fluidfft, fluidsim}.
-%
-With the project FluidDyn, we demonstrate that specialized scientific code can
-be written with methods and good practices of the open-source community. The
-Mercurial repositories are available in Bitbucket
-(\url{https://bitbucket.org/fluiddyn/}). All codes are documented using Sphinx
-and Read the Docs, and tested with continuous integration run on Bitbucket
-Pipelines and Travis.
-%
-To improve the reuse potential, the codes are as modular as possible, leveraging
-the simple object-oriented programming model of Python.
-%
-All codes are also written to be highly efficient, using C++, Cython and
-Pythran to speedup the performance of critical functions.
+% \href{http://fluiddyn.readthedocs.io}{FluidDyn} is a project to foster
+% open-science and open-source in the fluid dynamics community.  It is thought of as
+% a research project to channel open-source dynamics, methods and tools to do
+% science.
+% %
+% We propose a set of Python packages forming a framework to study fluid dynamics
+% with different methods, in particular laboratory experiments (package
+% \fluidpack{lab}), simulations (packages \fluidpack{fft}, \fluidpack{sim} and
+% \fluidpack{foam}) and data processing (package \fluidpack{image}).
+% %
+% In the present article, we give an overview of the specialized packages of the
+% project and then focus on the base package called \fluidpack{dyn}, which contains
+% common code used in the specialized packages.  Packages \fluidpack{fft} and
+% \fluidpack{sim} are described with greater detail in two companion papers,
+% \citet{fluidfft, fluidsim}.
+% %
+% With the project FluidDyn, we demonstrate that specialized scientific code can
+% be written with methods and good practices of the open-source community. The
+% Mercurial repositories are available in Bitbucket
+% (\url{https://bitbucket.org/fluiddyn/}). All codes are documented using Sphinx
+% and Read the Docs, and tested with continuous integration run on Bitbucket
+% Pipelines and Travis.
+% %
+% To improve the reuse potential, the codes are as modular as possible, leveraging
+% the simple object-oriented programming model of Python.
+% %
+% All codes are also written to be highly efficient, using C++, Cython and
+% Pythran to speedup the performance of critical functions.
 
-\section*{Keywords}
+% \section*{Keywords}
 
-% \textcolor{blue}{keyword 1; keyword 2; etc. \\
-% Keywords should make it easy to identify who and what the software will be
-% useful for.}
+% % \textcolor{blue}{keyword 1; keyword 2; etc. \\
+% % Keywords should make it easy to identify who and what the software will be
+% % useful for.}
 
-Fluid dynamics research with Python. Numerical simulations. Laboratory
-experiments. Free and open-source software, modular, object-oriented,
-collaborative, efficient, tested, documented.
+% Fluid dynamics research with Python. Numerical simulations. Laboratory
+% experiments. Free and open-source software, modular, object-oriented,
+% collaborative, efficient, tested, documented.
 
@@
 Science is mainly a collective activity.  We can go further only by
-\href{https://en.wikipedia.org/wiki/%
-Standing_on_the_shoulders_of_giants}{``standing on the shoulders of giants''}
+\href{https://en.wikipedia.org/wiki/Standing\_on\_the\_shoulders\_of\_giants}{``standing
+on the shoulders of giants''}
 (and of a huge number of technicians and scientists).  Science is a lot about
@@
 
-\section*{Implementation and architecture}
+\section{Implementation and architecture}
 
@@
 
-\section*{Software location:}
-
-% {\bf Archive} \textcolor{blue}{(e.g. institutional repository, general
-% repository) (required – please see instructions on journal website for
-% depositing archive copy of software in a suitable repository)}
-
-\begin{description}[noitemsep,topsep=0pt]
-\item[Name:] PyPI
-\item[Persistent identifier:] https://pypi.org/project/fluiddyn
-\item[Licence:] CeCILL-B, a BSD compatible French licence.
-\item[Publisher:] Pierre Augier
-\item[Version published:] 0.2.4
-\item[Date published:] 02/07/2018
-\end{description}
-
-{\bf Code repository}
-% \textcolor{blue}{(e.g. SourceForge, GitHub etc.) (required)}
-
-\begin{description}[noitemsep,topsep=0pt]
-\item[Name:] Bitbucket
-\item[Persistent identifier:] https://bitbucket.org/fluiddyn/fluiddyn
-\item[Licence:] CeCILL-B
-\item[Date published:] 2015
-\end{description}
-
+% \section*{Software location:}
+%
+% % {\bf Archive} \textcolor{blue}{(e.g. institutional repository, general
+% % repository) (required – please see instructions on journal website for
+% % depositing archive copy of software in a suitable repository)}
+%
+% \begin{description}[noitemsep,topsep=0pt]
+% \item[Name:] PyPI
+% \item[Persistent identifier:] https://pypi.org/project/fluiddyn
+% \item[Licence:] CeCILL-B, a BSD compatible French licence.
+% \item[Publisher:] Pierre Augier
+% \item[Version published:] 0.2.4
+% \item[Date published:] 02/07/2018
+% \end{description}
+%
+% {\bf Code repository}
+% % \textcolor{blue}{(e.g. SourceForge, GitHub etc.) (required)}
+%
+% \begin{description}[noitemsep,topsep=0pt]
+% \item[Name:] Bitbucket
+% \item[Persistent identifier:] https://bitbucket.org/fluiddyn/fluiddyn
+% \item[Licence:] CeCILL-B
+% \item[Date published:] 2015
+% \end{description}
+%
 % {\bf Emulation environment} \textcolor{blue}{(if appropriate)}
@@
 
-\section*{Language}
-
-% \textcolor{blue}{Language of repository, software and supporting files.}
-
-English
+% \section*{Language}
+%
+% % \textcolor{blue}{Language of repository, software and supporting files.}
+%
+% English
 
@@
 
-\bibliographystyle{../agsm}
-\bibliography{../bib}
-
 \rule{\textwidth}{1pt}
@@
 will apply to this submission if and when it is published by this journal.
-
-\end{document}

### 2019-02-21 paper_02_fluidsim/article.tex
Update template and python dependencies
sed 's/section\*/section/g' article.tex

@@
 
-\documentclass{../jors}
-
-\begin{document}
+% \documentclass{../jors}
 
@@
 
-\section*{(1) Overview}
+\section{(1) Overview}
 
@@
 
-\section*{Title}
+\section{Title}
 
@@
 
-\section*{Paper Authors}
+\section{Paper Authors}
 
@@
 
-\section*{Paper Author Roles and Affiliations}
+\section{Paper Author Roles and Affiliations}
 1. Ph.D. student, Linn\'e Flow Centre, KTH Royal Institute of Technology,
@@
 
-\section*{Abstract}
+\section{Abstract}
 
@@
 
-\section*{Keywords}
+\section{Keywords}
 
@@
 
-\section*{Introduction}
+\section{Introduction}
 
@@
 
-\subsubsection*{Balance between runtime efficiency and cost of development}
+\subsubsection{Balance between runtime efficiency and cost of development}
 In today's world where clusters are moving from petascale to exascale
@@
 
-\subsubsection*{Target audiences}
+\subsubsection{Target audiences}
 \fluidpack{sim} is designed to cater to the needs of three kinds of audience.
@@
 
-\subsubsection*{Package organization}
+\subsubsection{Package organization}
 
@@
 
-\subsubsection*{The simulation object}
+\subsubsection{The simulation object}
 
@@
 
-\subsection*{Performance}
+\subsection{Performance}
 
@@
 
-\subsubsection*{Profiling}
+\subsubsection{Profiling}
 
@@
 
-\subsubsection*{Scalability}
+\subsubsection{Scalability}
 
@@
 
-\subsubsection*{2D benchmarks}\label{sec:bench2d}
+\subsubsection{2D benchmarks}\label{sec:bench2d}
 
@@
 
-\subsubsection*{3D benchmarks}
+\subsubsection{3D benchmarks}
 
@@
 
-\subsection*{CFD pseudo-spectral code comparisons}
+\subsection{CFD pseudo-spectral code comparisons}
 
@@
 
-\section*{Quality control}
+\section{Quality control}
 
@@
 
-\section*{(2) Availability}
+\section{(2) Availability}
 \vspace{0.5cm}
-\section*{Operating system}
+\section{Operating system}
 
@@
 
-\section*{Programming language}
+\section{Programming language}
 
@@
 
-\section*{Dependencies}
+\section{Dependencies}
 
@@
 
-\section*{List of contributors}
+\section{List of contributors}
 
@@
 
-\section*{Software location:}
+\section{Software location:}
 
@@
 
-\section*{Language}
+\section{Language}
 
@@
 
-\section*{(3) Reuse potential}
+\section{(3) Reuse potential}
 
@@
 
-\section*{Acknowledgements}
+\section{Acknowledgements}
 
@@
 
-\section*{Funding statement}
+\section{Funding statement}
 
@@
 
-\section*{Competing interests}
+\section{Competing interests}
 
@@
 
-% \section*{References}
+% \section{References}
 
@@
 % \\ http://dx.doi.org/10.1371/journal.pone.0018657.}
-
-\bibliographystyle{../agsm}
-\bibliography{../bib}
 

### 2019-02-21 paper_02_fluidsim/article.tex
Subsection -> paragraphbf sed -i 's/subsubsection/paragraphbf/g' */article.tex

@@
 
-\subsubsection{Balance between runtime efficiency and cost of development}
+\paragraphbf{Balance between runtime efficiency and cost of development}
 In today's world where clusters are moving from petascale to exascale
@@
 
-\subsubsection{Target audiences}
+\paragraphbf{Target audiences}
 \fluidpack{sim} is designed to cater to the needs of three kinds of audience.
@@
 
-\subsubsection{Package organization}
+\paragraphbf{Package organization}
 
@@
 
-\subsubsection{The simulation object}
+\paragraphbf{The simulation object}
 
@@
 
-\subsubsection{Profiling}
+\paragraphbf{Profiling}
 
@@
 
-\subsubsection{Scalability}
+\paragraphbf{Scalability}
 
@@
 
-\subsubsection{2D benchmarks}\label{sec:bench2d}
+\paragraphbf{2D benchmarks}\label{sec:bench2d}
 
@@
 
-\subsubsection{3D benchmarks}
+\paragraphbf{3D benchmarks}
 

### 2019-02-21 paper_01_fluidfft/article.tex
Update template and python dependencies
sed 's/section\*/section/g' article.tex
Subsection -> paragraphbf sed -i 's/subsubsection/paragraphbf/g' */article.tex
Overview etc. as sections and sub+other divisions sed -i 's/section\{/subsection\{/g' */article.tex sed -i 's/subsection[{].[[:alnum:]]. /section\{/g' */article.tex
Include fluidfft paper and prettify

@@
 
-\documentclass{../jors}
-
-\begin{document}
-
-{\bf Software paper for submission to the Journal of Open Research Software} \\
-
-To complete this template, please replace the blue text with your own. The
-paper has three main sections: (1) Overview; (2) Availability; (3) Reuse
-potential.
-
-Please submit the completed paper to: editor.jors@ubiquitypress.com
-
-\rule{\textwidth}{1pt}
-
-\section*{(1) Overview}
-
-\vspace{0.5cm}
-
-\section*{Title}
-
-% \textcolor{blue}{The title of the software paper should focus on the software,
-% e.g. “Text mining software from the X project”. If the software is closely
-% linked to a specific research paper, then “Software from Paper Title” is
-% appropriate. The title should be factual, relating to the functionality of the
-% software and the area it relates to rather than making claims about the
-% software, e.g. “Easy-to-use”.}
-
-FluidFFT: common API (C++ and Python) for Fast Fourier Transform HPC libraries
-
-\section*{Paper Authors}
-
-% \textcolor{blue}{1. Last name, first name; (Lead/corresponding author first) \\
-% 2. Last name, first name; etc.}
-
-1. MOHANAN Ashwin Vishnu$^a$\\
-2. BONAMY Cyrille$^b$\\
-3. AUGIER Pierre$^b$\\
-
-\smallskip
-
-$^a$ Linn\'e Flow Centre, Department of Mechanics, KTH, 10044 Stockholm, Sweden.
-$^b$ Univ. Grenoble Alpes, CNRS, Grenoble INP\footnote{Institute of Engineering
-Univ. Grenoble Alpes}, LEGI, 38000 Grenoble, France.\\
-
-\section*{Paper Author Roles and Affiliations}
-% \textcolor{blue}{1. First author role and affiliation \\
-% 2. Second author role and affiliation etc.}
-
-1. Ph.D. student, Linn\'e Flow Centre, KTH Royal Institute of Technology,
-Sweden; \\
-2. Research Engineer, LEGI, Universit\'e Grenoble Alpes, CNRS, France; \\
-3. Researcher, LEGI, Universit\'e Grenoble Alpes, CNRS, France
-
-\section*{Abstract}
+% \documentclass{../jors}
+
+% {\bf Software paper for submission to the Journal of Open Research Software} \\
+%
+% To complete this template, please replace the blue text with your own. The
+% paper has three main sections: (1) Overview; (2) Availability; (3) Reuse
+% potential.
+%
+% Please submit the completed paper to: editor.jors@ubiquitypress.com
+%
+% \rule{\textwidth}{1pt}
+%
+\section{Overview}
+%
+% \vspace{0.5cm}
+%
+% \subsection{Title}
+%
+% % \textcolor{blue}{The title of the software paper should focus on the software,
+% % e.g. “Text mining software from the X project”. If the software is closely
+% % linked to a specific research paper, then “Software from Paper Title” is
+% % appropriate. The title should be factual, relating to the functionality of the
+% % software and the area it relates to rather than making claims about the
+% % software, e.g. “Easy-to-use”.}
+%
+% FluidFFT: common API (C++ and Python) for Fast Fourier Transform HPC libraries
+%
+% \subsection{Paper Authors}
+%
+% % \textcolor{blue}{1. Last name, first name; (Lead/corresponding author first) \\
+% % 2. Last name, first name; etc.}
+%
+% 1. MOHANAN Ashwin Vishnu$^a$\\
+% 2. BONAMY Cyrille$^b$\\
+% 3. AUGIER Pierre$^b$\\
+%
+% \smallskip
+%
+% $^a$ Linn\'e Flow Centre, Department of Mechanics, KTH, 10044 Stockholm, Sweden.
+% $^b$ Univ. Grenoble Alpes, CNRS, Grenoble INP\footnote{Institute of Engineering
+% Univ. Grenoble Alpes}, LEGI, 38000 Grenoble, France.\\
+%
+% \subsection{Paper Author Roles and Affiliations}
+% % \textcolor{blue}{1. First author role and affiliation \\
+% % 2. Second author role and affiliation etc.}
+%
+% 1. Ph.D. student, Linn\'e Flow Centre, KTH Royal Institute of Technology,
+% Sweden; \\
+% 2. Research Engineer, LEGI, Universit\'e Grenoble Alpes, CNRS, France; \\
+% 3. Researcher, LEGI, Universit\'e Grenoble Alpes, CNRS, France
+
+% \subsection{Abstract}
 
@@
 
-The Python package \fluidpack{fft} provides a common Python API for performing
-Fast Fourier Transforms (FFT) in sequential, in parallel and on GPU with different
-FFT libraries (FFTW, P3DFFT, PFFT, cuFFT). \fluidpack{fft} is a comprehensive FFT
-framework which allows Python users to easily and efficiently perform FFT and the
-associated tasks, such as as computing linear operators and energy spectra.
-%
-We describe the architecture of the package composed of C++ and Cython FFT
-classes, Python ``operator'' classes and Pythran functions.
-%
-The package supplies utilities to easily test itself and benchmark the different
-FFT solutions for a particular case and on a particular machine.
-%
-We present a performance scaling analysis on three different computing clusters
-and a microbenchmark showing that \fluidpack{fft} is an interesting solution to
-write efficient Python applications using FFT.
-
-\section*{Keywords}
-
-% \textcolor{blue}{keyword 1; keyword 2; etc. \\
-% Keywords should make it easy to identify who and what the software will be
-% useful for.}
-
-Free and open-source library; Python; Fast Fourier Transform; Distributed; MPI;
-GPU; High performance computing%
-
-\section*{Introduction}
+% The Python package \fluidpack{fft} provides a common Python API for performing
+% Fast Fourier Transforms (FFT) in sequential, in parallel and on GPU with different
+% FFT libraries (FFTW, P3DFFT, PFFT, cuFFT). \fluidpack{fft} is a comprehensive FFT
+% framework which allows Python users to easily and efficiently perform FFT and the
+% associated tasks, such as as computing linear operators and energy spectra.
+% %
+% We describe the architecture of the package composed of C++ and Cython FFT
+% classes, Python ``operator'' classes and Pythran functions.
+% %
+% The package supplies utilities to easily test itself and benchmark the different
+% FFT solutions for a particular case and on a particular machine.
+% %
+% We present a performance scaling analysis on three different computing clusters
+% and a microbenchmark showing that \fluidpack{fft} is an interesting solution to
+% write efficient Python applications using FFT.
+%
+% \subsection{Keywords}
+%
+% % \textcolor{blue}{keyword 1; keyword 2; etc. \\
+% % Keywords should make it easy to identify who and what the software will be
+% % useful for.}
+%
+% Free and open-source library; Python; Fast Fourier Transform; Distributed; MPI;
+% GPU; High performance computing%
+%
+\subsection{Introduction}
 
@@
 
-\section*{Implementation and architecture}
+\subsection{Implementation and architecture}
 
@@
 
-\subsection*{C++ API}
+\subsubsection{C++ API}
 
@@
 
-\subsection*{Python API} Similar to other packages in the FluidDyn project,
+\subsubsection{Python API} Similar to other packages in the FluidDyn project,
 \fluidpack{fft} also uses an object-oriented approach, providing FFT classes.
@@
 
-\subsection*{Code organization}
+\subsubsection{Code organization}
 \begin{figure}[htp]
@@
 
-\subsection*{Performance}
-
-\subsubsection*{Scalability tests using \codeinline{fluidfft-bench}}
+\subsubsection{Performance}
+
+\paragraphbf{Scalability tests using \codeinline{fluidfft-bench}}
 
@@
 
-\subsubsection*{Microbenchmark of critical ``operator'' functions}
+\paragraphbf{Microbenchmark of critical ``operator'' functions}
 
@@
 
-\section*{Quality control}
+\subsection{Quality control}
 
@@
 
-\section*{(2) Availability}
+\section{Availability}
 \vspace{0.5cm}
-\section*{Operating system}
+\subsection{Operating system}
 
@@
 
-\section*{Programming language}
+\subsection{Programming language}
 
@@
 
-\section*{Dependencies}
+\subsection{Dependencies}
 
@@
 % compatibility.}
-C++ API:
+\paragraph{C++ API:}
 \begin{itemize}
@@
 
-\section*{List of contributors}
+\subsection{List of contributors}
 
@@
 
-\section*{Software location:}
-
+\subsection{Software location}
+
+\paragraph{\bf Archive}
 % {\bf Archive} \textcolor{blue}{(e.g. institutional repository, general
@@
 
-\section*{Language}
+\subsection{Language}
 
@@
 
-\section*{(3) Reuse potential}
+\section{Reuse potential}
 
@@
 
-\section*{Acknowledgements}
+\subsection*{Acknowledgements}
 
@@
 
-\section*{Funding statement}
+\subsection*{Funding statement}
 
@@
 
-\section*{Competing interests}
+\subsection*{Competing interests}
 
@@
 
-% \section*{References}
+% \subsection{References}
 
@@
 
-\bibliographystyle{../agsm}
-\bibliography{../bib}
-
-\rule{\textwidth}{1pt}
-
-{\bf Copyright Notice} \\
-Authors who publish with this journal agree to the following terms: \\
+% \rule{\textwidth}{1pt}
+\vspace{10pt}
+\hrule
+\vspace{10pt}
+
+\paragraphbf{Copyright Notice}
+
+Authors who publish with this journal agree to the following terms:
 

### 2019-02-21 chapter_00_open_science.tex
Add todo notes and cleanup chapter 00

@@
 \chapter{Towards reproducible, open science through open research software}
-
-\section*{(1) Overview}
-
-\vspace{0.5cm}
-
-\section*{Title}
+%TODO: Fix table of contents
+\section{Introduction}
+
+% \vspace{0.5cm}
+
+% \section*{Title}
 
@@
 
-Taking advantage of recent advances in open-source software engineering to study
-fluid dynamics: the example of the FluidDyn project
-
-\section*{Abstract}
-
-% \textcolor{blue}{A short (ca. 100 word) summary of the software being
-% described: what problem the software addresses, how it was implemented and
-% architected, where it is stored, and its reuse potential.}
-
-\href{http://fluiddyn.readthedocs.io}{FluidDyn} is a project to foster
-open-science and open-source in the fluid dynamics community.  It is thought of
-as a research project to channel open-source dynamics, methods and tools to do
-better science.
-%
-We propose a set of Python packages forming a framework to study fluid dynamics
-with different methods, in particular laboratory experiments (package
-\fluidpack{lab}), simulations (packages \fluidpack{fft}, \fluidpack{sim} and
-\fluidpack{foam}) and data processing (package \fluidpack{image}).
-%
-In the present article, we review the recent advancements in web-based software
-development tools and Python which facilitated the execution of this project,
-and establish a picture of software engineering in fluid mechanics.
-%
-We give an overview of the specialized packages of the project and then
-focus on the base package called \fluidpack{dyn}, which contains common code
-used in the specialized packages.  Packages \fluidpack{fft} and \fluidpack{sim}
-are described with greater detail in two companion papers, \citet{fluidfft,
-fluidsim}.
-%
-With the project FluidDyn, we demonstrate that specialized scientific code can
-be written with methods and good practices of the open-source community. The
-Mercurial repositories are available in Bitbucket
-(\url{https://bitbucket.org/fluiddyn/}). All codes are documented using Sphinx
-and Read the Docs, and tested with continuous integration run on Bitbucket
-Pipelines and Travis.
-%
-To improve the reuse potential, the codes are as modular as possible, leveraging
-the simple object-oriented programming model of Python.
-%
-All codes are also written to be highly efficient, using C++, Cython and
-Pythran to speedup the performance of critical functions.
-
-\section*{Keywords}
-
-% \textcolor{blue}{keyword 1; keyword 2; etc. \\
-% Keywords should make it easy to identify who and what the software will be
-% useful for.}
-
-Fluid dynamics research with Python. Free and open-source software, modular,
-object-oriented, collaborative, efficient, tested, documented.
-
-\section*{Introduction}
+% Taking advantage of recent advances in open-source software engineering to study
+% fluid dynamics: the example of the FluidDyn project
 
@@
 
-\subsection*{Science, software, open-source and the computer revolution}
+\subsection{Science, software, open-source and the computer revolution}
 
@@
 
-The section ``Overview'' is completed by a second subsection called
+%TODO: Fix names
+The section ``Introduction'' is completed by a second subsection called
 ``Implementation and architecture'' devoted to present the organization of the
@@
 
-\subsection*{Methods and tools for open-source software engineering}
+\subsection{Methods and tools for open-source software engineering}
 
@@
 
-\subsection*{Python, a programming language adapted for open-science}
+\subsection{Python, a programming language adapted for open-science}
 
@@
 
-\subsubsection*{An incredible success}
+\paragraphbf{An incredible success}
 
@@
 
-\subsubsection*{Some Python issues}
+\paragraphbf{Some Python issues}
 
@@
 
-\subsubsection*{Rich landscape of open-source languages useful for science}
+\paragraphbf{Rich landscape of open-source languages useful for science}
 
@@
 
-\subsubsection*{Summary: Python, a versatile Swiss Army Knife for the
+\paragraphbf{Summary: Python, a versatile Swiss Army Knife for the
 scientists}
@@
 
-\subsection*{Productivity at individual, group and community levels}
+\subsection{Productivity at individual, group and community levels}
 
@@
 
-\subsection*{Programming in the field of fluid mechanics}
+\subsection{Programming in the field of fluid mechanics}
 
@@
 
-\subsubsection*{Discussion on Matlab, Labview and Mathematica}
+\paragraphbf{Discussion on Matlab, Labview and Mathematica}
 
@@
 
-\subsubsection*{Different models for software development in fluid mechanics}
+\paragraphbf{Different models for software development in fluid mechanics}
 
@@
 
-\subsection*{FluidDyn project: Breaking the status quo}
+\section{FluidDyn project: Breaking the status quo}
+% \section{Implementation and architecture}
 
@@
 
-\section*{Implementation and architecture}
-
 % \textcolor{blue}{How the software was implemented, with details of the
@@
 
-\subsection*{Organization of the code in packages}
+\subsection{Organization of the code in packages}
 
@@
 
-\subsection*{API of the Python library \fluiddyn}
+\subsection{API of the Python library \fluiddyn}
 
@@
 \fluidpack{image}.
-%
-For the sake of brevity, we shall only describe here some of the most important
-modules.
-
-\subsubsection*{Module \codeinline{fluiddyn.util.paramcontainer}}
-
-Sub-package \codeinline{paramcontainer} defines class
-\codeinline{ParamContainer} which is a hierarchical container for any type of
-parameters. As shown in
-\href{http://fluiddyn.readthedocs.io/en/latest/ipynb/tuto_paramscontainer.html}{%
-this tutorial}, various strengths of an object of this class include:
-\begin{itemize}
-\item Support containing and printing documentation on the parameters.
-\item Support printing to console as XML and saving as XML, hdf5 and netcdf4
-files.
-\item Evaluate data types automatically while loading from saved files.
-\item Easy exploration in an interactive console.
-\item Allows modification of default parameters through simple Python script
-files. An error is raised if the user attempts to use an invalid parameter.
-\item Graphical User Interface (GUI) frontend with PyQt.
-\end{itemize}
-
-Thus, it makes it a much more robust implementation for saving key parameters,
-compared to conventional methods which rely on text or CSV files.
-
-\subsubsection*{Module \codeinline{fluiddyn.util.serieofarrays}}
-
-This modules provides classes to iterate over files. It is a common task in data
-processing to understand how to organized a \textit{serie} of files from filenames
-and formats. Then one has to form smaller sets of arrays contained in the
-files. For example, we can have such \textit{serie}: \codeinline{im1\_1.png,
-im1\_2.png, im1\_3.png, im2\_1.png, im2\_2.png, im2\_3.png} and we could create
-subsets like \codeinline{((im1\_1.png, im1\_2.png, im1\_2.png), (im2\_1.png,
-im2\_2.png, im2\_3.png))}. From the same \textit{serie} of images, we can also
-create other subsets \codeinline{((im1\_1.png, im2\_1.png), (im1\_2.png,
-im2\_2.png), (im1\_3.png, im2\_3.png))}.
-
-The classes of this module allows one to do it with a quite simple and general
-API as shown in
-\href{http://fluiddyn.readthedocs.io/en/latest/ipynb/tuto_serieofarrays.html}{%
-the tutorial on the module}.
-
-\subsubsection*{Module \codeinline{fluiddyn.util.mpi}}
-
-This simple module makes simulataneous sequential and MPI programming a breeze
-by providing number of processes, \codeinline{nb\_proc = 1} and \codeinline{rank
-= 0}, when used in sequential mode, otherwise providing the appropriate values
-provided by \pack{mpi4py} package. Use of this module thwarts coding several if-else
-clauses. If the program was using MPI, also defines the variable
-\codeinline{comm} as an alias for the \codeinline{MPI.COMM\_WORLD} communicator.
-
-\subsubsection*{Module \codeinline{fluiddyn.calcul.easyfft}}
-
-Thin wrapper for an unified API using classes around the packages \pack{pyfftw}
-and \pack{scipy.fftpack}.  It is very easy to perform forward and inverse Fast
-Fourier Transforms (FFT) in one-, two- and three-dimensions.
-The FFT can be multithreaded if the environment variable
-\codeinline{OMP\_NUM\_THREADS} is defined.
-
-\subsection*{The \fluiddyn command-line utilities}
-
-The package \fluiddyn also provides few command-line utilities to perform
-simple tasks useful for a scientist developing with Python and using the
-FluidDyn packages.
-\begin{itemize}
-
-\item \codeinline{fluidinfo}
-
-Displays important information related to software and hardware. It
-includes detailed information such as currently installed FluidDyn packages,
-other third-party packages, C compiler, MPI and \Numpy configuration.
-
-\item \codeinline{fluiddump}
-
-Utility to print the hierarchy of HDF5 and NetCDF files. It does not depend on
-the NetCDF4 library.
-
-\item \codeinline{fluidnbstripout}
-
-Very simple layer to stripout Jupyter notebooks of output, which is useful to
-keep the notebooks lightweight when included in a repository.
-%
-This tool is based on \href{https://github.com/kynan/nbstripout}{nbstripout}
-but, in contrast to nbstripout,
-%
-by default the notebooks with a file name ending as `.nbconvert.ipynb' are
-excluded.
-
-\item \codeinline{fluidmat2py}
-
-Utility to produce a strange code which is no longer Matlab and not yet Python.
-%
-This strange code is then much easier to translate into correct Python than the
-original Matlab code.
-
-\end{itemize}
-
-\section*{Quality control}
-
-% \textcolor{blue}{Detail the level of testing that has been carried out on the
-% code (e.g. unit, functional, load etc.), and in which environments. If not
-% already included in the software documentation, provide details of how a user
-% could quickly understand if the software is working (e.g. providing examples
-% of running the software with sample input and output data). }
-
-The package \fluiddyn\ currently supplies unit tests covering around 70\% of its
-code.  These unit tests are run regularly through continuous integration on Travis
-CI with the most recent releases of \fluiddyn's dependencies and on Bitbucket
-Pipelines inside a static \href{https://hub.docker.com/u/fluiddyn}{Docker
-container}.  The tests are run using standard Python interpreter with all
-supported versions.
-
-For \fluiddyn, the code coverage results are displayed at
-\href{https://codecov.io/gh/fluiddyn/fluiddyn}{Codecov}.  Using third-party
-packages \pack{coverage} and \pack{tox}, it is straightforward to bootstrap the
-installation with dependencies, test with multiple Python versions and combine the
-code coverage report, ready for upload. It is also possible to run similar
-isolated tests using \pack{tox} or coverage analysis using \pack{coverage} in a
-local machine.  Up-to-date build status and coverage status are displayed on the
-landing page of the Bitbucket repository.
-
-We also try to follow a consistent code style as recomended by PEP (Python
-enhancement proposals) 8 and 257. This is also inspected using lint checkers
-such as \codeinline{flake8} and \codeinline{pylint} among the developers.  The
-code is regularly cleaned up using the Python code formatter \codeinline{black}.
-
-\section*{(2) Availability}
-\vspace{0.5cm}
-\section*{Operating system}
-
-% \textcolor{blue}{Please include minimum version compatibility.}
-
-Windows and any POSIX based OS, such as GNU/Linux and macOS.
-
-\section*{Programming language}
-
-% \textcolor{blue}{Please include minimum version compatibility.}
-
-Python 2.7, 3.4 or above.
-
-% \section*{Additional system requirements}
-
-% \textcolor{blue}{E.g. memory, disk space, processor, input devices, output devices.}
-
-\section*{Dependencies}
-
-% \textcolor{blue}{E.g. libraries, frameworks, incl. minimum version
-% compatibility.}
-
-We list here only the dependencies of the base package \fluidpack{dyn}.
-
-\begin{itemize}
-\item {\bf Minimum:} \Numpy, \pack{Matplotlib}, \pack{psutil}, \pack{future},
-\pack{subprocess32} (for Python 2.7 only).
-
-\item {\bf Full functionality:} \pack{h5py}, \pack{h5netcdf}, \pack{mpi4py},
-\pack{Scipy}, \pack{pyfftw} (requires FFTW library), \pack{pillow}.
-
-\item {\bf Optional:} OpenCV with Python bindings, \pack{scikit-image}.
-\end{itemize}
-\section*{List of contributors}
-
-% \textcolor{blue}{Please list anyone who helped to create the software (who may
-% also not be an author of this paper), including their roles and affiliations.}
-
-\begin{itemize}
-\item Pierre Augier (LEGI): creator of the FluidDyn project, developer of
-majority of the FluidDyn packages, future-proofing with Python 3 compatibility
-and documentation.
-
-\item Ashwin Vishnu Mohanan (KTH): developer of the packages
-\fluidpack{sim}, \fluidpack{fft}, \fluidpack{image} and
-\fluiddyn; documentation, code coverage and continuous integration (Docker,
-Bitbucket pipeline and Travis).
-
-\item Cyrille Bonamy (LEGI): developer of \fluidpack{fft}, \fluidpack{sim} and
-\fluidpack{foam}. Main maintainer of \fluidpack{foam}.
-
-\item Antoine Campagne (LEGI): developer of \fluidpack{image} and
-\fluidpack{lab}.
-
-\item Miguel Calpe Linares (LEGI): developer of \fluidpack{sim} and
-\fluidpack{lab}.
-
-\item Julien Salort (Laboratoire de physique, ENS de Lyon): developer of
-\fluidpack{lab}.
-
-\end{itemize}
-%
-% \section*{Software location:}
-%
-% % {\bf Archive} \textcolor{blue}{(e.g. institutional repository, general
-% % repository) (required – please see instructions on journal website for
-% % depositing archive copy of software in a suitable repository)}
-%
-% \begin{description}[noitemsep,topsep=0pt]
-% \item[Name:] PyPI
-% \item[Persistent identifier:] https://pypi.org/project/fluiddyn
-% \item[Licence:] CeCILL-B, a BSD compatible French licence.
-% \item[Publisher:] Pierre Augier
-% \item[Version published:] 0.2.3
-% \item[Date published:] 02/05/2018
-% \end{description}
-%
-% {\bf Code repository}
-% % \textcolor{blue}{(e.g. SourceForge, GitHub etc.) (required)}
-%
-% \begin{description}[noitemsep,topsep=0pt]
-% \item[Name:] Bitbucket
-% \item[Persistent identifier:] https://bitbucket.org/fluiddyn/fluiddyn
-% \item[Licence:] CeCILL-B
-% \item[Date published:] 2015
-% \end{description}
-
-% {\bf Emulation environment} \textcolor{blue}{(if appropriate)}
-
-% \begin{description}[noitemsep,topsep=0pt]
-% \item[Name:] \textcolor{blue}{The name of the archive.}
-% \item[Persistent identifier:] \textcolor{blue}{e.g. DOI, handle, PURL, etc.}
-% \item[Licence:] \textcolor{blue}{Open license under which the software is licensed.}
-% \item[Date published:] \textcolor{blue}{dd/mm/yy}
-% \end{description}
-
-% \section*{Language}
-
-% \textcolor{blue}{Language of repository, software and supporting files.}
-
-% English
-
-\section*{(3) Reuse potential}
-%
-% \textcolor{blue}{Please describe in as much detail as possible the ways in
-% which the software could be reused by other researchers both within and outside
-% of your field. This should include the use cases for the software, and also
-% details of how the software might be modified or extended (including how
-% contributors should contact you) if appropriate. Also you must include details
-% of what support mechanisms are in place for this software (even if there is no
-% support).}
-%
-
-As a library, \fluiddyn\ has been used in the project's specialized packages.
-The common code base for packages with such varied applications is a proof of
-\fluiddyn's versatility and generality. It can be used in other packages outside
-the FluidDyn project easily depending on the need. The command-lines tools can
-be useful to all scientists working with Python. Other use cases could be:
-\begin{itemize}
-	\item \codeinline{ParamContainer} as a generic parameter storage
-		standard.
-	\item Modules within \codeinline{fluiddyn.util} for miniature tasks,
-		printing in terminal with colours, getting memory usage
-		information, detecting if the current Python session is inside
-		IPython, creating a string with time and date, etc.
-	\item Modules within \codeinline{fluiddyn.io} to read
-		and save images of various formats including TIFF files with
-		multiple images; easily ask user with yes/no queries; handle
-		different file formats using classes, including CSV, HDF5,
-		Digiflow, Dantec formats.
-	\item Modules inside \codeinline{fluiddyn.clusters} subpackage to
-		make job submission scriptable in HPC clusters with OAR or
-		SLURM job schedulers. It could be even used for non-Python
-		jobs.
-\end{itemize}
-
-There is no formal support mechanism. However, bug reports can be submitted at the
-\href{https://bitbucket.org/fluiddyn/fluiddyn/issues}{Issues page on Bitbucket}.
-Discussions and questions can be aired on instant messaging channels in Riot (or
-equivalent with Matrix protocol) at
-\url{%
-  https://matrix.to/#/#fluiddyn-users:matrix.org}
-or via IRC protocol on Freenode at \codeinline{\#fluiddyn-users}.
-
-\subsection*{Conclusions}
-
-FluidDyn is an attempt to set off collaborative dynamics based on open-source
-development in fluid dynamics research.
-%
-We shall try, with this project, to explore the possibilities of open-source in
-science and fluid dynamics, by fully exploiting the new open-source tools and
-methods presented in the first subsections of the introduction.
-
-% todo: again mention some success of community-driven subject-area research
-% libraries? See the file ./communitydriven_research_soft.md
-
-The project is right now in a very preliminary stage.  Packages are actively
-evolving with interesting features and a framework for collaborative
-development, packaging, documenting and testing is now well set.
-%
-However, the community around the project is currently tiny and now, we have to
-work on attracting users and developers since an active community is a
-criteria for success and sustenance of an open-source project.
-
-Will people use the FluidDyn tools and collaborate through the project
-FluidDyn?
-%
-There are clearly many challenges and potential barriers:
-\begin{itemize}
-\item Some habits in the community.
-\item Knowledge and skills in the community. For example, only few people use
-issue tracker and pull requests.
-% from Julien Salort: bad!
-% \item Misunderstanding between colleagues and in particular colleagues with
-% responsibilities and power.
-\item Lack of a business model for open-source software in science.  A good
-quality open-source software has a cost that institutions should be willing
-to fund and support.
-\item Lack of recognition of the work spent in open-source, in particular for
-precarious scientists.
-\end{itemize}
-
-On our side, we also have positive points.
-%
-The quality of the tools we use (Python and its scientific ecosystem,
-Mercurial, Read the Docs, Jupyter, ...) is impressive.
-%
-Scientific code is done to be read and to transmit ideas. To this effect, Python
-is among the best languages today.
-%
-We can benefit from the dynamics of Python and of emerging subjects like deep
-learning and the Internet of Things.
-%
-Finally, if we manage to gather a community of users and of developers, the
-collective efficiency related to open-source methods and tools can be a strong
-booster.
-
-\section*{Acknowledgements}
-
-% \textcolor{blue}{Please add any relevant acknowledgements to anyone else who
-% supported the project in which the software was created, but did not work
-% directly on the software itself.}
-
-We thank the CNRS to finance the work of Pierre Augier while giving freedom in
-terms of scientific project.
-%
-Similarly, Ashwin Vishnu Mohanan could not have been as involved in this
-project without the kindness of Erik Lindborg.
-%
-We thank Cyrille Bonamy, Antoine Campagne, Miguel Calpe Linares and Julien
-Salort for their implications in the development of some of the FluidDyn
-packages.
-%
-We thank Bitbucket for providing us for free a high quality forge compatible with
-Mercurial.
-%
-Finally, we thank Gabriel Moreau and Jo\"el Sommeria for their constant willing
-to discuss and share their knowledge.
-
-\section*{Funding statement}
-
-% \textcolor{blue}{If the software resulted from funded research please give the
-% funder and grant number.}
-
-This project has indirectly benefit from funding from the foundation Simone et
-Cino Del Duca de l'Institut de France, the European Research Council (ERC)
-under the European Union's Horizon 2020 research and innovation program (grant
-agreement No 647018-WATU and Euhit consortium) and the Swedish Research Council
-(Vetenskapsr{\aa}det): 2013-5191.
-%
-We have also been able to use supercomputers of CIMENT/GRICAD, CINES/GENCI and
-the Swedish National Infrastructure for Computing (SNIC).
-
-% \section*{Competing interests}
-
-% \textcolor{blue}{If any of the authors have any competing interests then these
-% must be declared. The authors’ initials should be used to denote differing
-% competing interests. For example: “BH has minority shares in [company name],
-% which part funded the research grant for this project. All other authors have
-% no competing interests."
+
+\section{Conclusion}
+
+% FluidDyn is an attempt to set off collaborative dynamics based on open-source
+% development in fluid dynamics research.
 % %
-% If there are no competing interests, please add the statement: “The authors
-% declare that they have no competing interests.” }
-
-% The authors declare that they have no competing interests.
-
-% already included by \bibliography{../bib}
-% \section*{References}
-
-% \textcolor{blue}{Please enter references in the Harvard style and include a DOI
-% where available, citing them in the text with a number in square brackets,
-% e.g.}
-
-% \textcolor{blue}{[1] Piwowar, H A 2011 Who Shares? Who Doesn't? Factors
-% Associated with Openly Archiving Raw Research Data. PLoS ONE 6(7): e18657. DOI:
-% \\ http://dx.doi.org/10.1371/journal.pone.0018657.}
-%
-% \bibliographystyle{../agsm}
-% \bibliography{../bib}
-
-% \rule{\textwidth}{1pt}
-%
-% { \bf Copyright Notice} \\
-% Authors who publish with this journal agree to the following terms: \\
-%
-% Authors retain copyright and grant the journal right of first publication with
-% the work simultaneously licensed under a
-% \href{http://creativecommons.org/licenses/by/3.0/}{Creative Commons Attribution
-% License} that allows others to share the work with an acknowledgement of the
-% work's authorship and initial publication in this journal.
-%
-% Authors are able to enter into separate, additional contractual arrangements
-% for the non-exclusive distribution of the journal's published version of the
-% work (e.g., post it to an institutional repository or publish it in a book),
-% with an acknowledgement of its initial publication in this journal.
-%
-% By submitting this paper you agree to the terms of this Copyright Notice, which
-% will apply to this submission if and when it is published by this journal.
-%
-%
+% We shall try, with this project, to explore the possibilities of open-source in
+% science and fluid dynamics, by fully exploiting the new open-source tools and
+% methods presented in the first subsections of the introduction.
+%
+% % todo: again mention some success of community-driven subject-area research
+% % libraries? See the file ./communitydriven_research_soft.md
+%
+% The project is right now in a very preliminary stage.  Packages are actively
+% evolving with interesting features and a framework for collaborative
+% development, packaging, documenting and testing is now well set.
+% %
+% However, the community around the project is currently tiny and now, we have to
+% work on attracting users and developers since an active community is a
+% criteria for success and sustenance of an open-source project.
+%
+% Will people use the FluidDyn tools and collaborate through the project
+% FluidDyn?
+% %
+% There are clearly many challenges and potential barriers:
+% \begin{itemize}
+% \item Some habits in the community.
+% \item Knowledge and skills in the community. For example, only few people use
+% issue tracker and pull requests.
+% % from Julien Salort: bad!
+% % \item Misunderstanding between colleagues and in particular colleagues with
+% % responsibilities and power.
+% \item Lack of a business model for open-source software in science.  A good
+% quality open-source software has a cost that institutions should be willing
+% to fund and support.
+% \item Lack of recognition of the work spent in open-source, in particular for
+% precarious scientists.
+% \end{itemize}
+%
+% On our side, we also have positive points.
+% %
+% The quality of the tools we use (Python and its scientific ecosystem,
+% Mercurial, Read the Docs, Jupyter, ...) is impressive.
+% %
+% Scientific code is done to be read and to transmit ideas. To this effect, Python
+% is among the best languages today.
+% %
+% We can benefit from the dynamics of Python and of emerging subjects like deep
+% learning and the Internet of Things.
+% %
+% Finally, if we manage to gather a community of users and of developers, the
+% collective efficiency related to open-source methods and tools can be a strong
+% booster.

### 2019-02-21 paper_01_fluidfft/microbench/bench_proj_fortran_inplace.f90
Fix/compilation with papers

@@ -1,0 +1,71 @@
+! benchmark of the function proj
+
+program benchmark_proj
+  implicit none
+  integer, parameter :: N0=128, N1=128, N2=64, N=100
+  double precision, allocatable :: vx(:,:,:,:), vy(:,:,:,:), vz(:,:,:,:)
+  double precision, allocatable :: kx(:,:,:), ky(:,:,:), kz(:,:,:)
+  double precision, allocatable :: inv_k_square_nozero(:,:,:)
+  real :: start, finish, cumtime
+  integer :: i
+
+  allocate(vx(2, N2, N1, N0), vy(2, N2, N1, N0), vz(2, N2, N1, N0))
+  allocate(kx(N2, N1, N0), ky(N2, N1, N0), kz(N2, N1, N0))
+  allocate(inv_k_square_nozero(N2, N1, N0))
+
+  call random_number(vx)
+  call random_number(vy)
+  call random_number(vz)
+  call random_number(kx)
+  call random_number(ky)
+  call random_number(kz)
+  call random_number(inv_k_square_nozero)
+
+  cumtime = 0
+
+  print*, "This program make some calculations."
+  do i = 1, N
+     call cpu_time(start)
+     call proj(vx, vy, vz, kx, ky, kz, inv_k_square_nozero, N0, N1, N2)
+     call cpu_time(finish)
+     cumtime = cumtime + finish - start
+  enddo
+  print '("Mean Time = ",f6.3," ms")', 1000*cumtime/N
+
+
+  deallocate(vx, vy, vz)
+  deallocate(kx, ky, kz)
+  deallocate(inv_k_square_nozero)
+end program benchmark_proj
+
+
+subroutine proj(vx, vy, vz, kx, ky, kz, inv_k_square_nozero, N0, N1, N2)
+
+  implicit none
+
+  ! Input/Output
+  integer, intent(in) :: N0, N1, N2
+  double precision, intent(inout) :: vx(2, N2, N1, N0), vy(2, N2, N1, N0), vz(2, N2, N1, N0)
+  double precision, intent(in) :: kx(N2, N1, N0), ky(N2, N1, N0), kz(N2, N1, N0)
+  double precision, intent(in) :: inv_k_square_nozero(N2, N1, N0)
+
+  ! Locals
+  double precision :: tmp(2)
+  integer:: i, j, k
+
+  do k = 1, N0
[…]

### 2019-02-21 paper_01_fluidfft/microbench/proj_numba.py

@@ -1,0 +1,63 @@
+
+import numpy as np
+
+from numba import jit
+
+@jit
+def proj(vx, vy, vz, kx, ky, kz, inv_k_square_nozero):
+    tmp = (kx * vx + ky * vy + kz * vz) * inv_k_square_nozero
+
+    return (vx - kx * tmp,
+            vy - ky * tmp,
+            vz - kz * tmp)
+
+
+@jit
+def proj_loop(vx, vy, vz, kx, ky, kz, inv_k_square_nozero):
+
+    rx = np.empty_like(vx)
+    ry = np.empty_like(vx)
+    rz = np.empty_like(vx)
+
+    n0, n1, n2 = kx.shape
+
+    for i0 in range(n0):
+        for i1 in range(n1):
+            for i2 in range(n2):
+                tmp = (kx[i0, i1, i2] * vx[i0, i1, i2]
+                       + ky[i0, i1, i2] * vy[i0, i1, i2]
+                       + kz[i0, i1, i2] * vz[i0, i1, i2]
+                ) * inv_k_square_nozero[i0, i1, i2]
+
+                rx[i0, i1, i2] = vx[i0, i1, i2] - kx[i0, i1, i2] * tmp
+                ry[i0, i1, i2] = vz[i0, i1, i2] - kx[i0, i1, i2] * tmp
+                rz[i0, i1, i2] = vy[i0, i1, i2] - kx[i0, i1, i2] * tmp
+
+    return rx, ry, rz
+
+
+@jit
+def proj_inplace(vx, vy, vz, kx, ky, kz, inv_k_square_nozero):
+    tmp = (kx * vx + ky * vy + kz * vz) * inv_k_square_nozero
+
+    vx -= kx * tmp
+    vy -= ky * tmp
+    vz -= kz * tmp
+
+
+@jit
+def proj_inplace_loop(vx, vy, vz, kx, ky, kz, inv_k_square_nozero):
+
+    n0, n1, n2 = kx.shape
+
+    for i0 in range(n0):
+        for i1 in range(n1):
+                for i2 in range(n2):
+                    tmp = (kx[i0, i1, i2] * vx[i0, i1, i2]
+                           + ky[i0, i1, i2] * vy[i0, i1, i2]
+                           + kz[i0, i1, i2] * vz[i0, i1, i2]
+                    ) * inv_k_square_nozero[i0, i1, i2]
+
[…]

### 2019-02-21 paper_01_fluidfft/python/makefile_figures.py

@@ -1,0 +1,71 @@
+
+import os
+import getpass
+from subprocess import call
+
+import matplotlib.pyplot as plt
+
+from fluiddyn.util import modification_date
+from fluiddyn.output.rcparams import set_rcparams
+
+from fluidfft.bench_analysis import plot_scaling
+
+set_rcparams(fontsize=14, for_article=True, for_beamer=False)
+plt.rc('legend',fontsize=10)
+
+
+here = os.path.abspath(os.path.dirname(__file__))
+here_tmp = os.path.join(here, '../tmp')
+
+
+if not os.path.exists(here_tmp):
+    os.mkdir(here_tmp)
+
+path_tmp = os.path.join('/tmp', getpass.getuser())
+if not os.path.exists(path_tmp):
+    os.mkdir(path_tmp)
+
+repo_fluidfft_bench_results = \
+    'https://bitbucket.org/fluiddyn/fluidfft-bench-results'
+path_fluidfft_bench_results = os.path.join(path_tmp, 'fluidfft-bench-results')
+
+has_to_pull = True
+
+if os.path.exists(path_fluidfft_bench_results):
+    if has_to_pull:
+        os.chdir(path_fluidfft_bench_results)
+        call(['hg', 'pull', '-u'])
+else:
+    os.chdir(path_tmp)
+    call(['hg', 'clone', repo_fluidfft_bench_results])
+
+os.chdir(here)
+
+
+def save_fig_scaling(dir_name, dim, n0, n1, n2=None):
+    path_dir = os.path.join(path_fluidfft_bench_results, dir_name)
+    path_fig = os.path.join(here_tmp, 'fig_' + dir_name + '.pdf')
+
+    if not os.path.exists(path_fig) or \
+       modification_date(path_dir) > modification_date(path_fig):
+        print('make fig', path_fig)
+        fig = plot_scaling(path_dir, None, dim, n0, n1, n2, show=False,
+                           for_latex=True)
+
+        for ax in fig.axes:
+            ax.set_xlabel(r"number of processes $n_p$")
+            ax.set_ylabel(r"speedup $S_\alpha(n_p)$")
+        fig.set_size_inches(10, 5)
+        fig.suptitle('')
+        fig.tight_layout(
+            pad=1.02, h_pad=None, w_pad=None, rect=(-0.01, -0.02, 1, 1)
+        )
[…]

### 2019-02-21 paper_02_fluidsim/Python/make_fig_bench.py

@@ -1,0 +1,65 @@
+#!/usr/bin/env python
+
+from pathlib import Path
+import matplotlib.pyplot as plt
+import numpy as np
+
+import bench_analysis as ba
+from base import matplotlib_rc
+from sync import path_fluidsim_bench_results
+
+
+matplotlib_rc()
+save = True
+root = Path(path_fluidsim_bench_results) / 'benchmarks'
+# pattern2d = root.glob('beskow_1024x1024')
+pattern2d = sorted(root.glob('beskow_?0??x????'))
+pattern3d = sorted(root.glob('beskow_*x*x*'))[::-1]
+
+
+def increase_lims(ax, fac=2):
+    fac = np.array([1/fac, fac])
+    xlim = np.array(ax.get_xlim()) * fac
+    ylim = np.array(ax.get_ylim()) * fac
+    ax.set_xlim(*xlim)
+    ax.set_ylim(*ylim)
+
+
+def plot_fig(pattern, solver, output_file=None, exclude=[]):
+    fig, axes = plt.subplots(1, 2)
+    ax0, ax1 = axes.ravel()
+    markers = ("x", ".")
+    for ipath, path_dir in enumerate(pattern):
+        t_min = 1e10
+        path_dir = str(path_dir)
+        N = path_dir.rpartition("_")[-1].split("x")
+        N = [int(n) for n in N]
+        fig, t_min = ba.plot_scaling(
+            path_dir, solver, 'any', *N,
+            show=False, type_plot='strong', fig=fig, ax0=ax0, ax1=ax1,
+            t_min_cum=t_min, default_marker=markers[ipath],
+            type_fft_exclude=exclude)
+
+    ax0.set_title('')
+    ax0.set_ylabel(r'$S$')
+    ax0.set_xlabel(r'$n_p$')
+    ax1.set_ylabel(r'Walltime per iteration (s)')
+    ax1.set_xlabel(r'$n_p$')
+
+    increase_lims(ax0)
+    increase_lims(ax1)
+    fig.set_size_inches(6, 3)
+    fig.suptitle('')
+    fig.tight_layout(
+        pad=1.02, h_pad=None, w_pad=None, rect=(-0.01, -0.02, 1, 1)
+    )
+    if output_file is None or not save:
+        plt.show()
+    else:
+        fig.savefig(output_file)
+
+    return fig, axes
+
+
[…]

### 2019-02-21 paper_02_fluidsim/Python/sync.py

@@ -1,0 +1,61 @@
+
+import os
+import getpass
+from subprocess import call
+
+# import matplotlib.pyplot as plt
+
+from fluiddyn.util import modification_date
+
+
+here = os.path.abspath(os.path.dirname(__file__))
+here_tmp = os.path.join(here, '../tmp')
+
+
+if not os.path.exists(here_tmp):
+    os.mkdir(here_tmp)
+
+path_tmp = os.path.join('/tmp', getpass.getuser())
+if not os.path.exists(path_tmp):
+    os.mkdir(path_tmp)
+
+repo_fluidsim_bench_results = \
+    'https://bitbucket.org/fluiddyn/fluidsim-bench-results'
+path_fluidsim_bench_results = os.path.join(path_tmp, 'fluidsim-bench-results')
+
+has_to_pull = True
+
+if os.path.exists(path_fluidsim_bench_results):
+    if has_to_pull:
+        os.chdir(path_fluidsim_bench_results)
+        call(['hg', 'pull', '-u'])
+else:
+    os.chdir(path_tmp)
+    call(['hg', 'clone', repo_fluidsim_bench_results])
+
+os.chdir(here)
+
+
+# def save_fig_scaling(dir_name, dim, n0, n1, n2=None):
+#     path_dir = os.path.join(path_fluidsim_bench_results, dir_name)
+#     path_fig = os.path.join(here_tmp, 'fig_' + dir_name + '.pdf')
+
+#     if not os.path.exists(path_fig) or \
+#        modification_date(path_dir) > modification_date(path_fig):
+#         print('make fig', path_fig)
+#         fig = plot_scaling(path_dir, None, dim, n0, n1, n2, show=False,
+#                            for_latex=True)
+#         fig.set_size_inches(10, 5)
+#         fig.suptitle('')
+#         fig.tight_layout(
+#             pad=1.02, h_pad=None, w_pad=None, rect=(-0.01, -0.02, 1, 1)
+#         )
+#         fig.savefig(path_fig, dpi=800)
+
+#
+#  save_fig_scaling('legi_cluster8_320x640x640', '3d', 320, 640, 640)
+#  save_fig_scaling('legi_cluster8_2160x2160', '2d', 2160, 2160)
+#  save_fig_scaling('occigen_384x1152x1152', '3d', 384, 1152, 1152)
[…]

### 2019-02-21 paper_02_fluidsim/compare_codes/make_fig_bar.py

@@ -1,0 +1,54 @@
+import os
+import sys
+
+import numpy as np
+import matplotlib.pyplot as plt
+
+from fluiddyn.output.rcparams import set_rcparams
+
+set_rcparams(fontsize=14, for_article=True, for_beamer=False)
+
+here = os.path.abspath(os.path.dirname(__file__))
+here_tmp = os.path.join(here, '../tmp')
+
+if not os.path.exists(here_tmp):
+    os.mkdir(here_tmp)
+
+keys = (
+    'total',
+    'FFT',
+    'RK4',
+    'curl',
+    'vector\nproduct',
+    'projection',
+)
+
+times_ns3d = np.array([9.52, 6.18, 1.91, 0.49, 0.35, 0.46])
+times_fluidsim = np.array([9.45, 6.72, 1.57, 0.44, 0.34, 0.30])
+left = 2.2*np.arange(len(times_ns3d))
+
+fig, ax = plt.subplots(1, figsize=(8, 4))
+
+shift = 0.9
+ax.bar(left, height=times_ns3d, color='b', edgecolor='k', yerr=0.07)
+ax.bar(left+shift, height=times_fluidsim, color='y', edgecolor='k', yerr=0.07)
+
+ax.set_xticks([])
+ax.set_xticklabels([])
+ax.set_ylabel('time (s)')
+# ax.set_title('outplace (with memory allocation)')
+
+y = -0.8
+for x, s in zip(left, keys):
+    ax.text(x+shift/2, y, s, rotation=0,
+            horizontalalignment='center',
+            verticalalignment='center')
+
+fig.tight_layout(rect=(0, 0.05, 1, 1))
+
+path_fig = os.path.join(here_tmp, 'fig_compare_with_ns3d.pdf')
+
+if 'save' in sys.argv:
+    fig.savefig(path_fig, dpi=800)
+else:
+    plt.show()

### 2019-02-21 paper_02_fluidsim/compare_codes/make_table.py

@@ -1,0 +1,18 @@
+"""
+
+pip install tabulate
+
+
+"""
+
+
+from tabulate import tabulate
+
+table = [
+    ["512$^2$",  0.51, 8.01, 0.92, 0.82],
+    ["1024$^2$", 2.69, 43.0, 3.48, 3.96]
+]
+
+headers = ["", r"\fluidpack{sim}", 'Dedalus', 'SpectralDNS', 'NS3D']
+
+print(tabulate(table, headers, tablefmt="latex_raw"))

### 2019-02-21 scripts/get_bib.py

@@ -1,0 +1,22 @@
+"""Receive file served by Zotero + Better BibTeX plugin."""
+import requests
+import time
+
+
+FORMAT = "biblatex"
+SRC = "thesis." + FORMAT
+URL = "http://localhost:23119/better-bibtex/collection?/1/" + SRC
+SAVE_AS = "thesis.bib"  # + FORMAT
+
+print("Fetching", SRC, end="... ")
+tstart = time.time()
+try:
+    response = requests.get(URL, timeout=2)
+except requests.exceptions.Timeout:
+    raise IOError(f"Cannot fetch {SRC}. Is Zotero open?")
+
+with open(SAVE_AS, "w") as bib:
+    bib.write(response.content.decode("utf8"))
+
+tend = time.time()
+print(f"Saved as {SAVE_AS} in {(tend - tstart):.3f} seconds")

### 2019-02-21 templates/utils_render.py

@@ -1,0 +1,124 @@
+#!/usr/bin/env python
+"""Render a latex jinja template
+
+Usage:
+    utils_render.py <options_file> <template_file>
+
+Notes:
+    See https://github.com/AKuederle/Py-Tex-automation-example
+
+"""
+import re
+import jinja2
+import os
+import sys
+from pathlib import Path
+
+
+def get_options_from_file(path):
+    """Read a options file.
+
+    The pattern in the file is '%<key>:<value>'
+    """
+    with open(path) as f:
+        content = f.read()
+        keys = re.findall(r"%(.+):", content)
+        values = re.findall(r":\s*([\w\W]+?)\s*(?:%|$)", content)
+
+    options = dict(zip(keys, values))
+    return options
+
+
+def name_in_initials(name):
+    """John Doe -> J. Doe"""
+    name = name.split()
+    first_name = name[:-1]
+    last_name = name[-1]
+    name = [n[0] + "." for n in first_name]
+    name.append(last_name)
+    return " ".join(name)
+
+
+def get_options_from_yaml(path):
+    """Read a options YAML file."""
+    import ruamel.yaml as yaml
+
+    yml = yaml.YAML()
+    with open(path) as f:
+        options = yml.load(f.read())
+
+    if "dir" not in options or options["dir"] is None:
+        options["dir"] = path.parent
+
+    options["authors_short"] = [
+        name_in_initials(name) for name in options["authors"]
+    ]
+
+    uniq_affiliations = list(set(options["affiliations"]))  # unique
+    options["authors_aff"] = []
+    for author, affiliation in zip(options["authors"], options["affiliations"]):
+        options["authors_aff"].append(
+            f"{author}$^{uniq_affiliations.index(affiliation) + 1}$"
+        )
+    options["enum_affiliations"] = [
+        f"$^{i+1}$ {aff}" for i, aff in enumerate(uniq_affiliations)
+    ]
+    
+
+    return options
+
+
+def get_template(template_file):
+    """Get a jinja template with latex tags.
+
+    modified from http://eosrei.net/articles/2015/11/latex-templates-python-and-jinja2-generate-pdfs
+    """
+    latex_jinja_env = jinja2.Environment(
+        block_start_string=r"%:",
+        block_end_string=":%",
+        variable_start_string=r"\VAR{",
+        variable_end_string="}",
+        comment_start_string=r"\#{",
+        comment_end_string="}",
+        line_statement_prefix="%%",
+        line_comment_prefix="%#",
+        trim_blocks=True,
[…]

### 2019-02-21 paper_00_fluiddyn_meta/article.tex
sed 's/section\*/section/g' article.tex
Subsection -> paragraphbf sed -i 's/subsubsection/paragraphbf/g' */article.tex
Overview etc. as sections and sub+other divisions sed -i 's/section\{/subsection\{/g' */article.tex sed -i 's/subsection[{].[[:alnum:]]. /section\{/g' */article.tex

@@
 
-% \section*{(1) Overview}
+\section{Overview}
 
@@
 
-% \section*{Title}
+% \subsection{Title}
 
@@
 
-% \section*{Paper Authors}
+% \subsection{Paper Authors}
 
@@
 %
-% \section*{Paper Author Roles and Affiliations}
+% \subsection{Paper Author Roles and Affiliations}
 % \textcolor{blue}{1. First author role and affiliation \\
@@
 
-% \section*{Abstract}
+% \subsection{Abstract}
 
@@
 
-% \section*{Keywords}
+% \subsection{Keywords}
 
@@
 
-\section*{Introduction}
+\subsection{Introduction}
 
@@
 
-\section{Implementation and architecture}
+\subsection{Implementation and architecture}
 
@@
 
-\subsection*{Organization of the code in packages}
+\subsubsection{Organization of the code in packages}
 
@@
 
-\subsection*{API of the Python library \fluiddyn}
+\subsection{of the Python library \fluiddyn}
 
@@
 
-\subsubsection*{Module \codeinline{fluiddyn.util.paramcontainer}}
+\paragraphbf{Module \codeinline{fluiddyn.util.paramcontainer}}
 
@@
 
-\subsubsection*{Module \codeinline{fluiddyn.util.serieofarrays}}
+\paragraphbf{Module \codeinline{fluiddyn.util.serieofarrays}}
 
@@
 
-\subsubsection*{Module \codeinline{fluiddyn.util.mpi}}
+\paragraphbf{Module \codeinline{fluiddyn.util.mpi}}
 
@@
 
-\subsubsection*{Module \codeinline{fluiddyn.calcul.easyfft}}
+\paragraphbf{Module \codeinline{fluiddyn.calcul.easyfft}}
 
@@
 
-\subsection*{The \fluiddyn command-line utilities}
+\subsection{\fluiddyn command-line utilities}
 
@@
 
-% \section*{Additional system requirements}
+% \subsection{Additional system requirements}
 
@@
 \end{itemize}
-\section*{List of contributors}
+\subsection{List of contributors}
 
@@
 
-% \section*{Software location:}
+\subsection{Software location}
+
+\paragraph{\bf Archive}
+% \textcolor{blue}{(e.g. institutional repository, general
+% repository) (required – please see instructions on journal website for
+% depositing archive copy of software in a suitable repository)}
+
+\begin{description}[noitemsep,topsep=0pt]
+\item[Name:] PyPI
+\item[Persistent identifier:] https://pypi.org/project/fluiddyn
+\item[Licence:] CeCILL-B, a BSD compatible French licence.
+\item[Publisher:] Pierre Augier
+\item[Version published:] 0.2.4
+\item[Date published:] 02/07/2018
+\end{description}
+
+{\bf Code repository}
+% \textcolor{blue}{(e.g. SourceForge, GitHub etc.) (required)}
+
+\begin{description}[noitemsep,topsep=0pt]
+\item[Name:] Bitbucket
+\item[Persistent identifier:] https://bitbucket.org/fluiddyn/fluiddyn
+\item[Licence:] CeCILL-B
+\item[Date published:] 2015
+\end{description}
+
+% {\bf Emulation environment} \textcolor{blue}{(if appropriate)}
 %
-% % {\bf Archive} \textcolor{blue}{(e.g. institutional repository, general
-% % repository) (required – please see instructions on journal website for
-% % depositing archive copy of software in a suitable repository)}
-%
-% \begin{description}[noitemsep,topsep=0pt]
-% \item[Name:] PyPI
-% \item[Persistent identifier:] https://pypi.org/project/fluiddyn
-% \item[Licence:] CeCILL-B, a BSD compatible French licence.
-% \item[Publisher:] Pierre Augier
-% \item[Version published:] 0.2.4
-% \item[Date published:] 02/07/2018
-% \end{description}
-%
-% {\bf Code repository}
-% % \textcolor{blue}{(e.g. SourceForge, GitHub etc.) (required)}
-%
-% \begin{description}[noitemsep,topsep=0pt]
-% \item[Name:] Bitbucket
-% \item[Persistent identifier:] https://bitbucket.org/fluiddyn/fluiddyn
-% \item[Licence:] CeCILL-B
-% \item[Date published:] 2015
-% \end{description}
-%
-% {\bf Emulation environment} \textcolor{blue}{(if appropriate)}
-
 % \begin{description}[noitemsep,topsep=0pt]
@@
 
-% \section*{Language}
-%
-% % \textcolor{blue}{Language of repository, software and supporting files.}
-%
-% English
+\subsection{Language}
 
-\section*{(3) Reuse potential}
+% \textcolor{blue}{Language of repository, software and supporting files.}
+
+English
+
+\section{Reuse potential}
 %
@@
 
-\subsection*{Conclusions}
+\subsection{Conclusions}
 
@@
 % already included by \bibliography{../bib}
-% \section*{References}
+% \subsection{References}
 
@@
 
-\rule{\textwidth}{1pt}
+% \rule{\textwidth}{1pt}
+\vspace{10pt}
+\hrule
+\vspace{10pt}
 
-{ \bf Copyright Notice} \\
-Authors who publish with this journal agree to the following terms: \\
+\paragraphbf{Copyright Notice}
+
+Authors who publish with this journal agree to the following terms:
 

### 2019-02-21 paper_02_fluidsim/article.tex
Overview etc. as sections and sub+other divisions sed -i 's/section\{/subsection\{/g' */article.tex sed -i 's/subsection[{].[[:alnum:]]. /section\{/g' */article.tex

@@
 
-{\bf Software paper for submission to the Journal of Open Research Software} \\
-
-\rule{\textwidth}{1pt}
-
-\section{(1) Overview}
-
-\vspace{0.5cm}
-
-\section{Title}
-
-FluidSim: modular, object-oriented Python package for high-performance CFD
-simulations
-
-\section{Paper Authors}
-
-1. MOHANAN, Ashwin Vishnu$^a$ \\
-2. BONAMY, Cyrille$^b$ \\
-3. CALPE LINARES, Miguel$^b$ \\
-4. AUGIER, Pierre$^b$
-
-\smallskip
-
-$^a$ Linn\'e Flow Centre, Department of Mechanics, KTH, 10044 Stockholm, Sweden.\\
-$^b$ Univ. Grenoble Alpes, CNRS, Grenoble INP\footnote{Institute of Engineering
-Univ. Grenoble Alpes}, LEGI, 38000 Grenoble, France.
-
-\section{Paper Author Roles and Affiliations}
-1. Ph.D. student, Linn\'e Flow Centre, KTH Royal Institute of Technology,
-Sweden; \\
-2. Research Engineer, LEGI, Universit\'e Grenoble Alpes, CNRS, France; \\
-3. Ph.D. student, LEGI, Universit\'e Grenoble Alpes, CNRS, France; \\
-4. Researcher, LEGI, Universit\'e Grenoble Alpes, CNRS, France
-
-\section{Abstract}
-
+% {\bf Software paper for submission to the Journal of Open Research Software} \\
+%
+% \rule{\textwidth}{1pt}
+
+\section{Overview}
+%
+% \vspace{0.5cm}
+%
+% \subsection{Title}
+%
+% FluidSim: modular, object-oriented Python package for high-performance CFD
+% simulations
+%
+% \subsection{Paper Authors}
+%
+% 1. MOHANAN, Ashwin Vishnu$^a$ \\
+% 2. BONAMY, Cyrille$^b$ \\
+% 3. CALPE LINARES, Miguel$^b$ \\
+% 4. AUGIER, Pierre$^b$
+%
+% \smallskip
+%
+% $^a$ Linn\'e Flow Centre, Department of Mechanics, KTH, 10044 Stockholm, Sweden.\\
+% $^b$ Univ. Grenoble Alpes, CNRS, Grenoble INP\footnote{Institute of Engineering
+% Univ. Grenoble Alpes}, LEGI, 38000 Grenoble, France.
+%
+% \subsection{Paper Author Roles and Affiliations}
+% 1. Ph.D. student, Linn\'e Flow Centre, KTH Royal Institute of Technology,
+% Sweden; \\
+% 2. Research Engineer, LEGI, Universit\'e Grenoble Alpes, CNRS, France; \\
+% 3. Ph.D. student, LEGI, Universit\'e Grenoble Alpes, CNRS, France; \\
+% 4. Researcher, LEGI, Universit\'e Grenoble Alpes, CNRS, France
+%
+% \subsection{Abstract}
+%
 %\textcolor{blue}{A short (ca. 100 word) summary of the software being
@@
 
-The Python package \fluidpack{sim} is introduced in this article as an extensible
-framework for Computational Fluid Mechanics (CFD) solvers.
-%
-It is developed as a part of \href{https://fluiddyn.readthedocs.io}{FluidDyn}
-project \citep{fluiddyn}, an effort to promote open-source and open-science
-collaboration within fluid mechanics community and intended for both educational
-as well as research purposes.
-%
-Solvers in \fluidpack{sim} are scalable, High-Performance Computing (HPC) codes
-which are powered under the hood by the rich, scientific Python ecosystem and the
-Application Programming Interfaces (API) provided by \fluidpack{dyn} and
-\fluidpack{fft} packages \citep{fluidfft}.
-%
-The present article describes the design aspects of \fluidpack{sim}, which
-includes use of Python as the main language; focus on the ease of use, reuse
-and maintenance of the code without compromising performance.
-%
-The implementation details including optimization methods, modular organization
-of features and object-oriented approach of using classes to implement solvers
-are also briefly explained.
-%
-Currently, \fluidpack{sim} includes solvers for a variety of physical problems
-using different numerical methods (including finite-difference methods).
-%
-However, this metapaper shall dwell only on the implementation and performance
-of its pseudo-spectral solvers, in particular the two- and three-dimensional
-Navier-Stokes solvers.
-%
-We investigate the performance and scalability of \fluidpack{sim} in a
-state of the art HPC cluster.
-%
-Three similar pseudo-spectral CFD codes based on Python (Dedalus, SpectralDNS) and
-Fortran (NS3D) are presented and qualitatively and quantitatively compared to
-\fluidpack{sim}.
-%
-The source code is hosted at Bitbucket as a Mercurial repository
-\href{https://bitbucket.org/fluiddyn/fluidsim}{bitbucket.org/fluiddyn/fluidsim}
-and the documentation generated using Sphinx can be read online at
-\href{https://fluidsim.readthedocs.io}{fluidsim.readthedocs.io}.
-
-\section{Keywords}
-
-% Python;\ CFD;\ HPC;\ MPI;\ modular;\ object-oriented;\ tested;\ documented;\
-% open-source
-0D; 1D; 2D; 3D; pseudo-spectral; finite difference; computational fluid
-dynamics; sequential; MPI; GPU
-
-\section{Introduction}
+% The Python package \fluidpack{sim} is introduced in this article as an extensible
+% framework for Computational Fluid Mechanics (CFD) solvers.
+% %
+% It is developed as a part of \href{https://fluiddyn.readthedocs.io}{FluidDyn}
+% project \citep{fluiddyn}, an effort to promote open-source and open-science
+% collaboration within fluid mechanics community and intended for both educational
+% as well as research purposes.
+% %
+% Solvers in \fluidpack{sim} are scalable, High-Performance Computing (HPC) codes
+% which are powered under the hood by the rich, scientific Python ecosystem and the
+% Application Programming Interfaces (API) provided by \fluidpack{dyn} and
+% \fluidpack{fft} packages \citep{fluidfft}.
+% %
+% The present article describes the design aspects of \fluidpack{sim}, which
+% includes use of Python as the main language; focus on the ease of use, reuse
+% and maintenance of the code without compromising performance.
+% %
+% The implementation details including optimization methods, modular organization
+% of features and object-oriented approach of using classes to implement solvers
+% are also briefly explained.
+% %
+% Currently, \fluidpack{sim} includes solvers for a variety of physical problems
+% using different numerical methods (including finite-difference methods).
+% %
+% However, this metapaper shall dwell only on the implementation and performance
+% of its pseudo-spectral solvers, in particular the two- and three-dimensional
+% Navier-Stokes solvers.
+% %
+% We investigate the performance and scalability of \fluidpack{sim} in a
+% state of the art HPC cluster.
+% %
+% Three similar pseudo-spectral CFD codes based on Python (Dedalus, SpectralDNS) and
+% Fortran (NS3D) are presented and qualitatively and quantitatively compared to
+% \fluidpack{sim}.
+% %
+% The source code is hosted at Bitbucket as a Mercurial repository
+% \href{https://bitbucket.org/fluiddyn/fluidsim}{bitbucket.org/fluiddyn/fluidsim}
+% and the documentation generated using Sphinx can be read online at
+% \href{https://fluidsim.readthedocs.io}{fluidsim.readthedocs.io}.
+%
+% \subsection{Keywords}
+%
+% % Python;\ CFD;\ HPC;\ MPI;\ modular;\ object-oriented;\ tested;\ documented;\
+% % open-source
+% 0D; 1D; 2D; 3D; pseudo-spectral; finite difference; computational fluid
+% dynamics; sequential; MPI; GPU
+
+\subsection{Introduction}
 
@@
 
-\subsection{Performance}
+\subsubsection{Performance}
 
@@
 
-\subsection{CFD pseudo-spectral code comparisons}
+\subsection{pseudo-spectral code comparisons}
 
@@
 
-\section{Quality control}
+\subsection{Quality control}
 
@@
 
-\section{(2) Availability}
+\section{Availability}
 \vspace{0.5cm}
-\section{Operating system}
+\subsection{Operating system}
 
@@
 
-\section{Programming language}
+\subsection{Programming language}
 
@@
 
-\section{Dependencies}
+\subsection{Dependencies}
 
@@
 
-\section{List of contributors}
+\subsection{List of contributors}
 
@@
 
-\section{Software location:}
-
-{\bf Archive}
+\subsection{Software location}
+
+\paragraph{\bf Archive}
 
@@
 
-\section{Language}
+\subsection{Language}
 
@@
 
-\section{(3) Reuse potential}
+\section{Reuse potential}
 
@@
 
-\section{Acknowledgements}
+\subsection*{Acknowledgements}
 
@@
 
-\section{Funding statement}
+\subsection*{Funding statement}
 
@@
 
-\section{Competing interests}
+\subsection*{Competing interests}
 
@@
 
-% \section{References}
+% \subsection{References}
 
@@
 
-\rule{\textwidth}{1pt}
-
-{\bf Copyright Notice} \\
+% \rule{\textwidth}{1pt}
+\vspace{10pt}
+\hrule
+\vspace{10pt}
+
+\paragraphbf{Copyright Notice}
+
 Authors who publish with this journal agree to the following terms: \\

### 2019-02-22 imgs/dependency/plot.py
Add sources for figures dependency and open_science - add pdf when finalized

@@ -1,0 +1,51 @@
+import matplotlib.pyplot as plt
+import networkx as nx
+from networkx.readwrite.adjlist import read_adjlist
+
+colors = {
+    "royalblue": (
+        "Python",
+        "Cython",
+        "Pythran",
+        "numpy.f2py",
+        "mpi4py",
+        "transonic",
+        "h5py",
+        "h5netcdf",
+        "NumPy",
+        "SciPy",
+        "IPython",
+        "Jupyter",
+        "SymPy",
+        "xarray",
+        "pandas",
+        "matplotlib",
+        "fluiddyn",
+        "fluidsim",
+        "fluidfft",
+    ),
+    "pink": (
+        "C",
+        "C++",
+        "xproject",
+        "xsimd",
+        "HDF5",
+        "FFTW",
+        "PFFT",
+    ),
+    "purple": ("Fortran", "P3DFFT"),
+    "slategrey": ("CUDA", "cuFFT", "OpenMP", "MPI", "SIMD"),
+}
+
+G = read_adjlist("graph.txt")
+node_color = []
+for node in G.nodes():
+    color = [k for k, v in colors.items() if node in v]
+    if color:
+        node_color.append(color[0])
+    else:
+        raise ValueError("No color assigned for %s" % node)
+
+nx.draw(G, with_labels=True, node_shape="s", node_color=node_color,
+        node_size=1000,)
+plt.show()

### 2019-02-22 imgs/dependency/plot_nxviz.py

@@ -1,0 +1,63 @@
+import matplotlib.pyplot as plt
+import networkx as nx
+from networkx.readwrite.adjlist import read_adjlist
+
+colors = {
+    "royalblue": (
+        "Python",
+        "Cython",
+        "Pythran",
+        "numpy.f2py",
+        "mpi4py",
+        "transonic",
+        "h5py",
+        "h5netcdf",
+        "NumPy",
+        "SciPy",
+        "IPython",
+        "Jupyter",
+        "SymPy",
+        "xarray",
+        "pandas",
+        "matplotlib",
+        "fluiddyn",
+        "fluidsim",
+        "fluidfft",
+    ),
+    "pink": (
+        "C",
+        "C++",
+        "xproject",
+        "xsimd",
+        "HDF5",
+        "FFTW",
+        "PFFT",
+    ),
+    "purple": ("Fortran", "P3DFFT"),
+    "slategrey": ("CUDA", "cuFFT", "OpenMP", "MPI", "SIMD"),
+}
+
+G = read_adjlist("graph.txt")
+node_color = []
+for node in G.nodes():
+    color = [k for k, v in colors.items() if node in v]
+    if color:
+        # node_color.append(color[0])
+        G.node[node]["value"] = color[0]
+    else:
+        raise ValueError("No color assigned for %s" % node)
+
+#  nx.draw(G, with_labels=True, node_shape="s", node_color=node_color,
+#          node_size=1000,)
+from nxviz.plots import CircosPlot
+c = CircosPlot(
+    G,
+    node_color="value",
+    node_order="value",
+    node_labels=True,
+    #, node_color=node_color
+)
+c.draw()
+plt.tight_layout()
+plt.savefig("../dependency.pdf")
+plt.show()

### 2019-02-22 frontmatter.tex
Update front matter and title etc

%-------------------------------------------------------------------------------
% Cover page
%-------------------------------------------------------------------------------
%
\makecoverpage

%-------------------------------------------------------------------------------
% Dedication (comment if none)
%-------------------------------------------------------------------------------
%
\dedication{
	{\it ``You must unlearn what you have learnt.''}\\[20pt]
	Master Yoda
}

%-------------------------------------------------------------------------------
% Abstract (English)
%-------------------------------------------------------------------------------
%
% COMMENT: the abstract should be limited to 2000 characters and the keywords
% to 250 characters because otherwise they will not fit in the mailing sheet.
%
% TODO: abstract in english and swedish
\begin{abstract}
        To do
	% Access points  must work. In fact, few cryptographers would disagree with
	% the emulation of robots, which embodies the technical principles of
	% complexity theory. Hyp, our new application for lossless information, is the
	% solution to all of these challenges.
	%         %
	%         \keywords{super-lasers, space-stations, The Force, cookies.}
	%         %
\end{abstract}

%-------------------------------------------------------------------------------
% Abstrakt (Swedish) 
%-------------------------------------------------------------------------------
%
% COMMENT: the abstract should be limited to 2000 characters and the keywords
% to 250 characters because otherwise they will not fit in the mailing sheet.
%
\begin{abstrakt}
        To do
	% \r{A}tkomstpunkter m\r{a}ste arbeta. I sj\"{a}lva verket skulle f\r{a}
	% kryptografer h\r{a}ller inte med emulering av robotar, som
	% f\"{o}rkroppsligar de tekniska principer av komplexitetsteori. Hyp, v\r{a}r
	% nya applikation f\"{o}r f\"{o}rlustfri information är l\"{o}sningen på alla
	% dessa utmaningar.
	%         %
	%         \nyckelord{super-lasrar, rymdstationer, Kraften, kakor.}
	%         %
\end{abstrakt}

%-------------------------------------------------------------------------------
% Preface page
%-------------------------------------------------------------------------------
%
\begin{preface}
        % TODO: preface
        To do.
	% This thesis deals with high energy super-lasers. A brief introduction on the
	% basic concepts and methods is presented in the first part. The second part
	% contains two articles.
	% The papers are adjusted to comply with the present thesis format for
	% consistency, but their contents have not been altered as compared with their
	% original counterparts.
\end{preface}

%-------------------------------------------------------------------------------
% Division of work between authors
%-------------------------------------------------------------------------------
%
%TODO: division of work
\begin{divisionofwork}
	The main advisor for the project is Erik Lindborg (EL).
	Pierre Augier (PA) acts as co-advisor.

        To do: paper wise list
	% \paperitem
	%         The code has been developed by Anakin Skywalker (AS).
	%         The paper has been written by AS and Darth Vader (DV).
        %
	% \paperitem
	%         The experimental set-up has been designed by DS.
	%         The simulations have been performed by AS using the control-code
	%         developed by DS.
	%         The paper has been written by AS and DV with feedback from OK and DS.

\end{divisionofwork}

%TODO: fill in additional publications and conferences
%-------------------------------------------------------------------------------
% Additional publications (comment if none)
%-------------------------------------------------------------------------------
%
% \begin{otherpublications}
%         The following papers, although related, are not included in this thesis.
%
%   \paperitem%
%     {Anakin Skywalker \& Master Obi-Wan Kenobi}% Authors
%     {3639}% Year
%     {The light sabre: an elegant weapon for a more civilized age}% Title
%     {Jedi Journal of Weapons}% Journal
%     {33}% Volume
%     {2}% Number
%     {pp. 55--60}% Pages
%
% \end{otherpublications}

%-------------------------------------------------------------------------------
% Conferences (comment if none)
%-------------------------------------------------------------------------------
%
% \begin{conferences}
%         Part of the work in this thesis has been presented at the following
%         international conferences. The presenting author is underlined.
%
%   \conferenceitem%
%     {Anakin Skywalker, \underline{Darth Vader} \& Admiral Motti}% Authors
%     {3639}% Year
%     {Effects on breathing of underestimation of the power of The Force}% Title
%     {Intergalactic Conference on Weapons and Peace}% Conference
%     {Kamino}% Location
%
% \end{conferences}

%-------------------------------------------------------------------------------
% Table of contents
%-------------------------------------------------------------------------------
%
\tableofcontents

### 2019-02-22 chapter_00_open_science.tex
Describe open science

@@
-%% Journal of Open Research Software Latex template -- Created By Stephen
-%% Bonner and John Brennan, Durham Universtiy, UK.
-%% see http://openresearchsoftware.metajnl.com
-
 \chapter{Towards reproducible, open science through open research software}
-%TODO: Fix table of contents
 \section{Introduction}
 
-% \vspace{0.5cm}
-
-% \section*{Title}
-
-% \textcolor{blue}{The title of the software paper should focus on the software,
-% e.g. “Text mining software from the X project”. If the software is closely
-% linked to a specific research paper, then “Software from Paper Title” is
-% appropriate. The title should be factual, relating to the functionality of the
-% software and the area it relates to rather than making claims about the
-% software, e.g. “Easy-to-use”.}
-
-% old title
-
-% FluidDyn: a Python open-source framework for research and teaching in fluid
-% dynamics
-
-% pa: this title may not be really good for this paper...
-% - "research and teaching" may not be very important in the title
-% - Recent progress in open-source software engineering and consequences on
-%   science
-
-% todo: find a better title!
-
-% Taking advantage of recent advances in open-source software engineering to study
-% fluid dynamics: the example of the FluidDyn project
-
-% \textcolor{blue}{An overview of the software, how it was produced, and the
-% research for which it has been used, including references to relevant research
-% articles. A short comparison with software which implements similar
-% functionality should be included in this section. }
-
-\subsection{Science, software, open-source and the computer revolution}
-
-Science is mainly a collective activity.  We can go further only by
-\href{https://en.wikipedia.org/wiki/%
-Standing_on_the_shoulders_of_giants}{``standing on the shoulders of giants''}
-(and of a huge number of technicians and scientists).  Science is a lot about
-how to build new knowledge from the work of others and hence, the exchange
-of ideas is a fundamental aspect.
-%
-In the last decades, we have lived through a revolution on how people exchange
-ideas.
-%
-Computers of all kinds (from smartphones to HPC clusters) connected by a world
-wide web are used for human communication and also for many other applications.
-%
-It has become nearly effortless to reproduce and exchange ideas and data.
-%
-The set of intangibles that grows when shared and degrades when hoarded, such as
-knowledge and love, has been somehow extended.
-% Relevant quote: 
-% O Goddess Saraswati, your Treasure is unique in nature.  It increases when it
-% is freely spent and gets destroyed if it is hoarded (and not shared with
-% others).
-%
-The information technology revolution open doors to fantastic opportunities for
-human collaboration, and of course, for science.
-
-Web related activities alone account for at least 5\% of GDP in the USA and the
-European Union\footnote{See the report by Internet Association titled
-\href{https://internetassociation.org/reports/refreshing-understanding-internet-economy-ia-report/}
-{``Refreshing Our Understanding of the Internet Economy''}}.
-% \cite{hooton_christopher_refreshing_2017}.
-A huge amount of money (and work) is invested on the related technologies. We
-are familiar with the most prominent companies involved in this dynamics
-(Google, Facebook, etc.), but there are also several smaller and sometimes
-lesser known organizations.  Most of these companies base part of their work on
-the open-source paradigm contributing to open-source languages, libraries,
-software and operating systems, while also using them --- a win-win situation
-for both the corporations and the community.
-%
-This has lead to deep changes in software engineering, with a massive use of
-open-source methods and tools, for example
-\href{https://en.wikipedia.org/wiki/Distributed_version_control}{distributed
-version control systems (DVCS)} and web-based source development platforms.
-
-The computer performance continues to
-\href{https://en.wikipedia.org/wiki/Moore\%27s_law}{increase exponentially},
-now also with the help of Graphical Processing Units (GPU). This gave way to a
-big boom in practical uses of data science and machine learning, which drives a
-strong research on artificial intelligence.  This dynamics contributes to
-progresses in open-source software.
-
-These changes in our world also reflect in the way science is done.
-%
-Software and programming in science occupy a much bigger place than before.
-%
+% \subsection{Science, software, open-source and the computer revolution}
+
+% FIXME: Deleted
+% Science: giants, collective, grows when shared
+% Web: money, open-source, tools esp. DVCS
+% Computing: performance, GPU
+% Changes: dynamics, way science is done, coding inferior, open science
+
 % see http://sciencecodemanifesto.org/
@@
 %
-The role of software in science has changed. In the past, coding was sometimes
-considered as an inferior activity by some scientists.  The focus was on the
-theory and the mathematical demonstration, which had to be elegant as it gets
-included in the articles.  In contrast, it was normal to write crude code
-% av: crappy = shitty!
-% alternatives: shabby, inelegant, amateurish, unpolished, clumsy
-% https://en.oxforddictionaries.com/thesaurus/amateurish
-and to just show the results. Nowadays, codes tend to be at the heart of
-research.
-
-``Open-science'' is a new trend taking advantage of these new facts.  Pioneering
-attempts are being made to do better science, improving reproducibility and
-collective efficiency, by using the open-source methods and tools for science
-and sharing and collaborating via the world wide web.
-
 % http://www.nature.com/news/interactive-notebooks-sharing-the-code-1.16261
-
-This paper presents FluidDyn, a project to foster open-science and open-source in
-fluid mechanics. This first section ``Overview'' is organized into two large
-parts.  A long introduction presents the context and describes our motivations and
-the concept behind the project FluidDyn.  It is organized as follows:
+% We provide examples of solutions for: Good coding practice, source control
+% management, packaging, licenses documentation, testing,CI
+
+% Fluiddyn: foster open-science, source. Framework for collaborative dev, open
+% methods
+
+This chapter presents FluidDyn, a project and an ecosystem of packages to
+facilitate open-science and open-source in the field of fluid mechanics.  The
+chapter is organized into several parts to present the context and describes
+our motivations and the concept behind the project FluidDyn as follows:
+
 \begin{itemize}
@@
 
-\item FluidDyn is based on the language Python. Incidently, it is the defining
+\item FluidDyn is based on the language Python. Incidentally, it is the defining
 characteristics of Python which make the eventual success of such a
 project realizable.
-
+%
 Python is one of the most important tool in recent open-source dynamics and in
@@
 \item The introduction finally ends with a summary of the motivations for the
-project FluidDyn and a presentation of its principles.
+project FluidDyn and a summary of the packages used\footnote{%
+  We use FluidDyn (with capital letters) to name the project and \fluiddyn for
+the base package.}.
 \end{itemize}
 
-%TODO: Fix names
-The section ``Introduction'' is completed by a second subsection called
-``Implementation and architecture'' devoted to present the organization of the
-packages of the project and to describe in more detail the base package
-\fluiddyn\footnote{We use FluidDyn (with capital letters) to name the project and
-\fluiddyn for the base package.}.
+\section{A modern approach to computational sciences}
+
+Computational sciences, and especially computational fluid dynamics (CFD) as a
+discipline has flourished a lot in the latter half of the twentieth century.
+Incidentally the origin of this domain can be traced back to the ideas
+in \citet{richardson_weather_1922} to achieve numerical weather prediction.
+With the arrival of first computer general-purpose computer, ENIAC, the first
+simulations were made using a simple barotropic model to make 24 hour
+forecasts \citep{lynch_richardson_2010}. In those early days when the use and
+availability of computers were limited, such methods were neither widespread
+nor feasible. Scientific investigations were dominated by theoretical and
+experimental methods.
+We have come a long way since then and today one can say that computational
+methods now complement theoretical and experimental studies and are equally
+important.
+Thus to recreate a particular scientific study, the mathematical representation
+or even an algorithms leaves a lot to be desired. With the connectivity
+provided by the world wide web it is possible to do more and reinvent the way
+we perform sciences.
+
+\begin{figure}[h]
+  \includegraphics[width=0.8\textwidth]{open_science}
+  \caption{A schematic representation of various concepts and methodologies
+  involved in achieving reproducibility in sciences.}\label{fig:opensci}
+\end{figure}
+
+In recent years there has been a move towards the open science
+\citep{royal_society_great_britain_science_2012}. Depicted in figure
+\ref{fig:opensci} are some of steps needed to ensure reproducibility and
+transparency in sciences. \emph{Knowledge} has been traditionally disseminated
+in the form of courses and books from universities, and to the wider scientific
+community through research articles, conference proceedings and workshops.
+While this is nothing new, knowledge has become more accessible to the public
+through public-domain and Creative Commons (or similar) licensed course
+materials and massively open online courses (MOOCs). Moreover, there is also an
+emerging push to publish in open-access journals either through systematic
+reforms in academia on how researchers are assessed on their productivity or
+through regulations and offering incentives, such as waiving the application
+processing charge for publishing \citep{nosek_promoting_2015}.
+
+As mentioned earlier there is a lot of work involved in condensing and
+implicating this knowledge into working applications. The \emph{know-how} can
+be made accessible and open to scrutiny by making research software open-source
+with an appropriate license. To make the codes usable it has to be complemented
+by documentation which typically includes tutorials, examples and a detailed
+commentary of different components of the code. Furthermore, to inspire
+confidence in and to ensure reliability of the code throughout the process of
+software development, it needs unit-tests with which a code is continuously
+deployed, built and tested --- a process known as Continuous Integration (CI),
+which is described in detail later in this chapter. Depending on how much
+matured a software is, these can be easy or cumbersome to implement, but
+nevertheless, the result is a cleaner and reliable code.
+
+Finally, the \emph{results} generated from the such codes together with the
+workflow (i.e. scripts to run the codes and to postprocess the data) when made
+open in the form of datasets can ensure that research would be reproducible and
+preserved in archives for several years to come. Open datasets can be shared
+and made citeable through services such as \href{https://zenodo.org}{Zenodo}
+and \href{https://figshare.com}{Figshare}.
 
@@
 
-\paragraph{Free and Open-Source Software (FOSS).} The term \emph{free} in
+\paragraphbf{Free and Open-Source Software (FOSS)} The term \emph{free} in
 FOSS is a misnomer, as it actually stands for \emph{freedom} (to use, modify
@@
 
-\paragraph{Distributed Version Control Systems (DVCS) and web-based source
-development platforms.}  Collaboration was achieved in the early years of FOSS
+\paragraphbf{Distributed Version Control Systems (DVCS) and web-based source
+development platforms} Collaboration was achieved in the early years of FOSS
 development through emailing ``patches'', and centralized VCS near the turn of
@@
 
-\paragraph{Package management and third-party software repositories.} Python
+\paragraphbf{Package management and third-party software repositories} Python
 as a medium for scientific computing owes its success to a rich ecosystem of
@@
 The resulting extensions are usually as fast as Fortran or C++ written by
-non-specialists \cite[see, for example, benchmarks for the packages
+non-specialists \citep[see, for example, benchmarks for the packages
 \fluidpack{fft} and \fluidpack{sim},][]{fluidfft, fluidsim}.
@@
 A detailed presentation on the above packages can be found in their respective
-documentations on the web and for \fluidpack{fft} and \fluidpack{sim} in the
-two companion papers \cite[]{fluidfft, fluidsim}.  This article will now focus
-on the base package \fluiddyn.
-
-All functions and classes defined in \fluiddyn are pure Python elements, meaning
-that no extensions are implemented in \fluiddyn.  Thus the package \fluiddyn is
-extremely easy to install with just a \codeinline{pip install} command and no
-compilation.
-
-The code base was designed to follow Python 2.7 syntax during its genesis. Now,
-it has been made forward compatible with Python 3 through the use of external
-package future.
-
-\subsection{API of the Python library \fluiddyn}
-
-The package \fluiddyn is organized into five sub-packages: 
-\begin{itemize}
-\item \codeinline{fluiddyn.io}: This subpackage provides utilities for
-input/output to different file formats.
-\item \codeinline{fluiddyn.util}: Miscellaneous utilities.
-\item \codeinline{fluiddyn.calcul}: Modules for numerical computations.
-\item \codeinline{fluiddyn.clusters}: Classes to launch jobs on HPC clusters.
-\item \codeinline{fluiddyn.output}: Utilities to produce scientific outputs
-(figures and videos).
-\end{itemize}
-
-Sub-packages \codeinline{io}, \codeinline{util} and \codeinline{calcul} are the
-largest in terms of lines of code and provides the Application Programming
-Interfaces (API) to support \fluidpack{sim}, \fluidpack{lab} and
-\fluidpack{image}.
+documentations on the web and for \fluidpack{dyn}, \fluidpack{fft} and
+\fluidpack{sim} in the respective software metapapers
+\citep[]{fluiddyn, fluidfft, fluidsim}.
 

### 2019-02-22 chapter_00_open_science.tex
Formatting and insert dependecy figure

@@
 \item The introduction finally ends with a summary of the motivations for the
-project FluidDyn and a summary of the packages used\footnote{%
+project FluidDyn and an overview of the packages used\footnote{%
   We use FluidDyn (with capital letters) to name the project and \fluiddyn for
@@
 Thus to recreate a particular scientific study, the mathematical representation
-or even an algorithms leaves a lot to be desired. With the connectivity
+or even an algorithm leaves a lot to be desired. With the connectivity
 provided by the world wide web it is possible to do more and reinvent the way
@@
 \begin{figure}[h]
+  \centering
   \includegraphics[width=0.8\textwidth]{open_science}
@@
 open in the form of datasets can ensure that research would be reproducible and
-preserved in archives for several years to come. Open datasets can be shared
-and made citeable through services such as \href{https://zenodo.org}{Zenodo}
-and \href{https://figshare.com}{Figshare}.
+preserved in archives for several years to come \citep{gewin_data_2016}. Open
+datasets can be shared and made citeable through services such as
+\href{https://zenodo.org}{Zenodo} and \href{https://figshare.com}{Figshare}.
+
+In the present work, through FluidDyn project, we have managed to implement all
+the aspects of making the know-how open through open-source software. We shall
+now take a look at the tools which have made it possible.
 
@@
 non-profit organizations and government institutions. This could be regarded as
-the second-age of FOSS movement~\cite{fitzgerald_transformation_2006}.
+the second-age of FOSS movement~\citep{fitzgerald_transformation_2006}.
 
@@
 
-\paragraph{Continuous integration (CI).} The need for strong reliability of
+\paragraphbf{Continuous integration (CI)} The need for strong reliability of
 programs with increasing complexity has forced developers to use continuous
@@
 
-\paragraph{Websites to share knowledge.} Mutual aid and knowledge sharing
+\paragraphbf{Websites to share knowledge} Mutual aid and knowledge sharing
 have been important aspects of the web since its inception and these practices
@@
 
-\paragraph{Automatic web documentations built on servers.}  Documentation
+\paragraphbf{Automatic web documentations built on servers} Documentation
 generators such as Sphinx (for Python) and Doxygen (for many languages
@@
 %
-As a scientists, we can wonder how to fully exploit these new possibilities to do
+As scientists, we can wonder how to fully exploit these new possibilities to do
 science.  The project FluidDyn is an attempt to help the fluid dynamics community
@@
 
-\subsection{Python, a programming language adapted for open-science}
+\section{Python, a programming language adapted for open-science}
 
@@
 
-\paragraphbf{Some Python issues}
-
-\paragraph{Since the standard Python interpreter does not do any proper
-compilation,} Python code in some CPU bounded cases can be too slow.  Thus, one
+\subsection{Some Python issues and their solutions}
+
+Since the standard Python interpreter does not do any proper
+compilation, Python code in some CPU bounded cases can be too slow. Thus, one
 has to use special tools to get optimized machine instructions to speedup
@@
 
-\textbf{Ahead Of Time compilation.}  Since Python is primarily a ``glue''
+\paragraphbf{Ahead Of Time compilation}  Since Python is primarily a ``glue''
 language, a common solution is to use code written in C, C++ or Fortran.
@@
 
-We present in the two companion papers \citep{fluidfft, fluidsim} examples of
+We present in \citet{fluidfft, fluidsim} examples of
 highly efficient Python codes, optimized using a combination of Cython and
@@
 
-\textbf{Just In Time (JIT) compilation.} Another way to obtain optimized
+\paragraphbf{Just In Time (JIT) compilation} Another way to obtain optimized
 machine instructions is by using JIT compilation, i.e.\ by compiling only the
@@
 
-\paragraph{No type-checking.} In contrast to languages using Ahead Of Time
+\paragraphbf{No type-checking} In contrast to languages using Ahead Of Time
 compilation, there is no type-checking in Python (and more generally in many
@@
 
-\paragraph{A lively, huge and thus complicated ecosystem.} As already
+\paragraphbf{A lively, huge and thus complicated ecosystem} As already
 mentioned, Python is a versatile language. For many applications, the standard
@@
 
-\paragraph{Absence of multicore computational parallelism using threads.}
+\paragraphbf{Absence of multicore computational parallelism using threads}
 Threads behave like light subprograms. In principle, they can use the different
@@
 
-\paragraphbf{Rich landscape of open-source languages useful for science}
+\subsection{Rich landscape of open-source languages useful for science}
 
@@
 for the project FluidDyn.
+
+\section{On software development methodologies}
 
@@
 
-\paragraphbf{Discussion on Matlab, Labview and Mathematica}
+\subsection{Discussion on the use of proprietary software for research}
+% \paragraphbf{Discussion on Matlab, Labview and Mathematica}
 
@@
 
-Matlab is a closed-source proprietary numerical computing environment.
+Matlab is an example of a closed-source proprietary numerical computing
+environment.
 %
@@
 people do need to pay to use the software. Even the authors of the software give
-money to the same company to be able to use their software. Do they realize that
-when they work on these programs, they work (often, for free) for the company that
-sells the proprietary tool?
+money to the same company to be able to use their software. 
+%
+% Do they realize that when they work on these programs,
+Ironically, when people develop programs using these tools, they work (often,
+for free) for the company that sells the proprietary tool.
 
@@
 
-\section{FluidDyn project: Breaking the status quo}
+\section{FluidDyn project}
+%: Breaking the status quo}
 % \section{Implementation and architecture}
@@
 %
+What we have now is not by ``reinventing the wheel'', but rather by ``standing
+on the shoulders of giants''. In other words we have utilized and extended
+existing and mature packages and libraries to create packages in FluidDyn as
+demonstrated in figure \ref{fig:dependency}. We have designed them such that
+the packages are generic enough for a wider audience and are maintainable in
+the long run.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=0.8\textwidth]{dependency}
+  \caption{An illustration of the various C/C++ (green), Fortran (gray), Python
+  (pink) packages and language agnostic frameworks (yellow) utilized in
+  FluidDyn packages \fluidpack{dyn}, \fluidpack{fft} and \fluidpack{sim} and
+  how they depend on each other, denoted by the lines connecting the
+  circles.}\label{fig:dependency}
+\end{figure}
+
 % It first started with one package.  Soon we had to prefer a decentralized
@@
 
-\section{Conclusion}
+% \section{Conclusion}
 

### 2019-03-03 chapter_00_open_science.tex
Cleanup and add todos

@@
 This chapter presents FluidDyn, a project and an ecosystem of packages to
-facilitate open-science and open-source in the field of fluid mechanics.  The
-chapter is organized into several parts to present the context and describes
-our motivations and the concept behind the project FluidDyn as follows:
+facilitate open-science and open-source development in the field of fluid
+mechanics.
+%
+The chapter is organized into several parts to present the context and to
+describe our motivations and the concept behind the project FluidDyn as
+follows:
 
@@
 tools for open-source software engineering. We start by presenting these
-developments and discuss the implications for the scientists.
+developments and discuss the implications for scientists.
 
@@
 %
-Python is one of the most important tool in recent open-source dynamics and in
-particular in science.
+Python is one of the most important tool in recent open-source dynamics,
+particularly in science.
 %
@@
 in \citet{richardson_weather_1922} to achieve numerical weather prediction.
-With the arrival of first computer general-purpose computer, ENIAC, the first
+With the arrival of first electronic general-purpose computer, ENIAC, the first
 simulations were made using a simple barotropic model to make 24 hour
@@
 experimental methods.
-We have come a long way since then and today one can say that computational
-methods now complement theoretical and experimental studies and are equally
+We have come a long way since then and today one can say that, computational
+methods now complement theoretical and experimental studies, and are equally
 important.
@@
 In recent years there has been a move towards the open science
-\citep{royal_society_great_britain_science_2012}. Depicted in figure
-\ref{fig:opensci} are some of steps needed to ensure reproducibility and
-transparency in sciences. \emph{Knowledge} has been traditionally disseminated
-in the form of courses and books from universities, and to the wider scientific
-community through research articles, conference proceedings and workshops.
+\citep[see the report by][]{royal_society_great_britain_science_2012}.
+%
+Depicted in figure \ref{fig:opensci} are some of steps needed to ensure
+reproducibility and transparency in sciences.
+%
+\emph{Knowledge} has been traditionally disseminated in the form of courses and
+books from universities, and to the wider scientific community through research
+articles, conference proceedings and workshops.
+
 While this is nothing new, knowledge has become more accessible to the public
@@
 materials and massively open online courses (MOOCs). Moreover, there is also an
-emerging push to publish in open-access journals either through systematic
+emerging push to publish in open-access journals\footnote{In the European
+  Union, a proposal to implement Plan S (\url{https://www.coalition-s.org}) was
+  been set in motion in September 2018 to fast track adoption of open-access.
+} either through systematic
 reforms in academia on how researchers are assessed on their productivity or
@@
 As mentioned earlier there is a lot of work involved in condensing and
-implicating this knowledge into working applications. The \emph{know-how} can
-be made accessible and open to scrutiny by making research software open-source
-with an appropriate license. To make the codes usable it has to be complemented
-by documentation which typically includes tutorials, examples and a detailed
-commentary of different components of the code. Furthermore, to inspire
-confidence in and to ensure reliability of the code throughout the process of
-software development, it needs unit-tests with which a code is continuously
-deployed, built and tested --- a process known as Continuous Integration (CI),
-which is described in detail later in this chapter. Depending on how much
-matured a software is, these can be easy or cumbersome to implement, but
-nevertheless, the result is a cleaner and reliable code.
+implementing this knowledge in the form of working applications.
+%
+The \emph{know-how} can be made accessible and open to scrutiny by making
+research software open-source with an appropriate license.  To make the codes
+usable it has to be complemented by documentation which typically includes
+tutorials, examples and a detailed commentary of different components of the code.
+%
+Furthermore, to inspire confidence in and to ensure reliability of the code
+throughout the process of software development, it needs unit-tests with
+which a code is continuously deployed, built and tested --- a process known as
+Continuous Integration (CI), which is described in detail later in this
+chapter. Depending on how much matured a software is, these can be easy or
+cumbersome to implement, but nevertheless, the result is a cleaner and reliable
+code.
 
@@
 preserved in archives for several years to come \citep{gewin_data_2016}. Open
-datasets can be shared and made citeable through services such as
+datasets can be shared and made citable through services such as
 \href{https://zenodo.org}{Zenodo} and \href{https://figshare.com}{Figshare}.
@@
 the second-age of FOSS movement~\citep{fitzgerald_transformation_2006}.
-
-%todo: DONE
-% Here, we have to add a paragraph on free and open-source (libre)
-% software:
-
-% The following facts can be mentioned:
-
-% - tex (started in 1977), latex (1985)
-
-% - GNU (started 1983), FSF in 1985, licence and philosophy...
-% first complete system (with the Linux kernel) in 1991
 
@@
 of questions and answers on programming each week.
-% todo: remark from Julien Salort: it's old.
-% av: So are technologies like git and hg, but we are putting together what
-% characterizes our workflow, aren't we? We can emphasize that these forums
-% often work only when there is a "thriving community" like the Python
-% userbase.
 IRC protocol has continued to sustain the test of time. Instant messaging
@@
 such incredible success.
-
-% todo: DONE
-% We need an introductory paragraph on how big is Python today
-% + the fact that several scientists miss this fact.
-
-% See the file ./python-one-main-languages.rst
 
@@
 language, a common solution is to use code written in C, C++ or Fortran.
-%
-% todo: DONE
-% add a reference to ctypes and cffi to call C with pure Python code
 %
@@
 
-% todo: DONE
+% TODO: perhaps include the commented out sections in the beginning of the chapter
 % pa: I am really uncomfortable with this presentation which gives the
@@
 % \section{Implementation and architecture}
-
+% TODO: a lot of self-plagiarism below - needs rewrite
 We have seen that there is a strong dynamics in play around the use of
@@
 %
-% % todo: again mention some success of community-driven subject-area research
+% % again mention some success of community-driven subject-area research
 % % libraries? See the file ./communitydriven_research_soft.md

### 2019-04-24 paper_03_toy_model/article.tex
Add paper 3 toy model - published version

%% ****** Start of file aiptemplate.tex ****** %
%%
%%   This file is part of the files in the distribution of AIP substyles for REVTeX4.
%%   Version 4.1 of 9 October 2009.
%%
%
% This is a template for producing documents for use with 
% the REVTEX 4.1 document class and the AIP substyles.
% 
% Copy this file to another name and then work on that file.
% That way, you always have this original template file to use.

% \documentclass[aip,pof,reprint]{revtex4-1}
% %\documentclass[aip,graphicx]{revtex4-1}
% %\documentclass[aip,reprint]{revtex4-1}
%
% \usepackage{graphicx}
% \usepackage[paperwidth=210mm,
%             paperheight=297mm,
%             centering,
%             hmargin=2cm,
%             vmargin=2.5cm]{geometry}
% \usepackage[colorlinks=true,
%             linkcolor=blue,
%             urlcolor=blue,
%             citecolor=blue]{hyperref}
% \graphicspath{{../Pyfig/to_convert/}}
% \usepackage{color}  % Probably not required, and was used for revisions
% \draft % marks overfull lines with a black rule on the right

% \begin{document}

% Use the \preprint command to place your local institutional report number 
% on the title page in preprint mode.
% Multiple \preprint commands are allowed.
%\preprint{}

% \title{A two-dimensional toy model for geophysical turbulence} %Title of paper

% repeat the \author .. \affiliation  etc. as needed
% \email, \thanks, \homepage, \altaffiliation all apply to the current author.
% Explanatory text should go in the []'s, 
% actual e-mail address or url should go in the {}'s for \email and \homepage.
% Please use the appropriate macro for the type of information

% \affiliation command applies to all authors since the last \affiliation command. 
% The \affiliation command should follow the other information.

\author{Erik Lindborg}
%\email[]{Your e-mail address}
%\homepage[]{Your web page}
%\thanks{}
%\altaffiliation{}
\affiliation{Department of Mechanics, Osquars Backe 18, Royal Institute of Technology, Stockholm, Sweden}
\author{Ashwin Vishnu Mohanan}
%\email[]{Your e-mail address}
%\homepage[]{Your web page}
%\thanks{}
%\altaffiliation{}
% \affiliation{Department of Mechanics, Osquars Backe 18, Royal Institute of Technology, Stockholm, Sweden}

% Collaboration name, if desired (requires use of superscriptaddress option in \documentclass). 
% \noaffiliation is required (may also be used with the \author command).
%\collaboration{}
%\noaffiliation

% \date{\today}
%
% \begin{abstract}
% A toy model for large scale geophysical turbulence is constructed by making two modifications of the shallow water model. Unlike the shallow water model the toy model has a quadratic expression for total energy, which is the sum of Available Potential Energy (APE) and Kinetic Energy (KE). More importantly, in contrast to the shallow water model the toy model does not produce any shocks. Three numerical simulations with different forcing are presented and compared with the simulation of a full General Circulation Model (GCM).  The energy which is injected cascades in a similar way as  in the GCM.
% First, some of the energy is converted from APE to KE at large scales.  The wave field is then undergoing a forward energy cascade displaying shallow spectra, close to $ k^{-5/3} $, for both APE and KE,  {while the vortical field
% either displays a $ k^{-3} $-spectrum or a more shallow spectrum, close to $ k^{-5/3} $, depending on the forcing. }
% In a simulation with medium forcing wave number, some of the energy which is converted from APE to KE is undergoing an inverse energy cascade which is produced by nonlinear interactions only involving the rotational component of the velocity field.  The inverse energy cascade builds up a vortical field at larger scales than the forcing scale. At these scales coherent vortices emerge with a strong dominance of anticyclonic vortices. The relevance of the simulation results to the dynamics of the atmosphere is discussed as is possible continuations of the investigation.
% \end{abstract}
%
% \pacs{}% insert suggested PACS numbers in braces on next line
%
% \maketitle %\maketitle must follow title, authors, abstract and \pacs

% Body of paper goes here. Use proper sectioning commands. 
% References should be done using the \cite, \ref, and \label commands
\section{Introduction}

Large scale geophysical flows are often mentioned as possible applications in the context of two-dimensional (2D) turbulence whose fundamental features were first predicted in the seminal paper  by \citet{Kraichnan1967}, which this Focus Issue is commemorating.
\citet{Gage:1979} and \citet{Lilly:1983} suggested that the $ k^{-5/3} $ energy spectrum at atmospheric mesoscales, that are wave lengths from a few km up to five hundred km, measured by aircraft \citep{Nastrom-Gage:1985, Lindborg:1999}, is produced by an inverse energy cascade of similar type as described by \citet{Kraichnan1967}, a  hypothesis which recently was revived by \citet{Xia2011}. 
In the introduction of their excellent review article on 2D turbulence \citet{Boffetta2012} write that "large-scale motions in the atmosphere and oceans are described, to first approximation, as 2D turbulent fluids". There are almost countless examples of articles on 2D turbulence in which large scale geophysical flows are mentioned as possible applications. 
However, the two-dimensional Navier-Stokes equations is  a very crude model for geophysical turbulence. Fundamental features of geophysical flows such as inertia-gravity waves are absent. Moreover, the inverse energy cascade of two-dimensional turbulence is far from telling the whole story  about energy exchanges between different scales of motion in the atmosphere. In Fig.~\ref{Flux} we have reproduced  a plot from \citet{Augier-Lindborg:2013} showing spectral energy fluxes calculated in the framework of spherical harmonics, $ Y_l^m $, from a General Circulation Model (GCM) called AFES \citep{Hamilton-Takahashi-Ohfuchi:2008}.
Using these base functions the spectral energy flux can be calculated from solutions to the primitive equations \citep{Vallis:book} on a sphere as a function of wave number $ l $, after summation over wave number $ m $, in a corresponding way as the spectral energy flux can be calculated in Fourier space from solutions of Navier-Stokes equations on a plane. 
As shown by  \citet{Augier-Lindborg:2013} the flux can be calculated in a consistent way at each pressure level from a full  GCM. In the figure  we see spectral energy fluxes measured in $ W/m^2 $, integrated over all 24 levels of the model. In contrast to the case of  two-dimensional Navier-Stokes turbulence, the total spectral energy flux, $ \Pi (l) $  (black curve), of the primitive equations is the sum of the fluxes of two different forms of energy, the Kinetic Energy (KE) flux, $ \Pi_K(l) $, (red solid curve) and the flux, $ \Pi_A(l) $, of Available Potential Energy (APE) (blue curve). As shown by \citet{Lorenz:1955}, under certain approximations APE is quadratic in potential temperature and the sum of KE and APE is conserved by the primitive equations. For readers not familiar with Geophyscial Fluid Dynamics, potential temperature is the temperature a fluid element would have in case it would be adiabatically displaced to a reference pressure level. 
In Fig.~\ref{Flux}, the fast increase of $ \Pi_A $  at  $ l \in [1, 3] $ means that  APE is forced at planetary scales of motions. The decrease of $ \Pi_A $ at $ l \in [8, 30] $ is matched by a corresponding increase of $ \Pi_K $. In this wave number range APE is converted into KE by baroclinic instability \citep{Vallis:book} which is the mechanism producing large scale weather systems.  The green curve, $ C_{cum}(l) $,  is the cumulative spectral conversion from APE to KE. A decrease of $ C_{cum}(l) $ in a certain wave number range corresponds to the amount of energy which is converted from APE to KE in that range.

\begin{figure}[h]
\centerline{\includegraphics[angle=0,width=11.5cm]{./fig1.eps}}
 \caption{Spectral energy fluxes as functions of spherical harmonic wave number $ l $, calculated from the AFES GCM. The red solid curve is the KE flux, the blue curve is the APE flux, the black curve is the sum of KE and APE fluxes and the green curve is the accumulated APE-KE conversion. The dotted red curve is the 2D KE flux, calculated only from the rotational part of the velocity field, while the dashed curve is the difference between the total and the 2D KE flux. 
Reprinted
with permission from P. Augier \& E. Lindborg, 
"A new formulation of the spectral energy budget of the atmosphere, with application to two high-resolution general circulation models", 
J. Atmos. Sci. \textbf{70}, 2293--2308 (2013).
Copyright 2013 American Meterological Society.
}
 \label{Flux}
 \end{figure}

Some of the energy which is converted from APE to KE goes into an inverse energy cascade, as is clear from the fact that $ \Pi_K < 0 $ for $ l < 10 $. This inverse cascade is the energy source of the large scale circulation of the atmosphere. According to the model, approximately $ 0.5 \; W/m^2 $ is going into the inverse energy cascade. This cascade is produced by nonlinear interactions of the same type as is giving rise to the inverse energy cascade of the two-dimensional incompressible Navier-Stokes equations. To explain this further we make a Helmholtz decomposition of the horizontal velocity field,
\begin{equation} 
{\bf u} = \bf{u}_r + \bf{u}_d \, ,
\end{equation} 
where
\begin{equation}  {\bf u}_r  = -\nabla \times ( {\bf e_z} \Psi) \, , \end{equation}  
is the rotational component and
\begin{equation}  \bf {u}_d = \nabla \chi  \, , \end{equation}
is the divergent component of the velocity field,  $ \Psi  $ is the stream function,  $ \chi $ is the velocity potential and $ {\bf e}_z $ is the vertical unit vector.  In this paper we will not go into the problem of uniqueness of the Helmholtz decomposition, but assume that it can be carried out in an unambiguous way, as is the case  with a flow field on a sphere or on a square domain if we apply periodic boundary conditions.  The KE flux in Fig.~\ref{Flux} is decomposed into two parts,  the 2D-part (dotted red curve) that contains contributions from nonlinear interactions only involving $ {\bf u}_r $ and the remaining part (red dashed curve) that contains contributions from interactions also involving $ {\bf u}_d $. As seen from the figure, it is the 2D-part that produces the inverse energy cascade, while the other part is actually producing a forward energy cascade of approximately the same magnitude. 
In the model, this forward energy cascade is giving rise to a KE spectrum with an approximate  $ -5/3 $ power law in the spherical harmonic wave number $ l $.    The spectral flux of APE is also positive in the same range and the APE spectrum also shows an approximate  $-5/3 $ power law dependence. When velocity and temperature are decomposed into Fourier modes with wave number $ k $, both KE- and APE-spectra exhibit an extended $ k^{-5/3} $ range at atmospheric mesoscales as  demonstrated by \citet{Nastrom-Gage:1985}.  As mentioned previously, \citet{Gage:1979} and \citet{Lilly:1983}    suggested that the $ k^{-5/3} $ mesoscale range is produced by an inverse or upscale energy cascade of similar type as in 2D turbulence. However, there is an emerging consensus that the mesoscale spectrum is associated with a downscale energy cascade of Kolmogorov type, as first suggested by \citet{Dewan:1979}. There is still a lively debate whether the three-dimensional motions behind this cascade should be characterised as weakly nonlinear waves \citep{Callies-Ferrari-Buhler:2014, Waite-Snyder:2013}, or as turbulence \citep{Lindborg:2015}. In this paper we will not go into this debate but take a completely different approach.

Theories that are trying to explain the mesoscale $ k^{-5/3} $ spectrum predict that the three-dimensional structures behind the forward energy cascade are very thin and elongated. \citet{Lindborg:2006} and 
\citet{Waite-Bartello:2004}  emphasise that  it is necessary to have a vertical resolution which is finer than the buoyancy scale, $ \l_b= u/N $, in order to simulate the downscale energy cascade. Here, $ u $ is a characteristic horizontal velocity and $ N $ is the Brunt-V\"sisal\"a frequency. In the upper troposphere and the lower stratosphere $ l_b $ can be estimated to be of the order of one km at most.  According to the hypothesis  of \citet{Callies-Buhler-Ferrari:2016} the mesoscale energy spectrum is built up by inertia-gravity waves with frequencies very close to the inertial frequency $ f $. The ratio between the vertical and the horizontal wave length of such waves is extremely small. With a horizontal wave length $ \lambda_{h} \sim 100 $ km, the corresponding vertical wave length would be a few hundred meters. Despite the prediction of these theories that a very fine resolution would be needed in order to simulate the downscale mesoscale energy cascade it is a fact that GCMs such as the AFES-model and the MPAS-model \citep{Skamarock-Park-Klemp-Snyder:2014} are able to do this although they lack the required vertical resolution. This fact has lead us to the following questions: What is the minimum number of levels that are needed in order to simulate the mesoscale downscale energy cascade? Is it even possible to do this with a single level model, in other words with a two-dimensional model? It is the aim of the present paper to construct such a model.

\section{Development of the toy model}
It is clear from the previous discussion that  a  two-dimensional model that is able to produce both an inverse and a forward energy  cascade must allow for a divergent component of the two-dimensional velocity field. 
To mimic geophysical turbulence the model should also contain a scalar playing the role of potential temperature. Preferably, total energy should be the sum of KE and APE, KE should be quadratic in the velocity, APE should be quadratic in the scalar and the equations for KE and APE should contain an exchange term allowing for a velocity field to develop if the model is forced in the scalar field.

We first consider the shallow water model \citep{Vallis:book} and discuss why this  is not a fully satisfactory model for Geophysical Turbulence, despite the fact that it has several attractive features. The shallow water set of equations can be written as
\begin{eqnarray} \label{Velocity}
\frac{\partial {\bf u}} {\partial t} + {\bf u} \cdot \nabla {\bf u} +  f {\bf e}_z \times {\bf u} = -c^2 \nabla \eta \, , \\ \label {Eta}
\frac{\partial \eta}{\partial t}
+ {\bf u} \cdot \nabla \eta   = - (1+\eta) \nabla \cdot {\bf u} \, ,
\end{eqnarray}
where $ \eta \ll 1 $ is the non-dimensional surface displacement of a shallow water layer with unity depth, $ c = \sqrt{gD} $ is the phase velocity of surface gravity waves, $ g $ being the acceleration due to gravity and $ D $ the dimensional depth of the layer, and $ f  $ is the Coriolis parameter. In this paper we will not consider the case with differential rotation, that is a $ \beta $-plane on which $ f = f_{0} + \beta y $, where  $ y $ is the coordinate in the meridional direction, but restrict the study to an $ f $-plane on which $ f $ is constant. 
The shallow water model conserves the sum of kinetic energy and enthalpy, defined as 
\begin{eqnarray}
E_K = (1+\eta) \frac{{\bf u} \cdot {\bf u}}{2} \, , \\  \label{potential}
H = c^{2} \left (\alpha + \beta \eta  + \frac{(1+\eta)^2 }{2} \right) \, ,
\end{eqnarray} 
where $ \alpha $ and $ \beta $ are two arbitrary constants. 
We define APE by putting $  \alpha = -1/2 $ and $ \beta = - 1 $  in the expression for $ H $, that is
 \begin{equation}
E_{A} = \frac{c^2  \eta^{2} }{2} \, .
\end{equation}
The equations for $ E_K $ and $ E_A $ are
\begin{eqnarray} \label{KE}
\frac{\partial{E_K}} {\partial t} + \nabla \cdot ({\bf u}E_K) = -{\bf u} \cdot \nabla \left (\frac{c^2(1+\eta)^2}{2} \right ) \, , \\
\frac{\partial{E_A}} {\partial t} + \nabla \cdot({\bf u} (2E_A + \eta c^{2})) = {\bf u} \cdot \nabla \left (\frac{c^2(1+\eta)^2}{2} \right ) \, .
\end{eqnarray}
The shallow water model also has another interesting property. Potential vorticity, defined as 
\begin{equation} \label{PV}
Q = \frac{f +\zeta }{1+\eta} \, ,
\end{equation}
where $ \zeta =  \nabla^{2} \Psi $ is the vorticity, is a materially conserved quantity. 
This property is extensively explored in text books on Geophysical Fluid Dynamics (e.g., \citet{Vallis:book}) to illustrate interesting features of geophysical flows. { However, as discussed by \citet{Warn1986}  conservation of $ Q $ is not a property that is very useful from the point of view of turbulence theory, since potential enstrophy, defined as the square of $ Q $,  is not a quadratic quantity in the flow quantities.  As shown by \citet{Warn1986} it is only in the limit when $ Q $ becomes a linear quantity that 
 conservation of $ Q $ will not lead to any constraints on cascade directions. }  In the limit of strong rotation, or the Quasi Geostrophic (QG) limit \citep{Charney1971}, potential enstrophy becomes a quadratic quantity and so does total energy.  In this limit,  shallow water turbulence is very similar to 2D turbulence, since there are two related quadratic invariants and $ {\bf u}_d $ is zero. This limit will be discussed below.

At first glance the shallow water model seems to be a promising candidate for a two-dimensional model of geophysical turbulence that might be capable of reproducing the same type of energy exchanges and energy cascades as displayed in Fig.~\ref{Flux}.
The velocity field has both a rotational and a divergent component. Moreover, the model contains a scalar playing a similar role as potential temperature, APE is quadratic in the scalar, 
the sum of KE and APE is conserved and there is a term associated with exchange between the two forms of energy.  From the point of view of turbulence theory it is a drawback that kinetic energy is not a quadratic quantity, but since it must be assumed that $ \eta \ll 1 $, kinetic energy is approximately quadratic.

The shallow water model has been studied from the point of view of turbulence theory by several investigators \citep{FargeSadourny1989, YuanHamilton1994, Dritschel2001}.
As shown in these studies shallow water turbulence can be decomposed into a vortical field and a wave field. The vortical field shares some properties with two-dimensional Navier-Stokes incompressible turbulence while the  wave field undergoes a forward energy cascade. However, strong discontinuities or shocks develop in the wave field, as observed by \citet{SpallMcWilliams1992}. 
In a companion paper \citep{AugierMohananLindborg2017} we show that the forward energy cascade of the shallow water wave field is dominated by such shocks, 
giving rise to a $ k^{-2} $-spectrum, which is  characteristic of shock dominated turbulence \citep{Kuznetsov2004}, such as turbulence arising from Burger's equation  \citep{BouchaudMezardParisi1995}.  It is obvious that this type of turbulence has little to do with mesoscale dynamics in the atmosphere where no shocks are observed. \citet{Buhler1998} has shown that the shallow water model can be modified in such a way that shocks are avoided. This is accomplished by replacing the term  $ -c^{2} \nabla \eta $ on the right hand side of equation (\ref{Velocity}) by  $ -c^{2}(1+\eta)^{-3} \nabla \eta $. However, the  expression for enthalpy will in this case take the form $ H = c^2( \alpha \eta + \beta \eta+ (1+\eta)^{-1}/2) $, and there is no way to define APE such that it becomes quadratic in $ \eta $. 
We will instead introduce two other modifications of the shallow water model that will remove the shocks and also give an expression for total energy which is quadratic. The first modification we introduce is to replace the term
$ -(1+\eta) \nabla \cdot  {\bf u} $ on the right hand side of (\ref{Eta}) by $ -\nabla \cdot {\bf u} $. This modification can be motivated from the assumption $ \eta \ll 1 $, which is already inherent in the shallow water model. 
{ Equation (\ref{Eta}) of the shallow water model is derived from mass conservation. If this interpretation also is made of the modified equation,  mass conservation is not perfectly fulfilled, but only to the extent that $ \eta $ is small.}  The second modification is motivated by introducing the assumption that the large scale motions are dominated by the rotational component of the velocity field.  We define the fields in square domain $ 
[0 \; 2\pi]^2 $, apply periodic boundary conditions and denote the 2D Fourier transform of a quantity by putting a hat on it. We decompose the rotational and divergent velocity fields into large and small scale fields as
\begin{eqnarray}
{\bf u}_r = {\bf u}_r^{l} + {\bf u}_r^{s} \, , \;\;\;\; 
{\bf u}_r^{l} = \sum_{k < k_{d}} -{\mbox{i}} {\bf k} \times {\bf e}_z \widehat{\psi}  \exp({\mbox{i}}{\bf k}\cdot {\bf r}) \, , \\
{\bf u}_d = {\bf u}_d^{l} + {\bf u}_d^{s} \, , \;\;\;\; 
{\bf u}_d^{l} = \sum_{k < k_{d}} {\mbox{i}} {\bf k} \widehat{\chi}  \exp({\mbox{i}}{\bf k}\cdot {\bf r}) \, ,
\end{eqnarray} 
where $ k = |{\bf k} | $ and $ k_d = f/c $ is the wave number corresponding to the deformation radius $ r_d = c/f $.  We also decompose the vorticity, $ \zeta = \nabla^{2} \psi $, and the divergence , $ d = \nabla^2 \chi $, in the same way, 
\begin{eqnarray}
\zeta = \zeta^l + \zeta^s \, , \;\;\;\;  \zeta ^{l} = \sum_{k < k_{d}} -k^2 \widehat{\Psi} \exp({\mbox{i}}{\bf k}\cdot {\bf r}) \, , \\
d = d^l + d^s \, , \;\;\;\;  d ^{l} = \sum_{k < k_{d}} -k^2 \widehat{\chi} \exp({\mbox{i}}{\bf k}\cdot {\bf r}) \, , 
\end{eqnarray} 
and make the following assumptions  
\begin{eqnarray} \label{LargeAssumption}
 | {\bf u}_r^{l} | \gg  | {\bf u}_r ^{s} | \, , \;\; | {\bf u}_d^{l} | \gg  | {\bf u}_d ^{s} | \, , \;\; 
 | {\bf u}_r^{l} | \gg | {\bf u}_d^{l} | \, , \\  \label{SmallAssumption} 
 | \zeta^{l} | \ll | \zeta^{s} | \, , \;\;   | d^{l} | \ll | d^{s} | \, , \;\; 
 | \zeta^{s} | \sim | d^{s} | \, .
\end{eqnarray}
Indeed, these assumptions are consistent with atmospheric dynamics, where the velocity field is dominated by large scale rotational motions, while the vorticity and divergence fields are dominated by contributions from the high wave number end of the $ k^{-5/3} $-spectrum range and are of the same order of magnitude 
\citep{Lindborg:2015}. 
Under the assumption (\ref{LargeAssumption})   we can approximate the advective operator $ {\bf u} \cdot \nabla $ by 
$ {\bf u}_r \cdot \nabla $. This is the same type of approximation that is made in the QG approximation \citep{Charney1971}.
However, there is an important difference. The assumption (\ref{SmallAssumption}) does not permit us to neglect the derivatives of $ {\bf u}_d $ that arise when the advective operator is acting on the velocity field.  
Introducing these modifications in the shallow water equations the toy model equations can be written as 
\begin{eqnarray} \label{V2}
\frac{\partial {\bf u}} {\partial t} + {\bf u}_r \cdot \nabla {\bf u} + f {\bf e}_z \times {\bf u} = -c \nabla \theta + (-1)^{p+1} \nu_p \nabla^{2p} {\bf u} , \\ \label{Theta}
\frac{\partial \theta}{\partial t}
+ {\bf u}_r \cdot \nabla \theta   = -  c\nabla \cdot {\bf u} + (-1)^{p+1} \nu_p \nabla^{2p} {\theta}\, ,
\end{eqnarray}
where we have introduced the notation $ \theta = c \eta $ for the scalar and also added diffusion terms.  
Although it is very difficult to prove, it is quite easily argued that no shocks will arise from equations (\ref{V2}) and (\ref{Theta}), even in the absence of diffusion terms.  Locally, shock formation can be viewed as a one-dimensional process in which a gradient along a specific direction, say the $ x $-direction,  is amplified through nonlinear interactions. In the case  when $ {\bf u} = u {\bf e}_x + v {\bf e}_y $  and $ \theta $ are only functions of $ x $, we find that $ {\bf u}_r = v {\bf e}_y $ and the nonlinear terms in (\ref{V2}) and (\ref{Theta}) are zero. Therefore, no shocks will form in the one-dimensional case and it is therefore arguable that no shocks will be formed in the general two-dimensional case.

The toy model equations can be expressed as three coupled equations for $ \Psi $, $ \chi $ and $ \theta $. Unfortunately, we have found no elegant way in which they can be expressed in vector notation, why we write them as
\begin{eqnarray} \label{PsiEq}
\frac{\partial{\Psi_{ii}}}{\partial t} +\epsilon_{ijk}e_k \Psi_i \Psi_{jll} + \Psi_{ii} \chi_{jj} - \Psi_{ij} \chi_{ij} = -f \chi_{ii} + (-1)^{p+1} \nu_p \nabla^{2p} \Psi_{ii} \, , \\ \label{ChiEq}
\frac{\partial{\chi_{ii}}}{\partial t} +\epsilon_{ijk}e_k (\Psi_i \chi_{jll} - \Psi_{jl} \chi_{il}) + \Psi_{ij} \Psi_{ij} - \Psi_{ii} \Psi_{jj} =  f \Psi_{ii} - c \theta_{ii} + (-1)^{p+1} \nu_p \nabla^{2p} \chi_{ii} \, , \\ \label{ThetaEq} 
\frac{\partial{\theta}}{\partial t} + \epsilon_{ijk}e_k \Psi_i \theta_{j} = - c \chi_{ii} + (-1)^{p+1} \nu_p \nabla^{2p} \theta \, ,
\end{eqnarray} 
where $ e_k $ is the vertical unit vector expressed in three-dimensional Cartesian tensor notation,  subscript $ i $  has the same meaning as $ {\partial }/{\partial x_i }$ when used on $ \Psi $,  $ \chi $ or $ \theta $ and repeated indices means summation.

The toy model expression for  KE and APE are quadratic,   $ E_K = {\bf u}\cdot {\bf u}/2 $ and $ E_A = \theta^2/2 $. 
In the absence of diffusion terms the equations for $ E_K $ and $ E_A $ can be written as
\begin{eqnarray} \label{EK}
\frac{\partial{E_K}} {\partial t} + \nabla \cdot ({\bf u}_rE_K+ c{\bf u}_r \theta) = - c \nabla \theta \cdot {\bf u}_d  \, , \\ \label{EA}
\frac{\partial{E_A}} {\partial t} + \nabla \cdot ({\bf u}_rE_A + c{\bf u}_d  \theta) =  c \nabla \theta \cdot {\bf u}_d \, .
\end{eqnarray}
 The term on the right hand side of  (\ref{EK}) is the same as the term on the right hand side of (\ref{EA}) apart from the sign. In other words, this term is conversion from APE to KE. 
The potential vorticity, $ Q $,  as defined in (\ref{PV}), is not conserved by the toy model. This is the price that we have to pay in order to obtain a model that does not produce any shocks and has a quadratic expression for total energy. As we discussed above, since $ Q $ is not a linear quantity in the flow quantities conservation of $ Q $  is not a very interesting property of the shallow water model if we look at it strictly from the point of view of turbulence theory. 
We are therefore willing to pay this price. Instead of studying $ Q $ we study  linear potential vorticity  \citep{Dritschel2001} defined as 
\begin{equation} \label{Linear}
q = \zeta - \frac{f}{c}\theta .
\end{equation}
In the QG limit $ q $ is a materially conserved quantity in the shallow water model \citep{Vallis:book}. This is also the case in the toy model. 
The equation for $ q $ is derived as a linear combination of  (\ref{PsiEq}) and (\ref{ThetaEq}),
\begin{equation}
\frac{\partial q }{\partial t} + {\bf u}_r \cdot \nabla q  =  \Psi_{ij} \chi_{ij} - \Psi_{ii} \chi_{jj} \, ,
\end{equation}  
where we have excluded diffusion terms. A small parameter expansion of  the equations (\ref{PsiEq} - \ref{ThetaEq}) in the Rossby number \citep{Vallis:book} $ Ro $ gives that $ \chi \sim Ro \Psi $ and in the limit $ Ro \rightarrow 0 $, $ q $ will be a materially conserved quantity.

We define the KE and APE spectra as 
\begin{eqnarray} 
E_K(k) =\sum_{B(k)}  \frac{1}{2}  \widehat{\bf u} \cdot \widehat{\bf u}^{\star}  \, , \\
E_A(k) = \sum_{B(k)} \frac{1}{2} \widehat{\theta} \widehat{\theta}^{\star} \, ,
\end{eqnarray} 
where $ ^{\star} $ denotes the complex conjugate and the summation is over 
all $ k_x $ and $ k_y $ such that $ \sqrt{k_x^2+k_y^{2}} $ lies within the wavenumber band,  $ B(k) $,   centered at $ k $ and with unity width. The equation for the energy spectra can be written as
\begin{eqnarray}
\frac{\partial{E_K}(k)} {\partial t}  = T_K(k) + C (k) - \nu_p k^{2p} E_K(k) \, , \\
\frac{\partial{E_A}(k)} {\partial t}  = T_A(k) - C (k) - \nu_p k^{2p} E_A(k) \, ,
\end{eqnarray} 
where 
\begin{eqnarray}
T_{K}(k) = \sum_{B(k)} Im [  k_j \widehat{u_{rj} u_i} \widehat{u}_i ^{\star} ]  \, , \\
T_{A}(k)  = \sum_{B(k)} Im [ k_j \widehat{u_{rj} \theta} \widehat{\theta} ^{\star}] \, , \\
C (k) = \sum_{B(k)} Re[ c k^{2}\widehat{\theta} \widehat{\chi}^{\star}] \, .
\end{eqnarray} 
are spectral transfer and conversion terms. The spectral flux of kinetic and potential energy are defined as
\begin{eqnarray}
\Pi_K(k) = \sum_{k^{\prime} = k}^{k_{m}} T_K(k^{\prime}) \, , \\
\Pi_A(k) = \sum_{k^{\prime} = k}^{k_{m}} T_A(k^{\prime}) \, ,
\end{eqnarray} 
where $ k_{m} $ is the maximum wave number in the Fourier decomposition. The cumulative spectral conversion is defined as
\begin{equation} \label{Cum}
C_{cum} (k)  = \sum_{k^{\prime} = k}^{k_{m}} C (k^{\prime}) \, .
\end{equation} 
Unlike the energy fluxes, $ C_{cum} $ does not go to zero in the limit $ k =0 $, but approaches the total mean conversion of APE to KE. 
Potential enstrophy,  defined as $ q^2/2 $,  will be an inviscidly conserved quantity in the QG limit. Therefore, it may be interesting to study spectral transfer and spectral flux of potential enstrophy. We define these quantities as
\begin{eqnarray}
T_{q} =  \sum_{B(k)} Im [ k_j \widehat{u_{rj} q} \widehat{q}^{\star}]  \, , \\
\Pi_{q} = \sum_{k^{\prime} = k}^{k_{m}} T_{q} (k^{\prime}) \, .
\end{eqnarray} 
 Following \citet{Bartello:1995} and \citet{Dritschel2001} we decompose the flow field into normal  modes which we define as the three normalised eigenmodes of the linear inviscid part of the system (\ref{PsiEq}-\ref{ThetaEq}) in Fourier space. These modes are
\begin{eqnarray}
\widehat{a}_{1} = \frac{k}{\sqrt{2} \omega} ( c \widehat{\theta} + f \widehat{\Psi} - {\mbox{i}} \omega \widehat{\chi} )  \\
\widehat{a}_{2} =  \frac{k}{\sqrt{2} \omega} ( c \widehat{\theta} - f \widehat{\Psi} + {\mbox{i}} \omega \widehat{\chi} ) \\
\widehat{b} =  \frac{1}{\omega} (c k^2 \widehat{\Psi}-f \widehat{\theta}   ) \\
\end{eqnarray} 
where
\begin{equation} \label{Dispersion}
\omega =  \sqrt{f^2+k^2c^2}  \, .
\end{equation} 
The modes $ \widehat{a}_{1} $ and $ \widehat{a}_{2} $ are associated with inertia-gravity waves with frequencies $ \pm \omega $. 
The mode $ \widehat{b} $, on the other hand,  is associated with linear potential vorticity (\ref{Linear}). 
We define the wave and the vortical energy spectra as
\begin{eqnarray}
E_{W}(k) = \sum_{B(k)} \frac{1}{2}(\widehat{a}_{1} \widehat{a}_{1}^{\star} + \widehat{a}_{2} \widehat{a}_{2}^{\star} ) \, , \\
E_{V}(k) =  \sum_{B(k)} \frac{1}{2} \widehat{b} \widehat{b} ^{\star} \, .
\end{eqnarray} 
{ \noindent Following \citet{Deusebio-Augier-Lindborg:2013} the total energy transfer can be decomposed into four terms as
\begin{equation}
T_K(k) + T_A(k) = T_{VVV}(k) + T_{VVW}(k) + T_{VWW} + T_{WWW} \, ,
\end{equation} 
where 
$ T_{VVV} $ represents the contribution from triad interactions between three vortical modes, $ T_{VVW} $  two vortical modes and one wave mode,  $ T_{VWW} $ one vortical mode and two wave modes and $ T_{WWW} $ three wave modes.  It can be shown that each of these four  terms integrates to zero and  thus conserves energy. Accordingly, the total energy flux may  also be decomposed into four terms as
\begin{equation} \label{Decomposition}
\Pi_K(k) + \Pi_A(k) = \Pi_{VVV}(k)+ \Pi_{VVW}(k)+\Pi_{VWW}(k)+ \Pi_{WWW}(k) \, .
\end{equation} }
\noindent We are now in the possession of the tools that are needed in order to explore a numerical simulation of the toy model.

\section{Numerical simulations} 
In this section we will present { three}  simulations of the toy model showing that it does not produce any shocks, unlike the shallow water model, and that it is capable of reproducing energy exchanges and cascades that are similar to what is displayed in Fig.~\ref{Flux}.  A more comprehensive numerical study will be presented in a future paper.

The equations (\ref{PsiEq}-\ref{ThetaEq}) are solved in a square domain  using a standard pseudo spectral method and a fourth order Runge-Kutta scheme, in which the time step is set by a CFL-condition guaranteeing both numerical stability and that fast gravity waves are resolved.   A random forcing is  injecting energy into the system at a rate $ P $, which is approximately equal to unity. 
The forcing is white noice in time and restricted to a narrow range in Fourier space centered at wave number $ k_f $. Initially, the wave field is set to zero while the vortical field is initialised with a very low amplitude random noise at the forcing wave number to facilitate for the vortical field to build up.
The parameters $ f $ and $ c $ are specified in terms of the Rossby and the Froude numbers defined as
\begin{eqnarray}
Ro = \frac{P^{1/3} k_f^{2/3}}{f} \, , \\
Fr = \frac{P^{1/3}}{k_f^{1/3} c}  \, .
\end{eqnarray} 
We use resolution $ n = 1920 $ in both directions and dealiasing with truncation in Fourier modes  lying outside the square $ | k_y | <  k_{max} $,  $ | k_x | < k_{max} $,  where $ k_{max} = 4n/9 $.  
We chose  $ p = 4 $ as the order of the diffusion terms in (\ref{PsiEq}-\ref{ThetaEq}).  { Assuming that the simulations will reach a quasi stationary state in which $ \epsilon \approx P $, where $ \epsilon $ is the dissipation, $ \nu_4 $  is set so that $ k_{diss}/k_{max} \approx 1$, where $ k_{diss} = \epsilon^{1/3}/\nu_4^{3/22} $. With this choice of $ \nu_4 $ it was found that the peak of the dissipation spectrum is resolved. We present three simulations. In run 1 the forcing is injecting energy at a small wave number $ k_f = 6 $ directly into the wave field. In run 2, on the other hand, the forcing is primarily injecting energy into the vortical field at the same wave number. In run 3 we force the wave field at a medium wave number, $ k_f = 30 $.
In all three simulations we use $ Ro= Fr = 0.055 $. } That 
the Rossby number is equal to the Froude number means that we force at the deformation scale,  in other words that  $ k_f  = f/c $.  { All  three simulations are run for quite a long time in terms of inertial periods.  
For the three runs, $ t_{end}/\tau = 720, \; 480, \;  1224 $, respectively,  where $ \tau = 2 \pi /f  $. Run 1 and run 2 both reach an almost perfectly stationary state, while run 3 reaches a quasi stationary state with a slow build up of vortical energy. Spectra and fluxes are calculated as time averages over a time interval during the end of the simulations. }

{ We start by presenting the fluxes and the spectra from run 1. }Essentially all energy which is injected at $ k_f  =6 $, goes into a forward energy cascade, which can be seen in the plot of the energy fluxes in Fig.~\ref{F1}. The fluxes are normalised by the total energy dissipation, $ \epsilon $, calculated as the time averaged mean value in the same interval.  The total spectral normalised energy flux, $ \Pi(k) $, (black curve), is positive and constant throughout a broad range starting at the forcing wave number,  which is a signature of a forward energy cascade. The normalised flux is approximately equal to unity, which means that the energy which is going into the forward energy cascade is subsequently dissipated at high wave numbers, as expected. There is no doubt that the forward energy cascade is a cascade of wave motions. Although all energy is injected into the scalar field, the fluxes of KE and APE, $ \Pi_K $ (red line) and $ \Pi_A $ (blue line), attain equal values that are almost equal at wave numbers that are slightly larger than $ k_f $ and almost perfectly equal at even larger wave numbers.  In other words, the fluxes are equipartioned between KE and APE. This is characteristic of a gravity wave field with equipartion between KE and APE. 
Since we force at the deformation wave number, the waves are only slightly affected by rotation at wave numbers that are considerably larger than $ k_f $, which can be seen from the dispersion relation (\ref{Dispersion}).

\begin{figure}[h]
\centerline{\includegraphics[angle=0,width=10.5cm]{./fig2.eps}}
 \caption{Spectral energy fluxes from run 1, normalised by the time averaged total energy dissipation $ \epsilon $.  }
 \label{F1}
 \end{figure}

Looking closer to Fig.~\ref{F1} we see that the kinetic energy flux, $ \Pi_K $, attains negative values at wave numbers around $ k_f $, which is a sign of an inverse energy cascade, although this is obviously very weak. A decomposition of  $ \Pi_K $ in the 2D part, containing contributions from interactions only involving $ {\bf u}_r $ and a remaining part, reveals that the 2D part (lower dashed red curve) is negative and thus produces an inverse energy cascade  while the remaining part (upper dashed red curve) is positive. The two contributions almost cancel each other at wave numbers close to $ k_f $. This is similar as seen in Fig.~\ref{Flux} from the GCM, although the inverse energy cascade is very weak in run 1.

\begin{figure}[h]
\centerline{\includegraphics[angle=0,width=8.5cm]{./fig3.eps}}
 \caption{ Energy spectra from run 1. Upper plot is showing KE and APE spectra, while the lower plot is showing wave and vortical energy spectra. }
 \label{S1}
 \end{figure}

 In Fig.~\ref{S1} we see  time averaged energy spectra from run 1. The upper plot is
 showing the KE spectrum, $ E_K $,  and the APE spectrum, $ E_A$, while the lower plot
 is showing wave energy and the vortical energy spectra, $ E_W $ and $ E_V$.  In the
 forward cascade range, $ E_K $ and $ E_A $ are slightly more shallow than $ k^{-5/3} $,
 as is $ E_W $.  As can be seen from the upper plot, KE and APE are equipartioned at
 large wave numbers. Again, this is a characteristic feature of gravity waves. In the
 lower plot we see that $ E_V $  falls of approximately as $ k^{-3} $. Thus, at high
 wave numbers the wave field is totally dominant. At very low wave numbers, however, $
 E_V $ is slightly larger than $ E_W $, due to a slow build  up of vortical energy at
 large scales. { As we will argue in the following the approximate $ k^{-3} $-range of
 the vortical spectrum may be interpreted as an enstrophy or potential enstrophy cascade
 range of the same type as is found in 2D turbulence \citep{Kraichnan1967, Kraichnan1971}.}

{ It is not entirely clear why the wave energy spectrum is more shallow than $ k^{-5/3}
$ in run 1. One hypothesis would be that this can be explained by the so called
"bottleneck effect". Energy has a tendency to build up in the high wave number end of
the spectrum and this tendency becomes stronger with a higher degree of the diffusion
operator \citep{Haugen2004}. We have explored this hypothesis by carrying out a series
of simulations in which the only difference is the order of the diffusion operator. In
Fig.~\ref{Comp1} we see a plot of the compensated wave energy spectrum from three such
simulations,  using $ \Delta^{4} $ (run 1), another using $ \Delta^{2} $ and a third
using the Navier-Stokes diffusion operator $ \Delta $. Indeed, $ E_W $ drops a little
bit when we use $ \Delta^2 $ instead of $ \Delta^{4} $ and when we use $ \Delta $ we
obtain a pretty clean $ k^{-5/3} $-range, although it is limited. However, in this case
we also observed a clear decrease of the energy flux in the same range, which makes it
questionable if the deviation from $ k^{-5/3} $ entirely can be explained as a
bottleneck effect.  Another hypothesis is that  deviations from $ k^{-5/3} $ can be
generated as a consequence of the type of nonlinear interactions which are involved,
which in turn may be dependent on the type of forcing which is used. To explore this
hypothesis we carried out run 2, which is similar to run 1 in everything except in the
forcing scheme. While run  1 is primarily forced in  wave modes, run 2 is primarily
forced in vortical modes. In Fig.~\ref{Flux2} we see the energy fluxes from run 1 and
run 2 decomposed into the four contributions given in equation (\ref{Decomposition}). In
both runs $ \Pi_{VWW} $ is dominant. This flux is produced by triad interactions between
one vortical mode and two wave modes. However, in run 1 $  \Pi_{WWW} $ also makes an
important contribution, especially at high wave numbers, while $ \Pi_{VVW} $ is close to
zero. In run 2,  on the other hand,  $ \Pi_{VVW} $ makes an important contribution while
$ \Pi_{WWW} $ is close to zero. In both runs, $ \Pi_{VVV} $ is negative, which means
that triad interactions including three vortical modes produce an inverse cascade, as
expected.  In the left plot of Fig.~\ref{C2} we  see compensated wave energy spectra from
run 1 and run 2. Comparing the two runs we see that there is a range where the
compensated spectrum is flat in run 2, which means that we have a pretty clean $
k^{-5/3} $ range, while the compensated spectrum of run 1 is exhibiting a ``bump'' in the
high wave number range.  In all likelihood the difference is associated with the
difference in the type of triad interactions that is producing the cascade. In run 2,
VWW-interactions are totally dominant at high wave number, while WWW-interactions also
make important contributions in run 1. It is likely that the ``bump'' is associated with
these WWW-interactions. In the right plot of Fig.~\ref{C2} we see the uncompensated
vortical energy spectra from run 1 and run 2. Here, there is an even larger difference.
The vortical spectrum from run 2 is much more shallow, close to $ k^{-5/3} $,  than the
$ k^{-3} $ vortical spectrum from run 2. The difference between the two spectra is
associated with a difference in potential energy flux. In Fig.~\ref{EF1} we see the
potential energy flux, $ \Pi_q $,  from run 1 (left) and run 2 (right). In run 1,  there
is a range  which is reminiscent of a constant flux range, although there is slow
increase of  $ \Pi_q $. In run 2, on the other hand, there is a rapid increase of $
\Pi_q $ at high wave numbers where it reaches value which is eighty times larger than in
run 1. These results suggest that the $ k^{-3} $-spectrum of run 1 can be interpreted as
an enstrophy or potential enstrophy inertial range spectrum, although $ \Pi_q $ is not
perfectly constant. Whether or not such a range is obtained should mainly depend on $ Ro
$ and $ Fr $. In the quasigeostrophic limit, $ Ro \rightarrow 0 \, , Fr \rightarrow 0 $,
such a range should always be obtained, since potential enstrophy is an inviscidly
conserved quantity in this limit. At finite $ Ro $ and $ Fr $, different initial
conditions or different forcing mechanisms, may evidently also be important as to
whether such a range is obtained.}

\begin{figure}[h]
\centerline{\includegraphics[angle=0,width=10cm]{./fig4.eps}}
 \caption{Wave energy spectra from simulations with different degrees of the diffusion operator.  }
 \label{Comp1}
 \end{figure}

 \begin{figure}[h]
\centerline{\includegraphics[angle=0,width=17cm]{./fig5.eps}}
 \caption{Energy fluxes from run 1 (left) and run 2 (right) decomposed into different contributions. }
 \label{Flux2}
 \end{figure}

  \begin{figure}[h]
\centerline{\includegraphics[angle=0,width=17 cm]{./fig6.eps}}
 \caption{Compensated wave energy spectra (left) and uncompensated vortical spectra (right) from run 1 and run 2. }
 \label{C2}
 \end{figure}

\begin{figure}[h]
\centerline{\includegraphics[angle=0,width=17cm]{./fig7.eps}}
 \caption{Potential enstrophy flux from run 1 (left) and run 2 (right). }
 \label{EF1}
  \end{figure}

\begin{figure}[h]
\centerline{\includegraphics[angle=0,width=18.5cm]{./fig8.pdf}}
\caption{Divergence field from a simulation of the shallow water equations (left) and from run 1 of the toy model (right). }
\label{Shocks}
 \end{figure}

\begin{figure}[h]
\centerline{\includegraphics[angle=0,width=18.5cm]{./fig9.pdf}}
\caption{Potential vorticity field (left) and wave field (right) from the end state of run 1.}
\label{Vis1}
\end{figure}

 We now turn to the question regarding possible shock generation in the toy model.  We have found that the most reliable method of detecting shocks visually is to plot the divergence field. Shocks are displayed as thin lines of high amplitude negative divergence in such plots. The higher the resolution is, the thinner are the lines, since the width of the shocks scales with the smallest length scale that is resolved if there is sufficient viscosity to diffuse the shocks over this scale.
That the divergence is negative along the shocks  can be understood from the fact that the velocity component perpendicular to the shock has a negative jump in the direction of shock propagation.  Consider a shock which is parallel to the $ y $-axis and is traveling in the positive $ x $-direction. The velocity component in the $ x $-direction, $ u $, is positive on both sides of the shock, but is  smaller behind than in front of the shock. This jump in $ u $ is seen as a high amplitude negative value of $ \partial_x u $, which is the divergence. In Fig.~\ref{Shocks} we have plotted the divergence field from a simulation of the shallow water equations (to the left) together with the divergence field from run 1 (to the right). 
The simulation of the shallow water equations has a similar set up as run 1. 
{ Comparing the two plots one should not compare the absolute values of divergence, which scale differently in the two simulations, but  the pattern. In the left plot, long bands of high amplitude negative divergence are displayed against a background of almost uniform low amplitude divergence. These bands are signatures of shocks. In the right plot, on the other hand, we see tightly packed alternating bands of high amplitude negative and positive divergence. These bands are signatures of a continuous wave field.}
We can quite safely conclude that the toy model does not produce any shocks, just as we argued in the previous section.

In Fig.~\ref{Vis1} we see the potential vorticity field (left plot) and the wave field (right plot) corresponding to one of the wave field modes, namely $ \widehat{a_1} $. There is a huge difference in the visual appearance of the two fields. The vortical field develops structures at much larger length scales and with very different shapes than the elongated fronts of the wave field. A blow up display the wave structures in more detail.

We now turn our attention to run 3, where the only change we did in comparison to run 1, was to force in $ k_f = 30 $ instead of $ k_f = 6 $. The idea is that the inverse energy cascade can develop over a larger span of wave numbers if we use a larger $ k_f $ and may therefore become stronger.  In Fig.~\ref{F2} we see the spectral energy fluxes normalised by the total energy dissipation from run 3.  In this plot we have also included the cumulative spectral APE-KE conversion, $ C_{cum} $,  defined in (\ref{Cum}). In the forward  cascade range the fluxes are similar to what we saw in run 1, with $ \Pi_K = \Pi_A $ at high wave numbers --  a sign of gravity waves with equipartition between $ E_K $ and $ E_A $. At wave numbers $ k < k_f $ the total flux (black curve) is clearly negative which means that a substantial amount of energy is transferred upscale. This upscale cascade 
 is produced by interactions involving only the rotational component  of the velocity field. The lower red dashed curve, $ \Pi_{2D} $, is the negative flux produced by these interactions and the upper dashed red curve is the positive flux produced by the remaining type of interactions, also involving the divergent component.  The cumulative spectral conversion, $ C_{cum} $ (green curve), displays a strong decrease around the forcing wave number. The decrease of $ C_{cum} $ is equal to the APE-KE conversion in the corresponding wave number range and is approximately matched by a corresponding increase of $ \Pi_K $ in the same wave number range. This is similar to what is seen in Fig.~\ref{Flux} showing energy fluxes from a GCM. An interesting feature of $ C_{cum} $ is the strong positive gradient seen at a lower wave number than $ k_f $. That the gradient is positive means that there is conversion from KE to APE at the corresponding wave number. The energy which is converted  from KE to APE at this low wave number starts to cascade downscale which is seen from the positive APE flux at  $ k < k_f $.  At very low wave numbers, $ k \rightarrow 0 $, $ C_{cum} $ approaches the total APE-KE conversion.

\begin{figure}[h]
\centerline{\includegraphics[angle=0,width=10.5cm]{./fig10.eps}}
 \caption{Spectral energy fluxes from run 3, normalised by the time averaged total energy dissipation $ \epsilon $. }
 \label{F2}
 \end{figure}

 \begin{figure}[h]
\centerline{\includegraphics[angle=0,width=8.5cm]{./fig11.eps}}
 \caption{ Time averaged energy spectra from run 3. Upper plot is showing KE and APE spectra, while the lower plot is showing wave and vortical energy spectra. Averaging was made over the intervall $ t/\tau \in [756 \; 1224] $.}
 \label{S2}
 \end{figure}

 \begin{figure}[h]
 \centerline{\includegraphics[angle=0,width=18.5cm]{./fig12.eps}}
\caption{Potential vorticity field (left) and wave field (right) from the end state of run 3.}
 \label{Vis2}
 \end{figure}

\begin{figure}[h]
\centerline{\includegraphics[angle=0,width=18.5cm]{./fig13.eps}}
\caption{Blow up of the potential vorticity field (left) of a strong anticyclonic vortex and the corresponding wave field (right). }
\label{Vis3}
\end{figure}

In Fig.~\ref{S2} we see the time averaged energy spectra averaged over a time
interval during the end of run 3.  In the upper plot we see that $ E_K $ and $ E_A $ {
are equal or equipartioned} at high wave numbers, just as in run 1. At low wave
numbers, there is a stronger build up of $ E_A $ than $ E_K $, despite the fact that
the inverse energy cascade is restricted to KE. This can be explained by the KE-APE
conversion at small wave numbers that we just discussed.  In the lower plot we see that
$ E_W $ is falling off very close to $ k^{-5/3} $ in a limited range at wave numbers
larger than $ k_f $. At very high wave numbers  $ E_W $  is decreasing faster than in
run 2 and no bottleneck is observed. This may be due to the fact that the effective
diffusion coefficient is larger in run 3 as compared to run 1, since the total
dissipation is lower while we use the same numerical value of $ \nu_4 $.  Just as in
run 1, $ E_{V} $ has an approximate $ k^{-3} $-dependence in a broad range. However,
in Fig.~\ref{S2} we see some wiggles of $ E_W $ which were not present in run 1,
neither  in earlier stages of run 3, but only in the late stage. Taking a closer look
at the upper plot, we can see that these wiggles are also present in $ E_K $. It may be
suggested that these wiggles somehow are produced by the forcing. However, since they
were neither observed in run 1 nor in the earlier stages of run 3, we find it more
likely that there is another explanation. { As we saw in run 2, the vortical spectrum
may also display a more shallow dependence, close to $ k^{-5/3} $, if forcing is
applied in vortical modes. It may be the case that the late stage of run 3 has reached
a state in which the $ k^{-3} $ vortical spectrum is unstable and subsequently will be
replaced by a more shallow spectrum. }

The  potential enstrophy flux, $ \Pi_q $,  in run 3 had a very similar appearance as in run 1, why we don't show this plot.

In Fig.~\ref{Vis2} we show visualisations of the vortical field (left plot) and the wave field (right plot) from the end state of run 3. In the left plot we see that a multitude of coherent vortices has emerged in the vortical field. There is a strong dominance of anti-cyclonic vortices, displayed as dark,  almost circular,  disks of negative $ q $. An interesting feature of these vortices is that they are surrounded by  thin filaments of positive $ q $, displayed in yellow. The strong dominance of anti-cyclonic vortices has also been observed in shallow water  simulations by several investigators  \citep{Cushman1990, Arai1994, Polvani1994, Cho1996} and there have been many attempts to explain the phenomenon. An excellent review on the matter is given by \citet{Graves2006}, together with a plausible explanation. According to these authors, anti-cyclonic dominance is not only a feature of the shallow water equations but is also seen in natural flows in which the influence of boundaries is secondary. In the right plot we see that the wave field in run 3 displays much smaller structures than the vortical field, just as in run 1.  As can be seen from the blow up of the wave field inserted in the right plot, vortices are strong centres of high wave activity and the wave field within a vortex displays patterns that follow the organisation of the vortex filaments.  In Fig.~\ref{Vis3} we see a blow up of the potential vorticity field (left) of a strong anticyclonic vortex and the corresponding wave field in the same vortex (right). The high level of wave activity within the vortex suggests that there is a high level of energy dissipation in the vortex.  It is one of the most fascinating mysteries of fluid dynamics how  long live vortices may be sustained in the presence of dissipation.

 \section{Conclusions}

Comparing the spectral energy fluxes from a full GCM shown in Fig.~\ref{Flux} with the corresponding fluxes in Fig.~\ref{F2} from run 3 of the toy model we see that there is a striking similarity. A considerable amount of APE which is forced at large scales is converted into KE. A part of the KE is going into a downscale cascade at $ k > k_f $, while the other part, which is of the same order of magnitude, is going into an upscale cascade at $ k < k_f $. The upscale cascade is restricted to KE and is produced by interactions only involving the rotational component of the velocity field. The downscale energy cascade, on the other hand, is a cascade of both APE and KE. Here we also see an important difference. In the toy model there is perfect equipartition between KE and APE at high wave numbers and also between the fluxes $ \Pi_K $ and $ \Pi_A $. In the GCM, on the other hand, $ \Pi_K > \Pi_A $ at corresponding wave numbers. Clearly, the downscale energy cascade in the GCM is not a clean wave energy cascade with equipartition between KE and APE. \citet{Callies-Ferrari-Buhler:2014} have suggested that the mesoscale $ k^{-5/3} $ spectrum is produced by inertia-gravity waves with frequencies that are close to $ f $. For such waves, $ E_K > E_A $. However, the kinetic energy content in the rotational component of the velocity field is necessarily smaller than in the divergent component, and that is at odds with observational evidence \citep{Lindborg:2015, QiangLi2017}.  There is also numerical evidence \citep{Deusebio-Augier-Lindborg:2013, Asselin2017} suggesting that the forward energy mesoscale energy cascade cannot be described as a cascade of waves at high wave numbers, although it may start as a wave cascade at low wave numbers. In any case, as we pointed out earlier, the forward energy cascade in the GCM cannot be produced by inertia-gravity waves with frequencies close to $ f $, since it would require much finer resolution in the vertical to simulate such motions. Neither can it be fully accounted for by stratified turbulence as suggested by \citet{Lindborg:2006}, since that would also require much finer resolution. In the toy model simulation, on the other hand, we have been able to simulate a downscale energy cascade using only a single level. All nonlinear interactions that are present in the toy-model are also present in the dynamical equations of a full GCM, at each level. Moreover, in the GCM both $ E_A $   and $ E_K $ can be expressed as quadratic quantities to lowest order and the  sum of $ E_A $ and $ E_K $ is approximately conserved. Therefore, it seems likely that the toy model catches the essential cascade dynamics of the GCM. An important difference is, of course, that  inertia-gravity waves in the full GCM, as well as in nature, are three-dimensional, anisotropic and much more dispersive than in the toy model.

{The toy model is a tool by which energy and enstrophy cascades can be studied from the point of view of turbulence theory. Since the expressions for $ E_A $  and $ E_K $ are quadratic, the dynamics can be investigated by means of triadic interactions in Fourier space, just as 2D turbulence.  As we have shown in this study different types of interactions between vortical and wave modes may result in different cascade pictures. What types of interactions that are activated is not only dependent on $ Ro $ and $ Fr $ but also on the type of forcing which is used. Different types of forcing may lead to different types of cascades. If forcing is primarily applied in  wave modes, as in run 1, the vortical energy spectrum displays a $ k^{-3} $-range associated with a positive enstrophy flux which is limited, although not perfectly constant.  On the other hand, if forcing is primarily applied in vortical modes, as in run 2, the vortical energy spectrum is much more shallow, close to $ k^{-5/3} $, and the potential enstrophy flux is much larger. Evidently, vortical energy may either display an enstrophy cascade range or an energy cascade range depending on what types of interaction are activated. In a future study we will investigate this further.
}

A limitation of this study is that we have carried out the simulations on an $ f $-plane without including the $ \beta $-effect. The $ \beta $-effect may easily be included in the model. Another possible continuation is to implement the model on a rotating sphere and compare the results with simulations of the shallow water equations which were carried out by \citet{Cho1996}. Since the toy model shows the same type of cyclonic/anticyclonic asymmetry as the shallow water equations it may also be used as a tool to further investigate this issue. A definite advantage in this context, is that the model does not produce any shocks.

Indeed,  the toy model is constructed to meet certain mathematical desiderata -- quadratic expression for total energy and absence of shocks -- rather than as a model of a physical system that can be set up in an experiment.  In vain, we have tried to derive the model by constructing a gedanken experiment of a physical system that  to lowest order would evolve in accordance with the model equations. Until such a derivation is presented, we have to be content in using it as nothing more, nothing less,  than a toy.

%\begin{figure}[h]
%\centerline{\includegraphics[angle=0,width=10.5cm]{./Fig12.eps}}
% \caption{Time averaged potential enstrophy flux from run 2, displayed in a log-log plot. The sign has been reversed in the low wave number range, where the flux is negative (dashed line), implying that potential enstrophy is going both upscale and downscale.  Averaging was made over the intervall $ t/\tau \in [756 \; 1224] $.}
 %\label{EF2}
 % \end{figure}

% If in two-column mode, this environment will change to single-column format so that long equations can be displayed. 
% Use only when necessary.
%\begin{widetext}
%$$\mbox{put long equation here}$$
%\end{widetext}p

% Figures should be put into the text as floats. 
% Use the graphics or graphicx packages (distributed with LaTeX2e).
% See the LaTeX Graphics Companion by Michel Goosens, Sebastian Rahtz, and Frank Mittelbach for examples. 
%
% Here is an example of the general form of a figure:
% Fill in the caption in the braces of the \caption{} command. 
% Put the label that you will use with \ref{} command in the braces of the \label{} command.
%

% Tables may be be put in the text as floats.
% Here is an example of the general form of a table:
% Fill in the caption in the braces of the \caption{} command. Put the label
% that you will use with \ref{} command in the braces of the \label{} command.
% Insert the column specifiers (l, r, c, d, etc.) in the empty braces of the
% \begin{tabular}{} command.
%
% \begin{table}
% \caption{\label{} }
% \begin{tabular}{}
% \end{tabular}
% \end{table}

% If you have acknowledgments, this puts in the proper section head.
\section*{Acknowledgments}
We thank  Pierre Augier  for interesting discussions and for helping us with the development of the code. Financial support from the Swedish Research Council (VR) is gratefully acknowledged. We thank two anonymous reviewers for useful comments and suggestions. 
The simulations were performed on resources provided by the Swedish National Infrastructure for Computing (SNIC)  at  PDC.

% Create the reference section using BibTeX:

% \bibliography{references}

% \end{document}
%
% ****** End of file aiptemplate.tex ******

### 2019-04-25 paper_04_shallow_water/Latex/article.tex
Add shallow water paper with all references (some duplicates)
Figure path fixed as:

sed -i 's|../Pyfig|paper_04_shallow_water/Pyfig|g' paper_04_shallow_water/Latex/article.tex

% This is file JFM2esam.tex
% first release v1.0, 20th October 1996
%       release v1.01, 29th October 1996
%       release v1.1, 25th June 1997
%       release v2.0, 27th July 2004
%   (based on JFMsampl.tex v1.3 for LaTeX2.09)
% Copyright (C) 1996, 1997 Cambridge University Press

% \NeedsTeXFormat{LaTeX2e}
%
% \documentclass{jfm}
%\documentclass[referee]{jfm} %for double spaced output for submission

% \usepackage{graphicx}
% \usepackage{natbib}
%
% \usepackage{color}
% \usepackage{soul}
% \usepackage{booktabs}
% \pdfoptionpdfminorversion 4

% \pdfcompresslevel0

% See if the author has AMS Euler fonts installed: If they have, attempt2
% to use the 'upmath' package to provide upright math.
% \ifCUPmtlplainloaded \else
%   \checkfont{eurm10}
%   \iffontfound
%     \IfFileExists{upmath.sty}
%       {\typeout{^^JFound AMS Euler Roman fonts on the system,
%                    using the 'upmath' package.^^J}%
%        \usepackage{upmath}}
%       {\typeout{^^JFound AMS Euler Roman fonts on the system, but you
%                    dont seem to have the}%
%        \typeout{'upmath' package installed. JFM.cls can take advantage
%                  of these fonts,^^Jif you use 'upmath' package.^^J}%
%        \providecommand\upi{\pi}%
%       }
%   \else
%     \providecommand\upi{\pi}%
%   \fi
% \fi

% See if the author has AMS symbol fonts installed: If they have, attempt
% to use the 'amssymb' package to provide the AMS symbol characters.

% \ifCUPmtlplainloaded \else
%   \checkfont{msam10}
%   \iffontfound
%     \IfFileExists{amssymb.sty}
%       {\typeout{^^JFound AMS Symbol fonts on the system, using the
%                 'amssymb' package.^^J}%
%        \usepackage{amssymb}%
%        \let\le=\leqslant  \let\leq=\leqslant
%        \let\ge=\geqslant  \let\geq=\geqslant
%       }{}
%   \fi
% \fi
%
% See if the author has the AMS 'amsbsy' package installed: If they have,
% use it to provide better bold math support (with \boldsymbol).

% \ifCUPmtlplainloaded \else
%   \IfFileExists{amsbsy.sty}
%     {\typeout{^^JFound the 'amsbsy' package on the system, using it.^^J}%
%      \usepackage{amsbsy}}
%     {\providecommand\boldsymbol[1]{\mbox{\boldmath $##1$}}}
% \fi
%
%%% Example macros (some are not used in this sample file) %%%

% % For units of measure
% \newcommand\dynpercm{\nobreak\mbox{$\;$dyn\,cm$^{-1}$}}
% \newcommand\cmpermin{\nobreak\mbox{$\;$cm\,min$^{-1}$}}
%
% % Various bold symbols
% \providecommand\bnabla{\boldsymbol{\nabla}}
% \providecommand\bcdot{\boldsymbol{\cdot}}
% \newcommand\biS{\boldsymbol{S}}
% \newcommand\etb{\boldsymbol{\eta}}
%
% % For multiletter symbols
% \newcommand\Real{\mbox{Re}} % cf plain TeX's \Re and Reynolds number
% \newcommand\Imag{\mbox{Im}} % cf plain TeX's \Im
% \newcommand\Rey{\mbox{\textit{Re}}}  % Reynolds number
% \newcommand\Pran{\mbox{\textit{Pr}}} % Prandtl number, cf TeX's \Pr product
% \newcommand\Pen{\mbox{\textit{Pe}}}  % Peclet number
% \newcommand\Ai{\mbox{Ai}}            % Airy function
% \newcommand\Bi{\mbox{Bi}}            % Airy function
%
% % For sans serif characters:
% % The following macros are setup in JFM.cls for sans-serif fonts in text
% % and math.  If you use these macros in your article, the required fonts
% % will be substitued when you article is typeset by the typesetter.
% %
% % \textsfi, \mathsfi   : sans-serif slanted
% % \textsfb, \mathsfb   : sans-serif bold
% % \textsfbi, \mathsfbi : sans-serif bold slanted (doesnt exist in CM fonts)
% %
% % For san-serif roman use \textsf and \mathsf as normal.
% %
% \newcommand\ssC{\mathsf{C}}    % for sans serif C
% \newcommand\sfsP{\mathsfi{P}}  % for sans serif sloping P
% \newcommand\slsQ{\mathsfbi{Q}} % for sans serif bold-sloping Q
%
% % Hat position
% \newcommand\hatp{\skew3\hat{p}}      % p with hat
% \newcommand\hatR{\skew3\hat{R}}      % R with hat
% \newcommand\hatRR{\skew3\hat{\hatR}} % R with 2 hats
% \newcommand\doubletildesigma{\skew2\tilde{\skew2\tilde{\Sigma}}}
% %       italic Sigma with double tilde
%
% % array strut to make delimiters come out right size both ends
% \newsavebox{\astrutbox}
% \sbox{\astrutbox}{\rule[-5pt]{0pt}{20pt}}
% \newcommand{\astrut}{\usebox{\astrutbox}}
%
% \newcommand\GaPQ{\ensuremath{G_a(P,Q)}}
% \newcommand\GsPQ{\ensuremath{G_s(P,Q)}}
% \newcommand\p{\ensuremath{\partial}}
% \newcommand\tti{\ensuremath{\rightarrow\infty}}
% \newcommand\kgd{\ensuremath{k\gamma d}}
% \newcommand\shalf{\ensuremath{{\scriptstyle\frac{1}{2}}}}
% \newcommand\sh{\ensuremath{^{\shalf}}}
% \newcommand\smh{\ensuremath{^{-\shalf}}}
% \newcommand\squart{\ensuremath{{\textstyle\frac{1}{4}}}}
% \newcommand\thalf{\ensuremath{{\textstyle\frac{1}{2}}}}
% \newcommand\Gat{\ensuremath{\widetilde{G_a}}}
% \newcommand\ttz{\ensuremath{\rightarrow 0}}
% \newcommand\ndq{\ensuremath{\frac{\mbox{$\partial$}}{\mbox{$\partial$} n_q}}}
% \newcommand\sumjm{\ensuremath{\sum_{j=1}^{M}}}
% \newcommand\pvi{\ensuremath{\int_0^{\infty}%
%   \mskip \ifCUPmtlplainloaded -30mu\else -33mu\fi -\quad}}
%
% \newcommand\etal{\mbox{\textit{et al.}}}
% \newcommand\etc{etc.\ }
% \newcommand\eg{e.g.\ }
%
%
% \newcommand{\uu}{\textbf{u}}
% \newcommand{\xx}{\textbf{x}}
% \newcommand{\kk}{\textbf{k}}
% \newcommand{\JJ}{\textbf{J}}
% \newcommand{\rr}{\textbf{r}}
% \newcommand{\ff}{\textbf{f}}
% \newcommand{\FF}{\textbf{F}}
% \newcommand{\baa}{\textbf{a}}
% \newcommand{\bb}{\textbf{b}}
% \newcommand{\NN}{\textbf{N}}
%
%
%
%
%
% \newcommand{\ErtelPV}{\mathcal{Q}}
%
% \newcommand{\fdiss}{f_{\mbox{\scriptsize diss}}}
%
% \newcommand{\D}{\mbox{D}}
%
% \newcommand{\eez}{\boldsymbol{e_z}}
% \newcommand{\scalarprod}[2]{\big( #1 \, , \ #2 \big)_{\kk}}
%
% \newcommand{\mean}[1]{\langle #1 \rangle}
%
% \newcommand{\meane}[1]{\langle #1 \rangle}
% % \newcommand{\meane}[1]{{\langle #1 \rangle_{\hspace{-0.4mm}e}}}
%
% \newcommand{\meanx}[1]{{\langle #1 \rangle_{\hspace{-0.4mm}\mbox{\scriptsize$\xx$}}}}
%
% \newcommand{\meant}[1]{{\langle #1 \rangle_{\hspace{-0.4mm}\theta}}}
%
% \newcommand{\means}[1]{{\langle #1 \rangle_{\hspace{-0.4mm}s}}}
%
% \newcommand{\shocks}{ {\mbox{\tiny shocks}} }
%
% \newcommand{\kmax}{k_{\mbox{\scriptsize max}}}
% \newcommand{\kdiss}{k_{\mbox{\scriptsize diss}}}
%
%
% \newcommand{\mA}{\mathcal{A}}
%
%
% \newlength{\halfwidth}
% \setlength{\halfwidth}{2.6in}
%
% \newcommand{\PA}[1]{{\color{green}#1}}
%
% \newcommand{\Add}[1]{{\color{blue}#1}}
% % \newcommand{\Add}[1]{#1}
% \newcommand{\Remove}[1]{{\color{red}\st{#1}}}
% % \newcommand{\Remove}[1]{}
%
% \newtheorem{lemma}{Lemma}
% \newtheorem{corollary}{Corollary}
%
% \title[Shallow water wave turbulence]%
% { Shallow water wave turbulence}
%
% \author[P. Augier, A.V. Mohanan and E. Lindborg]%
% {Pierre Augier$^{1}$,   Ashwin Vishnu Mohanan$^{2}$
% and Erik Lindborg$^2$ \ns }
%
% \affiliation{$^2$ KTH Mechanics,
% SE-100 44 Stockholm, Sweden\\[\affilskip]
% $^1$ LEGI, BP53,
% 38041 Grenoble Cedex, France}
%
% \pubyear{2013}
% \volume{???}
% \pagerange{???--???}
% % Do not enter received and revised dates. These will be entered by
% % the editorial office.
% \date{?; revised ?; accepted ?. - To be entered by editorial office}
% %\setcounter{page}{1}
% \begin{document}
%
% \maketitle
%
%
% % \baselineskip 8mm

% \begin{abstract}
%
% The dynamics of irrotational shallow water wave turbulence forced in large
% scales and dissipated at small scales is investigated. First, we derive the
% shallow water analogue of the `four-fifths law' of Kolmogorov turbulence
% for a third order structure function involving velocity and displacement
% increments. Using this relation and assuming that the flow is dominated by
% shocks we develop a simple model predicting that the shock amplitude scales as
% $ (\epsilon d)^{1/3} $, where $ \epsilon $ is the mean dissipation rate and $ d $
% the mean distance between the shocks, and that the $ p $:th order displacement
% and velocity structure functions scale as $ (\epsilon d)^{p/3} r/d $, where $ r $
% is the separation. Then we carry out a series of forced simulations with
% resolutions up to $7680^2$, varying the Froude number, $F_{f} = (\epsilon L_f)^{1/3}/
% c $, where $ L_f $ is the forcing length scale  and $c$ is the wave
% speed. In all simulations a stationary state is reached in which there is a
% constant spectral energy flux and equipartition between kinetic and potential
% energy in the constant flux range. The third order structure function relation
% is satisfied with a high degree of accuracy. Mean energy is found to scale approximately as $
% E \sim \sqrt{\epsilon  L_f  c} $, and is also dependent on resolution, indicating
% that shallow water wave turbulence does not fit into the paradigm of a
% Richardson-Kolmogorov cascade. In all simulations shocks develop, displayed as
% long thin bands of negative divergence in flow visualisations. The mean
% distance between the shocks is found to scale as $ d \sim F_f^{1/2} L_f $.
% Structure functions of second and higher order are found to scale in good
% agreement with the model. We conclude that in the weak limit, $ F_f \rightarrow
% 0 $, shocks will become denser and weaker and finally disappear for a finite
% Reynolds number. On the other hand, for a given $ F_{f} $, no matter how small,
% shocks will prevail if the Reynolds number is sufficiently large.
%
% \end{abstract}
%
% \begin{keywords}
% \end{keywords}

\section{Introduction}

The shallow water (SW) equations  have been widely used to study basic
mechanisms occurring in geophysical flows \cite[see for
example][]{VallisLIVRE2006}.
%
The equations are based on the hydrostatic approximation which is
extremely well satisfied for a very wide range of scales in the oceans
and in the atmosphere.  They also rely on a stronger hypothesis
regarding the density stratification.  The fluid is assumed to be
structured in a limited number of thin homogeneous layers, which leads
to equations involving only two-dimensional operators. The two-layer
SW equations capture the baroclinic instability which is of
primary importance for the dynamics of the oceans and of the
atmosphere \cite[]{VallisLIVRE2006, Wirth2013}.
%
The single layer SW equations are very quite to
the two-dimensional compressible Navier-Stokes equations.  The
equations do not capture baroclinic instability but they constitute
one of the simplest hydrodynamic models with coexisting eigenmodes of
the linearized operator with zero linear frequency and with finite
non-zero linear frequency, $\omega = \pm\sqrt{f^2 + c^2
k^2}$, where $f$ is the Coriolis parameter, $c$ the gravity wave speed and $k
= |\kk|$ the wave number.
%

%
Due to their relative simplicity, the SW equations have
been used to study issues like the projection on the balanced manifold
\cite[]{Lorenz1980, MohebalhojehDritschel2000} and the production of
waves by balanced flows \cite[]{FargeSadourny1989,
LahayeZeitlin2012,Vanneste2013}. Balanced flows have a dynamics which is very similar to
two-dimensional turbulence, with an upscale energy cascade. %
Small scale energy dissipation vanishes in the limit of zero
viscosity.
%
In contrast to this, statistical mechanics indicates that wave energy
should be transferred towards small scales \cite[]{Warn1986}.
Numerical simulations have shown that long waves lose their energy by
transferring it to shorter waves \cite[]{Sadourny1975,
FargeSadourny1989}.   \cite{YuanHamilton1994} showed that
statistically stationary shallow-water flows can be obtained by forcing
quasi-geostrophic modes at large scales and with dissipation only at
small scales. They suggested that the $ k^{-5/3} $ energy spectrum observed at   atmospheric mesoscales  \cite[]{NastromGage1985,LiLindborg2018} is generated by a downscale cascade of gravity waves which may be captured by the SW equations.

The $ k^{-5/3} $ spectrum has been reproduced 
 by some General Circulation Models (GCM:s) based on the primitive
equations \cite[]{KoshykHamilton2001, Skamarock2004,
Hamilton_etal2008}.  \cite{AugierLindborg2013}  formulated a spectral energy budget of kinetic and available potential energy in terms of spherical harmonics and used this formulation to investigated two GCM:s. It was found that one of the models -- named the AFES model -- was able to reproduce a downscale energy cascade with realistic $ k^{-5/3} $ mesoscale spectra. The result was almost too good, since the AFES model simulation only contained 21 vertical levels, while theoretical arguments based on stratified turbulence theory \cite[]{Lindborg2006} suggest that a much finer vertical resolution would be needed. This lead us to ask what number of vertical levels a model needs in order  to reproduce a downscale energy cascade. Was \cite{YuanHamilton1994}  right  when they suggested that even a single layer model, such as the SW equations, could do the job? 
The original motivation of this study was to investigate this suggestion. A first set of simulations of the SW equations, containing both balanced and gravity wave motions, showed some encouraging results. Indeed, gravity waves were transferring energy towards small scales. However, after a while it became clear that the cascade was not as strong as in the GCM simulation.  Moreover, in all the SW simulations we observed shocks with an associated $ k^{-2} $-spectrum rather than a $ k^{-5/3} $-spectrum. These observations lead us to dismiss the SW equations as a model for the atmospheric downscale cascade. Based on the SW equations we developed another two-dimensional toy model  \cite[]{LindborgMohanan} that is absent of shocks and can reproduce the energy cascade in the GCM quite well, including a $ k^{-5/3}  $-spectrum.

The shocks we observed in the SW simulations left us in a state of curiosity. In some simulations they appeared to be pretty strong and in others rather weak. In some simulations they were densely packed and in others sparsely distributed. 
In what way  do the input parameters regulate these variations? Will there always be shocks or will they disappear in the limit of weak nonlinearities?  Such questions made us to shift our theoretical focus and pursue the study, despite the fact that the answer to the question that originally motived it turned out to be negative.

Some predictions  based on the weak wave turbulence formalism
\cite[]{ZakharovLvovFalkovich1992, Nazarenko2011} may be of relevance in the limit of weak nonlinearities. In the case of
inertia-gravity waves, \cite{FalkovichMedvedev1992} showed that the exact
solutions of the kinetic equation corresponding to a constant downscale energy
flux is associated with a spectrum scaling as $E(k) \sim k^{-8/3}$, while
\cite{ZakharovSagdeev1970} used weak turbulence formalism to derive a spectrum
of the form $ E(k) \sim k^{-3/2} $ for three-dimensional acoustic turbulence. This prediction is
of particular interest, since the acoustic equations and the SW equations are similar 
to the equations describing acoustic turbulence. It may be questioned, however, that weak turbulence formalism is
applicable to acoustic and SW wave turbulence, no matter how weak the
nonlinearities are supposed to be.  \cite{KadomtsevPet1973} argued that shocks always will always develop in acoustic turbulence, also in the limit of weak nonlinearities, due to the fact that acoustic waves are nondispersive.

Shocks have been been observed in many simulations of the SW equations \cite[]{FargeSadourny1989,
PolvaniMcWilliamsSpallFord1994, LahayeZeitlin2012}. As shown by \cite{Burgers1948} in the one-dimensional case and \cite{Kuznetsov2004} in the two- and three-dimensional cases, the energy spectrum of a flow dominated by
shocks will scale as $ k^{-2} $, a prediction which has been confirmed by a
great number of simulations of Burgers equation \cite[for a review, see ]
[]{Frisch2001}. The shocks also lead to very strong intermittency, as
demonstrated by \cite{BouchaudMezardParisi1995} and
\cite{WeinanKhaninMazelSinai1997} for Burgers turbulence.
%
\cite{FalkovichMeyer1996} simulated acoustic turbulence dominated by shocks
and reported spectra scaling as $ k^{-2} $.
\cite{YuanHamilton1994} carried out
forced dissipative simulations of the SW equations showing similar spectra.
However, the resolution in both these studies was quite coarse, ($386^2$
and $101^2$ grid points, respectively) so that the scaling range was very
narrow.  Moreover, \cite{FalkovichMeyer1996} used a very particular
type of strongly anisotropic forcing with waves in only one
direction while \cite{YuanHamilton1994} forced in geostrophic modes
which severely complicates the interpretation. Clearly, a new study of two-dimensional shock dominated wave turbulence can bring some more light on the problem.

In this paper, we study SW wave turbulence as a generic case of a forced-dissipative wave breaking system. Although there may not be any straightforward meteorological or laboratory applications,  we believe that the investigation has interesting implications for a general understanding of wave turbulence and that there may be long term applications in different fields, such as acoustics and cosmology.  To start with, we present some theory and then we report on a series of simulations of irrotational SW wave turbulence.

\section{Theory}

In a non-rotating system of reference the SW equations for a thin layer of fluid over a flat bottom and with a free
surface can be written as  \cite[see for example][]{VallisLIVRE2006}
\begin{eqnarray} \label{UEq}
\p_t \uu + \uu\cdot \bnabla  \uu   
&=& - c^2 \bnabla h + \frac{\nu}{h}  \bnabla^2 \uu,   \label{eq_uu}\\ \label{HEq}
\p_t   h    &=& - \bnabla \cdot (h \uu).  \label{eq_h}
\end{eqnarray}
where $\uu$ is the horizontal velocity, %
$h = 1 + \eta $ the non-dimensional thickness of the fluid layer,  $ \eta $ being the surface displacement, 
$\nu$ the kinematic viscosity and $ c $ is the gravity wave speed. The SW equations conserve mass, $ h $, and momentum $ {\bf J} = h{\bf u} $. As a matter of fact, we have constructed the viscous term in (\ref{UEq}) in such a way that momentum is conserved and energy dissipation is positive definite. This can only be accomplished by permitting the term to be nonlinear in the flow variables. There are different forms of nonlinear terms that fulfil both conditions and it may be disputed which is the best one. However, in the simulations we will use diffusion terms in both (\ref{UEq}) and (\ref{HEq}) that will not fulfil any of these two conditions in a strict sense and we will therefore not discuss this matter further. The inviscid SW equations conserve a local
quantity along trajectories of the fluid particles, the Ertel
potential vorticity $\ErtelPV = \zeta /h$, where $\zeta  $ is the vorticity, which in a rotating system should be interpreted as absolute vorticity.  This conservation law will be of no relevance in the present study in which we only consider
irrotational flows for which the velocity can be written as $ \uu = \bnabla \phi $ , where $ \phi $ is the velocity potential. In a domain with no net fluxes through the boundaries the inviscid equations conserve energy, $ E = E_K + E_P $, where $ E_K =   {\bf J}  \cdot {\uu} /2 $ is kinetic energy and $ E_P = c^{2} h^2/2 $ is potential energy. The latter can be split into three terms as \begin{equation}
E_P = c^2/2 + c^2\eta+ c^2 \eta^2/2, 
\end{equation}
where the first term corresponds to the potential energy of a state with no
surface displacement.  The third term, which is quadratic in the surface displacement, may be named available potential energy (APE),
$E_A = c^2\eta^2/2$, a concept introduced by  \cite{Lorenz1955}. As a consequence of mass and energy conservation the SW equations also conserve the sum of kinetic and available potential energy, $ E_K + E_A $.
 The velocity potential and the displacement of the linearised inviscid SW equations both satisfy the wave equation,
\begin{equation}
\frac{\partial ^2 \eta}{\partial t^2} = c^2 \nabla^2 \eta \, ,
\end{equation}
with equipartition between KE and APE over a wave period in each Fourier mode.

In order to analyze the spectral flux of energy, we derive the spectral energy
budget, i.e.\ the governing equations for spectral KE and APE
functions. This is not as straightforward as in incompressible turbulence since
the expression of the kinetic energy  is not
quadratic. To define the spectral KE function, we use the relation
\begin{equation}
\mean { E_K } = \sum_\kk \scalarprod{\uu}{\JJ}/2,
\end{equation}
where
\begin{equation}
\scalarprod{\baa}{\bb} \equiv \Re\{ \hat \baa(\kk)^* \cdot \hat \bb(\kk) \},
\end{equation}
and $\Re$ denotes the real part, $\kk$ is the wave number and the
hat denotes the Fourier transform.
%
The spectral KE function can therefore be defined as
\begin{equation}
E_K(\kk) \equiv \scalarprod{\uu}{\JJ}/2.
\end{equation}
Similarly, the APE can be written as the sum over all wave numbers of
the spectral APE function 
\begin{equation}
E_A(\kk) =  c^2 \scalarprod{\eta}{\eta}/2 \, .
\end{equation}
The inviscid equation for the spectral KE and APE functions can be written as
\begin{eqnarray}
\p_t E_K(\kk) &=& T_K(\kk) - C (\kk),\\
\p_t E_A(\kk) &=& T_A(\kk) + C (\kk),
\end{eqnarray}
where
\begin{equation}
T_K(\kk) 
= -\scalarprod{\bnabla \cdot \uu}{E_K} 
-\scalarprod{\uu }{\bnabla E_K } \, ,
\end{equation}
\begin{equation}
T_A(\kk) = -c^2 \scalarprod{\bnabla \eta }{\JJ}- c^2\scalarprod{\eta}{\bnabla \cdot \JJ}  \, ,
\end{equation}
are the spectral transfer functions of KE and APE, respectively, and
\begin{equation}
C (\kk) 
= \scalarprod{\uu}{\bnabla E_P}
\end{equation}
is the spectral KE-APE conversion function.
The spectral energy flux functions of KE and APE are defined as
\begin{eqnarray}
\Pi_K (k) = \sum_{| {\bf k}^{\prime} | > k} T_K ( {\bf k}^\prime ) \, , \;\;\;  \Pi_A (k) = \sum_{| {\bf k}^{\prime} | > k} T_A ( {\bf k}^\prime ) 
\end{eqnarray}

We will now derive an analogue of the so called `four-fifths  law' \cite[]{Kolmogorov1941} for the the third order velocity structure function of incompressible isotropic turbulence.   In the case of irrotational SW wave turbulence the analogous law contains a third order structure function involving both velocity and displacement increments.  In order for such a law to exist it is necessary that the velocity field is either non-divergent (incompressible) or irrotational. To derive the law it is convenient to start from the following equations
\begin{eqnarray} \label{IRR}
\p_t \uu &=& -\bnabla  (\uu \cdot \uu /2) - c^2 \bnabla h  \\
\p_t \JJ &=& -\p_j (u_j \JJ) - \bnabla E_p,\\
\p_t h   &=& -\bnabla \cdot \JJ \, ,
\end{eqnarray}
where we have used the  assumption that the velocity field is irrotational in (\ref{IRR}). We consider two points, positioned at $ \bf x $ and $ {\bf x}^\prime $, separated by the vector $ {\bf r} = {\bf x}^\prime - {\bf x} $.
Using Cartesian tensor notation we denote derivates with respect to $ {\bf x} $ as $ \partial_i $ and derivatives with respect to $ {\bf x}^\prime $ as $ \partial^{\prime}_i $. Treating $ \bf x $ and $ {\bf x}^\prime $ as independent variables we find

\begin{equation}
\p_t ( J_i' u_i ) = - \p_i( J_i'|\uu|^2/2) - \p_i'(u_j J_j' u_i')
-\p_i (J_i' c^2 h)  - \p_i' (u_i E_p')
\end{equation}
and
\begin{equation}
\p_t ( h' h ) = - \p_i( h'J_i ) - \p_i'( hJ_i' ),
\end{equation}
which gives
\begin{eqnarray}
\p_t ( \JJ'\cdot\uu + \JJ\cdot\uu' + 2c^2 h'h )
&=& - \p_i( J_i'|\uu|^2/2) - \p_i'( J_i|\uu'|^2/2) %\nonumber \\&& 
- \p_i'(u_j J_j' u_i') - \p_i(u_j' J_j u_i) \nonumber\\
&& -c^2\p_i (J_i'h) - c^2\p_i' (J_ih') %\nonumber\\&& 
- \p_i' (u_i E_p') - \p_i (u_i' E_p) \nonumber\\
&& -2c^2\p_i' (J_i'h) - 2c^2\p_i (J_ih').
\end{eqnarray}
We then assume homogeneity and take the average denoted by
$\meane{}$.   For
every function $g(\xx, \xx')$, we have $\p_i' \meane{g} =
-\p_i\meane{g} = \bnabla_{\rr} \meane{g}|_i$. We thus obtain
\begin{eqnarray}
\p_t \meane{ \JJ'\cdot\uu + \JJ\cdot\uu' + 2c^2 h'h }
&=& + \bnabla_{\rr}\cdot\meane{ \JJ'|\uu|^2/2 - \JJ|\uu'|^2/2 } \nonumber\\
&& - \bnabla_{\rr}\cdot\meane{ u_j J_j' \uu' - u_j' J_j \uu } \nonumber\\
&& - \bnabla_{\rr}\cdot\meane{ \uu E_p'- \uu' E_p } \nonumber\\
&& -c^2\bnabla_{\rr}\cdot\meane{ \JJ'h - \JJ h' }
\end{eqnarray}
%
We then introduce the structure functions and the operator $\delta$
returning the increment of a variable between two points separated by $\rr$, 
for example $\delta h (\xx, \rr) = h(\xx+\rr)- h(\xx)$.
%
Using homogeneity again, we obtain
\begin{equation}
\meane{ (\delta h)^2 \delta\uu } = 
-\meane{ h'^2 \uu } + \meane{ h^2 \uu' }
+2\meane{ hh' \uu } - 2\meane{ hh' \uu' }
\end{equation}
and 
\begin{equation}
\meane{ |\delta \uu|^2 \delta\JJ } = 
\meane{  |\uu|^2 \JJ' } - \meane{ |\uu'|^2\JJ }
+\meane{ u_ju_j' \JJ' } - \meane{  u_ju_j' \JJ },
\end{equation}
which gives
\begin{equation} \label{FluxT}
2\p_t \meane{ \JJ'\cdot\uu + \JJ\cdot\uu' + 2c^2 h'h }
= \bnabla_{\rr} \cdot ( \meane{ |\delta \uu|^2 \delta\JJ } 
+ c^2\meane{ (\delta h)^2 \delta\uu } ).
\end{equation}
For separations that are considerably larger than dissipative scales of motion and considerably smaller than forcing scales a standard argument \cite[see for example][]{Frisch} gives that the left hand side is approximately equal to $ -8 \epsilon $, where $ \epsilon $ is the mean dissipation rate. It is interesting to note that a factor of $ 8 $ appears in the case when the velocity field is irrotational while a corresponding factor of $ 4 $ appears in the case when the velocity field is non-divergent. 
Assuming isotropy and integrating yield the Kolmogorov law for
irrotational shallow water wave  turbulence
\begin{equation}
\meane{ |\delta \uu|^2 \delta J_L } 
+ c^2\meane{ (\delta h)^2 \delta u_L } = -4 \epsilon r, \label{eq_Kolmo}
\end{equation}
where $J_L \equiv
\JJ\cdot\rr / |\rr|$ and $u_L \equiv \uu\cdot\rr / |\rr|$ are
longitudinal increments.

We will now use (\ref{eq_Kolmo}) together with straightforward geometrical and statistical arguments to construct a model for velocity and displacement structure functions of any order above a certain minimum.
Our analysis has some similarities with the analyses by \cite{BouchaudMezardParisi1995}, \cite{WeinanKhaninMazelSinai1997} and \cite{Weinan}
for Burgers equation. 
We consider a domain with total area  $ {\mA} $ occupied by shocks which we 
consider as smooth lines. We denote the total length of all shocks by $ L_s $.  The mean distance between the shocks within the domain can be estimated as
\begin{equation}
d \simeq \frac{\mA} {L_s} .
\end{equation} 
A structure function of a flow variable is defined as the area average of increments  between two points separated by a distance $ r $.
Assuming that the contribution from increments where the two points are not separated by a shock is negligible for higher order structure functions we obtain
\begin{eqnarray}
\meane{|\delta h|^p} 
=  \frac{1}{\mA} \int_{\mA}  d^2\xx |\delta h(\xx)|^p 
 \simeq  \frac{r}{\mA} \int_\shocks ds |\cos\theta| |\Delta h (s)|^p,
\end{eqnarray}
where $\theta$ is the angle between the shock normal unit vector and the separation
vector $\rr$ %
and $\Delta h (s)$ is the shock displacement amplitude.
%
Assuming isotropy, the integral can be split and we get
\begin{equation} \label{StepH}
\meane{|\delta h|^p} 
\simeq 
r \frac{L_s}{\mA} \meant{|\cos\theta|} \means{|\Delta h |^p}
\simeq  \frac{2}{\pi} \means{|\Delta h |^p} \frac{r}{d} ,
\end{equation}
where $\meant{}$ is the mean over $\theta$
and $\means{}$ is the mean over all shocks.

For velocity increments, we can also use a characteristic of the
velocity singularities.  The step in velocity is confined to  the component
perpendicular to the shocks, which implies that the longitudinal and
transverse increments can be written as  $\delta u_L = \delta u
\cos\theta$, $\delta u_T = \delta u \sin\theta$.
%
Applying the same averaging method, we obtain
\begin{equation} \label{StepL}
\meane{|\delta u_L|^p} 
\simeq 
\meant{|\cos\theta|^{1+p}}
\means{|\Delta u |^p} \frac{r}{d}
\end{equation}
and
\begin{equation} \label{StepT} 
\meane{|\delta u_T|^p} 
\simeq 
\meant{|\cos\theta||\sin\theta|^{p}}
\means{|\Delta u |^p} \frac{r}{d},
\end{equation}
where $\Delta u $ is the shock velocity amplitude.
%
These assumptions completely determine the ratio between the
longitudinal and transverse structure functions, which can be computed as
\begin{eqnarray} \label{Ratio}
R_p(r) \equiv \frac{\meane{|\delta u_L|^p}}{\meane{|\delta u_T|^p}} = 
\frac{\meant{|\cos\theta|^{1+p}}}{\meant{|\cos\theta||\sin\theta|^p}},
\end{eqnarray}
giving the numerical values $R_2 = 2$, $R_3 = 6\pi/8$ and $R_4 = 8/3$.
Interestingly, for an isotropic and irrotational velocity field, the longitudinal and transverse second order structure functions
are exactly related by $\meane{(\delta u_L)^2} = \p_r( r\meane{(\delta
u_T)^2} )$ \cite[]{Lindborg2007jas}. If we further assume than both second order structure functions follow
the same scaling law $\meane{ (\delta u_L)^2} \propto \meane{ (\delta
u_T)^2} \propto r^\alpha$, then we have $\meane{ (\delta u_L)^2} =
(1+\alpha) \meane{ (\delta u_T)^2} $, which gives $ R_2 = 2 $ for 
$  \alpha = 1$.  In the limit $ r \rightarrow 0 $ we have $ \alpha = 2 $ and $ R_2 = 3 $.
%

In order for (\ref{StepH}-\ref{StepT}) to match with the third order structure function law (\ref{eq_Kolmo})  the shock amplitudes should scale as
\begin{equation} \label{Strength}
 | \Delta u | \sim | c \Delta h | \sim (\epsilon d)^{1/3} \, ,
\end{equation} 
and the structure functions of order $ p $ will consequently scale as
\begin{equation} \label{StrucFunc}
\meane{|\delta u |^p}  \sim \meane{(c|\delta h |)^p} \sim  (\epsilon  d)^{p/3} \,  \frac{r}{d} \, .
\end{equation} 
Assuming that (\ref{StrucFunc}) is valid for second order moments and that the non-quadratic contribution to the KE spectrum is negligible the corresponding scaling law for the energy spectra reads
\begin{equation} \label{Spectra}
E_K(k)  \sim  E_A(k) \sim \epsilon ^{2/3} d^{-1/3} k^{-2} \, .
\end{equation} 
The  model  provides the following predictions for the flatness factors of velocity increments
\begin{eqnarray} \label{FL}
F_L &=&  \frac{\meane{|\delta u_L|^4}}{\meane{|\delta u_L|^2}^2}
\propto 
\frac{\meant{|\cos\theta|^5}}{\meant{|\cos\theta|^3}^2} \frac{d}{r}   = \frac{9\pi}{15}  \frac{d}{r} \\ \label{FT}
F_T &=&  \frac{\meane{|\delta u_T|^4}}{\meane{|\delta u_T|^2}^2} 
\propto
\frac{\meant{|\cos\theta||\sin\theta|^4}}{\meant{|\cos\theta||\sin\theta|^2}^2} =  \frac{9\pi}{10}  \frac{d}{r} 
\end{eqnarray}
The ratio between the two flatness factors is equal  to
$F_T/F_L = 3/2$ according to the model.

We now turn to the question of how the shock width, $ \delta x $, can be estimated. One suggestion would be that $ \delta x $  is equal to the Kolmogorov length scale, $ \nu^{3/4}/\epsilon^{1/4} $. However, exact solutions of Burgers equation \cite[see for example][]{Whitham}
suggests that  $ \delta x \propto \nu $ rather than $ \delta x \propto \nu ^{3/4} $. The scaling of the energy spectra (\ref{Spectra}) also suggests that $ d $ should enter into the expression for $ \delta x$, apart from $ \nu $ and $ \epsilon $. Dimensional analysis thus gives
\begin{equation} \label{Width}
\delta x \sim \frac{\nu}{\; (d\epsilon)^{1/3}} \, ,
\end{equation} 
which is considerably smaller than the Kolmogorov length scale.
This result can also be derived in the following way. Assuming that all dissipation
takes place at the shocks, whose total area is equal to $ L_s \delta x $,  and
remembering that $ \epsilon $ is defined as the mean over the domain area $ \mathcal{A}$, we find
\begin{equation} \label{Estimate}
\epsilon \sim \nu \frac{ \mid \Delta u \mid ^{2} }{\delta x^2} \frac{L_s \delta x}{\mathcal{A}} \sim \nu \frac{\epsilon^{2/3}} {\delta x d^{1/3}} \, ,
\end{equation} 
again giving (\ref{Width}). 
Applying the same averaging method we obtain the following estimates for the skewness and flatness of the velocity derivative, $ \partial_x u $,
\begin{eqnarray}
S_{\partial_x u}  & \sim & \left ( \frac{d}{\delta x} \right )^{1/2} \sim \frac{d^{2/3} \epsilon ^{1/6}}{\nu^{1/2}} \, , \\
\label{FD} F_{\partial_x u} & \sim &  \frac{d}{\delta x} \sim \frac{d^{4/3} \epsilon^{1/3}} {\nu} \, ,
\end{eqnarray} 
where the skewness is negative. For Kolmogorov turbulence observations indicate that $ F_{\partial_x u} \propto \nu^{-1/6} $  \cite[]{VanAtta_Antonia1980}, a dependence which is well capture by the multifractal model of
\cite{Meneveau_Streenivasan1991}.  In SW wave turbulence there is a much stronger increase of  the flatness factor with decreasing  $ \nu $, which can be explained the fact that dissipation takes place only at the shocks.

Finally, we estimate the Mach number, $ M = U/c $, where $ U $ is the shock propagation velocity in the frame of reference where the fluid velocity in front of the shock is zero. 
Mass and momentum conservation give \cite[see][]{Baines1998} 
\begin{equation}
M = \sqrt{\frac{h_+}{h_-} \frac{h_+  + h_-}{2}} \, ,
\end{equation} 
where $ h_- $ and $ h_+ $ are the nondimensional layer thickness before and after the shock, respectively. Noting that $ h_+ - h_- =  \;  \Delta h  $, expanding in small $ \Delta h  $ and using (\ref{Strength}) the Mach number can be estimated as
\begin{equation} \label{Mach}
M \sim 1 + \frac{3}{4} \frac{\;\; (\epsilon d)^{1/3}}{c} \, ,
\end{equation} 
and is only slightly larger than unity for all relevant applications.

The analysis we have developed has, of course, some limitations. First, it is obvious that  the linear scaling of the structure functions only can be valid for separations  that fulfil the condition $ \delta x \ll r \ll d $.  Second, it is obvious that it  cannot be valid for low order moments or structure functions. According to arguments given by  \cite{BouchaudMezardParisi1995}  and \cite{Weinan}, it should be valid for structure functions of all orders $ p  \ge 1 $, in the asymptotic limit of large scale separation.  It is quite easily argued, however, that the shocks can not determine the scaling of the structure functions of order $ p=1 $. 
To see this, consider the first order moment of the absolute value of $ \delta u_L $. 
The longitudinal velocity increment is always negative over a shock, that is to say that the difference of the projection of the velocity on the direction over which the shock is crossed is always negative. In a domain with periodic boundary conditions the mean value of $ \delta u_L $   is zero. The negative values of $ \delta u_L $ for separations where a shock is crossed must therefore be compensated by positive values for separations where no shock is crossed. There is no reason to believe that these values are all positive,  but they are only positive on average. Thus, 
$ \means{ |\delta u_L | } < \langle | \delta u_L | \rangle _{ns} $, where the left hand side is the mean value over all separations where a shock is crossed and the right hand side is the mean value over separations where no shock is crossed.  Therefore, the analysis is not valid for first order moments. In the following, we will compare the predictions with structure functions of second and higher orders calculated from simulations.

\section{Simulations}

We have carried out a set of twenty forced-dissipative simulations of the SW
equations in a quadratic domain with periodic boundary conditions using the
standard open-source pseudo-spectral code FluidSim \cite[]{fluiddyn, fluidfft,
fluidsim}. In these twenty simulations we used an eight order diffusion operator to act as the agent of dissipation. 
 On the request from the reviewers we have also carried out a set of thirteen simulation with  Laplacian diffusion. The results from the latter set, which are similar to the results from the first set,  are presented in Appendix A.

The simulations were started from zero initial
conditions and were run till a statistically stationary state was reached. The
equations were reformulated in terms of the three eigenmodes of the linearized
equations in Fourier space. One of these eigenmodes is zero for an irrotational
flow. We first carried out a few test simulations in which all three eigenmodes
were allowed to evolve in time. However, when only the two wave modes are
forced vorticity remains very small throughout the simulations and does not
affect the wave dynamics. Therefore, we chose to solve the equations only for
the two wave modes. As discussed in the previous section, a linear viscous term
in (\ref{eq_uu}) will not conserve momentum exactly and does not give a
strictly positive definite expression for energy dissipation. We have relaxed
the requirements of exact momentum conservation and positive definite energy
dissipation. Instead of using the nonlinear form of the viscous term in
(\ref{eq_uu}) we have used a higher order linear operator $ \nu_8 \nabla^{8} $
both in (\ref{eq_uu}) and (\ref{eq_h}), where $ \nu_8 $ is a `hyperdiffusion'
or a `hyperviscosity'. There is no strong physical motivation for this
modification but, as discussed by \cite{FargeSadourny1989}, a molecular viscous
operator of the form $\nu \nabla^2 $ is not necessarily a relevant model of the
actual small-scale dissipation for SW flows, which should rather be described
as a transition from two-dimensional to three-dimensional motions before
reaching the scales where dissipation actually occurs.
From a numerical point of view it is very beneficial to introduce a direct sink of APE of equal magnitude as the sink of KE.
The addition of a hyperdiffusion term in (\ref{eq_h}) will provide such a sink without seriously affecting mass conservation, which will be 
extremely well fulfilled, even when a high order hyperdiffusion term is added.

%
Time advancement is carried out by a classical fourth-order Runge-Kutta scheme
for the nonlinear term and an exact integration for the linear and dissipative
terms. This explicit integration is especially useful at very high resolution
and very large $c$ when the shortest waves are very fast. We use an adaptable
time step method which maximizes the time step over a standard
Courant-Friedrichs-Lewy condition \cite[]{Lundbladh1999, AugierChomazBillant2012}.
%
Most of the aliasing is removed by truncating 8/9 of the modes along each
direction \cite[for a detail discussion on the issues of the non-conservation
of the non-quadratic energy and the aliasing errors in the truncated one-layer
shallow water model, see][]{FargeSadourny1989}. The resolution is characterised
by the number of nodes, $n$, in each direction and has been varied from 960
to 7680.
%
The wave speed $c$ has been varied over almost two orders of magnitude from 10
to 700. The wave modes are forced in a shell in spectral space corresponding to
relatively small wave numbers, $ 5 \delta k < |\kk| < 8 \delta k$, where $
\delta k = 2 \pi /L $ and $ L $ is the length of the side of the quadratic
domain.  We define the forcing wave number as $ k_f = 6 \delta k $. To obtain as broad scaling range as possible it is preferable to choose 
the forcing scale, $ L_f = \pi/k_f $, as large as possible. On the other hand, if $ L_f $ is too large, there is a risk that standing waves will be induced by the periodic boundary conditions and also that isotropy will be broken due to alignment of the shocks with the boundaries. 
We have found that $ L_f/L =1/12 $ is sufficient to avoid these two problems. 
The forcing is white noise in time and designed so that the energy
injection rate is normalised to unity at each time step. When the flow has reached a stationary state dissipation should therefore be equal to unity. 
The strength of the stratification is characterised by a
Froude number, defined as
\begin{equation} \label{Fr}
F_f = \frac{(\epsilon L_f)^{1/3}}{c} \, ,
\end{equation}
a parameter that can be determined from the start of each simulations.
Energy is dissipated at the smallest resolved scales by hyperviscosity. An intricate question is how the value of $ \nu_{8} $ should be chosen so that the shocks are resolved. The standard resolution requirement in turbulence simulations of this type is that the hyperviscosity analogue to the Kolmogorov scale
\begin{equation}  \label{KolmEight}
\eta_{8} = \left( \frac{{\nu_8}^3}{\epsilon} \right)^{\frac{1}{22}},
\end{equation}
should be resolved. However, our estimate (\ref{Width}) of the shock width suggests that this may not be a sufficient condition, since $ d $ is likely to enter the expression for $ \delta x $ also in the case when we use hyperviscosity.
A similar estimate as in (\ref{Estimate}) gives
\begin{equation} \label{EtaEight}
\delta x \sim \eta_{8} \left ( \frac{\eta_8}{d} \right)^{1/21} \, .
\end{equation} 
Thus, the shock width dependence on $ d $ can be expected to be extremely weak when hyperviscosity is used. Therefore we simply use the standard resolution requirement that (\ref{KolmEight}) should be resolved. We have found that setting
$  k_{d} \simeq \kmax/ 2.5$, where $ k_{d} = 1/\eta_8 $ and $\kmax =
(8/9)\pi n/L $ is the maximum wave number, is  sufficient to resolve the shocks. Since the energy injection rate is normalised to unity and all simulations reach a stationary state in which dissipation is equal to the injection rate, $ \nu_8 $ can be chosen already from the start so that this condition will be fulfilled in the stationary state. 
 The  simulation parameters are listed in table 1.

\begin{table}
\begin{center}

\begin{tabular}{lrrrrrrr}
\toprule
{} &   $n$ &  $c$ &  $\nu_8$ &  $\epsilon$ &  $\frac{k_{diss}}{k_f}$ &   $F_f$ &  $t_{\max}$ \\
\midrule
W1  &   960 &   10 & 1.56e-10 &        1.04 &                    28.8 &   0.16 &          50.0 \\
W2  &  1920 &   10 & 9.68e-13 &        1.06 &                    57.7 &   0.16 &        49.9 \\
W3  &  3840 &   10 &    6e-15 &         1.10 &                     115 &   0.17 &          50.0 \\
W4  &  7680 &   10 & 3.72e-17 &         1.20 &                     231 &   0.17 &        25.2 \\
W5  &   960 &   20 & 1.56e-10 &           1.00 &                    28.8 &  0.08 &          50.0 \\
W6  &  1920 &   20 & 9.69e-13 &        1.02 &                    57.7 &  0.081 &         120 \\
W7  &  3840 &   20 & 6.01e-15 &        1.08 &                     115 &  0.082 &         120 \\
W8  &  7680 &   20 & 3.72e-17 &        1.21 &                     231 &  0.086 &         120 \\
W9  &   960 &   40 & 1.56e-10 &       0.996 &                    28.8 &  0.040 &        49.9 \\
W10 &  1920 &   40 & 9.68e-13 &       0.998 &                    57.7 &  0.040 &        49.9 \\
W11 &  3840 &   40 &    6e-15 &        1.06 &                     115 &  0.041 &        49.9 \\
W12 &   960 &  100 & 1.56e-10 &       0.999 &                    28.8 &  0.016 &        49.9 \\
W13 &  1920 &  100 & 9.69e-13 &        1.01 &                    57.7 &  0.016 &         120 \\
W14 &  3840 &  100 & 6.01e-15 &        1.06 &                     115 &  0.016 &         120 \\
W15 &  7680 &  100 & 3.72e-17 &        1.22 &                     231 &  0.017 &        95.9 \\
W16 &   960 &  400 & 1.56e-10 &        1.14 &                    28.8 &  0.0042 &         120 \\
W17 &  1920 &  400 & 9.69e-13 &        1.15 &                    57.7 & 0.0042 &         120 \\
W18 &  3840 &  400 & 6.01e-15 &        1.12 &                     115 & 0.0042 &         120 \\
W19 &   960 &  700 & 1.56e-10 &        1.27 &                    28.8 & 0.0025 &        49.9 \\
W20 &  1920 &  700 & 9.68e-13 &        1.05 &                    57.7 & 0.0023 &        49.9 \\
\bottomrule
\end{tabular}

\vskip 3mm
\caption{Parameters for all simulations. $ n $: number of nodes in each
direction, $ c $: wave speed, $ \nu_8 $: hyperviscosity, $ \epsilon $: time
averaged mean energy dissipation in the stationary state, $ k_{d}/ k_f $: ratio
between dissipation wave number and forcing wave number, $ F_f $: Froude
number, $ t_{max} $: end time of simulation.}
\end{center}
\end{table}

\begin{figure}
\centerline{\includegraphics[width=5.12in]{paper_04_shallow_water/Pyfig/fig_Emean_time}}
\caption{Space averaged energy, $E = E_K + E_A $, versus time for all
simulations. }
\label{fig_Evstime}
\end{figure}

\begin{figure}
\centerline{\includegraphics[width=5.8in]{paper_04_shallow_water/Pyfig/fig_energy_w}}
\caption{Left: Mean energy in the stationary state versus $ c $ for different
resolutions. The lines have slope $ 0.5 $, corresponding to $ E \propto c^{1/2}
$. Right: Normalised mean energy in the stationary state versus resolution $ n
$ for different $ c $.}
\label{MeanE}
\end{figure}

The time evolution of mean energy is shown in figure~\ref{fig_Evstime} for all
twenty simulations. After an initial period of growth, mean energy is leveling
out at different values for different runs. This is not the same picture as we
see in simulations of three-dimensional incompressible turbulence or stratified
turbulence, where all curves would have leveled out more or less at the same
value. We find that mean energy in the stationary state is a function of both $
c $ (or the Froude number) and resolution $ n $ and that the functional
dependence can be approximately written as
\begin{equation} \label{MeanEnergy}
E \sim \sqrt{\epsilon L_f c } \, n^{\alpha}  \, .
\end{equation}
 In figure~\ref{MeanE} to the
left, we see $ E/\sqrt{\epsilon} $ as function of $ c $ for different resolutions $
n $. For each resolution the points are approaching a power law $ C_n c^{1/2} $
with increasing values of $ C_n $ for increasing $ n $. In the same figure to
the right we see $ E/\sqrt{\epsilon L_f c } $ for different values of $ c $. For
each $ c $ the points follow a power law $ n^{\alpha} $, where $ \alpha $ is
becoming slightly larger with increasing $ c $, approaching almost the same
value $ \alpha \approx 0.33 $ for the three highest values of $ c $. The
relation (\ref{MeanEnergy}) can be reformulated as
\begin{equation} \label{Dissipation}
\frac{\epsilon} {(E^{3/2} / L_f)} \sim \frac{E^{1/2}}{c} n^{-2\alpha} \, .
\end{equation}
Since the simulations are well resolved the dependence on resolution should not
be interpreted as consequence of the numerics but rather as a Reynolds number
effect, although it is not hundred percent clear how a Reynolds number should
be defined in our case. We may note, however, that $ n \sim L_f/\delta x $ and in the case of Navier-Stokes viscosity
 we should have $ \delta x \propto \nu $, according to (\ref{Width}). Therefore,  $ n $ may be substituted by a Reynolds number. The normalised energy dissipation on the left hand side of
(\ref{Dissipation}) will thus go to zero in the limit of large Reynolds number.
This is very different from three-dimensional turbulence where there is a local
Richardson-Kolmogorov cascade, in which energy is successively transfered from
large to small scales of motion. That the cascade is local means that eddies of
a particular scale are not directly influenced by eddies of widely different
scales. As a consequence, the normalised dissipation is of the order of unity
independent of Reynolds number \cite[]{Pope, TennekesLumley}. A review of the
experimental evidence and theoretical implications of this fundamental law of
Kolmogorov turbulence is given by \cite{Vassilicos2015}. Our result
(\ref{Dissipation}) suggests that SW wave turbulence does not fit into the
paradigm of a local Richardson-Kolmogorov cascade. Moreover, it is not only in
the limit of large Reynolds number the normalised dissipation will go to zero
but also in the limit of strong stratification, $ E^{1/2}/c \rightarrow 0 $.
This is different from three-dimensional stratified turbulence where the ratio
on the left hand side of (\ref{Dissipation}) is of the order of unity in the
limit of strong stratification \cite[]{Lindborg2006, Brethouwer2007}.

That the energy transfer is not local in a system where the agent of dissipation is wave breaking or shock formation may be expected. According to the theory of ocean waves by  \cite{Phillips}, 
`wave interactions are usually incapable of transferring energy from a given wavenumber
band as rapidly as it is supplied by the wind'. Instead, energy is fed into each wave mode by the
wind field and sucked out from the same mode in `intermittent patches of foaming'
or `white horses'.  This may be a too simplified picture of SW water turbulence, since a broad band energy spectrum develops between the forcing and the dissipation wave numbers. 
Nevertheless,  shock formation suggests that there is a considerable direct energy transfer from large to very small wave lengths, without any intermediate steps.

The `four-fifths law' for the third order structure function of Kolmogorov
turbulence is often derived using the assumption of a local cascade
\cite[]{Vassilicos2015} or the assumption that dissipation stays finite in the
limit of large Reynolds number \cite[]{Frisch}. As we just discussed, these
assumptions seem to be violated in the case of SW wave turbulence.
Nevertheless, the analogous law (\ref{eq_Kolmo}) is satisfied with a high degree of accuracy in our simulations. In
figure~\ref{Flux} we see the spectral energy flux from run W7 to the left and
the third order structure function to the right. There is a broad constant flux
range with equipartition between KE and APE flux and an almost equally broad
range where (\ref{eq_Kolmo}) is satisfied, with equipartition between the two
third order structure functions associated with KE and APE. The assumptions of
a local cascade and finite dissipation do not seem to be necessary in order for
the third order structure function law (\ref{eq_Kolmo}) to be valid.

\begin{figure}
\centerline{\includegraphics[width=5.8in]{paper_04_shallow_water/Pyfig/fig_flux_struct_combined}}
\caption{Left: Time averaged normalised spectral energy fluxes versus $ k/k_f
$. Right: Time averaged normalised third order structure functions versus $
r/L_f $. From run W7. }
\label{Flux}
\end{figure}

\begin{figure}
\centerline{\includegraphics[width=3.15in]{paper_04_shallow_water/Pyfig/fig_spatiotempspectra}}
\caption{Frequency spectra of KE and APE for different $ k = \mid {\bf k} \mid
$, from run W7. The spectra are plotted as functions of $\omega/\omega_l$,
where $\omega_l = c k$. From top to bottom: $ k /\delta k = 12$, 27, 62, 143,
327, and 746. }
\label{fig_spatiotemp_spectra}
\end{figure}

\begin{figure}
\centerline{\includegraphics[width=6.0in]{paper_04_shallow_water/Pyfig/fig_phys_fields_wave}}
\caption{Left: divergence $ \bnabla \cdot {\bf u} $. Right: velocity component in the $ y $-direction: $ u_y $. Top: run W6. Bottom: run W17.  }
\label{Physical}
\end{figure}

\begin{figure}
\centerline{\includegraphics[width=8cm]{paper_04_shallow_water/Pyfig/fig_shock_sep.pdf}}
\caption{Mean distance between the shocks as function of Froude number.  }
\label{fig_distance}
\end{figure}

The equipartition of KE and APE spectral fluxes suggests that the flow to
leading order may be described as a collection of linear gravity waves with energy
equipartition in each mode. To investigate this further we have
collected time series of the flow variables from Fourier modes whose magnitude
fall within certain shells, $ \mid {\bf k} \mid \in [ k -\delta k/2, \; k+
\delta k/2] $, and computed temporal power law spectra from these time series
and averaged over all wave numbers in each shell. In other words, we have
computed KE and APE frequency spectra corresponding to different magnitudes of
modes. Figure~\ref{fig_spatiotemp_spectra} presents such spectra plotted as 
functions of the normalised frequency $\omega/\omega_l$, where $ \omega_l = kc
$ is the linear wave frequency. The spectra are strongly dominated by peaks at
$\omega = \omega_l$. As can be seen, there is equipartition between KE and APE
in each shell. Although the widening of the peaks around the linear frequency
most likely is an effect of nonlinearities, we can quite safely conclude that
the time evolution of the flow variables in each Fourier mode to leading order
can be described as a linear wave.

Visualisations of the flow field give a completely different picture from what
may be expected from a collection of linear gravity waves. They are totally
dominated by the appearance of shock waves. In figure~\ref{Physical} we see
the divergence, $ \nabla \cdot {\bf u} $, to the left and the velocity
component in the $ y $-direction, $ u_y $, to the right.  Note that total area of the computational domain is sixteen times larger than the area of the outcuts shown in the figures, since the side length of the figures is $ 3 L_f $, while the side length of the computational domain is $ 12 L_f $.
The two figures at the
top are from run W6 while the two figures at the bottom are from run W17. The
difference between the two runs is that $ c $ is larger by a factor of $ 20 $
in W17 as compared to W6. To the left we see the shocks displayed as elongated
bands of negative divergence. That the divergence is negative at the shocks can
be understood from the fact that the velocity component perpendicular to the
shock always has a negative jump in the direction of shock propagation.
Apparently, the mean distance between the shocks is considerably smaller in run
W17 with the larger value of $ c $ compared to run $ W6 $ with the smaller
value of $ c $. 
The shocks are also visible in the figures to right where $ u_y
$ is displayed. It may be interesting to note that the shocks along the $ x
$-axis appear as extra sharp in this plot, since there is a strong jump of $
u_{y} $ over these shocks, while the shocks along the $ y $-axis are hardly
visible, since there is no jump of $ u_{y} $ over these. In the two figures to
the right we see a variation of the flow field over the length scale $ L_f $
which is not seen in the two figures to the left. This variation is evidently a
footprint of the random forcing. Although the forcing length scale $ L_f $ is
the same in the two simulations the mean distance, $ d $, between the shocks is
much smaller in run $ W17 $ than in run $ W6 $, which may come as a surprise. Another interesting observation is that the shocks in run W17 (lower figures) appear to have smaller curvature on average than the shocks in run W6 (upper figures). In both runs, the shocks are weak in the sense that the Mach number is close to unity and the interaction between the shocks should therefore be weak \cite[see][]{ApazidisEliasson2018}. However, as measured by the Mach number  (\ref{Mach}), the shocks are much weaker in run W17, where $ c $ is larger by a factor of 20 as compared to run W6 and $ d $ also is smaller. The shock interaction should therefore be weaker in run W17, which may explain the observation that the curvature of the shocks is smaller in run W17 than in run W6.
We have calculated  the mean distance, $ d $, between the shocks by counting the number of negative spikes
of the divergence along two hundred lines parallel to the $ x $-axis and equally many 
parallel to the $ y $-axis and then divided total length of all the lines by
the number of spikes. Generally, we found that $ d \propto c^{-1/2} $, or
\begin{equation} \label{MeanDistance}
d \sim L_f F_f ^{1/2} \, ,
\end{equation}
independent of resolution $ n $. In figure~\ref{fig_distance} we have plotted
$ d/L_f $ versus Froude number for all twenty runs. As can be seen, the points
follow the $ F_f^{1/2} $-curve quite well, although there seems to be a slight
systematic deviation at the highest $ F_f $.
\begin{figure}
\centerline{
\includegraphics[width=5.0in]{paper_04_shallow_water/Pyfig/fig_struct_order_246}}
\caption{From top to bottom: second, fourth and sixth order compensated and
normalised longitudinal structure functions. Left: runs with $ n=3840 $: W3,
W7, W11, W14 and W18. Right: runs with $ n = 7860 $: W4, W8 and W15. }
\label{fig_StrucFunc}
\end{figure}

\begin{figure}
\centerline{\includegraphics[width=11cm]{paper_04_shallow_water/Pyfig/fig_ratio_strfct}}
\caption{
Ratio of the structure functions
$R_p(r) = \mean{|\delta u_L|^p} / \mean{|\delta u_T|^p}$ from run W8,
for $ p = 2, \, 3, \, 4, \, 5, \, 6. $
The dotted straight lines indicate the values predicted by (\ref{Ratio}):
$R_2 = 2$, $R_3 = 3\pi/4$,  $R_4 = 8/3$, $ R_5 = 15 \pi /16 $ and $ R_6 = 16/5 $.}
\label{fig_ratio}
\end{figure}

\begin{figure}
\centerline{\includegraphics[width=12cm]{paper_04_shallow_water/Pyfig/fig_flatness}}
\caption{ Top: Flatness factor of transverse velocity increments. Bottom: Ratio between transverse and longitudinal flatness factor. Left: runs with $ n=3840 $: W3,
W7, W11,  W14 and W18. Right: runs with $ n = 7860 $: W4, W8 and W15.  The dotted straight lines indicate the value 1.5 predicted by the shock model.  }
\label{fig_flatness}
\end{figure}

\begin{figure}
\centerline{\includegraphics[width=11cm]{paper_04_shallow_water/Pyfig/fig_spectra}}
\caption{Time averaged compensated and normalised energy spectra
from runs with $ n = 7680 $: W4, W8 and W15.}
\label{fig_spectra_c40}
\end{figure}

\begin{figure}
\centerline{\includegraphics[width=13cm]{paper_04_shallow_water/Pyfig/fig_diss_spectra}}
\caption{Time averaged normalised dissipation spectra. Left: Constant $ c = 20 $ and varying $ n $, runs  W5, W6, W7 and W8. Right: Constant $ n= 1920 $ and varying $ c $, runs W2, W6, W10, W13, W17 and W20. The vertical lines indicate the wave numbers of the maxima.}
\label{fig_disspectra}
\end{figure}

We will now present the results for the structure functions. Substituting
(\ref{MeanDistance}) into (\ref{StrucFunc}) we find that the structure
functions are expected to scale as
\begin{equation} \label{StrucFunc2}
\meane{|\delta u |^p}  \sim \meane{(c|\delta h |)^p} \sim  (L_f F_f^{1/2})^{p/3-1}  \epsilon^{p/3}  r  \, .
\end{equation}
for separation distances that fulfil the condition $ \delta x \ll r \ll d $.
The scaling (\ref{StrucFunc2}) is
expected to become better and better fulfilled with decreasing $ c $ (or
increasing $ F_f $). According to (\ref{MeanDistance})
$ d $ is increasing with $ F_f $, and the scaling range is therefore becoming
wider for decreasing $ c $, if $ \delta x $ or the resolution  is kept
constant. The scaling
(\ref{StrucFunc2}) is also expected to become better and better fulfilled with
increased resolution $ n $, since $ \delta x $ is decreasing with increasing $
n $.

As expected, we have found that the displacement structure functions scale
exactly in the same way as the velocity structure functions, and for this reason we only plot
velocity structure functions. In figure~\ref{fig_StrucFunc} we see the
compensated and normalised second, fourth and sixth order longitudinal
structure functions from the five simulations with $ n = 3840 $ to the left and
the three simulations with $ n = 7860 $ to the right. At small separation, $ r
< \delta x $, the structure functions have the dependence $ \sim r^{p} $, which
is characteristic of a smooth velocity field, indicating that the shocks are
reasonably well resolved. There is a small bump around the shock width, $ r
\approx \delta x $, and at $ r > \delta x $, there is a range where the
compensated structure functions are flat, or almost flat. As expected, this
range is becoming broader for smaller $ c $ (or larger $ F_f $). The range is
also broader in the higher resolution runs to the right as compared to the
lower resolution simulations to the left. There is clearly a better collapse of
the curves for the sixth order structure functions compared to the second order
structure functions. This is not surprising, since the shock contribution to
the structure functions should become increasingly dominant with increasing
order. Plots of the transverse structure functions are very similar to the
plots of the longitudinal structure functions, with the difference that the
transverse structure functions are somewhat smoother at $ r \approx \delta x $.
In figure~\ref{fig_ratio} we see the ratio $ R_{p}(r) = \langle \mid \delta
u_L \mid ^{p} \rangle / \langle \mid \delta u_T \mid ^{p} \rangle$ for $ p=
2,3,4,5,6 $ from run W8, together with dotted straight lines, indicating the
values predicted by (\ref{Ratio}). As can be seen, there is a quite good
agreement with the predicted values in a limited range of separations, $ r >
\delta x $. Somewhat unexpectedly, the range in which there is a good agreement
becomes more narrow with increasing $ p $, an observation that we cannot
explain. It may also be noted that $ R_2 \rightarrow 3 $ in the limit of small
$ r $, which is the theoretical single point limit for an irrotational
isotropic field. In figure~\ref{fig_flatness} we see the flatness factor, $
F_T $, of the transverse structure functions at the top and the ratio, $
F_T/F_L $, at the bottom. The figures to the left show the results from the
five simulations with $ n = 3840 $ and the figures to the right show the
results from the three simulations with $ n = 7860 $. The flatness factor shows
increasingly large values for decreasing $ r $ with a dependence that is rather
close to but somewhat steeper than $ r^{-1} $. At small $ r $ the curves are
leveling out at very large values. For $ n = 7869 $ and $ c = 10 $ we see that
$ F_{T} \approx 400 $ at small $ r $. The degree to which the flatness factor
is increasing with decreasing scale is by some authors defined as the most
appropriate measure of spatial intermittency \cite[see for example][]{Frisch}.
If this measure is used, SW wave turbulence is, indeed, extremely intermittent.
In the two figures at the bottom we see that $ F_{T}/F_{L} \approx 1.5 $ in a
limited range of separations, $ r > \delta x $, which is becoming somewhat
broader as $ c $ is increasing and is also somewhat broader in the higher
resolution runs shown to the right as compared to the lower resolution runs
shown to the left. This is in accordance with the shock model predictions
(\ref{FL}) and (\ref{FT}).

Finally, we present energy and dissipation spectra. As expected, we found that APE and KE
spectra were almost identical, and therefore, we only plot the
total spectra. In figure~\ref{fig_spectra_c40} we see  compensated and
normalised energy spectra from the three highest resolution runs. The
compensated spectra are almost flat in a broad range. In the high wave number
end the compensated spectra show a `bottleneck', which we interpret as a
signature of the shocks. The $ k^{-2} $-dependence and the associated scaling predicted
by (\ref{Spectra}) are very well satisfied in a limited range of high wave
numbers close to the bottleneck. At smaller wave numbers the spectra are
becoming somewhat steeper than $ k^{-2} $ and the spectrum from the run with
largest $ c $ does not collapse perfectly onto the other two spectra.  In figure \ref{fig_disspectra}
 we see normalised dissipation spectra for fixed $ c= 20 $ and different $ n $ to the left, and fixed $ n = 1920 $
and different $ c $ to the right. First we note that the spectra go to zero quite nicely for large $ k $, indicating that the shock width $ \delta x $ is well resolved. Second, we note that there is a slight shift of the maximum in both figures, even though we have normalised the wave number by $ k_d = 1/\eta_8 $.   These shifts may be explained by the estimate (\ref{EtaEight}) predicting that $ \delta x $ should vary slightly with $ d $. If the maximum of the dissipation spectra is associated with $ \delta x $, there should be a shift of the maximum  to the right with increasing $ n $ and to the left with increasing $ c $, since $ d \propto c^{-1/2} $.
 The ratio between the maximum and the minimum $ n $ in the left figure is equal to 8 and  according to (\ref{EtaEight}) the corresponding ratio between the wave numbers at which the maxima are observed should be $ 8^{1/21} \approx 1.1 $, which is exactly the value we observe. The ratio between the maximum and minimum $ c $ in the right figure is equal to $ 70 $. Assuming that $ d \propto c^{-1/2} $ the ratio between the wave numbers at which the maxima are observed should be $ 70^{1/42} \approx 1.1 $, while the observed ratio is equal to $ 1.16 $. In the latter case, there is no exact quantitative agreement between the predicted trend and the numerical results.

\section{Conclusions and discussion}

We have carried out a theoretical and numerical investigation of SW wave
turbulence. First, we derived the SW analogue (\ref{eq_Kolmo}) of the
`four-fifths' law of Kolmogorov turbulence. Using this relation and
straightforward statistical and geometrical arguments we developed a simple
shock model predicting that the shock amplitude scales as $ (\epsilon d)^{1/3}
$, where $ d $ is the mean distance between the shocks, and that the $ p $:th
order structure function above a certain minimum will scale as $ (\epsilon
d)^{p/3} r/d $. Then, we carried out a series of  forced dissipative
simulations, varying the Froude number and the resolution. In all simulations,
a statistically stationary state was reached and in all simulations we
observed shocks. The flow variables in each Fourier mode were found to evolve
in accordance with linear wave dynamics, with equipartition between APE and KE
over a period. The third order structure function relation was fulfilled with a
high degree of accuracy. From the simulations we made two interesting
observations that fall outside the predictive scope of the model. The first
observation was that mean energy in the stationary state approximately scales
as (\ref{MeanEnergy}), suggesting that the normalised dissipation
(\ref{Dissipation}) will go to zero both in the limit of small viscosity and
large wave speed. A similar observation was made when we used a second order diffusion operator instead of hyper diffusion (see Appendix A).
This is different from three-dimensional Kolmogorov
turbulence as well as strongly stratified turbulence, where, in both cases,
dissipation is finite in the limit of small viscosity, and in the latter case
also is independent of the degree of stratification. This observation suggests
that SW wave turbulence does not fit into the paradigm of a local
Richardson-Kolmogorov cascade.  It remains a theoretical challenge to give a quantitative explanation of the growth of energy with increased Reynolds number. 
As far as wee can see, such an explanation would require a completely novel theoretical development.
The second observation we made was that the mean
distance between the shocks scales as $ c^{-1/2} $, if the other parameters are
held constant. As we increased $ c $, the shocks were thus becoming denser.
Using this observation we tested the shock model by plotting the normalised and
compensated structure functions of different orders. Generally we found that
there was a quite good agreement between the simulation results and the model
predictions, becoming better with decreasing wave speed and increasing
resolution.

Combining the model and the observations we can get a quite good picture of the
dynamics in the limit of weak nonlinearities, $ F_{f} \rightarrow 0 $. The
picture may become more vivid if we say `wave breaking' were we previously have
talked about shock formation, and also say `wave lengths' were we previously
have talked about length scales. We identify the longest wave length, $
\lambda_l $, with the forcing scale $ L_f $ and the breaking wave length, $
\lambda_b $, with the mean distance, $ d $, between the shocks. There are two
dynamically different regimes, the short wave regime, $ [\delta x, \;
\lambda_b] $, and the long wave regime, $ [\lambda_b, \, \lambda_l] $. We first
consider the latter, whose width scales $ \lambda_{l} /\lambda_b \sim
F_f^{-1/2} $. Identifying the amplitude of the breaking waves with the shock
amplitude, $ (\epsilon d)^{1/3} $, we find that the ratio between the amplitude of
the longest waves and the breaking waves scales as $ E^{1/2}/(\epsilon d)^{1/3}
\sim F_{f}^{-5/12} n^{\alpha/2} $, where we may substitute $ n $ with a
Reynolds number. Thus, for small $ F_f $ and large Reynolds
number the longest waves are not directly influenced by wave breaking. The
Reynolds number dependence of the amplitude ratio makes it unlikely that a
general similarity theory can be easily formulated for this range. Therefore,
predictions from weak turbulence theory \cite[for
example][]{ZakharovSagdeev1970} are not likely to hold for this range, although
it is not directly influenced by wave breaking. The exact values of the power
law exponents in the argument leading to this conclusion are, of course, not
important. As for the short wave regime, the width scales as $ \lambda_b
/\delta x \sim F_f^{1/2} $ for fixed $ \delta x $ or fixed Reynolds number. As
$ F_{f} \rightarrow 0 $ the range will eventually become so narrow so that
breaking will disappear. As argued in Appendix A, the condition $ \lambda_b \gg \delta x $ leads to a wave breaking condition involving both the Reynolds number,  $ Re  \equiv \epsilon^{1/3} L_f^{4/3} /  \nu $, and  the Froude number. If $ \lambda_b \propto F_f^{m} $ we obtain the condition $ Re F_f^{4m/3} \gg 1 $ and in the special case when $ m = 1/2 $, the condition is 
\begin{equation} \label{WC}
Re F_f^{2/3} \gg 1 \, .
\end{equation} 
Thus,  for a given Reynolds number, no matter how large, wave breaking will be absent provided that $ F_f $ is sufficiently small. On the other hand, if (\ref{WC}) not only is a necessary but also a sufficient condition,  wave breaking will be present for any given $ F_f $, no matter how small, provided that $ Re $ is sufficiently large. The condition (\ref{WC}) is analogous to the condition for the presence of turbulence in a stratified fluid, also involving both the Reynolds number and a Froude number \cite[]{Brethouwer2007}. The horizontal and vertical length scales of stratified turbulence scale as $ l_h \sim u^{3}/\epsilon $ and $ l_v \sim u/N $, respectively, where $ u $ is a characteristic horizontal velocity and $ N $ is the Brunt-V\"ais\"al\"a frequency. As stratification becomes stronger ($ N $ is increasing), $ l_v $ is becoming smaller.  However, in order for turbulence to be present $ l_v $ must be much larger than the Kolmogorov scale, $ \nu^{3/4}/\epsilon^{1/4} $, which leads to the condition $ Re F_h^2 \gg 1 $, where $ F_h = u/(l_h N) $ is a Froude number based on the horizontal length scale.

 The present study makes some advancement of the understanding of compressible turbulence. Previously, energy flux relations  have 
 been derived  for a compressible and barotropic fluid by \cite{Falkovich2010} and \cite{Galtier2011}.  Unlike (\ref{FluxT}) these flux relations are not exclusively expressed in terms of third order structure functions, {\em i. e.} third order moments of increments of flow variables. 
As is clear from our analysis, in the general case of a compressible and irrotational fluid the KE flux can be expressed in this way.  It seems to be more of a specific property that the APE flux also can be expressed in terms of a third order structure function, a property that the SW system shares with the Boussinesq system with constant Brunt-V\"ais\"al\"a frequency, in which case APE also is quadratic in buoyancy \cite[]{Augier2012}. 
In this context,  a comparison between our analysis and the analysis  by \cite{Galtier2011} of  isothermal compressible turbulence  is revealing. In the latter case, the pressure is linear in density, $ p = c_t^2 \rho $, where $ c_t $ is the isothermal speed of sound, and the internal energy per unit volume is  equal to $ c_t^2 \rho \ln (\rho/ \rho_0) $, where $ \rho_0 $ is an ambient reference density. Due to this  logarithmic dependence, the internal energy flux is not expressible in terms of a structure function. However, an equation system which is equivalent to the SW equations can be obtained by replacing the pressure term, $ \rho^{-1} \nabla p = \rho^{-1} c_t^{2} \nabla \rho $, in the isothermal momentum equation, by $ \rho_{0}^{-1} c_t^{2} \nabla \rho $, which may be motivated if density fluctuations are small. The expression for the internal energy per unit volume of the modified system is  $ c_t^2 \rho^2/2\rho_{0} $, which is equivalent to the SW potential energy. \footnote{To see how the logarithmic and quadratic expressions for the internal energy are 
 related, write $ \rho = \rho_0(1+\eta) $, expand in $ \eta $ and keep  terms up to second order: 
 $ c_t^2\rho \ln(\rho/\rho_0) \approx c_t^2 \rho_0(1+\eta)(\eta - \eta^2/2) \approx   c_t^2 \rho^2/2 \rho_0 - c_t^2 \rho_0/2 $. 
For small density fluctuations the two expressions only differ by a constant.} As a consequence of this quadratic dependence, the internal energy flux is expressible in terms of a structure function. It may be conjectured that the modified equation system is a good model for isothermal acoustic turbulence and that a flux relation which  equivalent to (\ref{FluxT}) also should hold for this case. In the three-dimensional case we then obtain
 \begin{equation}
\meane{ \delta (\rho u_L)    |\delta \uu|^2  }
+ c_t^2\meane{ \delta u_L(\delta \rho)^2  } /\rho_0 = - \frac{8}{3} \epsilon r \, ,\label{eq_Kolmo3}
\end{equation}
where  $ \epsilon $ here is the mean energy dissipation rate per unit volume.

While isothermal acoustic turbulence may have some interesting application for a radiating gas \cite[]{Stein1967}, standard text book acoustics is not isothermal but isentropic. The  condition of constant entropy puts us in a dilemma when we analyse an acoustic field as a forced-dissipative system, since dissipation inevitably is associated with entropy production.  According to arguments of \cite{KadomtsevPet1973}, an acoustic field is always dissipated by weak shocks if the Reynolds number is sufficiently large,  and this should be true in two as well as in three dimensions.  Given the numerical results of the present study such a hypothesis seems reasonable, since shock waves have similar stability properties 
in two and three dimensions \cite[]{ApazidisEliasson2018, LivertsApazidis2016}.  The shock wave hypothesis offers an elegant way out of the isentropic dilemma, since it implies that entropy production is concentrated at the shocks, while the rest of the field can be treated as isentropic. In a companion paper \cite[]{Lindborg2019}, scaling relations for such an acoustic field are derived, using the weak shock relations and the entropy equation. It turns out that these scaling relations are analogous to the relations for shallow water wave turbulence, with the  difference 
that the mean energy dissipation rate, $ \epsilon $, is replaced by $ \epsilon + \chi $, where $ \chi $ is a quantity associated with entropy production due to heat conduction. If  the Prandtl number is of the order of unity, $ \chi $ is of the same order as $ \epsilon $.  The shock amplitude scales as $ \Delta u \sim (\epsilon + \chi)^{1/3} d^{1/3} $, where $ d $ is the characteristic distance between the shocks and the third order longitudinal structure function scales as $ \langle \delta u_L^{3} \rangle = -C (\epsilon + \chi)r $, where $ C $ is a constant of the order of unity, which can be determined exactly in the weak shock limit.

Acoustic turbulence is an example of a wave breaking system where the concepts and methods we have developed to analyse SW water turbulence can be applied in a fruitful way.  We believe that there may be other such examples.

\vskip 0.5 cm

\noindent The authors would like to thank Gregory Falkovich and Sergey Nazaranko for  fruitful criticism of an early version of the manuscript and three anonymous reviewers for useful comments. Financial
support from the Swedish Research Council (VR) is gratefully
acknowledged. The simulations were performed
on resources provided by the Swedish National Infrastructure
for Computing (SNIC) at PDC.

\appendix

\section{Simulations with Laplacian diffusion}
%\label{app_comp}
On the request from the reviewers we have carried out a set of simulations using Laplacian diffusion, $ \nu \nabla^{2} $, to check that similar results are obtained in this case as when we use higher order diffusion $ \nu_8 \nabla^{8} $. 
A number of test simulations with resolution $ n = 960 $ and $ c= 10 $ was first carried out using different values of the viscosity, till we determined a value which was sufficiently large for the shocks to be resolved, but not unnecessarily large. 
This value was then used for all simulations with $ n = 960 $. For the higher resolved simulations the viscosity was then set as $ \nu \propto n^{-1} $, in accordance with the prediction  (\ref{Width}) for the shock width.
We define the Reynolds number as $ Re \equiv \epsilon^{1/3} L_f^{4/3}/\nu $, which is proportional to $ n $, since $ \nu \propto n^{-1} $.
The simulations are listed in table 2.

\begin{table}
\begin{center}

\label{Table2}

\begin{tabular}{lrrrrrrrrr}
\toprule
{} &  $n$ &  $c$ &  $\nu $ &  $\epsilon$ &  $\frac{k_{diss}}{k_f}$ &   $F_f$ &  $\; Re$ &  $ReF_f^{2/3} $ &  $t_{\max}$ \\
\midrule
WL1  &  960 &   10 &   0.0985 &       0.998 &                    22.6 &   0.16 &  68.0 &           20 &          25 \\
WL2  & 1920 &   10 &   0.0492 &       0.985 &                    45.3 &    0.16 &  136 &           40 &          25 \\
WL3  & 2880 &   10 &   0.0328 &       0.984 &                    67.9 &    0.16 &    204 &           60 &          25 \\
WL4  & 3840 &   10 &   0.0246 &       0.993 &                    90.5 &   0.16 &   273 &            80 &          25 \\
WL5  & 7680 &   10 &   0.0123 &       1.000 &                     181 &   0.16 &   545 &           160 &          25 \\
WL6  &  960 &   20 &   0.0985 &       0.994 &                    22.6 &  0.08 &  68.1&            13 &          25 \\
WL7  & 2880 &   20 &   0.0328 &       0.982 &                    67.9 &    0.08 &  204 &           38 &          25 \\
WL8  &  960 &   40 &   0.0985 &       0.985 &                    22.6 &    0.04 &  68.0 &           8.0 &          25 \\
WL9  & 2880 &   40 &   0.0328 &       0.981 &                    67.9 &    0.04 &  204 &            24 &          25 \\
WL10 &  960 &  100 &   0.0985 &       0.976 &                    22.6 &   0.016 &  67.5 &            4.3 &          25 \\
WL11 & 2880 &  100 &   0.0328 &       0.961 &                    67.9 &  0.016 &  202 &           13 &          25 \\
WL12 &  960 &  200 &   0.0985 &       0.988 &                    22.6 & 0.0080 &  67.8 &          2.7 &          25 \\
WL13 & 2880 &  200 &   0.0328 &       0.928 &                    67.9 & 0.0079 &  200 &           7.9&          25 \\
\bottomrule
\end{tabular}

\caption{Parameters for all the simulations executed with laplacian viscosity.
$ n $: number of nodes in each direction, $ c $: wave speed, $ \nu  $:  viscosity, $ \epsilon $: time
averaged mean energy dissipation in the stationary state, $ k_{d}/ k_f $: ratio
between dissipation wave number and forcing wave number, $F_f $: Froude
number, $Re$: Reynolds number, $ t_{max} $: end time of simulation.}
\end{center}
\end{table}

\begin{figure}
\centerline{\includegraphics[width=5.8in]{paper_04_shallow_water/Pyfig/fig_energy_lap}}
\caption{Left: Mean energy in the stationary state versus $ c $ for different
resolutions. The lines have slope $ 0.5 $, corresponding to $ E \propto c^{1/2}
$. Right: Normalised mean energy in the stationary state versus resolution $ n
$ for different $ c $.}
\label{MeanE2}
\end{figure}

\begin{figure}
\centerline{\includegraphics[width=8cm]{paper_04_shallow_water/Pyfig/fig_shock_sep_lap.pdf}}
\caption{Mean distance between the shocks as function of Froude number.  }
\label{fig_distance_lap}
\end{figure}

\begin{figure}
\centerline{\includegraphics[width= 8cm]{paper_04_shallow_water/Pyfig/fig_flatness_div_norm_nu.pdf}}
\caption{Flatness of divergence $ \nabla \cdot {\bf u} $ times viscosity $ \nu $, versus $ F_f $. Dotted straight line indicates $ F_f^{2/3} $. }
\label{FlatnessDiv}
\end{figure}

\begin{figure}
\centerline{\includegraphics[width=6.0in]{paper_04_shallow_water/Pyfig/fig_phys_fields_wave_lap}}
\caption{$ \nabla \cdot {\bf u} $ for four different runs. Upper left: WL1, $ Re = 68 $, $ F_f = 0.16 $; Upper right: WL4,  $ Re = 273 $, $ F_f = 0.16 $; Lower left: WL12, $ Re = 68 $, $ F_f = 0.008 $; Lower right: WL13,  $ Re = 200 $, $ F_f = 0.008 $.}
\label{Physical_lap}
\end{figure}

The dynamics of the runs with Laplacian diffusion were similar to the runs with higher order diffusion. In figure \ref{MeanE2} we see mean energy in the stationary state versus $ c $ to the left and versus $ n $ to the right. 
The variation of mean energy with $ c $ seems to be a little bit weaker for these runs as compared to the runs with hyper diffusion. However, the series of runs with $ n = 2280 $ (green crosses) is pretty close to approaching the $ c^{1/2} $-scaling.
The variation of mean energy with resolution, or Reynolds number, is quite similar to what is seen in the corresponding plot \ref{MeanE}. For each $ c $ there is a monotonic increase of mean energy with resolution, and the increase is becoming stronger as $ c $ is increased, with comparable but not equal values of the exponent $ \alpha $.

In figure \ref{fig_distance_lap} we see the mean distance between the shocks as a function of $ F_f $, together with a straight line indicating $ F_f^{1/2} $. There is a reasonable agreement between $ F_f ^{1/2} $ and each of the two series of simulations at $ Re \approx 70 $ (blue dots) and $ Re \approx 200$ (yellow triangles). However, there is a gap between the two series. This gap may be explained by the fact that the shock detection algorithm is much more sensitive to the threshold value which is used for $ \nabla \cdot {\bf u} $ in these simulations than in the simulations with hyperviscosity.  In particular, this is true in the lower Reynolds number runs. In figure \ref{FlatnessDiv} we see the flatness of the divergence, multiplied by $ \nu $, versus $ F_f $. According to (\ref{FD}) we should have $ F_{\partial u} \nu \propto  d^{4/3} $ and if $ d \propto Fr^{1/2} $, the points should collapse on a single line, $ F_f^{2/3} $.  The three runs at highest $ Re $ and the lowest value of $ F_f $ fall on top of each other in the upper right corner, supporting the scaling $ \nu^{-1} $ in the high Reynolds number limit. For the three highest $ F_f $ there is a reasonable agreement with $ F_f^{2/3} $, although the lower Reynolds number runs (blue dots) fall below the higher Reynolds number runs, which we interpret as a low Reynolds number effect. The points corresponding to the two lowest values of $ F_f $ fall below the prediction. This can be explained by the fact that the shocks are on the verge of fading away in these runs. Clearly, a necessary condition for shocks to be present is that the mean distance between them is much larger than their  width, that is
\begin{equation} \label{Condition} 
\frac{d}{\delta x} \gg  1\, .
\end{equation} 
Since $ F_{\partial u} \propto d/\delta x $, this means that the flatness of divergence must be much larger than unity. Substituting (\ref{Width}) into (\ref{Condition})  and assuming that (\ref{MeanDistance}) is valid we can formulate the condition as
\begin{equation}  \label{SCond}
Re F_f^{2/3} \gg 1 \, .
\end{equation} 
If this condition is not fulfilled the shocks will fade away.  In figure \ref{Physical_lap} we have plotted the divergence for four different runs with different $ Re $ and $ F_f $. It should be pointed out again, that the area of the computational domain is sixteen times larger than area of the outcuts shown in the figures.  The two upper plots are from simulations with  $ F_f = 0.16 $ but different $ Re $ ($ Re = 68 $ to the left and $ Re = 273 $ to the right). As can be seen, the shocks are wider in the lower Reynolds number run to the left than in the higher Reynolds number run to the right, as expected. The lower left plot is from a simulation with the same $ Re $ as in the upper left plot, but with a much smaller $ F_f $. In this simulation we have $ Re F_f^{2/3} = 2.7 $, so the condition (\ref{SCond}) is not fulfilled, and the shocks have almost faded away. The lower right plot is from a simulation with the same $ F_f $ as in the lower left plot, but with a higher $ Re $. Here, $ Re F_f^{2/3} = 7.8 $ and  the shocks are clearly discernible.

% \bibliographystyle{jfm}
% \bibliography{biblio}
%
% \end{document}

### 2019-04-25 paper_04_shallow_water/Latex/article.tex
Fix duplicate refs

@@
 Was \cite{YuanHamilton1994}  right  when they suggested that even a single layer model, such as the SW equations, could do the job?
-The original motivation of this study was to investigate this suggestion.
-A first set of simulations of the SW equations, containing both balanced and gravity wave motions, showed some encouraging results.
-Indeed, gravity waves were transferring energy towards small scales.
-However, after a while it became clear that the cascade was not as strong as in the GCM simulation.
-Moreover, in all the SW simulations we observed shocks with an associated $ k^{-2} $-spectrum rather than a $ k^{-5/3} $-spectrum.
-These observations lead us to dismiss the SW equations as a model for the atmospheric downscale cascade.
-Based on the SW equations we developed another two-dimensional toy model  \cite[]{LindborgMohanan} that is absent of shocks and can reproduce the energy cascade in the GCM quite well, including a $ k^{-5/3}  $-spectrum.
+The original motivation of this study was to investigate this suggestion. A first set of
+simulations of the SW equations, containing both balanced and gravity wave motions,
+showed some encouraging results. Indeed, gravity waves were transferring energy towards
+small scales. However, after a while it became clear that the cascade was not as strong
+as in the GCM simulation.  Moreover, in all the SW simulations we observed shocks with
+an associated $ k^{-2} $-spectrum rather than a $ k^{-5/3} $-spectrum. These
+observations lead us to dismiss the SW equations as a model for the atmospheric
+downscale cascade. Based on the SW equations we developed another two-dimensional toy
+model  \cite[]{LindborgMohanan2017} that is absent of shocks and can reproduce the
+energy cascade in the GCM quite well, including a $ k^{-5/3}  $-spectrum.
 
@@
 time step method which maximizes the time step over a standard
-Courant-Friedrichs-Lewy condition \cite[]{Lundbladh1999, AugierChomazBillant2012}.
+Courant-Friedrichs-Lewy condition \cite{Lundbladh1999, AugierChomazBillant2012}.
 %

### 2019-04-28 imgs/cascade/plot.py
Stop Kraichnan description

@@ -1,0 +1,92 @@
+# coding: utf-8
+import matplotlib.pyplot as plt
+import numpy as np
+
+plt.style.use("ggplot")
+plt.rc("text", usetex=True)
+
+fig, axes = plt.subplots(1, 2, dpi=200, figsize=(7, 3))
+ax1, ax2 = axes.ravel()
+
+km = 1.5
+k1 = np.logspace(0, km, dtype=float, endpoint=True)
+k2 = np.logspace(km, 4, dtype=float, endpoint=True)
+c = "k-"
+km = 10 ** km
+coef = km ** (-6.7 / 5)
+
+print("coef =", coef)
+ax1.set_title("Energy spectra predicted by Kraichnan's theory", fontsize=9)
+ax1.loglog(k1, k1 ** -(5 / 3) * coef, c)
+ax1.loglog(k2, k2 ** (-3), c)
+
+ax2.set_title("Observed atmospheric energy spectra", fontsize=9)
+ax2.loglog(k2, k2 ** (-5 / 3) * coef, c)
+ax2.loglog(k1, k1 ** -3, c)
+
+ax1.annotate(
+    r"Forcing, $P=\epsilon+\eta$",
+    (km, km ** -3),
+    (km, km ** -1.5),
+    arrowprops=dict(
+        #  arrowstyle="fancy",
+        shrink=0.05,
+        # posB=(km, km ** -1.5)
+    ),
+    ha="left",
+    fontsize=8,
+)
+ax1.text(
+    0.1*km, (0.8*km)**(-5/3),
+    r"$\epsilon^{2/3}k^{-5/3}$",
+    fontdict=dict(size=8),
+)
+ax1.text(
+    100*km, (80*km)**-3,
+    r"$\eta^{2/3}k^{-3}$",
+    fontdict=dict(size=8),
+)
+ax2.text(
+    0.1*km, (0.08*km)**(-3),
+    r"$k^{-3}$",
+    fontdict=dict(size=8),
+)
+ax2.text(
+    100*km, (800*km)**-(5/3),
+    r"$k^{-5/3}$",
+    fontdict=dict(size=8),
+)
+
+
+def arrow(ax, x, y, arrowroot=0.1, exp=-3):
+    return ax.annotate(
+        "",
+        (x*km, (y*km)**exp),
+        (x*arrowroot*km, (y*arrowroot*km)**exp),
+        arrowprops=dict(
+            arrowstyle='fancy',
[…]

### 2019-05-15 chapter_00_open_science.tex
New command fnref to add footnote and cite them

@@
 It is normal that Mathworks, the company which makes money with Matlab, spreads
-\href{https://www.mathworks.com/products/matlab/matlab-vs-python.html}{arguments
+\fnref{https://www.mathworks.com/products/matlab/matlab-vs-python.html}{arguments
 against Python and its scientific ecosystem}.  Some of them are valid, for
@@
 application. For example, Volkswagen group (Volkswagen, Audi, Seat, Porsche,
-Skoda, ...) has used OpenFOAM since
-2006\footnote{\url{https://www.nas.nasa.gov/assets/pdf/ams/%
-2014/AMS_20141202_Othmer.pdf}}.
+Skoda, ...) \fnref{%
+  https://www.nas.nasa.gov/assets/pdf/ams/2014/AMS_20141202_Othmer.pdf}{%
+  has used OpenFOAM since 2006}.
 %
 EDF, the main French electric utility company has also a interesting
-\href{http://linuxfr.org/news/strat\%C3\%A9gie-open-source-\%C3\%A0-edf-rd}{%
+\fnref{http://linuxfr.org/news/strat\%C3\%A9gie-open-source-\%C3\%A0-edf-rd}{%
 open-source strategy}. It has produced many programs and in particular the CFD
-solver \href{https://www.code-saturne.org}{Code\_Saturn}.
-%
-EDF is also involved in the development of
-\href{https://www.paraview.org/}{Paraview}\footnote{\url{https://blog.kitware.com/introducing-some-new-paraview-5-0-features/}}.
+solver \fnref{https://www.code-saturne.org}{Code\_Saturn}.
+%
+\fnref{https://blog.kitware.com/introducing-some-new-paraview-5-0-features/}{%
+EDF is also involved in the development of \fnref{https://www.paraview.org/}{Paraview}}.
 
 Finally, scientists have started to use Python to study fluid mechanics,
-especially through CFD codes (\href{http://dedalus-project.org/}{Dedalus},
-\href{https://github.com/spectralDNS}{SpectralDNS},
-\href{https://github.com/mikaem/Oasis}{Oasis}, \href{http://pyfr.org/}{PyFR}
-and \href{https://fenicsproject.org/}{FEniCS}) or through data analysis
-(\href{http://www.openptv.net/}{OpenPTV},
-\href{https://github.com/jr7/pypiv}{PyPIV}).
+especially through CFD codes (\fnref{http://dedalus-project.org/}{Dedalus},
+\fnref{https://github.com/spectralDNS}{SpectralDNS},
+\fnref{https://github.com/mikaem/Oasis}{Oasis}, \href{http://pyfr.org/}{PyFR}
+and \fnref{https://fenicsproject.org/}{FEniCS}) or through data analysis
+(\fnref{http://www.openptv.net/}{OpenPTV},
+\fnref{https://github.com/jr7/pypiv}{PyPIV}).
 %

### 2019-05-19 chapter_00_open_science.tex
Ch 0: Rewrite intro

@@
 \section{Introduction}
-
-% \subsection{Science, software, open-source and the computer revolution}
 
@@
 
-This chapter presents FluidDyn, a project and an ecosystem of packages to
-facilitate open-science and open-source development in the field of fluid
-mechanics.
-%
-The chapter is organized into several parts to present the context and to
-describe our motivations and the concept behind the project FluidDyn as
-follows:
-
-\begin{itemize}
-\item A scientific project based on the collaboration through open-source, such
-as FluidDyn, is only possible thanks to the recent developments in methods and
-tools for open-source software engineering. We start by presenting these
-developments and discuss the implications for scientists.
-
-\item FluidDyn is based on the language Python. Incidentally, it is the defining
-characteristics of Python which make the eventual success of such a
-project realizable.
+In order to attain the research goals in this thesis, a path off the beaten
+track was pursued -- i.e. to create a handful of easy to maintain, rigorously
+tested, and most-importantly, open-source scientific packages which does not
+compromise on performance.
+%
+This resulted in the creation of FluidDyn, a project to facilitate open-science
+in the field of fluid mechanics.
+%
+The focus of the project was to develop a set of packages to implement a
+framework for simulations \citep{fluidsim}, experiments
+\citep{augier_fluidimage_2016,augier_fluidlab_nodate}, Fast Fourier Transforms
+\citep{fluidfft} -- all depending on a base package to reuse certain utilities
+and data structures \citep{fluiddyn}.
+%
+The project owes a lot to the recent developments in methods and tools for
+open-source software engineering.
+%
+Several good and modern coding practices were adopted while developing these
+packages.
+%
+An understanding of the design choices and the workflow involved would be of
+general interest to the reader.
+
+FluidDyn packages are primarily written in Python language. Incidentally, it is
+the defining characteristics of Python which make the eventual success of such
+a project realizable.
 %
@@
 %
-Python in fluid mechanics is growing fast, but is yet to reach widespread
-adoption as the language of choice. Therefore, it is useful to present its
-characteristics for readers who are unacquainted with its nuances.
-
-\item We will discuss about the possible contradiction between productivity of
-individuals and productivity at the community level.
-
-\item We will establish a picture of software engineering in fluid mechanics
-and present some of the main tools commonly used.
-
-\item The introduction finally ends with a summary of the motivations for the
-project FluidDyn and an overview of the packages used\footnote{%
-  We use FluidDyn (with capital letters) to name the project and \fluiddyn for
-the base package.}.
-\end{itemize}
-
-\section{A modern approach to computational sciences}
+The popularity of the language in fluid mechanics is growing, but is yet
+to reach widespread adoption as the language of choice.
+
+We shall now discuss these topics in greater detail. The chapter is organized
+into two sections. The first section focusses on the technical aspects, with
+particular focus on emerging developments in programming languages, tools and
+services suited for open-source software development. We shall present Python's
+strengths and weaknesses as a language for scientific computing.
+%
+The section ends with a note on the capabilities of FluidDyn
+packages\footnote{We use FluidDyn (with capital letters) to name the project
+and \fluiddyn for the base package.}.
+%
+In the second section a comparison of various software development
+methodologies are made. Here, we will also discuss about the possible
+contradiction between productivity of individuals and productivity at the
+community level.
+%
+% The chapter finally ends with a summary of the motivations for the project
+% FluidDyn and an overview of the packages used.
+
+\section{Science, software, open-source and the computer revolution}
 

### 2019-05-21 chapter_00_open_science.tex
Revamp until just before section 1.3 Python, a programming

@@
 
-\section{Science, software, open-source and the computer revolution}
+% \section{Science, software, open-source and the computer revolution}
+\section{Open-source software and open-science}
 
 Computational sciences, and especially computational fluid dynamics (CFD) as a
-discipline has flourished a lot in the latter half of the twentieth century.
-Incidentally the origin of this domain can be traced back to the ideas
-in \citet{richardson_weather_1922} to achieve numerical weather prediction.
-With the arrival of first electronic general-purpose computer, ENIAC, the first
-simulations were made using a simple barotropic model to make 24 hour
-forecasts \citep{lynch_richardson_2010}. In those early days when the use and
-availability of computers were limited, such methods were neither widespread
-nor feasible. Scientific investigations were dominated by theoretical and
+discipline, have flourished a lot in the latter half of the twentieth century.
+Incidentally the origin of this discipline can be traced back to the ideas in
+\citet{richardson_weather_1922} to achieve numerical weather prediction.
+Following the arrival of first electronic general-purpose computer, ENIAC, the
+first simulations were made using a simple barotropic model to make 24 hour
+forecasts \citep{lynch_richardson_2010}. In those early days when the
+availability of computers was limited, such methods were neither widespread nor
+feasible. Scientific investigations were dominated by theoretical and
 experimental methods.
-We have come a long way since then and today one can say that, computational
-methods now complement theoretical and experimental studies, and are equally
-important.
-Thus to recreate a particular scientific study, the mathematical representation
-or even an algorithm leaves a lot to be desired. With the connectivity
-provided by the world wide web it is possible to do more and reinvent the way
-we perform sciences.
+%
+We have come a long way since then and today computational methods now
+complement theoretical and experimental studies, and are equally important.
+%
+% Thus, in a typical scientific study, there is significant effort involved in
+% translating the mathematical representation to a working research software.
+%
+With the connectivity provided by the world wide web it is possible to achieve
+more and reinvent the way we perform sciences.
 
@@
 %
-\emph{Knowledge} has been traditionally disseminated in the form of courses and
-books from universities, and to the wider scientific community through research
-articles, conference proceedings and workshops.
-
-While this is nothing new, knowledge has become more accessible to the public
+Traditionally, \emph{knowledge} has been disseminated in the form of courses
+and books from universities, and to the wider scientific community through
+research articles, conference proceedings and workshops.
+%
+In the last decade, knowledge has become more accessible to the public
 through public-domain and Creative Commons (or similar) licensed course
@@
 } either through systematic
-reforms in academia on how researchers are assessed on their productivity or
-through regulations and offering incentives, such as waiving the application
-processing charge for publishing \citep{nosek_promoting_2015}.
-
-As mentioned earlier there is a lot of work involved in condensing and
-implementing this knowledge in the form of working applications.
+reforms in academia on how researchers' productivity are assessed or through
+regulations and offering incentives, such as waiving the application processing
+charge for publishing \citep{nosek_promoting_2015}.
+
+A lot of work is required to condense and implement this knowledge in the
+form of working applications.
 %
 The \emph{know-how} can be made accessible and open to scrutiny by making
-research software open-source with an appropriate license.  To make the codes
-usable it has to be complemented by documentation which typically includes
-tutorials, examples and a detailed commentary of different components of the code.
-%
-Furthermore, to inspire confidence in and to ensure reliability of the code
-throughout the process of software development, it needs unit-tests with
-which a code is continuously deployed, built and tested --- a process known as
+research software open-source with an appropriate license\footnote{For a
+  detailed comparison of licenses, visit either
+  \url{https://choosealicense.com} or
+  \url{https://www.gnu.org/licenses/license-list.html}.
+}.
+%
+To make codes usable they have to be complemented by documentation which
+typically includes tutorials, examples and a detailed commentary of different
+components.
+%
+Furthermore, to inspire confidence in and to ensure reliability of a code
+throughout the process of software development, it needs unit-tests through
+which a code is continuously built, tested and deployed --- a process known as
 Continuous Integration (CI), which is described in detail later in this
-chapter. Depending on how much matured a software is, these can be easy or
+chapter. Depending on how matured a software is CI can be easy or
 cumbersome to implement, but nevertheless, the result is a cleaner and reliable
@@
 workflow (i.e. scripts to run the codes and to postprocess the data) when made
-open in the form of datasets can ensure that research would be reproducible and
+open in the form of datasets can ensure that research is reproducible and
 preserved in archives for several years to come \citep{gewin_data_2016}. Open
@@
 
-In the present work, through FluidDyn project, we have managed to implement all
-the aspects of making the know-how open through open-source software. We shall
-now take a look at the tools which have made it possible.
+In FluidDyn project, we have strived to implement all the aspects of making the
+know-how open through open-source software. We shall now take a look at the
+tools which have made it possible.
 
@@
 
-\paragraphbf{Free and Open-Source Software (FOSS)} The term \emph{free} in
-FOSS is a misnomer, as it actually stands for \emph{freedom} (to use, modify
-and distribute). The FOSS movement has dramatically decreased the cost of using
-computers, and is evident from the widespread use of GNU/Linux systems in
-desktops, computing clusters and web-servers in academia and beyond.
-%
-The beginnings of FOSS culture can be traced back to the success for scientific
-publication of the free type-setting standards \TeX\ (1977) and \LaTeX\ (1985).
+\subsubsection{Free and Open-Source Software}
+
+Open-source software development that we know today has been built on top of
+the firm foundations of the Free and Open-Source Software (FOSS) movement. The
+term \emph{free} in FOSS is a misnomer, as it actually stands for
+\emph{freedom} (to use, modify and distribute). The FOSS movement has
+dramatically decreased the cost of using computers, which is evident from the
+widespread use of GNU/Linux systems in desktops, computing clusters and
+web-servers in academia and beyond.
+%
+The FOSS culture can be traced back to the success for scientific publication
+of the free type-setting standards \TeX\ (1977) and \LaTeX\ (1985).
 %
@@
 in 1983 by \href{https://en.wikipedia.org/wiki/Richard_Stallman}{Richard
-Stallman}, to create a Unix-like computer operating system composed entirely of
-free software\footnote{For completeness, see also the work done on
-\href{https://www.levenez.com/unix/}{other Unix operating systems,
-for example BSD}.}.
+  Stallman}, to create a Unix-like computer operating system composed entirely
+  of free software\footnote{For completeness, see also the work done on
+  \href{https://www.levenez.com/unix/}{other Unix operating systems, for
+example BSD}.}.
 %
 GNU is today known for its compiler collection (GCC) and a multitude of tools
-which when combined with the Linux kernel (created by
+which combined with the Linux kernel (created by
 \href{https://en.wikipedia.org/wiki/Linus_Torvalds}{Linus Torvalds} in 1991)
-forms the GNU/Linux operating systems that we are familiar today.
-%
-FOSS movement has had huge successes in many frontiers --- Apache could be
-termed as the ``first killer-app of Linux'' and now with an ever-increasing
-suite of softwares including Firefox, LibreOffice, Gimp as solid alternatives
-to proprietary offerings.
+forms the GNU/Linux operating system.
+% and its several distributions (Debian, Red Hat, Slackware, Arch Linux and
+% Gentoo to name a few)
+that we are familiar with today.
+%
+% FOSS movement has had huge successes in many frontiers --- Apache could be
+% termed as the ``first killer-app of Linux'' and now with an ever-increasing
+% suite of softwares including Firefox, LibreOffice, Gimp as solid alternatives
+% to proprietary offerings.
 %
 Linux has become the most widely used kernel, being deployed on servers,
-personal computers, embedded devices, and also smart phones (with Android).
+personal computers, embedded devices, and also smart phones (with Android). The
+success of the Linux kernel project was attributed to the \emph{bazaar} model of
+software development in \citet{raymond_cathedral_1999}, wherein a code can stay
+structured despite releasing frequently and delegating tasks to community with
+myriad agendas.
 
@@
 volunteers, towards an organized system with participation from industries,
-non-profit organizations and government institutions. This could be regarded as
-the second-age of FOSS movement~\citep{fitzgerald_transformation_2006}.
+non-profit organizations and government institutions.
+%
+Such a model results in a win-win scenario -- users benefit from transparency and
+ability to tinker, and the organizations profit with more contributions.
+%
+The FOSS movement has thus entered into a second era
+\citep{fitzgerald_transformation_2006}.
 
@@
 
-\paragraphbf{Distributed Version Control Systems (DVCS) and web-based source
-development platforms} Collaboration was achieved in the early years of FOSS
-development through emailing ``patches'', and centralized VCS near the turn of
-the 21\textsuperscript{st} century.  The collective collaboration on
-development is today streamlined with the emergence of DVCS (especially, Git
-and Mercurial) and web-based source development platforms like GitHub,
-Bitbucket and more recently GitLab.
-%
-These tools are designed to work on any kind of text files, not just codes, so
-they could even be utilized to write scientific documents with multiple authors.
-%
-These tools are so useful for work, that using DVCS in science shall reap
-benefits in terms of productivity in the long run~\cite{wilson_best_2014}.
-
+\subsubsection{Source code management and packaging}
+
+Collaboration was achieved in the early years of FOSS development through
+emailing ``patches'', and centralized Version Control Systems (VCS) near the
+turn of the 21\textsuperscript{st} century.  The collective collaboration on
+development is today streamlined with the emergence of Distributed Version
+Control Systems (DVCS), especially, Git and Mercurial). In web-based source
+development platforms like GitHub, Bitbucket and GitLab DVCS repositories are
+enhanced with code-review tools, which typically consists of a issue tracker
+and pull-request mechanism.
+%
 Git has become the \textit{de facto} standard for source management. However,
@@
 extension \href{http://hg-git.github.io/}{hg-git}.}.
-
-\paragraphbf{Package management and third-party software repositories} Python
-as a medium for scientific computing owes its success to a rich ecosystem of
-third-party packages.  Smooth deployment of these packages requires a package
-manager and a supported repository. A credible hub to upload source codes as
-compressed archives or as pre-compiled binary packages was therefore
-indispensable. The use of repositories started with the conception of CPAN and
-CRAN repositories for the languages Perl and R in 1993 --- based on the
-Comprehensive TeX Archive Network (CTAN) model for \TeX\ packages.
-
-Today, \href{https://pypi.org}{Python Package Index (PyPI)} and
+%
+Version control tools are designed to work on any kind of text files, not just
+codes, so they could even be utilized to write scientific documents with
+multiple authors.
+%
+The use of DVCS in science will lead to long term productivity benefits
+\citep{wilson_best_2014}.
+
+%\paragraphbf{Package management and software repositories} 
+Python as a medium for scientific computing owes its success to a rich
+ecosystem of third-party packages.  Smooth deployment of these packages
+requires a package manager and a supported repository. A credible hub to upload
+source codes as compressed archives or as pre-compiled binary packages was
+therefore indispensable. The use of repositories started with the conception of
+CPAN and CRAN repositories for the languages Perl and R in 1993 --- based on
+the Comprehensive TeX Archive Network (CTAN) model for \TeX\ packages. 
+%
+Packaging process depends on the language used in a project, however some good
+practices are universal, such as supplying documentation (a README file at the
+bare minimum), keeping installation paths flexible and relying on official
+build tools \citep{taschuk_ten_2017}.
+%
+For Python, \href{https://pypi.org}{Python Package Index (PyPI)} and
 \href{https://anaconda.org}{Anaconda Cloud} are major repositories for
@@
 % list-of-top-repositories-by-programming-language}{https://stackoverflow.com/%
-% questions/1693529/list-of-top-repositories-by-programming-language}}
-% have similar repositories, with exceptions such as Fortran and Matlab.
-Specialized package managers have also emerged to install scientific
-libraries and compilers such as \pack{easybuild} and \pack{spack}.
-
-\paragraphbf{Continuous integration (CI)} The need for strong reliability of
-programs with increasing complexity has forced developers to use continuous
-integration.
-%
-In CI, firstly the code is \emph{built} against a known, repeatable emulation
-environment on a server. Thereafter, a set of \emph{unit tests} are run for each
-commit (modification of the code) to check that no bugs have been introduced. It
-is also important to analyse what portion of the source code is covered by these
-unit tests --- a method referred to as \emph{code coverage}. In FluidDyn project
-we stop at this level, but it is possible to further enhance it by adding code
-style checks by running linters alongside unit tests; and also to add
-\emph{continuous delivery} of documentation and package releases.
-
+% questions/1693529/list-of-top-repositories-by-programming-language}} have
+% similar repositories, with exceptions such as Fortran and Matlab.
+Specialized package managers have also emerged to install scientific libraries
+and compilers such as \pack{easybuild} and \pack{spack}.
+
+\subsubsection{Continuous Integration and documentation} The need for
+strong reliability of packages with increasing complexity has forced developers
+to use Continuous Integration (CI).
+%
+When the code depends on a wide variety of external dependencies, CI is useful
+to avoid \fnref{https://en.wikipedia.org/wiki/Software\_rot}{software rot},
+i.e.\ to avoid obsolescence with newer versions of the compiler or its
+dependencies. CI is specifically important for dynamical languages and
+open-source code since it allows open-source projects to welcome contributions
+from other developers while maintaining a high degree of reliability.
+%
+Firstly the code is \emph{built} against a known, repeatable emulation
+environment on a server. Thereafter, a set of \emph{unit tests} are run for
+each commit (modification of the code) to check that no bugs have been
+introduced. It is also important to analyse what portion of the source code is
+covered by these unit tests --- a method referred to as \emph{code coverage}.
+%
 Almost all CI platforms are free for open-source projects --- Travis CI,
-CircleCI and Appveyor % (sic) av: I have double checked the official names
-to name a few. Bitbucket and GitLab have built-in CI
+CircleCI and Appveyor to name a few. Bitbucket and GitLab have built-in CI
 implementations. Websites such as Codecov and Coveralls help to chart and
-analyse the code coverage history, viewable by anyone.
-
-When the code depends on a wide variety of external dependencies, CI is useful
-to avoid \href{https://en.wikipedia.org/wiki/Software\_rot}{software rot},
-i.e.\ to avoid becoming obsolete with newer versions of the compiler or its
-dependencies.  CI is specifically important for dynamical
-languages and open-source code since it allows open-source projects to welcome
-participation from many people while maintaining a high degree of reliability.
-
-\paragraphbf{Websites to share knowledge} Mutual aid and knowledge sharing
-have been important aspects of the web since its inception and these practices
-are now of foremost importance for many human activities.
-%
-In particular, learning programming and computer science is now greatly aided
-by websites like the community driven encyclopedia
-\href{https://www.wikipedia.org/}{Wikipedia} or the community driven forum
-\href{https://stackoverflow.com/}{Stack Overflow}, where people write thousands
-of questions and answers on programming each week.
-IRC protocol has continued to sustain the test of time. Instant messaging
-channels such as \codeinline{\#python} on Freenode facilitate interactive
-discussions and a form of community-level support system --- a role which used
-to be filled by mailing lists alone in the past. Alternatives such as Riot
-(Matrix protocol), Gitter, and Slack have gained traction offering different
-integrations solutions on top of instant messaging.
-%
-Through instant messaging a geographically separated team, as is the case with
-FluidDyn project, can achieve near real-time collaboration without the hassle
-of setting up video conferences or meetings.
-
-\paragraphbf{Automatic web documentations built on servers} Documentation
-generators such as Sphinx (for Python) and Doxygen (for many languages
-including C, C++, Fortran and Python) ``automagically'' convert in-line
-comments into documentations as PDF, HTML and more. Read the Docs has in the
-recent years become the main host for building and publishing documentation on
-servers as web pages, thus easing the process for developers and considerably
-improving access to information.
+analyse the code coverage history, viewable by anyone. To use these services
+one simply has to write a simple file in YAML format and authenticate the
+respective services with ``read'' permissions.
+
+New users and developers often rely on good documentation and examples to start
+working with an unfamiliar piece of software. Documentation generators such as
+Sphinx (for Python) and Doxygen (for many languages including C, C++, Fortran
+and Python) parses in-line comments into documentations as PDF, HTML and
+more \citep{lee_ten_2018}. In principle, the can be executed online and deployed
+using any CI. ``Read the Docs'' has, in the recent years, become the hub for
+building and hosting documentation online, thus facilitating the process for
+developers and considerably improving access to information for users.
+%
+In the FluidDyn project we this stop at automating tests and documentation of the
+latest version, but it is possible to further enhance it by adding code style
+checks by running ``linters'' alongside unit tests, \emph{continuous
+delivery} of package releases and preserving older versions of
+documentation.
+
+\subsubsection{Communication}
+
+% Mutual aid and knowledge sharing have been important aspects of the web since
+% its inception and these practices are now of foremost importance for many
+% human activities.
+%
+%In particular, 
+Learning programming and computer science is now greatly aided by websites like
+the community driven encyclopedia \href{https://www.wikipedia.org/}{Wikipedia}
+or the community driven forum \href{https://stackoverflow.com/}{Stack
+Overflow}, where people write thousands of questions and answers on programming
+each week. IRC protocol has continued to sustain the test of time. Instant
+messaging channels such as \codeinline{\#python} on Freenode facilitate
+interactive discussions and a form of community-level support system --- a role
+which used to be filled by mailing lists alone in the past. Alternatives such
+as Riot (Matrix protocol), Gitter, and Slack have gained popularity offering
+different integrations solutions on top of instant messaging.
+%
+Through instant messaging a geographically separated team, as in the case of
+the FluidDyn project, can achieve near real-time collaboration without having
+to arrange video conferences or meetings.
+
+%\paragraphbf{Automatic web documentations built on servers}
 
@@
 
-We have presented new methods and tools used in software development.  Taken
-all together, they greatly improve our efficiency for collective work and open
-possibilities of interactions that were unthinkable only few years ago.
+Thus, there is a number of new methods and tools in software development.
+Taken all together, they greatly improve our collective efficiency and
+open possibilities of interactions that were unthinkable a few years ago.
 %
 For example, the success of community-driven software like \pack{astropy} (a
-subject-area research library with \href{https://github.com/astropy/astropy}{more
-than 240 contributors}) and \pack{scikit-learn} (a machine learning toolkit with
-\href{https://github.com/scikit-learn/scikit-learn}{more than 1000 contributors})
-would not have been possible without these new software development methods.
+subject-area research library with
+\href{https://github.com/astropy/astropy}{more than 240 contributors}) and
+\pack{scikit-learn} (a machine learning toolkit with
+\href{https://github.com/scikit-learn/scikit-learn}{more than 1000
+contributors}) would not have been possible without these new software
+development methods.
 %
@@
 As scientists, we can wonder how to fully exploit these new possibilities to do
-science.  The project FluidDyn is an attempt to help the fluid dynamics community
-to advance in this direction.
+science.  The project FluidDyn is an attempt to help the fluid dynamics
+community to advance in this direction.
+
+% TODO:
 

### 2019-05-22 chapter_00_open_science.tex
Reorder sentences in package management section, watchdog make directive

@@
 
-\subsubsection{Source code management and packaging}
+\subsubsection{Source code management}
 
@@
 
-%\paragraphbf{Package management and software repositories} 
+\subsubsection{Package management and software repositories}
+
+Packaging process depends on the language used in a project, however some good
+practices are universal, such as licensing, supplying documentation (a README
+file at the bare minimum), keeping installation paths flexible and relying on
+official build tools \citep{taschuk_ten_2017}.
+%
+A credible repository to packages as compressed archives or as pre-compiled binary
+packages is indispensable to promote code reuse and reproducibility of results.
+%
+The use of repositories started with the conception of
+CPAN and CRAN repositories for the languages Perl and R in 1993 --- based on
+the Comprehensive \TeX\ Archive Network (CTAN) model for \TeX\ packages.
+
 Python as a medium for scientific computing owes its success to a rich
-ecosystem of third-party packages.  Smooth deployment of these packages
-requires a package manager and a supported repository. A credible hub to upload
-source codes as compressed archives or as pre-compiled binary packages was
-therefore indispensable. The use of repositories started with the conception of
-CPAN and CRAN repositories for the languages Perl and R in 1993 --- based on
-the Comprehensive TeX Archive Network (CTAN) model for \TeX\ packages. 
-%
-Packaging process depends on the language used in a project, however some good
-practices are universal, such as supplying documentation (a README file at the
-bare minimum), keeping installation paths flexible and relying on official
-build tools \citep{taschuk_ten_2017}.
-%
-For Python, \href{https://pypi.org}{Python Package Index (PyPI)} and
-\href{https://anaconda.org}{Anaconda Cloud} are major repositories for
-delivering and downloading Python packages, assisted by package managers
-\pack{pip} and \pack{conda} respectively.
+ecosystem of third-party packages. \href{https://pypi.org}{Python Package
+Index (PyPI)} and \href{https://anaconda.org}{Anaconda Cloud} are major
+repositories for delivering and downloading Python packages, assisted by
+package managers \pack{pip} and \pack{conda} respectively. The
+\fnref{https://packaging.python.org}{``Python Packaging User Guide''} is an
+excellent guide to install, package and distribute projects.
 % Almost all major
@@
 % similar repositories, with exceptions such as Fortran and Matlab.
-Specialized package managers have also emerged to install scientific libraries
-and compilers such as \pack{easybuild} and \pack{spack}.
+Specialized package managers, such as \pack{easybuild} and \pack{spack}, have
+also emerged to install scientific libraries and compilers.
 
@@
 implementations. Websites such as Codecov and Coveralls help to chart and
-analyse the code coverage history, viewable by anyone. To use these services
+analyse the code coverage history, viewable by anyone. To use these services,
 one simply has to write a simple file in YAML format and authenticate the
@@
 Sphinx (for Python) and Doxygen (for many languages including C, C++, Fortran
-and Python) parses in-line comments into documentations as PDF, HTML and
-more \citep{lee_ten_2018}. In principle, the can be executed online and deployed
-using any CI. ``Read the Docs'' has, in the recent years, become the hub for
-building and hosting documentation online, thus facilitating the process for
-developers and considerably improving access to information for users.
-%
-In the FluidDyn project we this stop at automating tests and documentation of the
-latest version, but it is possible to further enhance it by adding code style
-checks by running ``linters'' alongside unit tests, \emph{continuous
-delivery} of package releases and preserving older versions of
-documentation.
+and Python) parses in-line comments into documentations as PDF, HTML and more
+\citep{lee_ten_2018}. The documentation generation can be automated online and
+deployed using any CI. ``Read the Docs'' has, in the recent years, become the
+hub for building and hosting documentation online, thus facilitating the
+process for developers and considerably improving access to information for
+users.
+%
+In the FluidDyn project we this stop at automating tests and documentation of
+the latest version, but it is possible to further exploit the CI by adding code
+style checks by running ``linters'' alongside unit tests, \emph{continuous
+delivery} of package releases and preserving older versions of documentation.
 
@@
 Learning programming and computer science is now greatly aided by websites like
-the community driven encyclopedia \href{https://www.wikipedia.org/}{Wikipedia}
+the community driven encyclopaedia \href{https://www.wikipedia.org/}{Wikipedia}
 or the community driven forum \href{https://stackoverflow.com/}{Stack

### 2019-05-22 chapter_00_open_science.tex
Bullets in python

@@
 We shall now discuss these topics in greater detail. The chapter is organized
-into two sections. The first section focusses on the technical aspects, with
+into three sections. The first section focusses on the technical aspects, with
 particular focus on emerging developments in programming languages, tools and
-services suited for open-source software development. We shall present Python's
-strengths and weaknesses as a language for scientific computing.
+services suited for open-source software development. In the next section, we
+present Python's strengths and weaknesses as a language for scientific
+computing.
 %
@@
 %
-In the second section a comparison of various software development
-methodologies are made. Here, we will also discuss about the possible
-contradiction between productivity of individuals and productivity at the
-community level.
+The last section compares various software development methodologies. Here, we
+will also discuss about the possible contradiction between productivity of
+individuals and productivity at the community level.
 %
@@
 
-\subsubsection{Communication}
+\subsubsection{Knowledge hubs and communication}
 
@@
 
-%\paragraphbf{Automatic web documentations built on servers}
-
-% We need transitions and some words at the end of the subsections to help the
-% readers to understand why we present what we present and what is the link
-% with the FluidDyn project...
-
 Thus, there is a number of new methods and tools in software development.
@@
 %
-% pa: done: we need to mention in this paragraph existing community-driven
-% subject-area research libraries (see the file
-% ./communitydriven_research_soft.md).
-%
 As scientists, we can wonder how to fully exploit these new possibilities to do
@@
 community to advance in this direction.
-
-% TODO:
 
@@
 
-Python community has thrived remarkably in the last few years, evident from the
-fact that by the end of 2017, Python became the second most popular language in
-terms of pull-requests in GitHub\footnote{See GitHub's annual report titled
-\href{https://octoverse.github.com/}{ ``The State of the Octoverse 2017''}.} and
-questions on StackOverflow\footnote{See
+The Python community has thrived remarkably in the last few years, which is
+evident from the fact that by the end of 2017, Python became the second most
+popular language in terms of pull-requests in GitHub\footnote{See GitHub's
+	annual report titled \href{https://octoverse.github.com/}{ ``The State
+of the Octoverse 2017''}.} and questions on StackOverflow\footnote{See
 \href{https://stackoverflow.blog/2017/09/06/incredible-growth-python/}{ ``The
@@
 third-party vendors, ranks Python at number four. IEEE Spectrum\footnote{See
-the analysis titled \href{https://spectrum.ieee.org/computing/software/%
+	the analysis titled
+	\href{https://spectrum.ieee.org/computing/software/%
 the-2017-top-programming-languages}{``The 2017 Top Programming Languages''} on
@@
 based on trends, demand by employers and users on open-source hubs.  All the
-rankings do agree that Python usage has been steadily growing and is here to
-stay.  Python largely outclasses languages like Fortran and Matlab according to
-these statistics, and is now clearly one of the mainstream languages with C,
-Java, C++ and Javascript.
+rankings agree that Python usage has been steadily growing and is here to stay.
+Python largely outclasses languages like Fortran and Matlab according to these
+statistics, and is now clearly one of the mainstream languages with C, Java,
+C++ and JavaScript.
+%
+The language has been designed to boost the communication of technical ideas
+between humans. 
 %
 Let us summarize some characteristics of the Python languages that have lead to
-such incredible success.
-
-The Python language has been designed to boost the communication of technical
-ideas between humans. The resulting code is as simple as possible due to the
-inherent emphasis on readability. Humans can focus on the ideas. Python differs
-from other mainstream languages for a variety of reasons:
+such an incredible success.
+%
+Python differs from other mainstream languages for a variety of reasons:
 
 \begin{itemize}
-\item \textbf{Aesthetics.} The syntax is nice and elegant. Blocks of code are
-defined with the indentation.
-\item \textbf{Code style.} There are explicit guidelines
-(\href{https://www.python.org/dev/peps/pep-0008/}{PEP~8}) supporting the
-regularity (and thereby readability) of the code and tools
-(linters) to evaluate adherence to the style.
-\item \textbf{Dynamic typing.} The variables (the ``names'' in the code) are
-not attached for life to an object in memory\footnote{Opposite behavior
-compared to static languages as Fortran, C and C++.}. The types of the objects
-are inferred at run time from the code and the context, and in many situations,
-one does not need to declare them explicitly.
-%
-% Being a strongly typed language\footnote{In contrast to weakly typed languages,
-% such as JavaScript}, types of the objects are respected and implicit type
-% conversions are rarely performed.
-% av: "rarely / never"?
-% eg: 'hello' * 2 -> 'hellohello'; 1/3 -> 0.3333333
-%
-% pa: I think we can not cover such complicate notion in this document... It is
-% anyway not so important for our demonstration.
-\item \textbf{Automatic memory management.}  The user need not handle the
-memory allocation and deallocation by hand.  There are, of course, ways to
-control the memory usage of a program, though not as precisely as for low-level
-languages.
+\item \textbf{Aesthetics.} The syntax is expressive, with an inherent emphasis
+on readability. Blocks of code are defined with the indentation.
+
+\item \textbf{Code style.} There are official (but, optional) guidelines, named
+\fnref{https://www.python.org/dev/peps/pep-0008/}{PEP~8}, advocating for
+regularity of the code. Linters and code-formatters can evaluate adherence to
+this style.
+
+\item \textbf{Automatic memory management.}  A garbage collector handles the
+memory allocation and deallocation, which needs no user intervention. There
+are, of course, ways to control the memory usage of a program, though not as
+precisely as for low-level languages.
+
+\item \textbf{Dynamic typing.}
+The types of the objects are inferred at run time from the code and the
+context, and in many situations, one does not need to declare them explicitly.
+
 \item \textbf{Interpreted.} The standard way\footnote{There are also tools to
@@
 %
-We will discuss shortly the disadvantages of this characteristic. Two evident
-advantages are rapid development cycle and the possibility of interactive
-workflow (usually with IPython or Jupyter notebooks).
-
-Note that there are many implementations of Python interpreters\footnote{We can
-cite CPython (written in C), Jython (Java), IronPython (C\#), PyPy (Rpython, a
-subset of Python) and MicroPython/CircuitPython (C, targeted to
-micro-controllers).}. The default and most widely used implementation is written
-in C and is called CPython.
-
-\item \textbf{Simple, yet powerful.}
-
-\begin{itemize}
-\item Python supports multiple programming paradigms, including imperative,
-object-oriented and functional.
-
-\item It is easy to interface with code written in other languages (in particular
-C, C++ and Fortran).
-
-\item Python comes with a large and high quality \href{https://docs.python.org%
-/3/library/index.html}{standard library}.
-
-\item Python can be deployed on many different machines with different operating
-systems (Linux, Windows, macOS, Android) and architectures (from a microcontroller
---- with MicroPython or CircuitPython --- to a Blue Gene supercomputer).
+Two evident advantages are rapid development cycle and the possibility of
+interactive workflow (usually with IPython or Jupyter notebooks).
+%
+Note that there are many implementations of Python
+interpreters\footnote{Notable ones are CPython (written in C), Jython (Java),
+IronPython (C\#), PyPy (Rpython, a subset of Python) and
+MicroPython/CircuitPython (C, targeted to micro-controllers).}. The default and
+most widely used implementation is written in C and is called CPython.
+
+\item \textbf{Multi-paradigm.} Supports imperative, object-oriented, functional
+and aspect-oriented programming.
+
+\item \textbf{Extension support}. It is easy to interface with code written in
+several other languages (in particular C, C++ and Fortran).
+
+\item \textbf{Batteries included.} Python comes with a large and high quality
+\href{https://docs.python.org% /3/library/index.html}{standard library}.
+
+\item \textbf{Cross platform.} Python can be deployed on many different
+machines with different operating systems (Linux, Windows, macOS, Android) and
+architectures (from a microcontroller --- with MicroPython or CircuitPython ---
+to a Blue Gene supercomputer).
 
 \end{itemize}
-\end{itemize}
 
 Thus, Python is one of the best ``glue'' languages for fast prototyping.
-Developing with Python is much faster and easier than with many other
+Code development in Python is much faster and easier than with many other
 languages. Less bugs are introduced by the developers simply because there are
-much less lines of code \cite{nanz_comparative_2015}.
+fewer lines of code \citep{nanz_comparative_2015}.
 %
@@
 readability associated with the language.
-
+%
 Another consequence of the apparent simplicity of Python is that the learning
@@
 evolve.
-\begin{itemize}
-\item A hard and controversial decision was taken to clean up Python from its
+%
+A hard and controversial decision was taken to clean up Python from its
 historical incoherences by having a new series of backwards-incompatible
@@
 open doors to very clean, coherent and potentially faster Python
-experience. Therefore, one should not use Python 2 for science anymore and
-instead, adopt the newer versions ($>=3.5$ in 2018).
-
-\item New versions of Python have nice new features to adapt to new usages and to
+experience. Therefore, one should not use Python 2 for science any more and
+instead, adopt the newer versions.
+
+New versions of Python have nice new features to adapt to new usages and to
 recent trends in computer science.  For example, three new features were
@@
 \href{https://www.python.org/dev/peps/pep-0484/}{PEP~484}).
+
+Being a very versatile language, Python is widely used for many different
+applications\footnote{\href{https://www.jetbrains.com/research/devecosystem-2018/python/}{See
+\emph{The State of Developer Ecosystem Survey in 2018} by JetBrains}}.
+
+\begin{itemize}
+\item Simple scripting
+
+\item Web development
+
+\item Data science
+
+\item System, database and network administration
+
+\item GNU/Linux distribution software
+
+\item Python is now one of the main scripting languages to add programmability
+to applications (for example Paraview, Visit, QGIS, Blender, ...)
+
+\item Web scraping
+
+\item Animation, gaming and film industry
+
+\item Education
+
+\item Science
+
 \end{itemize}
 
-\paragraphbf{An incredible success}
-
-Being a very versatile language, Python is widely used for many different
-applications:
-\begin{itemize}
-\item Simple scripting.
-
-\item System, database and network administration.
-
-\item GNU/Linux distribution software.
-
-\item Python is now one of the main scripting languages to add programmability
-to applications (for example Paraview, Visit, QGIS, Blender, ...).
-
-\item Web servers.
-
-\item Web scraping and data analysis.
-
-\item Animation movies, game development and gaming.
-
-\item Education. Python has recently become one of the most popular language
+Python has recently become one of the most popular language
 for teaching programming and computer science.
-
-\item Science!
 Python is now widely used in scientific applications. It has now a mature and
@@
 structures) and several more specialized packages (to name a few, \pack{h5py},
-\pack{mpi4py}, \pack{skimage}, \pack{sklearn}, ...). There are now great (and
-often shiny) tools for most of the applications. For example, the
-\href{https://github.com/rougier/python-visualization-landscape}{%
-landscape of visualization tools usable through Python} is impressive.
+\pack{mpi4py}, \pack{skimage}, \pack{sklearn}).
+The \href{https://github.com/rougier/python-visualization-landscape}{%
+landscape of visualization tools usable through Python} makes it possible to
+integrate post-processing and analysis capabilities into a single application.
 %
@@
 
-There are now ready-to-use ``batteries-included'' Python installers
-(similar to Matlab), like in particular the
+There are now ready-to-use ``batteries-included'' Python installers (similar to
+Matlab), like in particular the
 \href{https://en.wikipedia.org/wiki/Anaconda_(Python_distribution)}{%
-freemium open-source distribution Anaconda}, so it is very easy to start using
-Python for scientific purposes.
-
-There are now nice Python integrated development environments (IDE) adapted for
+open-source distribution Anaconda}, so it is very easy to start using Python
+for scientific purposes. Integrated development environments (IDE) adapted for
 scientists, for example \href{https://github.com/spyder-ide/spyder}{Spyder},
-\href{http://www.pyzo.org/}{Pyzo} and
-\href{https://jupyterlab.readthedocs.io}{Jupyterlab}.
-
-\end{itemize}
+and \href{https://jupyterlab.readthedocs.io}{JupyterLab}.
 

### 2019-05-24 chapter_00_open_science.tex
Corrections and coherence to Python section

@@
 \item \textbf{Dynamic typing.}
-The types of the objects are inferred at run time from the code and the
+The types of objects are inferred at run time from the code and the
 context, and in many situations, one does not need to declare them explicitly.
@@
 %
-In contrast to other compiled languages such as Fortran, C or C++, there is no
+In contrast to other compiled languages, such as Fortran, C or C++, there is no
 proper compilation step: the code is not translated to optimized machine
 instructions, but only to Python bytecode.
-%
-Two evident advantages are rapid development cycle and the possibility of
-interactive workflow (usually with IPython or Jupyter notebooks).
+
+\item \textbf{Multi-paradigm.} Supports imperative, object-oriented, functional
+and aspect-oriented programming.
+
+\item \textbf{Extension support}. It is easy to interface with code written in
+several other languages (in particular C, C++ and Fortran).
+
+\item \textbf{Batteries included.} Python comes with a large and high quality
+\href{https://docs.python.org% /3/library/index.html}{standard library}.
+
+\item \textbf{Cross platform.} Python can be deployed on many different
+machines with different operating systems (Linux, Windows, macOS, Android) and
+architectures (from a microcontroller --- with MicroPython or CircuitPython ---
+to a Blue Gene supercomputer).
+
+\end{itemize}
+
+Thus, Python is one of the best ``glue'' languages for fast prototyping.
+Code development in Python is much faster and easier than with many other
+languages. Less bugs are introduced by the developers simply because there are
+fewer lines of code \citep{nanz_comparative_2015}.
+%
+Maintaining and modifying existing code is also eased by the very good code
+readability associated with the language.
+%
+Another consequence of the apparent simplicity of Python is that the learning
+curve has a nice shape. Learning Python is very easy at the beginning and the
+complexity gradually increases as the level of the developer increases.
+%
+Its interpreted nature is advantageous for a rapid development cycle and the
+possibility of interactive workflow (usually with IPython or Jupyter
+notebooks).
+%
+Therefore, Python is good for developers of all levels: very gentle for
+beginners and very powerful for advanced users.
+
+Python is an old language (first implementation in 1991!) but it continues to
+evolve.
 %
@@
 most widely used implementation is written in C and is called CPython.
-
-\item \textbf{Multi-paradigm.} Supports imperative, object-oriented, functional
-and aspect-oriented programming.
-
-\item \textbf{Extension support}. It is easy to interface with code written in
-several other languages (in particular C, C++ and Fortran).
-
-\item \textbf{Batteries included.} Python comes with a large and high quality
-\href{https://docs.python.org% /3/library/index.html}{standard library}.
-
-\item \textbf{Cross platform.} Python can be deployed on many different
-machines with different operating systems (Linux, Windows, macOS, Android) and
-architectures (from a microcontroller --- with MicroPython or CircuitPython ---
-to a Blue Gene supercomputer).
-
-\end{itemize}
-
-Thus, Python is one of the best ``glue'' languages for fast prototyping.
-Code development in Python is much faster and easier than with many other
-languages. Less bugs are introduced by the developers simply because there are
-fewer lines of code \citep{nanz_comparative_2015}.
-%
-Maintaining and modifying existing code is also eased by the very good code
-readability associated with the language.
-%
-Another consequence of the apparent simplicity of Python is that the learning
-curve has a nice shape. Learning Python is very easy at the beginning and the
-complexity gradually increases as the level of the developer increases.
-%
-Therefore, Python is good for developers of all levels: very gentle for
-beginners and very powerful for advanced users.
-
-Python is an old language (first implementation in 1991!) but it continues to
-evolve.
 %
@@
 historical incoherences by having a new series of backwards-incompatible
-versions of the language, Python 3.  Python 3.0 was released in December 2008.
+versions of the language, starting with Python 3.0, released in December 2008.
 After a long and difficult process, the transition from Python 2 to Python 3 is
-now (2018) behind us for nearly all important packages\footnote{Close to
+now (201) behind us for nearly all important packages\footnote{Close to
 \href{https://python3wos.appspot.com/}{95\% of the most downloaded packages}
@@
 instead, adopt the newer versions.
-
+%
 New versions of Python have nice new features to adapt to new usages and to
@@
 applications\footnote{\href{https://www.jetbrains.com/research/devecosystem-2018/python/}{See
-\emph{The State of Developer Ecosystem Survey in 2018} by JetBrains}}.
+\emph{The State of Developer Ecosystem Survey in 2018} by JetBrains}}:
 
@@
 
-\item Python is now one of the main scripting languages to add programmability
-to applications (for example Paraview, Visit, QGIS, Blender, ...)
+\item GUI desktop applications
+
+\item Scripting layer for applications (for example Paraview, Visit, QGIS, Blender, ...)
 
@@
 
-Python has recently become one of the most popular language
-for teaching programming and computer science.
-Python is now widely used in scientific applications. It has now a mature and
-powerful scientific ecosystem with well-established based packages (\Numpy for
-N-dimensional homogeneous arrays, \Scipy as the fundamental toolkit for
-scientific computing, \pack{Matplotlib} for plotting and \pack{Pandas} for data
-structures) and several more specialized packages (to name a few, \pack{h5py},
-\pack{mpi4py}, \pack{skimage}, \pack{sklearn}).
+Python
+% has recently become one of the most popular language for teaching programming
+% and developing scientific applications. It 
+has now a mature and powerful scientific ecosystem with well-established based
+packages (\Numpy for N-dimensional homogeneous arrays, \Scipy as the
+fundamental toolkit for scientific computing, \pack{Matplotlib} for plotting
+and \pack{Pandas} for data structures) and several more specialized packages
+(to name a few, \pack{h5py}, \pack{mpi4py}, \pack{skimage}, \pack{sklearn}).
 The \href{https://github.com/rougier/python-visualization-landscape}{%
@@
 Python is one of the main languages for data science with packages such as
-\pack{Pandas}, \pack{statmodels}, \pack{sklearn}, \pack{Keras}, \pack{TensorFlow}
-and \pack{PyTorch}.
-
-There are now ready-to-use ``batteries-included'' Python installers (similar to
-Matlab), like in particular the
+\pack{Pandas}, \pack{statmodels}, \pack{sklearn}, \pack{Keras},
+\pack{TensorFlow} and \pack{PyTorch}.
+%
+There user-friendly and ready-to-use Python installers (similar to Matlab), in
+particular the
 \href{https://en.wikipedia.org/wiki/Anaconda_(Python_distribution)}{%
-open-source distribution Anaconda}, so it is very easy to start using Python
-for scientific purposes. Integrated development environments (IDE) adapted for
+open-source distribution Anaconda} to streamline the installation in 
+desktop platforms.
+Integrated development environments (IDE) adapted for
 scientists, for example \href{https://github.com/spyder-ide/spyder}{Spyder},
-and \href{https://jupyterlab.readthedocs.io}{JupyterLab}.
-
-Python has a very large and supportive community (see for example
-the \href{http://stackoverflow.com/tags}{Stack Overflow tags}).
-%
-But Python is also supported by companies using Python and the open-source
-dynamics.  Most of these companies that support Python are specialized in
-web technologies and data-driven research.  But, recently, companies basing
-their business
-% av: earlier "computer and web technologies". Computer is quite obvious, isn't
-% it?
-on Python for science have started to become quite powerful in terms of
-development. For example Anaconda Inc.\ is strongly implied in the development of
-tools like \pack{Numba} and \pack{dask}. Hardware giants such as Intel and Nvidia
-also contribute to the Python scientific ecosystem.
+and \href{https://jupyterlab.readthedocs.io}{JupyterLab} are often useful for
+interactive development.
+
+Python has a very large and supportive community (see for example the
+\href{http://stackoverflow.com/tags}{Stack Overflow tags}).
+%
+But Python development is also supported by companies specialized in web
+technologies and data-driven research. However, recently, companies involved in
+scientific software development have started to become quite influential. For
+example Anaconda Inc.\ is strongly engaged in the development of tools like JIT
+Python accelerator \pack{Numba}, and parallelism framework \pack{dask}.
+Hardware giants such as Intel and Nvidia also contribute to the Python
+scientific ecosystem.
+%
+To summarize, it is practical to use and extend Python for scientific purposes.
 
@@
 Since the standard Python interpreter does not do any proper
-compilation, Python code in some CPU bounded cases can be too slow. Thus, one
+compilation, Python code for some CPU bounded tasks can be too slow. Thus, one
 has to use special tools to get optimized machine instructions to speedup
@@
 
-\paragraphbf{Ahead Of Time compilation}  Since Python is primarily a ``glue''
-language, a common solution is to use code written in C, C++ or Fortran.
+\paragraphbf{Ahead Of Time compilation} Python is primarily a ``glue''
+language, and a common solution is to use code written in C, C++ or Fortran.
 %

### 2019-05-24 chapter_00_open_science.tex
Correct Python section and start the workflow section

@@
 
-\item \textbf{Multi-paradigm.} Supports imperative, object-oriented, functional
-and aspect-oriented programming.
+\item \textbf{Multi-paradigm.} Imperative, object-oriented, functional
+and aspect-oriented programming are supported.
 
@@
 
-\item Scripting layer for applications (for example Paraview, Visit, QGIS, Blender, ...)
+\item Scripting layer for applications (for example Paraview, Visit, QGIS,
+	Blender)
 
@@
 
-Since the standard Python interpreter does not do any proper
-compilation, Python code for some CPU bounded tasks can be too slow. Thus, one
-has to use special tools to get optimized machine instructions to speedup
-performance-critical code.
-
-\paragraphbf{Ahead Of Time compilation} Python is primarily a ``glue''
-language, and a common solution is to use code written in C, C++ or Fortran.
+In the context of scientific computing there are some common issues that Python
+developers would face. The issues are listed below and the solutions are
+specified briefly. The solutions are explained in further detail in subsequent
+paragraphs.
+
+\begin{itemize}
+
+\item \textbf{Performance in CPU bounded tasks.} Since the standard Python
+	interpreter does not do any proper compilation, pure Python code for
+	some CPU bounded tasks can be too slow. Thus, one has to avoid writing
+	loops as much as possible and use vectorized array operations as much
+	as possible, or call \emph{foreign functions}, or use special tools to
+	build extensions using \emph{Ahead Of Time} or \emph{Just In Time}
+	compilation to speedup performance-critical code.
+
+\item \textbf{Absence of type checking.} In contrast to compiled languages,
+	there is no type-checking in Python (and more generally in many
+	interpreted languages). Types are respected however, and type errors,
+	if any, would occur during runtime.
+	\emph{Type hints} and \emph{static type checking} can solve this issue.
+
+\item \textbf{Concurrent, but not parallel multi-threading.} Threads behave
+	like light sub-programs. In principle, they can use the different cores
+	of the CPU at the same time, but the CPython interpreter forbid threads
+	to interpret python code at the same time\footnote{See
+	\href{https://opensource.com/article/17/4/grok-gil}{this blog article
+on ``How to grok the GIL''} and
+\href{https://faster-cpython.readthedocs.io/cpython37.html}{this technical
+presentation by a CPython core developer}.}. In other words, threads can
+coexist concurrently, by not execute in parallel.
+%
+This limitation comes from an important detail in the implementation of the
+CPython interpreter: a Global Interpreter Lock (GIL) to prevent race conditions
+that could corrupt data.
+%
+GIL greatly simplifies the implementation of CPython and that
+it is very difficult to remove it while keeping other nice technical properties
+of CPython intact\footnote{See
+\url{https://wiki.python.org/moin/GlobalInterpreterLock}. Note that PyPy
+also uses a GIL, while two other implementations of the Python language, Jython
+and IronPython, do not have this limitation}.
+
+\item \textbf{Lively, and thus, complicated ecosystem of packages.} For many
+	applications, the standard library is insufficient and one needs
+	external packages. Sometimes there are multiple options and the
+	user/developer has to make choices regarding the tools to use.
+	Especially for the beginners who are not used to the open-source realm,
+	it can be confusing prospect to make good technological choices.
+
+\end{itemize}
+
+\subsubsection{Foreign functions}
+%
+Python is primarily a ``glue'' language, and a common solution to achieve
+performance is to use existing code written in C, C++ or Fortran.
 %
@@
 %
+Similarly \href{https://cppyy.readthedocs.io}{cppyy} can interpret C++ code.
 Libraries with Python bindings can also be used directly (for e.g., wrapped
@@
 
-Python developers do not actually have to write the compiled extensions in C,
-C++ or Fortran, since there are tools to generate them automatically from
-Python code (Pythran) or from Cython code. Cython \citep{behnel_cython2011} is a
-smooth blend of two languages with possibility to add type declarations like in
-C/C++, but with a syntax similar to Python.
+\subsubsection{Ahead Of Time compilation}
+Ahead Of Time compilation is process of generating compiled extensions which
+can be dynamically executed by Python.
+%
+Python developers do not actually have to write the extensions in C, C++ or
+Fortran, since there are tools to generate them automatically from Python code
+(Pythran) or from Cython code. Cython \citep{behnel_cython2011} is a smooth
+blend of two languages with possibility to add type declarations like in C/C++,
+but with a syntax similar to Python.
+%
+There are also other projects to create extensions in C++ with
+\href{https://github.com/pybind/pybind11/}{pybind11}, and in Rust with
+\href{https://github.com/PyO3}{PyO3}.
 
@@
 %
-This two-step compilation provides room for very clever optimizations as
-demonstrated, for example, in
-\href{http://serge-sans-paille.github.io/pythran-stories/%
-being-more-than-a-translator.html}{%
-this blog post by the main developer of Pythran}.
+This two-step compilation provides room for very clever optimizations
+\footnote{As demonstrated in this blog post by the main developer of
+Pythran \url{http://serge-sans-paille.github.io/pythran-stories/%
+being-more-than-a-translator.html}}.
 %
@@
 \citep{guelton2018pythran}.
-
+%
 We present in \citet{fluidfft, fluidsim} examples of
@@
 
-\paragraphbf{Just In Time (JIT) compilation} Another way to obtain optimized
-machine instructions is by using JIT compilation, i.e.\ by compiling only the
-critical code at run time.
+\subsubsection{Just In Time (JIT) compilation}
+Another way to obtain optimized machine instructions is by using JIT
+compilation, i.e.\ by compiling only the critical code at run time.
 %
@@
 %
-Pypy, an alternative interpreter written in Python, has a JIT compiler. However,
-Pypy is not widely used for scientific applications mainly because of
+PyPy, an alternative interpreter written in Python, has a JIT compiler. However,
+PyPy is not widely used for scientific applications mainly because of
 compatibility problems with the extensions written for CPython.
 %
-However this could change since a recent version of Pypy (6.0) supports the main
-packages of the Python scientific stack (\Numpy, \Scipy, \pack{Matplotlib},
-\pack{Pandas}, etc.).
-
-Adding a JIT to the interpreter CPython has been notoriously difficult (see
-\href{https://faster-cpython.readthedocs.io/}{faster-cpython}). One of the
+However this could change since recent versions, starting with PyPy 6.0,
+supports the main packages of the Python scientific stack (\Numpy, \Scipy,
+\pack{Matplotlib}, \pack{Pandas}, etc.).
+
+Adding a JIT to the interpreter CPython has been notoriously difficult
+\footnote{\url{https://faster-cpython.readthedocs.io/}}. One of the
 problem seems to be the C API provided in CPython and used by many extensions.
 %
-Python being first a glue language, a faster CPython with a JIT is actually not
-so necessary for many purposes. Therefore, not as much money and work has been
-put in accelerating Python than for example for Java and JavaScript.
+A faster CPython with a JIT is actually not so necessary for many purposes,
+since Python extensions are enough. Therefore, not as much money and work
+has been put in accelerating Python than for example for Java and JavaScript.
 
 Another strategy is to add a JIT to CPython through an external package
-(\pack{Numba}) and to only compile the critical functions.  Moreover,
+(\pack{Numba}).
+%
 \pack{Numba} is particularly interesting because it can take advantage of the
@@
 
-\paragraphbf{No type-checking} In contrast to languages using Ahead Of Time
-compilation, there is no type-checking in Python (and more generally in many
-dynamical languages).  Type-checking is useful since it can avoid simple bugs.
+\subsubsection{Type hints and static type checking}
+%
+Type-checking is useful since it can avoid simple bugs.
+\fnref{https://docs.python.org/3/whatsnew/3.5.html\#whatsnew-pep-484}{PEP-484}
+introduced a syntax and a related \codeinline{typing} module to add
+type-hints, typically for function parameters and return types.
+%
+Static type checkers like \href{http://mypy-lang.org/}{Mypy} and
+\href{https://github.com/google/pytype}{pytype}
+%
+\footnote{We plan to add type hinting in the most important FluidDyn modules to
+investigate how we can take advantage of type-checking with Mypy.}
+%
+allows to process type-hints and report inconsistencies in the code.
+
 However, it is not a good practice even with static languages to rely only on
@@
 least a little bit of time and energy to write a decent battery of tests.
-%
-We have to mention the young and promising project
-\href{http://mypy-lang.org/}{Mypy}\footnote{Guido van Rossum, the author and
-``Benevolent Dictator For Life'' of Python, is involved in the project Mypy.},
-which allows for Optional Static Typing and thus type-checking for
-Python\footnote{We plan to add type hitting in the most important FluidDyn modules
-to investigate how we can take advantage of type-checking with Mypy.}.
-
-\paragraphbf{A lively, huge and thus complicated ecosystem} As already
-mentioned, Python is a versatile language. For many applications, the standard
-library is insufficient and one needs external packages.
-%
-% For most applications there are several projects usable through Python.
-Most of the time there are several Python packages for a certain application.
-A user/developer has to make choices regarding the tools to use.  Especially
-for the beginners who are not used to the open-source realm, it can be
-difficult to understand a ``landscape of projects'' for a particular
-application and to make good technological choices.
-
-It is important to get a good introduction on how to work with the open-source
-dynamics and the scientific Python ecosystem.  It can also be very useful to ask
-to more advanced users and attend Python courses or workshops\footnote{This could
-be regular academic courses, MOOCs or workshops like ones organized by
-\href{https://software-carpentry.org/}{Software Carpentry}.}.
-%
-Of course, such abundance of projects can also be seen as an advantage.
-
-\paragraphbf{Absence of multicore computational parallelism using threads}
-Threads behave like light subprograms. In principle, they can use the different
-cores of the CPU at the same time but the CPython interpreter forbids that they
-interpret python code at the same time\footnote{See
-\href{https://opensource.com/article/17/4/grok-gil}{this blog article on ``How
-to grok the GIL''} and
-\href{https://faster-cpython.readthedocs.io/cpython37.html}{this technical
-presentation by a CPython core developer}.}.
-%
-This limitation comes from an important detail in the implementation of the
-CPython interpreter: a Global Interpreter Lock (GIL) to prevent race conditions
-that could corrupt data.
-%
-It seems that the GIL greatly simplies the implementation of CPython and that it
-is very difficult to remove it while keeping other nice technical properties of
-CPython intact\footnote{See \url{https://wiki.python.org/moin/GlobalInterpreterLock}.}.
-Note that Pypy also uses a GIL, while two other implementations of the Python
-language, Jython and IronPython, do not have this limitation.
+
+\subsubsection{Achieving computational parallelism}
 
 At any rate, threads using the interpreter do not use the CPU at the same time
-so it is not possible to use threads to do multicore computational parallelism
+so it is not possible to use threads to do multi-core computational parallelism
 with pure python code to accelerate CPU bounded tasks.
@@
 %
-To do proper computational parallelism for CPU bounded tasks with the GIL, one has
-to use other strategies: extensions (fine grain parallelism), multiprocessing
-(coarse grain parallelism) and inter-process communication, for example, with MPI
-(with mpi4py) or ZeroMQ.
-
-\subsection{Rich landscape of open-source languages useful for science}
+To do proper computational parallelism for CPU bounded tasks with the GIL, one
+has to use other strategies: extensions (fine grain parallelism),
+multiprocessing (coarse grain parallelism) and inter-process communication, for
+example, with MPI (with mpi4py) or ZeroMQ.
+
+\subsubsection{Choosing the right package}
+%
+For a beginner in Python development, it is important to get a good
+introduction on how to work with the open-source workflows and understanding of
+the scientific Python stack.
+%
+It can also be very useful to ask to more advanced users and attend Python
+courses or workshops\footnote{This could be regular academic courses, MOOCs or
+workshops like ones organized by
+\href{https://software-carpentry.org/}{Software Carpentry}.}.
+%
+Of course, such abundance of projects can also be seen as an advantage.
+
+\subsection{The state of languages for scientific computing}
 
@@
 JavaScript, Java, Scala, Smalltalk, Haskell, R, Julia, Perl and Lua are used for
-some scientific codes. Go and Rust are quite young languages and are not (yet)
-widely used for scientific applications.
-
-It is of course very difficult to predict which technologies will be widely
+some scientific codes. Go and Rust are quite young languages with emphasis on
+memory-safety and are not (yet) widely used for scientific applications.
+%
+It is, of course, very difficult to predict which technologies will be widely
 used in even the near future (10 or 20 years for example).
 %
-From what we see, it seems that the idea of ``one language to do everything for
-science'' will not succeed, at least soon.  This was maybe the project of Julia
-creators, but despite the qualities of this language, it also has
-limitations\footnote{For example C-like include instead of Python-like
-import. From our point of view, a Pythonic flavor of Julia would be very
-interesting.}.
+It seems that the idea of ``one language to do everything for science'' will
+not succeed, at least not in the near future. 
+%
+The limited scope of the language is likely the reason why the community
+around Fortran and Julia are not widespread outside academia.
 %
@@
 between coexisting tools\footnote{See for example the cross-language development
-platform \href{https://arrow.apache.org/}{Apache Arrow}.}.
-
-So what about Python?  As already mentioned, it is today the language with the
-strongest dynamics in science, data analysis and much more.  Several students
-learn and will learn it. Many scientists in the future would, hopefully, be
-able to appreciate these strengths and favour the Pythonic approaches.  With
-such dynamics today, we can anticipate that the Python world will be able to
-embrace new technological trends and to overcome some of its current
-limitations.
-%
-It is reasonable to predict that Python will continue to shine as a language
-(i) particularly good for human communication and to write ideas, (ii) for fast
-prototyping and (iii) for interoperability with code written in other languages
-(see for example the recent projects \href{https://cppyy.readthedocs.io}{cppyy}
-and \href{https://github.com/pybind/pybind11/}{pybind11}, or
-\href{https://github.com/PyO3}{the possibility to interact with Rust code}).
-
-\paragraphbf{Summary: Python, a versatile Swiss Army Knife for the
-scientists}
-
-Python has become a great versatile tool for scientists.  The scientific
-ecosystem of Python is nowadays really impressive.  The scientific Python
-community is strong and quickly growing.
-%
-A scientist with good skills in Python can do most of what she/he needs to do in
-programming (which does not mean that Python is the best language to do
-anything!).
-%
-In contrast, no skill in Python is a real disadvantage for many tasks and for
-employability.
-%
-Therefore, in most fields today, if a scientist has to acquire good skills in
-one language, it is in Python.
-
+platform for in-memory data, \href{https://arrow.apache.org/}{Apache Arrow}.}.
+%
+Knowledge of Python is also an advantage in terms of employability these days.
+%
 Considering what we have presented on Python, we think that it is reasonable to
 bet that it will become one of the \textit{lingua franca} in science and in
-fluid dynamics.  This explains why we have chosen it to be the main language
-for the project FluidDyn.
+fluid dynamics.
+%
+We have, therefore, chosen it to be the main language for the project FluidDyn.
 
@@
 
-% remark Julien Salort "It is well known that": bad
-% pa: common sense says ?
-% It is well known 
+% remark Julien Salort "It is well known that": bad pa: common sense says ?  It
+% is well known 
 It is not surprising
@@
 %
-that short-term efficiency and long-term efficiency are
-sometimes incompatible.
-%
-Quick and dirty scripts can be efficient in the short term, but such a workflow
-has a negative impact on a longer time scale.
-
-Similarly, we can also differentiate productivity at different scales of
+that short-term efficiency and long-term efficiency are sometimes incompatible.
+%
+Quick and dirty scripts can be efficient in the short term, but has a negative
+impact on a longer time scale.
+%
+Similarly, we can also differentiate productivity at different levels of
 collaboration.
@@
 An individual can be very efficient with a particular tool, but the same tool
-can be very inefficient to collaborate even with co-workers.
-%
-A group can be very efficient with a code which is closed-source.  If this
-strategy can be efficient for the group (at least in the short term), it leads
-to waste for the community. Other groups will need to develop codes with the
-same features.  Good ideas will be spread in different codes and since the
-number of users and developers are smaller for each of these codes, the code
-quality will not improve as fast as it would with an open-source strategy. In a
-nutshell, as this proverb says, \textit{``If you want to go fast, go alone. But
-if you want to go far, go together.''}
+can be very inefficient when used in a collaboration.
+%
+A group can be very efficient with a code which is closed-source. Although such
+a strategy can be efficient for the group (at least in the short term), it may
+lead to waste for the community. Other groups will need to develop codes with
+the same features.
+%
+Good ideas will be spread in separate codes and since the number of users and
+developers is small for each of these codes, the code quality will not
+improve as fast as it would with an open-source strategy. In a nutshell, as
+this proverb says, \textit{``If you want to go fast, go alone. But if you want
+to go far, go together.''}
 
@@
 tools. However, some of its members have to spend more energy to bear this
-dynamics, which implies that they somehow sacrifice their individual efficiency.
+dynamics, at the risk of sacrificing their individual efficiency.
 We see that short-term efficiency, long-term efficiency and productivity at
@@
 We study fluid mechanics via laboratory experiments, in situ measurements,
-analytic computation, numerical simulations and data processing.
+analytic tools, numerical simulations and data processing.
 %
@@
 %
-Even today, it is quite commonplace to start a PhD without any serious training in
-GNU/Linux and modern programming tools.
+Even today, it is quite commonplace to start a Ph.D. without any serious
+training in GNU/Linux and modern programming tools.
 %

### 2019-05-25 chapter_00_open_science.tex
Correct till end of chapter opensci

@@
 We shall now discuss these topics in greater detail. The chapter is organized
-into three sections. The first section focusses on the technical aspects, with
+into four sections. The first section focusses on the technical aspects, with
 particular focus on emerging developments in programming languages, tools and
-services suited for open-source software development. In the next section, we
-present Python's strengths and weaknesses as a language for scientific
+services suited for open-source software development. The subsequent section
+presents Python's strengths and weaknesses as a language for scientific
 computing.
 %
-The section ends with a note on the capabilities of FluidDyn
-packages\footnote{We use FluidDyn (with capital letters) to name the project
-and \fluiddyn for the base package.}.
-%
-The last section compares various software development methodologies. Here, we
+The third section compares various software development methodologies. Here, we
 will also discuss about the possible contradiction between productivity of
@@
 %
-% The chapter finally ends with a summary of the motivations for the project
-% FluidDyn and an overview of the packages used.
+The chapter finally ends with a summary of the motivations for the project
+FluidDyn and an overview of the packages we have developed.
+% The chapter ends with a note on the capabilities of FluidDyn
+% packages\footnote{We use FluidDyn (with capital letters) to name the project
+% and \fluiddyn for the base package.}.
+%
 
@@
 %
-We have come a long way since then and today computational methods now
-complement theoretical and experimental studies, and are equally important.
+We have come a long way since then and today computational methods complement
+theoretical and experimental studies, and are equally important.
 %
@@
 With the connectivity provided by the world wide web it is possible to achieve
-more and reinvent the way we perform sciences.
+more and reinvent the way we do sciences.
 
@@
 
-In recent years there has been a move towards the open science
+In recent years, there has been a move towards open science
 \citep[see the report by][]{royal_society_great_britain_science_2012}.
 %
-Depicted in figure \ref{fig:opensci} are some of steps needed to ensure
+Depicted in figure~\ref{fig:opensci} are some of steps needed to ensure
 reproducibility and transparency in sciences.
@@
 
-A lot of work is required to condense and implement this knowledge in the
-form of working applications.
+A lot of work is required to condense and implement this knowledge in the form
+of working applications.
 %
@@
 	\url{https://choosealicense.com} or
-  \url{https://www.gnu.org/licenses/license-list.html}.
-}.
+	\url{https://www.gnu.org/licenses/license-list.html}. }.
 %
@@
 which a code is continuously built, tested and deployed --- a process known as
-Continuous Integration (CI), which is described in detail later in this
-chapter. Depending on how matured a software is CI can be easy or
-cumbersome to implement, but nevertheless, the result is a cleaner and reliable
-code.
+Continuous Integration (CI). Depending on how matured a software is, CI can be
+easy or cumbersome to implement, but nevertheless, the result is a cleaner and
+reliable code.
 
 Finally, the \emph{results} generated from the such codes together with the
-workflow (i.e. scripts to run the codes and to postprocess the data) when made
+workflow (i.e.\ scripts to run the codes and to post-process the data) when made
 open in the form of datasets can ensure that research is reproducible and
@@
 \href{https://en.wikipedia.org/wiki/Linus_Torvalds}{Linus Torvalds} in 1991)
-forms the GNU/Linux operating system.
+forms the GNU/Linux operating system
 % and its several distributions (Debian, Red Hat, Slackware, Arch Linux and
@@
 success of the Linux kernel project was attributed to the \emph{bazaar} model of
-software development in \citet{raymond_cathedral_1999}, wherein a code can stay
-structured despite releasing frequently and delegating tasks to community with
-myriad agendas.
+software development, as described in \citet{raymond_cathedral_1999}, wherein a
+code can stay structured despite releasing frequently and delegating tasks to
+community with myriad agendas.
 
@@
 %
-The FOSS movement has thus entered into a second era
+The FOSS movement has now entered into a second era
 \citep{fitzgerald_transformation_2006}.
@@
 Control Systems (DVCS), especially, Git and Mercurial). In web-based source
-development platforms like GitHub, Bitbucket and GitLab DVCS repositories are
-enhanced with code-review tools, which typically consists of a issue tracker
+development platforms like GitHub, Bitbucket and GitLab, DVCS repositories are
+enhanced with code-review tools, typically consisting of a issue tracker
 and pull-request mechanism.
@@
 scientists, for example \href{https://github.com/spyder-ide/spyder}{Spyder},
+\href{https://pyzo.org}{Pyzo},
 and \href{https://jupyterlab.readthedocs.io}{JupyterLab} are often useful for
@@
 %
-Similarly \href{https://cppyy.readthedocs.io}{cppyy} can interpret C++ code.
-Libraries with Python bindings can also be used directly (for e.g., wrapped
-using SWIG or Boost.Python) or one has to write extensions, i.e.\ Python
-modules written in C or C++ using the CPython API that can be imported directly
-from Python like any Python module.
+Similarly \href{https://cppyy.readthedocs.io}{\pack{cppyy}} can interpret C++
+code. Libraries with Python bindings can also be used directly (for e.g.,
+wrapped using SWIG or Boost.Python) or one has to write extensions, i.e.\
+Python modules written in C or C++ using the CPython API that can be imported
+directly from Python like any Python module.
 %
 However, to rely on such techniques to call C libraries would inhibit
-development of extending functionalities, as they have a steeper learning
-curve compared to Python.
-%
-Nevertheless, this strategy has given rise to the base modules of the scientific
-Python stack, namely \Numpy, \Scipy and \pack{Matplotlib}.
+development of extending functionalities, as they have a steeper learning curve
+compared to Python.
+%
+Nevertheless, this strategy has given rise to the base modules of the
+scientific Python stack, namely \Numpy, \Scipy and \pack{Matplotlib}.
 
@@
 pays for a license or a new development. The group does not learn how to develop
-what has been paid for. The company, on contrary, stands to gain from profits and
-the feedback generated from users, through which it improves the product sold.
+what has been paid for. The company, on the contrary, gains from
+the feedback generated from users, through which it improves the product.
 The group produces codes, books, courses using the closed-source product or
-acquire knowledge on how to use the closed-source product, so that the group is
-increasingly dependent on the product and is ready to pay more for it.
-%
-Such vicious circles are difficult to break, but not impossible. This can be
-done by introducing alternative open-source solutions, bringing in positivity to
-this circle.
-
-We have seen that it was very common to rely on commercial companies and
-closed-source software for most of the new developments in fluid mechanics.
-%
-Nevertheless there are also a lot of codes produced by scientists in the
-field. But without the technical knowledge on how to work collectively on code,
-this production is often of bad quality (compared to the information technology
-industry standard) and it is basically doomed to be abandoned after its use.
-%
-However, it is natural to try to reuse code or at least to build the next steps
-on what has been done before.  Thousands of hours of highly qualified people
-are spent in trying to understand and reuse codes badly written with
+acquire knowledge on how to use the closed-source product, making the group
+increasingly dependent on the product and ready to pay more for it.
+%
+It is difficult to break such vicious circles, but not impossible. This can be
+done by introducing alternative open-source solutions.
+
+Despite the reliance on commercial software for research in fluid mechanics,
+%
+there are also a lot of codes produced by scientists in fluid mechanics. But
+without the technical knowledge on how to work collectively, such codes
+are often of bad quality (compared to the information technology
+industry standard) and are basically doomed to be abandoned after their use.
+%
+However, it is natural to try to reuse codes or at least to build the next steps
+on what has been done already. Thousands of hours of highly qualified people
+are spent in trying to understand and reuse badly written codes in
 inappropriate languages!
 
-On a technical point of view, the majority of the coding in the field involve a
-mix of Fortran/C or C++, shell languages (as Bash) and Matlab. For experiments,
-the graphical programming environment
+From a technical perspective, the majority of the coding in the field involves
+a mix of Fortran/C or C++, shell languages (as Bash) and Matlab. For
+experiments, the graphical programming environment
 \href{http://www.ni.com/en-us/shop/labview.html}{Labview} is dominant for
-control of physical objects and acquisition and Matlab is used for data
-processing.
+control of physical objects and acquisition. Fortran is often used to implement
+solvers and post-processing algorithms. Matlab is used for data processing and
+visualization.
 %
@@
 %
-Similarly, using Matlab for developing complex programs is technically a
-questionable choice.
-%
-These technical solutions also represent a problem for code reuse, sharing and
+Use of compiled languages to develop new algorithms hinders the development
+process and ability to debug.
+%
+Similarly, using Matlab for developing complex programs is a questionable
+choice.
+%
+These technical solutions are also problematic for code reuse, sharing and
 collaborative development.
@@
 % \paragraphbf{Discussion on Matlab, Labview and Mathematica}
-
-% see http://www.pyzo.org/python_vs_matlab.html
-% see https://www.linkedin.com/pulse/matlab-vs-python-jan-rhebergen
 
@@
 fast, especially now that it has a JIT compiler.
-
+%
 However, the language suffers from serious technical issues which make Matlab
@@
 for doing more than simple processing and data plotting. The comparison with
-Python often debated upon. We list a few striking Matlab weaknesses:
+Python is often debated upon . We list a few striking Matlab weaknesses:
 
@@
 \item No real organization of the standard library. All built-in and
-user-defined functions are available in a huge flat namespace. No import
-mechanism and as a result we cannot deduce, simply by looking at the code,
-where a function comes from.
+user-defined functions are available in a huge flat namespace. Due to this
+absence of an import mechanism we cannot deduce, simply by looking at the code,
+where a function is defined.
 
@@
 modify and define global variables (see for example
-\href{https://www.damtp.cam.ac.uk/user/jrt51/files/diablo\_mat.tar.gz}{%
-this Matlab version of the code Diablo}).  It is so simple to do this that we
+\fnref{https://www.damtp.cam.ac.uk/user/jrt51/files/diablo\_mat.tar.gz}{%
+the Matlab version of the code Diablo}).  It is so simple to do this that we
 can assert the language strongly encourages this practice.  Matlab files are
@@
 \item Very bad default argument mechanism.
-
+%
 Default arguments for a function is a very common feature is many programming
@@
 \end{minted}
-
-An implementation for approximately the same behavior in Matlab could be (as
+%
+An implementation for approximately the same behaviour in Matlab could be (as
 advised in the official Matlab documentation):
@@
 
-\item Less intuitive syntax for string operations in Matlab, compared to
+\item Non-intuitive syntax for string operations in Matlab, compared to
 Python. Take for example for string comparison, \codeinline{strcmp(s1, s2)} in
@@
 in Python.
-% \item Less intuitive and clear syntax for exception
-% handling. \citationneeded[opinionated].
-% pa: Matlab has now a MException class. The syntax is not nice, but it is
-% not so important.
-\item Less clear and intuitive model and syntax for object oriented
-programming, compared to Python. As for functions, one file per user-defined
-class is necessary.
-% \citationneeded[opinionated].
-\item Parenthesis used for both function calls and indexing, causing ambiguity.
+\item Complicated syntax for object oriented programming, compared to Python.
+As for functions, one file per user-defined class is necessary.
+\item Parentheses used for both function calls and indexing, causing ambiguity.
 \item Matlab codes are usually full of \codeinline{;}, \codeinline{.*},
 \codeinline{./}, \codeinline{\&\&} and \codeinline{||}, which make them quite
-``noisy''.  Moreover, there is nothing like the Python PEP~8. This makes a
-large proportion of the Matlab codes difficult to read and understand, due to
-lack of a consistent code style.
-\item The syntax \codeinline{a(100, 100) = 1;} to create and extend matrix is a
+``noisy''.  Moreover, due to a lack of a consistent code style like, Python's
+PEP~8, a
+large proportion of the Matlab codes are difficult to read and understand.
+\item The syntax \codeinline{a(100, 100) = 1;} to create and extend a matrix is a
 very good way to hide bugs. The code \codeinline{a = eye(2); a(i0, i1) = 1;}
@@
 \end{itemize}
-
+%
 Now we turn our attention, from structural issues of Matlab syntax, to more
-practical constraints its use brings about.  By definition of a closed-source
-software, Matlab is a big black box: it is impossible to study the
+practical constraints brings about. The arguments are equally applicable for
+any proprietary software used in research.  Closed-source
+software are, by definition, black boxes: it is impossible to study the
 implementation of its functions.
-
+%
 Matlab is not free. The price for one license for non-commercial use is not
@@
 Matlab quickly escalates when deployed on a cluster with MPI, since a licence
-is required per user, per node!
-% av: not per process.
-% see: https://www.nas.nasa.gov/hecc/support/kb/running-matlab-applications-in-parallel_522.html
-
-It would be interesting to estimate what could be done with open-source
-solutions with a part of the amount paid by public research institutes. How
-many developers could be paid to improve the alternative open-source solutions?
-%
-This question raises the issue of funding scientific open-source software.
-
+is \fnref{%
+	https://www.nas.nasa.gov/hecc/support/kb/running-matlab-applications-in-parallel_522.html}{%
+required per user, per node}.
+%
 Moreover, it is well known that the lack of license for the personal computers
-of staff is sometimes solved with cracked versions of Matlab. How can serious
+of staff are solved with pirated versions. How can serious
 research projects rely on such illegal methods?
 %
+It is also unfair in education when students have to work on course projects and
+assignments, with a Matlab code as a starting point and not a free alternative.
+
+The same remarks can be made for Labview. It is a graphical programming
+language and the programs are saved as binary files. As a result, it is
+impossible to read a Labview program without Labview and to apply version
+control, crippling collaborative possibilities.
+%
+Matlab and Labview also lead to problems for students, who learn bad coding
+habits and languages that are much less in demand by employers than, for example
+\fnref{https://insights.stackoverflow.com/survey/2018\#technology}{C++ or
+Python}.
+
 Note that open-source Matlab interpreters do exist, such as Octave and Scilab.
 %
-% pa: warning! Octave has a JIT compiler!
-%
-% Unfortunately, these are not serious alternative today. Their user
-% community is too small and they are not very efficient, especially now that
-% Matlab uses a JIT compiler and that optimized Matlab codes rely on it.
-
-The recent open-source and Python dynamics represent a hard punch for Matlab,
-which is losing the dominant position it has in some domains.
-%
-Note that this is a very recent development. The Python scientific ecosystem can
-now compete in nearly all domains, but this was not the case typically 5 years
-ago.
-%
-There are still a large number of researchers and teachers in sciences who are
-strongly attached to Matlab mainly because they have used it a lot.
-%
-It is, of course, difficult to change habits especially when those are efficient
-(for a short time-span and on an individual or group scale) enough for what you
-do.
+With some minimal effort, one can port existing Matlab codes using either
+Julia which by design uses Matlab-like syntax, or Python's \codeinline{pylab}
+package provided via \pack{matplotlib} and the command-line tool
+\codeinline{fluidmat2py} in \pack{fluiddyn}.
+%
+For symbolic computation, some specialists use Maple or Mathematica, which are
+good but expensive proprietary closed-source programs.
+%
+Sage and Sympy are two complementary open-source alternative both based on
+Python.
+%
+The \fluidpack{lab} project in FluidDyn project is an attempt to orchestrate
+experiments without relying on Labview.
 
@@
 that we have already discussed. Other arguments are unfair or no longer valid
-nowadays.
-% Here, we remove a long argument with Mathworks (see the file
-% old_arguments_mathworks.tex)
-% av: Less provocative, that is good. But we can still add a footnote linking
-% the Pyzo page / LinkedIn page comparing python and matlab.
-%
-% We replace by this:
-The scientific Python ecosystem is now very mature, robust, well documented and
-fully integrated.  Many of this components are supported by professional
-developers paid by companies or research institutes, for example \pack{Numba}
-(Anaconda, Nvidia), \pack{TensorFlow} (Google), \pack{PyTorch} (Facebook, Twitter,
-Nvidia), Scikit-learn (INRIA), Mercurial (Facebook).
-%
-The support by the Python community is both very reactive and of very good
-level (for example in Stack Overflow, and issues page on major repositories)
-and it is now possible to pay companies to get professional support on
-scientific Python (for example
-\href{https://software.intel.com/en-us/distribution-for-python/get-help}{Intel}
-and \href{https://www.anaconda.com/enterprise/}{Anaconda}).
-%
-Finally, with the tools available today, with little effort we can get a very
-efficient program using only Python code.
-
-To conclude on Matlab, some people may use it efficiently for their work, but it
-starts to be a problem for the collective efficiency of the community. Except
-from the existence of a large code base difficult to translate in better
-languages or for some very specific applications, it seems difficult today to
-find a good technical reason to continue using Matlab.
-
-The same remarks can be made for Labview. It is a graphical programming
-language and the programs are saved as binary files. This is a problem because
-first it is impossible to read a Labview program without Labview and, more
-importantly, it forbids the use of source management tools which are so
-important for collective work and open-source.
-
-Matlab and Labview also lead to problems for students, who learn bad coding
-habits and languages that are much less in demand by employers than for example
-Python.
-
-For symbolic computation, some specialists use Mathematica, which is a good and
-very expensive proprietary closed-source mathematical symbolic computation
-program.
-%
-Sage and Sympy are two complementary open-source alternative both based on
-Python.
-
-\paragraphbf{Different models for software development in fluid mechanics}
+nowadays\footnote{See also \url{http://www.pyzo.org/python\_vs\_matlab.html}
+or \url{https://www.linkedin.com/pulse/matlab-vs-python-jan-rhebergen}}.
+%
+The Python scientific ecosystem can now compete in nearly all domains, but this
+was not the case typically 5 years ago.
+%
+It would be interesting to estimate what could be done with open-source
+solutions with a part of the amount paid by public research institutes. How
+many developers could be paid to improve the alternative open-source solutions?
+%
+This question raises the issue of funding scientific open-source software.
+
+\subsection{Different models for software development in fluid mechanics}
 
@@
 It is quite common to share only for ``friends'' to control the dissemination
-of the software.
-%
-This model can be institutionalized with an official group, for example with
-the code \href{https://www.coria-cfd.fr/index.php/YALES2}{Yales2}.
+of the software, as is the case with the codes such as
+\fnref{https://www.coria-cfd.fr/index.php/YALES2}{Yales2}.
 %
 % av: Institutionalized? I don't understand.
+%
 With such a model, the lack of a proper open-source license can naturally lead
-to problems and is a limitation for users/developers.
+to restrictions in collaboration.
 %
 Another common practice is to share without repository (for example
-\href{http://www.damtp.cam.ac.uk/user/jrt51/files.html}{Diablo} or
-\href{http://yakari.polytechnique.fr/people/deloncle/ns3d.html}{NS3D}) and/or
-without license (\href{https://choosealicense.com/no-permission/}{meaning that the
-code is actually under exclusive copyright}).
+% \fnref{http://www.damtp.cam.ac.uk/user/jrt51/files.html}{Diablo} or
+\fnref{http://turbulencehub.org/index.php/codes/tarang}{Tarang} or
+\fnref{http://yakari.polytechnique.fr/people/deloncle/ns3d.html}{NS3D}) and/or
+without license (\fnref{https://choosealicense.com/no-permission/}{meaning that
+the code is actually under exclusive copyright}).
 
@@
 (typically written in Matlab, see for example
-\href{http://pivlab.blogspot.de/}{PIVlab},
-\href{http://servforge.legi.grenoble-inp.fr/projects/soft-uvmat/}{UVmat} and
-\href{https://de.mathworks.com/matlabcentral/fileexchange/10902-pivmat-4-00}{PIVmat}).
-%
-Let us first say that these codes are a great wealth for the community at least
-for three reasons: (i) they can be used as they are, (ii) people using these
-codes can be more open to the open-source dynamics, and (iii) open-source
-research projects using open-source tools can exploit the ideas in these codes.
-%
-Let us also point out a drawback of such a model. We do not pay the authors, but
-people do need to pay to use the software. Even the authors of the software give
-money to the same company to be able to use their software. 
+\fnref{http://pivlab.blogspot.de/}{PIVlab},
+\fnref{http://servforge.legi.grenoble-inp.fr/projects/soft-uvmat/}{UVmat} and
+\fnref{https://de.mathworks.com/matlabcentral/fileexchange/10902-pivmat-4-00}{PIVmat}).
+%
+Let us first say that these codes are a great wealth for the community, as they
+can be used as they are, and other open-source research projects can reuse 
+the ideas expressed in these codes.
+%
+Although we do not pay the authors, but people do need to pay to use the
+software. 
 %
 % Do they realize that when they work on these programs,
-Ironically, when people develop programs using these tools, they work (often,
-for free) for the company that sells the proprietary tool.
+Ironically, when people develop programs using these tools, unintentionally
+they work (often, for free) for the company that sells the proprietary tool.
 
 Some proper open-source codes have emerged, for example
-\href{https://nek5000.mcs.anl.gov/}{NEK5000} (Fortran),
-\href{https://www.openfoam.com/}{OpenFOAM} (C++),
-\href{http://basilisk.fr}{Basilisk} (C) and
-\href{http://channelflow.org/}{Channelflow} (C++) and
-\href{https://www.code-saturne.org}{Code\_Saturn} (C/Fortran).
+\fnref{https://nek5000.mcs.anl.gov/}{NEK5000} (Fortran),
+\fnref{https://www.openfoam.com/}{OpenFOAM} (C++),
+\fnref{http://basilisk.fr}{Basilisk} (C) and
+\fnref{http://channelflow.org/}{Channelflow} (C++) and
+\fnref{https://www.code-saturne.org}{Code\_Saturn} (C/Fortran).
 %
@@
 %
-What we have now is not by ``reinventing the wheel'', but rather by ``standing
+What have not been ``reinventing the wheel'', but rather have been ``standing
 on the shoulders of giants''. In other words we have utilized and extended
@@
 
-% It first started with one package.  Soon we had to prefer a decentralized
-% organization with one common package and other specialized packages.
-% av: mentioned in the next subsection
-We provide examples of solutions for:
-\begin{itemize}
-\item Good coding practice with readable and easy to comprehend Python code
-(PEP~8).
-\item Source control management (Mercurial) and forge (Bitbucket) simple for
-the new comers.
-\item Packaging and installation procedure. All packages are available from
-\href{https://pypi.org/}{Pypi} and can be install simply with
-\href{https://pypi.org/project/pip/}{pip}, the standard Python installation
-tool.
-\item Licenses: depending on the packages, we choose to use the CeCILL-B or the
-CeCILL licenses. These licenses are a BSD compatible and a GPL compatible
-licenses adapted to both international and French legal
-matters\footnote{\url{http://www.cecill.info/licences.en.html}}.
-\item Documentation produced with standard and up-to-date tools: Sphinx,
-Anaconda and Jupyter. Built and hosted online at
-\href{https://readthedocs.org/}{Read the Docs}. Can also be generated offline.
-\item Unittest and continuous integration with Bitbucket Pipelines and Travis.
-\end{itemize}
-We hope that such clean framework will allow contributions from scientists in
-the field and that we can build together a nice, user-friendly and efficient
-ecosystem specialized in research and teaching in fluid dynamics.
-
-% \textcolor{blue}{How the software was implemented, with details of the
-% architecture where relevant. Use of relevant diagrams is appropriate. Please
-% also describe any variants and associated implementation differences.}
-
-\subsection{Organization of the code in packages}
-
-FluidDyn was originally intended to be a single package to perform experiments
-and simulations. Since a typical user may not be involved in both experiments
-and simulations and also, due to increasing complexity as a virtue of rapid
-development cycle made the need to decentralize FluidDyn evident.
-Now, FluidDyn project hosts a number of specialized packages, namely:
-
-\begin{itemize}
-\item \fluidpack{dyn}: The base package which contains pure-python code that can
-be reused in scripts or in specialized FluidDyn packages. It also contains codes
-for miscellaneous command-line utilities useful for a typical fluid dynamics user.
-
-The code of this package is presented in further detail in its documentation
-(\url{https://fluiddyn.readthedocs.io}) and some prominent features are presented
-in the following subsection.
-
-\item \fluidpack{fft}~\cite[see the companion paper][]{fluidfft}: a package which
-provides C++ and Python classes unifying various libaries to perform Fast Fourier
-Transform (FFT) in sequential and in parallel.
-
-\item \fluidpack{sim}~\cite[see the companion paper][]{fluidsim}: Numerically
-oriented framework to run sequential and parallel Computational Fluid Dynamics
-(CFD) simulations and on-the-fly post-processing for a variety of problems
-(Navier-Stokes, Shallow Water, F\"oppl von K\'arm\'an equations, to name a few).
-A study using \fluidpack{sim} has just been published in Physics of Fluids
-\cite[]{LindborgMohanan2017}.
-
-\item \fluidpack{lab}: Package to handle laboratory experiments. Primarily used to
-communicate with various hardware devices such as motors and pumps, to handle I/O
-between sensors, and to store data.
-%
-Experiments using \fluidpack{lab} have been carried out in the DAMTP fluid
-laboratory \cite[Cambridge, UK. cf.][]{LeclercqPartridgeAugierDalzielKerswell2016}, in ENS Lyon laboratory
-\cite[Lyon, France][]{salort2018} and at LEGI \cite[Grenoble, France][]{ISSF2016}.
-
-\item \fluidpack{image}: Scalable image processing package which implements
-various algorithms to calibrate cameras, to preprocess images, to do Particle
-Image Velocimetry (PIV) and to postprocess data.
-%
-Fluidimage is used to process images taken during experiments performed in the
-Coriolis platform at LEGI \cite[]{ISSF2016}.
-
-\item \fluidpack{foam}: Small package to load OpenFoam data and plot them.
-
-\item \fluidpack{coriolis}: Small package used to carry out experiments in the
-\href{http://www.legi.grenoble-inp.fr/web/spip.php?article757}{Coriolis platform}
-(a large rotating platform participating in the European consortiums Euhit and
-Hydralab) and open the data obtained \cite[see, for example,][]{ISSF2016}. One of
-the motivations behind creating this package is to study how to use open-source to
-create and share open-data.
-
-\end{itemize}
-
-A detailed presentation on the above packages can be found in their respective
-documentations on the web and for \fluidpack{dyn}, \fluidpack{fft} and
-\fluidpack{sim} in the respective software metapapers
-\citep[]{fluiddyn, fluidfft, fluidsim}.
-
-% \section{Conclusion}
-
-% FluidDyn is an attempt to set off collaborative dynamics based on open-source
-% development in fluid dynamics research.
-% %
-% We shall try, with this project, to explore the possibilities of open-source in
-% science and fluid dynamics, by fully exploiting the new open-source tools and
-% methods presented in the first subsections of the introduction.
-%
-% % again mention some success of community-driven subject-area research
-% % libraries? See the file ./communitydriven_research_soft.md
-%
-% The project is right now in a very preliminary stage.  Packages are actively
-% evolving with interesting features and a framework for collaborative
-% development, packaging, documenting and testing is now well set.
-% %
-% However, the community around the project is currently tiny and now, we have to
-% work on attracting users and developers since an active community is a
-% criteria for success and sustenance of an open-source project.
-%
-% Will people use the FluidDyn tools and collaborate through the project
-% FluidDyn?
-% %
-% There are clearly many challenges and potential barriers:
-% \begin{itemize}
-% \item Some habits in the community.
-% \item Knowledge and skills in the community. For example, only few people use
-% issue tracker and pull requests.
-% % from Julien Salort: bad!
-% % \item Misunderstanding between colleagues and in particular colleagues with
-% % responsibilities and power.
-% \item Lack of a business model for open-source software in science.  A good
-% quality open-source software has a cost that institutions should be willing
-% to fund and support.
-% \item Lack of recognition of the work spent in open-source, in particular for
-% precarious scientists.
-% \end{itemize}
-%
-% On our side, we also have positive points.
-% %
-% The quality of the tools we use (Python and its scientific ecosystem,
-% Mercurial, Read the Docs, Jupyter, ...) is impressive.
-% %
-% Scientific code is done to be read and to transmit ideas. To this effect, Python
-% is among the best languages today.
-% %
-% We can benefit from the dynamics of Python and of emerging subjects like deep
-% learning and the Internet of Things.
-% %
-% Finally, if we manage to gather a community of users and of developers, the
-% collective efficiency related to open-source methods and tools can be a strong
-% booster.
+FluidDyn project has adopted a decentralized
+organization with one common package and other specialized packages, to make
+maintenance easier and also the target audience (students, experimentalists
+or computational scientists) may not use the full package. FluidDyn project now
+offers the following packages: \fluidpack{dyn},
+%~\citep{fluiddyn}
+the base package; \fluidpack{fft}%
+%\citep[see ][]{fluidfft}
+, an API for efficient Fast Fourier Transforms (FFT) in sequential and in
+parallel; \fluidpack{sim}
+%~\citep[see ][]{fluidsim},
+a CFD framework; \fluidpack{lab}, a package meant for
+facilitating experiments; \fluidpack{image}, an asynchronously parallelized
+image processing package and much more.
+%
+A short introduction on the above packages can be found in their respective
+documentations on the web. The packages \fluidpack{dyn}, \fluidpack{fft}
+and \fluidpack{sim} are described in detail in their respective software
+metapapers
+\citep[]{fluiddyn, fluidfft, fluidsim} which are included in this thesis.

### 2019-05-25 chapter_00_open_science.tex
Final proofread

@@
 development is today streamlined with the emergence of Distributed Version
-Control Systems (DVCS), especially, Git and Mercurial). In web-based source
-development platforms like GitHub, Bitbucket and GitLab, DVCS repositories are
-enhanced with code-review tools, typically consisting of a issue tracker
-and pull-request mechanism.
+Control Systems (DVCS), especially, Git and Mercurial).
+%
+Version control tools are designed to work on any kind of text files, not just
+codes, so they could even be utilized to write scientific documents with
+multiple authors.
+%
+In web-based source development platforms like GitHub, Bitbucket and GitLab,
+DVCS repositories are enhanced with code-review tools, typically consisting of
+a issue tracker and pull-request mechanism.
 %
@@
 %
-Version control tools are designed to work on any kind of text files, not just
-codes, so they could even be utilized to write scientific documents with
-multiple authors.
-%
 The use of DVCS in science will lead to long term productivity benefits
@@
 the Comprehensive \TeX\ Archive Network (CTAN) model for \TeX\ packages. 
-
+%
 Python as a medium for scientific computing owes its success to a rich
@@
 \fnref{https://packaging.python.org}{``Python Packaging User Guide''} is an
-excellent guide to install, package and distribute projects.
+excellent guide to learn how to install, package and distribute projects.
 % Almost all major
@@
 dependencies. CI is specifically important for dynamical languages and
-open-source code since it allows open-source projects to welcome contributions
+open-source code since, it allows open-source projects to welcome contributions
 from other developers while maintaining a high degree of reliability.
@@
 as Riot (Matrix protocol), Gitter, and Slack have gained popularity offering
-different integrations solutions on top of instant messaging.
+different integration solutions on top of instant messaging.
 %
@@
 \fnref{https://www.python.org/dev/peps/pep-0008/}{PEP~8}, advocating for
-regularity of the code. Linters and code-formatters can evaluate adherence to
-this style.
+regularity of the code. Linters and code-formatters can evaluate and enforce
+adherence to this style.
 
@@
 
-\item \textbf{Extension support}. It is easy to interface with code written in
+\item \textbf{Extensible}. It is easy to interface with code written in
 several other languages (in particular C, C++ and Fortran).
@@
 \end{itemize}
-
+%
 Thus, Python is one of the best ``glue'' languages for fast prototyping.
@@
 %
-Maintaining and modifying existing code is also eased by the very good code
+Maintaining and modifying existing code is also facilitated by the very good code
 readability associated with the language.
@@
 Another consequence of the apparent simplicity of Python is that the learning
-curve has a nice shape. Learning Python is very easy at the beginning and the
+curve has a nice shape. Learning Python is very easy from the start and the
 complexity gradually increases as the level of the developer increases.
@@
 After a long and difficult process, the transition from Python 2 to Python 3 is
-now (201) behind us for nearly all important packages\footnote{Close to
+behind us for nearly all important packages\footnote{Close to
 \href{https://python3wos.appspot.com/}{95\% of the most downloaded packages}
-are Python 3 compatible now.}. Especially for scientific applications, we can
-now work only in Python 3. For instance, \href{https://github.com/%
+are Python 3 compatible as of 2018.}. Especially for scientific applications,
+we can now work only in Python 3. For instance, \href{https://github.com/%
 numpy/numpy/blob/master/doc/neps/dropping-python2.7-proposal.rst}{future
@@
 \end{itemize}
-
+%
 Python
@@
 % and developing scientific applications. It 
-has now a mature and powerful scientific ecosystem with well-established based
-packages (\Numpy for N-dimensional homogeneous arrays, \Scipy as the
+has now a mature and powerful scientific ecosystem with well-established
+foundational packages
+%
+(\Numpy for N-dimensional homogeneous arrays, \Scipy as the
 fundamental toolkit for scientific computing, \pack{Matplotlib} for plotting
@@
 
-Python has a very large and supportive community (see for example the
-\href{http://stackoverflow.com/tags}{Stack Overflow tags}).
-%
-But Python development is also supported by companies specialized in web
-technologies and data-driven research. However, recently, companies involved in
-scientific software development have started to become quite influential. For
-example Anaconda Inc.\ is strongly engaged in the development of tools like JIT
-Python accelerator \pack{Numba}, and parallelism framework \pack{dask}.
-Hardware giants such as Intel and Nvidia also contribute to the Python
-scientific ecosystem.
+Python has a very large and supportive community (see, for example,
+\fnref{http://stackoverflow.com/tags/python}{Stack Overflow tags}).
+%
+Python development is also supported by companies, and most of them specialize
+in web technologies and data-driven research. However, recently, companies
+involved in scientific software development have started to become quite
+influential. For example, Anaconda Inc.\ is strongly engaged in the development
+of tools like JIT Python accelerator \pack{Numba}, and parallelism framework
+\pack{dask}. Hardware giants such as Intel and Nvidia also contribute to the
+Python scientific ecosystem.
 %
@@
 	loops as much as possible and use vectorized array operations as much
-	as possible, or call \emph{foreign functions}, or use special tools to
-	build extensions using \emph{Ahead Of Time} or \emph{Just In Time}
-	compilation to speedup performance-critical code.
+	as possible. One can also call \emph{foreign functions}, or use special
+	tools to build extensions using \emph{Ahead Of Time} or \emph{Just In
+	Time} compilation to speedup performance-critical code.
 
@@
 	interpreted languages). Types are respected however, and type errors,
-	if any, would occur during runtime.
+	if any, would be raised during runtime.
 	\emph{Type hints} and \emph{static type checking} can solve this issue.
@@
 %
-There are also other projects to create extensions in C++ with
-\href{https://github.com/pybind/pybind11/}{pybind11}, and in Rust with
-\href{https://github.com/PyO3}{PyO3}.
+There are also other projects to create extensions in C++ with \pack{pybind11},
+and in Rust with \href{https:///pyo3.rs}{PyO3}.
 
@@
 %
-This two-step compilation provides room for very clever optimizations
+This two-step compilation provides room for very clever optimizations%
 \footnote{As demonstrated in this blog post by the main developer of
@@
 %
+Adding a JIT to the interpreter CPython has been notoriously difficult
+\footnote{\url{https://faster-cpython.readthedocs.io/}}. One of the
+problem seems to be the C API provided in CPython and used by many extensions.
+%
+A faster CPython with a JIT is actually not so necessary for many purposes,
+since Python extensions are enough. Therefore, not as much money and work
+has been put in accelerating Python than for example for Java and JavaScript.
+
 PyPy, an alternative interpreter written in Python, has a JIT compiler. However,
@@
 \pack{Matplotlib}, \pack{Pandas}, etc.).
-
-Adding a JIT to the interpreter CPython has been notoriously difficult
-\footnote{\url{https://faster-cpython.readthedocs.io/}}. One of the
-problem seems to be the C API provided in CPython and used by many extensions.
-%
-A faster CPython with a JIT is actually not so necessary for many purposes,
-since Python extensions are enough. Therefore, not as much money and work
-has been put in accelerating Python than for example for Java and JavaScript.
-
+%
 Another strategy is to add a JIT to CPython through an external package
@@
 Static type checkers like \href{http://mypy-lang.org/}{Mypy} and
-\href{https://github.com/google/pytype}{pytype}
-%
+\href{https://github.com/google/pytype}{pytype}%
 \footnote{We plan to add type hinting in the most important FluidDyn modules to
@@
 
-\subsubsection{Achieving computational parallelism}
+\subsubsection{Multi-core computational parallelism}
 
@@
 Note however, that threads can be (and are widely) used in Python for
-concurrency, i.e.\ to perform I/O tasks (which does not need CPU) concurrently.
+concurrency, i.e.\ to perform I/O tasks, which does not need CPU, concurrently.
 %
@@
 %
-It seems that a very important aspect in the future will be interoperability
+An important aspect in the future would very likely be interoperability
 between coexisting tools\footnote{See for example the cross-language development
 platform for in-memory data, \href{https://arrow.apache.org/}{Apache Arrow}.}.
-%
-Knowledge of Python is also an advantage in terms of employability these days.
 %
@@
 %
-Nowadays, programming is involved in all methods, and thus it is difficult to do
+Nowadays, programming is involved in all methods, and is usually difficult to do
 fluid mechanics without software development.
@@
 without the technical knowledge on how to work collectively, such codes
-are often of bad quality (compared to the information technology
-industry standard) and are basically doomed to be abandoned after their use.
+are often substandard (compared to the information technology
+industry standards) and are basically doomed to be abandoned after their use.
 %
@@
 %
-Often, languages are used for things for which they are not adapted.
+Often, languages are used for purposes which they are not adapted for.
 %
@@
 start to add up the price of the toolboxes necessary to run most codes, it
-starts to become substantially large for some institutions.  The price of
+starts to become substantial for some institutions.  The price of
 Matlab quickly escalates when deployed on a cluster with MPI, since a licence
@@
 
-The same remarks can be made for Labview. It is a graphical programming
+The same remarks are applicable for Labview. It is a graphical programming
 language and the programs are saved as binary files. As a result, it is
@@
 With some minimal effort, one can port existing Matlab codes using either
-Julia which by design uses Matlab-like syntax, or Python's \codeinline{pylab}
-package provided via \pack{matplotlib} and the command-line tool
+Julia which, by design, uses a Matlab-like syntax, or Python's \codeinline{pylab}
+package provided via \pack{Matplotlib} and the command-line tool
 \codeinline{fluidmat2py} in \pack{fluiddyn}.
@@
 that we have already discussed. Other arguments are unfair or no longer valid
-nowadays\footnote{See also \url{http://www.pyzo.org/python\_vs\_matlab.html}
-or \url{https://www.linkedin.com/pulse/matlab-vs-python-jan-rhebergen}}.
+nowadays\footnote{See also \url{http://www.pyzo.org/python\_vs\_matlab.html} or
+\url{https://www.linkedin.com/pulse/matlab-vs-python-jan-rhebergen}}.
 %
@@
 %
-It would be interesting to estimate what could be done with open-source
-solutions with a part of the amount paid by public research institutes. How
-many developers could be paid to improve the alternative open-source solutions?
+It would be interesting to estimate what open-source solutions could benefit
+from using the amount paid by public research institutes. How many
+developers could be paid to improve the alternative open-source solutions?
 %
@@
 %
-What have not been ``reinventing the wheel'', but rather have been ``standing
-on the shoulders of giants''. In other words we have utilized and extended
+We have not been ``reinventing the wheel'', but rather have been ``standing
+on the shoulders of giants''. In other words, we have utilized and extended
 existing and mature packages and libraries to create packages in FluidDyn as

### 2019-05-27 scripts/pandoc_filters.py
Image autoheight using panflute filter

@@ -1,0 +1,33 @@
+import panflute as pf
+
+
+def detect_height(url):
+    if url.endswith(".pdf"):
+        from PyPDF2 import PdfFileReader
+        with open(url, "rb") as fp:
+            pdf = PdfFileReader(fp)
+            w, h = pdf.getPage(0).mediaBox[-2:]
+    else:
+        import imageio
+        w, h, colors = imageio.imread(url).shape
+
+    height = f"{100*h/w}%"
+    return height
+
+
+def action_auto_height(elem, doc):
+    """Automatically determines height attribute for images."""
+    if isinstance(elem, pf.Image):
+        # pf.debug(elem)
+        if "width" in elem.attributes and "height" not in elem.attributes:
+            elem.attributes["height"] = detect_height(elem.url)
+
+        return elem
+
+
+def main(doc=None):
+    return pf.run_filter(action_auto_height, doc=doc)
+
+
+if __name__ == '__main__':
+    main()

### 2019-05-27 scripts/pandoc_filters.py
Set width cascade figure, move first para to end, up deps

@@ -1,12 +1,26 @@
 import panflute as pf
 
 
+def extract_image_from_pdf_page(xObject):
+    """https://gist.github.com/gstorer/f6a9f1dfe41e8e64dcf58d07afa9ab2a"""
+    xObject = xObject['/Resources']['/XObject'].getObject()
+
+    for obj in xObject:
+        o = xObject[obj]
+        if o['/Subtype'] == '/Image':
+            return o
+        else:
+            return extract_image_from_pdf_page(o)
+
+
 def detect_height(url):
     if url.endswith(".pdf"):
         from PyPDF2 import PdfFileReader
         with open(url, "rb") as fp:
             pdf = PdfFileReader(fp)
             w, h = pdf.getPage(0).mediaBox[-2:]
+            # o = extract_image_from_pdf_page(pdf.getPage(0))
+            # w, h = (o['/Width'], o['/Height'])
     else:
         import imageio
         w, h, colors = imageio.imread(url).shape

### 2019-05-27 imgs/cascade/plot.py
Add makefile imgs

@@ -89,4 +89,4 @@ for ax in ax1, ax2:
 
 fig.tight_layout()
 fig.savefig("../cascade.pdf")
-plt.show()
+# plt.show()

### 2019-05-27 imgs/dependency/plot.py

@@ -42,10 +42,23 @@ node_color = []
 for node in G.nodes():
     color = [k for k, v in colors.items() if node in v]
     if color:
-        node_color.append(color[0])
+        # node_color.append(color[0])
+        G.node[node]["value"] = color[0]
     else:
         raise ValueError("No color assigned for %s" % node)
 
-nx.draw(G, with_labels=True, node_shape="s", node_color=node_color,
-        node_size=1000,)
+#  nx.draw(G, with_labels=True, node_shape="s", node_color=node_color,
+#          node_size=1000,)
+from nxviz.plots import CircosPlot
+c = CircosPlot(
+    G,
+    node_color="value",
+    node_order="value",
+    node_labels=True,
+    # , node_color=node_color
+    fontfamily="monospace",
+)
+c.draw()
+plt.tight_layout()
+plt.savefig("../dependency.pdf")
 plt.show()

### 2019-05-27 imgs/dependency/plot_nxviz.py

@@ -1,64 +1,0 @@
-import matplotlib.pyplot as plt
-import networkx as nx
-from networkx.readwrite.adjlist import read_adjlist
-
-colors = {
-    "royalblue": (
-        "Python",
-        "Cython",
-        "Pythran",
-        "numpy.f2py",
-        "mpi4py",
-        "transonic",
-        "h5py",
-        "h5netcdf",
-        "NumPy",
-        "SciPy",
-        "IPython",
-        "Jupyter",
-        "SymPy",
-        "xarray",
-        "pandas",
-        "matplotlib",
-        "fluiddyn",
-        "fluidsim",
-        "fluidfft",
-    ),
-    "pink": (
-        "C",
-        "C++",
-        "xproject",
-        "xsimd",
-        "HDF5",
-        "FFTW",
-        "PFFT",
-    ),
-    "purple": ("Fortran", "P3DFFT"),
-    "slategrey": ("CUDA", "cuFFT", "OpenMP", "MPI", "SIMD"),
-}
-
-G = read_adjlist("graph.txt")
-node_color = []
-for node in G.nodes():
-    color = [k for k, v in colors.items() if node in v]
-    if color:
-        # node_color.append(color[0])
-        G.node[node]["value"] = color[0]
-    else:
-        raise ValueError("No color assigned for %s" % node)
-
-#  nx.draw(G, with_labels=True, node_shape="s", node_color=node_color,
-#          node_size=1000,)
-from nxviz.plots import CircosPlot
-c = CircosPlot(
-    G,
-    node_color="value",
-    node_order="value",
-    node_labels=True,
-    # , node_color=node_color
-    fontfamily="monospace",
-)
-c.draw()
-plt.tight_layout()
-plt.savefig("../dependency.pdf")
-plt.show()

### 2019-05-27 imgs/cascade/plot.py
Make imgs in correct path

@@ -1,6 +1,10 @@
 # coding: utf-8
 import matplotlib.pyplot as plt
 import numpy as np
+from pathlib import Path
+
+
+root = Path(__file__).parent.parent
 
 plt.style.use("ggplot")
 plt.rc("text", usetex=True)
@@ -88,5 +92,5 @@ for ax in ax1, ax2:
     #  ax.get_yaxis().set_visible(False)
 
 fig.tight_layout()
-fig.savefig("../cascade.pdf")
+fig.savefig(root / "cascade.pdf")
 # plt.show()

### 2019-05-27 imgs/dependency/plot.py

@@ -1,6 +1,10 @@
 import matplotlib.pyplot as plt
 import networkx as nx
 from networkx.readwrite.adjlist import read_adjlist
+from pathlib import Path
+
+
+root = Path(__file__).parent.parent
 
 colors = {
     "royalblue": (
@@ -37,7 +41,7 @@ colors = {
     "slategrey": ("CUDA", "cuFFT", "OpenMP", "MPI", "SIMD"),
 }
 
-G = read_adjlist("graph.txt")
+G = read_adjlist(root / "dependency/graph.txt")
 node_color = []
 for node in G.nodes():
     color = [k for k, v in colors.items() if node in v]
@@ -60,5 +64,5 @@ c = CircosPlot(
 )
 c.draw()
 plt.tight_layout()
-plt.savefig("../dependency.pdf")
-plt.show()
+plt.savefig(root / "dependency.pdf")
+# plt.show()

### 2019-05-28 imgs/cascade/plot.py
Fix cascade

@@ -77,17 +77,13 @@ def arrow(ax, x, y, arrowroot=0.1, exp=-3):
 arrow(ax1, 0.1, 0.8, 6, -5/3)
 arrow(ax1, 100, 80)
 
-ax1.set_ylabel("$E(k)$")
-ax1.set_xlabel("$k$")
-
-
 for ax in ax1, ax2:
     #  ax.get_xaxis().set_ticks([])
     #  ax.get_yaxis().set_ticks([])
     ax.get_xaxis().set_ticklabels([])
     ax.get_yaxis().set_ticklabels([])
-    ax.set_ylabel("$E(k)$")
-    ax.set_xlabel("$k$")
+    ax.set_ylabel(r"$E(k)$")
+    ax.set_xlabel(r"$k$")
     #  ax.get_xaxis().set_visible(False)
     #  ax.get_yaxis().set_visible(False)
 

### 2019-05-28 paper_03_toy_model/article.tex
Safe commands in pandoc preamble to allow for auto width of images, use  to remove width

@@
 \begin{figure}[h]
-\centerline{\includegraphics[angle=0,width=11.5cm]{./fig1.eps}}
+\centerline{\includegraphics[angle=0,width=8cm]{./fig1.eps}}
  \caption{Spectral energy fluxes as functions of spherical harmonic wave number $ l $, calculated from the AFES GCM.
@@
 Theories that are trying to explain the mesoscale $ k^{-5/3} $ spectrum predict that the three-dimensional structures behind the forward energy cascade are very thin and elongated.
-\citet{Lindborg:2006} and
+\citet{Lindborg2006} and
 \citet{Waite-Bartello:2004}  emphasise that  it is necessary to have a vertical resolution which is finer than the buoyancy scale, $ \l_b= u/N $, in order to simulate the downscale energy cascade.
@@
 \begin{figure}[h]
-\centerline{\includegraphics[angle=0,width=10.5cm]{./fig2.eps}}
+\centerline{\includegraphics[angle=0]{./fig2.eps}}
  \caption{Spectral energy fluxes from run 1, normalised by the time averaged total energy dissipation $ \epsilon $.  }
@@
 \begin{figure}[h]
-\centerline{\includegraphics[angle=0,width=10cm]{./fig4.eps}}
+\centerline{\includegraphics[angle=0]{./fig4.eps}}
  \caption{Wave energy spectra from simulations with different degrees of the diffusion operator.  }
@@
  \begin{figure}[h]
-\centerline{\includegraphics[angle=0,width=17cm]{./fig5.eps}}
+\centerline{\includegraphics[angle=0]{./fig5.eps}}
  \caption{Energy fluxes from run 1 (left) and run 2 (right) decomposed into different contributions. }
@@
   \begin{figure}[h]
-\centerline{\includegraphics[angle=0,width=17 cm]{./fig6.eps}}
+\centerline{\includegraphics[angle=0]{./fig6.eps}}
  \caption{Compensated wave energy spectra (left) and uncompensated vortical spectra (right) from run 1 and run 2. }
@@
 \begin{figure}[h]
-\centerline{\includegraphics[angle=0,width=17cm]{./fig7.eps}}
+\centerline{\includegraphics[angle=0]{./fig7.eps}}
  \caption{Potential enstrophy flux from run 1 (left) and run 2 (right). }
@@
 \begin{figure}[h]
-\centerline{\includegraphics[angle=0,width=18.5cm]{./fig8.pdf}}
+\centerline{\includegraphics[angle=0]{./fig8.pdf}}
 \caption{Divergence field from a simulation of the shallow water equations (left) and from run 1 of the toy model (right). }
@@
 \begin{figure}[h]
-\centerline{\includegraphics[angle=0,width=18.5cm]{./fig9.pdf}}
+\centerline{\includegraphics[angle=0]{./fig9.pdf}}
 \caption{Potential vorticity field (left) and wave field (right) from the end state of run 1.}
@@
 \begin{figure}[h]
-\centerline{\includegraphics[angle=0,width=10.5cm]{./fig10.eps}}
+\centerline{\includegraphics[angle=0]{./fig10.eps}}
  \caption{Spectral energy fluxes from run 3, normalised by the time averaged total energy dissipation $ \epsilon $. }
@@
  \begin{figure}[h]
- \centerline{\includegraphics[angle=0,width=18.5cm]{./fig12.eps}}
+ \centerline{\includegraphics[angle=0]{./fig12.eps}}
 \caption{Potential vorticity field (left) and wave field (right) from the end state of run 3.}
@@
 \begin{figure}[h]
-\centerline{\includegraphics[angle=0,width=18.5cm]{./fig13.eps}}
+\centerline{\includegraphics[angle=0]{./fig13.eps}}
 \caption{Blow up of the potential vorticity field (left) of a strong anticyclonic vortex and the corresponding wave field (right). }
@@
 In any case, as we pointed out earlier, the forward energy cascade in the GCM cannot be produced by inertia-gravity waves with frequencies close to $ f $, since it would require much finer resolution in the vertical to simulate such motions.
-Neither can it be fully accounted for by stratified turbulence as suggested by \citet{Lindborg:2006}, since that would also require much finer resolution.
+Neither can it be fully accounted for by stratified turbulence as suggested by \citet{Lindborg2006}, since that would also require much finer resolution.
 In the toy model simulation, on the other hand, we have been able to simulate a downscale energy cascade using only a single level.
@@
 %\begin{figure}[h]
-%\centerline{\includegraphics[angle=0,width=10.5cm]{./Fig12.eps}}
+%\centerline{\includegraphics[angle=0]{./Fig12.eps}}
 % \caption{Time averaged potential enstrophy flux from run 2, displayed in a log-log plot.

### 2019-05-28 paper_04_shallow_water/Latex/article.tex
Safe commands in pandoc preamble to allow for auto width of images, use  to remove width

@@
 
-We now turn to the question of how the shock width, $ \delta x $, can be estimated.
-One suggestion would be that $ \delta x $  is equal to the Kolmogorov length scale, $ \nu^{3/4}/\epsilon^{1/4} $.
-However, exact solutions of Burgers equation \cite[see for example][]{Whitham}
+We now turn to the question of how the shock ]{Whitham}
 suggests that  $ \delta x \propto \nu $ rather than $ \delta x \propto \nu ^{3/4} $.
@@
 \begin{figure}
-\centerline{\includegraphics[width=5.12in]{paper_04_shallow_water/Pyfig/fig_Emean_time}}
+\centerline{\includegraphics[]{paper_04_shallow_water/Pyfig/fig_Emean_time}}
 \caption{Space averaged energy, $E = E_K + E_A $, versus time for all
@@
 \begin{figure}
-\centerline{\includegraphics[width=5.8in]{paper_04_shallow_water/Pyfig/fig_energy_w}}
+\centerline{\includegraphics[]{paper_04_shallow_water/Pyfig/fig_energy_w}}
 \caption{Left: Mean energy in the stationary state versus $ c $ for different
@@
 \begin{figure}
-\centerline{\includegraphics[width=5.8in]{paper_04_shallow_water/Pyfig/fig_flux_struct_combined}}
+\centerline{\includegraphics[]{paper_04_shallow_water/Pyfig/fig_flux_struct_combined}}
 \caption{Left: Time averaged normalised spectral energy fluxes versus $ k/k_f
@@
 \begin{figure}
-\centerline{\includegraphics[width=3.15in]{paper_04_shallow_water/Pyfig/fig_spatiotempspectra}}
+\centerline{\includegraphics[width=3in]{paper_04_shallow_water/Pyfig/fig_spatiotempspectra}}
 \caption{Frequency spectra of KE and APE for different $ k = \mid {\bf k} \mid
@@
 \begin{figure}
-\centerline{\includegraphics[width=6.0in]{paper_04_shallow_water/Pyfig/fig_phys_fields_wave}}
+\centerline{\includegraphics[]{paper_04_shallow_water/Pyfig/fig_phys_fields_wave}}
 \caption{Left: divergence $ \bnabla \cdot {\bf u} $. Right: velocity component in the $ y $-direction: $ u_y $. Top: run W6. Bottom: run W17.  }
@@
 \begin{figure}
-\centerline{\includegraphics[width=8cm]{paper_04_shallow_water/Pyfig/fig_shock_sep.pdf}}
+\centerline{\includegraphics[width=9cm]{paper_04_shallow_water/Pyfig/fig_shock_sep.pdf}}
 \caption{Mean distance between the shocks as function of Froude number.  }
@@
 systematic deviation at the highest $ F_f $.
-\begin{figure}
+
+\begin{figure}[htp!]
 \centerline{
-\includegraphics[width=5.0in]{paper_04_shallow_water/Pyfig/fig_struct_order_246}}
+\includegraphics[]{paper_04_shallow_water/Pyfig/fig_struct_order_246}}
 \caption{From top to bottom: second, fourth and sixth order compensated and
@@
 \begin{figure}
-\centerline{\includegraphics[width=12cm]{paper_04_shallow_water/Pyfig/fig_flatness}}
+\centerline{\includegraphics[]{paper_04_shallow_water/Pyfig/fig_flatness}}
 \caption{ Top: Flatness factor of transverse velocity increments. Bottom: Ratio between transverse and longitudinal flatness factor. Left: runs with $ n=3840 $: W3,
@@
 \begin{figure}
-\centerline{\includegraphics[width=11cm]{paper_04_shallow_water/Pyfig/fig_spectra}}
+\centerline{\includegraphics[]{paper_04_shallow_water/Pyfig/fig_spectra}}
 \caption{Time averaged compensated and normalised energy spectra
@@
 \begin{figure}
-\centerline{\includegraphics[width=13cm]{paper_04_shallow_water/Pyfig/fig_diss_spectra}}
+\centerline{\includegraphics[]{paper_04_shallow_water/Pyfig/fig_diss_spectra}}
 \caption{Time averaged normalised dissipation spectra.
@@
 \begin{figure}
-\centerline{\includegraphics[width=5.8in]{paper_04_shallow_water/Pyfig/fig_energy_lap}}
+\centerline{\includegraphics[]{paper_04_shallow_water/Pyfig/fig_energy_lap}}
 \caption{Left: Mean energy in the stationary state versus $ c $ for different
@@
 \begin{figure}
-\centerline{\includegraphics[width=8cm]{paper_04_shallow_water/Pyfig/fig_shock_sep_lap.pdf}}
+\centerline{\includegraphics[width=9cm]{paper_04_shallow_water/Pyfig/fig_shock_sep_lap.pdf}}
 \caption{Mean distance between the shocks as function of Froude number.  }
@@
 \begin{figure}
-\centerline{\includegraphics[width= 8cm]{paper_04_shallow_water/Pyfig/fig_flatness_div_norm_nu.pdf}}
+\centerline{\includegraphics[width=9cm]{paper_04_shallow_water/Pyfig/fig_flatness_div_norm_nu.pdf}}
 \caption{Flatness of divergence $ \nabla \cdot {\bf u} $ times viscosity $ \nu $, versus $ F_f $. Dotted straight line indicates $ F_f^{2/3} $. }
@@
 \begin{figure}
-\centerline{\includegraphics[width=6.0in]{paper_04_shallow_water/Pyfig/fig_phys_fields_wave_lap}}
+\centerline{\includegraphics[]{paper_04_shallow_water/Pyfig/fig_phys_fields_wave_lap}}
 \caption{$ \nabla \cdot {\bf u} $ for four different runs.

### 2019-05-31 paper_05_milestone_issf/article.tex
Add files from milestone_issf paper as is

\documentclass[a4paper, 12pt, times]{article}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{natbib}
\usepackage{fancyhdr}
\usepackage{titlesec}
\usepackage[font=footnotesize]{caption}

\oddsidemargin 0mm \topmargin 0mm \textheight 247mm \textwidth 160mm
\voffset 0mm \hoffset 0mm \headheight = 0mm \headsep = 0mm
\pagestyle{fancy}
\pagenumbering{arabic}
\fancyhead{}
\cfoot{\footnotesize ${\rm VIII^{th}}$ Int. Symp. on Stratified Flows, San Diego, USA, Aug. 29 - Sept. 1, 2016}
\rfoot{\footnotesize \thepage}
\renewcommand{\headrulewidth}{0pt}
\setcounter{secnumdepth}{2}

\newfont{\tit}{ptmb at 14pt}      %% Commands to use required fonts.
\newfont{\auth}{ptmr at 12pt}
\newfont{\sect}{ptmb at 12pt}
\newfont{\subsect}{ptmb at 11pt}

\titleformat{\section}{\sect}{\thesection}{1em}{}
\titleformat{\subsection}{\subsect}{\thesubsection}{1em}{}

\newcommand{\acknowledgements}[0]{\vspace{16pt} \noindent {\sect Acknowledgements} \vspace{10pt}\\}
\newcommand{\references}[0]{\vspace{16pt} \noindent {\sect References} \vspace{10pt}\\}
\renewcommand{\abstract}[0]{\noindent {\sect Abstract} \vspace{5pt}\\}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newcommand{\R}{\mathcal{R}}

\newcommand{\eps}{\varepsilon}
\newcommand{\epsK}{\varepsilon_{\!\scriptscriptstyle K}}
\newcommand{\epsA}{\varepsilon_{\!\scriptscriptstyle A}}

\newcommand{\pathfigures}{/fsnet/project/coriolis/2016/16MILESTONE/figure}
% \newcommand{\pathfigures}{/home/pierre/16MILESTONE/Figures}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}
\begin{center}

{\tit First report of the MILESTONE experiment: strongly stratified turbulence
and mixing efficiency in the Coriolis platform}\\ %% Title

\vspace{10pt}

{\auth \underline{Antoine Campagne}$^{1}$, Henrik Alfredsson$^{2}$, R\'emi Chassagne$^{1}$, Diane
Micard$^{3}$, Nicolas Mordant$^{1}$, Antonio Segalini$^{2}$, Joel
Sommeria$^{1}$, Samuel Viboud$^{1}$, Ashwin Vishnu Mohanan$^{2}$,\\ Erik
Lindborg$^{2}$ and Pierre Augier$^{1}$}\\ %%Authors

\vspace{10pt}

%%Addresses & emails
{$^{1}$LEGI, Universit\'{e} Grenoble Alpes, CNRS, France}\\
{$^{2}$Royal Institute of Technology (KTH), Department of Mechanics, Stockholm, Sweden}\\
{$^{3}$LMFA, \'Ecole Centrale de Lyon, France}\\

Antoine.Campagne@univ-grenoble-alpes.fr
\end{center}

\abstract

\noindent Strongly stratified turbulence is a possible interpretation of oceanic and
atmospheric measurements. However, this regime has never been produced in a
laboratory experiment because of the two conditions of very small horizontal
Froude number $F_h$ and large buoyancy Reynolds number $\R$ which require a verily large experimental facility. We
present a new attempt to study strongly stratified turbulence experimentally in
the Coriolis platform. The flow is forced by a slow periodic movement of an
array of six vertical cylinders of 25\ cm diameter with a mesh of 75\ cm.
%
Five cameras are used for 3D-2C scanned horizontal particles image velocimetry
(PIV) and stereo 2D vertical PIV.  Five density-temperature probes are used to
measure vertical and horizontal profiles and signals at fixed positions.
%
The first preliminary results indicate that we manage to produce strongly
stratified turbulence at very small $F_h$ and large $\R$ in a laboratory
experiment.

\section{Introduction}

\noindent
Our understanding of the dynamics of turbulence strongly influenced by a stable
density stratification has been deeply changed at the beginning of the 21st
century.
%
Studies based on large-resolution numerical simulations have shown that there
is a strongly stratified regime associated with a downscale energy cascade
\cite[]{RileyDeBruynKops2003, Lindborg2006}.
%
It has been understood that the stratified turbulence is characterised by two
important non-dimensional numbers and that the regime called ``strongly
stratified turbulence'' is obtained only for very small horizontal Froude
number and large buoyancy Reynolds number \cite[]{Billant2001,
BrethouwerBillantLindborg2007},
\begin{equation}
 F_h = \frac{\epsK}{N U^2} \ll 1,\quad \R = \frac{\epsK}{\nu N^2} > 10,
\label{eq:FhR}
\end{equation}
where $\epsK$ is the kinetic energy dissipation, $N$ the
Brunt-V\"{a}iss\"{a}l\"{a} frequency, $U$ the root-mean-square horizontal
velocity and $\nu$ the kinematic viscosity.
%
\cite{RileyLindborg2008} showed that strongly stratified turbulence is a
possible interpretation for some geophysical turbulent measurements.

\noindent In a laboratory experiment, it is very difficult to fulfill the two conditions $F_h \ll 1$ and $\R > 10$ at the same time, since this requires very large
facilities.
%
Most of stratified flows studied experimentally
correspond to a regime very different than strongly stratified geophysical
flows \cite[see for example][]{PraudFinchamSommeria2005,
AugierBillantNegrettiChomaz2013, AugierBillantChomaz2015}.
%
\cite{BrethouwerLindborg2009} and \cite{Maffioli2016} investigated numerically
the mixing produced by strongly stratified turbulence and showed that at very
small horizontal Froude number ($F_h < 0.03$) and large buoyancy Reynolds
number, the mixing efficiency tends to be independent of these parameters with
a mixing coefficient $\Gamma = \epsA/\epsK$ of the order of 0.35, where $\epsA$
and $\epsK$ are the dissipation rates of available potential energy and kinetic
energy, respectively. This value is nearly twice larger that the canonical
value 0.2 used for mixing parametrization in oceanic models. Moreover, the
effect of rotation on mixing efficiency of strongly stratified
turbulence has to be investigated.

\noindent We present the first report of the EuHIT project experiment (called
MILESTONE) dedicated to better understand these issues by careful measurements
of strongly stratified turbulence in a large-scale experiment in the Coriolis
platform.

\section{Experimental setup}

\begin{figure}[htb]
\centerline{
\includegraphics[width=10cm]{Figures/scheme_exp_grid_MILESTONE_Euhit.pdf}}
\vspace{5mm}
\centerline{
\includegraphics[width=80mm]{Figures/GOPR1453.pdf}
\includegraphics[width=80mm]{Figures/fig_movement_carriage.pdf}}
\caption{Top: scheme of the experimental setup (see text for
description). Bottom left: photography of the experiment. Two of our cameras
are visible (highlighted in red), the third one being below the window
highlighted by the black rectangle. Three of our five probes are highlighted in
blue, $P_c$ is embarked on the carriage and $P_p^1$ and $P_p^2$ are attached to
profilers. The vertical PIV field of view and its associated cameras are not
visible in the photo. Bottom right: chronogram of the position of the carriage
$x_c(t)$.}
\label{fig:exp}
\end{figure}

\begin{figure}
\centerline{\includegraphics[width=16cm]{Figures/fig_profiles.pdf}}
\vspace{-2mm}
\caption{Left: density profiles for $N=0.8$ rad/s and $U_c = 6$ cm/s. Right:
variations of density with respect to the profile at rest
$\Delta\rho=\rho-\rho(t/T=-0.13)$.}
\label{fig_profiles}
\end{figure}

\subsection{Overview of the experiment}
\noindent The experimental setup, installed in the large Coriolis platform, is
presented in figure~\ref{fig:exp}: a tank with a rectangle base of
9$\times$6~m$^2$ area is filled with an approximately linear salt
stratification of $81$~cm depth. The flow is generated with an oscillating
comb consisting of 6 vertical cylinders of $25$~cm diameter attached to a
carriage with a mesh of $M=75$~cm. We impose on the carriage the following
periodic motion of period $T$ (see figure~\ref{fig:ken} for a chronogram of its
position $x_c$): it accelerates with a constant acceleration over 25~cm to a
velocity $U_c$, travels at the uniform speed a distance of 550~cm and
decelerates with a constant deceleration over 25~cm.  It them comes back to its
initial position with the opposite movement and immediately restarts this
cycle. In the following, we use cartesian coordinates with the origin
centered with respect to the vertical walls and at the bottom of the tank. The
vertical coordinate is given by $z$ and the horizontal coordinate along the
direction of displacement of the carriage by $x$ such that $(x,y,z)$ form right
handed coordinates.

\noindent We performed a large number of experiments varying the velocity $U_c$
of the carriage and the linear stratification characterized by the
Brunt-V\"aiss\"al\"a frequency
\begin{equation}
 N =  \sqrt{-\frac{g}{\langle\rho\rangle_z}~\frac{\partial \rho}{\partial z}},
\end{equation}
where $g$ denotes the gravitational acceleration, $\rho$ the volumetric mass
density of the fluid and $\langle\cdot\rangle_z$ the vertical average. In
practice, $N$ is evaluated from the density profile $\rho(z)$ measured at rest,
the slope $\partial\rho/\partial z$ being computed by a linear fit over
$\rho(z)$ on 20~cm around the mid-height of water. A single experiment can be
characterized by two dimensionless parameters based on the forcing control
parameters, a horizontal Froude number and a buoyancy Reynolds number, defined
as
\begin{equation}
 F_{hc} = \frac{U_c}{NM},\quad \R_c = \frac{{U_c}^3}{\nu N^2 M}.
\label{eq:c}
\end{equation}

\input{table.tex}

\noindent Table~\ref{table} presents the parameters spanned with the
experiments performed without rotation. Thanks to the large dimensions of our
facility, we obtain small to moderate Froude numbers $F_{hc}\in[0.03,~ 0.53]$
and small to large buoyancy Reynolds numbers $\R_c\in[15,~ 34133]$.
%
\noindent Analysis on scaling laws with respect to $F_{hc}$ and $\mathcal{R}_c$
is of course needed for a careful understanding of the flows. However, as this
project is at its earliest stage, a thorough analysis has not yet been
done. For the sake of simplicity, we only present here preliminary results from
the case for $F_{hc} = 0.1$ and $\mathcal{R}_c=450$ (highlighted by bold font
in table~\ref{table}).

\subsection{Measurements}
\noindent This experiment benefits from a well-furnished instrumentation which is
described in the following.

\paragraph{Density measurements}
Five calibrated conductimetric probes $P_i^j$ measure the volumetric
mass density (see figure~\ref{fig:exp}).
%
The probe $P_c$ is embarked on the carriage at the position $(x_c(t)+0.48$,
0.71, 0.395)~m. Since the position of this probe breaks the symetry with
respect to the $x$-axis, its signal has to be separated into two phases. When
the probe is downstream, the measurement can be interpreted as the
statistically stationary signal of a forced turbulence. When the probe is
upstream, we can suppose that the probe is not affected by the
cylinders. The measurement can then be interpreted as the signal of the
decaying turbulence that has been generated during the previous phase.

\noindent The probes $P_f^1$ and $P_f^2$ are respectively placed near the bottom
$(2.73, 2.18, 0.025)$~m and near the surface $(2.73, 2.18, 0.785)$~m in order
to measure the time evolution of the mixed layers.
%
\noindent Finally, the probes $P_p^1$ and $P_p^2$ are fixed on well controlled
vertical profilers at positions $(-1.01, 2.24, z_p(t))$ and
$(1.66, 2.14, z_p(t))$, respectively. The short time response of the probes and
the rapidity of the profilers allow us to get well resolved vertical density
profiles over 70~cm height on a time of $8$~s. Figure~\ref{fig_profiles}
represents three density profiles during an experiment consisting of 7 periods
of oscillation of the carriage. The blue line corresponds to the background
profile with fluid at rest. It shows a well defined linear dependence below  $z=65$~cm
and a mixed layer near the surface. The expected mixed layer at the bottom is not present here since the probes does not approach enough the bottom of the
tank.
%
Profiles at later times show density fluctuations with respect to the
background profile of the order of 1~kg/m$^3$ which corresponds to a
typical vertical displacement of isopycnals of 1~cm. In the profile
for $t/T = 9.56$ (yellow curves), we see that part of these fluctuations persist a long
time after the carriage stops. Importantly, we see that we can easily measure
the density variation in time at the surface that will allow us to compute
the mixing efficiency for every experiment.

\paragraph{PIV measurements}
 We measure vertical and horizontal velocity fields using two PIV
systems. The vertical measurement domain is a $0.45 \times 0.45$~m$^2$ vertical
square with a resolution of 5~mm centered at the position
$(1.88, 2.20, 0.35)$~m. We acquire up to 1300 images per period of oscillation
$T$ with a $1024^2$ pixel camera.

\noindent Horizontal fields are measured with 2D scanned PIV system with a
rapid vertically oscillating horizontal laser sheet. The vertical position of
the laser sheet is given by a staircase profile with up to 11 levels on a time
less than 100~ms. Oscillation of the laser sheet is achieved by an
oscillating mirror placed 6~m from the region of interest. Half of the depth can be scanned with quasi-horizontal laser sheets, with the angle of the
laser sheet with respect to the horizontal being less than $1.5^{\circ}$. In
practice, two sets of measurement are done: one with a small scan amplitude
(five planes with $z\in[0.483,~0.514]$~m) in order to resolve small vertical
scale structures. The second set is performed with a large scan amplitude
(typically eight planes with $z\in[0.252,~,0.554]$~m) that provides more
vertically decorrelated fields.
%
Images are acquired by three $2560\times2160$ pixel cameras synchronized with
the oscillating mirror, each camera looking on three different regions of
interest (see figure~\ref{fig:exp}). Two of them are centered at $y=0$ with a
large and a small field. The larger field ($2.2\times2.5$~m$^2$) is acquired
from the top through the water surface and the small field
($1.18\times0.53$~m$^2$) through a glass from the bottom of the tank. The third
view ($1.3\times1.0 $~m$^2$) is taken at the vicinity of a side wall. We
acquire up to 1300 images per vertical level of the laser sheet and per period
of the movement of the cylinders.

\noindent Acquisition of images begins after one period of oscillation of the
carriage and ends one period after the carriage stops. The forcing consists of
3 periods of oscillation for the small amplitude scan case and 12 periods for
large scan amplitude case.

\noindent Figure~\ref{fig:field} shows instantaneous horizontal and vertical velocity
fields for $t/T = [$1.31, 1.5, 1.68$]$. 
%
The resolution is equal to $5.3$~mm for the vertical cross-section and $2.3$~cm
for the horizontal cross-section.
%
Note that, as the turbulence is inhomogeneous and positions of measurement of
horizontal and vertical fields are different, these former fields cannot be
mutually described. Just after the carriage quit the field of view of the
camera ($t/T=1.31$), horizontal field exhibits energetic vortices of typical
sizes of $M$. At larger times, we observe the persistence of large horizontal
scale structures. Furthermore, vertical fields show strongly horizontally
elongated structures, which is a robust feature of turbulence in stratified
fluids.

\begin{figure}[htb]
\centerline{
\includegraphics[width=5.18cm]{\pathfigures/exp21/vh_400.pdf}
\includegraphics[width=4.3cm]{\pathfigures/exp21/vh_655.pdf}
\includegraphics[width=5.9cm]{\pathfigures/exp21/vh_890.pdf}}
\vspace{0mm}
\centerline{
\includegraphics[width=5.2cm]{\pathfigures/exp21/vv_890.pdf}
\includegraphics[width=4.35cm]{\pathfigures/exp21/vv_655.pdf}
\includegraphics[width=5.8cm]{\pathfigures/exp21/vv_400.pdf}
}
\vspace{-2mm}
\caption{Instantaneous horizontal large fields (top, $z=40$~cm) and vertical
fields (bottom) for $t/T = [1.31,~1.5,~1.68]$, $F_{hc} = 0.1$ and
$\mathcal{R}_c=450$. Background colors indicate the norm of the 2D field
normalized by $U_c$, which is $(u_x^2 + u_y^2)/U_c$ and $(u_x^2 + u_z^2)/U_c$,
respectively.}
\label{fig:field}
\end{figure}

\section{Preliminary results for $F_{hc} = 0.1$ and $\mathcal{R}_c=450$}

With this setup, energy is injected locally by an oscillating comb of
cylinders. Turbulence in a given region of interest is therefore inhomogeneous
and unstationary. Evolution of kinetic energy with time is then an important
feature of this flow. Figure~\ref{fig:ken} represents the normalized horizontal
kinetic energy $\langle u_x^2 + u_y^2 \rangle/2U_c^2$ as a function of the
fractional part of normalized time $\{t/T\}$ for 5 particular periods, with
$\langle\cdot\rangle$ denoting 3D spatial average. Kinetic energy is computed
from large horizontal PIV fields for each time when the carriage does not
appear in the field of view of the camera. The carriage is stopped at the end
of the 12th period. The flow undergoes successive turbulent decays. The energy
injects is of the order of $U^2 \simeq 0.08 U_c^2$.
%
We estimate the mean kinetic energy dissipation rate to
$\epsK \simeq 2 \times 10^{-2} U_c^3/M$ by computing the time average of
$-\partial\langle u_x^2+u_y^2\rangle/\partial t$ over $0.3 < \{t/T\} < 0.35$.

\begin{figure}[htb]
\centerline{\includegraphics[width=9.cm]%
{\pathfigures/exp28/ken_fct_t_T_exp28.pdf}}
\vspace{-2mm}
\caption{Spatial average of normalized horizontal kinetic energy
$\langle u_x^2 + u_y^2\rangle/2U_c^2$ as a function of the fractional part of
the normalized time $\{t/T\}$ for different periods of measurement and for
$F_{hc} = 0.1$ and $\mathcal{R}_c=450$. Times where no data appears corresponds
to those when the carriage is in the field of view and hides a part of the
field of view. The gray line represents the position of the carriage $x_c$ and
the three red circles mark the times chosen for the figure~\ref{fig:field}. The
pair of symbols $\vartriangleleft-\vartriangleright$ delimits the ranges of
time where the carriage appears in the field of view.}
\label{fig:ken}
\end{figure}

\noindent We now characterize the distribution of energy among scales by second-order
structure function, defined as
\begin{eqnarray}
S_h(r) = (S_{xx} + S_{yy} + S_{xy} + S_{yx})/2,\\
\text{with} \quad S_{ij} = \langle\overline{(u_i({\bf x}+r{\bf e}_j)-u_i({\bf x}))^2}\rangle,
\end{eqnarray}
and $\overline{\cdot}$ time average when the carriage is moving. For a
developed turbulence, and for scales larger than the dissipative scale and
lower than injection scale, we expect
$S_h(r) = C (\varepsilon r)^{2/3} = C' (rU_c^3/M)^{2/3}$, with $C$ a coefficient
approximately equal to 6 \cite[see figure 15 in][]{AugierBillantChomaz2015}.
\begin{figure}[htb]
\centerline{\includegraphics[height=7.5cm]%
{\pathfigures/exp28/normalized_S2_exp28.pdf}}
\vspace{-2mm}
\caption{Normalized second-order structure $S_h$ function as a function of
$r/M$ for $F_{hc} = 0.1$ and $\mathcal{R}_c=450$.}
\label{fig:S2}
\end{figure}
Figure~\ref{fig:S2} represents normalized second-order structure function
computed from large horizontal fields for a 12 periods forcing
experiment. Interestingly, the curve exhibits a plateau corresponding to a
$r^{2/3}$ scaling law on half a decade at a value going from $0.4$ at
$\{t/T\} = 0.31$ to $0.2$ at $\{t/T\} = 0.31$. Supposing that this range
corresponds to an inertial range, we evaluate a surrogate of the mean
dissipation rate at $\epsK \sim 0.01 U_c^3/M$. This order of magnitude is
consistent with the previous estimation by direct computation of the energy
decay.

\noindent By extrapolation of the energy dissipation rate and energy injected
at the vicinity of the carriage, we define surrogates of the turbulent
horizontal Froude and buoyancy Reynolds numbers based on the measured flow as
\begin{equation}
F_{he} \simeq \frac{\epsK}{N U^2} \simeq 0.25 F_{hc},\quad 
\R_e \simeq \frac{\epsK}{\nu N^2} \simeq 2\times10^{-2} \R_c.
\label{eq:e}
\end{equation}
For this case for $F_{hc} = 0.1$ and $\mathcal{R}_c=450$, we find
$F_{he} \simeq 0.02$ and $\mathcal{R}_e \simeq 9$. These orders of magnitude
correspond to flows for which a downscale energy cascade and horizontal spectra
with a -5/3 scaling law have been observed numerically
\cite[]{BrethouwerBillantLindborg2007, AugierBillantChomaz2015}.
%
This strongly indicates that the $r^{2/3}$ scaling law for $S_h$ is indeed
related to strongly stratified turbulence.

\noindent The values of these dimensionless parameters given in table~\ref{table} for the
other cases show that we were able with this experiment setup to span very
interesting ranges of parameters, from very strongly stratified
($F_{he} \simeq 0.01$) to weakly stratified ($F_{he} \simeq 0.1$) and from
small to large buoyancy Reynolds numbers.
%
Although the analysis of the data is still at a very preliminary stage, the
reached values of the turbulent numbers $F_{he}$ and $\R_e$ will enable us to
address the question of mixing efficiency in the regime relevant of the ocean
dynamics.

\section*{Acknowledgements}

This project has received funding from the foundation Simone et Cino Del Duca de
l'Institut de France and the European Research Council (ERC) under the European
Union's Horizon 2020 research and innovation program (grant agreement No
647018-WATU and Euhit consortium).
%
We thank Mile Kusulja, Thomas Valran, Julie Germinario and Gabriel Moreau for
their help to design, build and carry out this experiment.

\references
\def\refname{}
\vspace{-0.9in}
\bibliographystyle{apalike}
\bibliography{references}

\end{document}

### 2019-05-31 paper_05_milestone_issf/table.tex
Add files from milestone_issf paper as is

\begin{table}
\caption{Parameters explored in our experiments without system rotation.  The
line written in bold font indicates the experiment described in the last
section of the present paper. $F_{hc}$ and $\R_c$ are based on the carriage
velocity $U_c$ and the mesh $M$ as defined in equations (\ref{eq:c}).  $F_{he}$
and $\R_e$ are estimations of the turbulent Froude and buoyancy Reynolds
numbers as defined in equations (\ref{eq:FhR}) and (\ref{eq:e}), i.e.\ with the
estimations $E_K \simeq 0.08 U_c^2$ and $\epsK \simeq 0.02 {U_c}^3/M$.}
\centerline{\begin{tabular}{| c | c | r | r | r | r |}
\hline
{$N$~(rad/s)} & {$U_c$~(cm/s)} & {\bf $F_{hc}$} & {\bf $\R_c$} &
{\bf $F_{he}$} & {\bf $\R_e$} \\
\hline
0 & [2, 4, 6,  &  $\infty$ & $\infty$ &  $\infty$ & $\infty$  \\
&  8, 12, 16] & ~~~~~~~~~~ & ~~~~~~~~~~ & ~~~~~~~~~~ & ~~~~~~~~~~\\
\hline
    &    2 & 0.07 &    67 &  0.02 &   1.3 \\ \cline{2-6}
    &    4 & 0.13 &   533 &  0.03 &  10.7 \\ \cline{2-6}
0.4 &    6 & 0.20 &  1800 &  0.05 &  36.0 \\ \cline{2-6}
    &    8 & 0.27 &  4267 &  0.07 &  85.3 \\ \cline{2-6}
    &   12 & 0.40 & 14400 &  0.10 & 288.0 \\ \cline{2-6}
    &   16 & 0.53 & 34133 &  0.13 & 682.7 \\
\hline
    &    2 & 0.03 &    17 &  0.01 &   0.3 \\ \cline{2-6}
    &    4 & 0.07 &   133 &  0.02 &   2.7 \\ \cline{2-6}
0.8 & {\bf 6.00} & {\bf 0.10} & {\bf  450} & {\bf  0.02} & {\bf  9.0} \\ \cline{2-6}
    &    8 & 0.13 &  1067 &  0.03 &  21.3 \\ \cline{2-6}
    &   12 & 0.20 &  3600 &  0.05 &  72.0 \\ \cline{2-6}
    &   16 & 0.27 &  8533 &  0.07 & 170.7 \\
\hline
\end{tabular}}
\label{table}
\end{table}

### 2019-05-31 paper_05_milestone_issf/article.tex
Comment out paper 5 template boilerplate

@@
-\documentclass[a4paper, 12pt, times]{article}
-\usepackage{graphicx}
-\usepackage{amsmath}
-\usepackage{amssymb}
-\usepackage{natbib}
-\usepackage{fancyhdr}
-\usepackage{titlesec}
-\usepackage[font=footnotesize]{caption}
-
-\oddsidemargin 0mm \topmargin 0mm \textheight 247mm \textwidth 160mm
-\voffset 0mm \hoffset 0mm \headheight = 0mm \headsep = 0mm
-\pagestyle{fancy}
-\pagenumbering{arabic}
-\fancyhead{}
-\cfoot{\footnotesize ${\rm VIII^{th}}$ Int. Symp. on Stratified Flows, San Diego, USA, Aug. 29 - Sept. 1, 2016}
-\rfoot{\footnotesize \thepage}
-\renewcommand{\headrulewidth}{0pt}
-\setcounter{secnumdepth}{2}
-
-\newfont{\tit}{ptmb at 14pt}      %% Commands to use required fonts.
-\newfont{\auth}{ptmr at 12pt}
-\newfont{\sect}{ptmb at 12pt}
-\newfont{\subsect}{ptmb at 11pt}
-
-\titleformat{\section}{\sect}{\thesection}{1em}{}
-\titleformat{\subsection}{\subsect}{\thesubsection}{1em}{}
-
-\newcommand{\acknowledgements}[0]{\vspace{16pt} \noindent {\sect Acknowledgements} \vspace{10pt}\\}
-\newcommand{\references}[0]{\vspace{16pt} \noindent {\sect References} \vspace{10pt}\\}
-\renewcommand{\abstract}[0]{\noindent {\sect Abstract} \vspace{5pt}\\}
+%
+% \documentclass[a4paper, 12pt, times]{article}
+% \usepackage{graphicx}
+% \usepackage{amsmath}
+% \usepackage{amssymb}
+% \usepackage{natbib}
+% \usepackage{fancyhdr}
+% \usepackage{titlesec}
+% \usepackage[font=footnotesize]{caption}
+%
+% \oddsidemargin 0mm \topmargin 0mm \textheight 247mm \textwidth 160mm
+% \voffset 0mm \hoffset 0mm \headheight = 0mm \headsep = 0mm
+% \pagestyle{fancy}
+% \pagenumbering{arabic}
+% \fancyhead{}
+% \cfoot{\footnotesize ${\rm VIII^{th}}$ Int. Symp. on Stratified Flows, San Diego, USA, Aug. 29 - Sept. 1, 2016}
+% \rfoot{\footnotesize \thepage}
+% \renewcommand{\headrulewidth}{0pt}
+% \setcounter{secnumdepth}{2}
+%
+% \newfont{\tit}{ptmb at 14pt}      %% Commands to use required fonts.
+% \newfont{\auth}{ptmr at 12pt}
+% \newfont{\sect}{ptmb at 12pt}
+% \newfont{\subsect}{ptmb at 11pt}
+%
+% \titleformat{\section}{\sect}{\thesection}{1em}{}
+% \titleformat{\subsection}{\subsect}{\thesubsection}{1em}{}
+%
+% \newcommand{\acknowledgements}[0]{\vspace{16pt} \noindent {\sect Acknowledgements} \vspace{10pt}\\}
+% \newcommand{\references}[0]{\vspace{16pt} \noindent {\sect References} \vspace{10pt}\\}
+% \renewcommand{\abstract}[0]{\noindent {\sect Abstract} \vspace{5pt}\\}
 
@@
 
-\newcommand{\R}{\mathcal{R}}
-
-\newcommand{\eps}{\varepsilon}
-\newcommand{\epsK}{\varepsilon_{\!\scriptscriptstyle K}}
-\newcommand{\epsA}{\varepsilon_{\!\scriptscriptstyle A}}
-
-\newcommand{\pathfigures}{/fsnet/project/coriolis/2016/16MILESTONE/figure}
-% \newcommand{\pathfigures}{/home/pierre/16MILESTONE/Figures}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{document}
-\begin{center}
-
-{\tit First report of the MILESTONE experiment: strongly stratified turbulence
-and mixing efficiency in the Coriolis platform}\\ %% Title
-
-\vspace{10pt}
-
-{\auth \underline{Antoine Campagne}$^{1}$, Henrik Alfredsson$^{2}$, R\'emi Chassagne$^{1}$, Diane
-Micard$^{3}$, Nicolas Mordant$^{1}$, Antonio Segalini$^{2}$, Joel
-Sommeria$^{1}$, Samuel Viboud$^{1}$, Ashwin Vishnu Mohanan$^{2}$,\\ Erik
-Lindborg$^{2}$ and Pierre Augier$^{1}$}\\ %%Authors
-
-\vspace{10pt}
-
-%%Addresses & emails
-{$^{1}$LEGI, Universit\'{e} Grenoble Alpes, CNRS, France}\\
-{$^{2}$Royal Institute of Technology (KTH), Department of Mechanics, Stockholm, Sweden}\\
-{$^{3}$LMFA, \'Ecole Centrale de Lyon, France}\\
-
-Antoine.Campagne@univ-grenoble-alpes.fr
-\end{center}
-
-\abstract
-
-\noindent Strongly stratified turbulence is a possible interpretation of oceanic and
-atmospheric measurements. However, this regime has never been produced in a
-laboratory experiment because of the two conditions of very small horizontal
-Froude number $F_h$ and large buoyancy Reynolds number $\R$ which require a verily large experimental facility. We
-present a new attempt to study strongly stratified turbulence experimentally in
-the Coriolis platform. The flow is forced by a slow periodic movement of an
-array of six vertical cylinders of 25\ cm diameter with a mesh of 75\ cm.
+% \newcommand{\R}{\mathcal{R}}
 %
-Five cameras are used for 3D-2C scanned horizontal particles image velocimetry
-(PIV) and stereo 2D vertical PIV.  Five density-temperature probes are used to
-measure vertical and horizontal profiles and signals at fixed positions.
+% \newcommand{\eps}{\varepsilon}
+% \newcommand{\epsK}{\varepsilon_{\!\scriptscriptstyle K}}
+% \newcommand{\epsA}{\varepsilon_{\!\scriptscriptstyle A}}
 %
-The first preliminary results indicate that we manage to produce strongly
-stratified turbulence at very small $F_h$ and large $\R$ in a laboratory
-experiment.
+% \newcommand{\pathfigures}{/fsnet/project/coriolis/2016/16MILESTONE/figure}
+% % \newcommand{\pathfigures}{/home/pierre/16MILESTONE/Figures}
+%
+% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+% \begin{document}
+% \begin{center}
+%
+% {\tit First report of the MILESTONE experiment: strongly stratified turbulence
+% and mixing efficiency in the Coriolis platform}\\ %% Title
+%
+% \vspace{10pt}
+%
+% {\auth \underline{Antoine Campagne}$^{1}$, Henrik Alfredsson$^{2}$, R\'emi Chassagne$^{1}$, Diane
+% Micard$^{3}$, Nicolas Mordant$^{1}$, Antonio Segalini$^{2}$, Joel
+% Sommeria$^{1}$, Samuel Viboud$^{1}$, Ashwin Vishnu Mohanan$^{2}$,\\ Erik
+% Lindborg$^{2}$ and Pierre Augier$^{1}$}\\ %%Authors
+%
+%
+% \vspace{10pt}
+%
+% %%Addresses & emails
+% {$^{1}$LEGI, Universit\'{e} Grenoble Alpes, CNRS, France}\\
+% {$^{2}$Royal Institute of Technology (KTH), Department of Mechanics, Stockholm, Sweden}\\
+% {$^{3}$LMFA, \'Ecole Centrale de Lyon, France}\\
+%
+% Antoine.Campagne@univ-grenoble-alpes.fr
+% \end{center}
+%
+% \abstract
+%
+% \noindent Strongly stratified turbulence is a possible interpretation of oceanic and
+% atmospheric measurements. However, this regime has never been produced in a
+% laboratory experiment because of the two conditions of very small horizontal
+% Froude number $F_h$ and large buoyancy Reynolds number $\R$ which require a verily large experimental facility. We
+% present a new attempt to study strongly stratified turbulence experimentally in
+% the Coriolis platform. The flow is forced by a slow periodic movement of an
+% array of six vertical cylinders of 25\ cm diameter with a mesh of 75\ cm.
+% %
+% Five cameras are used for 3D-2C scanned horizontal particles image velocimetry
+% (PIV) and stereo 2D vertical PIV.  Five density-temperature probes are used to
+% measure vertical and horizontal profiles and signals at fixed positions.
+% %
+% The first preliminary results indicate that we manage to produce strongly
+% stratified turbulence at very small $F_h$ and large $\R$ in a laboratory
+% experiment.
 
@@
 
-\references
-\def\refname{}
-\vspace{-0.9in}
-\bibliographystyle{apalike}
-\bibliography{references}
-
-\end{document}
+% \references
+% \def\refname{}
+% \vspace{-0.9in}
+% \bibliographystyle{apalike}
+% \bibliography{references}
+%
+%
+% \end{document}

### 2019-06-03 templates/utils_render.py
Fix for long author list add et al.

@@ -50,9 +50,17 @@ def get_options_from_yaml(path):
     if "dir" not in options or options["dir"] is None:
         options["dir"] = path.parent
 
-    options["authors_short"] = [
-        name_in_initials(name) for name in options["authors"]
-    ]
+    if len(options["authors"]) < 4:
+        options["authors_short"] = [
+            name_in_initials(name) for name in options["authors"][:-1]
+        ]
+        options["authors_short"].extend([
+            r'\&',
+            name_in_initials(options["authors"][-1])
+        ])
+    else:
+        options["authors_short"] = [name_in_initials(options["authors"][0]), "et al."]
+
 
     uniq_affiliations = list(set(options["affiliations"]))  # unique
     options["authors_aff"] = []
@@ -63,7 +71,6 @@ def get_options_from_yaml(path):
     options["enum_affiliations"] = [
         f"$^{i+1}$ {aff}" for i, aff in enumerate(uniq_affiliations)
     ]
-    
 
     return options
 

### 2019-06-03 paper_04_shallow_water/Latex/article.tex

@@
 
-We now turn to the question of how the shock ]{Whitham}
+We now turn to the question of how the shock width, $ \delta x $, can be estimated.
+One suggestion would be that $ \delta x $  is equal to the Kolmogorov length scale, $ \nu^{3/4}/\epsilon^{1/4} $.
+However, exact solutions of Burgers equation \cite[see for example][]{Whitham}
 suggests that  $ \delta x \propto \nu $ rather than $ \delta x \propto \nu ^{3/4} $.

### 2019-06-04 chapter_00_open_science.tex
Revise chapter opensci

@@
 track was pursued -- i.e. to create a handful of easy to maintain, rigorously
-tested, and most-importantly, open-source scientific packages which does not
+tested, and most-importantly, open-source scientific packages which do not
 compromise on performance.
@@
 %
-Several good and modern coding practices were adopted while developing these
-packages.
-%
-An understanding of the design choices and the workflow involved would be of
-general interest to the reader.
-
-FluidDyn packages are primarily written in Python language. Incidentally, it is
-the defining characteristics of Python which make the eventual success of such
-a project realizable.
-%
-Python is one of the most important tool in recent open-source dynamics,
+Several modern coding practices were adopted while developing these packages.
+
+The FluidDyn packages are primarily written in Python language. Incidentally,
+it is the defining characteristics of Python which make the eventual success of
+such a project realizable.
+%
+Python is one of the most important tools in recent open-source dynamics,
 particularly in science.
 %
-The popularity of the language in fluid mechanics is growing, but is yet
+The popularity of the language in fluid mechanics is growing, but Python is yet
 to reach widespread adoption as the language of choice.
@@
 Computational sciences, and especially computational fluid dynamics (CFD) as a
-discipline, have flourished a lot in the latter half of the twentieth century.
-Incidentally the origin of this discipline can be traced back to the ideas in
+discipline, have flourished in the latter half of the twentieth century.
+Incidentally, the origin of this discipline can be traced back to the ideas in
 \citet{richardson_weather_1922} to achieve numerical weather prediction.
-Following the arrival of first electronic general-purpose computer, ENIAC, the
-first simulations were made using a simple barotropic model to make 24 hour
-forecasts \citep{lynch_richardson_2010}. In those early days when the
-availability of computers was limited, such methods were neither widespread nor
-feasible. Scientific investigations were dominated by theoretical and
-experimental methods.
-%
-We have come a long way since then and today computational methods complement
-theoretical and experimental studies, and are equally important.
+Following the arrival of first electronic general-purpose computer, ENIAC in
+1945, the first simulations were made by \citet{charney_numerical_1950} using a
+simple barotropic model to make 24 hour forecasts \citep[see
+also][]{lynch_richardson_2010}. In those early days when the availability of
+computers was limited, such methods were neither widespread nor feasible.
+Scientific investigations were dominated by theoretical and experimental
+methods.
+%
+We have come a long way and today computational methods complement theoretical
+and experimental studies, and are equally important.
 %
@@
 } either through systematic
-reforms in academia on how researchers' productivity are assessed or through
+reforms in academia on how researchers' productivity are assessed, or through
 regulations and offering incentives, such as waiving the application processing
@@
 throughout the process of software development, it needs unit-tests through
-which a code is continuously built, tested and deployed --- a process known as
+which it is continuously built, tested and deployed --- a process known as
 Continuous Integration (CI). Depending on how matured a software is, CI can be
@@
 
-Finally, the \emph{results} generated from the such codes together with the
+Finally, the \emph{results} generated from such codes together with the
 workflow (i.e.\ scripts to run the codes and to post-process the data) when made
@@
 
-Open-source software development that we know today has been built on top of
-the firm foundations of the Free and Open-Source Software (FOSS) movement. The
-term \emph{free} in FOSS is a misnomer, as it actually stands for
+The Free and Open-Source Software (FOSS) movement has given rise to the tools
+that enable open-source software development today.
+%
+The term \emph{free} in FOSS is a misnomer, as it actually stands for
 \emph{freedom} (to use, modify and distribute). The FOSS movement has
@@
 %
-The FOSS culture can be traced back to the success for scientific publication
-of the free type-setting standards \TeX\ (1977) and \LaTeX\ (1985).
+The FOSS culture can be traced back to the success of the type-setting
+standards \TeX\ (1977) and \LaTeX\ (1985) for authoring scientific publications.
 %
@@
   \href{https://www.mercurial-scm.org/wiki/GitConcepts}{``a detailed
-comparison''} on Mercurial website.}. Simplicity and lowering the learning
-barrier for scientists who are not experts in software development is crucial
-for our project.
+comparison''} on Mercurial website.}. Simplicity and low learning barriers for
+scientists who are not experts in software development is crucial for our
+project.
 %
@@
 %
-Firstly the code is \emph{built} against a known, repeatable emulation
+Firstly, the code is \emph{built} against a known, repeatable emulation
 environment on a server. Thereafter, a set of \emph{unit tests} are run for
@@
 
-New users and developers often rely on good documentation and examples to start
-working with an unfamiliar piece of software. Documentation generators such as
-Sphinx (for Python) and Doxygen (for many languages including C, C++, Fortran
-and Python) parses in-line comments into documentations as PDF, HTML and more
-\citep{lee_ten_2018}. The documentation generation can be automated online and
-deployed using any CI. ``Read the Docs'' has, in the recent years, become the
-hub for building and hosting documentation online, thus facilitating the
-process for developers and considerably improving access to information for
-users.
-%
-In the FluidDyn project we this stop at automating tests and documentation of
-the latest version, but it is possible to further exploit the CI by adding code
-style checks by running ``linters'' alongside unit tests, \emph{continuous
-delivery} of package releases and preserving older versions of documentation.
+New users and developers often rely on good documentation and examples.
+Documentation generators such as Sphinx (for Python) and Doxygen (for many
+languages including C, C++, Fortran and Python) parses in-line comments into
+documentations as PDF, HTML and more \citep{lee_ten_2018}.
+%
+The documentation generation can be automated online and deployed using any CI.
+``Read the Docs'' has, in recent years, become the hub for building and hosting
+documentation online, thus facilitating the process for developers and
+considerably improving access to information for users.
+%
+In the FluidDyn project we this limit ourselves to automating tests and
+documentation of the latest version. It is possible to further exploit the
+CI by adding code style checks by running ``linters'' alongside unit tests,
+\emph{continuous delivery} of package releases and preserving older versions of
+documentation.
 
@@
 Overflow}, where people write thousands of questions and answers on programming
-each week. IRC protocol has continued to sustain the test of time. Instant
+each week. The IRC protocol has continued to sustain the test of time. Instant
 messaging channels such as \codeinline{\#python} on Freenode facilitate
@@
 Taken all together, they greatly improve our collective efficiency and
-open possibilities of interactions that were unthinkable a few years ago.
+open possibilities for interactions that were unthinkable a few years ago.
 %
@@
 As scientists, we can wonder how to fully exploit these new possibilities to do
-science.  The project FluidDyn is an attempt to help the fluid dynamics
-community to advance in this direction.
+science. The project FluidDyn is an attempt to advance the fluid dynamics
+community in this direction.
 
@@
 evident from the fact that by the end of 2017, Python became the second most
-popular language in terms of pull-requests in GitHub\footnote{See GitHub's
-	annual report titled \href{https://octoverse.github.com/}{ ``The State
-of the Octoverse 2017''}.} and questions on StackOverflow\footnote{See
-\href{https://stackoverflow.blog/2017/09/06/incredible-growth-python/}{ ``The
-Incredible Growth of Python''}, a blog post by Stack Overflow.}.
+popular language in terms of pull-requests in GitHub\footnote{%
+%See GitHub's annual report titled \href
+\url{https://octoverse.github.com/}
+% { ``The State of the Octoverse 2017''}.
+%
+} and
+\fnref%
+% [See the blog post ``The Incredible Growth of Python'']%
+{https://stackoverflow.blog/2017/09/06/incredible-growth-python/}%
+{questions on StackOverflow}.
 %
@@
 popularity of a language based on number of skilled users, courses and
-third-party vendors, ranks Python at number four. IEEE Spectrum\footnote{See
-	the analysis titled
-	\href{https://spectrum.ieee.org/computing/software/%
-the-2017-top-programming-languages}{``The 2017 Top Programming Languages''} on
-IEEE spectrum.} ranks Python as the most popular language in 2017 with a metric
+third-party vendors, ranks Python at number four. IEEE Spectrum\footnote{
+%See the analysis titled 
+\url{https://spectrum.ieee.org/computing/software/the-2017-top-programming-languages}
+%{``The 2017 Top Programming Languages''} on IEEE spectrum.
+}
+ranks Python as the most popular language in 2017 with a metric
 based on trends, demand by employers and users on open-source hubs.  All the
@@
 
-\item \textbf{Batteries included.} Python comes with a large and high quality
-\href{https://docs.python.org% /3/library/index.html}{standard library}.
+\item \textbf{Batteries included.} Python comes with a large
+\fnref{https://docs.python.org/3/library/index.html}{standard library}.
 
@@
 %
-Therefore, Python is good for developers of all levels: very gentle for
+Therefore, Python is good for developers at all levels: very gentle for
 beginners and very powerful for advanced users.
@@
 A hard and controversial decision was taken to clean up Python from its
-historical incoherences by having a new series of backwards-incompatible
+historical incoherences by introducing a new series of backwards-incompatible
 versions of the language, starting with Python 3.0, released in December 2008.
 After a long and difficult process, the transition from Python 2 to Python 3 is
-behind us for nearly all important packages\footnote{Close to
+completed for nearly all important packages\footnote{Close to
 \href{https://python3wos.appspot.com/}{95\% of the most downloaded packages}
@@
 numpy/numpy/blob/master/doc/neps/dropping-python2.7-proposal.rst}{future
-releases of \Numpy} and of \href{http://www.python3statement.org/}{the main
-scientific packages} will soon drop compatibility with Python 2. These changes
-open doors to very clean, coherent and potentially faster Python
+versions of \Numpy} and \fnref{http://www.python3statement.org/}{the main
+scientific packages} will not be compatible with Python 2. These changes
+open doors to a very clean, coherent and potentially faster Python
 experience. Therefore, one should not use Python 2 for science any more and
@@
 %
-New versions of Python have nice new features to adapt to new usages and to
+New versions of Python have nice features to adapt to new usages and to
 recent trends in computer science.  For example, three new features were
@@
 %
-There user-friendly and ready-to-use Python installers (similar to Matlab), in
+There are user-friendly and ready-to-use Python installers (similar to Matlab), in
 particular the
@@
 %
-To summarize, it is practical to use and extend Python for scientific purposes.
+To summarize, Python's many features and its adoption in several domains
+makes it a sensible language choice for scientific purposes.
 
@@
 	to interpret python code at the same time\footnote{See
-	\href{https://opensource.com/article/17/4/grok-gil}{this blog article
-on ``How to grok the GIL''} and
-\href{https://faster-cpython.readthedocs.io/cpython37.html}{this technical
-presentation by a CPython core developer}.}. In other words, threads can
-coexist concurrently, by not execute in parallel.
+	\url{https://opensource.com/article/17/4/grok-gil}
+	%{a blog article ``How to grok the GIL''}
+	% and \url{https://faster-cpython.readthedocs.io/cpython37.html}{a
+	% technical presentation by a CPython core developer}.
+	}. In other words, threads can coexist concurrently, by not execute in
+	parallel.
 %
@@
 \item \textbf{Lively, and thus, complicated ecosystem of packages.} For many
-	applications, the standard library is insufficient and one needs
-	external packages. Sometimes there are multiple options and the
+	applications, the standard library is insufficient and external
+	packages are needed. Sometimes there are multiple options and the
 	user/developer has to make choices regarding the tools to use.
 	Especially for the beginners who are not used to the open-source realm,
-	it can be confusing prospect to make good technological choices.
+	it can be confusing prospect to make good technological choices; but
+	this can be solved through proper guidance and training.
+	Of course, such an abundance of projects can also be seen as an advantage.
 
@@
 %
-Python is primarily a ``glue'' language, and a common solution to achieve
+Python is primarily a ``glue'' language. A common solution to achieve
 performance is to use existing code written in C, C++ or Fortran.
@@
 \subsubsection{Ahead Of Time compilation}
-Ahead Of Time compilation is process of generating compiled extensions which
+Ahead Of Time compilation is a way of generating compiled extensions which
 can be dynamically executed by Python.
@@
 %
-This two-step compilation provides room for very clever optimizations%
+This two-step compilation opens possibilities for very clever optimizations%
 \footnote{As demonstrated in this blog post by the main developer of
@@
 %
-We present in \citet{fluidfft, fluidsim} examples of
+In \citet{fluidfft, fluidsim}, we present examples of
 highly efficient Python codes, optimized using a combination of Cython and
@@
 %
-Type-checking is useful since it can avoid simple bugs.
+Type-checking is useful since it can detect simple bugs.
 \fnref{https://docs.python.org/3/whatsnew/3.5.html\#whatsnew-pep-484}{PEP-484}
@@
 
-\subsubsection{Choosing the right package}
-%
-For a beginner in Python development, it is important to get a good
-introduction on how to work with the open-source workflows and understanding of
-the scientific Python stack.
-%
-It can also be very useful to ask to more advanced users and attend Python
-courses or workshops\footnote{This could be regular academic courses, MOOCs or
-workshops like ones organized by
-\href{https://software-carpentry.org/}{Software Carpentry}.}.
-%
-Of course, such abundance of projects can also be seen as an advantage.
+% \subsubsection{Choosing the right package}
+% %
+% For a beginner in Python development, it is important to get a good
+% introduction on how to work with the open-source workflows and understanding of
+% the scientific Python stack.
+% %
+% It can also be very useful to ask to more advanced users and attend Python
+% courses or workshops\footnote{This could be regular academic courses, MOOCs or
+% workshops like ones organized by
+% \href{https://software-carpentry.org/}{Software Carpentry}.}.
+% %
 
@@
 
-So far, we have discussed in detail about Python. However, there are many
+So far, we have discussed the details of Python. However, there are many
 interesting open-source languages that can be used for scientific applications.
@@
 %
-Quick and dirty scripts can be efficient in the short term, but has a negative
+Quick and dirty scripts can be efficient in the short term, but have negative
 impact on a longer time scale.
@@
 
-Thus, a community using open-source methods can be efficient to co-develop its
-tools. However, some of its members have to spend more energy to bear this
-dynamics, at the risk of sacrificing their individual efficiency.
-We see that short-term efficiency, long-term efficiency and productivity at
-different levels can often be incompatible and that we have to consider these
-conflicts when choosing between different technologies.
+Thus, a community using open-source methods can be efficient in co-developing
+its tools. However, some of its members have to spend more energy to bear this
+dynamics, at the risk of sacrificing their individual efficiency.  We see that
+short-term efficiency, long-term efficiency and productivity at different
+levels can often be incompatible and that we have to consider these conflicts
+when choosing between different technologies.
 
@@
 %
-Nowadays, programming is involved in all methods, and is usually difficult to do
+Nowadays, programming is involved in all methods. It is usually difficult to do
 fluid mechanics without software development.
 %
-However, the level in software engineering is on average very low in the
+However, on average, the level in software engineering is very low in the
 community.
@@
 
-Since fluid dynamics is an engineering science, closed-source commercial software
-have historically been dominant in the field.
-%
-Few people are aware of the vicious circle of the closed-source model for people
-and groups relying on it.
-%
-This mechanism is nevertheless very damaging. In the closed-source model, a group
-pays for a license or a new development. The group does not learn how to develop
-what has been paid for. The company, on the contrary, gains from
-the feedback generated from users, through which it improves the product.
-The group produces codes, books, courses using the closed-source product or
-acquire knowledge on how to use the closed-source product, making the group
-increasingly dependent on the product and ready to pay more for it.
+Since fluid dynamics is an engineering science, closed-source commercial
+software have been dominant in the field historically.
+%
+Few people are aware of the vicious circle of the closed-source model for
+people and groups relying on it.
+%
+In the closed-source model, a group pays for a license or a new development.
+The group does not learn how to develop what has been paid for. The company, on
+the other hand, gains from the feedback generated from users, through which it
+improves the product.  The group produces codes, books, courses using the
+closed-source product or acquire knowledge on how to use the closed-source
+product, making the group increasingly dependent on the product and ready to
+pay more for it.
 %
@@
 Despite the reliance on commercial software for research in fluid mechanics,
-%
-there are also a lot of codes produced by scientists in fluid mechanics. But
-without the technical knowledge on how to work collectively, such codes
+there are also a lot of codes produced by scientists in fluid mechanics.
+%
+Without the technical knowledge on how to work collectively, such codes
 are often substandard (compared to the information technology
@@
 these tasks. The development environment is nice and the interpreter is quite
-fast, especially now that it has a JIT compiler.
+fast, especially now, when it has a JIT compiler.
 %
@@
 for doing more than simple processing and data plotting. The comparison with
-Python is often debated upon . We list a few striking Matlab weaknesses:
+Python is often debated upon. We list a few striking Matlab weaknesses:
 
@@
 Now we turn our attention, from structural issues of Matlab syntax, to more
-practical constraints brings about. The arguments are equally applicable for
+practical constraints. The arguments are equally applicable for
 any proprietary software used in research. Closed-source
@@
 research projects rely on such illegal methods?
-%
-It is also unfair in education when students have to work on course projects and
-assignments, with a Matlab code as a starting point and not a free alternative.
 
@@
 The Python scientific ecosystem can now compete in nearly all domains, but this
-was not the case typically 5 years ago.
+was not the case typically five years ago.
 %
@@
 %
-Although we do not pay the authors, but people do need to pay to use the
-software. 
+Although the codes are meant to be free, users need to pay to obtain the
+compilers/interpreters which execute them. 
 %
@@
 %
-The potential of these tools is not fully exploited yet, in the field of fluid
+The potential of these tools is not fully exploited yet in the field of fluid
 mechanics.
@@
 %
-We have not been ``reinventing the wheel'', but rather have been ``standing
+We have not been ``reinventing the wheel'', but rather we have been ``standing
 on the shoulders of giants''. In other words, we have utilized and extended
@@
 
-FluidDyn project has adopted a decentralized
-organization with one common package and other specialized packages, to make
-maintenance easier and also the target audience (students, experimentalists
-or computational scientists) may not use the full package. FluidDyn project now
-offers the following packages: \fluidpack{dyn},
+The FluidDyn project has adopted a decentralized organization with one common
+package and other specialized packages, to make maintenance easier.
+%
+The project now offers the following packages:
+%
+\fluidpack{dyn},
 %~\citep{fluiddyn}

### 2019-06-05 chapter_00_open_science.tex
Minor corrections, exec black and isort on python scripts

@@
 %
-The FOSS culture can be traced back to the success of the type-setting
+The FOSS culture can be traced back to the success of typesetting
 standards \TeX\ (1977) and \LaTeX\ (1985) for authoring scientific publications.
@@
 %See GitHub's annual report titled \href
-\url{https://octoverse.github.com/}
+\url{https://octoverse.github.com/2017}
 % { ``The State of the Octoverse 2017''}.
@@
 %
-To summarize, Python's many features and its adoption in several domains
+To summarize, Python's numerous features and its adoption in several domains
 makes it a sensible language choice for scientific purposes.
@@
 	% technical presentation by a CPython core developer}.
-	}. In other words, threads can coexist concurrently, by not execute in
+	}. In other words, threads can coexist concurrently, by not executing in
 	parallel.

### 2019-06-05 imgs/pythonrc.py

@@ -1,4 +1,3 @@
 import matplotlib
 
-
 matplotlib.use("Agg")

### 2019-06-05 scripts/get_bib.py

@@ -1,7 +1,7 @@
 """Receive file served by Zotero + Better BibTeX plugin."""
-import requests
 import time
 
+import requests
 
 FORMAT = "biblatex"
 SRC = "thesis." + FORMAT

### 2019-06-05 templates/utils_render.py

@@ -8,11 +8,12 @@ Notes:
     See https://github.com/AKuederle/Py-Tex-automation-example
 
 """
-import re
-import jinja2
 import os
+import re
 import sys
 from pathlib import Path
+
+import jinja2
 
 
 def get_options_from_file(path):
@@ -54,13 +55,14 @@ def get_options_from_yaml(path):
         options["authors_short"] = [
             name_in_initials(name) for name in options["authors"][:-1]
         ]
-        options["authors_short"].extend([
-            r'\&',
-            name_in_initials(options["authors"][-1])
-        ])
+        options["authors_short"].extend(
+            [r"\&", name_in_initials(options["authors"][-1])]
+        )
     else:
-        options["authors_short"] = [name_in_initials(options["authors"][0]), "et al."]
-
+        options["authors_short"] = [
+            name_in_initials(options["authors"][0]),
+            "et al.",
+        ]
 
     uniq_affiliations = list(set(options["affiliations"]))  # unique
     options["authors_aff"] = []
@@ -115,6 +117,7 @@ def render_template(template, insert_variables, out_path):
 if __name__ == "__main__":
     try:
         from docopt import docopt
+
         argv = docopt(__doc__)
         options_file = Path(argv["<options_file>"])
         template_file = Path(argv["<template_file>"])
@@ -122,7 +125,6 @@ if __name__ == "__main__":
         argv = sys.argv
         options_file = Path(argv[1])
         template_file = Path(argv[2])
-
 
     options = get_options_from_yaml(options_file)
     template = get_template(template_file)

### 2019-06-05 scripts/render_paper_yml.py
Move python files to scripts/

@@ -2,7 +2,7 @@
 """Render a latex jinja template
 
 Usage:
-    utils_render.py <options_file> <template_file>
+    render_paper_yml.py <options_file> <template_file>
 
 Notes:
     See https://github.com/AKuederle/Py-Tex-automation-example

### 2019-06-05 chapter_01_decomposition.tex
Section on helmholtz normal mode decomposition + glossary

\section{Helmholtz and normal mode decomposition}

\subsection{Governing Equations}

The governing equations for a shallow layer of fluid are:
\begin{align}
  \partial_t \mathbf u 
      = &- (\mathbf{u}.\nabla) \mathbf{u} - c^2 \nabla h - f\mathbf{e_z} \times 
\mathbf u \label{dtu0} \\
  \partial_t h 
      =& - \nabla. (h \mathbf u) \label{dth}
\end{align}
Equation \eqref{dtu0} may also be written in the \textit{rotational form} as:
\begin{equation}
 \label{dtu}
 \partial_t \mathbf u 
 = - \nabla |\mathbf u|^2/2 - c^2 \nabla h - \zeta \times \mathbf u
\end{equation}
where, \gls{zeta}
represents absolute vorticity, i.e. the sum of vorticity and system rotation.
Furthermore, equations \eqref{dtu0} and \eqref{dth} can be combined to form an 
equation for the \textit{total mass flux}, \gls{J} $= h\mathbf{u}$:
\begin{equation}
  \label{dtJ}
 \partial_t \mathbf J = -(\mathbf{u}.\nabla)\mathbf{J} - \nabla(c^2h^2)/2 -
\zeta \times \mathbf J - (\nabla. \mathbf J)\mathbf u
\end{equation}
Similarly, an equation for the \textit{displaced mass flux}, \gls{M} $ = 
\eta\mathbf{u}$ is:
\begin{equation}
  \label{dtM}
 \partial_t \mathbf M = -(\mathbf{u}.\nabla)\mathbf{M} - \nabla(c^2\eta^2)/2 - 
\zeta \times \mathbf M - (\nabla.\mathbf{u} + \nabla.\mathbf{M})\mathbf u
\end{equation}
For a divergence free flow, a Poisson equation for \emph{h} can be formulated.
Taking divergence of \eqref{dtu},
yields the Poisson equation:
\begin{equation}
  \label{poisson}
  \nabla^2 h = \frac{1}{c^2} \left[ \nabla.(\zeta \times \mathbf u )
	      - \nabla^2 \frac{|u|^2}{2} \right]
\end{equation}
Applying fourier transform, the spectral counterpart for \eqref{poisson} in 
tensor notation is:
\begin{equation}
\label{poisson_fft}
  -\kappa^2 \hat{h} = \frac{1}{c^2} \left[ ik_i (\widehat{\epsilon_{ijk} \zeta_j 
u_k})
	      + \kappa^2 \frac{\widehat{u_i u_i}}{2} \right]
\end{equation}

\subsection{Helmholtz Decomposition}
\subsubsection{For velocity field, u}
The Helmholtz decomposition theorem, or the fundamental theorem of vector 
calculus, 
states that any well-behaved vector field can be decomposed into
the sum of a longitudinal (diverging, non-curling, irrotational) vector field 
and 
a transverse (solenoidal, curling, rotational, non-diverging) vector field. 
This allows us to express the velocity field as:
\begin{align}
\label{helm_u}
\mathbf u & = -\nabla \times (e_z \Psi) + \nabla \Phi\\
	& =  -\nabla \times \Psi_z + \nabla \Phi
\end{align}

For the sake of clarity,
we shall denote the rotational (vortical) and divergent (wave) parts of the 
velocity with the suffix \emph{r} and \emph{d} respectively. Thus,
\begin{align}
\mathbf u_r & = -\nabla \times \Psi_z ; &\mathbf u_d =   \nabla \Phi
\end{align}
And therefore, $\mathbf u  = \mathbf u_r + \mathbf u_d$.

To find the projection operators for the divergent part, taking divergence of 
\eqref{helm_u} gives $\nabla .\mathbf{u} = \nabla^2 \Phi
$.
This transforms in the spectral plane as, $ik_j \hat{u}_j = -\kappa^2 
\hat{\Phi}$, implying:
\begin{align}
\mathbf{\hat{u}}_d =& ik_i \hat{\Phi} = \frac{k_i k_j}{\kappa^2} \hat{u}_j\\
\mathbf{\hat{u}}_r =& \mathbf{\hat{u}} - \mathbf{\hat{u}}_d = \left( \delta_{ij} 
- \frac{k_i k_j}{\kappa^2} \right) \hat{u}_j
\end{align}

Thus, for two-dimensions the decomposed velocity are represented as follows,

\begin{align}
\mathbf{\hat{u}}_d^x =& \frac{k_x (k_x \hat{u}_x + k_y \hat{u}_y)}{\kappa^2} \\
\mathbf{\hat{u}}_r =& \hat{u}_{x} - \frac{k_x  (k_x \hat{u}_x + k_y \hat{u}_y)}{\kappa^2} 
\end{align}

\subsubsection{For fluid depth, h}
To obtain a similar decomposition, $h = h_r + h_d$, one should use the Poisson 
equation \eqref{poisson_fft}.
Since Poisson equation requires a divergence free flow, the LHS of 
\eqref{poisson_fft} would correspond to the rotational part of the flow in the 
transformed plane, i.e. $\hat{h}_r$. Henceforth, the divergent part, $\hat{h}_d$ 
can be obtained by subtracting $\hat{h}_r$ from $\hat{h}$.

\subsection{Normal mode decomposition}

\subsubsection{Following Bartello 1995}
In order to isolate oscillating fast modes that display no nonlinearity, one 
can adopt a linearization approach followed by a normal mode decomposition. 
As a result of linearization, from \eqref{dtu} and \eqref{dth}:
  \begin{align}
    \partial_t \mathbf u =& - c^2 \nabla \eta - f\mathbf{e_z} \times \mathbf u 
  \label{dtu_l} \\
    \partial_t \eta =& - \nabla.  \mathbf u \label{dth_l}
  \end{align}
Taking curl and divergence of the linearized equation for momentum 
\eqref{dtu_l} gives the following evolution equations with change of variables:
  \begin{align}
    \partial_t \zeta =& - f \delta
  \label{dtcurl_l} \\ 
    \partial_t \delta =&  f \zeta - c^2 \nabla^2 \eta \label{dtdiv_l} \\
    \partial_t \eta =& - \delta \label{dteta_l}
  \end{align}
where $\zeta$ and $\delta$ are relative vorticity and divergence as functions 
of $\mathbf{r}$ and $t$ respectively. Representing the dependent flow 
quantities in terms of Fourier modes:
  \begin{align}
  \begin{pmatrix}
      u \\ v \\ \eta \\ \zeta \\ \delta
  \end{pmatrix} (\mathbf{r},t)
  = \int  
  \begin{pmatrix}
      \hat{u} \\ \hat{v} \\ \hat{\eta} \\ \hat{\zeta} \\ \hat{\delta}
  \end{pmatrix} e^{i(\mathbf k . \mathbf{r} - \omega t)} \mathbf{dk} d\omega
  \end{align}
allows to rewrite the system of equations \eqref{dtcurl_l} to \eqref{dteta_l} 
as an eigenvalue problem:
\begin{align}
 i\omega
   \begin{Bmatrix}
       \hat{\zeta} \\ \hat{\delta} \\c\kappa\hat{\eta}
  \end{Bmatrix}
  = i
  \begin{bmatrix}
     0 & if & 0 \\
   -if &  0 & -ic\kappa \\
     0 & ic\kappa & 0 
  \end{bmatrix}
  \begin{Bmatrix}
       \hat{\zeta} \\ \hat{\delta} \\c\kappa\hat{\eta}
  \end{Bmatrix}
\end{align}
where, $\kappa = |\mathbf{k}|$.

Let us define $A$ as the Hermitian matrix operating on the vector 
$\mathbf{W} = \{\hat{\zeta}, \hat{\delta} ,c\kappa \hat{\eta} \}^T $.which 
yields the 
familiar dispersion relation 
for the slow geostrophic mode 
and fast Poincar\'e wave modes:
\begin{equation}
 \omega^{(0)} = 0;\quad \omega^{(\pm)}=\pm \sigma
\end{equation}
where, $\sigma = \sqrt{f^2 + c^2\kappa^2 }$. These are the eigenvalues of 
the matrix operator \emph{A}. Since \emph{A} is Hermitian, the corresponding 
eigenvectors are orthogonal and these are normalized as follows
\begin{align}
  \mathbf X^{(0)}_n = 
    \frac{1}{\sigma}
    \begin{Bmatrix}
       -c\kappa \\ 0 \\ f
     \end{Bmatrix}; \quad
  \mathbf X^{(\pm)}_n = 
    \frac{1}{\sqrt{2} \sigma}
    \begin{Bmatrix}
       f \\ \mp i\sigma \\ c\kappa
     \end{Bmatrix}
\end{align}
Let $X_n$ be the eigenvector matrix, which follows the 
property $X_n \bar{X}_n^{T}=I$. It can, then, be applied to diagonalize the 
system of 
equations as follows:
\begin{align}
 \partial_t \mathbf{W} =& [A] \mathbf{W}\\
 \partial_t (\bar{X}_n^T \mathbf{W}) =& \bar{X}_n^T[A]X_n 
\bar{X}_n^T\mathbf{W}\\
			    =& [\Lambda] (\bar{X}_n^T\mathbf{W})
\end{align}
Thus, the alternate diagonalized system of equations for the normal modes are 
given by:
\begin{align}
 \partial_t
\mathbf{N}
  = &
  \begin{bmatrix}
     0 & 0 & 0 \\
     0 & -i\sigma & 0 \\
     0 & 0 & i\sigma 
  \end{bmatrix}
  \mathbf{N}\\
\text{where},
   \mathbf{N} = \bar{X}_n^T \mathbf{W}
   = &\frac{1}{\sqrt{2}\sigma}
   \begin{Bmatrix}
       -\sqrt{2}c\kappa(\hat{\zeta} -f \hat{\eta}) \\ 
        f\hat{\zeta} + c^2\kappa^2\hat{\eta} - i\sigma\hat{\delta} \\
       f\hat{\zeta} + c^2\kappa^2\hat{\eta} + i\sigma\hat{\delta}
   \end{Bmatrix}
\end{align}

The first mode represents \emph{linearised potential vorticity} which is 
conserved in this framework. The remaining two are the \emph{linearised 
ageostrophic or wave modes}. In the absence of system rotation, i.e. $f=0$, 
the normal modes are simply,
\begin{equation}
\label{nmode}
 \mathbf{N} = 
 \begin{Bmatrix}
  \hat \zeta \\
  \frac{1}{\sqrt{2}} (c\kappa \hat{\eta} - i\hat{\delta})\\
  \frac{1}{\sqrt{2}} (c\kappa \hat{\eta} + i\hat{\delta})
 \end{Bmatrix}
\end{equation}

%---------------------------------------------%
\subsubsection{Following Farge \& Sadourny 1989}
Instead of finding the normal modes for the vorticity, divergence and 
displacement field of the flow, we shall make use of the Helmholtz 
decomposition described in \eqref{helm_u}. The shallow water equations then 
transform to:
 \begin{align}
    \partial_t \psi =&  f \phi
  \label{dtpsi_l} \\ 
    \partial_t \phi =&  -f \psi - c^2 \eta \label{dtphi_l} \\
    \partial_t \eta =& - \nabla^2 \phi \label{dteta_l2}
  \end{align}
where $\psi$ and $\psi$ are stream function and velocity potential as functions 
of $\mathbf{r}$ and $t$ respectively. By substituting the dependent 
variables with the respective Fourier transform, this reduces to the eigenvalue 
problem:
\begin{align}
 i\omega
   \begin{Bmatrix}
       \hat{\psi} \\ \hat{\phi} \\ \hat{\eta}
  \end{Bmatrix}
  = i
  \begin{bmatrix}
     0 & if & 0 \\
   -if &  0 & ic^2 \\
     0 & -i\kappa^2 & 0 
  \end{bmatrix}
  \begin{Bmatrix}
       \hat{\psi} \\ \hat{\phi} \\ \hat{\eta}
  \end{Bmatrix}
\end{align}
the square matrix is not Hermitian and this would result in complex 
eigenvalues. By adopting the following change of variables:
\begin{equation}
 \hat{\psi} \to \kappa^2\hat \psi = \hat{\zeta} ; \quad
 \hat{\phi} \to -\kappa^2\hat \phi = \hat{\delta}; \quad
 \hat{\eta} \to c\kappa\hat \eta
\end{equation}
it falls back to the previous eigenvalue problem as demonstrated in the  
previous section. In other words, we can use the same eigenvector matrix, $X_n$ 
to find the normal modes of:

$$\mathbf{H} = \{\hat \psi,\; \hat \phi,\;  \eta  \}^T$$
which is closely related to:
$$\mathbf{W} 
= \{\kappa^2\hat \psi,\; -\kappa^2\hat 
\phi,\; c\kappa\hat \eta  \}^T 
= \{\hat \zeta;\; \hat \delta;\;  c\kappa \hat \eta \}^T $$

%---------------------------------------------%
\subsubsection{Normal mode inversion to primitive variables}
To begin with, we shall form a new vector $\mathbf{B} = \mathbf{N}/\kappa$ 
which will have the same dimensions as velocity.
It has been shown before that the normal modes can be represented by $\mathbf{N} 
= \bar{X}_n^T \mathbf{W}$. Now,  we shall form a new vector $\mathbf{B} = 
\mathbf{N}/\kappa$ which will have the same dimensions as velocity. Thus,
\begin{align}
   \mathbf{B}
   = &\frac{1}{\sqrt{2}\sigma}
    \begin{Bmatrix} \sqrt{2} c\left(- 
    \kappa^{2} \hat \psi +  f\hat\eta \right)\\
    \kappa \left(c^{2} \eta + f \psi + i \phi \sigma\right)\\
    \kappa \left(c^{2} \eta + f \psi - i \phi \sigma\right)
    \end{Bmatrix}
\end{align}

The $\mathbf{B}$ vector can also be related to the primitive variable vector, 
$\mathbf{U} = \{\hat{u},\hat{v},c\hat{\eta}\}^T$ using transformation matrices 
as follows:
\begin{align}
 \mathbf{B} =&\frac{1}{\kappa}[\bar{X}_n]^T W\\
	    =&\frac{1}{\kappa}[\bar{X}_n]^T [P] \mathbf{U} 
\end{align}
where,
\begin{equation}P=
\begin{bmatrix}- i k_{y} &  i k_{x} & 0\\ i k_{x} &  i 
k_{y} & 
0\\0 & 0 & \kappa\end{bmatrix}
\end{equation}
It is also straightforward to show that the magnitude of the normal modes 
equals the total energy as:
\begin{align}
 \mathbf{\bar B}^T\mathbf{B} 
 =&\frac{1}{\kappa^2} \mathbf{\bar U}^T 
     [\bar P]^T[{X}_n]  [\bar{X}_n]^T [P] \mathbf{U} \\
 =& \mathbf{\bar U}^T\mathbf{U}
\end{align}
implying, $$2(E_K+E_P)=\sum_i B^{(i)}\bar{B}^{(i)}=\kappa^{2} \psi 
\overline{\psi} +\kappa^{2} \phi \overline{\phi} + c^{2} 
\eta \overline{\eta}
= u\bar{u} + v\bar{v} + c^2\eta\bar{\eta}$$
This allows us to represent the primitive variables in terms of normal modes as 
$$\mathbf{U} = [Q]\mathbf{B}$$
where, the inversion matrix is
\begin{align}
 Q 
 = & \kappa [P]^{-1}[\bar{X}_n]^{T^{-1}}\\
 =& \frac{1}{\sqrt{2}\sigma\kappa}
 \begin{bmatrix}
      -i\sqrt{2}c\kappa  k_{y} & 
      \left(i f k_{y} + 
	  k_{x}   \sigma\right) &
      \left(i f k_{y} - k_{x} \sigma\right)\\
      i \sqrt{2}c\kappa k_{x} & 
      \left(-  i f k_{x} +  k_{y}   \sigma\right) &
      - \left(i f  k_{x} + k_{y} \sigma\right)\\
      \sqrt{2}\kappa f & 
       c\kappa^{2} &
      c\kappa^{2}
\end{bmatrix}
\end{align}
In tensor notation,
\begin{align}
 \hat{u}_l =& \epsilon_{lm3} ik_m\left[ -\frac{c}{\sigma} B^{(0)} + 
	      \frac{f}{\sqrt{2}\sigma\kappa} (B^{(+)} + B^{(-)})
	      \right]
	      + k_l \frac{1}{\sqrt{2}\kappa}  (B^{(+)} - B^{(-)})\\
c\hat{\eta} =& \frac{f}{\sigma} B^{(0)} + \frac{c\kappa}{\sqrt{2}\sigma} 
(B^{(+)} + B^{(-)})    
\end{align}

\subsection{Spectral energy budget}
\subsubsection{Kinetic energy}
%---------------------------------------------%
\paragraphbf{Expression for mean kinetic energy}
Kinetic energy being cubic for shallow water equations, it can be split into 
quadratic and non-quadratic parts as follows:
\begin{align*}
  E_K(\mathbf{r},t)
    =& h\mathbf{u}.\mathbf{u}\\
    =& (1+\eta)\mathbf{u}.\mathbf{u}\\
    =& \mathbf{u}.\mathbf{u} + \eta\mathbf{u}.\mathbf{u}\\
    =& E_K^Q + E_K^{NQ}
\end{align*}

%---------------------------------------------%
\paragraphbf{K.E. budget}
Let the rate of change of kinetic energy, without any approximations, be 
written as:
\begin{equation}\label{dtKE}
 \pder{t}E_K(\mathbf{k},t) = T_K + C_K
\end{equation}
where, $T_K$ and $C_K$ represents the transfer and conversion spectral 
functions respectively. Equation 
\eqref{dtKE} splits into:
\begin{align*}
 \pder{t}E_K^Q + \pder{t}E_K^{NQ}
 =&  (T_K^{Q} + C_K^{Q}) + (T_K^{NQ} + C_K^{NQ})
%  =& \frac{1}{2}\partial_t(\mathbf{\hat u}.\widehat{h\mathbf{u}}^*)\\
%  =& \frac{1}{2}\partial_t(\mathbf{\hat u}.\hat{\mathbf{J}}^*)
\end{align*}
%-------------------------------------------------%
\paragraph{Quadratic K.E. budget:}
Consider the rate of change of quadratic kinetic energy,
\begin{equation}\label{dtKE2}
 \pder{t}E_K^{Q}(\mathbf{k},t) = T_K^{Q} + C_K^{Q}
\end{equation}
Following \eqref{dtu0}, equation \eqref{dtKE2} expands into:
\begin{align*}
 \partial_t E_K^{Q}(\mathbf{k},t)
 =& \frac{1}{2}\partial_t(\mathbf{\hat u}.\mathbf{\hat u^*})\\
 =& \frac{1}{2}\left[ \mathbf{\hat u} .\pder[\mathbf{\hat u^*}]{t} 
		      + \mathbf{\hat u^*}. \pder[\mathbf{\hat u}]{t}\right]\\
 =& \frac{1}{2}\left[ -\hat{u}_i (\widehat{ u_j\partial_j u_i })^* 
    - c^2 \hat{u}_i (ik_i \hat{\eta})^* 
    - \hat{u}_i (\epsilon_{i3k} f \hat{u}_k)^*
    - ... \text{conjugate terms}
    \right]\\
 =& -\text{Re}\left[ \hat{u}_i (\widehat{ u_j\partial_j u_i })^* 
    + c^2 \hat{u}_i (ik_i \hat{\eta})^* 
    + \hat{u}_i (\epsilon_{i3k} f \hat{u}_k)^* \right]
\end{align*}
% since, $u_i (\epsilon_{ijk} \zeta_j u_k) = 0$
% \footnote{ \textbf{Note}
% $\epsilon_{ijk} \zeta_j u_k 
% = \epsilon_{jki}  (\epsilon_{jlm} \partial_l u_m +   f_j )u_k$}
% =& (\delta_{kl}\delta_{im} - \delta_{km} \delta_{il})\partial_l u_m +    
% \epsilon_{ijk} f_j u_k\\
% =& u_k\pder{x_k}u_i - \pder{x_i}u_ku_k + \epsilon_{ijk} f_j u_k
we have,
\begin{align}
 T_K^{Q}= &-\text{Re}\left[\hat{u}_i (\widehat{ u_j\partial_j u_i })^* 
	   \right]\\
 C_K^{Q}= &-\text{Re}\left[   c^2 \hat{u}_i (ik_i \hat{\eta})^*  \right]\\
 	= &-\text{Re}\left[   c^2 \hat{u}_i^* ik_i \hat{\eta} \right]
\end{align}

%-------------------------------------------------%
\paragraph{Non-quadratic K.E. budget:}
\begin{align*}
  \pder{t}E_K^{NQ}(\mathbf{k},t) =& T_K^{NQ} + C_K^{NQ}
\end{align*}
Following equations \eqref{dtu0} and \eqref{dtM} for rate of change of velocity 
$\mathbf{u}$ and the displaced mass flux, $\mathbf M = \eta \mathbf{u}$ 
respectively,
 \begin{align*}
 \partial_t E_K^{NQ}(\mathbf{k},t)
 =& \frac{1}{2}\partial_t(\mathbf{\hat M}.\mathbf{\hat u^*})\\
 =& \frac{1}{2}\left[ \mathbf{\hat M} .\pder[\mathbf{\hat u^*}]{t} 
		      + \mathbf{\hat u^*}. \pder[\mathbf{\hat M}]{t}\right]\\
 =& \frac{1}{2}\left[-\hat{M}_i(\widehat{ u_j\partial_j u_i })^*
		      - c^2 \hat{M}_i (ik_i \hat{\eta})^*
		      - \hat{M}_i (\epsilon_{i3k} f \hat{u}_k)^* \right]\\
  &+\frac{1}{2}\left[ - \hat{u}_i^* \widehat{(u_j\partial_j M_i) }
		      - c^2\hat{u}_i^*ik_i\widehat{\eta\eta}/2
		      - \hat{u}_i^* (\epsilon_{i3k} f \hat{M}_k) 
		      - \hat{u_i}^* \widehat{(u_i \partial_j({u}_j+ 
M_j))}
			\right]
\end{align*}
Hence,
\begin{align}
 T_K^{NQ}= & -\frac{1}{2}\left[\hat{M}_i(\widehat{ u_j\partial_j u_i })^*
			      + \hat{u}_i^* \widehat{(u_j\partial_j M_i) }
			      + \hat{u_i}^* \widehat{(u_i \partial_j{M}_j)}
			      \right] \\
 T_{K,K}^{NQ,Q}= & -\frac{1}{2}\left[ \hat{u_i}^* \widehat{(u_i 
\partial_j{u}_j)}
			\right] \\
 T_{K,P}^{NQ,Q}= & -\frac{1}{2}\left[c^2 \hat{M}_i (ik_i \hat{\eta})^*
	      + c^2\hat{u}_i^*ik_i\widehat{\eta\eta} /2
	      \right]
\end{align}

\paragraphbf{Helmholtz decomposition of kinetic energy}

Furthermore, using Helmholtz decomposition to substitute for $\mathbf u$ in  
the expression 
for \emph{mean kinetic energy},

\begin{align}
\braket{ E_K} 
& = \frac{1}{2} \braket{ h\mathbf u.\mathbf u} \\
& = \frac{1}{2} \braket{h(\mathbf{u_r} + \mathbf{u_d} ).(\mathbf{u_r} + 
\mathbf{u_d} )}\\
& = \frac{1}{2}\braket{ h |\mathbf{u_r} | ^2} 
    -\braket{ \mathbf{u_r}.h\mathbf{u_d}}
    +  \frac{1}{2} \braket{ h | \mathbf{u_d}| ^2}
\end{align}
where, $h = 1 + \eta$ and, $\eta$ is the surface displacement of the shallow 
water layer. In the limit for small displacements i.e., $\eta \to 0$ or $ h \to 
1$, invoking orthogonality,

\begin{equation}
\lim_{h \to 1} \braket{E_K} = \braket{E_K^{Q}}
=  \frac{1}{2}\braket{|\mathbf{u_r}  | ^2 + | \mathbf{u_d} | ^2}
\end{equation}
The spectral equivalent of the above expression will be:
\begin{equation}
\braket{E_K^{Q} }
=  \frac{1}{2}\braket{ \mathbf{\hat{u}_r}.\mathbf{\hat{u}_r}^* + 
\mathbf{\hat{u}_d}.\mathbf{\hat{u}_d}^* }
\end{equation}

%---------------------------------------------------------%
\subsubsection{Available potential energy}
\paragraphbf{Expression for mean available potential energy}
\begin{align}
\braket{ E_P} 
& = \frac{c^2}{2} \braket{ \eta^2}
\end{align}
The spectral equivalent of which is:
\begin{align}
\braket{ E_P} 
& = \frac{c^2}{2} \braket{ \hat{\eta}\hat{\eta}^*}
\end{align}

\paragraphbf{A.P.E. budget}
Similar to equation \eqref{dtKE} we can form,
\begin{equation}\label{dtPE}
 \pder{t}E_P(\mathbf{k},t) = T_P+ C_P
\end{equation}
where, $T_P$ and $C_P$ represents the transfer and conversion spectral 
functions respectively.

Following \eqref{dth}, equation \eqref{dtPE} expands as:
\begin{align*}
 \pder{t}E_P(\mathbf{k},t)
 =& \frac{c^2}{2}\partial_t({\hat \eta}{\hat \eta^*})\\
 =& \frac{c^2}{2}\left[ \hat{\eta} .\pder[\hat{\eta}^*]{t}+ \hat 
\eta^*.\pder[\hat \eta]{t}\right]\\
 =& \frac{c^2}{2}\left[ - \hat{\eta} (ik_i\hat{ u_i } + ik_i\widehat{ 
      \eta u_i })^*  
	    - \hat{\eta}^* (ik_i\hat{ u_i } + ik_i\widehat{ 
      \eta u_i })  \right]\\
      =& -\text{Re}\left[c^2\hat{\eta} (ik_i\hat{ u_i } + ik_i\widehat{ 
      \eta u_i })^* \right]
\end{align*}
Thus,
\begin{align}
T_P= & -\text{Re}\left[c^2\hat{\eta} (ik_i\widehat{ 
      \eta u_i })^*  \right]\\
C_P=&  -\text{Re}\left[c^2\hat{\eta} (ik_i\hat{u_i })^*\right]\\
   =&  +\text{Re}\left[c^2\hat{\eta} ik_i\hat{u_i }^*\right]
\end{align}
thus, $C^Q_K = -C_P$ which, in turn, makes the assertion that  
equivalent conversion occurs between quadratic K.E. and A.P.E.
% 
% \begin{align*}
% 2\times T_K(\mathbf{k},t)
% & = \partial_t (\mathbf u.h\mathbf u)\\
% &= \partial_t \mathbf u .h\mathbf u  +|\mathbf u|^2 \partial_t h + h\mathbf u 
% .\partial_t \mathbf u \\
% & =   2 \left( -  \frac{\nabla|\mathbf u|^2}{2} - c^2 \nabla h - \zeta_a \times \mathbf u	\right).h\mathbf u - \left( |\mathbf u|^2 \nabla.(h \mathbf u )\right)\\ 
% & =   \left( -  h\frac{\nabla|\mathbf u|^3}{3} - c^2 \frac{\nabla h^2}{2} . \mathbf u \right)- \left( |\mathbf u|^2 \nabla.(h \mathbf u )\right) \\
% \end{align*}
% 
%

% \end{document}

### 2019-06-07 paper_05_milestone_issf/article.tex
Fix for long author list add et al.

@@
 \centerline{
-\includegraphics[width=80mm]{Figures/GOPR1453.pdf}
-\includegraphics[width=80mm]{Figures/fig_movement_carriage.pdf}}
+\includegraphics[width=0.45\textwidth]{Figures/GOPR1453.pdf}
+\includegraphics[width=0.45\textwidth]{Figures/fig_movement_carriage.pdf}}
 \caption{Top: scheme of the experimental setup (see text for
@@
 \begin{figure}
-\centerline{\includegraphics[width=16cm]{Figures/fig_profiles.pdf}}
+\centerline{\includegraphics[]{Figures/fig_profiles.pdf}}
 \vspace{-2mm}
@@
 
-\input{table.tex}
+\input{paper_05_milestone_issf/table.tex}
 
@@
 \centerline{
-\includegraphics[width=5.18cm]{\pathfigures/exp21/vh_400.pdf}
-\includegraphics[width=4.3cm]{\pathfigures/exp21/vh_655.pdf}
-\includegraphics[width=5.9cm]{\pathfigures/exp21/vh_890.pdf}}
+\includegraphics[width=4.14cm]{paper_05_milestone_issf/Figures/exp21/vh_400.pdf}
+\includegraphics[width=3.44cm]{paper_05_milestone_issf/Figures/exp21/vh_655.pdf}
+\includegraphics[width=4.72cm]{paper_05_milestone_issf/Figures/exp21/vh_890.pdf}}
 \vspace{0mm}
 \centerline{
-\includegraphics[width=5.2cm]{\pathfigures/exp21/vv_890.pdf}
-\includegraphics[width=4.35cm]{\pathfigures/exp21/vv_655.pdf}
-\includegraphics[width=5.8cm]{\pathfigures/exp21/vv_400.pdf}
+\includegraphics[width=4.16cm]{paper_05_milestone_issf/Figures/exp21/vv_890.pdf}
+\includegraphics[width=3.48cm]{paper_05_milestone_issf/Figures/exp21/vv_655.pdf}
+\includegraphics[width=4.63cm]{paper_05_milestone_issf/Figures/exp21/vv_400.pdf}
 }
@@
 \begin{figure}[htb]
-\centerline{\includegraphics[width=9.cm]%
-{\pathfigures/exp28/ken_fct_t_T_exp28.pdf}}
+\centerline{
+\includegraphics[width=9.cm]{paper_05_milestone_issf/Figures/exp28/ken_fct_t_T_exp28.pdf}}
 \vspace{-2mm}
@@
 \begin{figure}[htb]
-\centerline{\includegraphics[height=7.5cm]%
-{\pathfigures/exp28/normalized_S2_exp28.pdf}}
+\centerline{
+\includegraphics[height=7.5cm]{paper_05_milestone_issf/Figures/exp28/normalized_S2_exp28.pdf}}
 \vspace{-2mm}

### 2019-06-13 scripts/pandoc_auto_height.py
Pandoc filters and latexindent

@@ -29,9 +29,9 @@ def detect_height(url):
     return height
 
 
-def action_auto_height(elem, doc):
+def action(elem, doc):
     """Automatically determines height attribute for images."""
-    if isinstance(elem, pf.Image):
+    if doc.format == 'latex' and isinstance(elem, pf.Image):
         # pf.debug(elem)
         if "width" in elem.attributes and "height" not in elem.attributes:
             elem.attributes["height"] = detect_height(elem.url)
@@ -40,7 +40,7 @@ def action_auto_height(elem, doc):
 
 
 def main(doc=None):
-    return pf.run_filter(action_auto_height, doc=doc)
+    return pf.run_filter(action, doc=doc)
 
 
 if __name__ == '__main__':

### 2019-06-13 scripts/pandoc_math.py

@@ -1,0 +1,70 @@
+import re
+
+import panflute as pf
+
+
+def create_citation(label, prefix=""):
+    pandoc_label = f"[@{prefix}{label}]"
+    elem = pf.Cite(pf.Str(pandoc_label))
+    #  pf.debug(elem)
+    return elem
+
+
+def extract_latex_labels(text):
+    p = re.compile(
+        r"""
+        \\label{
+            ([0-9a-zA-Z:_]+)  # capture within the label
+        }
+        """,
+        re.VERBOSE,
+    )
+    labels = p.findall(text)
+    if labels:
+        return labels
+    else:
+        pf.debug(f"No label in {text}")
+
+
+def action(elem, doc):
+    r"""Handles \begin{align} \label \eqref and \ref while converting from LaTeX.
+
+    Why
+    ---
+    * Pandoc converts \begin{align} -> \begin{aligned} implicitly which does
+    not handle multiple labels
+
+    * Pandoc does not have a mechanism to handle \eqref or \ref, but
+    pandoc-crossref has one, so we re-engineer an expression which suits
+    the latter
+
+    """
+    if isinstance(elem, pf.Math) and elem.format == "DisplayMath":
+        text = pf.stringify(elem)
+        if r"\label" in text:
+            env = "align"
+        else:
+            env = "align*"
+            # labels = extract_latex_labels(text)
+
+        if r"\begin{aligned}" in text:
+            return pf.RawInline(
+                "\n" + text.replace("{aligned}", f"{{{env}}}") + "\n", format="tex"
+            )
+
+    if doc.format.startswith("markdown"):
+        if isinstance(elem, pf.RawInline) and elem.format == "tex":
+            label = pf.stringify(elem).lstrip(r"\eqref{").rstrip("}")
+            return create_citation(label)
+        elif isinstance(elem, pf.Link) and "reference-type" in elem.attributes:
+            if elem.attributes["reference-type"] in ("eqref", "ref"):
+                label = elem.url.lstrip("#")
+                return create_citation(label)
+
+
+def main(doc=None):
+    return pf.run_filter(action, doc=doc)
+
+
+if __name__ == "__main__":
+    main()

### 2019-06-13 scripts/pandoc_math.py
Polish pandoc_math filter and convert: pandoc -F ./scripts/pandoc_math.py chapter_01_decomposition.tex -o chapter_01_decomposition.md

@@ -5,7 +5,7 @@ import panflute as pf
 
 def create_citation(label, prefix=""):
     pandoc_label = f"[@{prefix}{label}]"
-    elem = pf.Cite(pf.Str(pandoc_label))
+    elem = pf.RawInline(pandoc_label)
     #  pf.debug(elem)
     return elem
 
@@ -23,7 +23,7 @@ def extract_latex_labels(text):
     if labels:
         return labels
     else:
-        pf.debug(f"No label in {text}")
+        return []  # "No label in {text}"
 
 
 def action(elem, doc):
@@ -41,24 +41,33 @@ def action(elem, doc):
     """
     if isinstance(elem, pf.Math) and elem.format == "DisplayMath":
         text = pf.stringify(elem)
-        if r"\label" in text:
-            env = "align"
-        else:
-            env = "align*"
-            # labels = extract_latex_labels(text)
+
+        for label in extract_latex_labels(text):
+            if not label.startswith("eq:"):
+                text = text.replace(rf'\label{{{label}}}', rf'\label{{eq:{label}}}')
 
         if r"\begin{aligned}" in text:
+            if r"\label" in text:
+                env = "align"
+            else:
+                env = "align*"
+
             return pf.RawInline(
                 "\n" + text.replace("{aligned}", f"{{{env}}}") + "\n", format="tex"
             )
+        else:
+            return pf.Math(text, format=elem.format)
 
     if doc.format.startswith("markdown"):
+        label = ""
         if isinstance(elem, pf.RawInline) and elem.format == "tex":
             label = pf.stringify(elem).lstrip(r"\eqref{").rstrip("}")
-            return create_citation(label)
+            return create_citation(label, 'eq:')
         elif isinstance(elem, pf.Link) and "reference-type" in elem.attributes:
-            if elem.attributes["reference-type"] in ("eqref", "ref"):
-                label = elem.url.lstrip("#")
+            label = elem.url.lstrip("#")
+            if elem.attributes["reference-type"] == "eqref":
+                return create_citation(label, "eq:")
+            elif elem.attributes["reference-type"] == "ref":
                 return create_citation(label)
 
 

### 2019-07-02 paper_04_shallow_water/Latex/article.tex
Rename and add figs of sw paper #!/bin/bash let num=1 git checkout -- article.tex for filelinenofig in $(rg --pcre2 -o --no-heading '(?<=Pyfig/)[a-zA-Z._0-9]*' article.tex) do   sed -i -E "s|$filelinenofig}|fig$num}|" article.tex   let num+=1 done git diff article.tex

@@
 \begin{figure}
-\centerline{\includegraphics[]{paper_04_shallow_water/Pyfig/fig_Emean_time}}
+\centerline{\includegraphics[]{paper_04_shallow_water/Pyfig/fig1}}
 \caption{Space averaged energy, $E = E_K + E_A $, versus time for all
@@
 \begin{figure}
-\centerline{\includegraphics[]{paper_04_shallow_water/Pyfig/fig_energy_w}}
+\centerline{\includegraphics[]{paper_04_shallow_water/Pyfig/fig2}}
 \caption{Left: Mean energy in the stationary state versus $ c $ for different
@@
 \begin{figure}
-\centerline{\includegraphics[]{paper_04_shallow_water/Pyfig/fig_flux_struct_combined}}
+\centerline{\includegraphics[]{paper_04_shallow_water/Pyfig/fig3}}
 \caption{Left: Time averaged normalised spectral energy fluxes versus $ k/k_f
@@
 \begin{figure}
-\centerline{\includegraphics[width=3in]{paper_04_shallow_water/Pyfig/fig_spatiotempspectra}}
+\centerline{\includegraphics[width=3in]{paper_04_shallow_water/Pyfig/fig4}}
 \caption{Frequency spectra of KE and APE for different $ k = \mid {\bf k} \mid
@@
 \begin{figure}
-\centerline{\includegraphics[]{paper_04_shallow_water/Pyfig/fig_phys_fields_wave}}
+\centerline{\includegraphics[]{paper_04_shallow_water/Pyfig/fig5}}
 \caption{Left: divergence $ \bnabla \cdot {\bf u} $. Right: velocity component in the $ y $-direction: $ u_y $. Top: run W6. Bottom: run W17.  }
@@
 \begin{figure}
-\centerline{\includegraphics[width=9cm]{paper_04_shallow_water/Pyfig/fig_shock_sep.pdf}}
+\centerline{\includegraphics[width=9cm]{paper_04_shallow_water/Pyfig/fig6}}
 \caption{Mean distance between the shocks as function of Froude number.  }
@@
 \centerline{
-\includegraphics[]{paper_04_shallow_water/Pyfig/fig_struct_order_246}}
+\includegraphics[]{paper_04_shallow_water/Pyfig/fig7}}
 \caption{From top to bottom: second, fourth and sixth order compensated and
@@
 \begin{figure}
-\centerline{\includegraphics[width=11cm]{paper_04_shallow_water/Pyfig/fig_ratio_strfct}}
+\centerline{\includegraphics[width=11cm]{paper_04_shallow_water/Pyfig/fig8}}
 \caption{
@@
 \begin{figure}
-\centerline{\includegraphics[]{paper_04_shallow_water/Pyfig/fig_flatness}}
+\centerline{\includegraphics[]{paper_04_shallow_water/Pyfig/fig9}}
 \caption{ Top: Flatness factor of transverse velocity increments. Bottom: Ratio between transverse and longitudinal flatness factor. Left: runs with $ n=3840 $: W3,
 W7, W11,  W14 and W18. Right: runs with $ n = 7860 $: W4, W8 and W15.  The dotted straight lines indicate the value 1.5 predicted by the shock model.  }
-\label{fig_flatness}
+\label{fig9}
 \end{figure}
@@
 \begin{figure}
-\centerline{\includegraphics[]{paper_04_shallow_water/Pyfig/fig_spectra}}
+\centerline{\includegraphics[]{paper_04_shallow_water/Pyfig/fig10}}
 \caption{Time averaged compensated and normalised energy spectra
@@
 \begin{figure}
-\centerline{\includegraphics[]{paper_04_shallow_water/Pyfig/fig_diss_spectra}}
+\centerline{\includegraphics[]{paper_04_shallow_water/Pyfig/fig11}}
 \caption{Time averaged normalised dissipation spectra.
@@
 $ r $, which is the theoretical single point limit for an irrotational
-isotropic field. In figure~\ref{fig_flatness} we see the flatness factor, $
+isotropic field. In figure~\ref{fig9} we see the flatness factor, $
 F_T $, of the transverse structure functions at the top and the ratio, $
@@
 \begin{figure}
-\centerline{\includegraphics[]{paper_04_shallow_water/Pyfig/fig_energy_lap}}
+\centerline{\includegraphics[]{paper_04_shallow_water/Pyfig/fig12}}
 \caption{Left: Mean energy in the stationary state versus $ c $ for different
@@
 \begin{figure}
-\centerline{\includegraphics[width=9cm]{paper_04_shallow_water/Pyfig/fig_shock_sep_lap.pdf}}
+\centerline{\includegraphics[width=9cm]{paper_04_shallow_water/Pyfig/fig13}}
 \caption{Mean distance between the shocks as function of Froude number.  }
@@
 \begin{figure}
-\centerline{\includegraphics[width=9cm]{paper_04_shallow_water/Pyfig/fig_flatness_div_norm_nu.pdf}}
+\centerline{\includegraphics[width=9cm]{paper_04_shallow_water/Pyfig/fig14}}
 \caption{Flatness of divergence $ \nabla \cdot {\bf u} $ times viscosity $ \nu $, versus $ F_f $. Dotted straight line indicates $ F_f^{2/3} $. }
@@
 \begin{figure}
-\centerline{\includegraphics[]{paper_04_shallow_water/Pyfig/fig_phys_fields_wave_lap}}
+\centerline{\includegraphics[]{paper_04_shallow_water/Pyfig/fig15}}
 \caption{$ \nabla \cdot {\bf u} $ for four different runs.

### 2019-07-02 paper_03_toy_model/article.tex

--- 2019-05-28 paper_03_toy_model/article.tex
@@
 Large scale geophysical flows are often mentioned as possible applications in the context of two-dimensional (2D) turbulence whose fundamental features were first predicted in the seminal paper  by \citet{Kraichnan1967}, which this Focus Issue is commemorating.
-\citet{Gage:1979} and \citet{Lilly:1983} suggested that the $ k^{-5/3} $ energy spectrum at atmospheric mesoscales, that are wave lengths from a few km up to five hundred km, measured by aircraft \citep{Nastrom-Gage:1985, Lindborg:1999}, is produced by an inverse energy cascade of similar type as described by \citet{Kraichnan1967}, a  hypothesis which recently was revived by \citet{Xia2011}.
+\citet{Gage:1979} and \citet{Lilly:1983} suggested that the $ k^{-5/3} $ energy spectrum at atmospheric mesoscales, that are wave lengths from a few km up to five hundred km, measured by aircraft \citep{Nastrom-Gage:1985, Lindborg1999}, is produced by an inverse energy cascade of similar type as described by \citet{Kraichnan1967}, a  hypothesis which recently was revived by \citet{Xia2011}.
 In the introduction of their excellent review article on 2D turbulence \citet{Boffetta2012} write that "large-scale motions in the atmosphere and oceans are described, to first approximation, as 2D turbulent fluids".

### 2019-07-09 paper_04_shallow_water/Latex/article.tex

@@
 The shocks we observed in the SW simulations left us in a state of curiosity.
-In some simulations they appeared to be pretty strong and in others rather weak.
+In some simulations they appeared to be rather strong and in others rather weak.
 In some simulations they were densely packed and in others sparsely distributed.
@@
 fluidsim}. In these twenty simulations we used an eight order diffusion operator to act as the agent of dissipation. 
- On the request from the reviewers we have also carried out a set of thirteen simulation with  Laplacian diffusion.
+ On the request from the reviewers we have also carried out a set of nineteen simulations with  Laplacian diffusion.
 The results from the latter set, which are similar to the results from the first set,  are presented in Appendix A.
@@
 W7, W11,  W14 and W18. Right: runs with $ n = 7860 $: W4, W8 and W15.  The dotted straight lines indicate the value 1.5 predicted by the shock model.  }
-\label{fig9}
+\label{fig_flatness}
 \end{figure}
@@
 $ r $, which is the theoretical single point limit for an irrotational
-isotropic field. In figure~\ref{fig9} we see the flatness factor, $
+isotropic field. In figure~\ref{fig_flatness} we see the flatness factor, $
 F_T $, of the transverse structure functions at the top and the ratio, $
@@
 
-\vskip 0.5 cm
-
-\noindent The authors would like to thank Gregory Falkovich and Sergey Nazaranko for  fruitful criticism of an early version of the manuscript and three anonymous reviewers for useful comments.
+\vskip 0.3 cm
+
+\noindent The authors would like to thank Gregory Falkovich and Sergey Nazarenko for  fruitful criticism of an early version of the manuscript, Sebastien Galtier for making a check of the final manuscript and three anonymous reviewers for useful comments.
 Financial
@@
 
-\label{Table2}
-
 \begin{tabular}{lrrrrrrrrr}
 \toprule
-{} &  $n$ &  $c$ &  $\nu $ &  $\epsilon$ &  $\frac{k_{diss}}{k_f}$ &   $F_f$ &  $\; Re$ &  $ReF_f^{2/3} $ &  $t_{\max}$ \\
+{} &   $n$ &  $c$ &  $\nu_2$ &  $\epsilon$ &  $\frac{k_{diss}}{k_f}$ &   $F_f$ &  $Re$ &  $ReF_f^{2/3}$ &  $t_{\max}$ \\
 \midrule
-WL1  &  960 &   10 &   0.0985 &       0.998 &                    22.6 &   0.16 &  68.0 &           20 &          25 \\
-WL2  & 1920 &   10 &   0.0492 &       0.985 &                    45.3 &    0.16 &  136 &           40 &          25 \\
-WL3  & 2880 &   10 &   0.0328 &       0.984 &                    67.9 &    0.16 &    204 &           60 &          25 \\
-WL4  & 3840 &   10 &   0.0246 &       0.993 &                    90.5 &   0.16 &   273 &            80 &          25 \\
-WL5  & 7680 &   10 &   0.0123 &       1.000 &                     181 &   0.16 &   545 &           160 &          25 \\
-WL6  &  960 &   20 &   0.0985 &       0.994 &                    22.6 &  0.08 &  68.1&            13 &          25 \\
-WL7  & 2880 &   20 &   0.0328 &       0.982 &                    67.9 &    0.08 &  204 &           38 &          25 \\
-WL8  &  960 &   40 &   0.0985 &       0.985 &                    22.6 &    0.04 &  68.0 &           8.0 &          25 \\
-WL9  & 2880 &   40 &   0.0328 &       0.981 &                    67.9 &    0.04 &  204 &            24 &          25 \\
-WL10 &  960 &  100 &   0.0985 &       0.976 &                    22.6 &   0.016 &  67.5 &            4.3 &          25 \\
-WL11 & 2880 &  100 &   0.0328 &       0.961 &                    67.9 &  0.016 &  202 &           13 &          25 \\
-WL12 &  960 &  200 &   0.0985 &       0.988 &                    22.6 & 0.0080 &  67.8 &          2.7 &          25 \\
-WL13 & 2880 &  200 &   0.0328 &       0.928 &                    67.9 & 0.0079 &  200 &           7.9&          25 \\
+WL1  &   960 &   10 &   0.0985 &       0.998 &                    22.6 &   0.161 &  68.0 &           20.1 &          25 \\
+WL2  &  1920 &   10 &   0.0492 &       0.985 &                    45.3 &   0.160 &   136 &           40.0 &          25 \\
+WL3  &  2880 &   10 &   0.0328 &       0.984 &                    67.9 &   0.160 &   203 &           59.9 &          25 \\
+WL4  &  3840 &   10 &   0.0246 &       0.993 &                    90.5 &   0.161 &   272 &           80.3 &          25 \\
+WL5  &  7680 &   10 &   0.0123 &        1.00 &                     181 &   0.161 &   545 &            161 &          25 \\
+WL6  &   960 &   20 &   0.0985 &       0.994 &                    22.6 &  0.0803 &  68.0 &           12.6 &          25 \\
+WL7  &  2880 &   20 &   0.0328 &       0.982 &                    67.9 &  0.0800 &   203 &           37.7 &          25 \\
+WL8  &  3840 &   20 &   0.0246 &       0.983 &                    90.5 &  0.0800 &   271 &           50.3 &          25 \\
+WL9  &  7680 &   20 &   0.0123 &        1.00 &                     181 &  0.0805 &   545 &            102 &          25 \\
+WL10 &   960 &   40 &   0.0985 &       0.985 &                    22.6 &  0.0400 &  67.7 &           7.93 &          25 \\
+WL11 &  2880 &   40 &   0.0328 &       0.981 &                    67.9 &  0.0400 &   203 &           23.7 &          25 \\
+WL12 &  3840 &   40 &   0.0246 &       0.978 &                    90.5 &  0.0399 &   270 &           31.6 &          25 \\
+WL13 &  7680 &   40 &   0.0123 &       0.991 &                     181 &  0.0401 &   543 &           63.6 &          25 \\
+WL14 &   960 &  100 &   0.0985 &       0.976 &                    22.6 &  0.0160 &  67.5 &           4.28 &          25 \\
+WL15 &  2880 &  100 &   0.0328 &       0.961 &                    67.9 &  0.0159 &   202 &           12.7 &          25 \\
+WL16 &  3840 &  100 &   0.0246 &       0.972 &                    90.5 &  0.0159 &   270 &           17.1 &          25 \\
+WL17 &   960 &  200 &   0.0985 &       0.988 &                    22.6 & 0.00801 &  67.8 &           2.72 &          25 \\
+WL18 &  2880 &  200 &   0.0328 &       0.928 &                    67.9 & 0.00785 &   199 &           7.87 &          25 \\
+WL19 &  3840 &  200 &   0.0246 &       0.934 &                    90.5 & 0.00786 &   266 &           10.5 &          25 \\
 \bottomrule
 \end{tabular}
+
+\label{Table2}
 
@@
 \caption{$ \nabla \cdot {\bf u} $ for four different runs.
-Upper left: WL1, $ Re = 68 $, $ F_f = 0.16 $; Upper right: WL4,  $ Re = 273 $, $ F_f = 0.16 $; Lower left: WL12, $ Re = 68 $, $ F_f = 0.008 $; Lower right: WL13,  $ Re = 200 $, $ F_f = 0.008 $.}
+Upper left: WL1, $ Re = 68 $, $ F_f = 0.16 $; Upper right:  { WL3,  $ Re = 203 $, $ F_f = 0.16 $;} Lower left: WL17, $ Re = 68 $, $ F_f = 0.008 $; Lower right: WL18,  $ Re = 199 $, $ F_f = 0.008 $.}
 \label{Physical_lap}
@@
 The variation of mean energy with $ c $ seems to be a little bit weaker for these runs as compared to the runs with hyper diffusion.
-However, the series of runs with $ n = 2280 $ (green crosses) is pretty close to approaching the $ c^{1/2} $-scaling.
+{ However, the higher resolution runs are relatively close to approaching the $ c^{1/2} $-scaling.}
 The variation of mean energy with resolution, or Reynolds number, is quite similar to what is seen in the corresponding plot \ref{MeanE}.
@@
 In figure \ref{fig_distance_lap} we see the mean distance between the shocks as a function of $ F_f $, together with a straight line indicating $ F_f^{1/2} $.
-There is a reasonable agreement between $ F_f ^{1/2} $ and each of the two series of simulations at $ Re \approx 70 $ (blue dots) and $ Re \approx 200$ (yellow triangles).
-However, there is a gap between the two series.
-This gap may be explained by the fact that the shock detection algorithm is much more sensitive to the threshold value which is used for $ \nabla \cdot {\bf u} $ in these simulations than in the simulations with hyperviscosity.
+{ As compared to the runs with hyperviscosity there is a larger spread of the points and the $ F_f^{1/2} $ scaling is not as clear.
+The lower Reynolds number series ($ Re \approx 70 $ blue dots) falls below the higher Reynolds number series.
+This may be explained by the fact that the shock detection algorithm is much more sensitive to the threshold value which is used for $ \nabla \cdot {\bf u} $ in these simulations than in the simulations with hyperviscosity.
 In particular, this is true in the lower Reynolds number runs.
-In figure \ref{FlatnessDiv} we see the flatness of the divergence, multiplied by $ \nu $, versus $ F_f $.
+} In figure \ref{FlatnessDiv} we see the flatness of the divergence, multiplied by $ \nu $, versus $ F_f $.
 According to (\ref{FD}) we should have $ F_{\partial u} \nu \propto  d^{4/3} $ and if $ d \propto Fr^{1/2} $, the points should collapse on a single line, $ F_f^{2/3} $.
-The three runs at highest $ Re $ and the lowest value of $ F_f $ fall on top of each other in the upper right corner, supporting the scaling $ \nu^{-1} $ in the high Reynolds number limit.
-For the three highest $ F_f $ there is a reasonable agreement with $ F_f^{2/3} $, although the lower Reynolds number runs (blue dots) fall below the higher Reynolds number runs, which we interpret as a low Reynolds number effect.
-The points corresponding to the two lowest values of $ F_f $ fall below the prediction.
+{For the three largest $ F_f $, there is a relatively good collapse of the points from the three highest  Reynolds numbers series, supporting both the $ \nu^{-1} $-scaling and the $ F_f^{2/3} $-scaling.
+}
+ The lower Reynolds number runs (blue dots) fall below the higher Reynolds number runs, which we interpret as a low Reynolds number effect.
+The points corresponding to the two lowest values of $ F_f $ fall below the prediction $ F_f^{2/3} $.
 This can be explained by the fact that the shocks are on the verge of fading away in these runs.
@@
 It should be pointed out again, that the area of the computational domain is sixteen times larger than area of the outcuts shown in the figures.
-The two upper plots are from simulations with  $ F_f = 0.16 $ but different $ Re $ ($ Re = 68 $ to the left and $ Re = 273 $ to the right).
+The two upper plots are from simulations with  $ F_f = 0.16 $ but different $ Re $ ($ Re = 68 $ to the left and { $ Re = 203 $ } to the right).
 As can be seen, the shocks are wider in the lower Reynolds number run to the left than in the higher Reynolds number run to the right, as expected.
@@
 In this simulation we have $ Re F_f^{2/3} = 2.7 $, so the condition (\ref{SCond}) is not fulfilled, and the shocks have almost faded away.
-The lower right plot is from a simulation with the same $ F_f $ as in the lower left plot, but with a higher $ Re $.
+The lower right plot is from a simulation with the same $ F_f $ as in the lower left plot, but with a higher $ Re = 199 $.
 Here, $ Re F_f^{2/3} = 7.8 $ and  the shocks are clearly discernible.
+{ Comparing the two right plots with each other, they have approximately the same $ Re $, while the lower one has a considerably lower $ F_f $.
+In this run the shocks are denser and straighter as compared to the higher $ F_f $-run.}
 

### 2019-07-23 latex/glossary.tex
Move stuff

\makeglossaries

%% ------------
%% Definitions
%% ------------
\newglossaryentry{zeta}{name=$\zeta$, description={absolute vorticity}}
\newglossaryentry{kappa}{name=$\kappa$, description={magnitude of wavenumber}}
\newglossaryentry{sigma}{name=$\sigma$, description={frequency of IG waves, $\sqrt{f^2+(c\kappa)^2}$}}
\newglossaryentry{J}{name=$\mathbf{J}$, description={mass flux, $h\mathbf{u}$}}
\newglossaryentry{M}{name=$\mathbf{M}$, description={displaced mass flux, $\eta\mathbf{u}$}}
\newglossaryentry{W}{name=$\mathbf{W}$, description={derived variable vector}}
\newglossaryentry{U}{name=$\mathbf{U}$, description={primitive variable vector}}
\newglossaryentry{B}{name=$\mathbf{B}$, description={normal mode vector, dimension: $LT^{-1}$}}
\newglossaryentry{N}{name=$\mathbf{N}$, description={normal mode vector, dimension: $T^{-1}$}}
\newglossaryentry{Xn}{name=$X_n$, description={normalized eigenvector matrix}}
\newglossaryentry{P}{name=$P$, description={transformation matrix from $\mathbf{W}$ to $\mathbf{U}$}}
\newglossaryentry{Q}{name=$Q$, description={transformation matrix from $\mathbf{U}$ to $\mathbf{B}$}}

% \glossarystyle{mcolindex}
\glsaddall

### 2019-07-27 imgs/spectral_flux/plot.py
Update seb for toy model

@@ -1,0 +1,60 @@
+import matplotlib
+import matplotlib.pyplot as plt
+from matplotlib.patches import Circle, Wedge, Polygon
+from matplotlib.collections import PatchCollection
+from math import pi
+import numpy as np
+from pathlib import Path
+
+root = Path(__file__).absolute().parent.parent
+
+plt.style.use("seaborn-paper")
+
+matplotlib.rcParams['text.latex.preamble'] = [r"\usepackage{amsmath}"]
+plt.rc('text', usetex=True)
+plt.rc('figure', dpi=300)
+
+SIDE_TEXT = False
+
+if SIDE_TEXT:
+    fig = plt.figure(figsize=(10,5))
+    fig.subplots_adjust(right=5/10)
+    ax = fig.add_subplot(111)
+else:
+    fig, ax = plt.subplots(figsize=(5,5))
+
+circle1 = Circle((0, 0), pi/2)
+circle2 = Circle((0, 0), pi/1.9)
+circle3 = Circle((0, 0), pi * 2)
+patches = [circle3, circle2, circle1]
+
+p = PatchCollection(patches, cmap="Pastel1")#, alpha=0.4)
+#  colors = np.array([54, 50, 20])
+p.set_array(np.array([0.1, 2, 9]))
+#  p.set_cmap(cmap)
+ax.add_collection(p)
+
+arror_r = pi / 2**1.5
+ax.arrow(arror_r, arror_r, 0.3, 0.3, head_width=0.05, head_length=0.1, fc='k', ec='k')
+tex = r"$\Pi(\mathbf{k}, t) = \sum\limits_{|\mathbf{k'}| \geq |\mathbf{k}|} T(\mathbf{k}', t)$"
+ax.text(0.9, 1.2, tex, fontsize=15, va='bottom', rotation=45)
+
+if SIDE_TEXT:
+    tex = '\n'.join([
+        r'$T=T_K + T_A$',
+        r'$T_K=-\frac{1}{2}ik_i(\hat{u_j}\widehat{u_iu_j}^* - \hat{u_j}^*\widehat{u_iu_j})$',
+        r'$T_A=-\frac{1}{2}ik_i(\hat{\phi}\widehat{u_i \phi}^* - \hat{\phi}^*\widehat{u_i\phi})$'
+    ])
[…]

### 2019-07-31 paper_06_milestone/1st/article.tex
Non-working try with milestone paper

\section {Introduction}

The meridional overturning circulation (MOC) of the ocean can be envisioned as
the product of two opposing processes. On the one hand, disruptions of
temperature and salt balance make surface water locally denser than deep water,
leading to isolated convective events of surface water sinking at high
latitudes. On the other hand, turbulent mixing, distributed throughout the
interior of the ocean, leads to a net upward transport of dense water. The
first process tends to lower the center of mass of the ocean while the second
process tends to lift it. Over time, the two processes must balance each other.
To understand and model the MOC it is essential to quantify the efficiency by
which turbulence is mixing the ocean \cite{Jayne}. As shown by \cite{Nilsson}
different parametrisations of the mixing efficiency in models can lead to very
different strengths of the MOC.

Generally, mixing efficiency should be understood as the ratio between the net
increase of potential energy produced by the mixing and the energy needed to
produce it \cite{Gregg}. To translate this understanding into a generally
accepted quantitative definition has been proven extremely difficult. Numerous
different definitions of mixing efficiency are used (e.g. \cite{Osborn,
OsbornCox, Caulfield}) in the large body of literature that has evolved in the
field. A good review of all different practices are given by [Gregg]. Following
[Gregg] we will use the measure \begin{equation} \Gamma =
\epsilon_P/\epsilon_K, \end{equation} where $ \epsilon_K $ and $ \epsilon_P $
are the mean kinetic energy dissipation and the mean dissipation of available
potential energy, respectively. Instead of calling $ \Gamma $ `mixing
efficiency' we call it the mixing coefficient. The canonical value $ \Gamma =
0.2 $ is often used in models. Values derived from observations, experiments
and numerical simulations are often in the range $ [0.1 \, \; 0.3] $, but
considerably smaller and larger values have also been reported. There is no
general consensus on the degree to which $ \Gamma $ should be regarded as a
constant and -- if not being a constant -- how it should be parametrised. As
discussed by \cite{IveyWinters}, apart from a possible dependence on Prandtl
number, $ \Gamma $ may depend on two independent parameters which can be taken
to be the buoyancy Reynolds number, $ {\cal{R}} $, and a turbulent Froude
number, $ F_h $, defined as \begin{equation} {\cal{R}} = \frac{\epsilon_K}{\nu
N^2} \, , \;\;\;\;\; F_h = \frac{U_h}{N L_h} \, , \end{equation} where $ U_h $
and $ L_h $ are characteristic horizontal turbulent velocity and length scales,
respectively, $ \nu $ is the kinematic viscosity and $ N $ the
Brunt-V\"ais\"al\"a frequency. Alternatively, one may assume that $ \Gamma $
depends on the Reynolds number, $ Re = U_h L_h / \nu $, and $ F_h $. The use of
$ {\cal{R}} $ instead of $ Re $ has a clear advantage in the limit of strong
stratification, where $ {\cal{R}} $ may become very small, while $ Re $ still
is large. It is generally agreed
(e.g. \cite{BrethouwerBillantLindborg2007}) that turbulence is very weak
in flows with $ {\cal{R}} $ of the order of unity and smaller. Therefore, $
\Gamma $ will go to zero in the limit of small $ {\cal{R}} $, even though $ Re
$ may be large. On the other hand, it can be argued that the use of $ {\cal{R}}
$ can lead to confusions when the limit of weak stratification is considered.
Analysing results from direct numerical simulations, Shih et al. \cite{Shih} argued
that $ \Gamma $ goes to zero as $ {\cal{R}} ^{-1/2} $ in the limit of large $
{\cal{R}} $, a hypothesis which has been widely referenced
(e.g. \cite{IveyWinters}). Plotting mixing efficiency calculated as a function
of $ {\cal{R}} $ using data from a tank experiments of grid generated
turbulence Barry et al. \cite{Barry} concluded that $ \Gamma \sim
{\cal{R}}^{-2/3} $ for large $ {\cal{R}} $. The problem with such
interpretations is that the supposed decrease of $ \Gamma $ at large $
{\cal{R}} $ is likely to be an effect of weak stratification (large $ F_h $)
rather than an effect of an increasing $ Re $. In general, turbulence
quantities are expected to exhibit Reynolds number similarity for large
Reynolds numbers, suggesting that $ \Gamma $ should not vary with $ {\cal{R}} $
if the degree of stratification is held constant, that is if $ F_h $ is held
constant. Maffioli et al. \cite{Maffioli2016} suggested that $ \Gamma $ should
become independent of the buoyancy Reynolds number as long as it is above a
certain limit (approximately equal to ten), in which case $ \Gamma $ should
only depend on $ F_h $. In the limit of small $ F_h $ (strong stratification) $
\Gamma $ should approach a constant value and in the opposite limit $ \Gamma $
should go to zero as $ F_h^{-2} $. Maffioli et al \cite{Maffioli2016} supported
these predictions by reporting on a series of direct numerical simulations
(DNS). Recently, these predictions have also gained support from an analysis of
other DNS data \cite{Garanaik2019}.

The regime of strongly stratified turbulence, in which $ F_h $ is small and $
{\cal{R}} $ is large, is highly relevant for applications in the
ocean\cite[]{RileyDeBruynKops2003, Lindborg2006}, but is notoriously difficult
to reproduce in experiments and simulations. In trying to reach the strongly
stratified regime experimentally by increasing the degree of stratification,
the buoyancy Reynolds number often becomes so low that turbulence is totally
surpressed. Observations made in this regime suggest that $ \Gamma $ is
generally very small. Ivey \& Imberger \cite{IveyImberger1991} plotted mixing
efficiency derived from different laboratory measurements
\cite{Stillinger1983, Itsweire1986, Rohr1988, Lienhard1990} against a turbulent Froude
number. In the limit of large Froude number, the $ F_h^{-2} $-dependence is
clearly visible in all plots. At Froude number of the order of unity the
measurements indicate that $ \Gamma \approx 0.2 $, while at lower Froude
numbers, $ \Gamma $, or the related Richardson flux coefficient [Osborn], is
rapidly dropping below unity. In all likelihood, such observations are
artefacts of the limitation which is present in all observational studies of
strongly stratified turbulence, namely the severe difficulty in reaching the
regime where $ F_h $ is small while $ {\cal{R}} $ still is above a certain
limit. In the present study, we make an effort to push the limits further and
measure the mixing coefficient in a parameter regime which previously has not
been reached in laboratory studies. In particular, we focus on the regime, $ 10
< {\cal R} < 200 $, with $ F_h $ being as small as possible.

\section{Introduction bis}

Why is the mixing coefficient $\Gamma$ important?

Ocean models are LES (scale filter $[]$).

\begin{itemize}

\item Approximation of a term similar to a Reynolds stress with a turbulent diffusivity

$$- \bnabla \cdot [\vv b_{tot}] \simeq \kappa_t \bnabla^2 [b_{tot}] $$

\item Approximation of the turbulent diffusivity from a flux law for the buoyancy flux:

$$\mean{w b_{tot}} \simeq - \kappa_t d_z \bar b_{tot} = - \kappa_t N^2  \Rightarrow \kappa_t = \frac{\CKA}{N^2}.$$

\item  Approximation of the energy conversion $\CKA$ by a proportionality relation

$$ \CKA = \Gamma \epsK, $$
with $\Gamma = 0.2$ a constant!
\end{itemize}

\section{Experimental methods}

\subsection{Experimental goal: estimate mixing coefficient in strongly
stratified turbulence}

Before presenting in details the experimental setup, let us present our main
experimental goal and some experimental consequences.

We want to estimate the the mixing coefficient $\Gamma = \eps_P/\eps_K$ for
flows closed to the geophysical regimes which is consistent with oceanic
dynamics. We now know \cite{BrethouwerBillantLindborg2007, Maffioli2016} that
we need to obtain flows associated with small horizontal Froude number $F_h$
and relatively large buoyancy Reynolds number $\R$. In a laboratory experiment,
we are limited in two aspects. In term of strength of stratification with water
and salt, we cannot reach a Brunt-V\"ais\"al\"a frequency larger than
$\sim1$~rad/s. We are of course strongly limited in term of size $L_h$.

The Coriolis platform is a huge rotating platform (13~m
diameter) designed to study rotating and stratified flows. Its large scale
allows us to study geophysical turbulence at large Reynolds number and
relatively small $F_h$ and $Ro$.

To estimate the mixing coefficient $\Gamma = \eps_P/\eps_K$, we would need to
measure the kinetic energy dissipation rate $\eps_K$ and the APE dissipation
rate $\eps_P$. These two quantities are associated with viscous dissipation and
salt diffusion happening at the smallest scales of the flow. It would be very
difficult to accurately measure these two quantities in strongly stratified
turbulence.

Instead of trying to evaluate the mixing coefficient from small-scale
quantities, we estimate large-scale quantities, namely the decay of kinetic
energy and the long-term global mixing, i.e. the increase of potential energy.

To evaluate the average decay of kinetic energy after one stroke, we need a
good representation of the large scale flow field and many realizations.
Therefore, we design an experiment for which the fluid is periodically stirred
with large cylinders. We focus on scanned horizontal PIV because it is adapted
to obtain a good evaluation of the averaged kinetic energy.

Regarding the long-term global mixing, we need to measure the long-term
evolution of the density profiles after many strokes, from which can be
computed the increase of potential energy.

Figure~\ref{fig:average:mixing} illustrates the effect of a mixing period on
the horizontally averaged density profiles. The values of depth and density are
representative of our experiments in the Coriolis platform. The tank is
initially filled with a stable linear density stratification due to a
stratification in salt concentration. Some mixing is triggered by stirring the
flow, i.e. injecting kinetic energy. After some time, the horizontally averaged
density profile has evolved and is no longer linear. There are two region at
the top and the bottom where the density is higher and lower, respectively,
than at the beginning of the experiment. There has been a net flux of salt from
the bottom to the top and the center of mass has been lifted by 2 millimeters.

In this case, the mixing coefficient can be evaluated as $\Gamma =
\frac{|\Delta E_{Pb}|}{\Delta E_K}$, where $\Delta E_K$ is the kinetic energy
dissipation and $\Delta E_{Pb}$ the increase background potential energy.

\begin{figure}
\includegraphics[width=\figwidth]{tmp/fig_scheme_mixing}
\caption{Scheme representing the effect of mixing on the density profile and the center of mass.}
\label{fig:average:mixing}
\end{figure}

\subsection{Overview of the experiment}

The experimental setup, installed in the large Coriolis platform, is presented
in figure~\ref{fig:exp}: a tank with a rectangle base of 9$\times$6~m$^2$ area
is filled with an approximately linear salt stratification of $81$~cm depth.

The flow is generated with an oscillating comb. Two different combs were used:
the first one in made of 6 vertical cylinders of $25$~cm diameter attached to a
carriage with a mesh of $M=75$~cm, and for the second one, there are 4
cylinders of $50$~cm diameter with a mesh of $M=1.5$~m. We impose on the
carriage the following periodic motion of period $T$ (see figure~\ref{fig:exp}
for a chronogram of its position $x_c$): it accelerates with a constant
acceleration over 25~cm to a velocity $U_c$, travels at the uniform speed a
distance of 650~cm and decelerates with a constant deceleration over 25~cm. It
them comes back to its initial position with the opposite movement and
immediately restarts this cycle. In the following, we use cartesian coordinates
with the origin centered with respect to the vertical walls and at the bottom
of the tank. The vertical coordinate is given by $z$ and the horizontal
coordinate along the direction of displacement of the carriage by $x$ such that
$(x,y,z)$ form right handed coordinates.

We use in this study results from two different experimental campains, carried
out in 2016 and 2017. In the following, the two sets of experiments are denoted
as MILESTONE 2016 (``M16") and MILESTONE 2017 (``M17"). The M16 campain has
already been described in \cite{campagne2016}. The two experimental setups are
nearly identical. They differ mainly on the precise locations of the density
and temperature probes and on the number and the precise locations of the
cameras used for Particle Image Velocity (PIV). For the second set of
experiments (M17), the carriage is accelerated over 0.5 cm and travels at
constant speed over 5~m. The constant speed is varied from 1 cm/s up to 24
cm/s.

\begin{figure}[htb]
\centering
\includegraphics[width=80mm]{../figs/scheme_milestone17}\\[5mm]
\includegraphics[width=80mm]{../figs/milestone17_top}
% \includegraphics[width=\figwidth]{../figs/real_setup_milestone}
\includegraphics[width=80mm]{tmp/fig_movement_carriage}

\caption{Top: scheme of the experimental setup for the MILESTONE 2017
experiments (see text for description). Center: photography of the experiment.
Bottom: chronogram of the position of the carriage $x_c(t)$.}

\label{fig:exp}
\end{figure}

Density is measured with temperature and conductivity probes which are either
static (to measure the density at the top and the bottom of the tank), attached
to the carriage or attached to traverses to obtained density profiles.

To be able to accurately measure the mixing, the slow experiments are very long
($\sim$ 5 hours, i.e. more than 10 periods of oscillation of the carriage). In
contrast, the mixing is much faster for the larger velocity so it is difficult
to maintain the forcing during more than 2 or 3 periods of oscillation and
therefore to get a lot of statistics.

The velocity is measured by Particle Image Velocity (PIV). While we used 5
cameras in the M16 campain, for M17 experiments, we use only 3 cameras. One is
dedicated to record images for horizontal scanning PIV with up to 16 vertical
levels. The two other cameras are used for stereoscopic vertical 2D PIV.

\input{tmp/table.tex}

\subsection{Open-data and open-source}

% Ashwin can write this part...

\input{sec_open_science.latex}
% An interesting particularity of the MILESTONE experiment is to be based on
% open-source method and to provide open datasets and open-source packages to
% load and analyze them.
%
% Two open-source Python packages called FluidLab and FluidCoriolis
% \cite{fluiddyn} were developed for these experiments and used to control most
% of the objects during the experiments.
%
% For example the movement of the carriage and of the probes are controlled with
% a graphical application and Python scripts provided by fluidlab and
% FluidCoriolis. Horizontal scanning PIV is also made possible by controling the
% rotating mirror and the triggers of the camera by functions provided in
% fluidlab. This allows anyone to perform horizontal scanning PIV with a good
% camera (here a PCO Edge), a rotating mirror and a quite cheap acquisition board
% (T7 LabJack) to trigger the camera.
%
% FluidLab is a generic API for orchestrating laboratory experiments. The
% software leverages the object-oriented programming features in Python to model
% real life instrumentation. An experiment in the simplest level can be thought
% of as a network of interconnected instruments awaiting commands and also
% sending and receiving data.
%
% The computation of PIV fields is performed on the cluster of the LEGI with an
% open-source Python package called fluidimage
%
% Data processing and data opening with FluidCoriolis.

\section{Results}

\subsubsection{Estimating the kinetic energy dissipation rate $\eps_K$}

As already stated, we have not been able to measure the local energy
dissipation. \cite{PraudFinchamSommeria2005} used 3D PIV to compute some terms
of the local kinetic dissipation from velocity fields but only for very slow
flows, corresponding to very small buoyancy Reynolds number. However, when the
velocity is increased, the flow becomes more turbulent and the Kolmogorov
length scale decreases so that it becomes much more difficult to compute
accurately the velocity gradient from PIV velocity field.

Therefore, we are forced to estimate the kinetic energy dissipation rate from
the decay of the kinetic energy, which is a large-scale quantity. By averaging
the energy computed from several horizontal PIV fields (different vertical
levels and periods of the carriage), we can evaluate the averaged time
evolution of the kinetic energy after one stroke and finally the averaged
kinetic energy dissipation.

\begin{figure}
\includegraphics[width=80mm]{tmp/fig_energy_vs_time}
\label{fig:energy:vs:time}

\caption{Kinetic energy decay for experiment M16-73.}

\end{figure}

Figure~\ref{fig:energy:vs:time} shows the time evolution of the mean kinetic
energy in the region scanned by the horizontal PIV. For this experiment
(M16-73), the carriage oscillates during 4 periods and is then stopped.

\begin{figure}
\includegraphics[width=80mm]{tmp/fig_fit_EK}
\label{fig:fit:EK}

\caption{Fits of the kinetic energy decay for M16-73 (same as in
figure~\ref{fig:energy:vs:time}). The time derivatives of the kinetic energy in
(b) are normalized by $\eps_c = U_c^3 / D_c$.}

\end{figure}

Figure~\ref{fig:fit:EK}(a) shows the space and phase averaged kinetic energy as
a function of $(t - t_0)/T$ and a fit of the data $a / (1 + t^\alpha)/\beta$.
Figure~\ref{fig:fit:EK}(b) presents the time derivative of the kinetic energy
and the same quantity obtained from the fit of the kinetic energy. We finally
compute by averaging over time for $t < 0.5 T$, $\eps_m= \mean{d_t E_{Kh}}$, an
estimation of the kinetic energy dissipation averaged over few turnover times
after the strokes.

\begin{figure}
\includegraphics[width=\figwidth]{tmp/fig_R_vs_Fh}
\label{fig:RvsFh}

\caption{Evaluation of the buoyancy Reynolds number $\R$ and the horizontal
Froude number $F_h$ from the horizontal kinetic energy $E_{Kh}$ and its decay
$\eps_m = \mean{d_t E_{Kh}}$ for the M16 experiments.}

\end{figure}

Together with the mean Brunt-V\"ais\"al\"a frequency computed from the density
profiles, we are able to estimate the values of $\R$ and $F_h$ for each
experiment. The values of $E_{Kh}$, $\eps_m$, $\R$ and $F_h$ are plotted in
figure~\ref{fig:RvsFh}. Horizontal and vertical grey bars have been added to
delimit the 3 regimes of stratified flows, namely strongly stratified
turbulence ($F_h < 0.03$), weakly stratified turbulence ($F_h > 0.03$), and
low-$\R$ stratified flows (affected by dissipation at large horizontal scale).

We see that with these estimations, none of the M16 experiments is really in
the strongly stratified turbulence regime. However, some experiments are close to
the thresholds and we have to stress that these estimates of $E_{Kh}$ and
$\eps_K$ are computed from a time average for $t < 0.5 T$.

\begin{figure}
\includegraphics[width=\figwidth]{tmp/fig_R_vs_Fh_other_studies_with_milestone17}
\label{fig:RvsFh:other}

\caption{Buoyancy Reynolds number $\R$ versus horizontal Froude number $F_h$
for few M16 and M17 experiments and for other experimental and numerical
studies of stratified turbulence.}

\end{figure}

In figure~\ref{fig:RvsFh:other}, we compare the typical values of the
non-dimensional numbers reached in these experiments with values evaluated for
other studies.

\subsubsection{Evaluation of the mixing coefficient}

Similarly as for the kinetic energy dissipation, we have not been able to
measure the different terms of the density gradient, which are needed to
compute the local available potential energy dissipation. In order to estimate
the mixing, we measure the rate of increase of the average background potential
energy.

\begin{figure}
\includegraphics[width=\figwidth]{tmp/fig_rho_vs_time}\\
\includegraphics[width=\figwidth]{tmp/fig_rho_vs_time_corrected}
\label{fig:rho:vs:time}

\caption{Density signals measured by the 5 density probes for experiment M17-21
corresponding to $D = 0.5$~m, $N=0.55$~rad/s and $U=12$~cm/s. Three probes are
attached to vertical profilers. The two other fixed probes are at the top and
at the bottom, respectively. (a) raw signals and (b) signals corrected to
compensate a derive of the probes.
}

\end{figure}

To evaluate this quantity, we need a good measurement of the evolution of the
space-average density profile. Figure~\ref{fig:rho:vs:time}(a) shows the
density signals of 5 probes. Three of these probes are attached to traverses
and the two others are attached at the top and the bottom of the tank. We see
that some probes do not give accurate measurements of the density, so we
correct these measurements with the more precise probes as show in
figure~\ref{fig:rho:vs:time}(b).

\begin{figure}
\includegraphics[width=\figwidth]{tmp/fig_profiles_mixing}\\
\includegraphics[width=\figwidth]{tmp/fig_profiles_probe_averaged}
\label{fig:profiles:mixing}

\caption{Evolution of the density profiles for experiment M17-21 corresponding
to $D = 0.5$~m, $N=0.55$~rad/s and $U=12$~cm/s.}

\end{figure}

The space-averaged density profiles computed from these measurements are
displayed in figure~\ref{fig:profiles:mixing}(a). The shape of these profiles
at the top and the bottom is especially important to compute the background
potential energy. However, the probes attached to profilers cannot measure the
density very close to the top and bottom boundaries. The profiles are extended
with values computed from their extrema values as presented in
figure~\ref{fig:profiles:mixing}(b).

\begin{figure}
\includegraphics[width=\figwidth]{tmp/fig_energy_pot_vs_time}
\label{fig:energy:pot:vs:time}

\caption{Evolution of the normalized potential energy for experiment M17-21
corresponding to $D = 0.5$~m, $N=0.55$~rad/s and $U=12$~cm/s.}

\end{figure}

The evolution of the background potential energy is shown in
figure~\ref{fig:energy:pot:vs:time} for the same experiment M17-21. The value
-1 would correspond to a linear stratification with the same initial
Brunt-V\"ais\"al\"a frequency. Since the profiles at the beginning of the
experiment are already eroded, the initial normalized potential energy is
slighly smaller than -0.6. We see that it increases approximately linearly with
time while the fluid is stirred and stabilize at a constant value after the
stop of the carriage. From this curve, we can evaluate the mean rate of
increasement of the average background potential energy $\eps_P$ during an
experiment.

\begin{figure}
\includegraphics[width=\figwidth]{tmp/fig_dt_pot_energy}
\label{fig:dt:pot:energy}

\caption{Normalized mixing coefficient $\eps_P / (3\times10^{-3} {U_c}^3/D_c)$
for some MILESTONE 17 experiments.}

\end{figure}

This quantity normalized by an estimation of the dissipation of kinetic energy
$3\times10^{-3} {U_c}^3/D_c$ (see figure~\ref{fig:RvsFh}) is plotted as a
function of the Froude number in figure~\ref{fig:dt:pot:energy}. This quantity
is approximately proportional to the mixing coefficient $\Gamma$. The colors
represent the buoyancy Reynolds number such that the yellow points correspond
to $\R > 15$. We see that the yellow points fall on a same curve, while there
are more variations for smaller values of buoyancy Reynolds number. The yellow
points are more consistent with a ${F_h}^{-1}$ than with the ${F_h}^{-2}$
scaling law observed and predicted with \cite{Maffioli2016}.

\section{Conclusions}

???

### 2019-07-31 paper_06_milestone/1st/py/make_fig_R_vs_Fh.py

@@ -1,0 +1,104 @@
+"""Make the figure fig_R_vs_Fh.png
+
+"""
+
+import numpy as np
+
+from util import plt, get_colors, iexps_norot, show, save
+
+from fluidcoriolis.milestone.results_energy_budget import ResultEnergyBudgetExp
+
+iexps = iexps_norot
+
+colors = get_colors(len(iexps))
+
+fig = plt.figure(figsize=(9, 8))
+
+rect = [0.1, 0.08, 0.56, 0.5]
+ax0 = fig.add_axes(rect)
+
+rect = [0.1, 0.67, 0.38, 0.31]
+ax1 = fig.add_axes(rect)
+
+rect[0] = 0.6
+ax2 = fig.add_axes(rect)
+
+axes = [ax0, ax1, ax2]
+
+xlim = [3e-3, 0.1]
+ylim = [1e-1, 2e2]
+
+Flim = 3e-2
+Rlim = 10
+color = [0.9] * 3
+linewidth = 20
+ax0.plot([Flim] * 2, [Rlim, ylim[1]], "-", color=color, linewidth=linewidth)
+ax0.plot(xlim, [Rlim] * 2, "-", color=color, linewidth=linewidth)
+
+
+print(iexps)
+
+for i, iexp in enumerate(iexps):
+    print("\niexp = ", iexp)
+    results = ResultEnergyBudgetExp(iexp, "PCO_top", extract=True)
+
+    exp = results.exp
+    if exp.N == 0.8:
+        symbol = "o"
+    elif exp.N == 0.56:
+        symbol = "s"
+    else:
+        symbol = "v"
+    ax0.plot(
+        results.Fht,
+        results.Rt,
+        symbol,
+        color=colors[i],
+        label=r"exp {}, $N =$ {:.2f} rad/s, $U_c =$ {:.0f} cm/s".format(
+            iexp, exp.N, 100 * exp.Uc
+        ),
+    )
+
+    if ~np.isnan(results.Fht):
+        ax0.text(results.Fht, results.Rt, str(iexps[i]), fontsize=8)
+        ax1.errorbar(
+            results.Fht,
+            results.eps / exp.epsc,
+            yerr=results.deltaeps / exp.epsc,
+            fmt=symbol,
+            ecolor=colors[i],
+        )
[…]

### 2019-07-31 paper_06_milestone/1st/py/make_fig_dt_pot_energy.py

@@ -1,0 +1,99 @@
+
+from util import show, save
+
+import numpy as np
+import matplotlib.pyplot as plt
+
+from fluidcoriolis.milestone17.time_signals import SignalsExperiment, comments
+
+from fluidcoriolis.milestone17.exp import iexps_strat, create_exp
+
+nu = 1e-6
+
+C_eps = 2e-3
+C_E = 4e-2
+
+C_eps = 3e-3
+C_E = 4e-2
+
+"""
+
+eps_K = C_eps * Uc**3 / Dc
+
+E_K = C_E * Uc**2
+
+
+"""
+
+
+iexps = [iexp for iexp in iexps_strat
+         if iexp in comments and comments[iexp].startswith('ok')]
+
+iexps_bad = (
+    24,  # too mixed at the beginning
+    57, 58,  # horizontal cylinders
+    15,  # not long enough
+)
+
+
+iexps = [iexp for iexp in iexps if iexp not in iexps_bad]
+
+print(iexps)
+
+exps = [create_exp(iexp) for iexp in iexps]
+
+exps = [exp for exp in exps if exp.Uc > 0.01]
+
+signals_exps = [SignalsExperiment(exp) for exp in exps]
+
+dt_energies = np.zeros([len(exps)])
+Fhs = np.zeros_like(dt_energies)
+Ds = np.zeros_like(dt_energies)
+Rs = np.zeros_like(dt_energies)
+for i, signals in enumerate(signals_exps):
+    _, _, _, _, _, dt_e = signals.calcul_energy_pot_vs_time()
+
+    dt_energies[i] = dt_e/C_eps
+    Fhs[i] = C_eps/C_E*signals.exp.Fhc
+    Ds[i] = signals.exp.Dc
+    Rs[i] = C_eps * signals.exp.epsc / (nu * signals.exp.N**2)
+
+fig = plt.figure(figsize=(9, 5))
+ax = fig.gca()
+
+plot = ax.scatter
+
+cond = Ds == 0.25
+plot(Fhs[cond], dt_energies[cond], c=Rs[cond], marker='o', vmin=0, vmax=20)
+
+cond = Ds == 0.5
+sc = plot(Fhs[cond], dt_energies[cond], c=Rs[cond], marker='s', vmin=0, vmax=20)
+
+for i, exp in enumerate(exps):
[…]

### 2019-07-31 paper_06_milestone/1st/py/make_fig_energy_vs_time.py

@@ -1,0 +1,29 @@
+
+from util import plt, show, save
+
+cam = 'PCO_top'  # MILESTONE16
+
+year = 2016
+
+assert year in (2016, 2017)
+
+if year == 2016:
+    from fluidcoriolis.milestone.results_energy_budget import ResultEnergyBudgetExp
+elif year == 2017:
+    from fluidcoriolis.milestone17.results_energy_budget import ResultEnergyBudgetExp
+
+if year == 2016:
+    cam = 'PCO_top'
+    iexp = 73
+else:
+    cam = 'Cam_horiz'
+    iexp = 18
+
+r = ResultEnergyBudgetExp(iexp, camera=cam)
+
+r.plot_energy_vs_time(vertical_grid=True, mean_over_levels=True)
+fig = plt.gcf()
+fig.tight_layout()
+
+save(fig, 'fig_energy_vs_time.png')
+show()

### 2019-07-31 paper_06_milestone/1st/py/make_fig_scheme_mixing.py

@@ -1,0 +1,68 @@
+from util import show, save
+
+import numpy as np
+import matplotlib.pyplot as plt
+
+g = 9.81
+rho_min = 1.0
+N = 0.5
+H = 0.8
+
+l_nomixed = 0.35
+
+zs = np.linspace(0, H, 40) - H
+zmid = H / 2
+
+# approximation...
+dz_rho = -N ** 2 * rho_min / g
+rhos = rho_min + dz_rho * zs
+
+rho0 = rho_min - dz_rho * zmid
+rhos1 = rho0 + l_nomixed * dz_rho * np.tanh((zs + zmid) / l_nomixed)
+
+z0 = np.trapz(rhos * zs, zs) / np.trapz(rhos, zs)
+z1 = np.trapz(rhos1 * zs, zs) / np.trapz(rhos1, zs)
+
+rho_max = rhos.max()
+rho_min = rhos.min()
+
+fig = plt.figure(figsize=(8, 3.5))
+ax0 = fig.add_subplot(121)
+ax1 = fig.add_subplot(122)
+
+ax0.set_title("initial condition (linear strat.)", {"color": "b"})
+ax1.set_title("effect of mixing", {"color": "r"})
+
+axes = [ax0, ax1]
+
+ax0.plot(rhos, zs, "b")
+# ax0.plot([rho_min, rho_max], [z0]*2, 'b')
+# ax0.plot([rho_min, rho_max], [-zmid]*2, 'k:')
+
+
+def addtext(ax, z, color):
+    ax.text(
+        1.005,
+        -0.6,
+        "center of mass\n" + f"{1000*z:.1f} mm",
+        {"size": 14.0, "color": color, "horizontalalignment": "center"},
+    )
+
+
+addtext(ax0, z0, "b")
+
+ax1.plot(rhos, zs, "b:")
+ax1.plot(rhos1, zs, "r")
+# ax1.plot([rho_min, rho_max], [z1]*2, 'r')
+# ax1.plot([rho_min, rho_max], [-zmid]*2, 'k:')
+
+addtext(ax1, z1, "r")
+
+for ax in axes:
+    ax.set_xlabel(r"$\bar\rho(z)$ (kg/l)")
+    ax.set_ylabel(r"$z$ (m)")
+
+fig.tight_layout(pad=0.2, w_pad=1.0)
+
+save(fig, "fig_scheme_mixing.png")
+show()

### 2019-07-31 paper_06_milestone/1st/py/make_figs.py

@@ -1,0 +1,73 @@
+from pathlib import Path
+import os
+import sys
+from datetime import datetime
+from runpy import run_path
+
+
+def modification_date(filename):
+    """Get the modification date of a file"""
+    return datetime.fromtimestamp(os.path.getmtime(str(filename)))
+
+
+def has_to_build(output_file: Path, input_file: Path):
+    """Check if a file has to be (re)built"""
+    output_file = Path(output_file)
+    input_file = Path(input_file)
+    if not output_file.exists():
+        return True
+    mod_date_output = modification_date(output_file)
+    if mod_date_output < modification_date(input_file):
+        return True
+    return False
+
+
+sys.argv.append("save_for_tex")
+
+here = Path(__file__).parent.absolute()
+
+scripts = sorted(here.glob("make_fig_*"))
+
+path_tmp = here.parent / "tmp"
+
+scripts = [
+    script for script in scripts if script.name != "make_fig_other_studies.py"
+]
+
+figure_names = {path: path.name[5:-3] + ".png" for path in scripts}
+
+for script, figure in figure_names.items():
+
+    if has_to_build(path_tmp / figure, script):
+        print("run", script)
+        run_path(str(script))
+
+
+script = here / "make_fig_other_studies.py"
+
+figures = [
+    "fig_R_vs_Fh_other_studies_with_milestone",
+    "fig_R_vs_Fh_other_studies_with_milestone17",
+]
+
+for figure in figures:
+    path_fig = path_tmp / (figure + ".png")
+    if has_to_build(path_fig, script):
+        sys.argv.append(figure.split("with_")[1])
+        print("run", script)
+        run_path(script)
+        sys.argv.pop()
+
+
+figures = [
+    "fig_rho_vs_time.png",
+    "fig_rho_vs_time_corrected.png",
+    "fig_profiles_mixing.png",
+    "fig_profiles_probe_averaged.png",
+    "fig_energy_pot_vs_time.png",
+]
+script = here / "make_figs_probes.py"
+
+if any(has_to_build(path_tmp / figure, script) for figure in figures):
+    print("run", script)
+    run_path(script)

### 2019-07-31 paper_06_milestone/1st/py/make_figs_probes.py

@@ -1,0 +1,39 @@
+from util import plt, show, save
+
+from fluidcoriolis.milestone17.time_signals import SignalsExperiment
+
+
+def unset_title(fig):
+    fig.axes[0].set_title(None)
+
+
+iexp = 21
+signals = SignalsExperiment(iexp)
+
+signals.plot_vs_times()
+fig = plt.gcf()
+unset_title(fig)
+save(fig, "fig_rho_vs_time.png")
+
+signals.plot_vs_times(corrected=1)
+fig = plt.gcf()
+unset_title(fig)
+save(fig, "fig_rho_vs_time_corrected.png")
+
+probe = signals.probes_profiles[0]
+probe.plot_profiles(corrected=1)
+fig = plt.gcf()
+save(fig, "fig_profiles_mixing.png")
+
+signals.plot_profiles_probe_averaged(
+    corrected=1, sort=1, extend=1, len_extend=0.005
+)
+fig = plt.gcf()
+save(fig, "fig_profiles_probe_averaged.png")
+
+signals.plot_energy_pot_vs_time()
+fig = plt.gcf()
+unset_title(fig)
+save(fig, "fig_energy_pot_vs_time.png")
+
+show()

### 2019-07-31 paper_06_milestone/1st/py/make_table.py

@@ -1,0 +1,69 @@
+from pathlib import Path
+
+from fluidcoriolis.milestone import Experiment
+from fluidcoriolis.milestone17.exp import iexps_strat, create_exp
+from fluidcoriolis.milestone17.time_signals import comments
+
+from util import path_tmp, iexps_norot
+
+path_tmp = Path(path_tmp)
+
+lines = []
+
+for iexp in iexps_norot:
+    exp = Experiment(iexp)
+    lines.append(
+        fr" M16-{exp.iexp:02d} &  {100 * exp.Uc:2.0f} &  "
+        fr"{100 * exp.Dc:2.0f} &  "
+        fr"{exp.N_measured:.2f} &  {exp.Fhc:.3f} &  {exp.Rec:5.0f} &  "
+        fr"{exp.Rc:5.0f}"
+    )
+
+
+iexps = [
+    iexp
+    for iexp in iexps_strat
+    if iexp in comments and comments[iexp].startswith("ok")
+]
+
+iexps_bad = (
+    24,  # too mixed at the beginning
+    57,
+    58,  # horizontal cylinders
+    15,  # not long enough
+)
+
+iexps = [iexp for iexp in iexps if iexp not in iexps_bad]
+
+for iexp in iexps:
+    exp = create_exp(iexp)
+    if exp.Uc == 0:
+        continue
+
+    lines.append(
+        fr" M17-{exp.iexp:02d} &  {100 * exp.Uc:2.0f} &  "
+        fr"{100 * exp.Dc:2.0f} &  "
+        fr"{exp.N:.2f} &  {exp.Fhc:.3f} &  {exp.Rec:5.0f} &  "
+        fr"{exp.Rc:5.0f}"
+    )
+
+code_line = " \\\\\n".join(lines)
+
+latex_code = rf"""
+\begin{{table}}
+\caption{{\label{{table:exp}} Experiments used for this study.}}
+\begin{{ruledtabular}}
+\begin{{tabular}}{{l{6 * 'c'}}}
+    &   $Uc$ &  $D_c$ &  $N$  &  $F_{{hc}}$ &  $Re_c$ &   $\R_c$  \\
+    & (cm/s) &   (cm) &(rad/s)&           &         &           \\
+
+{code_line} \\
+\end{{tabular}}
+\end{{ruledtabular}}
+\end{{table}}
+"""
+
+# print(latex_code)
+
+with open(path_tmp / "table.tex", "w") as file:
+    file.write(latex_code)

### 2019-07-31 paper_06_milestone/1st/py/other_studies/almakie2012.py

@@ -1,0 +1,37 @@
+
+import numpy as np
+
+label = r'Almakie et al. (2012, $\sim 4096^2\times 1024$)'
+symbol = 'go'
+
+epsK = np.array([0.87, 0.64, 0.64])
+
+epsKh = np.array([0.58, 0.43, 0.44])
+
+Eh = np.array([1.56, 1.54, 1.74])
+
+Uh = np.sqrt(Eh)
+
+F4 = np.array([0.0615, 0.0231, 0.0105])
+
+Fht = F4*epsK/epsKh
+
+Rt = np.array([222.6, 40.62, 10.19])
+
+Ret = Rt/Fht**2
+
+nu = Eh**2/(epsK*Ret)
+
+N = epsKh/(F4*Eh)
+
+ko = np.sqrt(N**3/epsK)
+
+kd = (epsK/nu**3)**(1./4)
+
+nh = 4096
+Lh = 2*np.pi
+Delta = Lh/nh
+C_dealiasing = 2./3
+kmax = C_dealiasing*np.pi*nh/Lh
+
+kb = N/Uh

### 2019-07-31 paper_06_milestone/1st/py/other_studies/gargett1981.py

@@ -1,0 +1,35 @@
+
+
+import numpy as np
+
+label = 'Gargett et al. (1981, ocean)'
+symbol = 'rv'
+
+# N (rad/s)
+N = np.array([3.94e-3, 3.96e-3, 4.59e-3])
+
+# epsK (cm^2 s^{-3})
+epsK = np.array([7.4e-6, 1.4e-5, 3.9e-5])
+epsK *= 1e-4 # (m^2 s^{-3})
+
+# ko (rad/m)
+ko0 = np.array([9.1, 6.7, 5.0])
+
+# Phib = sqrt(epsK*N) (s^{-2} m^{-1})
+Phib0 = np.array([1.7e-6, 2.3e-6, 4.2e-6])  
+
+Phib = np.sqrt(epsK*N)
+
+ko = np.sqrt(N**3/epsK)
+
+nu = 1.1e-6  # (m^2/s)
+
+Rt = epsK/(nu*N**2)
+
+kb = np.array([0.18351, 0.17514, 0.22024])
+
+Uh = N / kb
+
+Eh = Uh**2
+
+Fht = epsK / (N * Eh)

### 2019-07-31 paper_06_milestone/1st/py/other_studies/milestone.py

@@ -1,0 +1,10 @@
+
+import numpy as np
+
+
+label = 'Milestone (2016, exp.)'
+symbol = 'rs'
+
+Rt = np.array([2.4, 3.4, 9.5, 16, 60])
+
+Fht = np.array([1.3, 2.2, 3, 4.2, 9.8]) * 1e-2

### 2019-07-31 paper_06_milestone/1st/py/other_studies/riley2003.py

@@ -1,0 +1,18 @@
+
+import numpy as np
+
+label = 'Riley et al. (2003, $\sim 512^3$)'
+symbol = 'mo'
+
+F = 4.
+N = 2 * np.pi / F
+
+Re_l = np.array([1600., 3200, 6400])
+nu = 1./Re_l
+
+epsK = 3.e-3 * np.ones_like(Re_l)
+E = 0.06
+
+Fht = epsK / (N * E)
+
+Rt = epsK / (nu * N**2)

### 2019-07-31 paper_06_milestone/1st/py/pandoc_auto_height.py

@@ -1,0 +1,49 @@
+import panflute as pf
+
+
+def extract_image_from_pdf_page(xObject):
+    """https://gist.github.com/gstorer/f6a9f1dfe41e8e64dcf58d07afa9ab2a"""
+    xObject = xObject['/Resources']['/XObject'].getObject()
+
+    for obj in xObject:
+        o = xObject[obj]
+        if o['/Subtype'] == '/Image':
+            return o
+        else:
+            return extract_image_from_pdf_page(o)
+
+
+def detect_height(elem):
+    url = elem.url
+    width = int(elem.attributes["width"].rstrip('%'))
+    if url.endswith(".pdf"):
+        from PyPDF2 import PdfFileReader
+        with open(url, "rb") as fp:
+            pdf = PdfFileReader(fp)
+            w, h = pdf.getPage(0).mediaBox[-2:]
+            # o = extract_image_from_pdf_page(pdf.getPage(0))
+            # w, h = (o['/Width'], o['/Height'])
+    else:
+        import imageio
+        w, h, colors = imageio.imread(url).shape
+
+    height = f"{width*h/w}%"
+    return height
+
+
+def action(elem, doc):
+    """Automatically determines height attribute for images."""
+    if doc.format == 'latex' and isinstance(elem, pf.Image):
+        # pf.debug(elem)
+        if "width" in elem.attributes and "height" not in elem.attributes:
+            elem.attributes["height"] = detect_height(elem)
+
+        return elem
+
+
+def main(doc=None):
+    return pf.run_filter(action, doc=doc)
+
+
+if __name__ == '__main__':
+    main()

### 2019-07-31 paper_06_milestone/1st/article.tex
Paper 6 compiles

@@
 number, $ \Gamma $ may depend on two independent parameters which can be taken
-to be the buoyancy Reynolds number, $ {\cal{R}} $, and a turbulent Froude
-number, $ F_h $, defined as \begin{equation} {\cal{R}} = \frac{\epsilon_K}{\nu
+to be the buoyancy Reynolds number, $ {\mathcal{R}} $, and a turbulent Froude
+number, $ F_h $, defined as \begin{equation} {\mathcal{R}} = \frac{\epsilon_K}{\nu
 N^2} \, , \;\;\;\;\; F_h = \frac{U_h}{N L_h} \, , \end{equation} where $ U_h $
@@
 depends on the Reynolds number, $ Re = U_h L_h / \nu $, and $ F_h $. The use of
-$ {\cal{R}} $ instead of $ Re $ has a clear advantage in the limit of strong
-stratification, where $ {\cal{R}} $ may become very small, while $ Re $ still
+$ {\mathcal{R}} $ instead of $ Re $ has a clear advantage in the limit of strong
+stratification, where $ {\mathcal{R}} $ may become very small, while $ Re $ still
 is large. It is generally agreed
 (e.g. \cite{BrethouwerBillantLindborg2007}) that turbulence is very weak
-in flows with $ {\cal{R}} $ of the order of unity and smaller. Therefore, $
-\Gamma $ will go to zero in the limit of small $ {\cal{R}} $, even though $ Re
-$ may be large. On the other hand, it can be argued that the use of $ {\cal{R}}
+in flows with $ {\mathcal{R}} $ of the order of unity and smaller. Therefore, $
+\Gamma $ will go to zero in the limit of small $ {\mathcal{R}} $, even though $ Re
+$ may be large. On the other hand, it can be argued that the use of $ {\mathcal{R}}
 $ can lead to confusions when the limit of weak stratification is considered.
 Analysing results from direct numerical simulations, Shih et al. \cite{Shih} argued
-that $ \Gamma $ goes to zero as $ {\cal{R}} ^{-1/2} $ in the limit of large $
-{\cal{R}} $, a hypothesis which has been widely referenced
+that $ \Gamma $ goes to zero as $ {\mathcal{R}} ^{-1/2} $ in the limit of large $
+{\mathcal{R}} $, a hypothesis which has been widely referenced
 (e.g. \cite{IveyWinters}). Plotting mixing efficiency calculated as a function
-of $ {\cal{R}} $ using data from a tank experiments of grid generated
+of $ {\mathcal{R}} $ using data from a tank experiments of grid generated
 turbulence Barry et al. \cite{Barry} concluded that $ \Gamma \sim
-{\cal{R}}^{-2/3} $ for large $ {\cal{R}} $. The problem with such
+{\mathcal{R}}^{-2/3} $ for large $ {\mathcal{R}} $. The problem with such
 interpretations is that the supposed decrease of $ \Gamma $ at large $
-{\cal{R}} $ is likely to be an effect of weak stratification (large $ F_h $)
+{\mathcal{R}} $ is likely to be an effect of weak stratification (large $ F_h $)
 rather than an effect of an increasing $ Re $. In general, turbulence
 quantities are expected to exhibit Reynolds number similarity for large
-Reynolds numbers, suggesting that $ \Gamma $ should not vary with $ {\cal{R}} $
+Reynolds numbers, suggesting that $ \Gamma $ should not vary with $ {\mathcal{R}} $
 if the degree of stratification is held constant, that is if $ F_h $ is held
@@
 The regime of strongly stratified turbulence, in which $ F_h $ is small and $
-{\cal{R}} $ is large, is highly relevant for applications in the
+{\mathcal{R}} $ is large, is highly relevant for applications in the
 ocean\cite[]{RileyDeBruynKops2003, Lindborg2006}, but is notoriously difficult
@@
 strongly stratified turbulence, namely the severe difficulty in reaching the
-regime where $ F_h $ is small while $ {\cal{R}} $ still is above a certain
+regime where $ F_h $ is small while $ {\mathcal{R}} $ still is above a certain
 limit. In the present study, we make an effort to push the limits further and
@@
 been reached in laboratory studies. In particular, we focus on the regime, $ 10
-< {\cal R} < 200 $, with $ F_h $ being as small as possible.
+< \mathcal{R} < 200 $, with $ F_h $ being as small as possible.
 
@@
 
-\begin{figure}
+\begin{figure}[htp!]
 \includegraphics[width=\figwidth]{tmp/fig_scheme_mixing}
@@
 
-\begin{figure}[htb]
+\begin{figure}[htp!]
 \centering
@@
 
-\input{tmp/table.tex}
+\input{paper_06_milestone/1st/tmp/table.tex}
 
@@
 
-\input{sec_open_science.latex}
+\input{paper_06_milestone/1st/sec_open_science.latex}
 % An interesting particularity of the MILESTONE experiment is to be based on
@@
 
-\begin{figure}
+\begin{figure}[htp!]
 \includegraphics[width=80mm]{tmp/fig_energy_vs_time}
@@
 
-\begin{figure}
+\begin{figure}[htp!]
 \includegraphics[width=80mm]{tmp/fig_fit_EK}
@@
 
-\begin{figure}
+\begin{figure}[htp!]
 \includegraphics[width=\figwidth]{tmp/fig_R_vs_Fh}
@@
 
-\begin{figure}
+\begin{figure}[htp!]
 \includegraphics[width=\figwidth]{tmp/fig_R_vs_Fh_other_studies_with_milestone17}
@@
 
-\begin{figure}
+\begin{figure}[htp!]
 \includegraphics[width=\figwidth]{tmp/fig_rho_vs_time}\\
@@
 
-\begin{figure}
+\begin{figure}[htp!]
 \includegraphics[width=\figwidth]{tmp/fig_profiles_mixing}\\
@@
 
-\begin{figure}
+\begin{figure}[htp!]
 \includegraphics[width=\figwidth]{tmp/fig_energy_pot_vs_time}
@@
 
-\begin{figure}
+\begin{figure}[htp!]
 \includegraphics[width=\figwidth]{tmp/fig_dt_pot_energy}

### 2019-07-31 paper_06_milestone/1st/article.tex
Typeset paper6

@@
 \begin{figure}[htp!]
+\centering
 \includegraphics[width=\figwidth]{tmp/fig_scheme_mixing}
-\caption{Scheme representing the effect of mixing on the density profile and the center of mass.}
+\caption{Scheme representing the effect of mixing on the density profile and
+the center of mass.}%
 \label{fig:average:mixing}
@@
 experiments (see text for description). Center: photography of the experiment.
-Bottom: chronogram of the position of the carriage $x_c(t)$.}
-
+Bottom: chronogram of the position of the carriage $x_c(t)$.}%
 \label{fig:exp}
@@
 \begin{figure}[htp!]
-\includegraphics[width=80mm]{tmp/fig_energy_vs_time}
+\centering
+\includegraphics[width=0.5\textwidth]{tmp/fig_energy_vs_time}
+
+\caption{Kinetic energy decay for experiment M16-73.}%
 \label{fig:energy:vs:time}
-
-\caption{Kinetic energy decay for experiment M16-73.}
 
@@
 \begin{figure}[htp!]
+\centering
 \includegraphics[width=80mm]{tmp/fig_fit_EK}
-\label{fig:fit:EK}
 
@@
 figure~\ref{fig:energy:vs:time}). The time derivatives of the kinetic energy in
-(b) are normalized by $\eps_c = U_c^3 / D_c$.}
+(b) are normalized by $\eps_c = U_c^3 / D_c$.}%
+\label{fig:fit:EK}
 
@@
 \begin{figure}[htp!]
-\includegraphics[width=\figwidth]{tmp/fig_R_vs_Fh}
-\label{fig:RvsFh}
+\centering
+\includegraphics[width=\figwidth]{tmp/fig_R_vs_Fh}%
 
@@
 Froude number $F_h$ from the horizontal kinetic energy $E_{Kh}$ and its decay
-$\eps_m = \mean{d_t E_{Kh}}$ for the M16 experiments.}
+$\eps_m = \mean{d_t E_{Kh}}$ for the M16 experiments.}%
+\label{fig:RvsFh}
 
@@
 \begin{figure}[htp!]
-\includegraphics[width=\figwidth]{tmp/fig_R_vs_Fh_other_studies_with_milestone17}
-\label{fig:RvsFh:other}
+\centering
+\includegraphics[width=\figwidth]{tmp/fig_R_vs_Fh_other_studies_with_milestone17}%
 
@@
 for few M16 and M17 experiments and for other experimental and numerical
-studies of stratified turbulence.}
+studies of stratified turbulence.}%
+\label{fig:RvsFh:other}
 
@@
 \begin{figure}[htp!]
-\includegraphics[width=\figwidth]{tmp/fig_rho_vs_time}\\
-\includegraphics[width=\figwidth]{tmp/fig_rho_vs_time_corrected}
-\label{fig:rho:vs:time}
+\centering
+\includegraphics[width=0.7\textwidth]{tmp/fig_rho_vs_time}\\
+\includegraphics[width=0.7\textwidth]{tmp/fig_rho_vs_time_corrected}
 
@@
 compensate a derive of the probes.
-}
+}%
+\label{fig:rho:vs:time}
 
@@
 \begin{figure}[htp!]
-\includegraphics[width=\figwidth]{tmp/fig_profiles_mixing}\\
-\includegraphics[width=\figwidth]{tmp/fig_profiles_probe_averaged}
-\label{fig:profiles:mixing}
+\centering
+\includegraphics[width=0.7\figwidth]{tmp/fig_profiles_mixing}
+\includegraphics[width=0.7\figwidth]{tmp/fig_profiles_probe_averaged}
 
 \caption{Evolution of the density profiles for experiment M17-21 corresponding
-to $D = 0.5$~m, $N=0.55$~rad/s and $U=12$~cm/s.}
+to $D = 0.5$~m, $N=0.55$~rad/s and $U=12$~cm/s.}%
+\label{fig:profiles:mixing}
 
@@
 \begin{figure}[htp!]
-\includegraphics[width=\figwidth]{tmp/fig_energy_pot_vs_time}
-\label{fig:energy:pot:vs:time}
+\centering
+\includegraphics[width=0.7\figwidth]{tmp/fig_energy_pot_vs_time}
 
 \caption{Evolution of the normalized potential energy for experiment M17-21
-corresponding to $D = 0.5$~m, $N=0.55$~rad/s and $U=12$~cm/s.}
+corresponding to $D = 0.5$~m, $N=0.55$~rad/s and $U=12$~cm/s.}%
+\label{fig:energy:pot:vs:time}
 
@@
 \begin{figure}[htp!]
+\centering
 \includegraphics[width=\figwidth]{tmp/fig_dt_pot_energy}
-\label{fig:dt:pot:energy}
 
 \caption{Normalized mixing coefficient $\eps_P / (3\times10^{-3} {U_c}^3/D_c)$
-for some MILESTONE 17 experiments.}
+for some MILESTONE 17 experiments.}%
+\label{fig:dt:pot:energy}
 

### 2019-08-04 abstract.tex
Fix frontmatter

\noindent%
Two studies of two-dimensional models of flows influenced by stratification and stratification/rotation are carried out in
order to investigate whether a two-dimensional model can reproduce a downscale energy cascade with an associated $ k^{-5/3} $ wavenumber  spectrum.
 Firstly, a series of highly resolved numerical simulations of the classical shallow water model is carried out. A forward energy cascade is observed but the dynamics is dominated by shocks, with an associated $ k^{-2} $-spectrum.
A theory for shallow water wave turbulence is formulated and compared to the results from the simulations.
 Secondly, a series of simulations of a new two-dimensional toy model is carried out, showing that  the model is not generating  shocks and can reproduce a downscale energy cascade with
an associated $ k^{-5/3} $ spectrum. The energy transfer is studied in detail in Fourier space and is compared with results from a general circulation model.

An experimental study of strongly stratified turbulence at the Coriolis platform in Grenoble is carried out, with the aim of testing novel theories of stratified turbulence. Turbulence is generated by traversing an array of cylinders through a tank containing stratified salt water. Velocity is measured by Particle Image Velocimetry (PIV) and density is measured by conductivity probes. In particular, the author has developed the software system analysing the PIV images. Preliminary results from the experiment are presented.

 To realise the research objectives, a set of open-source software packages are developed in Python, under the
umbrella of the FluidDyn project. The
packages enable execution of simulations, experiments and processing of data.
The codes are well documented, tested and designed to
promote development and reuse.

\keywords{geophysical flows, shallow water wave turbulence, energy cascade,
stratified turbulence, waves and vortices, open source software}

### 2019-08-04 abstrakt.tex
Fix frontmatter

\noindent%
Tv\aa $ \; $  studier av tv\aa dimensionella modeller av str\"omningar
influerade av stratifiering respektive stratifiering och rotation genomf\"ors,
i syfte att unders\"oka huruvida en {tv\aa}dimensionell modell kan reproducera en energikaskad till mindre skalor med ett associerat v\aa gtalsspektrum av formen $ k^{-5/3} $.  F\"orst genomf\"ors en serie h\"oguppl\"osta simuleringar av den klassiska modellen f\"or str\"omningar i grunt vatten (shallow water model). En energikaskad till mindre skalor observeras, men str\"omningen domineras av st\"otar med ett associerat $ k^{-2} $-spektrum.
Sedan genomf\"ors en serie simuleringar av en ny tv\aa dimensionell leksaksmodell, som visar att modellen kan reproducera en energikaskad till mindre skalor med ett associerat $ k^{-5/3}$-spektrum. Energi\"overf\"oringen studeras i detalj i Fourier-rummet och j\"amf\"ors med resultat fr\aa n en global cirkulationsmodell.

En experimentell studie av starkt stratifierad turbulens genomf\"ors vid Coriolisplattformen i Grenoble, med m\aa let att testa nya teorier f\"or stratifierad turbulens. Turbulensen genereras genom att traversera en rad cylindrar genom en tank inneh\aa llande stratifierat saltvatten. Hastighet m\"ats med Particle Image Velocimetry (PIV) och densitet m\"ats med konduktivitetssonder. I synnerhet har f\"orfattaren utvecklat den programvara som analyserar bilder fr\aa n PIV-m\"atningarna. Prelimin\"ara resultat fr\aa n experimentet presenteras.

F\"or att realisera forskningsm\aa len  utvecklas en m\"angd mjukvarupaket f\"or \"oppet bruk i Python, i ett projekt med namnet FluidDyn. Paketen erbjuder m\"ojligheter att utf\"ora simuleringar, experiment och databehandling. Koderna \"ar v\"al dokumenterade, testade och designade f\"or att underl\"atta utveckling och \aa teranv\"andning.

\nyckelord{geofysikaliska str\"omningar, v\aa gturbulens i grunt vatten,
stratifierad turbulens, v\aa gor och virvlar, \"oppen mjukvara}

### 2019-08-04 frontmatter.tex
Fix frontmatter

@@
 \dedication{
-	{\it ``You must unlearn what you have learnt.''}\\[20pt]
-	Master Yoda
+	% {\it ``You must unlearn what you have learnt.''}\\[20pt]
+	% Master Yoda
+	{\it ``Essentially, all models are wrong, but some are useful.''}\\[20pt]
+	George E.~P.~Box
 }
@@
 \begin{abstract}
-        To do
+    \input{abstract}
 	% Access points  must work. In fact, few cryptographers would disagree with
@@
 \begin{abstrakt}
-        To do
+    \input{abstrakt}
 	% \r{A}tkomstpunkter m\r{a}ste arbeta. I sj\"{a}lva verket skulle f\r{a}
@@
 \begin{preface}
-        % TODO: preface
-        To do.
-	% This thesis deals with high energy super-lasers. A brief introduction on the
-	% basic concepts and methods is presented in the first part. The second part
-	% contains two articles.
-	% The papers are adjusted to comply with the present thesis format for
-	% consistency, but their contents have not been altered as compared with their
-	% original counterparts.
+    This thesis contains numerical, theoretical, experimental and
+    technical studies aiming at advancing the understanding of stratified flows.
+    Studies which motivated this thesis, along with theoretical
+    details and research highlights of the present work are summarized in the
+    first part. The second part contains five articles and one conference
+    paper.
+    The papers are adjusted to comply with the present thesis format for
+    consistency, but their contents have not been altered as compared with their
+    original counterparts.
 \end{preface}
@@
 %
-%TODO: division of work
 \begin{divisionofwork}
-	The main advisor for the project is Erik Lindborg (EL).
+	The main advisor for the thesis is Erik Lindborg (EL).
     Pierre Augier (PA) acts as co-advisor.
 
-        To do: paper wise list
-	% \paperitem
-	%         The code has been developed by Anakin Skywalker (AS).
-	%         The paper has been written by AS and Darth Vader (DV).
+	\paperitem
+        The code was developed by PA and Ashwin Vishnu
+        Mohanan (AVM). The simulations and post-processing were performed
+        by AVM. The theory was developed by PA and EL with input from AVM. A
+        first version of the paper was written by PA. The final version was
+        written by EL with input from AVM.
         %
-	% \paperitem
-	%         The experimental set-up has been designed by DS.
-	%         The simulations have been performed by AS using the control-code
-	%         developed by DS.
-	%         The paper has been written by AS and DV with feedback from OK and DS.
+	\paperitem
+        The code development, the simulations and post-processing was
+        handled by AVM. The paper was written by EL with input from
+        AVM.
+        %
+    \paperitem
+        The experiment was planned by the group from KTH together with the group from
+        LEGI. The experimental set-up was put together by
+        the group from LEGI. The experiment was carried out by the two groups
+        together.
+        The code for image-processing was developed by AVM, PA, Antoine
+        Campagne (AC), Joel Sommeria (JS) and the code for processing
+        probe data was developed by PA, RC and Diane Micard. The
+        post-processing was performed by AVM, PA, AC, AS and JS.
+        The paper was written by AC with input from PA, EL and AVM.
+        %
+    \paperitem
+        The code was developed by PA, Cyrille Bonamy (CB) and AVM. The paper
+        was authored by PA with input from AVM.
+    \paperitem
+        The code was developed by AVM, CB and PA.
+        The code was benchmarked for performance by PA and AVM.
+        The paper was written by AVM with inputs from PA and CB.
+    \paperitem
+        The code was developed by AVM, CB, Miguel Calpe Linares (MCL) and PA.
+        The code was benchmarked for performance by PA and AVM.
+        The paper was written by AVM with inputs from PA and CB.
 

### 2019-08-04 scripts/render_paper_yml.py
Corrections to chapter intro

@@ -40,6 +40,12 @@ def name_in_initials(name):
     return " ".join(name)
 
 
+def ordered_set(iterable):
+    from collections import OrderedDict
+
+    return OrderedDict.fromkeys(iterable).keys()
+
+
 def get_options_from_yaml(path):
     """Read a options YAML file."""
     import ruamel.yaml as yaml
@@ -51,11 +57,11 @@ def get_options_from_yaml(path):
     if "dir" not in options or options["dir"] is None:
         options["dir"] = path.parent
 
-    if len(options["authors"]) < 4:
-        options["authors_short"] = [
+    if len(options["authors"]) <= 3:
+        authors_short = ', '.join(
             name_in_initials(name) for name in options["authors"][:-1]
-        ]
-        options["authors_short"].extend(
+        )
+        options["authors_short"] = [authors_short] + (
             [r"\&", name_in_initials(options["authors"][-1])]
         )
     else:
@@ -64,7 +70,7 @@ def get_options_from_yaml(path):
             "et al.",
         ]
 
-    uniq_affiliations = list(set(options["affiliations"]))  # unique
+    uniq_affiliations = list(ordered_set(options["affiliations"]))  # unique
     options["authors_aff"] = []
     for author, affiliation in zip(options["authors"], options["affiliations"]):
         options["authors_aff"].append(

### 2019-08-05 frontmatter.tex
TODO and some last min corrections

@@
 %
-% TODO: abstract in english and swedish
 \begin{abstract}
@@
         Campagne (AC), Joel Sommeria (JS) and the code for processing
-        probe data was developed by PA, RC and Diane Micard. The
+        probe data was developed by PA, R\'emi Chassagne and Diane Micard. The
         post-processing was performed by AVM, PA, AC, AS and JS.
@@
     \paperitem
-        The code was developed by AVM, CB, Miguel Calpe Linares (MCL) and PA.
+        The code was developed by AVM, CB, Miguel Calpe Linares and PA.
         The code was benchmarked for performance by PA and AVM.
@@
 
-%TODO: fill in additional publications and conferences
+%TODO: fill in additional datasets
 %-------------------------------------------------------------------------------

### 2019-08-05 paper_03_toy_model/article.tex

@@
 \end{eqnarray}
-{ \noindent Following \citet{Deusebio-Augier-Lindborg:2013} the total energy transfer can be decomposed into four terms as
+{ \noindent Following \citet{Deusebio-Vallgren-Lindborg:2013} the total energy transfer can be decomposed into four terms as
 \begin{equation}
@@
 However, the kinetic energy content in the rotational component of the velocity field is necessarily smaller than in the divergent component, and that is at odds with observational evidence \citep{Lindborg:2015, QiangLi2017}.
-There is also numerical evidence \citep{Deusebio-Augier-Lindborg:2013, Asselin2017} suggesting that the forward energy mesoscale energy cascade cannot be described as a cascade of waves at high wave numbers, although it may start as a wave cascade at low wave numbers.
+There is also numerical evidence \citep{Deusebio-Vallgren-Lindborg:2013, Asselin2017} suggesting that the forward energy mesoscale energy cascade cannot be described as a cascade of waves at high wave numbers, although it may start as a wave cascade at low wave numbers.
 In any case, as we pointed out earlier, the forward energy cascade in the GCM cannot be produced by inertia-gravity waves with frequencies close to $ f $, since it would require much finer resolution in the vertical to simulate such motions.

### 2019-08-05 scripts/render_glossary_yml.py
Glossary jinja

@@ -1,0 +1,57 @@
+#!/usr/bin/env python
+"""Render a latex jinja template
+
+Usage:
+    render_glossary_yml.py <options_file> <template_file>
+
+Notes:
+    See https://github.com/AKuederle/Py-Tex-automation-example
+
+"""
+import os
+import sys
+from pathlib import Path
+
+from render_paper_yml import get_template, render_template
+
+
+
+def get_options_from_yaml(path):
+    """Read a options YAML file."""
+    import ruamel.yaml as yaml
+
+    yml = yaml.YAML()
+    with open(path) as f:
+        options = yml.load(f.read())
+
+    # post-process options
+    post_options = []
+    for label, description in sorted(options.items()):
+        if label.startswith("text_"):
+            label = name = label.lstrip('text_')
+        else:
+            name = f'${label}$'
+            label = label.lstrip('\\').lstrip('mathbf')
+
+        post_options.append((label, name, description))
+
+    return post_options
+
+
+if __name__ == "__main__":
+    try:
+        from docopt import docopt
+
+        argv = docopt(__doc__)
+        options_file = Path(argv["<options_file>"])
+        template_file = Path(argv["<template_file>"])
+    except ImportError:
+        argv = sys.argv
+        options_file = Path(argv[1])
+        template_file = Path(argv[2])
+
+    options = get_options_from_yaml(options_file)
+    template = get_template(template_file)
+
+    out_file = os.fspath(options_file.parent / options_file.stem) + ".tex"
+    render_template(template, {"entries": options}, out_file)

### 2019-08-06 scripts/render_glossary_yml.py
Glossary and acronyms

@@ -15,6 +15,57 @@ from pathlib import Path
 from render_paper_yml import get_template, render_template
 
 
+GREEK = [
+    "Alpha",
+    "Beta",
+    "Gamma",
+    "Delta",
+    "Epsilon",
+    "Zeta",
+    "Eta",
+    "Theta",
+    "Iota",
+    "Kappa",
+    "Lamda",
+    "Mu",
+    "Nu",
+    "Xi",
+    "Omicron",
+    "Pi",
+    "Rho",
+    "Sigma",
+    "Tau",
+    "Upsilon",
+    "Phi",
+    "Chi",
+    "Psi",
+    "Omega",
+    "alpha",
+    "beta",
+    "gamma",
+    "delta",
+    "epsilon",
+    "zeta",
+    "eta",
+    "theta",
+    "iota",
+    "kappa",
+    "lamda",
+    "mu",
+    "nu",
+    "xi",
+    "omicron",
+    "pi",
+    "rho",
+    "sigma",
+    "tau",
+    "upsilon",
+    "phi",
+    "chi",
+    "psi",
+    "omega",
+]
+
 
 def get_options_from_yaml(path):
     """Read a options YAML file."""
@@ -25,17 +76,26 @@ def get_options_from_yaml(path):
         options = yml.load(f.read())
 
     # post-process options
-    post_options = []
-    for label, description in sorted(options.items()):
-        if label.startswith("text_"):
-            label = name = label.lstrip('text_')
+    glossary = []
+    acronym = []
+    for label, description in options.items():
+        if label.startswith("acr_"):
+            label = name = label.replace("acr_", "")
+            acronym.append((label, name, description))
         else:
-            name = f'${label}$'
-            label = label.lstrip('\\').lstrip('mathbf')
+            name = f"${label}$"
+            for char in r"\{^},":
+                label = label.replace(char, "")
+            for word in ("mathbf", "mathcal", "hat"):
+                if word in label:
+                    label = label.replace(word, "") + "_" + word
+            if any(label.startswith(greek_alphabet) for greek_alphabet in
+                   GREEK):
+                label = "aa_" + label
 
-        post_options.append((label, name, description))
+            glossary.append((label, name, description))
 
-    return post_options
+    return sorted(glossary), sorted(acronym)
 
 
 if __name__ == "__main__":
@@ -50,8 +110,10 @@ if __name__ == "__main__":
         options_file = Path(argv[1])
         template_file = Path(argv[2])
 
[…]

### 2019-08-13 chapter_02_open_science.tex
Change figure captions and positions
More coreections and improved figures in exp chapter

--- 2019-06-05 chapter_00_open_science.tex
@@
 It is normal that Mathworks, the company which makes money with Matlab, spreads
-\fnref{https://www.mathworks.com/products/matlab/matlab-vs-python.html}{arguments
+\fnref{https://web.archive.org/web/20190807004347/https://www.mathworks.com/products/matlab/matlab-vs-python.html}{arguments
 against Python and its scientific ecosystem}.  Some of them are valid, for
@@
 
-% TODO: perhaps include the commented out sections in the beginning of the chapter
 % pa: I am really uncomfortable with this presentation which gives the
@@
 \fnref{http://pivlab.blogspot.de/}{PIVlab},
-\fnref{http://servforge.legi.grenoble-inp.fr/projects/soft-uvmat/}{UVmat} and
+\fnref{http://servforge.legi.grenoble-inp.fr/projects/soft-uvmat/}{UVMAT} and
 \fnref{https://de.mathworks.com/matlabcentral/fileexchange/10902-pivmat-4-00}{PIVmat}).
@@
 % \section{Implementation and architecture}
-% TODO: a lot of self-plagiarism below - needs rewrite
 We have seen that there is a strong dynamics in play around the use of

### 2019-08-13 paper_06_milestone/1st/py/make_figs_probes.py
More coreections and improved figures in exp chapter

@@ -33,6 +33,7 @@ save(fig, "fig_profiles_probe_averaged.png")
 
 signals.plot_energy_pot_vs_time()
 fig = plt.gcf()
+plt.gca().set_ylabel("normalized $E_P$")
 unset_title(fig)
 save(fig, "fig_energy_pot_vs_time.png")
 

### 2019-08-14 paper_03_toy_model/article.tex
Formatting equations

@@
 Unfortunately, we have found no elegant way in which they can be expressed in vector notation, why we write them as
-\begin{eqnarray} \label{PsiEq}
-\frac{\partial{\Psi_{ii}}}{\partial t} +\epsilon_{ijk}e_k \Psi_i \Psi_{jll} + \Psi_{ii} \chi_{jj} - \Psi_{ij} \chi_{ij} = -f \chi_{ii} + (-1)^{p+1} \nu_p \nabla^{2p} \Psi_{ii} \, , \\ \label{ChiEq}
-\frac{\partial{\chi_{ii}}}{\partial t} +\epsilon_{ijk}e_k (\Psi_i \chi_{jll} - \Psi_{jl} \chi_{il}) + \Psi_{ij} \Psi_{ij} - \Psi_{ii} \Psi_{jj} =  f \Psi_{ii} - c \theta_{ii} + (-1)^{p+1} \nu_p \nabla^{2p} \chi_{ii} \, , \\ \label{ThetaEq}
-\frac{\partial{\theta}}{\partial t} + \epsilon_{ijk}e_k \Psi_i \theta_{j} = - c \chi_{ii} + (-1)^{p+1} \nu_p \nabla^{2p} \theta \, ,
-\end{eqnarray}
+\begin{align} \label{PsiEq}
+&\frac{\partial{\Psi_{ii}}}{\partial t} +\epsilon_{ijk}e_k \Psi_i \Psi_{jll} + \Psi_{ii} \chi_{jj} - \Psi_{ij} \chi_{ij} = -f \chi_{ii} + (-1)^{p+1} \nu_p \nabla^{2p} \Psi_{ii} \, , \\ \label{ChiEq}
+&\frac{\partial{\chi_{ii}}}{\partial t} +\epsilon_{ijk}e_k (\Psi_i \chi_{jll} -
+\Psi_{jl} \chi_{il}) + \Psi_{ij} \Psi_{ij} - \Psi_{ii} \Psi_{jj} \nonumber\\
+&\quad=  f \Psi_{ii} - c \theta_{ii} + (-1)^{p+1} \nu_p \nabla^{2p} \chi_{ii} \, , \\ \label{ThetaEq}
+&\frac{\partial{\theta}}{\partial t} + \epsilon_{ijk}e_k \Psi_i \theta_{j} = - c \chi_{ii} + (-1)^{p+1} \nu_p \nabla^{2p} \theta \, ,
+\end{align}
 where $ e_k $ is the vertical unit vector expressed in three-dimensional Cartesian tensor notation,  subscript $ i $  has the same meaning as $ {\partial }/{\partial x_i }$ when used on $ \Psi $,  $ \chi $ or $ \theta $ and repeated indices means summation.

### 2019-08-14 paper_04_shallow_water/Latex/article.tex
Formatting equations
Proof read swe paper

@@
 % $ (\epsilon d)^{1/3} $, where $ \epsilon $ is the mean dissipation rate and $ d $
-% the mean distance between the shocks, and that the $ p $:th order displacement
+% the mean distance between the shocks, and that the $ p $th-order displacement
 % and velocity structure functions scale as $ (\epsilon d)^{p/3} r/d $, where $ r $
@@
 %
-The single layer SW equations are very quite to
+The single-layer SW equations are quite similar to
 the two-dimensional compressible Navier-Stokes equations.  The
@@
 one of the simplest hydrodynamic models with coexisting eigenmodes of
-the linearized operator with zero linear frequency and with finite
+the linearised operator with zero linear frequency and with finite
 non-zero linear frequency, $\omega = \pm\sqrt{f^2 + c^2
 k^2}$, where $f$ is the Coriolis parameter, $c$ the gravity wave speed and $k
-= |\kk|$ the wave number.
+= |\kk|$ the wavenumber.
 %
@@
 FargeSadourny1989}.   \cite{YuanHamilton1994} showed that
-statistically stationary shallow-water flows can be obtained by forcing
+statistically stationary shallow water flows can be obtained by forcing
 quasi-geostrophic modes at large scales and with dissipation only at
@@
 The $ k^{-5/3} $ spectrum has been reproduced
- by some General Circulation Models (GCM:s) based on the primitive
+ by some General Circulation Models (GCMs) based on the primitive
 equations \cite[]{KoshykHamilton2001, Skamarock2004,
-Hamilton_etal2008}.
-\cite{AugierLindborg2013}  formulated a spectral energy budget of kinetic and available potential energy in terms of spherical harmonics and used this formulation to investigated two GCM:s.
-It was found that one of the models -- named the AFES model -- was able to reproduce a downscale energy cascade with realistic $ k^{-5/3} $ mesoscale spectra.
-The result was almost too good, since the AFES model simulation only contained 21 vertical levels, while theoretical arguments based on stratified turbulence theory \cite[]{Lindborg2006} suggest that a much finer vertical resolution would be needed.
-This lead us to ask what number of vertical levels a model needs in order  to reproduce a downscale energy cascade.
-Was \cite{YuanHamilton1994}  right  when they suggested that even a single layer model, such as the SW equations, could do the job?
-The original motivation of this study was to investigate this suggestion. A first set of
+Hamilton_etal2008}.  \cite{AugierLindborg2013}  formulated a spectral energy
+budget of kinetic and available potential energy in terms of spherical
+harmonics and used this formulation to investigated two GCMs. It was found that
+one of the models -- named the Atmospheric GCM for the Earth Simulator (AFES)
+model -- was able to reproduce a downscale energy cascade with realistic $
+k^{-5/3} $ mesoscale spectra. The result was almost too good, since the AFES
+model simulation only contained 21 vertical levels, while theoretical arguments
+based on stratified turbulence theory \cite[]{Lindborg2006} suggest that a much
+finer vertical resolution would be needed.
+This leads us to ask what number of vertical levels a model needs in order  to reproduce a downscale energy cascade.
+Was \cite{YuanHamilton1994}  right  when they suggested that even a single-layer model, such as the SW equations, could do the job?
+The original motivation for this study was to investigate this suggestion. A first set of
 simulations of the SW equations, containing both balanced and gravity wave motions,
@@
 In some simulations they were densely packed and in others sparsely distributed.
-In what way  do the input parameters regulate these variations?
-Will there always be shocks or will they disappear in the limit of weak nonlinearities?
-Such questions made us to shift our theoretical focus and pursue the study, despite the fact that the answer to the question that originally motived it turned out to be negative.
+In what way  do the input parameters regulate these variations? Will there
+always be shocks or will they disappear in the limit of weak nonlinearities?
+Such questions made us to shift our theoretical focus and pursue this question, despite the fact that the answer to the question that originally motived it turned out to be negative.
 
@@
 solutions of the kinetic equation corresponding to a constant downscale energy
-flux is associated with a spectrum scaling as $E(k) \sim k^{-8/3}$, while
-\cite{ZakharovSagdeev1970} used weak turbulence formalism to derive a spectrum
+flux are associated with a spectrum scaling as $E(k) \sim k^{-8/3}$, while
+\cite{ZakharovSagdeev1970} used the weak turbulence formalism to derive a spectrum
 of the form $ E(k) \sim k^{-3/2} $ for three-dimensional acoustic turbulence. This prediction is
 of particular interest, since the acoustic equations and the SW equations are similar
-to the equations describing acoustic turbulence. It may be questioned, however, that weak turbulence formalism is
+to the equations describing acoustic turbulence. However, it is questionable that
+the weak turbulence formalism is
 applicable to acoustic and SW wave turbulence, no matter how weak the
 nonlinearities are supposed to be.
-\cite{KadomtsevPet1973} argued that shocks always will always develop in acoustic turbulence, also in the limit of weak nonlinearities, due to the fact that acoustic waves are nondispersive.
-
-Shocks have been been observed in many simulations of the SW equations \cite[]{FargeSadourny1989,
+\cite{KadomtsevPet1973} argued that shocks always will always develop in acoustic turbulence, also in the limit of weak nonlinearities, due to the fact that acoustic waves are non-dispersive.
+
+Shocks have been observed in many simulations of the SW equations \cite[]{FargeSadourny1989,
 PolvaniMcWilliamsSpallFord1994, LahayeZeitlin2012}.
@@
 shocks will scale as $ k^{-2} $, a prediction which has been confirmed by a
-great number of simulations of Burgers equation \cite[for a review, see ]
+great number of simulations of the Burgers equation \cite[for a review, see ]
 []{Frisch2001}. The shocks also lead to very strong intermittency, as
@@
 \cite{YuanHamilton1994} carried out
-forced dissipative simulations of the SW equations showing similar spectra.
+forced-dissipative simulations of the SW equations showing similar spectra.
 However, the resolution in both these studies was quite coarse, ($386^2$
@@
 $h = 1 + \eta $ the non-dimensional thickness of the fluid layer,  $ \eta $ being the surface displacement,
-$\nu$ the kinematic viscosity and $ c $ is the gravity wave speed.
-The SW equations conserve mass, $ h $, and momentum $ {\bf J} = h{\bf u} $.
-As a matter of fact, we have constructed the viscous term in (\ref{UEq}) in such a way that momentum is conserved and energy dissipation is positive definite.
-This can only be accomplished by permitting the term to be nonlinear in the flow variables.
-There are different forms of nonlinear terms that fulfil both conditions and it may be disputed which is the best one.
-However, in the simulations we will use diffusion terms in both (\ref{UEq}) and (\ref{HEq}) that will not fulfil any of these two conditions in a strict sense and we will therefore not discuss this matter further.
-The inviscid SW equations conserve a local
+$\nu$ the kinematic viscosity and $ c $ the gravity wave speed. The SW
+equations conserve mass, $ h $, and momentum $ {\bf J} = h{\bf u} $. As a
+matter of fact, we have constructed the viscous term in (\ref{UEq}) in such a
+way that momentum is conserved and energy dissipation is positive definite.
+This can only be accomplished by permitting the term to be nonlinear in the
+flow variables. There are different forms of nonlinear terms that fulfil both
+conditions and it may be disputed which is the best one. However, in the
+simulations we will use diffusion terms in both (\ref{UEq}) and (\ref{HEq})
+that will not fulfil either of these two conditions in a strict sense and we
+will therefore not discuss this matter further. The inviscid SW equations
+conserve a local
 quantity along trajectories of the fluid particles, the Ertel
@@
 \end{equation}
-with equipartition between KE and APE over a wave period in each Fourier mode.
-
-In order to analyze the spectral flux of energy, we derive the spectral energy
+with equipartition between kinetic energy (KE) and APE over a wave period in each Fourier mode.
+
+In order to analyse the spectral flux of energy, we derive the spectral energy
 budget, i.e.\ the governing equations for spectral KE and APE
@@
 \end{equation}
-and $\Re$ denotes the real part, $\kk$ is the wave number and the
+and $\Re$ denotes the real part, $\kk$ is the wavenumber and the
 hat denotes the Fourier transform.
@@
 \end{equation}
-Similarly, the APE can be written as the sum over all wave numbers of
+Similarly, the APE can be written as the sum over all wavenumbers of
 the spectral APE function
@@
 
-We will now derive an analogue of the so called `four-fifths  law' \cite[]{Kolmogorov1941} for the the third order velocity structure function of incompressible isotropic turbulence.
-In the case of irrotational SW wave turbulence the analogous law contains a third order structure function involving both velocity and displacement increments.
+We will now derive an analogue of the so called `four-fifths  law' \cite[]{Kolmogorov1941} for the the third-order velocity structure function of incompressible isotropic turbulence.
+In the case of irrotational SW wave turbulence the analogous law contains a third-order structure function involving both velocity and displacement increments.
 In order for such a law to exist it is necessary that the velocity field is either non-divergent (incompressible) or irrotational.
@@
 which gives
-\begin{eqnarray}
-\p_t ( \JJ'\cdot\uu + \JJ\cdot\uu' + 2c^2 h'h )
-&=& - \p_i( J_i'|\uu|^2/2) - \p_i'( J_i|\uu'|^2/2) %\nonumber \\&& 
-- \p_i'(u_j J_j' u_i') - \p_i(u_j' J_j u_i) \nonumber\\
-&& -c^2\p_i (J_i'h) - c^2\p_i' (J_ih') %\nonumber\\&& 
-- \p_i' (u_i E_p') - \p_i (u_i' E_p) \nonumber\\
-&& -2c^2\p_i' (J_i'h) - 2c^2\p_i (J_ih').
-\end{eqnarray}
+\begin{flalign}
+&\p_t ( \JJ'\cdot\uu + \JJ\cdot\uu' + 2c^2 h'h ) \nonumber\\
+&= - \p_i( J_i'|\uu|^2/2) - \p_i'( J_i|\uu'|^2/2) %\nonumber \\&&
+   - \p_i'(u_j J_j' u_i') - \p_i(u_j' J_j u_i) %\nonumber\\
+   -c^2\p_i (J_i'h) \nonumber\\
+&\quad- c^2\p_i' (J_ih') - \p_i' (u_i E_p') - \p_i (u_i' E_p) %\nonumber\\
+ -2c^2\p_i' (J_i'h) - 2c^2\p_i (J_ih').
+\end{flalign}
 We then assume homogeneity and take the average denoted by
@@
 \end{equation}
-For separations that are considerably larger than dissipative scales of motion and considerably smaller than forcing scales a standard argument \cite[see for example][]{Frisch} gives that the left hand side is approximately equal to $ -8 \epsilon $, where $ \epsilon $ is the mean dissipation rate.
+For separations that are considerably larger than dissipative scales of motion and considerably smaller than forcing scales a standard argument \cite[see for example][]{Frisch} gives that the left-hand side is approximately equal to $ -8 \epsilon $, where $ \epsilon $ is the mean dissipation rate.
 It is interesting to note that a factor of $ 8 $ appears in the case when the velocity field is irrotational while a corresponding factor of $ 4 $ appears in the case when the velocity field is non-divergent.
@@
 A structure function of a flow variable is defined as the area average of increments  between two points separated by a distance $ r $.
-Assuming that the contribution from increments where the two points are not separated by a shock is negligible for higher order structure functions we obtain
+Assuming that the contribution from increments where the two points are not separated by a shock is negligible for higher-order structure functions we obtain
 \begin{eqnarray}
@@
 giving the numerical values $R_2 = 2$, $R_3 = 6\pi/8$ and $R_4 = 8/3$.
-Interestingly, for an isotropic and irrotational velocity field, the longitudinal and transverse second order structure functions
+Interestingly, for an isotropic and irrotational velocity field, the longitudinal and transverse second-order structure functions
 are exactly related by $\meane{(\delta u_L)^2} = \p_r( r\meane{(\delta
-u_T)^2} )$ \cite[]{Lindborg2007jas}. If we further assume than both second order structure functions follow
+u_T)^2} )$ \cite[]{Lindborg2007jas}. If we further assume than both second-order structure functions follow
 the same scaling law $\meane{ (\delta u_L)^2} \propto \meane{ (\delta
@@
 
-In order for (\ref{StepH}-\ref{StepT}) to match with the third order structure function law (\ref{eq_Kolmo})  the shock amplitudes should scale as
+In order for (\ref{StepH}-\ref{StepT}) to match with the third-order structure function law (\ref{eq_Kolmo})  the shock amplitudes should scale as
 \begin{equation} \label{Strength}
@@
 \end{equation}
-Assuming that (\ref{StrucFunc}) is valid for second order moments and that the non-quadratic contribution to the KE spectrum is negligible the corresponding scaling law for the energy spectra reads
+Assuming that (\ref{StrucFunc}) is valid for second-order moments and that the non-quadratic contribution to the KE spectrum is negligible the corresponding scaling law for the energy spectra reads
 \begin{equation} \label{Spectra}
@@
 However, exact solutions of Burgers equation \cite[see for example][]{Whitham}
-suggests that  $ \delta x \propto \nu $ rather than $ \delta x \propto \nu ^{3/4} $.
+suggest that  $ \delta x \propto \nu $ rather than $ \delta x \propto \nu ^{3/4} $.
 The scaling of the energy spectra (\ref{Spectra}) also suggests that $ d $ should enter into the expression for $ \delta x$, apart from $ \nu $ and $ \epsilon $.
@@
 For Kolmogorov turbulence observations indicate that $ F_{\partial_x u} \propto \nu^{-1/6} $  \cite[]{VanAtta_Antonia1980}, a dependence which is well capture by the multifractal model of
-\cite{Meneveau_Streenivasan1991}.
-In SW wave turbulence there is a much stronger increase of  the flatness factor with decreasing  $ \nu $, which can be explained the fact that dissipation takes place only at the shocks.
+\cite{Meneveau_Streenivasan1991}.  In SW wave turbulence there is a much
+stronger increase of  the flatness factor with decreasing  $ \nu $, which can
+be explained by the fact that dissipation takes place only at the shocks.
 
@@
 \end{equation}
-where $ h_- $ and $ h_+ $ are the nondimensional layer thickness before and after the shock, respectively.
+where $ h_- $ and $ h_+ $ are the non-dimensional layer thickness before and after the shock, respectively.
 Noting that $ h_+ - h_- =  \;  \Delta h  $, expanding in small $ \Delta h  $ and using (\ref{Strength}) the Mach number can be estimated as
@@
 First, it is obvious that  the linear scaling of the structure functions only can be valid for separations  that fulfil the condition $ \delta x \ll r \ll d $.
-Second, it is obvious that it  cannot be valid for low order moments or structure functions.
+Second, it is obvious that it  cannot be valid for low-order moments or structure functions.
 According to arguments given by  \cite{BouchaudMezardParisi1995}  and \cite{Weinan}, it should be valid for structure functions of all orders $ p  \ge 1 $, in the asymptotic limit of large scale separation.
-It is quite easily argued, however, that the shocks can not determine the scaling of the structure functions of order $ p=1 $.
-To see this, consider the first order moment of the absolute value of $ \delta u_L $. 
+It is quite easily argued, however, that the shocks cannot determine the scaling of the structure functions of order $ p=1 $.
+To see this, consider the first-order moment of the absolute value of $ \delta u_L $.
 The longitudinal velocity increment is always negative over a shock, that is to say that the difference of the projection of the velocity on the direction over which the shock is crossed is always negative.
@@
 Thus,
-$ \means{ |\delta u_L | } < \langle | \delta u_L | \rangle _{ns} $, where the left hand side is the mean value over all separations where a shock is crossed and the right hand side is the mean value over separations where no shock is crossed.
-Therefore, the analysis is not valid for first order moments.
-In the following, we will compare the predictions with structure functions of second and higher orders calculated from simulations.
+$ \means{ |\delta u_L | } < \langle | \delta u_L | \rangle _{ns} $, where the left-hand side is the mean value over all separations where a shock is crossed and the right-hand side is the mean value over separations where no shock is crossed.
+Therefore, the analysis is not valid for first-order moments.
+In the following, we will compare the predictions with structure functions of second and higher-orders calculated from simulations.
 
@@
 standard open-source pseudo-spectral code FluidSim \cite[]{fluiddyn, fluidfft,
-fluidsim}. In these twenty simulations we used an eight order diffusion operator to act as the agent of dissipation. 
+fluidsim}. In these twenty simulations we used an eighth-order diffusion operator to act as the agent of dissipation.
  On the request from the reviewers we have also carried out a set of nineteen simulations with  Laplacian diffusion.
@@
 conditions and were run till a statistically stationary state was reached. The
-equations were reformulated in terms of the three eigenmodes of the linearized
+equations were reformulated in terms of the three eigenmodes of the linearised
 equations in Fourier space. One of these eigenmodes is zero for an irrotational
@@
 dissipation. Instead of using the nonlinear form of the viscous term in
-(\ref{eq_uu}) we have used a higher order linear operator $ \nu_8 \nabla^{8} $
+(\ref{eq_uu}) we have used a higher-order linear operator $ \nu_8 \nabla^{8} $
 both in (\ref{eq_uu}) and (\ref{eq_h}), where $ \nu_8 $ is a `hyperdiffusion'
@@
 The addition of a hyperdiffusion term in (\ref{eq_h}) will provide such a sink without seriously affecting mass conservation, which will be
-extremely well fulfilled, even when a high order hyperdiffusion term is added.
+extremely well fulfilled, even when a high-order hyperdiffusion term is added.
 
@@
 and very large $c$ when the shortest waves are very fast. We use an adaptable
-time step method which maximizes the time step over a standard
+time step method which maximises the time step over a standard
 Courant-Friedrichs-Lewy condition \cite{Lundbladh1999, AugierChomazBillant2012}.
@@
 Most of the aliasing is removed by truncating 8/9 of the modes along each
-direction \cite[for a detail discussion on the issues of the non-conservation
+direction \cite[for a detailed discussion on the issues of the non-conservation
 of the non-quadratic energy and the aliasing errors in the truncated one-layer
@@
 to 700. The wave modes are forced in a shell in spectral space corresponding to
-relatively small wave numbers, $ 5 \delta k < |\kk| < 8 \delta k$, where $
+relatively small wavenumbers, $ 5 \delta k < |\kk| < 8 \delta k$, where $
 \delta k = 2 \pi /L $ and $ L $ is the length of the side of the quadratic
-domain.  We define the forcing wave number as $ k_f = 6 \delta k $. To obtain as broad scaling range as possible it is preferable to choose 
+domain.  We define the forcing wavenumber as $ k_f = 6 \delta k $. To obtain as broad scaling range as possible it is preferable to choose
 the forcing scale, $ L_f = \pi/k_f $, as large as possible.
@@
 The forcing is white noise in time and designed so that the energy
-injection rate is normalised to unity at each time step. When the flow has reached a stationary state dissipation should therefore be equal to unity. 
+injection rate is normalised to unity at each time step.
+%When the flow has reached a stationary state dissipation should therefore be equal to unity.
+Therefore, when the flow has reached a stationary state dissipation should be equal to unity.
 The strength of the stratification is characterised by a
@@
 \end{equation}
-a parameter that can be determined from the start of each simulations.
+a parameter that can be determined from the start of each simulation.
 Energy is dissipated at the smallest resolved scales by hyperviscosity.
@@
 $  k_{d} \simeq \kmax/ 2.5$, where $ k_{d} = 1/\eta_8 $ and $\kmax =
-(8/9)\pi n/L $ is the maximum wave number, is  sufficient to resolve the shocks.
+(8/9)\pi n/L $ is the maximum wavenumber, is  sufficient to resolve the shocks.
 Since the energy injection rate is normalised to unity and all simulations reach a stationary state in which dissipation is equal to the injection rate, $ \nu_8 $ can be chosen already from the start so that this condition will be fulfilled in the stationary state.
@@
 \toprule
-{} &   $n$ &  $c$ &  $\nu_8$ &  $\epsilon$ &  $\frac{k_{diss}}{k_f}$ &   $F_f$ &  $t_{\max}$ \\
+{} &   $n$ &  $c$ &  $\nu_8$ &  $\epsilon$ &  $\frac{k_d}{k_f}$ &   $F_f$ &  $t_{\max}$ \\
 \midrule
-W1  &   960 &   10 & 1.56e-10 &        1.04 &                    28.8 &   0.16 &          50.0 \\
-W2  &  1920 &   10 & 9.68e-13 &        1.06 &                    57.7 &   0.16 &        49.9 \\
-W3  &  3840 &   10 &    6e-15 &         1.10 &                     115 &   0.17 &          50.0 \\
-W4  &  7680 &   10 & 3.72e-17 &         1.20 &                     231 &   0.17 &        25.2 \\
-W5  &   960 &   20 & 1.56e-10 &           1.00 &                    28.8 &  0.08 &          50.0 \\
-W6  &  1920 &   20 & 9.69e-13 &        1.02 &                    57.7 &  0.081 &         120 \\
-W7  &  3840 &   20 & 6.01e-15 &        1.08 &                     115 &  0.082 &         120 \\
-W8  &  7680 &   20 & 3.72e-17 &        1.21 &                     231 &  0.086 &         120 \\
-W9  &   960 &   40 & 1.56e-10 &       0.996 &                    28.8 &  0.040 &        49.9 \\
-W10 &  1920 &   40 & 9.68e-13 &       0.998 &                    57.7 &  0.040 &        49.9 \\
-W11 &  3840 &   40 &    6e-15 &        1.06 &                     115 &  0.041 &        49.9 \\
-W12 &   960 &  100 & 1.56e-10 &       0.999 &                    28.8 &  0.016 &        49.9 \\
-W13 &  1920 &  100 & 9.69e-13 &        1.01 &                    57.7 &  0.016 &         120 \\
-W14 &  3840 &  100 & 6.01e-15 &        1.06 &                     115 &  0.016 &         120 \\
-W15 &  7680 &  100 & 3.72e-17 &        1.22 &                     231 &  0.017 &        95.9 \\
-W16 &   960 &  400 & 1.56e-10 &        1.14 &                    28.8 &  0.0042 &         120 \\
-W17 &  1920 &  400 & 9.69e-13 &        1.15 &                    57.7 & 0.0042 &         120 \\
-W18 &  3840 &  400 & 6.01e-15 &        1.12 &                     115 & 0.0042 &         120 \\
-W19 &   960 &  700 & 1.56e-10 &        1.27 &                    28.8 & 0.0025 &        49.9 \\
-W20 &  1920 &  700 & 9.68e-13 &        1.05 &                    57.7 & 0.0023 &        49.9 \\
+W1  &   960 &   10 & \expn{1.56}{-10} &        1.04 &                    28.8 &   0.16 &          50.0 \\
+W2  &  1920 &   10 & \expn{9.68}{-13} &        1.06 &                    57.7 &   0.16 &        49.9 \\
+W3  &  3840 &   10 & \expn{6.00}{-15} &         1.10 &                     115 &   0.17 &          50.0 \\
+W4  &  7680 &   10 & \expn{3.72}{-17} &         1.20 &                     231 &   0.17 &        25.2 \\
+W5  &   960 &   20 & \expn{1.56}{-10} &           1.00 &                    28.8 &  0.08 &          50.0 \\
+W6  &  1920 &   20 & \expn{9.69}{-13} &        1.02 &                    57.7 &  0.081 &         120 \\
+W7  &  3840 &   20 & \expn{6.01}{-15} &        1.08 &                     115 &  0.082 &         120 \\
+W8  &  7680 &   20 & \expn{3.72}{-17} &        1.21 &                     231 &  0.086 &         120 \\
+W9  &   960 &   40 & \expn{1.56}{-10} &       0.996 &                    28.8 &  0.040 &        49.9 \\
+W10 &  1920 &   40 & \expn{9.68}{-13} &       0.998 &                    57.7 &  0.040 &        49.9 \\
+W11 &  3840 &   40 & \expn{6.00}{-15} &        1.06 &                     115 &  0.041 &        49.9 \\
+W12 &   960 &  100 & \expn{1.56}{-10} &       0.999 &                    28.8 &  0.016 &        49.9 \\
+W13 &  1920 &  100 & \expn{9.69}{-13} &        1.01 &                    57.7 &  0.016 &         120 \\
+W14 &  3840 &  100 & \expn{6.01}{-15} &        1.06 &                     115 &  0.016 &         120 \\
+W15 &  7680 &  100 & \expn{3.72}{-17} &        1.22 &                     231 &  0.017 &        95.9 \\
+W16 &   960 &  400 & \expn{1.56}{-10} &        1.14 &                    28.8 &  0.0042 &         120 \\
+W17 &  1920 &  400 & \expn{9.69}{-13} &        1.15 &                    57.7 & 0.0042 &         120 \\
+W18 &  3840 &  400 & \expn{6.01}{-15} &        1.12 &                     115 & 0.0042 &         120 \\
+W19 &   960 &  700 & \expn{1.56}{-10} &        1.27 &                    28.8 & 0.0025 &        49.9 \\
+W20 &  1920 &  700 & \expn{9.68}{-13} &        1.05 &                    57.7 & 0.0023 &        49.9 \\
 \bottomrule
@@
 averaged mean energy dissipation in the stationary state, $ k_{d}/ k_f $: ratio
-between dissipation wave number and forcing wave number, $ F_f $: Froude
+between dissipation wavenumber and forcing wavenumber, $ F_f $: Froude
 number, $ t_{max} $: end time of simulation.}
@@
 The time evolution of mean energy is shown in figure~\ref{fig_Evstime} for all
-twenty simulations. After an initial period of growth, mean energy is leveling
+twenty simulations. After an initial period of growth, mean energy is levelling
 out at different values for different runs. This is not the same picture as we
 see in simulations of three-dimensional incompressible turbulence or stratified
-turbulence, where all curves would have leveled out more or less at the same
+turbulence, where all curves would have levelled out more or less at the same
 value. We find that mean energy in the stationary state is a function of both $
@@
  In figure~\ref{MeanE} to the
-left, we see $ E/\sqrt{\epsilon} $ as function of $ c $ for different resolutions $
+left, we see $ E/\sqrt{\epsilon} $ as a function of $ c $ for different resolutions $
 n $. For each resolution the points are approaching a power law $ C_n c^{1/2} $
@@
 be interpreted as consequence of the numerics but rather as a Reynolds number
-effect, although it is not hundred percent clear how a Reynolds number should
+effect, although it is not hundred per cent clear how a Reynolds number should
 be defined in our case. We may note, however, that $ n \sim L_f/\delta x $ and in the case of Navier-Stokes viscosity
- we should have $ \delta x \propto \nu $, according to (\ref{Width}). Therefore,  $ n $ may be substituted by a Reynolds number. The normalised energy dissipation on the left hand side of
+ we should have $ \delta x \propto \nu $, according to (\ref{Width}). Therefore,  $ n $ may be substituted by a Reynolds number. The normalised energy dissipation on the left-hand side of
 (\ref{Dissipation}) will thus go to zero in the limit of large Reynolds number.
 This is very different from three-dimensional turbulence where there is a local
-Richardson-Kolmogorov cascade, in which energy is successively transfered from
+Richardson-Kolmogorov cascade, in which energy is successively transferred from
 large to small scales of motion. That the cascade is local means that eddies of
@@
 paradigm of a local Richardson-Kolmogorov cascade. Moreover, it is not only in
-the limit of large Reynolds number the normalised dissipation will go to zero
+the limit of large Reynolds numbers the normalised dissipation will go to zero
 but also in the limit of strong stratification, $ E^{1/2}/c \rightarrow 0 $.
 This is different from three-dimensional stratified turbulence where the ratio
-on the left hand side of (\ref{Dissipation}) is of the order of unity in the
+on the left-hand side of (\ref{Dissipation}) is of the order of unity in the
 limit of strong stratification \cite[]{Lindborg2006, Brethouwer2007}.
@@
 wind field and sucked out from the same mode in `intermittent patches of foaming'
-or `white horses'.  This may be a too simplified picture of SW water turbulence, since a broad band energy spectrum develops between the forcing and the dissipation wave numbers. 
-Nevertheless,  shock formation suggests that there is a considerable direct energy transfer from large to very small wave lengths, without any intermediate steps.
-
-The `four-fifths law' for the third order structure function of Kolmogorov
+or `white horses'.  This may be a too simplified picture of SW water turbulence, since a broad band energy spectrum develops between the forcing and the dissipation wavenumbers.
+Nevertheless,  shock formation suggests that there is a considerable direct energy transfer from large to very small wavelengths, without any intermediate steps.
+
+The `four-fifths law' for the third-order structure function of Kolmogorov
 turbulence is often derived using the assumption of a local cascade
@@
 figure~\ref{Flux} we see the spectral energy flux from run W7 to the left and
-the third order structure function to the right. There is a broad constant flux
+the third-order structure function to the right. There is a broad constant flux
 range with equipartition between KE and APE flux and an almost equally broad
 range where (\ref{eq_Kolmo}) is satisfied, with equipartition between the two
-third order structure functions associated with KE and APE. The assumptions of
+third-order structure functions associated with KE and APE. The assumptions of
 a local cascade and finite dissipation do not seem to be necessary in order for
-the third order structure function law (\ref{eq_Kolmo}) to be valid.
+the third-order structure function law (\ref{eq_Kolmo}) to be valid.
 
@@
 \caption{Left: Time averaged normalised spectral energy fluxes versus $ k/k_f
-$. Right: Time averaged normalised third order structure functions versus $
+$. Right: Time averaged normalised third-order structure functions versus $
 r/L_f $. From run W7. }
@@
 \centerline{\includegraphics[]{paper_04_shallow_water/Pyfig/fig5}}
-\caption{Left: divergence $ \bnabla \cdot {\bf u} $. Right: velocity component in the $ y $-direction: $ u_y $. Top: run W6. Bottom: run W17.  }
+\caption{$(a,c)$ Divergence $ \bnabla \cdot {\bf u} $.  $(b,d)$ Velocity
+component in the $ y $-direction: $ u_y $. $(a,b)$ Run W6. $(c, d)$ Run W17.
+}
 \label{Physical}
@@
 
-The equipartition of KE and APE spectral fluxes suggests that the flow to
+The equipartition of the KE and APE spectral fluxes suggests that the flow to
 leading order may be described as a collection of linear gravity waves with energy
 equipartition in each mode. To investigate this further we have
-collected time series of the flow variables from Fourier modes whose magnitude
-fall within certain shells, $ \mid {\bf k} \mid \in [ k -\delta k/2, \; k+
+collected time series of the flow variables from Fourier modes whose magnitudes
+fall within certain shells, $ |{\bf k}| \in [ k -\delta k/2, \; k+
 \delta k/2] $, and computed temporal power law spectra from these time series
-and averaged over all wave numbers in each shell. In other words, we have
+and averaged over all wavenumbers in each shell. In other words, we have
 computed KE and APE frequency spectra corresponding to different magnitudes of
 modes. Figure~\ref{fig_spatiotemp_spectra} presents such spectra plotted as
-functions of the normalised frequency $\omega/\omega_l$, where $ \omega_l = kc
+functions of the normalised frequency $\omega/\omega_l$, where $ \omega_l = ck
 $ is the linear wave frequency. The spectra are strongly dominated by peaks at
@@
 the divergence, $ \nabla \cdot {\bf u} $, to the left and the velocity
-component in the $ y $-direction, $ u_y $, to the right.
-Note that total area of the computational domain is sixteen times larger than the area of the outcuts shown in the figures, since the side length of the figures is $ 3 L_f $, while the side length of the computational domain is $ 12 L_f $.
+component in the $ y $-direction, $ u_y $, to the right.  Note that the total
+area of the computational domain is sixteen times larger than the area of the
+outcuts shown in the figures, since the side of the figures is $ 3 L_f
+$, while the side of the computational domain is $ 12 L_f $.
 The two figures at the
@@
 \includegraphics[]{paper_04_shallow_water/Pyfig/fig7}}
-\caption{From top to bottom: second, fourth and sixth order compensated and
-normalised longitudinal structure functions. Left: runs with $ n=3840 $: W3,
-W7, W11, W14 and W18. Right: runs with $ n = 7860 $: W4, W8 and W15. }
+\caption{From top to bottom: second-, fourth- and sixth-order compensated and
+    normalised longitudinal structure functions. $(a,c,e)$ Runs with $ n=3840 $: W3,
+W7, W11, W14 and W18. $(b,d,f)$ Runs with $ n = 7860 $: W4, W8 and W15. }
 \label{fig_StrucFunc}
@@
 Right: Constant $ n= 1920 $ and varying $ c $, runs W2, W6, W10, W13, W17 and W20.
-The vertical lines indicate the wave numbers of the maxima.}
+The vertical lines indicate the wavenumbers of the maxima.}
 \label{fig_disspectra}
@@
 velocity structure functions. In figure~\ref{fig_StrucFunc} we see the
-compensated and normalised second, fourth and sixth order longitudinal
+compensated and normalised second-, fourth- and sixth-order longitudinal
 structure functions from the five simulations with $ n = 3840 $ to the left and
@@
 lower resolution simulations to the left. There is clearly a better collapse of
-the curves for the sixth order structure functions compared to the second order
+the curves for the sixth-order structure functions compared to the second-order
 structure functions. This is not surprising, since the shock contribution to
@@
 close to but somewhat steeper than $ r^{-1} $. At small $ r $ the curves are
-leveling out at very large values. For $ n = 7869 $ and $ c = 10 $ we see that
+levelling out at very large values. For $ n = 7869 $ and $ c = 10 $ we see that
 $ F_{T} \approx 400 $ at small $ r $. The degree to which the flatness factor
@@
 normalised energy spectra from the three highest resolution runs. The
-compensated spectra are almost flat in a broad range. In the high wave number
+compensated spectra are almost flat in a broad range. At the high wavenumber
 end the compensated spectra show a `bottleneck', which we interpret as a
 signature of the shocks. The $ k^{-2} $-dependence and the associated scaling predicted
-by (\ref{Spectra}) are very well satisfied in a limited range of high wave
-numbers close to the bottleneck. At smaller wave numbers the spectra are
+by (\ref{Spectra}) are very well satisfied in a limited range of high wavenumbers close to the bottleneck. At smaller wavenumbers the spectra are
 becoming somewhat steeper than $ k^{-2} $ and the spectrum from the run with
@@
 First we note that the spectra go to zero quite nicely for large $ k $, indicating that the shock width $ \delta x $ is well resolved.
-Second, we note that there is a slight shift of the maximum in both figures, even though we have normalised the wave number by $ k_d = 1/\eta_8 $.
+Second, we note that there is a slight shift of the maximum in both figures, even though we have normalised the wavenumber by $ k_d = 1/\eta_8 $.
 These shifts may be explained by the estimate (\ref{EtaEight}) predicting that $ \delta x $ should vary slightly with $ d $.
 If the maximum of the dissipation spectra is associated with $ \delta x $, there should be a shift of the maximum  to the right with increasing $ n $ and to the left with increasing $ c $, since $ d \propto c^{-1/2} $.
- The ratio between the maximum and the minimum $ n $ in the left figure is equal to 8 and  according to (\ref{EtaEight}) the corresponding ratio between the wave numbers at which the maxima are observed should be $ 8^{1/21} \approx 1.1 $, which is exactly the value we observe.
+ The ratio between the maximum and the minimum $ n $ in the left figure is equal to 8 and  according to (\ref{EtaEight}) the corresponding ratio between the wavenumbers at which the maxima are observed should be $ 8^{1/21} \approx 1.1 $, which is exactly the value we observe.
 The ratio between the maximum and minimum $ c $ in the right figure is equal to $ 70 $.
-Assuming that $ d \propto c^{-1/2} $ the ratio between the wave numbers at which the maxima are observed should be $ 70^{1/42} \approx 1.1 $, while the observed ratio is equal to $ 1.16 $.
+Assuming that $ d \propto c^{-1/2} $ the ratio between the wavenumbers at which the maxima are observed should be $ 70^{1/42} \approx 1.1 $, while the observed ratio is equal to $ 1.16 $.
 In the latter case, there is no exact quantitative agreement between the predicted trend and the numerical results.
@@
 shock model predicting that the shock amplitude scales as $ (\epsilon d)^{1/3}
-$, where $ d $ is the mean distance between the shocks, and that the $ p $:th
-order structure function above a certain minimum will scale as $ (\epsilon
-d)^{p/3} r/d $. Then, we carried out a series of  forced dissipative
+$, where $ d $ is the mean distance between the shocks, and that the
+$ p $th-order structure function above a certain minimum will scale as
+$ (\epsilon d)^{p/3} r/d $. Then, we carried out a series of forced-dissipative
 simulations, varying the Froude number and the resolution. In all simulations,
@@
 in accordance with linear wave dynamics, with equipartition between APE and KE
-over a period. The third order structure function relation was fulfilled with a
+over a period. The third-order structure function relation was fulfilled with a
 high degree of accuracy. From the simulations we made two interesting
@@
 (\ref{Dissipation}) will go to zero both in the limit of small viscosity and
-large wave speed. A similar observation was made when we used a second order diffusion operator instead of hyper diffusion (see Appendix A).
+large wave speed. A similar observation was made when we used a second-order diffusion operator instead of hyper diffusion (see Appendix A).
 This is different from three-dimensional Kolmogorov
@@
 Richardson-Kolmogorov cascade.  It remains a theoretical challenge to give a quantitative explanation of the growth of energy with increased Reynolds number.
-As far as wee can see, such an explanation would require a completely novel theoretical development.
+As far as we can see, such an explanation would require a completely novel theoretical development.
 The second observation we made was that the mean
@@
 dynamics in the limit of weak nonlinearities, $ F_{f} \rightarrow 0 $. The
-picture may become more vivid if we say `wave breaking' were we previously have
-talked about shock formation, and also say `wave lengths' were we previously
-have talked about length scales. We identify the longest wave length, $
-\lambda_l $, with the forcing scale $ L_f $ and the breaking wave length, $
+picture may become more vivid if we say `wave breaking' where we previously have
+talked about shock formation, and also say `wavelengths' were we previously
+have talked about length scales. We identify the longest wavelength, $
+\lambda_l $, with the forcing scale $ L_f $ and the breaking wavelength, $
 \lambda_b $, with the mean distance, $ d $, between the shocks. There are two
@@
 Reynolds number. Thus, for small $ F_f $ and large Reynolds
-number the longest waves are not directly influenced by wave breaking. The
+numbers the longest waves are not directly influenced by wave breaking. The
 Reynolds number dependence of the amplitude ratio makes it unlikely that a
@@
  been derived  for a compressible and barotropic fluid by \cite{Falkovich2010} and \cite{Galtier2011}.
-Unlike (\ref{FluxT}) these flux relations are not exclusively expressed in terms of third order structure functions, {\em i.
-e.} third order moments of increments of flow variables.
+Unlike (\ref{FluxT}) these flux relations are not exclusively expressed in terms of third-order structure functions, {\em i.
+e.} third-order moments of increments of flow variables.
 As is clear from our analysis, in the general case of a compressible and irrotational fluid the KE flux can be expressed in this way.
-It seems to be more of a specific property that the APE flux also can be expressed in terms of a third order structure function, a property that the SW system shares with the Boussinesq system with constant Brunt-V\"ais\"al\"a frequency, in which case APE also is quadratic in buoyancy \cite[]{Augier2012}.
-In this context,  a comparison between our analysis and the analysis  by \cite{Galtier2011} of  isothermal compressible turbulence  is revealing.
-In the latter case, the pressure is linear in density, $ p = c_t^2 \rho $, where $ c_t $ is the isothermal speed of sound, and the internal energy per unit volume is  equal to $ c_t^2 \rho \ln (\rho/ \rho_0) $, where $ \rho_0 $ is an ambient reference density.
-Due to this  logarithmic dependence, the internal energy flux is not expressible in terms of a structure function.
-However, an equation system which is equivalent to the SW equations can be obtained by replacing the pressure term, $ \rho^{-1} \nabla p = \rho^{-1} c_t^{2} \nabla \rho $, in the isothermal momentum equation, by $ \rho_{0}^{-1} c_t^{2} \nabla \rho $, which may be motivated if density fluctuations are small.
-The expression for the internal energy per unit volume of the modified system is  $ c_t^2 \rho^2/2\rho_{0} $, which is equivalent to the SW potential energy.
-\footnote{To see how the logarithmic and quadratic expressions for the internal energy are
- related, write $ \rho = \rho_0(1+\eta) $, expand in $ \eta $ and keep  terms up to second order: 
+It seems to be more of a specific property that the APE flux also can be expressed in terms of a third-order structure function, a property that the SW system shares with the Boussinesq system with constant Brunt-V\"ais\"al\"a frequency, in which case APE also is quadratic in buoyancy \cite[]{Augier2012}.
+In this context,  a comparison between our analysis and the analysis  by
+\cite{Galtier2011} of  isothermal compressible turbulence  is revealing. In the
+latter case, the pressure is linear in density, $ p = c_t^2 \rho $, where $ c_t
+$ is the isothermal speed of sound, and the internal energy per unit volume is
+equal to $ c_t^2 \rho \ln (\rho/ \rho_0) $, where $ \rho_0 $ is an ambient
+reference density. Due to this  logarithmic dependence, the internal energy
+flux is not expressible in terms of a structure function. However, an equation
+system which is equivalent to the SW equations can be obtained by replacing the
+pressure term, $ \rho^{-1} \nabla p = \rho^{-1} c_t^{2} \nabla \rho $, in the
+isothermal momentum equation, by $ \rho_{0}^{-1} c_t^{2} \nabla \rho $, which
+may be motivated if density fluctuations are small. The expression for the
+internal energy per unit volume of the modified system is  $ c_t^2
+\rho^2/2\rho_{0} $, which is equivalent to the SW potential energy. (To see how the logarithmic and quadratic expressions for the internal energy are
+ related, write $ \rho = \rho_0(1+\eta) $, expand in $ \eta $ and keep  terms up to second-order:
  $ c_t^2\rho \ln(\rho/\rho_0) \approx c_t^2 \rho_0(1+\eta)(\eta - \eta^2/2) \approx   c_t^2 \rho^2/2 \rho_0 - c_t^2 \rho_0/2 $.
-For small density fluctuations the two expressions only differ by a constant.} As a consequence of this quadratic dependence, the internal energy flux is expressible in terms of a structure function.
+ For small density fluctuations the two expressions only differ by a constant.) As a consequence of this quadratic dependence, the internal energy flux is expressible in terms of a structure function.
 It may be conjectured that the modified equation system is a good model for isothermal acoustic turbulence and that a flux relation which  equivalent to (\ref{FluxT}) also should hold for this case.
@@
 If  the Prandtl number is of the order of unity, $ \chi $ is of the same order as $ \epsilon $.
-The shock amplitude scales as $ \Delta u \sim (\epsilon + \chi)^{1/3} d^{1/3} $, where $ d $ is the characteristic distance between the shocks and the third order longitudinal structure function scales as $ \langle \delta u_L^{3} \rangle = -C (\epsilon + \chi)r $, where $ C $ is a constant of the order of unity, which can be determined exactly in the weak shock limit.
+The shock amplitude scales as $ \Delta u \sim (\epsilon + \chi)^{1/3} d^{1/3} $, where $ d $ is the characteristic distance between the shocks and the third-order longitudinal structure function scales as $ \langle \delta u_L^{3} \rangle = -C (\epsilon + \chi)r $, where $ C $ is a constant of the order of unity, which can be determined exactly in the weak shock limit.
 
@@
 %\label{app_comp}
-On the request from the reviewers we have carried out a set of simulations using Laplacian diffusion, $ \nu \nabla^{2} $, to check that similar results are obtained in this case as when we use higher order diffusion $ \nu_8 \nabla^{8} $.
-A number of test simulations with resolution $ n = 960 $ and $ c= 10 $ was first carried out using different values of the viscosity, till we determined a value which was sufficiently large for the shocks to be resolved, but not unnecessarily large.
-This value was then used for all simulations with $ n = 960 $.
-For the higher resolved simulations the viscosity was then set as $ \nu \propto n^{-1} $, in accordance with the prediction  (\ref{Width}) for the shock width.
+On the request from the reviewers we have carried out a set of simulations using Laplacian diffusion, $ \nu \nabla^{2} $, to check that similar results are obtained in this case as when we use higher-order diffusion $ \nu_8 \nabla^{8} $.
+A number of test simulations with resolution $ n = 960 $ and $ c= 10 $ were
+first carried out using different values of the viscosity, until we determined a value which was sufficiently large for the shocks to be resolved, but not unnecessarily large.
+This value was then used for all simulations with $ n = 960 $. For the more
+highly resolved simulations the viscosity was then set as $ \nu \propto n^{-1} $, in accordance with the prediction  (\ref{Width}) for the shock width.
 We define the Reynolds number as $ Re \equiv \epsilon^{1/3} L_f^{4/3}/\nu $, which is proportional to $ n $, since $ \nu \propto n^{-1} $.
@@
 
+%TODO fix cite[] -> citep/parencite?
+
 \begin{table}
@@
 \toprule
-{} &   $n$ &  $c$ &  $\nu_2$ &  $\epsilon$ &  $\frac{k_{diss}}{k_f}$ &   $F_f$ &  $Re$ &  $ReF_f^{2/3}$ &  $t_{\max}$ \\
+{} &   $n$ &  $c$ &  $\nu_2$ &  $\epsilon$ &  $\frac{k_d}{k_f}$ &   $F_f$ &  $Re$ &  $ReF_f^{2/3}$ &  $t_{\max}$ \\
 \midrule
@@
 averaged mean energy dissipation in the stationary state, $ k_{d}/ k_f $: ratio
-between dissipation wave number and forcing wave number, $F_f $: Froude
+between dissipation wavenumber and forcing wavenumber, $F_f $: Froude
 number, $Re$: Reynolds number, $ t_{max} $: end time of simulation.}
@@
 
-The dynamics of the runs with Laplacian diffusion were similar to the runs with higher order diffusion.
+The dynamics of the runs with Laplacian diffusion was similar to the runs with higher-order diffusion.
 In figure \ref{MeanE2} we see mean energy in the stationary state versus $ c $ to the left and versus $ n $ to the right.
@@
 \end{equation}
-If this condition is not fulfilled the shocks will fade away.
-In figure \ref{Physical_lap} we have plotted the divergence for four different runs with different $ Re $ and $ F_f $.
-It should be pointed out again, that the area of the computational domain is sixteen times larger than area of the outcuts shown in the figures.
-The two upper plots are from simulations with  $ F_f = 0.16 $ but different $ Re $ ($ Re = 68 $ to the left and { $ Re = 203 $ } to the right).
-As can be seen, the shocks are wider in the lower Reynolds number run to the left than in the higher Reynolds number run to the right, as expected.
-The lower left plot is from a simulation with the same $ Re $ as in the upper left plot, but with a much smaller $ F_f $.
-In this simulation we have $ Re F_f^{2/3} = 2.7 $, so the condition (\ref{SCond}) is not fulfilled, and the shocks have almost faded away.
-The lower right plot is from a simulation with the same $ F_f $ as in the lower left plot, but with a higher $ Re = 199 $.
-Here, $ Re F_f^{2/3} = 7.8 $ and  the shocks are clearly discernible.
-{ Comparing the two right plots with each other, they have approximately the same $ Re $, while the lower one has a considerably lower $ F_f $.
-In this run the shocks are denser and straighter as compared to the higher $ F_f $-run.}
+If this condition is not fulfilled the shocks will fade away.  In figure
+\ref{Physical_lap} we have plotted the divergence for four different runs with
+different $ Re $ and $ F_f $. It should be pointed out again, that the area of
+the computational domain is sixteen times larger than area of the outcuts shown
+in the figures.  $(a,b)$ are from simulations with  $ F_f = 0.16 $ but
+different $ Re $ ($ Re = 68\, (a) $ and { $ Re = 203\, (b) $ }).
+As can be seen, the shocks are wider in the lower Reynolds number run $(a)$
+than in the higher Reynolds number run $(b)$, as expected. $(c)$
+is from a simulation with the same $ Re $ as in $(a)$, but with a much smaller
+$ F_f $. In this simulation we have $ Re F_f^{2/3} = 2.7 $, so the condition
+(\ref{SCond}) is not fulfilled, and the
+shocks have almost faded away. $(d)$ is from a simulation with
+the same $ F_f $ as in $(c)$, but with a higher $ Re = 199 $.
+Here, $ Re F_f^{2/3} = 7.8 $ and  the shocks are clearly discernible. {
+Comparing $(b)$ and $(d)$ with each other, they have approximately the same
+$ Re $, while the lower one has a considerably lower $ F_f $.  In this run the
+shocks are denser and straighter as compared to the higher $ F_f $-run.}
 

### 2019-08-14 paper_04_shallow_water/Latex/article.tex
Fix references cite[ -> citep[

@@
 The shallow water (SW) equations  have been widely used to study basic
-mechanisms occurring in geophysical flows \cite[see for
+mechanisms occurring in geophysical flows \citep[see for
 example][]{VallisLIVRE2006}.
@@
 primary importance for the dynamics of the oceans and of the
-atmosphere \cite[]{VallisLIVRE2006, Wirth2013}.
+atmosphere \citep[]{VallisLIVRE2006, Wirth2013}.
 %
@@
 been used to study issues like the projection on the balanced manifold
-\cite[]{Lorenz1980, MohebalhojehDritschel2000} and the production of
-waves by balanced flows \cite[]{FargeSadourny1989,
+\citep[]{Lorenz1980, MohebalhojehDritschel2000} and the production of
+waves by balanced flows \citep[]{FargeSadourny1989,
 LahayeZeitlin2012,Vanneste2013}. Balanced flows have a dynamics which is very similar to
@@
 In contrast to this, statistical mechanics indicates that wave energy
-should be transferred towards small scales \cite[]{Warn1986}.
+should be transferred towards small scales \citep[]{Warn1986}.
 Numerical simulations have shown that long waves lose their energy by
-transferring it to shorter waves \cite[]{Sadourny1975,
+transferring it to shorter waves \citep[]{Sadourny1975,
 FargeSadourny1989}.   \cite{YuanHamilton1994} showed that
@@
 small scales.
-They suggested that the $ k^{-5/3} $ energy spectrum observed at   atmospheric mesoscales  \cite[]{NastromGage1985,LiLindborg2018} is generated by a downscale cascade of gravity waves which may be captured by the SW equations.
+They suggested that the $ k^{-5/3} $ energy spectrum observed at   atmospheric mesoscales  \citep[]{NastromGage1985,LiLindborg2018} is generated by a downscale cascade of gravity waves which may be captured by the SW equations.
 
@@
  by some General Circulation Models (GCMs) based on the primitive
-equations \cite[]{KoshykHamilton2001, Skamarock2004,
+equations \citep[]{KoshykHamilton2001, Skamarock2004,
 Hamilton_etal2008}.  \cite{AugierLindborg2013}  formulated a spectral energy
@@
 model simulation only contained 21 vertical levels, while theoretical arguments
-based on stratified turbulence theory \cite[]{Lindborg2006} suggest that a much
+based on stratified turbulence theory \citep[]{Lindborg2006} suggest that a much
 finer vertical resolution would be needed.
@@
 downscale cascade. Based on the SW equations we developed another two-dimensional toy
-model  \cite[]{LindborgMohanan2017} that is absent of shocks and can reproduce the
+model  \citep[]{LindborgMohanan2017} that is absent of shocks and can reproduce the
 energy cascade in the GCM quite well, including a $ k^{-5/3}  $-spectrum.
@@
 Some predictions  based on the weak wave turbulence formalism
-\cite[]{ZakharovLvovFalkovich1992, Nazarenko2011} may be of relevance in the limit of weak nonlinearities. In the case of
+\citep[]{ZakharovLvovFalkovich1992, Nazarenko2011} may be of relevance in the limit of weak nonlinearities. In the case of
 inertia-gravity waves, \cite{FalkovichMedvedev1992} showed that the exact
@@
 
-Shocks have been observed in many simulations of the SW equations \cite[]{FargeSadourny1989,
+Shocks have been observed in many simulations of the SW equations \citep[]{FargeSadourny1989,
 PolvaniMcWilliamsSpallFord1994, LahayeZeitlin2012}.
@@
 shocks will scale as $ k^{-2} $, a prediction which has been confirmed by a
-great number of simulations of the Burgers equation \cite[for a review, see ]
+great number of simulations of the Burgers equation \citep[for a review, see ]
 []{Frisch2001}. The shocks also lead to very strong intermittency, as
@@
 In a non-rotating system of reference the SW equations for a thin layer of fluid over a flat bottom and with a free
-surface can be written as  \cite[see for example][]{VallisLIVRE2006}
+surface can be written as  \citep[see for example][]{VallisLIVRE2006}
 \begin{eqnarray} \label{UEq}
@@
 
-We will now derive an analogue of the so called `four-fifths  law' \cite[]{Kolmogorov1941} for the the third-order velocity structure function of incompressible isotropic turbulence.
+We will now derive an analogue of the so called `four-fifths  law' \citep[]{Kolmogorov1941} for the the third-order velocity structure function of incompressible isotropic turbulence.
 In the case of irrotational SW wave turbulence the analogous law contains a third-order structure function involving both velocity and displacement increments.
@@
 \end{equation}
-For separations that are considerably larger than dissipative scales of motion and considerably smaller than forcing scales a standard argument \cite[see for example][]{Frisch} gives that the left-hand side is approximately equal to $ -8 \epsilon $, where $ \epsilon $ is the mean dissipation rate.
+For separations that are considerably larger than dissipative scales of motion and considerably smaller than forcing scales a standard argument \citep[see for example][]{Frisch} gives that the left-hand side is approximately equal to $ -8 \epsilon $, where $ \epsilon $ is the mean dissipation rate.
 It is interesting to note that a factor of $ 8 $ appears in the case when the velocity field is irrotational while a corresponding factor of $ 4 $ appears in the case when the velocity field is non-divergent.
@@
 are exactly related by $\meane{(\delta u_L)^2} = \p_r( r\meane{(\delta
-u_T)^2} )$ \cite[]{Lindborg2007jas}. If we further assume than both second-order structure functions follow
+u_T)^2} )$ \citep[]{Lindborg2007jas}. If we further assume than both second-order structure functions follow
 the same scaling law $\meane{ (\delta u_L)^2} \propto \meane{ (\delta
@@
 One suggestion would be that $ \delta x $  is equal to the Kolmogorov length scale, $ \nu^{3/4}/\epsilon^{1/4} $.
-However, exact solutions of Burgers equation \cite[see for example][]{Whitham}
+However, exact solutions of Burgers equation \citep[see for example][]{Whitham}
 suggest that  $ \delta x \propto \nu $ rather than $ \delta x \propto \nu ^{3/4} $.
@@
 where the skewness is negative.
-For Kolmogorov turbulence observations indicate that $ F_{\partial_x u} \propto \nu^{-1/6} $  \cite[]{VanAtta_Antonia1980}, a dependence which is well capture by the multifractal model of
+For Kolmogorov turbulence observations indicate that $ F_{\partial_x u} \propto \nu^{-1/6} $  \citep[]{VanAtta_Antonia1980}, a dependence which is well capture by the multifractal model of
 \cite{Meneveau_Streenivasan1991}.  In SW wave turbulence there is a much
@@
 Finally, we estimate the Mach number, $ M = U/c $, where $ U $ is the shock propagation velocity in the frame of reference where the fluid velocity in front of the shock is zero.
-Mass and momentum conservation give \cite[see][]{Baines1998}
+Mass and momentum conservation give \citep[see][]{Baines1998}
 \begin{equation}
@@
 equations in a quadratic domain with periodic boundary conditions using the
-standard open-source pseudo-spectral code FluidSim \cite[]{fluiddyn, fluidfft,
+standard open-source pseudo-spectral code FluidSim \citep[]{fluiddyn, fluidfft,
 fluidsim}. In these twenty simulations we used an eighth-order diffusion operator to act as the agent of dissipation.
@@
 time step method which maximises the time step over a standard
-Courant-Friedrichs-Lewy condition \cite{Lundbladh1999, AugierChomazBillant2012}.
+Courant-Friedrichs-Lewy condition \citep{Lundbladh1999, AugierChomazBillant2012}.
 %
 Most of the aliasing is removed by truncating 8/9 of the modes along each
-direction \cite[for a detailed discussion on the issues of the non-conservation
+direction \citep[for a detailed discussion on the issues of the non-conservation
 of the non-quadratic energy and the aliasing errors in the truncated one-layer
@@
 scales. As a consequence, the normalised dissipation is of the order of unity
-independent of Reynolds number \cite[]{Pope, TennekesLumley}. A review of the
+independent of Reynolds number \citep[]{Pope, TennekesLumley}. A review of the
 experimental evidence and theoretical implications of this fundamental law of
@@
 on the left-hand side of (\ref{Dissipation}) is of the order of unity in the
-limit of strong stratification \cite[]{Lindborg2006, Brethouwer2007}.
+limit of strong stratification \citep[]{Lindborg2006, Brethouwer2007}.
 
@@
 turbulence is often derived using the assumption of a local cascade
-\cite[]{Vassilicos2015} or the assumption that dissipation stays finite in the
-limit of large Reynolds number \cite[]{Frisch}. As we just discussed, these
+\citep[]{Vassilicos2015} or the assumption that dissipation stays finite in the
+limit of large Reynolds number \citep[]{Frisch}. As we just discussed, these
 assumptions seem to be violated in the case of SW wave turbulence.
@@
 Another interesting observation is that the shocks in run W17 (lower figures) appear to have smaller curvature on average than the shocks in run W6 (upper figures).
-In both runs, the shocks are weak in the sense that the Mach number is close to unity and the interaction between the shocks should therefore be weak \cite[see][]{ApazidisEliasson2018}.
+In both runs, the shocks are weak in the sense that the Mach number is close to unity and the interaction between the shocks should therefore be weak \citep[see][]{ApazidisEliasson2018}.
 However, as measured by the Mach number  (\ref{Mach}), the shocks are much weaker in run W17, where $ c $ is larger by a factor of 20 as compared to run W6 and $ d $ also is smaller.
@@
 is increasing with decreasing scale is by some authors defined as the most
-appropriate measure of spatial intermittency \cite[see for example][]{Frisch}.
+appropriate measure of spatial intermittency \citep[see for example][]{Frisch}.
 If this measure is used, SW wave turbulence is, indeed, extremely intermittent.
@@
 general similarity theory can be easily formulated for this range. Therefore,
-predictions from weak turbulence theory \cite[for
+predictions from weak turbulence theory \citep[for
 example][]{ZakharovSagdeev1970} are not likely to hold for this range, although
@@
 On the other hand, if (\ref{WC}) not only is a necessary but also a sufficient condition,  wave breaking will be present for any given $ F_f $, no matter how small, provided that $ Re $ is sufficiently large.
-The condition (\ref{WC}) is analogous to the condition for the presence of turbulence in a stratified fluid, also involving both the Reynolds number and a Froude number \cite[]{Brethouwer2007}.
+The condition (\ref{WC}) is analogous to the condition for the presence of turbulence in a stratified fluid, also involving both the Reynolds number and a Froude number \citep[]{Brethouwer2007}.
 The horizontal and vertical length scales of stratified turbulence scale as $ l_h \sim u^{3}/\epsilon $ and $ l_v \sim u/N $, respectively, where $ u $ is a characteristic horizontal velocity and $ N $ is the Brunt-V\"ais\"al\"a frequency.
@@
 As is clear from our analysis, in the general case of a compressible and irrotational fluid the KE flux can be expressed in this way.
-It seems to be more of a specific property that the APE flux also can be expressed in terms of a third-order structure function, a property that the SW system shares with the Boussinesq system with constant Brunt-V\"ais\"al\"a frequency, in which case APE also is quadratic in buoyancy \cite[]{Augier2012}.
+It seems to be more of a specific property that the APE flux also can be expressed in terms of a third-order structure function, a property that the SW system shares with the Boussinesq system with constant Brunt-V\"ais\"al\"a frequency, in which case APE also is quadratic in buoyancy \citep[]{Augier2012}.
 In this context,  a comparison between our analysis and the analysis  by
@@
 
-While isothermal acoustic turbulence may have some interesting application for a radiating gas \cite[]{Stein1967}, standard text book acoustics is not isothermal but isentropic.
+While isothermal acoustic turbulence may have some interesting application for a radiating gas \citep[]{Stein1967}, standard text book acoustics is not isothermal but isentropic.
 The  condition of constant entropy puts us in a dilemma when we analyse an acoustic field as a forced-dissipative system, since dissipation inevitably is associated with entropy production.
@@
 Given the numerical results of the present study such a hypothesis seems reasonable, since shock waves have similar stability properties
-in two and three dimensions \cite[]{ApazidisEliasson2018, LivertsApazidis2016}.
+in two and three dimensions \citep[]{ApazidisEliasson2018, LivertsApazidis2016}.
 The shock wave hypothesis offers an elegant way out of the isentropic dilemma, since it implies that entropy production is concentrated at the shocks, while the rest of the field can be treated as isentropic.
-In a companion paper \cite[]{Lindborg2019}, scaling relations for such an acoustic field are derived, using the weak shock relations and the entropy equation.
+In a companion paper \citep[]{Lindborg2019}, scaling relations for such an acoustic field are derived, using the weak shock relations and the entropy equation.
 It turns out that these scaling relations are analogous to the relations for shallow water wave turbulence, with the  difference

### 2019-08-14 paper_02_fluidsim/article.tex
Cleanup and add todos
Update todos and comment out shallow water paper
Formatting equations
Fix more references cite[ -> citep[

@@
 The goal is to optimize such that ``80 percent of the runtime is spent in
-20 percent of the code''~\cite[]{meyers2012effective}.
+20 percent of the code''~\citep[]{meyers2012effective}.
 %
@@
 
-% todo: redo this figure (API changes)...
-
 \begin{figure}[htp]
@@
 \subsubsection{Performance}
-
-% todo: we need to go back over this part after the fluidfft paper will be
-% written and using the fluidfft paper. Some bits of what is here could go in
-% the fluidfft paper.
-%
-% It is not interesting here to multiply the FFT comparisons and to run these
-% benchmarks on many clusters. We have to be as simple as possible.
-%
-% We need to include the CFD code comparison here.
 
@@
 displayed within a single category, \emph{other}.
-
-% todo: redo these figures (API changes + better speed). Use with_pyfftw for
-% sequential runs.
 
@@
 
-\subsection{pseudo-spectral code comparisons}
+\subsection{CFD pseudo-spectral code comparisons}
 
@@
 Boussinesq approximation with stratification and system
-rotation~\cite[][]{DeloncleBillantChomaz2008}.It is in particular specialized in
-stability studies~\cite[][]{BillantDeloncleChomazOtheguy2010}.
+rotation~\citep[][]{DeloncleBillantChomaz2008}.It is in particular specialized in
+stability studies~\citep[][]{BillantDeloncleChomazOtheguy2010}.
 %
@@
 parallel simulations (using MPI and OpenMP).  However, the parallelization is
-limited to 1D decomposition for the FFT~\cite[][]{fluidfft}.
+limited to 1D decomposition for the FFT~\citep[][]{fluidfft}.
 %
@@
 code NS3D.  This is because this corresponds to Pythran functions written with
-explicit loops~\cite[see][]{fluidfft}.
+explicit loops~\citep[see][]{fluidfft}.
 
@@
 \item {\bf Minimum:} \fluidpack{dyn}, \Numpy, \pack{h5netcdf}, \fluidpack{fft}
-\cite[and FFT libraries, see][]{fluidfft}.
+\citep[and FFT libraries, see][]{fluidfft}.
 \item {\bf Optional:} \Scipy, \pack{mpi4py}, \pack{Cython} and
@@
 \fluidpack{sim} is used at LEGI and at KTH for studies on geophysical turbulence
-\cite[see for example][]{LindborgMohanan2017}.
+\citep[see for example][]{LindborgMohanan2017}.
 %

### 2019-08-14 paper_05_milestone_issf/article.tex
Fix more references cite[ -> citep[

@@
 is a strongly stratified regime associated with a downscale energy cascade
-\cite[]{RileyDeBruynKops2003, Lindborg2006}.
+\citep[]{RileyDeBruynKops2003, Lindborg2006}.
 %
@@
 stratified turbulence'' is obtained only for very small horizontal Froude
-number and large buoyancy Reynolds number \cite[]{Billant2001,
+number and large buoyancy Reynolds number \citep[]{Billant2001,
 BrethouwerBillantLindborg2007},
@@
 correspond to a regime very different than strongly stratified geophysical
-flows \cite[see for example][]{PraudFinchamSommeria2005,
+flows \citep[see for example][]{PraudFinchamSommeria2005,
 AugierBillantNegrettiChomaz2013, AugierBillantChomaz2015}.
@@
 $S_h(r) = C (\varepsilon r)^{2/3} = C' (rU_c^3/M)^{2/3}$, with $C$ a coefficient
-approximately equal to 6 \cite[see figure 15 in][]{AugierBillantChomaz2015}.
+approximately equal to 6 \citep[see figure 15 in][]{AugierBillantChomaz2015}.
 \begin{figure}[htb]
@@
 with a -5/3 scaling law have been observed numerically
-\cite[]{BrethouwerBillantLindborg2007, AugierBillantChomaz2015}.
+\citep[]{BrethouwerBillantLindborg2007, AugierBillantChomaz2015}.
 %

### 2019-08-14 paper_06_milestone/1st/article.tex
Rewrite chapter exp

@@
 {\mathcal{R}} $ is large, is highly relevant for applications in the
-ocean\cite[]{RileyDeBruynKops2003, Lindborg2006}, but is notoriously difficult
+ocean\citep[]{RileyDeBruynKops2003, Lindborg2006}, but is notoriously difficult
 to reproduce in experiments and simulations. In trying to reach the strongly
@@
 as MILESTONE 2016 (``M16") and MILESTONE 2017 (``M17"). The M16 campain has
-already been described in \cite{campagne2016}. The two experimental setups are
+already been described in \cite{campagne2016}. 
+The two experimental setups are
 nearly identical. They differ mainly on the precise locations of the density

### 2019-08-14 paper_00_fluiddyn_meta/article.tex
Formatting equations
Proof read swe paper

@@
 
-\item \fluidpack{fft}~\cite[see the companion paper][]{fluidfft}: a package which
+\item \fluidpack{fft}~\citep[see the companion paper][]{fluidfft}: a package which
 provides C++ and Python classes unifying various libraries to perform Fast Fourier
@@
 
-\item \fluidpack{sim}~\cite[see the companion paper][]{fluidsim}: Numerically
+\item \fluidpack{sim}~\citep[see the companion paper][]{fluidsim}: Numerically
 oriented framework to run sequential and parallel Computational Fluid Dynamics
@@
 A study using \fluidpack{sim} has just been published in Physics of Fluids
-\cite[]{LindborgMohanan2017}.
+\citep[]{LindborgMohanan2017}.
 
@@
 Experiments using \fluidpack{lab} have been carried out in the DAMTP fluid
-laboratory \cite[Cambridge,
+laboratory \citep[Cambridge,
 UK. cf.][]{LeclercqPartridgeAugierDalzielKerswell2016}, in ENS Lyon laboratory
-\cite[Lyon, France,][]{salort2018} and at LEGI \cite[Grenoble,
+\citep[Lyon, France,][]{salort2018} and at LEGI \citep[Grenoble,
 France,][]{ISSF2016}.
@@
 \fluidpack{image} is used to process images taken during experiments performed in the
-Coriolis platform at LEGI \cite[]{ISSF2016}.
+Coriolis platform at LEGI \citep[]{ISSF2016}.
 
@@
 (a large rotating platform participating in the European consortiums Euhit and
-Hydralab) and open the data obtained \cite[see, for example,][]{ISSF2016}. One of
+Hydralab) and open the data obtained \citep[see, for example,][]{ISSF2016}. One of
 the motivations behind creating this package is to study how to use open-source to
@@
 documentations on the web and for \fluidpack{fft} and \fluidpack{sim} in the
-two companion papers \cite[]{fluidfft, fluidsim}.
+two companion papers \citep[]{fluidfft, fluidsim}.
 %
@@
 
-\subsection{of the Python library \fluiddyn}
+\subsection{API of the Python library \fluiddyn}
 
@@
 
-\fbox{%
-\begin{minipage}{\textwidth}
+\noindent\fbox{%
+\begin{minipage}{0.97\textwidth}
   \codeinline{%
@@
 
-\fbox{%
-\begin{minipage}{0.5\textwidth}
+\noindent\fbox{%
+\begin{minipage}{0.55\textwidth}
 \codeinline{((im1\_1.png, im1\_2.png, im1\_3.png),\\
@@
 \fbox{%
-\begin{minipage}{0.47\textwidth}
+\begin{minipage}{0.385\textwidth}
 \codeinline{((im1\_1.png, im2\_1.png),\\

### 2019-08-16 imgs/wordcloud/plot.py
Cover and wordcloud

@@ -1,0 +1,93 @@
+import subprocess
+from itertools import chain
+from pathlib import Path
+
+import matplotlib.pyplot as plt
+import numpy as np
+from PIL import Image
+from wordcloud import STOPWORDS, ImageColorGenerator, WordCloud
+
+cwd = Path(__file__).absolute().parent
+root = cwd.parent.parent
+
+
+class args:
+    width = 600 * 4
+    height = 250 *4
+    max_words = 500
+    #  output = None
+    #  output = "plot"
+    output = cwd / 'wordcloud.png'
+    mask = cwd / "mask_darkblue_rect.png"
+    colormask = mask
+
+
+body = ""
+for filename in chain(
+    root.glob("chapter*.md"),
+    (root / "chapter_02_open_science.tex",),
+    root.glob("paper*/**/article.tex")
+):
+    if filename.suffix == ".tex":
+        plain_text = subprocess.check_output(("detex", "-n", filename), text=True)
+        body += plain_text
+    else:
+        with open(filename) as f:
+            body += f.read()
+
+
+blacklist = {
+    "code", "package", "language", "one", "use", "time", "tool",
+    "kappa", "using", "used", "Matlab", "emph", "many", "footnote",
+    "mathbf", "fnref", "eq", "https", "http", "href", "url", "hat", "theta",
+    "will", "Thus", "found", "F_h", "u_c", "two", "end", "org", "cite",
+    "item", "begin", "Bmatrix", "textbf", "label", "caption", "partial_t",
+    "see", "delta", "figure", "equation", "sim", "run", "different", "value",
+    "term", "itemize", "example", "may", "become", "three", "case", "eqnarray",
+    "now", "even", "function",
+    "citep", "shown", "pack", "several", "make", "codeinline", "src", "fig",
+    "raw", "html", "latex", "bf", "img", "div", "math", "png", "alt", "align",
+    "green", "Ding", "link", "nbsp", "cdot", "nabla", "frac", "partial", "eta",
+    "style", "width", "right", "left", "center", "Result", "drawing",
+}
+stopwords = set(STOPWORDS).union(blacklist)
+
+
+def imread(f):
+    return (np.array(Image.open(f)))
+
+
+mask = imread(args.mask)
+color_func = ImageColorGenerator(imread(args.colormask))
+wordcloud = WordCloud(
+    background_color="white",
+    stopwords=stopwords,
+    width=args.width,
+    height=args.height,
[…]

### 2019-08-19 paper_04_shallow_water/Latex/article.tex
Fix fig 15 in swe paper

@@
 \begin{figure}
-\centerline{\includegraphics[]{paper_04_shallow_water/Pyfig/fig15}}
-\caption{$ \nabla \cdot {\bf u} $ for four different runs.
-Upper left: WL1, $ Re = 68 $, $ F_f = 0.16 $; Upper right:  { WL3,  $ Re = 203 $, $ F_f = 0.16 $;} Lower left: WL17, $ Re = 68 $, $ F_f = 0.008 $; Lower right: WL18,  $ Re = 199 $, $ F_f = 0.008 $.}
+\centerline{\includegraphics[]{paper_04_shallow_water/Pyfig/fig15-eps-converted-to.pdf}}
+\caption{Value of $ \nabla \cdot {\bf u} $ for four different runs. $(a)$ WL1,
+$ Re = 68 $, $ F_f = 0.16 $; $(b)$ { WL3,  $ Re = 203 $, $ F_f = 0.16 $;} $(c)$
+WL17, $ Re = 68 $, $ F_f = 0.008 $; $(d)$ WL18,  $ Re = 199 $, $ F_f = 0.008 $.}
 \label{Physical_lap}

### 2019-08-19 scripts/pdf_separate.py
Try pdf separate, pause ci

@@ -1,0 +1,87 @@
+import os
+import shlex
+import subprocess
+from pathlib import Path
+
+import PyPDF2 as pypdf
+from PyPDF2 import PdfFileReader
+
+
+def _setup_page_id_to_num(pdf, pages=None, _result=None, _num_pages=None):
+    if _result is None:
+        _result = {}
+    if pages is None:
+        _num_pages = []
+        pages = pdf.trailer["/Root"].getObject()["/Pages"].getObject()
+    t = pages["/Type"]
+    if t == "/Pages":
+        for page in pages["/Kids"]:
+            _result[page.idnum] = len(_num_pages)
+            _setup_page_id_to_num(pdf, page.getObject(), _result, _num_pages)
+    elif t == "/Page":
+        _num_pages.append(1)
+    return _result
+
+
+def outlines_pg_zoom_info(outlines, pg_id_num_map, result=None):
+    if result is None:
+        result = dict()
+    if type(outlines) == list:
+        for outline in outlines:
+            outlines_pg_zoom_info(outline, pg_id_num_map, result)
+    elif type(outlines) == pypdf.pdf.Destination:
+        title = outlines["/Title"]
+        result[title] = dict(
+            title=outlines["/Title"],
+            top=outlines["/Top"],
+            left=outlines["/Left"],
+            page=(pg_id_num_map[outlines.page.idnum] + 1),
+        )
+    return result
+
+
+def get_page_numbers(pdf_name):
+    with open(pdf_name, "rb") as f:
+        pdf = PdfFileReader(f)
+        total_pages = pdf.numPages
+        # map page ids to page numbers
+        pg_id_num_map = _setup_page_id_to_num(pdf)
+        outlines = pdf.getOutlines()
+        bookmarks_info = outlines_pg_zoom_info(outlines, pg_id_num_map)
+
+    #  print(pg_id_num_map)
+    #  print(bookmarks_info)
+    pages = {meta["title"]: meta["page"] for meta in bookmarks_info.values()}
+
+    return pages, total_pages
+
+
+def pdf_separate(pdf_name, output_dir, version, dry_run=True):
+    pages, total_pages = get_page_numbers(pdf_name)
+    #  print("Pages:", pages)
+
+    pages_range = {"Overview": ["1"]}
+    for title, page in pages.items():
+        if title.startswith("Paper") and "Abstracts" not in title:
+            print(title)
+            article_num = int(title[6])  # 7th character
[…]

### 2019-08-19 chapter_02_open_science.tex
Typos

@@
 	Especially for the beginners who are not used to the open-source realm,
-	it can be confusing prospect to make good technological choices; but
+	it can be a confusing prospect to make good technological choices; but
 	this can be solved through proper guidance and training.

### 2019-08-20 imgs/spectral_flux/plot.py
Make figures grayscale friendly

@@ -8,7 +8,7 @@ from pathlib import Path
 
 root = Path(__file__).absolute().parent.parent
 
-plt.style.use("seaborn-paper")
+#  plt.style.use("seaborn-paper")
 
 matplotlib.rcParams['text.latex.preamble'] = [r"\usepackage{amsmath}"]
 plt.rc('text', usetex=True)
@@ -21,18 +21,20 @@ if SIDE_TEXT:
     fig.subplots_adjust(right=5/10)
     ax = fig.add_subplot(111)
 else:
-    fig, ax = plt.subplots(figsize=(5,5))
+    fig, ax = plt.subplots(figsize=(4,4))
 
-circle1 = Circle((0, 0), pi/2)
-circle2 = Circle((0, 0), pi/1.9)
-circle3 = Circle((0, 0), pi * 2)
+circle1 = Circle((0, 0), pi/2, facecolor="gray")
+circle2 = Circle((0, 0), pi/1.9, facecolor="black")
+circle3 = Circle((0, 0), pi * 2, facecolor="white")
 patches = [circle3, circle2, circle1]
 
-p = PatchCollection(patches, cmap="Pastel1")#, alpha=0.4)
+for circle in patches:
+    ax.add_patch(circle)
+#  p = PatchCollection(patches) # , cmap="gist_gray")  #,alpha=0.4)
 #  colors = np.array([54, 50, 20])
-p.set_array(np.array([0.1, 2, 9]))
+#  p.set_array(np.array([0.1, 2, 3]))
 #  p.set_cmap(cmap)
-ax.add_collection(p)
+#  ax.add_collection(p)
 
 arror_r = pi / 2**1.5
 ax.arrow(arror_r, arror_r, 0.3, 0.3, head_width=0.05, head_length=0.1, fc='k', ec='k')

### 2019-08-20 scripts/pdf_separate.py
PDF separate using pdftk

@@ -74,14 +74,16 @@ def pdf_separate(pdf_name, output_dir, version, dry_run=True):
     print("Pages range:", pages_range)
 
     output_dir = Path.cwd() / output_dir / version
+    if not dry_run:
+        os.makedirs(output_dir, exist_ok=True)
     for part, prange in pages_range.items():
         prange = "-".join(prange)
         output = output_dir / ("_".join((version, part)) + ".pdf")
-        cmd = f"pdftk {pdf_name} cat {prange} {output}"
+        cmd = f"pdftk {pdf_name} cat {prange} output {output}"
         print(cmd)
         if not dry_run:
-            os.makedirs(output_dir)
+            subprocess.run(shlex.split(cmd))
 
 
 if __name__ == "__main__":
-    pdf_separate("thesis.pdf", "print", "v1")
+    pdf_separate("thesis.pdf", "print", "v3.0a", False)

### 2019-08-26 chapter_02_open_science.tex
Redo figures and a final proofread

@@
 be simpler for the beginners and just as powerful for expert users\footnote{See
-  \href{https://www.mercurial-scm.org/wiki/GitConcepts}{``a detailed
-comparison''} on Mercurial website.}. Simplicity and low learning barriers for
+  {``a detailed comparison''} on Mercurial website,
+  \url{https://www.mercurial-scm.org/wiki/GitConcepts}.
+  }. Simplicity and low learning barriers for
 scientists who are not experts in software development is crucial for our
@@
 %
-Sage and Sympy are two complementary open-source alternative both based on
+Sage and Sympy are two complementary open-source alternatives both based on
 Python.
@@
 Big companies have started to use open-source development for fluid mechanics
-application. For example, Volkswagen group (Volkswagen, Audi, Seat, Porsche,
-Skoda, ...) \fnref{%
+application. For example, the automotive group Volkswagen \fnref{%
   https://www.nas.nasa.gov/assets/pdf/ams/2014/AMS_20141202_Othmer.pdf}{%
@@
 open-source strategy}. It has produced many programs and in particular the CFD
-solver \fnref{https://www.code-saturne.org}{Code\_Saturn}.
+solver \fnref{https://www.code-saturne.org}{Code\_Saturne}.
 %
 \fnref{https://blog.kitware.com/introducing-some-new-paraview-5-0-features/}{%
-EDF is also involved in the development of \fnref{https://www.paraview.org/}{Paraview}}.
+EDF is also involved in the development} of \fnref{https://www.paraview.org/}{Paraview}.
 
@@
 \fnref{https://github.com/spectralDNS}{SpectralDNS},
-\fnref{https://github.com/mikaem/Oasis}{Oasis}, \href{http://pyfr.org/}{PyFR}
+\fnref{https://github.com/mikaem/Oasis}{Oasis}, \fnref{http://pyfr.org/}{PyFR}
 and \fnref{https://fenicsproject.org/}{FEniCS}) or through data analysis

### 2019-08-26 scripts/pdf_separate.py
Pdf separate in parallel

@@ -1,6 +1,8 @@
 import os
 import shlex
 import subprocess
+# from concurrent.futures import ProcessPoolExecutor as Pool
+from multiprocessing import Pool
 from pathlib import Path
 
 import PyPDF2 as pypdf
@@ -56,6 +58,10 @@ def get_page_numbers(pdf_name):
     return pages, total_pages
 
 
+def run(cmd):
+    subprocess.run(shlex.split(cmd))
+
+
 def pdf_separate(pdf_name, output_dir, version, dry_run=True):
     pages, total_pages = get_page_numbers(pdf_name)
     #  print("Pages:", pages)
@@ -76,14 +82,19 @@ def pdf_separate(pdf_name, output_dir, version, dry_run=True):
     output_dir = Path.cwd() / output_dir / version
     if not dry_run:
         os.makedirs(output_dir, exist_ok=True)
+
+    cmds = []
     for part, prange in pages_range.items():
         prange = "-".join(prange)
         output = output_dir / ("_".join((version, part)) + ".pdf")
         cmd = f"pdftk {pdf_name} cat {prange} output {output}"
         print(cmd)
-        if not dry_run:
-            subprocess.run(shlex.split(cmd))
+        cmds.append(cmd)
+
+    if not dry_run:
+        with Pool() as pool:
+            pool.map(run, cmds)
 
 
 if __name__ == "__main__":
-    pdf_separate("thesis.pdf", "print", "v3.0a", False)
+    pdf_separate("thesis.pdf", "print", "v3.1a", False)

### 2019-08-26 scripts/pdf_separate.py
Version v3.2a

@@ -97,4 +97,4 @@ def pdf_separate(pdf_name, output_dir, version, dry_run=True):
 
 
 if __name__ == "__main__":
-    pdf_separate("thesis.pdf", "print", "v3.1a", False)
+    pdf_separate("thesis.pdf", "print", "v3.2a", False)

### 2019-08-26 scripts/pdf_separate.py
Rerender because table of contents was wrong

@@ -97,4 +97,4 @@ def pdf_separate(pdf_name, output_dir, version, dry_run=True):
 
 
 if __name__ == "__main__":
-    pdf_separate("thesis.pdf", "print", "v3.2a", False)
+    pdf_separate("thesis.pdf", "print", "v3.3a", False)

### 2019-08-27 paper_03_toy_model/article.tex

@@
 For such waves, $ E_K > E_A $.
-However, the kinetic energy content in the rotational component of the velocity field is necessarily smaller than in the divergent component, and that is at odds with observational evidence \citep{Lindborg:2015, QiangLi2017}.
+However, the kinetic energy content in the rotational component of the velocity field is necessarily smaller than in the divergent component, and that is at odds with observational evidence \citep{Lindborg:2015, LiLindborg2018}.
 There is also numerical evidence \citep{Deusebio-Vallgren-Lindborg:2013, Asselin2017} suggesting that the forward energy mesoscale energy cascade cannot be described as a cascade of waves at high wave numbers, although it may start as a wave cascade at low wave numbers.

### 2019-08-27 paper_00_fluiddyn_meta/article.tex

@@
 European Union\footnote{See the report by Internet Association titled
-\href{https://internetassociation.org/reports/refreshing-understanding-internet-economy-ia-report/}
-{``Refreshing Our Understanding of the Internet Economy''}}.
+{``Refreshing Our Understanding of the Internet Economy''}
+\url{https://internetassociation.org/reports/refreshing-understanding-internet-economy-ia-report/}
+}.
 % \cite{hooton_christopher_refreshing_2017}.

### 2019-08-27 paper_02_fluidsim/article.tex
Fix Table 1 in fluidsim paper

@@
 \centering
-\begin{tabular}{l p{8cm}}
+\begin{tabular}{l p{7cm}}
   \toprule
@@
 \caption{Comparison of the execution times for a 3D case (128$^3$, 10 time steps)
-between NS3D (blue bars) and \codeinline{fluidsim.solvers.ns3d} (yellow bars).
+between NS3D (blue / dark bars) and \codeinline{fluidsim.solvers.ns3d} (yellow / light bars).
 %
@@
 Fig.~\ref{fig:compare:with:ns3d} presents a more detailed comparison between NS3D
-(blue bars) and \fluidpack{sim} (yellow bars).
+(blue / dark bars) and \fluidpack{sim} (yellow / light bars).
 %

### 2019-09-18 scripts/pandoc_auto_height.py
Of path manipulations

@@ -1,3 +1,6 @@
+from pathlib import Path
+import os
+
 import panflute as pf
 
 
@@ -14,9 +17,19 @@ def extract_image_from_pdf_page(xObject):
 
 
 def detect_height(elem):
-    url = elem.url
+    url = Path(elem.url)
+    if not url.exists() and not url.is_absolute():
+        # Try parent dir
+        new_url = url.parent.parent / url
+        if new_url.exists():
+            url = new_url
+        else:
+            raise
+            pf.debug(f"WARNING: {url} not found")
+            return "100%"
+
     width = int(elem.attributes["width"].rstrip('%'))
-    if url.endswith(".pdf"):
+    if url.suffix == ".pdf":
         from PyPDF2 import PdfFileReader
         with open(url, "rb") as fp:
             pdf = PdfFileReader(fp)
@@ -33,7 +46,7 @@ def detect_height(elem):
 
 def action(elem, doc):
     """Automatically determines height attribute for images."""
-    if doc.format == 'latex' and isinstance(elem, pf.Image):
+    if doc.format in ('latex', 'beamer') and isinstance(elem, pf.Image):
         # pf.debug(elem)
         if "width" in elem.attributes and "height" not in elem.attributes:
             elem.attributes["height"] = detect_height(elem)

### 2019-09-18 scripts/pandoc_auto_height.py
Auto height handles more units

@@ -1,5 +1,5 @@
 from pathlib import Path
-import os
+import re
 
 import panflute as pf
 
@@ -24,11 +24,14 @@ def detect_height(elem):
         if new_url.exists():
             url = new_url
         else:
-            raise
             pf.debug(f"WARNING: {url} not found")
-            return "100%"
+            return
 
-    width = int(elem.attributes["width"].rstrip('%'))
+    r = re.compile(r'([0-9]*[.])?[0-9]+')
+    w = elem.attributes["width"]
+    width = float(r.match(w).group(0))
+    unit = r.sub("", w)
+
     if url.suffix == ".pdf":
         from PyPDF2 import PdfFileReader
         with open(url, "rb") as fp:
@@ -40,16 +43,17 @@ def detect_height(elem):
         import imageio
         w, h, colors = imageio.imread(url).shape
 
-    height = f"{width*h/w}%"
+    height = f"{width*h/w} {unit}"
     return height
 
 
 def action(elem, doc):
     """Automatically determines height attribute for images."""
-    if doc.format in ('latex', 'beamer') and isinstance(elem, pf.Image):
-        # pf.debug(elem)
+    if doc.format in ('latex', 'beamer', 'native') and isinstance(elem, pf.Image):
         if "width" in elem.attributes and "height" not in elem.attributes:
-            elem.attributes["height"] = detect_height(elem)
+            height = detect_height(elem)
+            if height:
+                elem.attributes["height"] = height
 
         return elem
 

### 2019-09-19 scripts/pandoc_auto_height.py
Script update

@@ -58,8 +58,16 @@ def action(elem, doc):
         return elem
 
 
+def prepare(doc):
+    pf.debug('Starting "pandoc_auto_height.py" ...')
+
+
+def finalize(doc):
+    pf.debug('Ending "pandoc_auto_height.py" ...')
+
+
 def main(doc=None):
-    return pf.run_filter(action, doc=doc)
+    return pf.run_filter(action, prepare, finalize, doc=doc)
 
 
 if __name__ == '__main__':

### 2019-09-19 slides/scripts/_pandoc_wrapfig.py

@@ -1,0 +1,63 @@
+#! /usr/bin/env python3
+# -*- coding: utf-8 -*-
+"""Pandoc filter to allow variable wrapping of LaTeX/pdf documents
+through the wrapfig package.
+
+Simply add a " {?}" tag to the end of the caption for the figure, where
+? is an integer specifying the width of the wrap in inches. 0 will
+cause the width of the figure to be used.
+
+"""
+import panflute as pf
+# FIXME: Not working yet
+
+
+def wrapfig(elem, doc):
+    attrs = elem.attributes
+    caption = ''.join(pf.stringify(e) for e in elem.content)
+    #  caption = elem.content
+    #  pf.debug(caption)
+    target = elem.url
+    fmt = "latex"
+
+    # Strip tag
+    size = attrs.get("width")
+    if "lineheight" in attrs:
+        lineheight = attrs.lineheight
+        latex_begin = r"\begin{wrapfigure}[" + lineheight + "]{l}{" + size + "}"
+    else:
+        latex_begin = r"\begin{wrapfigure}{l}{" + size + "}"
+
+    if len(caption) > 0:
+        latex_fig = r"\centering\includegraphics{" + target + "}\caption{"
+        latex_end = r"}\end{wrapfigure}"
+        return pf.RawInline(latex_begin + latex_fig + caption + latex_end, fmt)
+        # return list(pf.RawInline(latex_begin + latex_fig), caption, pf.RawInline(latex_end))
+    else:
+        latex_fig = r"\centering\includegraphics{" + target[0] + "}"
+        latex_end = r"\end{wrapfigure}"
+        return pf.RawInline(latex_begin + latex_fig + latex_end, fmt)
+
+
+def action(elem, doc):
+    """Use wrapfig instead of includegraphics."""
+    if (
+        doc.format in (
+            "latex", "beamer",
+            "native", "json",
+        ) and isinstance(
+            elem, pf.Image
+        )
+    ):
+        if "wrap" in elem.attributes:
+            return wrapfig(elem, doc)
+        else:
+            return elem
+
+
+def main(doc=None):
+    return pf.run_filter(action, doc=doc)
+
+
+if __name__ == "__main__":
+    main()

### 2019-09-19 slides/scripts/pandoc_moreblocks.py

@@ -40,7 +40,8 @@ blockmap = {
     'fact':         'Fakt',
     'facts':        'Fakten',
     'lemma':        'Lemma',
-    'solution':     'Loesung'
+    'solution':     'Loesung',
+    'cbox':         'mycolorbox',
 }
 
 """

### 2019-09-20 slides/scripts/pandoc_footnote.py
Pandoc footnote filter

@@ -1,0 +1,32 @@
+"""Pandoc filter for footnotes in beamer which function. Forces [frame] option,
+parses citation and returns a RawInline latex element instead."""
+import panflute as pf
+
+
+def cite(word):
+    """@Author -> \citet{Author}"""
+    if word.startswith("@"):
+        return rf"\citet{{{word.lstrip('@')}}}"
+    else:
+        return word
+
+
+def natbib_citations(txt):
+    return " ".join(cite(word) for word in txt.split(" "))
+
+
+def action(elem, doc):
+    if doc.format in ("beamer", "native") and isinstance(elem, pf.Note):
+        txt = pf.stringify(elem, newlines=False)
+        txt = natbib_citations(txt)
+        tex = rf"\footnote<.->[frame]{{{txt}}}"
+        elem = pf.RawInline(tex, format="latex")
+    return elem
+
+
+def main(doc=None):
+    return pf.run_filter(action, doc=doc)
+
+
+if __name__ == "__main__":
+    main()

### 2019-09-20 slides/scripts/pandoc_footnote.py
Improved footnote filter, fix unique in refs, quiet biber

@@ -3,23 +3,28 @@ parses citation and returns a RawInline latex element instead."""
 import panflute as pf
 
 
-def cite(word):
-    """@Author -> \citet{Author}"""
+def elem2cite(elem, doc):
+    """Safely extract a Cite element and convert to RawInline"""
+    if isinstance(elem, pf.Cite):
+        citation = elem.content[0]
+        tex = pf.RawInline(str2cite(pf.stringify(citation)), format="latex")
+        #  pf.debug(tex)
+        return tex
+
+
+def str2cite(word):
+    r"""@Author -> \citet{Author}"""
     if word.startswith("@"):
         return rf"\citet{{{word.lstrip('@')}}}"
-    else:
-        return word
 
-
-def natbib_citations(txt):
-    return " ".join(cite(word) for word in txt.split(" "))
+    return word
 
 
 def action(elem, doc):
     if doc.format in ("beamer", "native") and isinstance(elem, pf.Note):
+        elem = elem.walk(elem2cite)
         txt = pf.stringify(elem, newlines=False)
-        txt = natbib_citations(txt)
-        tex = rf"\footnote<.->[frame]{{{txt}}}"
+        tex = rf"\footnote<.->[frame]{{\tiny {txt}}}"
         elem = pf.RawInline(tex, format="latex")
     return elem
 

### 2019-09-20 scripts/pandoc_auto_height.py
Fix auto_height filter

@@ -20,12 +20,11 @@ def detect_height(elem):
     url = Path(elem.url)
     if not url.exists() and not url.is_absolute():
         # Try parent dir
-        new_url = url.parent.parent / url
+        new_url = Path("..") / url
         if new_url.exists():
             url = new_url
         else:
-            pf.debug(f"WARNING: {url} not found")
-            return
+            raise FileNotFoundError(f"Neither {url} nor {new_url} found")
 
     r = re.compile(r'([0-9]*[.])?[0-9]+')
     w = elem.attributes["width"]
@@ -43,7 +42,11 @@ def detect_height(elem):
         import imageio
         w, h, colors = imageio.imread(url).shape
 
-    height = f"{width*h/w} {unit}"
+    if url.suffix == ".eps":
+        height = f"{width*w/h}{unit}"  # TODO: wtf!
+    else:
+        height = f"{width*h/w}{unit}"
+
     return height
 
 
@@ -59,11 +62,11 @@ def action(elem, doc):
 
 
 def prepare(doc):
-    pf.debug('Starting "pandoc_auto_height.py" ...')
+    pf.debug(f'Starting {__file__} ...')
 
 
 def finalize(doc):
-    pf.debug('Ending "pandoc_auto_height.py" ...')
+    pf.debug(f'Ending {__file__} ...')
 
 
 def main(doc=None):

### 2019-09-20 slides/scripts/pandoc_footnote.py
Preface and frame number consistency

@@ -29,8 +29,16 @@ def action(elem, doc):
     return elem
 
 
+def prepare(doc):
+    pf.debug(f'Starting {__file__} ...')
+
+
+def finalize(doc):
+    pf.debug(f'Ending {__file__} ...')
+
+
 def main(doc=None):
-    return pf.run_filter(action, doc=doc)
+    return pf.run_filter(action, prepare, finalize, doc=doc)
 
 
 if __name__ == "__main__":

### 2019-09-21 slides/scripts/pandoc_moreblocks.py
New tcolorbox envs, remove Type 3 fonts

@@ -41,7 +41,9 @@ blockmap = {
     'facts':        'Fakten',
     'lemma':        'Lemma',
     'solution':     'Loesung',
-    'cbox':         'mycolorbox',
+    'cbox':         'bluecolorbox',
+    'rbox':         'redcolorbox',
+    'gbox':         'greencolorbox',
 }
 
 """

### 2019-09-23 scripts/pandoc_auto_height.py
Keep aspect ration of figures, rename template

@@ -42,6 +42,7 @@ def detect_height(elem):
         import imageio
         w, h, colors = imageio.imread(url).shape
 
+    w, h = (float(i) for i in (w, h))
     if url.suffix == ".eps":
         height = f"{width*w/h}{unit}"  # TODO: wtf!
     else:

### 2019-09-24 slides/scripts/pandoc_footnote.py
Increase sizes of footnote and all text

@@ -24,7 +24,8 @@ def action(elem, doc):
     if doc.format in ("beamer", "native") and isinstance(elem, pf.Note):
         elem = elem.walk(elem2cite)
         txt = pf.stringify(elem, newlines=False)
-        tex = rf"\footnote<.->[frame]{{\tiny {txt}}}"
+        #  tex = rf"\footnote<.->[frame]{{\tiny {txt}}}"
+        tex = rf"\footnote<.->[frame]{{{txt}}}"
         elem = pf.RawInline(tex, format="latex")
     return elem
 

### 2019-09-24 slides/scripts/pandoc_footnote.py
Add more images and avoid pandoc footnote filter inside image captions

@@ -20,8 +20,16 @@ def str2cite(word):
     return word
 
 
+def inside_image(elem):
+    return isinstance(elem.ancestor(1), pf.Image)
+
+
 def action(elem, doc):
-    if doc.format in ("beamer", "native") and isinstance(elem, pf.Note):
+    if (
+        doc.format in ("beamer", "native")
+        and isinstance(elem, pf.Note)
+        and not inside_image(elem)
+    ):
         elem = elem.walk(elem2cite)
         txt = pf.stringify(elem, newlines=False)
         #  tex = rf"\footnote<.->[frame]{{\tiny {txt}}}"
@@ -31,11 +39,11 @@ def action(elem, doc):
 
 
 def prepare(doc):
-    pf.debug(f'Starting {__file__} ...')
+    pf.debug(f"Starting {__file__} ...")
 
 
 def finalize(doc):
-    pf.debug(f'Ending {__file__} ...')
+    pf.debug(f"Ending {__file__} ...")
 
 
 def main(doc=None):

### 2019-09-24 imgs/cascade/plot.py
Fix figures font sizes in frames 2 and 3

@@ -4,12 +4,22 @@ import numpy as np
 from pathlib import Path
 
 
+BEAMER = True
+
+
 root = Path(__file__).parent.parent
 
 plt.style.use("ggplot")
-plt.rc("text", usetex=True)
 
-fig, axes = plt.subplots(2, 1, dpi=200, figsize=(3.5, 6))
+if BEAMER:
+    with plt.xkcd():
+        fig, axes = plt.subplots(1, 2, dpi=200, figsize=(9, 3.2))
+    font = {"size": 11}
+else:
+    plt.rc("text", usetex=True)
+    fig, axes = plt.subplots(2, 1, dpi=200, figsize=(3.5, 6))
+    font = {"size": 8}
+
 ax1, ax2 = axes.ravel()
 
 km = 1.5
@@ -20,11 +30,11 @@ km = 10 ** km
 coef = km ** (-6.7 / 5)
 
 print("coef =", coef)
-ax1.set_title("Energy spectra predicted by Kraichnan's theory", fontsize=9)
+ax1.set_title("Energy spectra predicted by Kraichnan's theory", fontsize=font["size"] + 1)
 ax1.loglog(k1, k1 ** -(5 / 3) * coef, c)
 ax1.loglog(k2, k2 ** (-3), c)
 
-ax2.set_title("Observed atmospheric energy spectra", fontsize=9)
+ax2.set_title("Observed atmospheric energy spectra", fontsize=font["size"] + 1)
 ax2.loglog(k2, k2 ** (-5 / 3) * coef, c)
 ax2.loglog(k1, k1 ** -3, c)
 
@@ -38,27 +48,27 @@ ax1.annotate(
         # posB=(km, km ** -1.5)
     ),
     ha="left",
-    fontsize=8,
+    fontsize=font["size"],
 )
 ax1.text(
     0.1*km, (0.8*km)**(-5/3),
     r"$\epsilon^{2/3}k^{-5/3}$",
-    fontdict=dict(size=8),
+    fontdict=font,
 )
 ax1.text(
     100*km, (80*km)**-3,
     r"$\eta^{2/3}k^{-3}$",
-    fontdict=dict(size=8),
+    fontdict=font,
 )
 ax2.text(
     0.1*km, (0.08*km)**(-3),
     r"$k^{-3}$",
-    fontdict=dict(size=8),
+    fontdict=font,
 )
 ax2.text(
     100*km, (800*km)**-(5/3),
[…]
